### 2016-06-22 notes/Fernanda/Jan19-2016/notes.tex
added notes

\documentclass[12pt]{article}
\usepackage[top=0.85in, bottom=0.93in, right=0.93in, left=0.93in,
paperwidth=8.5in, paperheight=11in, nohead]{geometry}
\geometry{letterpaper}
\usepackage[pdftex]{graphicx}
\usepackage{color}
\usepackage[normalem]{ulem}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage{setspace}
\usepackage{mdwlist}
\usepackage{verbatim}
\usepackage[numbers]{natbib}

%% make bibliography more compact \setlength{\bibsep}{0in}
\renewcommand\bibsection{\subsubsection*{\refname}}

\begin{document}

\begin{centering}
  \large {\bf The assembly of pollinator communities in response to
    restoration} \\
\end{centering}
\vspace{0.15in}

\section{Goals}
\begin{enumerate}
\item Explore mechanisms underlying the assembly of plant-pollinator
  communities on on-farm habitat restorations by comparing empirical
  observations with assembly theory and adaptive foraging
\item How specialization changes in a pollination network during the
  pollinator guild assembly (colonization, persistence and extirpation
  processes over time)
\end{enumerate}

\section{Data prep}
\begin{enumerate}
\item starting plant community for each hedgerow summed over the first
  three years following installment
\item the number of pollinators at each site in the initial three
  years, and those added/lost after the initial three years
\item the mean, minimum and maximum pollinator species degree, summed
  over years/for each year
\end{enumerate}

\section{Model}
\begin{enumerate}
\item Use the initial network summed over years 1-3 for each hedgerow,
  randomly add species (equal to the number added in the observed
  data) and allow the community to assembly using adaptive foraging
\item  Keep track of
  \begin{enumerate}
  \item number of species that go extinct at each hedgerow
  \item the specialization of each plant and pollinator
  \end{enumerate}
\end{enumerate}

\section{Predictions}
\begin{enumerate}
\item ``generalist'' pollinators early in succession will become more
  specialized
\item ``generalist'' pollinators early in succession will visit a
  subset of what they originally visited, mainly the originally
  ``specialist'' plants
\end{enumerate}

\section{Considerations}
\begin{enumerate}
\item some plant species that were planted were never visited by
  pollinators in the first few years. This is likely because they did
  not produce many/any flowers/nectar resources
\item compare observed data to a model with/without adaptive foraging?
\item all the pollinator species needed (bees, syrphids, wasps
  etc.?). Not all have been identified to species.
\item weedy plant species ``added'' to communities by invasion, lost
  by death
\end{enumerate}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-PDF-mode: t
%%% End:

### 2016-06-23 notes/Marillia/Jun23-2016/notes.tex
added Peel's code

--- 2016-06-22 notes/Fernanda/Jan19-2016/notes.tex
@@
 \begin{centering}
-  \large {\bf The assembly of pollinator communities in response to
-    restoration} \\
+  \large {\bf The assembly of plant-pollinator communities in response
+  to restoration8} \\
 \end{centering}
@@
 
-\section{Goals}
+\section{Questions}
 \begin{enumerate}
-\item Explore mechanisms underlying the assembly of plant-pollinator
-  communities on on-farm habitat restorations by comparing empirical
-  observations with assembly theory and adaptive foraging
-\item How specialization changes in a pollination network during the
-  pollinator guild assembly (colonization, persistence and extirpation
-  processes over time)
-\end{enumerate}
-
-\section{Data prep}
-\begin{enumerate}
-\item starting plant community for each hedgerow summed over the first
-  three years following installment
-\item the number of pollinators at each site in the initial three
-  years, and those added/lost after the initial three years
-\item the mean, minimum and maximum pollinator species degree, summed
-  over years/for each year
-\end{enumerate}
-
-\section{Model}
-\begin{enumerate}
-\item Use the initial network summed over years 1-3 for each hedgerow,
-  randomly add species (equal to the number added in the observed
-  data) and allow the community to assembly using adaptive foraging
-\item  Keep track of
-  \begin{enumerate}
-  \item number of species that go extinct at each hedgerow
-  \item the specialization of each plant and pollinator
-  \end{enumerate}
-\end{enumerate}
-
-\section{Predictions}
-\begin{enumerate}
-\item ``generalist'' pollinators early in succession will become more
-  specialized
-\item ``generalist'' pollinators early in succession will visit a
-  subset of what they originally visited, mainly the originally
-  ``specialist'' plants
-\end{enumerate}
-
-\section{Considerations}
-\begin{enumerate}
-\item some plant species that were planted were never visited by
-  pollinators in the first few years. This is likely because they did
-  not produce many/any flowers/nectar resources
-\item compare observed data to a model with/without adaptive foraging?
-\item all the pollinator species needed (bees, syrphids, wasps
-  etc.?). Not all have been identified to species.
-\item weedy plant species ``added'' to communities by invasion, lost
-  by death
+\item Determine whether there are change points in the assmebly of the
+  ``maturing'' hedgerow communties, and compare those to the
+  ``mature'' and ``unrestored''
+\item Examine the resileicen of the maturing communities through time,
+  as well as the different site types
+\item Simulate the extinction of plants affected by drought, and see
+  if the predict extinct pollinator species that are 
+\item Assess whether (and how) species roles/importance are changing
+  through time, as community matures. 
 \end{enumerate}

### 2016-06-23 analysis/changePoint/cptPeel/changepointDetection.py

@@ -1,0 +1,420 @@
+"""change-pointDetection.py - detects change-points within sequences of graphs
+    Copyright (C) 2014 Leto Peel
+
+    This program is free software; you can redistribute it and/or modify
+    it under the terms of the GNU General Public License as published by
+    the Free Software Foundation; either version 2 of the License, or
+    (at your option) any later version.
+
+    This program is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU General Public License for more details.
+
+    You should have received a copy of the GNU General Public License
+    along with this program; if not, write to the Free Software
+    Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA"""
+
+from sys import stdout
+import dendropy
+import numpy as np
+import os
+from scipy.special import gammaln
+from multiprocessing import Pool
+from subprocess import call
+from scipy.stats import norm
+
+
+def unwrap_self_sampleNullModel(arg, **kwarg):
+    return sampleNullModel(*arg, **kwarg)
+
+
+class Filenames(object):
+    def __init__(self,consensusFilename,path,currentNormalNetworks,namesFile):
+        self.consensusFilename = consensusFilename
+        self.path = path
+        self.currentNormalNetworks = currentNormalNetworks
+        self.namesFile = namesFile
+    
+
+
+def sampleNullModel(filenames,label):
+    ad = anomalyDetection()
+    ad.importNames(filenames.namesFile)
+    ng=len(filenames.currentNormalNetworks)
+    normal_model = ad.createDendroTreeFromConsensus(filenames.consensusFilename)####
+    edgelist=[]
+    for networkFile in filenames.currentNormalNetworks:###
+        with open(os.path.join(filenames.path,networkFile)) as f:
+            edgelist.extend([edge.split() for edge in f.readlines()])
+    ad.calculatePvals(normal_model,edgelist,ng,setPriors=True)
+    
+    #generate sample graphs
+    G=[]
+    for gi in xrange(ng):
+        G.append(ad.generateGraph(normal_model))
+    
+    ad.calculatePvals(normal_model,[edge for g in G for edge in g],ng,setPriors=True)
+    
+    loglikelihoodR = -np.inf
[…]

### 2016-06-23 analysis/changePoint/cptPeel/consensusHRG_GPL_Bayes/dendro_eq.h

@@ -1,0 +1,1805 @@
+// ****************************************************************************************************
+// *** COPYRIGHT NOTICE *******************************************************************************
+// dendro_eq.h - hierarchical random graph (hrg) data structure
+// Copyright (C) 2006-2008 Aaron Clauset
+// Copyright (C) 2014 Leto Peel
+//
+// This program is free software; you can redistribute it and/or modify
+// it under the terms of the GNU General Public License as published by
+// the Free Software Foundation; either version 2 of the License, or
+// (at your option) any later version.
+//
+// This program is distributed in the hope that it will be useful,
+// but WITHOUT ANY WARRANTY; without even the implied warranty of
+// MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+// GNU General Public License for more details.
+//
+// You should have received a copy of the GNU General Public License
+// along with this program; if not, write to the Free Software
+// Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
+// 
+// See http://www.gnu.org/licenses/gpl.txt for more details.
+// 
+// ****************************************************************************************************
+// Author       : Aaron Clauset  ( aaronc@santafe.edu | http://www.santafe.edu/~aaronc/ )
+// Collaborators: Cristopher Moore and Mark E.J. Newman
+// Project      : Hierarchical Random Graphs
+// Location     : University of New Mexico, Dept. of Computer Science AND Santa Fe Institute
+// Created      : 19 April 2006
+// Modified     : 19 May 2007
+//			 : 19 May 2008 (cleaned up for public consumption)
+//
+// ****************************************************************************************************
+// 
+// Maximum likelihood dendrogram data structure. This is the heart of the HRG algorithm: all
+// manipulations are done here and all data is stored here. The data structure uses the separate
+// graph data structure to store the basic adjacency information (in a dangerously mutable way).
+//
+// Note: This version (dendro_eq.h) differs from other versions because it includes methods for 
+//       doing the consensus dendrogram calculation.
+// 
+// ****************************************************************************************************
+
+#if !defined(dendro_INCLUDED)
+#define dendro_INCLUDED
+
+#include <iostream>
+#include <fstream>
+#include <stdio.h>
+#include <math.h>
+
+#include "MersenneTwister.h"
+#include "graph.h"
+#include "rbtree.h"
[…]

### 2016-06-23 analysis/changePoint/nodesPeel/activelearn.py

@@ -1,0 +1,160 @@
+"""activelearn.py - module to run active discovery of network roles
+    Copyright (C) 2014 Leto Peel
+
+    This program is free software; you can redistribute it and/or modify
+    it under the terms of the GNU General Public License as published by
+    the Free Software Foundation; either version 2 of the License, or
+    (at your option) any later version.
+
+    This program is distributed in the hope that it will be useful,
+    but WITHOUT ANY WARRANTY; without even the implied warranty of
+    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+    GNU General Public License for more details.
+
+    You should have received a copy of the GNU General Public License
+    along with this program; if not, write to the Free Software
+    Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA"""
+
+import supervisedBM 
+from random import sample
+import numpy as np
+import os
+
+
+def expInstance(args):
+    np.random.seed()
+    networkfile,classlist,iters,no = args
+    exp=os.path.split(networkfile)[1].split(".")[0]
+    print exp
+    #~ print asdsad
+    #~ networkfile="../data/%s_edges.txt" % exp
+    #~ classlist ="../data/%s_labels.txt" % exp
+    inG = supervisedBM.nx.read_adjlist(networkfile, create_using=supervisedBM.nx.DiGraph())
+    
+    with open(classlist) as f:
+        labels = f.readlines()
+    class_lut=dict((label.split()[0],label.split()[1].strip()) for label in labels)
+    
+    #select one of each class
+    classes=np.unique(class_lut.values())
+    print classes
+    for c in classes:
+        nodes_of_class_c = [k for k,v in class_lut.iteritems() if v==c]
+        v=nodes_of_class_c[int(np.floor(np.random.rand()*len(nodes_of_class_c)))]
+        inG.node[v]["label"]=class_lut[v]
+    
+    
+    #~ sb = supervisedBM.MixedMembership(inG)
+    sb = supervisedBM.MaxBM(inG,K=(2*len(classes)))
+    sb.id=no
+    
+    for i in range(iters):
+        sb.infer()
+        
+        #X-validation step
+        maxacc=0
+        maxr=64
+        print "*************"
+        training_nodes = dict((node,nodeData["cindx"]) for node,nodeData in sb.G.nodes_iter(data=True) if nodeData["cindx"] is not None)
+        if len(training_nodes)>5:
[…]

### 2016-06-23 analysis/changePoint/nodesPeel/svm_struct_api.c

@@ -1,0 +1,630 @@
+/***********************************************************************/
+/*                                                                     */
+/*   svm_struct_api.c                                                  */
+/*                                                                     */
+/*   Definition of API for attaching implementing SVM learning of      */
+/*   structures (e.g. parsing, multi-label classification, HMM)        */ 
+/*                                                                     */
+/*   Author: Thorsten Joachims                                         */
+/*   Date: 03.07.04                                                    */
+/*                                                                     */
+/*   Copyright (c) 2004  Thorsten Joachims - All rights reserved       */
+/*                                                                     */
+/*   This software is available for non-commercial use only. It must   */
+/*   not be modified and distributed without prior permission of the   */
+/*   author. The author is not responsible for implications from the   */
+/*   use of this software.                                             */
+/*                                                                     */
+/***********************************************************************/
+
+#include <stdio.h>
+#include <string.h>
+#include "svm_struct/svm_struct_common.h"
+#include "svm_struct_api.h"
+
+void        svm_struct_learn_api_init(int argc, char* argv[])
+{
+  /* Called in learning part before anything else is done to allow
+     any initializations that might be necessary. */
+}
+
+void        svm_struct_learn_api_exit()
+{
+  /* Called in learning part at the very end to allow any clean-up
+     that might be necessary. */
+}
+
+void        svm_struct_classify_api_init(int argc, char* argv[])
+{
+  /* Called in prediction part before anything else is done to allow
+     any initializations that might be necessary. */
+}
+
+void        svm_struct_classify_api_exit()
+{
+  /* Called in prediction part at the very end to allow any clean-up
+     that might be necessary. */
+}
+
+SAMPLE      read_struct_examples(char *file, STRUCT_LEARN_PARM *sparm)
+{
[…]

### 2016-06-27 notes/Marillia/Jun23-2016/notes.tex
ran leto's analysis on Barger and Butler

@@
   \large {\bf The assembly of plant-pollinator communities in response
-  to restoration8} \\
+    to restoration} \\
 \end{centering}
@@
 \begin{enumerate}
-\item Determine whether there are change points in the assmebly of the
-  ``maturing'' hedgerow communties, and compare those to the
-  ``mature'' and ``unrestored''
-\item Examine the resileicen of the maturing communities through time,
+\item Do the maturing hedgerows undergo a change point during
+  assembly? Is that change point associated with drought? Do the
+  ``controls'', unrestored and mature hedgerows undergo a change point
+  in responce to drought similar to the maturing hedgerows?
+\item Even if the network isn't shifting drastically, are the roles of
+  species changing? Assess whether (and how) species roles/importance
+  are changing through time, as community matures.
+\item Are the maturing hedgerows getting more resilient through
+  time? Examine the resilience of the maturing communities through time,
   as well as the different site types
-\item Simulate the extinction of plants affected by drought, and see
-  if the predict extinct pollinator species that are 
-\item Assess whether (and how) species roles/importance are changing
-  through time, as community matures. 
+\item can we predict the fate of pollinator species to ``exctinction''
+  of their plants in response to drought? Simulate the extinction of
+  plants affected by drought, and see if the predict extinct
+  pollinator species that are
 \end{enumerate}

### 2016-06-27 analysis/changePoint/cptPeel/changepointDetection.py

@@ -317,7 +317,7 @@ class anomalyDetection(object):
     
     def mapDendro(self,tree,G):
         print "MAP"
-        pdm = treemeasure.PatristicDistanceMatrix(tree)
+        pdm = treemeasure.PatrisiticDistanceMatrix(tree)
         
         tns = tree.taxon_namespace
         
@@ -384,7 +384,7 @@ class anomalyDetection(object):
         st=time.time()
         #~ print "Pvals"
         
-        pdm = treemeasure.PatristicDistanceMatrix(tree)
+        pdm = treemeasure.PatrisiticDistanceMatrix(tree)
         
         tns = tree.taxon_namespace
         

### 2016-07-07 ms/ms.tex
added abstract

\documentclass[12pt]{article}
\usepackage[pdftex]{graphicx}
\usepackage[numbers]{natbib}
% \usepackage{natbib}
\usepackage{color}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{verbatim}
\usepackage{mathpazo}
\usepackage{setspace}
\usepackage{multirow}
\usepackage{fullpage}
\usepackage{lscape}
\usepackage{array}
\usepackage{paralist}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{times}
\usepackage{textcomp}
\usepackage{xr}
\usepackage{etoolbox}
\usepackage{filecontents}
\usepackage{url}
\externaldocument{SI}

\makeatletter
\def\@biblabel#1{#1.\\}
\makeatother

\bibliographystyle{gcb}

\parindent=2em
\setlength{\parskip}{1ex}

\RequirePackage{lineno}

\def\title{The assembly of plant-pollinator networks}

\def\author{Lauren C.\ Ponisio$^{1}$, Marilia P.\ Gaiarsa$^2$
  and Asher Mullokandov$^3$}

\def\runninghead{Temporal network assembly}

\def\affiliation{
  \begin{enumerate}
  \item Department of Environmental Science, Policy, and Management\\
    University of California, Berkeley\\
    130 Mulford Hall\\
    Berkeley, California, USA\\
    94720\\
  \item Departamento de Ecologia\\
    Universidade de Sao Paulo\\
    Sao Paulo, SP, Brazil\\
    05508-900\\
  \item Department of Physics\\
    Boston University\\
    Boston, MA, USA\\
    02215\\
  \end{enumerate}
}

\newcommand{\mstitlepage}{
  \paragraph{Running head:} \textsc{\runninghead}
  % \parindent=0pt
  \begin{center}%
    {\LARGE \title \par}%
    \vskip 3em%
    {\large
      \lineskip .75em%
      \begin{tabular}[t]{c}%
        \author
      \end{tabular}\par}%
    \vskip 1.5em%
  \end{center}\par
  \affiliation
}
\clearpage

\begin{document}

\mstitlepage
\doublespacing
\linenumbers
\clearpage

\begin{abstract}
  The structure of networks is related to ability of communities to
  maintain function in the face of species extinction. Understanding
  network structure and how it relates to network disassembly,
  therefore, is a priority for system-level conservation biology.  We
  explore the assembly of plant-pollinator communities on native plant
  restorations in the Central Valley of California. The assembling
  communities are paired with unrestored field margins (controls) and
  mature (non-assembling) hedgerows. We determine whether there are
  change points in the assembly of the communities where the network
  undergoes significant reorganization. We are also ask how are the
  individual species changing their interaction patterns? What does
  this mean for the topology/resilience of the network? We also
  attempted to adapt a financial model to mutualistic networks. Our
  biggest difficulty with this approach was to translate the price
  term to mutualistic systems. We explored a range of approaches, such
  as number of visits a species performs. However, it seems that
  financial systems cannot be easily translated to mutualistic
  systems. In addition, we used a Changing Point Detection Algorithm
  to assess weather the different communities went through a critical
  reorganization on their interaction patterns. We were able to
  identify some changing points in the communities, and also to
  explore some general patterns commonly used to describe ecological
  networks. For example, on the network level, networks become
  increasingly modular and less nested, whereas on the species level,
  species become more specialized, as resources become more reliable.
\end{abstract}

Keywords: pollinators, community assembly, networks, changing points,
temporal networks
 \clearpage

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-PDF-mode: t
%%% End:

### 2016-07-23 ms/ms.tex
started ms

@@
 \usepackage[pdftex]{graphicx}
-\usepackage[numbers]{natbib}
-% \usepackage{natbib}
+\usepackage{natbib} 
 \usepackage{color}
@@
 \usepackage{lscape}
-\usepackage{array}
+\usepackage{fancyhdr}
+\usepackage[normalem]{ulem} 
+\usepackage{hyperref}
+\usepackage[parfill]{parskip}
+\usepackage{xr}
 \usepackage{paralist}
-\usepackage{fancyhdr}
-\usepackage{graphicx}
-\usepackage{longtable}
-\usepackage{times}
-\usepackage{textcomp}
 \usepackage{xr}
-\usepackage{etoolbox}
-\usepackage{filecontents}
-\usepackage{url}
 \externaldocument{SI}
-
-\makeatletter
-\def\@biblabel#1{#1.\\}
-\makeatother
-
-\bibliographystyle{gcb}
-
-\parindent=2em
-\setlength{\parskip}{1ex}
-
+\hypersetup{colorlinks=true, linkcolor=black, citecolor=black}
 \RequirePackage{lineno}
 
-\def\title{The assembly of plant-pollinator networks}
+\newcommand{\flagged}[1] {
+  \textcolor{blue}{#1}
+}
 
-\def\author{Lauren C.\ Ponisio$^{1}$, Marilia P.\ Gaiarsa$^2$
-  and Asher Mullokandov$^3$}
+\def\title{The temporal assembly of plant-pollinator networks
+  following restoration}
+\def\author{Lauren C.\ Ponisio$^1$, Marilia P. Gaiarsa$^2$, Claire Kremen$^1$}
 
-\def\runninghead{Temporal network assembly}
+\def\runninghead{Coevolution and network structure}
+\def\keywords{evolution, nestedness, modularity, bipartite}
+
+\def\extras{
+  \begin{itemize}
+  \item Submitted as a Letter
+  \item Abstract word count: 
+  \item Main text word count: 
+  \item Number of references: 
+  \item Number of figures:
+  \end{itemize}
+}
 
@@
     05508-900\\
-  \item Department of Physics\\
-    Boston University\\
-    Boston, MA, USA\\
-    02215\\
   \end{enumerate}
@@
 
-Keywords: pollinators, community assembly, networks, changing points,
-temporal networks
+Keywords: changing points, temporal networks, hedgerows, species
+interactions
+
 \clearpage
+
+\section*{Introduction}
+\label{sec:introduction}
+\begin{itemize}
+\item The structure of networks is related to ability of communities
+  to maintain function in the face of species extinction.
+\item A key restoration aim is to facilitate assembly of robust
+  networks; thus it is critical to study how restoration
+  influences the assembly of plant-pollinator interactions.
+\item few theories about how networks assemble, preferential
+  attachment 
+\item To date, only two field studies have examined how networks
+  assemble over time, often using space for time gradients.
+\item \cite{Olesen2008} was investigated day-to-day, temporal assmebly
+  of a plant-pollinator network within a season, taking advantage of
+  the extreme seasonality of pollinator communities in Greenland.
+  \cite{Olesen2008} found that within a season, the network assembly
+  was similar to preferential attachment. New species tended to
+  interact predominantly with already well-connected species, likely
+  because these species are either more abundant or more temporally
+  persistent.
+\item Studying primary succession along a glacier foreland,
+  \cite{albrecht2010plant} found a similar pattern where nestedness, a
+  pattern of interactions where a generalist core interacts with both
+  specialist and generalist species, increased as the community
+  aged. % Studying ``managed succession'' of a clear-cut pine forest,
+  % \cite{devoto2012understanding} found changes in network structure
+  % were explained by a combination of age, tree density and variation
+  % in tree diameter. 
+\item Even non-succesional temporal dyanmics suggest a stable core of
+  generalsits persiss despite high turnover of peripheral species
+  \citep{fang2012relative, diaz2010changes, alarcon2008year}.
+\item In contrast to the ordered notion of netowrk build-up described
+  by perferential attachment, assembly may be punctated by significant
+  reorganizations of interactions. For example, as new species are
+  added, resident species may change their interaction parteners to
+  minimize competition, or go extinct if they are unable to persist
+  with more competing species.
+\end{itemize}
+
+\section*{Materials \& Methods}
+\label{sec:methods}
+
+\subsection*{Study sites and collection methods}
+\label{sec:study-sites}
+
+\section*{Results}
+\label{sec:results}
+
+\section*{Discussion}
+\label{sec:discussion}
+
+\section*{Acknowledgments}
+\label{sec:acknowledge}
+
+We would like to thank Leto Peel and Aaron Clauset for their
+invaluable discussions and for help with the change point analysis.
+We thank the growers and land owners that allowed us to work on their
+property.  We also appreciate the identification assistance of expert
+taxonomists Robbin Thorp and Jason Gibbs.  This work was supported by
+funding from the Army Research Office (W911NF-11-1-0361 to CK), the
+Natural Resources Conservation Service (CIG-69-3A75-12-253,
+CIG-69-3A75-9-142, CIG-68-9104-6-101 and WLF-69-7482-6-277 to The
+Xerces Society), the National Science Foundation (DEB-0919128 to CK),
+The U.S.  Department of Agriculture (USDA-NIFA 2012-51181-20105 to
+Michigan State University).  Funding for LCP was provided by an NSF
+Graduate Research Fellowship and the USDA NIFA Graduate
+Fellowship. FUNING FOR MARILLIA.
+
+\bibliographystyle{ecol_let}
+
+\bibliography{refs}
 

### 2016-07-25 ms/ms.tex
added cv occurence

@@
   was similar to preferential attachment. New species tended to
-  interact predominantly with already well-connected species, likely
-  because these species are either more abundant or more temporally
-  persistent.
+  interact with already well-connected species, likely because these
+  species are either more abundant or more temporally persistent.
 \item Studying primary succession along a glacier foreland,
@@
   \citep{fang2012relative, diaz2010changes, alarcon2008year}.
-\item In contrast to the ordered notion of netowrk build-up described
-  by perferential attachment, assembly may be punctated by significant
+\item In contrast to the ordered netowrk build-up described by
+  perferential attachment, assembly may be punctated by significant
   reorganizations of interactions. For example, as new species are
-  added, resident species may change their interaction parteners to
-  minimize competition, or go extinct if they are unable to persist
-  with more competing species.
+  added, resident species change their interaction parteners to
+  minimize competition, or become extinct.
 \end{itemize}
@@
 property.  We also appreciate the identification assistance of expert
-taxonomists Robbin Thorp and Jason Gibbs.  This work was supported by
-funding from the Army Research Office (W911NF-11-1-0361 to CK), the
-Natural Resources Conservation Service (CIG-69-3A75-12-253,
-CIG-69-3A75-9-142, CIG-68-9104-6-101 and WLF-69-7482-6-277 to The
-Xerces Society), the National Science Foundation (DEB-0919128 to CK),
-The U.S.  Department of Agriculture (USDA-NIFA 2012-51181-20105 to
-Michigan State University).  Funding for LCP was provided by an NSF
-Graduate Research Fellowship and the USDA NIFA Graduate
-Fellowship. FUNING FOR MARILLIA.
+taxonomists Martin Hauser, Robbin Thorp and Jason Gibbs.  This work
+was supported by funding from the Army Research Office
+(W911NF-11-1-0361 to CK), the Natural Resources Conservation Service
+(CIG-69-3A75-12-253, CIG-69-3A75-9-142, CIG-68-9104-6-101 and
+WLF-69-7482-6-277 to The Xerces Society), the National Science
+Foundation (DEB-0919128 to CK), The U.S.  Department of Agriculture
+(USDA-NIFA 2012-51181-20105 to Michigan State University).  Funding
+for LCP was provided by an NSF Graduate Research Fellowship and the
+USDA NIFA Graduate Fellowship. FUNING FOR MARILLIA.
 

### 2016-07-25 analysis/variability/cv.R

@@ -43,12 +43,26 @@ dprime.closeness.sd <- cv.trait(spec,
                                 standard.cv=FALSE,
                                 na.rm=TRUE)
 
+## sig
 occ.k.sd <- cv.trait(spec,
                      specs[specs$speciesType =="pollinator",],
                      trait="occ.date",
                      method= "time", time.col="assem",
                      abund.col="k",
-                     cv.function=sd, zero2na=TRUE, standard.cv=FALSE,
+                     cv.function=sd,
+                     zero2na=TRUE,
+                     standard.cv=FALSE,
+                     na.rm=TRUE)
+
+## not sig
+occ.k.cv <- cv.trait(spec,
+                     specs[specs$speciesType =="pollinator",],
+                     trait="occ.date",
+                     method= "time", time.col="assem",
+                     abund.col="k",
+                     cv.function=cv,
+                     zero2na=TRUE,
+                     standard.cv=FALSE,
                      na.rm=TRUE)
 
 

### 2016-07-25 analysis/variability/src/cvCalc.R

@@ -4,6 +4,9 @@ corCv <- function(x){
   cv(x)*(1 + (1/4*length(x)))
 }
 
+
+## standard cv devides cv by 100 to avoid too large of numbers for the
+## lm
 cv.trait <- function(spec.dat,
                      byType,
                      trait,

### 2016-07-26 ms/ms.tex
added stuff

@@
 \item few theories about how networks assemble, preferential
-  attachment 
+  attachmen. According to this model, a new species is more likely to
+  interact with species that are already well-connected (''the
+  rich-get-richer'' principle).
 \item To date, only two field studies have examined how networks
@@
   \citep{fang2012relative, diaz2010changes, alarcon2008year}.
-\item In contrast to the ordered netowrk build-up described by
+\item In contrast to the ordered network build-up described by
   perferential attachment, assembly may be punctated by significant
@@
   added, resident species change their interaction parteners to
-  minimize competition, or become extinct.
-\end{itemize}
+  minimize competition, or become extinct. Such significant
+  reorganizations of interactions, or changing points, have been
+  observed in networks \citep{peel2014detecting}.
 

### 2016-07-26 analysis/pa/pa.R

@@ -1,0 +1,42 @@
+rm(list=ls())
+setwd('~/Dropbox/hedgerow_assembly/analysis/pa')
+
+load('../../data/networks/allSpecimens.Rdata')
+f.path <- "../changePoint/cptPeel/baci"
+load(file=file.path(f.path, "graphs.Rdata"))
+
+## total interactions
+plant.tot.int <- lapply(nets, rowSums)
+pol.tot.int <- lapply(nets, colSums)
+
+## degree
+plant.degree <- lapply(nets, function(x)
+                       apply(x, 1, function(x) sum(x >= 1)))
+                      
+pol.degree <- lapply(nets, function(x)
+                       apply(x, 2, function(x) sum(x >= 1)))
+## ypr 1 
+first.yrs <- spec[spec$ypr == 1,]
+first.yrs <- unique(data.frame(Site=first.yrs$Site,
+                  Year=first.yrs$Year))
+first.yrs <- first.yrs[!apply(first.yrs, 1, function(x) any(is.na(x))),]
+
+ypr1 <- names(nets) %in%
+                    paste(first.yrs$Site, first.yrs$Year, sep=".")
+
+
+sites <- sapply(strsplit(names(nets), "[.]"), function(x) x[1])
+years <- sapply(strsplit(names(nets), "[.]"), function(x) x[2])
+
+N <- 2
+lapply(unique(sites), function(x){
+  this.net <- nets[sites == x]
+  pres.sp <- lapply(this.net, function(y){
+    y[y > 1] <- 1
+    y2 <- y
+    y2[rowSums(y) != 0,] <- 1
+    y2[,colSums(y) != 0] <- 1
+    return(y2)
+  })
+  browser()
+})

### 2016-08-03 ms/ms.tex
changed pcoa analysis to be by change point, edited ms

--- 2016-07-25 ms/ms.tex
@@
 Keywords: changing points, temporal networks, hedgerows, species
-interactions
+interactions, preferential attachment
 
@@
   influences the assembly of plant-pollinator interactions.
-\item few theories about how networks assemble, preferential
-  attachment 
+\item In general, however, few mechanisms of network assembly have
+  been developed and analyzed.
+\item The mostly widely explored mechanism of network assembly,
+  preferential attachment \citep{barabasi1999emergence}, predicts that
+  a new species (or node more generally) is more likely to interact
+  with species that are already well-connected \citep[''the
+  rich-get-richer'' principle,][]{barabasi1999emergence}.
 \item To date, only two field studies have examined how networks
   assemble over time, often using space for time gradients.
-\item \cite{Olesen2008} was investigated day-to-day, temporal assmebly
+\item \cite{Olesen2008} was investigated day-to-day, temporal assembly
   of a plant-pollinator network within a season, taking advantage of
@@
   % in tree diameter. 
-\item Even non-succesional temporal dyanmics suggest a stable core of
-  generalsits persiss despite high turnover of peripheral species
+\item Even non-successional temporal dynamics suggest a stable core of
+  generalists persist despite high turnover of peripheral species
   \citep{fang2012relative, diaz2010changes, alarcon2008year}.
-\item In contrast to the ordered netowrk build-up described by
-  perferential attachment, assembly may be punctated by significant
+\item In contrast to the ordered network build-up described by
+  preferential attachment, assembly may be punctuated by significant
   reorganizations of interactions. For example, as new species are
-  added, resident species change their interaction parteners to
-  minimize competition, or become extinct.
+  added, resident species change their interaction partners to
+  minimize competition, or become extinct. Such significant
+  reorganizations of interactions, or changing points, have been
+  observed in networks \citep{peel2014detecting}.
 \end{itemize}
@@
 \label{sec:study-sites}
+
+We surveyed plant-pollinator interaction networks of assembling
+hedgerows (N=5), as well as in two types of non-assembling communities
+to serve as controls: unrestored, weedy field margins (N=XX) and
+established hedgerows (greater than 10 years since planting,
+N=XX). The sites were located in the Central Valley of California in
+Yolo, Colusa and Solano Counties. This area is comprised of
+intensively managed agriculture -- primarily monocultures of
+conventional row crops, vineyards and orchards. Hedgerows we planted
+along field margins where they do not remove valuable land from
+production, and are ca. 3--6m wide and approximately 350m long and
+border large (ca.\ 30--hectare) crop fields. Hedgerows consist of
+native, perennial, shrub and tree plantings including \textit{Rosa
+  californica}, \textit{Cercis occidentalis}, \textit{Ceanothus spp.},
+\textit{Heteromeles arbutifolia}, \textit{Sambucus mexicana},
+\textit{Eriogonum spp.}, \textit{Baccharis spp.}, \textit{Salvia
+  spp}. and others. The mean distance between monitoring sites was 15
+km, and the minimum distance between sites of the same type sampled in
+the same year was 2 km.  The entire area surveyed spanned almost 300
+km$^2$. The crop fields adjacent to all sites were similarly managed
+as intensive, high-input monoculture.
+
+Monitoring began in 2006 and continued through 2015. Sampling of the
+assembling hedgerows began the year before the area was restored. For
+logistical reasons, no sampling of assembling hedgerows was conducted
+in 2010. Sites were sampled between two and five times per year (Table
+XX). In each round of sampling, the order in which sites were sampled
+was randomized. Surveys were conducted under sunny conditions when the
+temperature was above $21^{\circ}\mathrm{C}$ and wind speed was below
+$2.5$ meters/second.
+
+Flower-visitors to plants in hedgerows and unrestored controls were
+netted for one hour of active search time (the timer was paused when
+handling specimens). Honeybees (\textit{Apis mellifera}) were not
+collected because their abundance is determined largely by the
+placement of hives throughout the region by bee-keepers. All other
+insect flower visitors that touched the reproductive parts of the
+flower were collected; however, here we focus only on wild bees and
+syrphids (representing XX and XX percent of records, respectively),
+the most abundant and effective pollinators in the system (C. Kremen,
+A. Klein and L. Morandin, unpublished data). Bee and syrphid specimens
+were identified to species (or morpho-species for some bee specimens
+in the genera \textit{Nomada} and \textit{Sphecodes}) by expert
+taxonomists.
+
+\subsection*{Change point analysis}
+
+We employed a change point detection method \citep{peel2014detecting}
+to identify fundamental changes in large-scale pattern of interactions
+of plants and pollinators. A change point is caused by a merge, split,
+fragmentation or formation of communities (also called modules or
+compartments). Following \cite{peel2014detecting}, we first defined a
+probability distribution over the networks using the generalized
+hierarchical random graph model (GHRG). The GHRG model is able to
+capture both assortative and disassortative community structure
+patterns at all scales in the network \citep{peel2014detecting}. A
+network $G$ is composed of vertices $V$ and edges $E \subseteq {V × V
+}$. The GHRG model decomposes the $N$ vertices into a series of nested
+groups, the relationships among which are represented by the
+dendrogram $T$.  The tips of $T$ are the vertices of $G$, and the
+probability that two vertices $u$ and $v$ connect is given by the
+parameter $p_r$. The probability distribution of the network $G$ thus
+defined as:
+
+\begin{equation}
+    \label{eq:lik}
+    P(G|T,{pr}) = p_r^{E_r}(1-p_r)^{N_r-E_r}
+\end{equation}
+% 
+Where $E_r$ is the observed number of edges between vertices with the
+common ancestor $r$, and $N_r$ is the total possible edges.
+
+Using Bayesian posterior inference and techniques from phylogenetic
+tree reconstruction, we fit the GHRG model to the networks
+\citep{peel2014detecting}. This is accomplished by using a Markov
+Chain Monte Carlo (MCMC) procedure to first sample the posterior
+distribution of bipartitions, from which a consensus tree is derived
+\citep{peel2014detecting}. ADD NEWMAN PAPER. We used $\beta$
+distributions with the hyperparameters $\alpha=\beta=1$ to define
+priors for $p_r$.
+
+Once the GHRG model has been fit to the networks, we determine whether
+a change point occurred between two time slices. To detect a change
+point, we compare the fir of two models -- one where a change point
+had occurred between two networks, and one where no change occurred --
+using posterior Bayes factors. We chose a sliding window of length,
+$w$, of four, within which to find change points. Larger windows allow
+for more gradual changes, and four was the maximum possible with our
+maximum of nine years of data. Lastly, to calculate a $p$-value for
+the Bayes factors, we use parametric bootstrapping to numerically
+estimate the null distribution \citep{peel2014detecting}. The change
+point analysis was carried out using code published online by L~.Peel.
+
+We next test whether the change points occurring in maturing hedgerows
+were a component of the assembly process or a product of environmental
+shifts that lead to network reorganizations in all types of
+communities. We model the number of change points as successes and the
+total number of years each site was sampled as trails, and use a
+generalized linear model with Binomial error to test whether the
+probability of a change point occurring varied by site type.
+
+To further elucidate the nature of the change points, we examine the
+characteristics of the species that contributed the reorganization of
+interactions. To do so, we first generate dendrograms using the GHRG
+model before and after each change point.  We then determined which
+species belonged to the ``core'' and ``peripheral'' network
+communities at each site. The ``core'' network communities contain the
+majority of species and are more basil than the more derived, less
+specious ``peripheral'' network communities. We next use a
+Permutational Multivariate Analysis of Variance (PERMANOVA)
+\citep{anderson-2013-557} to determine whether the species
+compositions of the species that belonged to the core and peripheral
+communities differed. We also test whether core or peripheral species
+had more variability in their species compositions \citep[i.e.,
+multivariate dispersion, ][]{anderson-2011-19,
+  anderson-2006-683}.
+
+%% keep trait analysis?
+% Lastly, we explored the functional diversity of the core and
+% peripheral pollinator species. The syrphids in our study area have
+% similar traits, so we focused on the trait diversity of the bee subset
+% of our interaction networks. We focus on resource capture and use
+% traits that collectively influence the distribution of bee species as
+% pollinators over space and time \citep{kremen-2015-602} including
+% resource specialization \citep[quantitative,
+% $d'$;][]{bluthgen-2006-9}, body size \citep[quantitative,
+% inter-tegular span, mm,][]{cane-1987-145}, sociality (categorical:
+% eusocial, solitary, cleptoparasitic), nest location (categorical:
+% above ground, below ground or mix), and nest construction
+% \citep[categorical: excavate or rent;][]{williams-2010-2280} as
+% described in more detail in \cite{kremen-2015-602}.  Each trait has
+% the same weight in trait diversity metric estimation
+% \citep{villeger-2008-2290, schleuter-2010-469}.  Pollinator
+% specialization was calculated using plant-pollinator interaction
+% observations from a more extensive data-set from Yolo County (18000
+% interaction records) that included both the data included in this
+% study and additional data from sites where we collected flower
+% visitors using the same methods \citep{mgonigle-2015-x}.  The
+% specialization metric measures the deviation of the observed
+% interaction frequency between a plant and pollinator from a null
+% expectation where all partners interact in proportion to their
+% abundances \citep{bluthgen-2006-9}.  It ranges from 0 for generalist
+% species to 1 for specialist species.
+
+% To characterize the functional diversity of the bee species in core
+% and peripherial communities, we computed three metrics that capture
+% diversity, uniqueness, and distribution of functional values in the
+% community: functional dispersion, divergence, and evenness
+% \citep{villeger-2008-2290, schleuter-2010-469}.  Functional
+% dispersion is a measure of trait diversity, corrected for species
+% richness \citep{schleuter-2010-469}; functional divergence measures
+% how species abundances are distributed within the trait space
+% \citep[i.e., a measure of functional
+% uniqueness,][]{villeger-2008-2290}; functional evenness measures the
+% regularity with which traits are distributed across functional
+% space, accounting for abundance \citep{villeger-2008-2290}.  In
+% combination, these metrics provide a relatively complete overview of
+% the different aspects of functional diversity
+% \citep{villeger-2008-2290, schleuter-2010-469}. To determine whether
+% trait evenness, dispersion, and divergence differed between controls
+% and hedgerows at different stages of maturation, we used the trait
+% diversity metrics as response variables in linear mixed models with
+% site type as a fixed effect and year and site as random effects
+% \citep{lme4,lmetest}.
+
+\subsection*{Network structure}
+
+The changing points in network structure may contribute to the
+reorganization of the assembling networks into predictable interaction
+patterns. Pollination networks exhibit two main topologies ---
+modularity \citep[e.g.,][]{Olesen2007} and nestedness
+\citep[e.g.,][]{Bascompte2006, Bascompte2003}.  Modular community
+interactions are insular, occurring within separate groups or
+``modules'' more often than between modules. Communities in the
+network may fragment as the network assembles, enhancing
+modularity. Conversely, nested networks are like pyramid of
+interactions, where there are some species that interact with many
+species, other species that interact with a subset of those species,
+and so on. If species entering the network tend to interact with the
+generalist base of the network pyramid, nestedness would increase
+through time. Lastly, if the network is accumulating specialist
+species or species are beginning to limit their interaction niche
+breath as the network assembles, this would lead to an increase in the
+network-level specialization \citep{bluthgen-2006-9}. To test whether
+network modularity, nestedness or specialization changed linearly with
+assembly, we used linear mixed models with the descriptive network
+metrics as the response variable, year of assembly as the explanatory
+variable, and random effects of site and year.
+
+\subsection*{Network robustness}
+Lastly, we test the changes in interaction patterns associated with
+network assembly affect the robustness of the network to species
+loss. Following \cite{Memmott2004}, we simulate the extinction of
+plant species the subsequent extinction cascades of pollinator
+species. Because the reproduction of plant species if facilitated by
+active restoration efforts, it is unlikely the extinction of
+pollinators would effect plant populations in the hedgerows. However,
+plants ceasing to bloom, for example in response to drought, will
+likely effect the pollinators that depend on them. Plants species were
+eliminated based on their degree or abundance, and the number of
+pollinators that secondarily went extinct is calculated. The area
+below the extinction curve is used as a measure of network robustness.
 
@@
 for LCP was provided by an NSF Graduate Research Fellowship and the
-USDA NIFA Graduate Fellowship. FUNING FOR MARILLIA.
+USDA NIFA Graduate Fellowship. FUNDING FOR MARILLIA.
 

### 2016-08-03 analysis/changePoint/comparisons.R

@@ -3,8 +3,38 @@ setwd("~/Dropbox/hedgerow_assembly/analysis/changePoint")
 load('cptPeel/baci/graphs.Rdata')
 load('../../data/networks/allSpecimens.Rdata')
 library(MASS)
+library(nlme)
 dats <- read.csv('cptPeel/changing_points.csv')
+BACI.site <- c('Barger', 'Butler', 'Hrdy', 'MullerB', 'Sperandio')
+## **********************************************************
+## binomial 
+## **********************************************************
+count.samples <- aggregate(spec$Year, list(Site=spec$Site),
+                           FUN=function(x) length(unique(x)))
+count.samples$x <- count.samples$x - 1
 
+change.points.site <- tapply(dats$cp, dats$sites, length)
+
+chpt.trial <- data.frame(chpts=change.points.site,
+                         trials=count.samples$x[match(
+                           rownames(change.points.site),
+                           count.samples$Site)])
+
+chpt.trial$status <- spec$SiteStatus[match(rownames(chpt.trial), spec$Site)]
+chpt.trial$status[rownames(chpt.trial) %in% BACI.site] <- "maturing"
+
+## binomial model with change points as successes
+mod.chpt <- glm(cbind(chpt.trial$chpts, chpt.trial$trials - chpt.trial$chpts) ~
+    chpt.trial$status, family="binomial")
+summary(mod.chpt)
+
+## maturing has more successes than mature and controls, and mature
+## and controls have about the same
+
+## **********************************************************
+## poisson likelihood
+## maybe not quite right given the number of trails differs
+## **********************************************************
 counts <- table(dats[,-3])
 
 chpts.yrs <- colSums(counts)
@@ -21,7 +51,6 @@ counts <- rbind(counts, add.chpt)
 chpts.sites <- rowSums(counts)
 
 statuses <- spec$SiteStatus[match(names(chpts.sites), spec$Site)]
-BACI.site <- c('Barger', 'Butler', 'Hrdy', 'MullerB', 'Sperandio')
 statuses[names(chpts.sites) %in% BACI.site] <- "maturing"
 
 layout(matrix(1:3, nrow=1))

### 2016-08-03 analysis/changePoint/cptPeel/convertfiles.py

@@ -11,7 +11,6 @@ cc.writeToFile('baci/Butler_2009_joint_s1-consensus.tree','gml')
 cc.writeToFile('baci/Butler_2015_joint_s1-consensus.tree','gml')
 
 ## three change points
-
 cc.writeToFile('baci/Hrdy_2007_joint_s1-consensus.tree','gml')
 cc.writeToFile('baci/Hrdy_2009_joint_s1-consensus.tree','gml')
 cc.writeToFile('baci/Hrdy_2011_joint_s1-consensus.tree','gml')

### 2016-08-03 analysis/changePoint/plotting/src/initialize.R

@@ -2,10 +2,13 @@ library(ape)
 library(vegan)
 library(FD)
 library(igraph)
+library(lme4)
+library(lmerTest)
 source('../networkLevel/src/misc.R')
 source('plotting/src/plotPcoa.R')
 load('../../data/networks/allSpecimens.Rdata')
 source('plotting/src/communities.R')
+source('src/calcCore.R')
 
 fig.path <- 'plotting/figures'
 f.path <- "cptPeel/baci"

### 2016-08-03 analysis/changePoint/plotting/src/plotPcoa.R

@@ -1,75 +1,13 @@
 
-getSpecies <- function(out, names.v, num.v){
-  out$names <- names.v[match(out$label, num.v)]
-  memb <- apply(out[, -c(1, ncol(out))], 2, table)
-  ## which are thecommunity names of the "cores"
-  core <- sapply(memb, function(x) names(x)[x == max(x)])
-  stayed <- list()
-  ## what are the species that are in the core
-  for(i in 1:length(core)){
-    stayed[[i]] <- out[, i+1] == core[i]
-  }
-  ## which species were always in the core
-  ind.stayed <- Reduce("*", stayed)
-  ## always in the core
-  stayed.core <- out$names[ind.stayed == 1]
-  ## left the core at least once
-  left.core <- out$names[ind.stayed == 0]
-  return(list(stayed=stayed.core, left=left.core))
-}
-
-
-makeCores <- function(site.nodes, l.path, names.v, num.v){
-  site.cores <- list()
-  for(i in 1:length(site.nodes)){
-    load(file.path(l.path, site.nodes[i]))
-    site.cores[[i]] <- getSpecies(out, names.v, num.v)
-  }
-
-  ## create data structure
-  names(site.cores) <- unique(sites.trees)
-  site.cores <- unlist(site.cores, recursive=FALSE)
-  num.sp <- sapply(site.cores, length)
-
-  site.cores <- as.data.frame(unlist(site.cores))
-  rownames(site.cores) <- NULL
-  colnames(site.cores) <- "GenusSpecies"
-  site.cores$SiteStat <- rep(names(num.sp), num.sp)
-  site.cores$speciesType <- ifelse(site.cores$GenusSpecies %in% 
-                                   unique(spec$GenusSpecies),
-                                   "pollinator", "plant")
-  site.cores$Site <- sapply(strsplit(site.cores$SiteStat, "[.]"), 
-                            function(x) x[1])
-  site.cores$comm <- sapply(strsplit(site.cores$SiteStat, "[.]"),
-                            function(x) x[2])
-
-  site.cores$Abund <- 1
-  ## split by species type
-  site.cores <- split(site.cores, site.cores$speciesType)
-  
-  return(site.cores)
-}
-
-## create community matrics
-
-makeComm <- function(site.core){
-  comm.mat <-  samp2site.spp(site.core$SiteStat,
-                             site.core$GenusSpecies,
-                             site.core$Abund)
-
[…]

### 2016-08-03 analysis/changePoint/src/calcCore.R

@@ -1,0 +1,82 @@
+
+getSpecies <- function(out, names.v, num.v){
+  out$names <- names.v[match(out$label, num.v)]
+  memb <- apply(out[, -c(1, ncol(out))], 2, table)
+  ## which are thecommunity names of the "cores"
+  core <- sapply(memb, function(x) names(x)[x == max(x)])
+  stayed <- list()
+  ## what are the species that are in the core
+  for(i in 1:length(core)){
+    stayed[[i]] <- out[, i+1] == core[i]
+  }
+  ## which species were always in the core
+  ind.stayed <- Reduce("*", stayed)
+  ## always in the core
+  stayed.core <- out$names[ind.stayed == 1]
+  ## left the core at least once
+  left.core <- out$names[ind.stayed == 0]
+  ## for each year, who was in the core
+  in.core <- lapply(stayed, function(x) out$names[x])
+  ## for each year, who was in the periphery
+  in.periphery <- lapply(stayed, function(x) out$names[!x])
+  return(list(across.yrs=list(stayed=stayed.core, left=left.core),
+              by.yr=list(core=in.core, periphery=in.periphery)))
+}
+
+prepDat <- function(sp.core, spec, by.year=FALSE){
+  if(by.year){
+    num.sp <- rapply(sp.core, length, how="unlist")
+  }else{
+    num.sp <- sapply(sp.core, length)
+  }
+  sp.core <- as.data.frame(unlist(sp.core))
+  rownames(sp.core) <- NULL
+  colnames(sp.core) <- "GenusSpecies"
+  sp.core$SiteStat <- rep(names(num.sp), num.sp)
+  sp.core$speciesType <- ifelse(sp.core$GenusSpecies %in% 
+                                unique(spec$GenusSpecies),
+                                "pollinator", "plant")
+  sp.core$Site <- sapply(strsplit(sp.core$SiteStat, "[.]"), 
+                         function(x) x[1])
+  sp.core$comm <- sapply(strsplit(sp.core$SiteStat, "[.]"),
+                         function(x) x[2])
+  if(by.year){
+    sp.core$ChangePoint <- gsub("[a-z]", "", sp.core$comm)
+    sp.core$comm <- sub("[0-9]", "", sp.core$comm)
+  }
+  sp.core$Abund <- 1
+  ## split by species type
+  sp.core <- split(sp.core, sp.core$speciesType)
+  return(sp.core)
+}
+
+makeCores <- function(site.nodes, l.path, names.v, num.v){
[…]

### 2016-08-03 ms/ms.tex
added figures to ms

@@
 
-\def\runninghead{Coevolution and network structure}
-\def\keywords{evolution, nestedness, modularity, bipartite}
+\def\runninghead{Plant-pollinator network assembly}
+\def\keywords{community assembly, change points, specialization,
+  nestedness, modularity, bipartite, preferential attachment}
 
@@
 \begin{abstract}
+TO BE RE-WRITTEN
   The structure of networks is related to ability of communities to
@@
 \subsection*{Change point analysis}
-
+\subsubsection*{Identifying change points}
 We employed a change point detection method \citep{peel2014detecting}
@@
 
+\subsubsection*{Characteristics of ``core'' and ``peripheral'' communities}
 To further elucidate the nature of the change points, we examine the
@@
 \label{sec:results}
+\subsection*{Change point analysis}
+\subsubsection*{Identifying change points}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.8\textwidth]{../analysis/changePoint/plotting/networks.pdf}
+  \caption{XXX}
+  \label{fig:changePoints}
+\end{figure}
+\clearpage
+
+\subsubsection*{Characteristics of ``core'' and ``peripheral''
+  communities}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.8\textwidth]{../analysis/changePoint/plotting/communities.pdf}
+  \caption{XXX}
+  \label{fig:communities}
+\end{figure}
+\clearpage
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.7\textwidth]{../analysis/changePoint/plotting/figures/pcoa/pcoa.pdf}
+  \caption{XXX}
+  \label{fig:pcoa}
+\end{figure}
+\clearpage
+
+\subsection*{Network structure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.4\textwidth]{../analysis/networkLevel/figures/baci.pdf}
+  \caption{XXX}
+  \label{fig:baci}
+\end{figure}
+\clearpage
+
+\subsection*{Network robustness}
+
+%% no need for robustness figs bec not sig
 

### 2016-08-03 analysis/changePoint/plotting/src/plotPcoa.R

@@ -9,7 +9,7 @@ plot.beta.div <- function(comms, statuses, fig.path, method,
       pcoa.comm <-  cmdscale(vegdist(comm, method=method))
       plot(pcoa.comm[status == treatments[1],], asp=1,
            col=cols[1], pch=16, cex=1.5,
-           ylim=range(pcoa.comm[,2]) + c(0,0.2),
+           ylim=range(pcoa.comm[,2]) + c(-0.1,0.1),
            xlim=range(pcoa.comm[,1]),
            xlab='',
            ylab='',
@@ -35,8 +35,8 @@ plot.beta.div <- function(comms, statuses, fig.path, method,
     layout(matrix(1:2, ncol=1, byrow=TRUE))
     par(oma=c(2,7,1,1), mar=c(0.5,0,1,0.5),
         mgp=c(2,1,0), cex.axis=1.5)
-    cols <- c("darkolivegreen3",
-              "darkgoldenrod1")
+    cols <- c("darksalmon",
+              "black")
     for(i in 1:length(comms)){
       pcoa(comms[[i]], statuses[[i]], i=i,
            method=method)

### 2016-08-03 analysis/networkLevel/plotting/src/plotPanels.R

@@ -1,6 +1,6 @@
 plot.panels <- function(){
   f <- function(){
-    col.lines <- brewer.pal(3, "Dark2")[2]
+    col.lines <- brewer.pal(4, "Greys")[3]
     col.fill <- add.alpha(col.lines, alpha=0.2)
     layout(matrix(1:3, ncol=1))
     par(oma=c(6, 7, 2, 1),
@@ -34,7 +34,7 @@ plot.panels <- function(){
     mtext("Specialization", 2, line=5, cex=1.5)
 
     axis(1, pretty(cor.dats$ypr), labels=pretty(cor.dats$ypr))
-    mtext("Years Post Restoration", 1, line=3.5, cex=1.5)
+    mtext("Year of assembly", 1, line=3.5, cex=1.5)
   }
   path <- 'figures'
   pdf.f(f, file=file.path(path,

### 2016-08-03 analysis/networkLevel/src/CalcMetrics.R

@@ -48,7 +48,7 @@ calc.metric <- function(dat.web) {
                        wbinary=TRUE)$statistic["NODF"]
     mets <-  c(nodf,
                networklevel(dat.web, index="H2",
-                            H2_integer=TRUE))
+                            H2_integer=FALSE))
   }
   mod.met <- calc.mod(dat.web)
   return(c(mets, mod.met= mod.met))

### 2016-08-05 ms/ms.tex
added methods to ms

@@
   following restoration}
-\def\author{Lauren C.\ Ponisio$^1$, Marilia P. Gaiarsa$^2$, Claire Kremen$^1$}
+\def\author{Lauren C.\ Ponisio$^{1,2}$, Marilia P. Gaiarsa$^3$, Claire
+  Kremen$^1$}
 
@@
     94720\\
+  \item Department of Entomology\\
+    University of California, Riverside\\
+    417 Entomology Bldg.\\
+    Riverside, California, USA\\
+    92521\\
   \item Departamento de Ecologia\\
@@
 \label{sec:introduction}
-\begin{itemize}
-\item The structure of networks is related to ability of communities
-  to maintain function in the face of species extinction.
-\item A key restoration aim is to facilitate assembly of robust
-  networks; thus it is critical to study how restoration
-  influences the assembly of plant-pollinator interactions.
-\item In general, however, few mechanisms of network assembly have
-  been developed and analyzed.
-\item The mostly widely explored mechanism of network assembly,
-  preferential attachment \citep{barabasi1999emergence}, predicts that
-  a new species (or node more generally) is more likely to interact
-  with species that are already well-connected \citep[''the
-  rich-get-richer'' principle,][]{barabasi1999emergence}.
-\item To date, only two field studies have examined how networks
-  assemble over time, often using space for time gradients.
-\item \cite{Olesen2008} was investigated day-to-day, temporal assembly
-  of a plant-pollinator network within a season, taking advantage of
-  the extreme seasonality of pollinator communities in Greenland.
-  \cite{Olesen2008} found that within a season, the network assembly
-  was similar to preferential attachment. New species tended to
-  interact with already well-connected species, likely because these
-  species are either more abundant or more temporally persistent.
+
+Global change has created a severe biodiversity crisis
+\citep{barnosky2011has}. As species are lost, so are their
+interactions, causing further coextinction cascades
+\citep{dunn2009sixth}.  Mutualistic systems are particualrly at risk.
+Ecological restoration plays an increasingly important role in
+recovering biodiversity (CITATIONS). Because the structure of
+interaction networks is related to ability of communities to maintain
+function in the face of species extinction \citep{Memmott2004}, a key
+restoration aim is to facilitate assembly of robust networks. Little
+is known, however, about how restoration influences the assembly of
+interaction networks, or ecological network assembly more generally.
+
+To date, only two field studies have examined how networks assemble
+over time, often using space for time gradients. \cite{Olesen2008} was
+investigated day-to-day, temporal assembly of a plant-pollinator
+network within a season, taking advantage of the extreme seasonality
+of pollinator communities in Greenland.  \cite{Olesen2008} found that
+within a season, the network assembly was similar to preferential
+attachment. New species tended to interact with already well-connected
+species, likely because these species are either more abundant or more
+temporally persistent.
+
+The mostly widely explored mechanism of network assembly, preferential
+attachment \citep{barabasi1999emergence}, predicts that a new species
+is more likely to interact with species that are already
+well-connected \citep[''the rich-get-richer''
+principle,][]{barabasi1999emergence}.
+
 \item Studying primary succession along a glacier foreland,
@@
 \textit{Eriogonum spp.}, \textit{Baccharis spp.}, \textit{Salvia
-  spp}. and others. The mean distance between monitoring sites was 15
-km, and the minimum distance between sites of the same type sampled in
-the same year was 2 km.  The entire area surveyed spanned almost 300
-km$^2$. The crop fields adjacent to all sites were similarly managed
-as intensive, high-input monoculture.
+  spp}. and others \citep{kremen-2015-602, mgonigle-2015-x}. The mean
+distance between monitoring sites was 15 km, and the minimum distance
+between sites of the same type sampled in the same year was 2 km.  The
+entire area surveyed spanned almost 300 km$^2$. The crop fields
+adjacent to all sites were similarly managed as intensive, high-input
+monoculture.
 
@@
 
+Quantitative networks were generative for each site through time. To
+account for the unequal number of surveys between years, the mean of
+the interactions between a pair of plants and pollinators across
+surveys within a year was used as a representation of the frequency of
+interactions.
+
 \subsection*{Change point analysis}
@@
 fragmentation or formation of communities (also called modules or
-compartments). Following \cite{peel2014detecting}, we first defined a
-probability distribution over the networks using the generalized
-hierarchical random graph model (GHRG). The GHRG model is able to
-capture both assortative and disassortative community structure
-patterns at all scales in the network \citep{peel2014detecting}. A
-network $G$ is composed of vertices $V$ and edges $E \subseteq {V × V
-}$. The GHRG model decomposes the $N$ vertices into a series of nested
-groups, the relationships among which are represented by the
-dendrogram $T$.  The tips of $T$ are the vertices of $G$, and the
-probability that two vertices $u$ and $v$ connect is given by the
-parameter $p_r$. The probability distribution of the network $G$ thus
-defined as:
+compartments). Change point detection methods have yet to be
+generalized to quantitative networks, so for this analysis we focused
+on qualitative (binary) networks. Following \cite{peel2014detecting},
+we first defined a probability distribution over the networks using
+the generalized hierarchical random graph model (GHRG). The GHRG model
+is able to capture both assortative and disassortative community
+structure patterns at all scales in the network
+\citep{peel2014detecting}. A network $G$ is composed of vertices $V$
+and edges $E \subseteq {V × V }$. The GHRG model decomposes the $N$
+vertices into a series of nested groups, the relationships among which
+are represented by the dendrogram $T$. The tips of $T$ are the
+vertices of $G$, and the probability that two vertices $u$ and $v$
+connect is given by the parameter $p_r$. The probability distribution
+of the network $G$ thus defined as:
 
@@
 a change point occurred between two time slices. To detect a change
-point, we compare the fir of two models -- one where a change point
+point, we compare the fit of two models -- one where a change point
 had occurred between two networks, and one where no change occurred --
@@
 estimate the null distribution \citep{peel2014detecting}. The change
-point analysis was carried out using code published online by L~.Peel.
+point analysis was carried out using code published online by L.~Peel.
 
@@
 generalized linear model with Binomial error to test whether the
-probability of a change point occurring varied by site type.
+probability of a change point occurring varied by site type. For the
+non-assembling hedgerows and weedy field margins, only sites with five
+or greater years of sampling was included in this analysis.
 
-\subsubsection*{Characteristics of ``core'' and ``peripheral'' communities}
-To further elucidate the nature of the change points, we examine the
-characteristics of the species that contributed the reorganization of
-interactions. To do so, we first generate dendrograms using the GHRG
-model before and after each change point.  We then determined which
-species belonged to the ``core'' and ``peripheral'' network
-communities at each site. The ``core'' network communities contain the
-majority of species and are more basil than the more derived, less
-specious ``peripheral'' network communities. We next use a
-Permutational Multivariate Analysis of Variance (PERMANOVA)
-\citep{anderson-2013-557} to determine whether the species
-compositions of the species that belonged to the core and peripheral
-communities differed. We also test whether core or peripheral species
-had more variability in their species compositions \citep[i.e.,
-multivariate dispersion, ][]{anderson-2011-19,
-  anderson-2006-683}.
+% \subsubsection*{Characteristics of species that contribute to change
+%   points}
 
-%% keep trait analysis?
+% To further elucidate the nature of the change points, we examine the
+% characteristics of the species that contributed the reorganization of
+% interactions. Some species remain in relativly similar network
+% positions through time, whereas others are more variable in their role
+% and thus contribute more strongly to network reorganization. If
+% assembly is following according to a mechanism of preferential
+% attachment, the more persistnet species with the higest degree would
+% be most stable in their network position ???
+
+% We test the hypothesis that the species that occupy the most stable
+% positions in the network are the most temporally persistent and or
+% generalized \citep{fang2012relative}.
+
+% We calculate species persistence as the proportion of the surveys a
+% pollinator was observed. Species observed consistently within and
+% between years are thus maximally persistent. Species degree is based
+% on plant-pollinator interaction observations from a more extensive
+% data-set from Yolo County (approx.~ 18000 interaction records) that
+% included both the data included in this study and additional data from
+% sites where we collected flower visitors using the same methods
+% \citep{mgonigle-2015-x, ponisio2015farm}.
+
+\subsection*{Temporal changes in interaction patterns}
+
+The changing points in network structure may contribute to the
+reorganization of the assembling networks into predictable interaction
+patterns. Pollination networks exhibit two main topologies ---
+modularity \citep[e.g.,][]{Olesen2007} and nestedness
+\citep[e.g.,][]{Bascompte2006, Bascompte2003}. Modular community
+interactions are insular, occurring within separate groups or
+``modules'' more often than between modules. Communities in the
+network may fragment as the network assembles, enhancing
+modularity. Conversely, nested networks are like pyramid of
+interactions, where there are some species that interact with many
+species, other species that interact with a subset of those species,
+and so on. If species entering the network tend to interact with the
+generalist base of the network pyramid (i.e., preferential
+attachment), nestedness would increase through time. Lastly, if the
+network is accumulating specialist species or species are beginning to
+limit their interaction niche breath as the network assembles, this
+would lead to an increase in the network-level specialization
+\citep{bluthgen-2006-9}. To test whether network modularity,
+nestedness or specialization changed linearly with assembly, we used
+linear mixed models with the descriptive network metrics as the
+response variable, year of assembly as the explanatory variable, and
+random effects of site and year. We use NODF (weighted or unweighted)
+to evaluate network nestedness \citep{nodf}. NODF evaluates whether
+species with fewer partners interact with subsets of partners with
+which more connected species interact \citep{nodf}. We used a
+hierarchical clustering algorithm to estimate modularity
+\citep{Newman2004, igraph}.
+
+We calculated standardized $z$-scores so that network metrics could be
+compared between communities. The $z$-scores were calculated by
+generating an ensemble of $999$ randomly assembled communities,
+subtracting the mean of the statistic calculated across these
+communities from the observed value, and then dividing by the standard
+deviation. To assemble random communities, we reshuffled the
+interactions between species but fixed the total number of
+interactions, species and the distribution of the interaction
+frequencies \citep{Galeano2009}.
+
+\subsection*{Core and peripherial communities}
+
+% To classify species as we first generate dendrograms using the GHRG
+% model before and after each change point.  We then determined which
+% species belonged to the ``core'' and ``peripheral'' network
+% communities at each site. The ``core'' network communities contain the
+% majority of species and are more basil than the more derived, less
+% specious ``peripheral'' network communities. We next use a
+% Permutational Multivariate Analysis of Variance (PERMANOVA)
+% \citep{anderson-2013-557} to determine whether the species
+% compositions of the species that belonged to the core and peripheral
+% communities differed. We also test whether core or peripheral species
+% had more variability in their species compositions \citep[i.e.,
+% multivariate dispersion][]{anderson-2011-19, anderson-2006-683}.
+
 % Lastly, we explored the functional diversity of the core and
-% peripheral pollinator species. The syrphids in our study area have
-% similar traits, so we focused on the trait diversity of the bee subset
-% of our interaction networks. We focus on resource capture and use
-% traits that collectively influence the distribution of bee species as
-% pollinators over space and time \citep{kremen-2015-602} including
-% resource specialization \citep[quantitative,
-% $d'$;][]{bluthgen-2006-9}, body size \citep[quantitative,
-% inter-tegular span, mm,][]{cane-1987-145}, sociality (categorical:
-% eusocial, solitary, cleptoparasitic), nest location (categorical:
-% above ground, below ground or mix), and nest construction
-% \citep[categorical: excavate or rent;][]{williams-2010-2280} as
-% described in more detail in \cite{kremen-2015-602}.  Each trait has
-% the same weight in trait diversity metric estimation
-% \citep{villeger-2008-2290, schleuter-2010-469}.  Pollinator
-% specialization was calculated using plant-pollinator interaction
-% observations from a more extensive data-set from Yolo County (18000
-% interaction records) that included both the data included in this
-% study and additional data from sites where we collected flower
-% visitors using the same methods \citep{mgonigle-2015-x}.  The
-% specialization metric measures the deviation of the observed
-% interaction frequency between a plant and pollinator from a null
-% expectation where all partners interact in proportion to their
-% abundances \citep{bluthgen-2006-9}.  It ranges from 0 for generalist
-% species to 1 for specialist species.
+% peripheral pollinator species
+
+% The syrphids in our study area have similar traits, so we focused on
+% the trait diversity of the bee subset of our interaction networks. We
+% focus on resource capture and use traits that collectively
+% \citep{kremen-2015-602} including resource specialization
+% \citep[quantitative, $d'$;][]{bluthgen-2006-9}, body size
+% \citep[quantitative, inter-tegular span, mm,][]{cane-1987-145},
+% sociality (categorical: eusocial, solitary, cleptoparasitic), nest
+% location (categorical: above ground, below ground or mix), and nest
+% construction \citep[categorical: excavator, mason or
+% renter;][]{williams-2010-2280} as described in more detail in
+% \cite{kremen-2015-602}.  Each trait has the same weight in trait
+% diversity metric estimation \citep{villeger-2008-2290,
+%   schleuter-2010-469}.
 
@@
 
-\subsection*{Network structure}
+\subsection*{Network robustness}
 
-The changing points in network structure may contribute to the
-reorganization of the assembling networks into predictable interaction
-patterns. Pollination networks exhibit two main topologies ---
-modularity \citep[e.g.,][]{Olesen2007} and nestedness
-\citep[e.g.,][]{Bascompte2006, Bascompte2003}.  Modular community
-interactions are insular, occurring within separate groups or
-``modules'' more often than between modules. Communities in the
-network may fragment as the network assembles, enhancing
-modularity. Conversely, nested networks are like pyramid of
-interactions, where there are some species that interact with many
-species, other species that interact with a subset of those species,
-and so on. If species entering the network tend to interact with the
-generalist base of the network pyramid, nestedness would increase
-through time. Lastly, if the network is accumulating specialist
-species or species are beginning to limit their interaction niche
-breath as the network assembles, this would lead to an increase in the
-network-level specialization \citep{bluthgen-2006-9}. To test whether
-network modularity, nestedness or specialization changed linearly with
-assembly, we used linear mixed models with the descriptive network
-metrics as the response variable, year of assembly as the explanatory
-variable, and random effects of site and year.
-
-\subsection*{Network robustness}
 Lastly, we test the changes in interaction patterns associated with
@@
 
-\subsubsection*{Characteristics of ``core'' and ``peripheral''
-  communities}
+\subsubsection*{Characteristics of species that contribute to change
+  points}
 
@@
   \centering
-  \includegraphics[width=.8\textwidth]{../analysis/changePoint/plotting/communities.pdf}
+  \includegraphics[width=.8\textwidth]{../analysis/variability/figures/cv/occ_degree.pdf}
   \caption{XXX}
-  \label{fig:communities}
+  \label{fig:cv}
 \end{figure}
 \clearpage
+
+% \subsubsection*{Characteristics of ``core'' and ``peripheral''
+%   communities}
+
+% \begin{figure}
+%   \centering
+%   \includegraphics[width=.8\textwidth]{../analysis/changePoint/plotting/communities.pdf}
+%   \caption{XXX}
+%   \label{fig:communities}
+% \end{figure}
+% \clearpage
 

### 2016-08-05 analysis/networkLevel/plotting/baci.R

@@ -40,11 +40,11 @@ mod.pi <- predict.int(mod= baci.mod.mod,
 ## ************************************************************
 load(file='saved/mods/baci_h2.Rdata')
 
-dd.h2 <- cbind(dd, zH2=0)
+dd.h2 <- cbind(dd, H2=0)
 
 h2.pi <- predict.int(mod= baci.h2.mod,
                         dd=dd.h2,
-                        y="zH2",
+                        y="H2",
                         family="gaussian")
 
 ## ************************************************************

### 2016-08-05 analysis/networkLevel/plotting/src/plotPanels.R

@@ -26,7 +26,7 @@ plot.panels <- function(){
     ## diversity
     plot.panel(new.dd=h2.pi,
                dats=cor.dats,
-               y1="zH2",
+               y1="H2",
                xs="ypr",
                col.fill=col.fill,
                col.lines=col.lines,

### 2016-08-05 analysis/networkLevel/src/CalcMetrics.R

@@ -1,8 +1,9 @@
+## used wine nestedness calculator and null model
 
-calc.metric <- function(dat.web) {
+calc.metric <- function(dat.web, H2_integer=FALSE) {
   ## calculates modularity
-  calc.mod <- function(dat.web){
-    ## converts a p-a matrix to a graph for modularity computation
+  calc.mod <- function(dat.web){ 
+    ## converts a p-a matrix to a graph for modularity computation 
     mut.adj <- function(x) {
       nr <- dim(x)[1]
       nc <- dim(x)[2]
@@ -21,7 +22,7 @@ calc.metric <- function(dat.web) {
                          membership(fastgreedy.community(graph,
                                                          weights=weights)),
                          weights=weights)
-
+    
     random.walk <-  modularity(graph,
                                membership(walktrap.community(graph,
                                                              weights=weights)),
@@ -39,23 +40,110 @@ calc.metric <- function(dat.web) {
   if(min(dat.web) == max(dat.web)){
     return(c(mets=rep(0, 1),
              mod.met=rep(0,3)))
-
-  }else{
-    ## wbinary=TRUE ensures binary matrices are calcualted like
-    ## unweighted
+  } else{
     nodf <- nestednodf(dat.web,
                        weighted=TRUE,
                        wbinary=TRUE)$statistic["NODF"]
     mets <-  c(nodf,
                networklevel(dat.web, index="H2",
-                            H2_integer=FALSE))
+                            H2_integer=H2_integer))
   }
   mod.met <- calc.mod(dat.web)
   return(c(mets, mod.met= mod.met))
-
+  
 }
 
+prob.null <- function(M) {
+  randiag <- function(vdims) {
+    mats <- diag(vdims)
+    ## what cell number are the interactions
+    posdiag <- 1:vdims + ((1:vdims - 1) * vdims)
+    desp <- matrix(rep(1:vdims, vdims), vdims, vdims, 
+                   byrow = TRUE) - (1:vdims)
+    fdesp <- sample(1:vdims)
+    move <- desp[cbind(1:vdims, fdesp)]
+    moved <- posdiag + move
+    mdesp <- matrix(0, vdims, vdims)
+    mdesp[moved] <- 1
+    return(mdesp)
+  }
[…]

### 2016-08-05 analysis/variability/cv.R

@@ -16,6 +16,11 @@ dprime <- cv.trait(spec, byYear, trait="d",
                    time.col="Date",
                    abund.col="Abund")
 
+degree <- cv.trait(spec, byYear, trait="degree",
+                   method= "time",
+                   time.col="Date",
+                   abund.col="Abund")
+
 itd <- cv.trait(spec, byYear, trait="ITD",
                 method= "time",
                 time.col="Date",
@@ -25,6 +30,9 @@ itd <- cv.trait(spec, byYear, trait="ITD",
 ## coefficient of variation of degree thingy through time
 ## ************************************************************
 
+## ************************************************************
+## dprime
+## not sig
 dprime.k.sd <- cv.trait(spec,
                         specs[specs$speciesType =="pollinator",],
                         trait="d",
@@ -33,16 +41,34 @@ dprime.k.sd <- cv.trait(spec,
                         cv.function=sd,
                         zero2na=TRUE, standard.cv=FALSE,
                         na.rm=TRUE)
+summary(dprime.k.sd$lm.nss)
 
-dprime.closeness.sd <- cv.trait(spec,
+## not sig
+dprime.k.cv <- cv.trait(spec,
+                        specs[specs$speciesType =="pollinator",],
+                        trait="d",
+                        method= "time", time.col="assem",
+                        abund.col="k",
+                        cv.function=cv,
+                        zero2na=TRUE,
+                        standard.cv=TRUE,
+                        na.rm=TRUE)
+summary(dprime.k.cv$lm.nss)
+
+## not sig
+dprime.closeness.cv <- cv.trait(spec,
                                 specs[specs$speciesType =="pollinator",],
                                 trait="d",
                                 method= "time", time.col="assem",
                                 abund.col="weighted.closeness",
-                                cv.function=sd, zero2na=TRUE,
+                                cv.function=cv,
+                                zero2na=TRUE,
                                 standard.cv=FALSE,
                                 na.rm=TRUE)
+summary(dprime.closeness.cv$lm.nss)
 
+## ************************************************************
+## occurrence
 ## sig
 occ.k.sd <- cv.trait(spec,
                      specs[specs$speciesType =="pollinator",],
@@ -53,6 +79,7 @@ occ.k.sd <- cv.trait(spec,
                      zero2na=TRUE,
                      standard.cv=FALSE,
[…]

### 2016-08-05 dataPrep/dataPrep.R

@@ -39,11 +39,12 @@ agg.spec <- aggregate(list(abund=spec$GenusSpecies),
                       length)
 
 nets.all <- samp2site.spp(agg.spec$PlantGenusSpecies,
-                      agg.spec$GenusSpecies, agg.spec$abund, FUN=sum)
+                          agg.spec$GenusSpecies, agg.spec$abund, FUN=sum)
 
 all.specializations <- specieslevel(nets.all,
                                     index=c("proportional generality",
-                                    "d"))
+                                      "degree",
+                                      "d"))
 
 traits <- data.frame(GenusSpecies= unlist(sapply(all.specializations,
                        rownames)),
@@ -58,7 +59,7 @@ spec <- spec[!spec$SiteStatus %in% to.drop.status,]
 ## add various traits
 ## specialization
 spec$d <- traits$d[match(spec$GenusSpecies, traits$GenusSpecies)]
-
+spec$degree <- traits$degree[match(spec$GenusSpecies, traits$GenusSpecies)]
 ## occurence
 load('~/Dropbox/hedgerow/data_sets/matrices/net/bee.syr.RData')
 occ <- apply(mat, c(3,1), function(x){
@@ -68,7 +69,7 @@ occ <- apply(mat, c(3,1), function(x){
 
 findOcc <- function(x){
   out <- try(occ[x["GenusSpecies"], paste(x["Site"], x["SiteStatusBACI"],
-                                sep=":")], silent=TRUE)
+                                          sep=":")], silent=TRUE)
   if(inherits(out, "try-error")) out <- NA
   return(out)
 }
@@ -166,11 +167,11 @@ species.lev <- lapply(networks, function(x){
 })
 
 specializations  <-  mapply(function(a, b)
-  getSpec(species.lev = a,
-          names.net = b),
-  a = species.lev,
-  b = names(networks),
-  SIMPLIFY = FALSE)
+                            getSpec(species.lev = a,
+                                    names.net = b),
+                            a = species.lev,
+                            b = names(networks),
+                            SIMPLIFY = FALSE)
 
 specializations <- do.call(rbind, specializations)
 rownames(specializations) <- NULL
@@ -188,23 +189,23 @@ save(specializations, file=file.path(f.path, 'specializations.Rdata'))
 
 ## plants
 diff.gens.plants <- getVisitChange(0, 0.2, "proportional.generality",
-                            "PlantGenusSpecies", "GenusSpecies")
[…]

### 2016-08-09 ms/ms.tex
finsihed first draft of introduction

@@
 Keywords: changing points, temporal networks, hedgerows, species
-interactions, preferential attachment
+interactions, preferential attachment, mutualisms
 
@@
 
-Global change has created a severe biodiversity crisis
-\citep{barnosky2011has}. As species are lost, so are their
-interactions, causing further coextinction cascades
-\citep{dunn2009sixth}.  Mutualistic systems are particualrly at risk.
-Ecological restoration plays an increasingly important role in
-recovering biodiversity (CITATIONS). Because the structure of
-interaction networks is related to ability of communities to maintain
-function in the face of species extinction \citep{Memmott2004}, a key
-restoration aim is to facilitate assembly of robust networks. Little
-is known, however, about how restoration influences the assembly of
-interaction networks, or ecological network assembly more generally.
+%% introduce mutualisms, restoration, network assembly
+Global change has created a severe biodiversity crisis, and as species
+are lost, so are their interactions \citep{dunn2009sixth,
+  barnosky2011has}. Because mutualistic interactions are essential for
+maintaining the diversity their component guilds of species, these
+systems are particularly at risk from coextinction cascades. The
+nature of these coextinction cascades depends on the interaction
+patterns within a community \citep{Memmott2004, Rezende2007,
+  Bascompte2009}. Recovering the lost biodiversity and interactions
+through ecological restoration has become increasingly imperative, and
+a key restoration aim is to facilitate assembly of robust interaction
+networks \citep{menz-2010-4}. We know little, however, about how to
+re-assemble interacting communities through restoration, or the
+process of ecological network assembly more generally.
 
-To date, only two field studies have examined how networks assemble
-over time, often using space for time gradients. \cite{Olesen2008} was
-investigated day-to-day, temporal assembly of a plant-pollinator
-network within a season, taking advantage of the extreme seasonality
-of pollinator communities in Greenland.  \cite{Olesen2008} found that
-within a season, the network assembly was similar to preferential
-attachment. New species tended to interact with already well-connected
-species, likely because these species are either more abundant or more
-temporally persistent.
-
+%% most people think network assemble via preferential attachment, and
+%% there is some evidence of that. (other studies on mutualisms?)
 The mostly widely explored mechanism of network assembly, preferential
@@
 well-connected \citep[''the rich-get-richer''
-principle,][]{barabasi1999emergence}.
+principle,][]{barabasi1999emergence}. In pollination systems --- a
+particularly ubiquitous mutualism \citep{ollerton-2011-321,
+  klein-2007-303} --- some studies have found support for this
+mechanism of assembly. Investigating the day-to-day, temporal assembly
+of a plant-pollinator network within a season, \cite{Olesen2008} found
+that new species tended to interact with already well-connected
+species, likely because these species are either more abundant or more
+temporally persistent \citep{Olesen2008}. In addition, using a
+space-for-time substitution to study primary succession along a
+glacier foreland, \cite{albrecht2010plant} found some indication
+assembly was occurring through preferential attachment. Network
+nestedness, a pattern of interactions where a generalist core
+interacts with both specialist and generalist species, increased as
+the community aged \citep{albrecht2010plant}. Increasing nestedness
+could result from a process like preferential attachment where
+specialist species attach to the well-connected, generalist core.
 
-\item Studying primary succession along a glacier foreland,
-  \cite{albrecht2010plant} found a similar pattern where nestedness, a
-  pattern of interactions where a generalist core interacts with both
-  specialist and generalist species, increased as the community
-  aged. % Studying ``managed succession'' of a clear-cut pine forest,
-  % \cite{devoto2012understanding} found changes in network structure
-  % were explained by a combination of age, tree density and variation
-  % in tree diameter. 
-\item Even non-successional temporal dynamics suggest a stable core of
-  generalists persist despite high turnover of peripheral species
-  \citep{fang2012relative, diaz2010changes, alarcon2008year}.
-\item In contrast to the ordered network build-up described by
-  preferential attachment, assembly may be punctuated by significant
-  reorganizations of interactions. For example, as new species are
-  added, resident species change their interaction partners to
+%% there are other ways to assemble that are not preferentia;
+%% attchment, particularly vis big reorganizations of interactions
+In contrast to the ordered network build-up described by preferential
+attachment, assembly can be punctuated by significant reorganizations
+of interactions \citep{peel2014detecting}. For example, as new species
+are added, resident species change their interaction partners to
 minimize competition, or become extinct. Such significant
 reorganizations of interactions, or changing points, have been
-  observed in networks \citep{peel2014detecting}.
-\end{itemize}
+observed in networks responding to abrupt shifts in the behavior of
+interacters \citep{peel2014detecting}. No studies, however, have
+examined whether changing points occur during ecological network
+assembly.
+
+%% restoration
+Understanding network assembly is particularly relevant to ecological
+restoration, which is essentially 'applied succession'
+\citep[e.g.,]{parker1997scale}. In pollination systems, the time since
+an area was restored has been shown to effect the structure of
+networks \citep{forup-2008-742, forup2008restoration,
+  devoto2012understanding}, suggesting interactions are evolving as the
+community develops. Understanding the mechanisms of network assembly
+will help to guide the restoration of particular communities.
+
+%% restoration of ecosystem services is really important, particularly
+%% in ag. but ag is a bad place for pollinators
+Facilitating effective restoration of networks is particularly
+imperative in areas where species interactions provide essential
+ecosystem services, such as crop pollination. In intensively managed
+agricultural landscapes, the demand for pollination services is the
+greatest \citep{kremen-2008-10}. However, honey bees, managed
+extensively around the world to provide crop pollination, are in
+global decline \citep{neumann-2010-1, van-engelsdorp-2009-e6481}. In
+addition, native pollinators, which have the capacity to provide
+sufficient crop pollination \citep{kremen-2002-16816,
+  winfree-2007-1105, kremen-2004-1109}, are in short supply because
+these landscapes make poor habitats for pollinator populations
+\citep{kremen-2002-16816}. To ensure provision the continued provision
+of ecosystem services and curb biodiversity loss, effective
+restoration of pollinators and their interactions in agricultural
+landscapes is critical.
+
+%% hedgerows!!  
+To promote pollinator services in agriculture, farmers are
+increasingly turning to the habitat restoration technique of planting
+strips of native plants along farm edges (hedgerows) to help provide
+habitat for pollinators without removing arable land from
+production. Hedgerows have been shown to augment the richness,
+abundance and trait diversity of pollinators in agricultural
+landscapes\citep{morandin-2013-829, mgonigle-2015-x, kremen-2015-602,
+  ponisio2015farm}. In addition, hedgerows promote the persistence and
+colonization of floral resource specialists
+\citep{mgonigle-2015-x}. Little is known however, about the assembly
+of the network following hedgerow restoration.
+
+Using a long-term data-set of plant-pollinator communities assembling
+following hedgerow restoration in the highly simplified and
+intensively managed agricultural landscape of California's Central
+Valley, we explore the process of network development. We first
+determine whether the mechanism underlying network assembly is a
+smooth build up of interactions as would be predicted by preferential
+attachment, or punctuated by significant reorganizations of
+interactions (i.e., changing points). Even with changing points in
+interaction organization, networks could still be assembling via
+preferential attachment if the network reorganizations were primarily
+driven the by peripheral, temporally variable species while a stable,
+well-connected core of species still persists. We thus examine whether
+the species are most variable in their network position --- and thus
+important contributors network reorganizations --- are less persistent
+and connected species. %% interaction beta div, time beta div??
+Lastly, we examine whether networks are assembling toward predictable
+interaction patterns, and the ramifications for the robustness of the
+networks to species extinction and perturbation.
+
+%% objectives
+% 1) Are hedgerows assembling by preferential attachment, or are there
+% more abrupt reorganizations of network structure?
+% 2) Who are the species responsible for interaction reorganizations?
+% how to incorporate this? 
+% 3) Are the changing points a product of turnover in species
+% composition, or are more substantial changes in the verticles of the
+% networks?
+% 4) Is the structure of interactions changing? Does this change the
+% robustness of the network to species extinction and/or perturbation?
+
+% Pollinators are one such key group of interactors: $75\%$ of all
+% crop species depend to some extent on pollinator visitation
+% \citep{klein-2007-303}, and animal-pollinated crops supply a large
+% proportion of essential nutrients to the human diet
+% \citep{eilers-2011-e21363}.
+
+% Even non-successional temporal dynamics suggest a stable core of
+% generalists persist despite high turnover of peripheral species
+% \citep{fang2012relative, diaz2010changes, alarcon2008year}.
 
@@
 distribution of bipartitions, from which a consensus tree is derived
-\citep{peel2014detecting}. ADD NEWMAN PAPER. We used $\beta$
-distributions with the hyperparameters $\alpha=\beta=1$ to define
-priors for $p_r$.
+\citep{peel2014detecting}. We used $\beta$ distributions with the
+hyperparameters $\alpha=\beta=1$ to define priors for $p_r$.
 
@@
 
-% \subsubsection*{Characteristics of species that contribute to change
-%   points}
+\subsubsection*{Characteristics of species that contribute to change
+  points}
 
-% To further elucidate the nature of the change points, we examine the
-% characteristics of the species that contributed the reorganization of
-% interactions. Some species remain in relativly similar network
-% positions through time, whereas others are more variable in their role
-% and thus contribute more strongly to network reorganization. If
-% assembly is following according to a mechanism of preferential
-% attachment, the more persistnet species with the higest degree would
-% be most stable in their network position ???
+To further elucidate the nature of the change points, we examine the
+characteristics of the species that contributed the reorganization of
+interactions. Some species remain in relatively similar network
+positions through time, whereas others are more variable in their role
+and thus contribute more strongly to network reorganization. We test
+the that more persistent species with the highest degree are the most
+stable in their network position, as would be expected is the networks
+were assembling via preferential attachment.
 
-% We test the hypothesis that the species that occupy the most stable
-% positions in the network are the most temporally persistent and or
-% generalized \citep{fang2012relative}.
-
-% We calculate species persistence as the proportion of the surveys a
-% pollinator was observed. Species observed consistently within and
-% between years are thus maximally persistent. Species degree is based
-% on plant-pollinator interaction observations from a more extensive
-% data-set from Yolo County (approx.~ 18000 interaction records) that
-% included both the data included in this study and additional data from
-% sites where we collected flower visitors using the same methods
-% \citep{mgonigle-2015-x, ponisio2015farm}.
+We calculate species persistence as the proportion of the surveys a
+pollinator was observed. Species observed consistently within and
+between years are thus maximally persistent. Weighted species degree
+is calculated from interaction observations observed in more extensive
+data-set from Yolo County (approx.~ 18000 interaction records) that
+included both the data included in this study and additional data from
+sites where we collected flower visitors using the same methods
+\citep{mgonigle-2015-x, ponisio2015farm}.%% does this make sense to
+                                %% calculate from an independent data set?
+To represent the the variability of species within networks, we
+computed the coefficient of variation of weighted closeness at each
+site through time. Closeness describes the centrality of a species in
+the network by calculating path lengths to other vertices (species) in
+the graph. We used linear mixed models to test whether the variability
+of species closeness values was related to the persistence or degree
+of that species. We included random effects for species, as well as
+site. We focused on the pollinator species because the hedgerow
+flowers are planted and thus are not directly assembling. Because
+degree and persistence were strongly correlated, ($\rho = 0.84$,
+$p$-value $<$ $2*10^{-16}$), each explanatory variable was included in
+the model separately.
 
 \subsection*{Temporal changes in interaction patterns}
-
 The changing points in network structure may contribute to the
@@
 response variable, year of assembly as the explanatory variable, and
-random effects of site and year. We use NODF (weighted or unweighted)
-to evaluate network nestedness \citep{nodf}. NODF evaluates whether
-species with fewer partners interact with subsets of partners with
-which more connected species interact \citep{nodf}. We used a
-hierarchical clustering algorithm to estimate modularity
-\citep{Newman2004, igraph}.
+random effects of site and year.
 
-We calculated standardized $z$-scores so that network metrics could be
-compared between communities. The $z$-scores were calculated by
-generating an ensemble of $999$ randomly assembled communities,
+We use NODF (weighted or unweighted) to evaluate network nestedness
+\citep{nodf}. NODF evaluates whether species with fewer partners
+interact with subsets of partners with which more connected species
+interact \citep{nodf}. To estimate modularity, we use a hierarchical
+clustering algorithm \citep{Newman2004, igraph}.  We calculated
+standardized $z$-scores so that nestedness and modularity metrics
+could be compared between communities. The $z$-scores were calculated
+by generating an ensemble of $999$ randomly assembled communities,
 subtracting the mean of the statistic calculated across these
@@
 interactions, species and the distribution of the interaction
-frequencies \citep{Galeano2009}.
+frequencies \citep{Galeano2009}. Lastly, Network specialization was
+measured using H2, which estimate the deviation of the observed
+interaction frequency between plants and pollinators from a null
+expectation where all partners interact in proportion to their
+abundances \citep{bluthgen-2006-9}. It ranges from 0 for generalized
+networks to 1 for specialized networks.
 
-\subsection*{Core and peripherial communities}
+\subsection*{Network robustness}
+
+Lastly, we test the changes in interaction patterns associated with
+network assembly affect the robustness. First, following
+\cite{Memmott2004}, we simulate the extinction of plant species the
+subsequent extinction cascades of pollinator species. Because the
+reproduction of plant species if facilitated by active restoration
+efforts, it is unlikely the extinction of pollinators would effect
+plant populations in the hedgerows. However, plants ceasing to bloom,
+for example in response to drought, will likely effect the pollinators
+that depend on them. Plants species were eliminated based on their
+degree or abundance, and the number of pollinators that secondarily
+went extinct is calculated. The area below the extinction curve is
+used as a measure of network robustness \citep{Memmott2004}.
+
+%% Marilia add perturbation methods here
+
+\section*{Results}
+\label{sec:results}
+\subsection*{Change point analysis}
+\subsubsection*{Identifying change points}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.8\textwidth]{../analysis/changePoint/plotting/networks.pdf}
+  \caption{XXX}
+  \label{fig:changePoints}
+\end{figure}
+\clearpage
+
+\subsubsection*{Characteristics of species that contribute to change
+  points}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.8\textwidth]{../analysis/variability/figures/cv/occ_degree.pdf}
+  \caption{XXX}
+  \label{fig:cv}
+\end{figure}
+\clearpage
+
+\subsection*{Network structure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.4\textwidth]{../analysis/networkLevel/figures/baci.pdf}
+  \caption{XXX}
+  \label{fig:baci}
+\end{figure}
+\clearpage
+
+\subsection*{Network robustness}
+
+\section*{Discussion}
+\label{sec:discussion}
+
+\section*{Acknowledgments}
+\label{sec:acknowledge}
+
+We would like to thank Leto Peel and Aaron Clauset for their
+invaluable discussions and for help with the change point analysis.
+We thank the growers and land owners that allowed us to work on their
+property.  We also greatly appreciate the identification assistance of
+expert taxonomists Martin Hauser, Robbin Thorp and Jason Gibbs.  This
+work was supported by funding from the Army Research Office
+(W911NF-11-1-0361 to CK), the Natural Resources Conservation Service
+(CIG-69-3A75-12-253, CIG-69-3A75-9-142, CIG-68-9104-6-101 and
+WLF-69-7482-6-277 to The Xerces Society), the National Science
+Foundation (DEB-0919128 to CK), The U.S.  Department of Agriculture
+(USDA-NIFA 2012-51181-20105 to Michigan State University).  Funding
+for LCP was provided by an NSF Graduate Research Fellowship and the
+USDA NIFA Graduate Fellowship. FUNDING FOR MARILLIA.
+
+\bibliographystyle{ecol_let}
+
+\bibliography{refs}
+
+\end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-PDF-mode: t
+%%% End:
+
+% \subsection*{Core and peripherial communities}
 
@@
 % \citep{lme4,lmetest}.
-
-\subsection*{Network robustness}
-
-Lastly, we test the changes in interaction patterns associated with
-network assembly affect the robustness of the network to species
-loss. Following \cite{Memmott2004}, we simulate the extinction of
-plant species the subsequent extinction cascades of pollinator
-species. Because the reproduction of plant species if facilitated by
-active restoration efforts, it is unlikely the extinction of
-pollinators would effect plant populations in the hedgerows. However,
-plants ceasing to bloom, for example in response to drought, will
-likely effect the pollinators that depend on them. Plants species were
-eliminated based on their degree or abundance, and the number of
-pollinators that secondarily went extinct is calculated. The area
-below the extinction curve is used as a measure of network robustness.
-
-\section*{Results}
-\label{sec:results}
-\subsection*{Change point analysis}
-\subsubsection*{Identifying change points}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.8\textwidth]{../analysis/changePoint/plotting/networks.pdf}
-  \caption{XXX}
-  \label{fig:changePoints}
-\end{figure}
-\clearpage
-
-\subsubsection*{Characteristics of species that contribute to change
-  points}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.8\textwidth]{../analysis/variability/figures/cv/occ_degree.pdf}
-  \caption{XXX}
-  \label{fig:cv}
-\end{figure}
-\clearpage
-
-% \subsubsection*{Characteristics of ``core'' and ``peripheral''
-%   communities}
-
-% \begin{figure}
-%   \centering
-%   \includegraphics[width=.8\textwidth]{../analysis/changePoint/plotting/communities.pdf}
-%   \caption{XXX}
-%   \label{fig:communities}
-% \end{figure}
-% \clearpage
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.7\textwidth]{../analysis/changePoint/plotting/figures/pcoa/pcoa.pdf}
-  \caption{XXX}
-  \label{fig:pcoa}
-\end{figure}
-\clearpage
-
-\subsection*{Network structure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.4\textwidth]{../analysis/networkLevel/figures/baci.pdf}
-  \caption{XXX}
-  \label{fig:baci}
-\end{figure}
-\clearpage
-
-\subsection*{Network robustness}
-
-%% no need for robustness figs bec not sig
-
-\section*{Discussion}
-\label{sec:discussion}
-
-\section*{Acknowledgments}
-\label{sec:acknowledge}
-
-We would like to thank Leto Peel and Aaron Clauset for their
-invaluable discussions and for help with the change point analysis.
-We thank the growers and land owners that allowed us to work on their
-property.  We also appreciate the identification assistance of expert
-taxonomists Martin Hauser, Robbin Thorp and Jason Gibbs.  This work
-was supported by funding from the Army Research Office
-(W911NF-11-1-0361 to CK), the Natural Resources Conservation Service
-(CIG-69-3A75-12-253, CIG-69-3A75-9-142, CIG-68-9104-6-101 and
-WLF-69-7482-6-277 to The Xerces Society), the National Science
-Foundation (DEB-0919128 to CK), The U.S.  Department of Agriculture
-(USDA-NIFA 2012-51181-20105 to Michigan State University).  Funding
-for LCP was provided by an NSF Graduate Research Fellowship and the
-USDA NIFA Graduate Fellowship. FUNDING FOR MARILLIA.
-
-\bibliographystyle{ecol_let}
-
-\bibliography{refs}
-
-\end{document}
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-PDF-mode: t
-%%% End:

### 2016-08-09 analysis/networkLevel/laplacian.R

@@ -1,0 +1,58 @@
+rm(list=ls())
+library(lme4)
+library(lmerTest)
+setwd('~/Dropbox/hedgerow_assembly/analysis/networkLevel')
+
+load('../../data/networks/allSpecimens.Rdata')
+f.path <- "../changePoint/cptPeel/baci"
+load(file=file.path(f.path, "graphs.Rdata"))
+
+source('src/laplacian_functions.R')
+
+fig.path <- '../speciesLevel/Figures'
+sites <- sapply(strsplit(names(nets), "[.]"), function(x) x[1])
+years <- sapply(strsplit(names(nets), "[.]"), function(x) x[2])
+
+status.table <- table(spec$Site, spec$SiteStatus)
+status.table <- as.data.frame(cbind(rownames(status.table),
+                                    colnames(status.table)[apply(status.table,
+                                                            1, which.max)]))
+colnames(status.table)<- c("Site", "SiteStatus")
+
+all.alg.Con <- t(do.call(cbind.data.frame, lapply(nets, algCone)))
+all.alg.Con <- as.data.frame(cbind(sites,years, all.alg.Con))
+colnames(all.alg.Con) <- c("Site", "Year","Ncomp", "AlgCon")
+
+## add status
+all.alg.Con.status <- merge(x=all.alg.Con,
+                            y= status.table,
+                            by.x="Site",
+                            by.y="Site")
+
+## add ypr
+all.alg.Con.status$ypr <- spec$ypr[match(paste(all.alg.Con.status$Site,
+                                               all.alg.Con.status$Year),
+                                         paste(spec$Site, spec$Year))]
+
+baci.sites <- c("MullerB", "Sperandio", "Barger", "Butler", "Hrdy")
+
+all.alg.Con.status$ypr[!all.alg.Con.status$Site %in% baci.sites] <- NA
+
+## change AlgCon to numeric (it is a factor for some reason..)
+all.alg.Con.status$AlgCon <- as.numeric(as.character(
+  all.alg.Con.status$AlgCon))
+
+alg.con.mod <- lmer(AlgCon ~ SiteStatus +
+                    (1|Site) + (1|Year),    
+                    data=all.alg.Con.status)
+
+summary(alg.con.mod)
+
+
+alg.con.mod.ypr <- lmer(AlgCon ~ ypr +
+                        (1|Site) + (1|Year),    
+                        data=all.alg.Con.status)
[…]

### 2016-08-09 analysis/networkLevel/plotting/resilience.R

@@ -6,7 +6,7 @@ source('plotting/src/plotPanels.R')
 source('src/initialize.R')
 
 ## ************************************************************
-## robustness
+## robustness to species extinction
 ## ************************************************************
 ## either "abund" or "degree"
 extinction.method <- "degree"
@@ -31,8 +31,32 @@ ypr.pi <- predict.int(mod= mod.ypr,
                         family="gaussian")
 
 plot.predict.ypr(new.dd=ypr.pi,
-                 ylabel="Robustness",
+                 ylabel="Robustness to \n species extinction",
                  dats=res,
                  y1="Robustness",
                  extinction.method=extinction.method)
+
+## ************************************************************
+## robustness to perturbation
+## ************************************************************
+
+load(file=file.path(save.path, 'mods/AlgCon.Rdata'))
+
+dd.ypr.alg <- expand.grid(ypr=seq(from= min(all.alg.Con.status$ypr,
+                                    na.rm=TRUE),
+                    to= max(all.alg.Con.status$ypr,
+                      na.rm=TRUE),
+                    length=10),
+                  AlgCon=0)
+
+ypr.pi.alg <- predict.int(mod= alg.con.mod.ypr,
+                        dd=dd.ypr.alg,
+                        y="AlgCon",
+                        family="gaussian")
+
+plot.predict.ypr(new.dd=ypr.pi.alg,
+                 ylabel="Robustness to perturbation",
+                 dats=all.alg.Con.status,
+                 y1="AlgCon",
+                 extinction.method="perturb")
 

### 2016-08-09 analysis/variability/cv.R

@@ -29,9 +29,8 @@ itd <- cv.trait(spec, byYear, trait="ITD",
 ## ************************************************************
 ## coefficient of variation of degree thingy through time
 ## ************************************************************
-
-## ************************************************************
 ## dprime
+## ************************************************************
 ## not sig
 dprime.k.sd <- cv.trait(spec,
                         specs[specs$speciesType =="pollinator",],
@@ -69,6 +68,7 @@ summary(dprime.closeness.cv$lm.nss)
 
 ## ************************************************************
 ## occurrence
+## ************************************************************
 ## sig
 occ.k.sd <- cv.trait(spec,
                      specs[specs$speciesType =="pollinator",],
@@ -117,6 +117,7 @@ cor.test(occ.k.sd$data$traits.ns, occ.k.sd$data$spec)
 
 ## ************************************************************
 ## degree
+## ************************************************************
 ## sig
 degree.k.sd <- cv.trait(spec,
                         specs[specs$speciesType =="pollinator",],

### 2016-08-09 analysis/variability/plotting/dissimilarity.R

@@ -31,7 +31,7 @@ plot.predict.div(new.dd=plants.k.pi,
                  dats=plants,
                  xs="k",
                  y1="Dist",
-                 xlabel="Plant network position",
+                 xlabel="Plant network position (k)",
                  legend.loc="topright",
                  height=5,
                  width=5,
@@ -40,6 +40,36 @@ plot.predict.div(new.dd=plants.k.pi,
                  treatments="all",
                  col.lines="black")
 
+
+## ************************************************************
+## interaction dispersion, closeness, plants
+## ************************************************************
+
+dd.plants.closeness <- expand.grid(closeness=seq(
+                             from= min(plants$closeness),
+                             to= max(plants$closeness),
+                             length=10),
+                           SiteStatus= c("all"),
+                           Dist= 0)
+
+plants.closeness.pi <- predict.int(mod= mod.plants.close,
+                           dd=dd.plants.closeness,
+                           y="Dist",
+                           family="gaussian")
+
+plot.predict.div(new.dd=plants.closeness.pi,
+                 ylabel="Interaction dispersion",
+                 dats=plants,
+                 xs="closeness",
+                 y1="Dist",
+                 xlabel="Plant closeness",
+                 legend.loc="topright",
+                 height=5,
+                 width=5,
+                 x.adj=0.5,
+                 scaled=FALSE,
+                 treatments="all",
+                 col.lines="black")
 
 
 ## ************************************************************
@@ -96,7 +126,7 @@ plot.predict.div(new.dd=pols.k.pi,
                  dats=pols,
                  xs="k",
                  y1="Dist",
-                 xlabel="Pollinator network position",
+                 xlabel="Pollinator network position (k)",
                  legend.loc="topright",
                  height=5,
                  width=5,
@@ -130,6 +160,40 @@ plot.predict.div(new.dd=pols.d.pi,
                  xs="d",
                  y1="Dist",
                  xlabel="Pollinator specialization",
+                 legend.loc="topright",
+                 height=5,
+                 width=5,
+                 x.adj=0.5,
+                 scaled=FALSE,
+                 treatments="all",
+                 col.lines="black")
+
+
+
+
+## ************************************************************
+## interaction dispersion, closeness, pols
+## ************************************************************
+
+dd.pols.closeness <- expand.grid(closeness=seq(
+                           from= min(pols$closeness),
+                           to= max(pols$closeness),
+                           length=10),
[…]

### 2016-08-09 analysis/variability/src/calcDis.R

@@ -41,9 +41,9 @@ calcDis <- function(site, ind, nets, spec){
                                  unlist(sapply(pp, names)), "[.]"),
                                  function(x) x[2])))
     out.dist$SiteStatus <- spec$SiteStatus[match(paste(out.dist$Site,
-                                                 out.dist$Year),
-                                           paste(spec$Site,
-                                                 spec$Year))]
+                                                       out.dist$Year),
+                                                 paste(spec$Site,
+                                                       spec$Year))]
     rownames(out.dist) <- NULL
     return(out.dist)
   }
@@ -54,9 +54,14 @@ getDis <- function(sites, ind, nets, specs.agg, traits, spec){
   pp <- lapply(unique(sites), calcDis, ind=ind, nets, spec)
   pp <- pp[!sapply(pp, function(x) all(is.na(x)))]
   pp <- do.call(rbind, pp)
-  pp$k <- specs.agg$k[match(pp$GenusSpecies, specs.agg$GenusSpecies)]
-  pp$d <- traits$d[match(pp$GenusSpecies, traits$GenusSpecies)]
-  pp$occ.date <- traits$occ.date[match(pp$GenusSpecies, traits$GenusSpecies)]
+  pp$k <- specs.agg$k[match(pp$GenusSpecies,
+                            specs.agg$GenusSpecies)]
+  pp$closeness <- specs.agg$closeness[match(pp$GenusSpecies,
+                                            specs.agg$GenusSpecies)]
+  pp$d <- traits$d[match(pp$GenusSpecies,
+                         traits$GenusSpecies)]
+  pp$occ.date <- traits$occ.date[match(pp$GenusSpecies,
+                                       traits$GenusSpecies)]
   pp$core <- "core"
   pp$core[pp$k <= 1] <- "peripheral"
   return(pp)

### 2016-08-11 ms/ms.tex
finished writing results and methods

@@
 \usepackage[parfill]{parskip}
+\usepackage{graphicx}
+\usepackage{longtable}
+\usepackage{times}
+\usepackage{textcomp}
 \usepackage{xr}
-\usepackage{paralist}
-\usepackage{xr}
+\usepackage{etoolbox}
+\usepackage{filecontents}
+\usepackage{url}
 \externaldocument{SI}
@@
 hedgerows (N=5), as well as in two types of non-assembling communities
-to serve as controls: unrestored, weedy field margins (N=XX) and
+to serve as controls: unrestored, weedy field margins (N=19) and
 established hedgerows (greater than 10 years since planting,
-N=XX). The sites were located in the Central Valley of California in
+N=29). The sites were located in the Central Valley of California in
 Yolo, Colusa and Solano Counties. This area is comprised of
@@
 logistical reasons, no sampling of assembling hedgerows was conducted
-in 2010. Sites were sampled between two and five times per year (Table
-XX). In each round of sampling, the order in which sites were sampled
-was randomized. Surveys were conducted under sunny conditions when the
+in 2010. Sites were sampled between two and five times per year
+(Tables S1-S3). In each round of
+sampling, the order in which sites were sampled was
+randomized. Surveys were conducted under sunny conditions when the
 temperature was above $21^{\circ}\mathrm{C}$ and wind speed was below
@@
 flower were collected; however, here we focus only on wild bees and
-syrphids (representing XX and XX percent of records, respectively),
+syrphids (representing 49 and 19 percent of records, respectively),
 the most abundant and effective pollinators in the system (C. Kremen,
@@
 estimate the null distribution \citep{peel2014detecting}. The change
-point analysis was carried out using code published online by L.~Peel.
+point analysis was carried out using code published online by
+L.~Peel. Analyses we conducted in Python 3.4.
 
@@
 non-assembling hedgerows and weedy field margins, only sites with five
-or greater years of sampling was included in this analysis.
+or greater years of sampling was included in this analysis. All
+statistical analysis were conducted in R 3.2.3 \citep{R}.
 
@@
 of species closeness values was related to the persistence or degree
-of that species. We included random effects for species, as well as
-site. We focused on the pollinator species because the hedgerow
-flowers are planted and thus are not directly assembling. Because
-degree and persistence were strongly correlated, ($\rho = 0.84$,
-$p$-value $<$ $2*10^{-16}$), each explanatory variable was included in
-the model separately.
+of that species \citep{lme4, lmetest}. We included random effects for
+species, as well as site. We focused on the pollinator species because
+the hedgerow flowers are planted and thus are not directly
+assembling. Because degree and persistence were strongly correlated,
+($\rho = 0.84$, $p$-value $<$ $2*10^{-16}$), each explanatory variable
+was included in the model separately. Because a linear increase in
+closeness, as might be expected with assembly by preferential
+attachment, would lead to a high variability in closeness scores, we
+also test whether closeness increases through time.
 
-\subsection*{Temporal changes in interaction patterns}
+\section*{Temporal changes in interaction patterns}
+\subsection*{Network structure}
 The changing points in network structure may contribute to the
@@
 
-Lastly, we test the changes in interaction patterns associated with
-network assembly affect the robustness. First, following
+Lastly, we test whether the changes in interaction patterns associated
+with network assembly affect the robustness. First, following
 \cite{Memmott2004}, we simulate the extinction of plant species the
@@
 \label{sec:results}
+
+Over eight years and $747$ samples, we collected and identified
+$19,547$ wild bees and syrphids comprising $173$ species from $50$
+genera. We observed $1,521$ unique interactions between plants and
+pollinators.
+
 \subsection*{Change point analysis}
 \subsubsection*{Identifying change points}
+
+The majority ($76\%$) of the sites tests underwent at least one
+significant reorganization of interactions (Fig.~
+\ref{fig:changePoints}). There were no consistent trends as to when
+change points occurred within assembling hedgerows or across all sites,
+except many site had changing points between year 2009 and 2011 (Fig.~
+\ref{fig:changePoints}). In California, 2011 marked the beginning of a
+multi year drought. In the assembling hedgerows were not sampled in
+2010, so disentangling whether the changing points are due to skipping
+a year of assembly or the drought is not impossible.
+
+All five of the assembling hedgerows experienced changing points,
+whereas only $40\%$ and $81\%$ of non-assembling hedgerows and field
+margins, respectively, underwent significant interaction
+reorganizations. Assembling hedgerows experienced significantly more
+changing points than the non-assembling networks (difference in the
+odds ratios between assembling and non-assembling networks $\pm$
+$95\%$ CI, $3.316 \pm 1.314 8.572$, $p$-value= $0.0117$). Network
+assembly following restoration is thus punctuated by more interaction
+reorganizations than would be expected by environmental shifts that
+would effect assembling and non-assembling networks equally.
 
@@
   \includegraphics[width=.8\textwidth]{../analysis/changePoint/plotting/networks.pdf}
-  \caption{XXX}
+  \caption{The network structure and changing points (vertical red
+    lines) in assembling hedgerows and a representative sample of
+    non-assembling hedgerows and weedy field margins. In each network,
+    plants and pollinators are represented by green and yellow
+    circles, respectively, weighted by their degree. Each species has
+    has a consistent position in the network across years. In the
+    assembling hedgerows, colored circles in the corner of each
+    network represent the years post restoration.}
   \label{fig:changePoints}
@@
 
+In contradiction to the predictions of assembly by preferential
+attachment, both pollinator persistence and degree were positively
+related to network position variability. (ADD STATS IF KEEPING
+RESULT).
+
 \begin{figure}
@@
   \includegraphics[width=.8\textwidth]{../analysis/variability/figures/cv/occ_degree.pdf}
-  \caption{XXX}
+  \caption{The coefficient of variation of network position, as
+    represented by closeness, plotted against pollinator persistence
+    and degree. Points represents means for each species across sites.
+    The solid line indicates the mean slope estimate and the dashed
+    lines are the $95\%$ confidence intervals around the estimate. }
   \label{fig:cv}
@@
 
+\section*{Temporal changes in interaction patterns}
 \subsection*{Network structure}
+Network nestedness significantly increased with assembly (estimate of
+the slope of nestedness through time $\pm$ standard error of the
+estimate, $1.834 \pm 0.6142$, $p$-value=$0.022$, Fig.~
+\ref{fig:baci}).  Modularity decreased (Fig.~ \ref{fig:baci}), though
+the slope was not significantly different from zero (estimate of the
+slope of modularity through time $\pm$ standard error of the estimate,
+$-0.524$ $\pm$ $0.295$, $p$-value=$0.124$). Specialization remained
+relatively constant (estimate of the slope of specialization through time
+$\pm$ standard error of the estimate, $0.003$ $\pm$ $0.015$,
+$p$-value=$0.827$).
 
@@
   \centering
-  \includegraphics[width=.4\textwidth]{../analysis/networkLevel/figures/baci.pdf}
-  \caption{XXX}
+  \includegraphics[width=.7\textwidth]{../analysis/networkLevel/figures/baci.pdf}
+  \caption{The change in nestedness, modularity, specialization and
+    connectance as the networks assemble. The left panels represent
+    $z$-scores. Scores greater than $\sim 2$ or less than $\sim -2$
+    are significantly more or less structured than randomly assembled
+    networks. Points are the value for each site at each year of
+    assembly. The solid line indicates the mean slope estimate and the
+    dashed lines are the $95\%$ confidence intervals around the
+    estimate.}
   \label{fig:baci}
@@
 \subsection*{Network robustness}
+Assembly did not effect the robustness of the networks to species
+extinction when species where removed incrementally by degree
+(estimate of the slope of robustness through time $\pm$ standard error
+of the estimate, $6*10^{-5} \pm 4*10^{-3}$, $p$-value=$0.987$) or
+abundance ($0.001 \pm 0.003$, $p$-value=$0.65$, Fig.~ \ref{fig:rob}).
+
+In contrast, the robustness of networks to perturbation, as measured
+by the algebraic connectivity of the network, increased as the network
+assembled (estimate of the slope of robustness through time $\pm$
+standard error of the estimate, $0.6814 \pm 0.272$, $p$-value=$0.042$,
+Fig.~ \ref{fig:rob}).
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.7\textwidth]{../analysis/networkLevel/figures/robustness.pdf}
+  \caption{The robustness of networks to species extinction and
+    perturbation. The robustness to species extinction is measured by
+    incrementally removing species by degree, through removing
+    species by abundance did not yield qualitatively different
+    results. Points are the value for each site at each year of
+    assembly. The solid line indicates the mean slope estimate and the
+    dashed lines are the $95\%$ confidence intervals around the
+    estimate.}
+  \label{fig:rob}
+\end{figure}
+\clearpage
 
@@
 \bibliography{refs}
-
-\end{document}
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-PDF-mode: t
-%%% End:
 
@@
 % \citep{lme4,lmetest}.
+
+\end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-PDF-mode: t
+%%% End:

### 2016-08-11 analysis/changePoint/comparisons.R

@@ -9,31 +9,48 @@ BACI.site <- c('Barger', 'Butler', 'Hrdy', 'MullerB', 'Sperandio')
 ## **********************************************************
 ## binomial 
 ## **********************************************************
-count.samples <- aggregate(spec$Year, list(Site=spec$Site),
+chpt.trial <- aggregate(spec$Year, list(Site=spec$Site),
                            FUN=function(x) length(unique(x)))
-count.samples$x <- count.samples$x - 1
+chpt.trial$x <- chpt.trial$x - 1
+
+chpt.trial <- chpt.trial[chpt.trial$x >= 4,]
 
 change.points.site <- tapply(dats$cp, dats$sites, length)
 
-chpt.trial <- data.frame(chpts=change.points.site,
-                         trials=count.samples$x[match(
-                           rownames(change.points.site),
-                           count.samples$Site)])
+chpt.trial$chpts <- change.points.site[match(chpt.trial$Site,
+                                                rownames(change.points.site))]
+chpt.trial$chpts[is.na(chpt.trial$chpts)] <- 0
+colnames(chpt.trial) <- c("Site", "trial", "chpts")
 
-chpt.trial$status <- spec$SiteStatus[match(rownames(chpt.trial), spec$Site)]
-chpt.trial$status[rownames(chpt.trial) %in% BACI.site] <- "maturing"
+chpt.trial$status <- spec$SiteStatus[match(chpt.trial$Site, spec$Site)]
+chpt.trial$status[chpt.trial$Site %in% BACI.site] <- "maturing"
 
 ## binomial model with change points as successes
-mod.chpt <- glm(cbind(chpt.trial$chpts, chpt.trial$trials - chpt.trial$chpts) ~
+mod.chpt <- glm(cbind(chpt.trial$chpts, chpt.trial$trial - chpt.trial$chpts) ~
     chpt.trial$status, family="binomial")
 summary(mod.chpt)
+
+exp(cbind(coef(mod.chpt), confint(mod.chpt)))  
+
+# controls with change points
+nrow(chpt.trial[chpt.trial$status == "control" &
+                chpt.trial$chpts != 0,])/
+  nrow(chpt.trial[chpt.trial$status == "control",])
+
+nrow(chpt.trial[chpt.trial$status == "mature" &
+                chpt.trial$chpts != 0,])/
+  nrow(chpt.trial[chpt.trial$status == "mature",])
+
[…]

### 2016-08-11 analysis/networkLevel/baci.R

@@ -3,27 +3,28 @@ setwd('~/Dropbox/hedgerow_assembly/analysis/networkLevel')
 source('src/initialize.R')
 load('../../data/networks/baci_networks_years.Rdata')
 N <- 999
-## ************************************************************
-## calculate metrics and zscores
-## ************************************************************
-mets <- lapply(nets, network.metrics, N)
 
-cor.dats <- prep.dat(mets,  spec)
+## ## ************************************************************
+## ## calculate metrics and zscores
+## ## ************************************************************
+## mets <- lapply(nets, network.metrics, N)
 
-## ************************************************************
-## niche overlap
-## ************************************************************
+## cor.dats <- prep.dat(mets,  spec)
 
-no <- t(sapply(nets, networklevel, index="niche overlap"))
+## ## ************************************************************
+## ## niche overlap
+## ## ************************************************************
 
-cor.dats$niche.overlap.pol <- no[, "niche.overlap.HL"][match(rownames(no),
-                                     paste(cor.dats$Site,
-                                     cor.dats$Year, sep="."))]
-cor.dats$niche.overlap.plants <- no[, "niche.overlap.LL"][match(rownames(no),
-                                     paste(cor.dats$Site,
-                                     cor.dats$Year, sep="."))]
+## no <- t(sapply(nets, networklevel, index="niche overlap"))
 
-save(cor.dats, file='saved/corMets.Rdata')
+## cor.dats$niche.overlap.pol <- no[, "niche.overlap.HL"][match(rownames(no),
+##                                      paste(cor.dats$Site,
+##                                      cor.dats$Year, sep="."))]
+## cor.dats$niche.overlap.plants <- no[, "niche.overlap.LL"][match(rownames(no),
+##                                      paste(cor.dats$Site,
+##                                      cor.dats$Year, sep="."))]
+
+## save(cor.dats, file='saved/corMets.Rdata')
 
 ## ************************************************************
 ## effect of years post restoration
@@ -39,7 +40,7 @@ summary(baci.nodf.mod)
[…]

### 2016-08-11 analysis/networkLevel/plotting/baci.R

@@ -28,11 +28,11 @@ nodf.pi <- predict.int(mod= baci.nodf.mod,
 ## ************************************************************
 load(file='saved/mods/baci_mod.Rdata')
 
-dd.mod <- cbind(dd, zmodD=0)
+dd.mod <- cbind(dd, zmod.met.D=0)
 
 mod.pi <- predict.int(mod= baci.mod.mod,
                         dd=dd.mod,
-                        y="zmodD",
+                        y="zmod.met.D",
                       family="gaussian")
 
 ## ************************************************************
@@ -68,7 +68,20 @@ dd.nopol <- cbind(dd, niche.overlap.pol =0)
 nopol.pi <- predict.int(mod= baci.no.pol.mod,
                         dd=dd.nopol,
                         y="niche.overlap.pol",
-                        family="gaussian")                        
+                        family="gaussian")
+
+## ************************************************************
+## connectance
+## ************************************************************
+load(file='saved/mods/baci_conn.Rdata')
+
+dd.conn <- cbind(dd, connectance =0)
+conn.pi <- predict.int(mod= baci.conn.mod,
+                        dd=dd.conn,
+                        y="connectance",
+                        family="gaussian")
+
+
 ## ************************************************************
 ## all panels
 ## ************************************************************

### 2016-08-11 analysis/networkLevel/plotting/src/plotPanels_resilence.R

@@ -1,0 +1,36 @@
+plot.panels <- function(){
+  f <- function(){
+    col.lines <- brewer.pal(4, "Greys")[3]
+    col.fill <- add.alpha(col.lines, alpha=0.2)
+    layout(matrix(1:2, ncol=1))
+    par(oma=c(4, 1, 2, 1),
+        mar=c(1, 6, 0.5, 1), cex.axis=1.5)
+    ## nodf
+    plot.panel(new.dd=ypr.pi,
+               dats=res,
+               y1="Robustness",
+               xs="ypr",
+               col.fill=col.fill,
+               col.lines=col.lines,
+               plotPoints=TRUE)
+    mtext("Robustness to \n species extinction", 2, line=4, cex=1.5)
+    ## modularity
+    plot.panel(new.dd=ypr.pi.alg,
+               dats=all.alg.Con.status,
+               y1="AlgCon",
+               xs="ypr",
+               col.fill=col.fill,
+               col.lines=col.lines,
+               plotPoints=TRUE)
+    mtext("Robustness to \n perturbation", 2, line=4, cex=1.5)
+
+    axis(1, pretty(res$ypr), labels=pretty(res$ypr))
+    mtext("Year of assembly", 1, line=3.5, cex=1.5)
+  }
+  path <- 'figures'
+  pdf.f(f, file=file.path(path,
+             sprintf("%s.pdf", "robustness")),
+        width=4, height=5.5)
+
+}
+

### 2016-08-11 analysis/networkLevel/resilience.R

@@ -4,7 +4,7 @@ source('src/initialize.R')
 load('../../data/networks/all_networks_years.Rdata')
 
 ## either "abund" or "degree"
-extinction.method <- "degree"
+extinction.method <- "abund"
 
 ## **********************************************************
 ## robustness

### 2016-08-11 analysis/networkLevel/src/CalcMetrics.R

@@ -32,8 +32,10 @@ calc.metric <- function(dat.web, H2_integer=FALSE) {
                                                                 weights=
                                                                 weights)),
                           weights=weights)
-    return(c(greedy, random.walk, dendro))
-  }
+    return(c(G=greedy,
+             R=random.walk,
+             D=dendro))
+  } 
 
   dat.web <- as.matrix(empty(dat.web))
   ## matrix of all the same number
@@ -46,8 +48,8 @@ calc.metric <- function(dat.web, H2_integer=FALSE) {
                        wbinary=TRUE)$statistic["NODF"]
     mets <-  c(nodf,
                networklevel(dat.web,
-                      index=c("H2", "connectance", "weighted connectance"),
-                      H2_integer=H2_integer))
+                            index=c("H2", "connectance", "weighted connectance"),
+                            H2_integer=H2_integer))
   }
   mod.met <- calc.mod(dat.web)
   return(c(mets, mod.met= mod.met))
@@ -137,15 +139,17 @@ network.metrics <- function (dat.web, N) {
         out.mets <- cbind(true.stat, null.stat)
         ## compute z scores
         zvalues <- zvals(out.mets)
+        names(zvalues) <- paste("z", names(true.stat), sep="")
         ## compute p-values
         pvalues <- pvals(out.mets, N)
+        names(pvalues) <- paste("p", names(true.stat), sep="")
         out <- c(true.stat, zvalues, pvalues)
-        names(out) <- c("NODF", "H2", "connectance", "weighted connectance",
-                        "modularityG", "modularityR","modularityD",
-                        "zNODF", "zH2", "zcon", "zWcon",
-                        "zmodG", "zmodR", "zmodD",
-                        "pNODF", "pH2", "pcon", "pWcon",
-                        "pmodG", "pmodR", "pmodD")
+        ## names(out) <- c("NODF", "connectance", "weighted.connectance", "H2",
+        ##                 "modularityG", "modularityR","modularityD",
+        ##                 "zNODF", "zcon", "zWcon", "zH2",
+        ##                 "zmodG", "zmodR", "zmodD",
+        ##                 "pNODF", "pcon", "pWcon", "pH2",
+        ##                 "pmodG", "pmodR", "pmodD")
         return(out)
       }
     }
@@ -157,9 +161,9 @@ prep.dat <- function(cor.stats, spec.dat){
   dats <- do.call(rbind, cor.stats)
   out <- data.frame(dats)
[…]

### 2016-08-11 ms/ms.tex
finished writing results and methods

@@
   explore the assembly of plant-pollinator communities on native plant
-  restorations in the Central Valley of California. The assembling
-  communities are paired with unrestored field margins (controls) and
-  mature (non-assembling) hedgerows. We determine whether there are
-  change points in the assembly of the communities where the network
-  undergoes significant reorganization. We are also ask how are the
-  individual species changing their interaction patterns? What does
-  this mean for the topology/resilience of the network? We also
-  attempted to adapt a financial model to mutualistic networks. Our
-  biggest difficulty with this approach was to translate the price
-  term to mutualistic systems. We explored a range of approaches, such
-  as number of visits a species performs. However, it seems that
-  financial systems cannot be easily translated to mutualistic
-  systems. In addition, we used a Changing Point Detection Algorithm
-  to assess weather the different communities went through a critical
-  reorganization on their interaction patterns. We were able to
-  identify some changing points in the communities, and also to
-  explore some general patterns commonly used to describe ecological
-  networks. For example, on the network level, networks become
-  increasingly modular and less nested, whereas on the species level,
-  species become more specialized, as resources become more reliable.
+  restorations in the Central Valley of California. 
 \end{abstract}
@@
 sites where we collected flower visitors using the same methods
-\citep{mgonigle-2015-x, ponisio2015farm}.%% does this make sense to
+\citep{mgonigle-2015-x, ponisio2015farm}. %% does this make sense to
 %% calculate from an independent data set?
@@
 changing points than the non-assembling networks (difference in the
-odds ratios between assembling and non-assembling networks $\pm$
-$95\%$ CI, $3.316 \pm 1.314 8.572$, $p$-value= $0.0117$). Network
-assembly following restoration is thus punctuated by more interaction
+odds ratios between assembling and non-assembling networks, $3.316$,
+$95\%$ CI $[1.314, 8.572]$, $p$-value= $0.0117$). Network assembly
+following restoration is thus punctuated by more interaction
 reorganizations than would be expected by environmental shifts that

### 2016-08-11 ms/ms.tex
finished writing results and methods

@@
 reorganization of the assembling networks into predictable interaction
-patterns. Pollination networks exhibit two main topologies ---
-modularity \citep[e.g.,][]{Olesen2007} and nestedness
-\citep[e.g.,][]{Bascompte2006, Bascompte2003}. Modular community
+patterns. Pollination networks exhibit two main structural patterns
+--- modularity \citep[e.g.,][]{Olesen2007} and nestedness
+\citep[e.g.,][]{Bascompte2006, Bascompte2003}. In modular networks,
 interactions are insular, occurring within separate groups or
@@
 and so on. If species entering the network tend to interact with the
-generalist base of the network pyramid (i.e., preferential
+generalist base of the network pyramid (i.e., via preferential
 attachment), nestedness would increase through time. Lastly, if the

### 2016-08-11 ms/ms_ma.tex
finished writing results and methods

--- 2016-08-11 ms/ms.tex
@@
 \usepackage[parfill]{parskip}
-\usepackage{graphicx}
-\usepackage{longtable}
-\usepackage{times}
-\usepackage{textcomp}
 \usepackage{xr}
-\usepackage{etoolbox}
-\usepackage{filecontents}
-\usepackage{url}
+\usepackage{paralist}
+\usepackage{xr}
 \externaldocument{SI}
@@
   following restoration}
-\def\author{Lauren C.\ Ponisio$^{1,2}$, Marilia P. Gaiarsa$^3$, Claire
-  Kremen$^1$}
+\def\author{Lauren C.\ Ponisio$^1$, Marilia P. Gaiarsa$^2$, Claire Kremen$^1$}
 
@@
     94720\\
-  \item Department of Entomology\\
-    University of California, Riverside\\
-    417 Entomology Bldg.\\
-    Riverside, California, USA\\
-    92521\\
   \item Departamento de Ecologia\\
@@
 Keywords: changing points, temporal networks, hedgerows, species
-interactions, preferential attachment, mutualisms
+interactions, preferential attachment
 
@@
 \label{sec:introduction}
-
-%% introduce mutualisms, restoration, network assembly
-Global change has created a severe biodiversity crisis, and as species
-are lost, so are their interactions \citep{dunn2009sixth,
-  barnosky2011has}. Because mutualistic interactions are essential for
-maintaining the diversity their component guilds of species, these
-systems are particularly at risk from coextinction cascades. The
-nature of these coextinction cascades depends on the interaction
-patterns within a community \citep{Memmott2004, Rezende2007,
-  Bascompte2009}. Recovering the lost biodiversity and interactions
-through ecological restoration has become increasingly imperative, and
-a key restoration aim is to facilitate assembly of robust interaction
-networks \citep{menz-2010-4}. We know little, however, about how to
-re-assemble interacting communities through restoration, or the
-process of ecological network assembly more generally.
-
-%% most people think network assemble via preferential attachment, and
-%% there is some evidence of that. (other studies on mutualisms?)
-The mostly widely explored mechanism of network assembly, preferential
-attachment \citep{barabasi1999emergence}, predicts that a new species
-is more likely to interact with species that are already
-well-connected \citep[''the rich-get-richer''
-principle,][]{barabasi1999emergence}. In pollination systems --- a
-particularly ubiquitous mutualism \citep{ollerton-2011-321,
-  klein-2007-303} --- some studies have found support for this
-mechanism of assembly. Investigating the day-to-day, temporal assembly
-of a plant-pollinator network within a season, \cite{Olesen2008} found
-that new species tended to interact with already well-connected
-species, likely because these species are either more abundant or more
-temporally persistent \citep{Olesen2008}. In addition, using a
-space-for-time substitution to study primary succession along a
-glacier foreland, \cite{albrecht2010plant} found some indication
-assembly was occurring through preferential attachment. Network
-nestedness, a pattern of interactions where a generalist core
-interacts with both specialist and generalist species, increased as
-the community aged \citep{albrecht2010plant}. Increasing nestedness
-could result from a process like preferential attachment where
-specialist species attach to the well-connected, generalist core.
-
-%% there are other ways to assemble that are not preferentia;
-%% attchment, particularly vis big reorganizations of interactions
-In contrast to the ordered network build-up described by preferential
-attachment, assembly can be punctuated by significant reorganizations
-of interactions \citep{peel2014detecting}. For example, as new species
-are added, resident species change their interaction partners to
+\begin{itemize}
+\item The structure of networks is related to ability of communities
+  to maintain function in the face of species extinction.
+\item A key restoration aim is to facilitate assembly of robust
+  networks; thus it is critical to study how restoration
+  influences the assembly of plant-pollinator interactions.
+\item In general, however, few mechanisms of network assembly have
+  been developed and analyzed.
+\item The mostly widely explored mechanism of network assembly,
+  preferential attachment \citep{barabasi1999emergence}, predicts that
+  a new species (or node more generally) is more likely to interact
+  with species that are already well-connected \citep[''the
+  rich-get-richer'' principle,][]{barabasi1999emergence}.
+\item To date, only a handufullof field studies have examined how networks
+  assemble over time, often using space for time gradients.
+  %I think it is better to saw few, because there are some others like Olesen et al. 2011 and refs in there (7-13).  
+\item \cite{Olesen2008} was investigated day-to-day, temporal assembly
+  of a plant-pollinator network within a season, taking advantage of
+  the extreme seasonality of pollinator communities in Greenland.
+  \cite{Olesen2008} found that within a season, the network assembly
+  was similar to preferential attachment. New species tended to
+  interact with already well-connected species, likely because these
+  species are either more abundant or more temporally persistent.
+\item Studying primary succession along a glacier foreland,
+  \cite{albrecht2010plant} found a similar pattern where nestedness, a
+  pattern of interactions where a generalist core interacts with both
+  specialist and generalist species, increased as the community
+  aged. % Studying ``managed succession'' of a clear-cut pine forest,
+  % \cite{devoto2012understanding} found changes in network structure
+  % were explained by a combination of age, tree density and variation
+  % in tree diameter. 
+\item Even non-successional temporal dynamics suggest a stable core of
+  generalists persist despite high turnover of peripheral species
+  \citep{fang2012relative, diaz2010changes, alarcon2008year}.
+\item In contrast to the ordered network build-up described by
+  preferential attachment, assembly may be punctuated by significant
+  reorganizations of interactions. For example, as new species are
+  added, resident species change their interaction partners to
   minimize competition, or become extinct. Such significant
   reorganizations of interactions, or changing points, have been
-observed in networks responding to abrupt shifts in the behavior of
-interacters \citep{peel2014detecting}. No studies, however, have
-examined whether changing points occur during ecological network
-assembly.
-
-%% restoration
-Understanding network assembly is particularly relevant to ecological
-restoration, which is essentially 'applied succession'
-\citep[e.g.,]{parker1997scale}. In pollination systems, the time since
-an area was restored has been shown to effect the structure of
-networks \citep{forup-2008-742, forup2008restoration,
-  devoto2012understanding}, suggesting interactions are evolving as the
-community develops. Understanding the mechanisms of network assembly
-will help to guide the restoration of particular communities.
-
-%% restoration of ecosystem services is really important, particularly
-%% in ag. but ag is a bad place for pollinators
-Facilitating effective restoration of networks is particularly
-imperative in areas where species interactions provide essential
-ecosystem services, such as crop pollination. In intensively managed
-agricultural landscapes, the demand for pollination services is the
-greatest \citep{kremen-2008-10}. However, honey bees, managed
-extensively around the world to provide crop pollination, are in
-global decline \citep{neumann-2010-1, van-engelsdorp-2009-e6481}. In
-addition, native pollinators, which have the capacity to provide
-sufficient crop pollination \citep{kremen-2002-16816,
-  winfree-2007-1105, kremen-2004-1109}, are in short supply because
-these landscapes make poor habitats for pollinator populations
-\citep{kremen-2002-16816}. To ensure provision the continued provision
-of ecosystem services and curb biodiversity loss, effective
-restoration of pollinators and their interactions in agricultural
-landscapes is critical.
-
-%% hedgerows!!  
-To promote pollinator services in agriculture, farmers are
-increasingly turning to the habitat restoration technique of planting
-strips of native plants along farm edges (hedgerows) to help provide
-habitat for pollinators without removing arable land from
-production. Hedgerows have been shown to augment the richness,
-abundance and trait diversity of pollinators in agricultural
-landscapes\citep{morandin-2013-829, mgonigle-2015-x, kremen-2015-602,
-  ponisio2015farm}. In addition, hedgerows promote the persistence and
-colonization of floral resource specialists
-\citep{mgonigle-2015-x}. Little is known however, about the assembly
-of the network following hedgerow restoration.
-
-Using a long-term data-set of plant-pollinator communities assembling
-following hedgerow restoration in the highly simplified and
-intensively managed agricultural landscape of California's Central
-Valley, we explore the process of network development. We first
-determine whether the mechanism underlying network assembly is a
-smooth build up of interactions as would be predicted by preferential
-attachment, or punctuated by significant reorganizations of
-interactions (i.e., changing points). Even with changing points in
-interaction organization, networks could still be assembling via
-preferential attachment if the network reorganizations were primarily
-driven the by peripheral, temporally variable species while a stable,
-well-connected core of species still persists. We thus examine whether
-the species are most variable in their network position --- and thus
-important contributors network reorganizations --- are less persistent
-and connected species. %% interaction beta div, time beta div??
-Lastly, we examine whether networks are assembling toward predictable
-interaction patterns, and the ramifications for the robustness of the
-networks to species extinction and perturbation.
-
-%% objectives
-% 1) Are hedgerows assembling by preferential attachment, or are there
-% more abrupt reorganizations of network structure?
-% 2) Who are the species responsible for interaction reorganizations?
-% how to incorporate this? 
-% 3) Are the changing points a product of turnover in species
-% composition, or are more substantial changes in the verticles of the
-% networks?
-% 4) Is the structure of interactions changing? Does this change the
-% robustness of the network to species extinction and/or perturbation?
-
-% Pollinators are one such key group of interactors: $75\%$ of all
-% crop species depend to some extent on pollinator visitation
-% \citep{klein-2007-303}, and animal-pollinated crops supply a large
-% proportion of essential nutrients to the human diet
-% \citep{eilers-2011-e21363}.
-
-% Even non-successional temporal dynamics suggest a stable core of
-% generalists persist despite high turnover of peripheral species
-% \citep{fang2012relative, diaz2010changes, alarcon2008year}.
+  observed in networks \citep{peel2014detecting}.
+\end{itemize}
 
@@
 hedgerows (N=5), as well as in two types of non-assembling communities
-to serve as controls: unrestored, weedy field margins (N=19) and
+to serve as controls: unrestored, weedy field margins (N=XX) and
 established hedgerows (greater than 10 years since planting,
-N=29). The sites were located in the Central Valley of California in
-Yolo, Colusa and Solano Counties. This area is comprised of
+N=XX). The sites were located in the Central Valley of California in
+Yolo, Colusa and Solano Counties. The area is comprised of
 intensively managed agriculture -- primarily monocultures of
-conventional row crops, vineyards and orchards. Hedgerows we planted
+conventional row crops, vineyards and orchards. Hedgerows were planted
 along field margins where they do not remove valuable land from
@@
 \textit{Eriogonum spp.}, \textit{Baccharis spp.}, \textit{Salvia
-  spp}. and others \citep{kremen-2015-602, mgonigle-2015-x}. The mean
-distance between monitoring sites was 15 km, and the minimum distance
-between sites of the same type sampled in the same year was 2 km.  The
-entire area surveyed spanned almost 300 km$^2$. The crop fields
-adjacent to all sites were similarly managed as intensive, high-input
-monoculture.
+  spp}. and others. The mean distance between monitoring sites was 15
+km, and the minimum distance between sites of the same type sampled in
+the same year was 2 km.  The entire area surveyed spanned almost 300
+km$^2$ and the crop fields adjacent to all sites were similarly managed
+as intensive, high-input monoculture.
 
@@
 logistical reasons, no sampling of assembling hedgerows was conducted
-in 2010. Sites were sampled between two and five times per year
-(Tables S1-S3). In each round of
-sampling, the order in which sites were sampled was
-randomized. Surveys were conducted under sunny conditions when the
+in 2010. Sites were sampled between two and five times per year (Table
+XX). In each round of sampling, the order in which sites were sampled
+was randomized. Surveys were conducted under sunny conditions when the
 temperature was above $21^{\circ}\mathrm{C}$ and wind speed was below
@@
 flower were collected; however, here we focus only on wild bees and
-syrphids (representing 49 and 19 percent of records, respectively),
+syrphids (representing XX and XX percent of records, respectively),
 the most abundant and effective pollinators in the system (C. Kremen,
@@
 taxonomists.
-
-Quantitative networks were generative for each site through time. To
-account for the unequal number of surveys between years, the mean of
-the interactions between a pair of plants and pollinators across
-surveys within a year was used as a representation of the frequency of
-interactions.
 
@@
 fragmentation or formation of communities (also called modules or
-compartments). Change point detection methods have yet to be
-generalized to quantitative networks, so for this analysis we focused
-on qualitative (binary) networks. Following \cite{peel2014detecting},
-we first defined a probability distribution over the networks using
-the generalized hierarchical random graph model (GHRG). The GHRG model
-is able to capture both assortative and disassortative community
-structure patterns at all scales in the network
-\citep{peel2014detecting}. A network $G$ is composed of vertices $V$
-and edges $E \subseteq {V × V }$. The GHRG model decomposes the $N$
-vertices into a series of nested groups, the relationships among which
-are represented by the dendrogram $T$. The tips of $T$ are the
-vertices of $G$, and the probability that two vertices $u$ and $v$
-connect is given by the parameter $p_r$. The probability distribution
-of the network $G$ thus defined as:
+compartments). Following \cite{peel2014detecting}, we first defined a
+probability distribution over the networks using the generalized
+hierarchical random graph model (GHRG). The GHRG model is able to
+capture both assortative and disassortative community structure
+patterns at all scales in the network \citep{peel2014detecting}. A
+network $G$ is composed of vertices $V$ and edges $E \subseteq {V × V
+}$. The GHRG model decomposes the $N$ vertices into a series of nested
+groups, the relationships among which are represented by the
+dendrogram $T$.  The tips of $T$ are the vertices of $G$, and the
+probability that two vertices $u$ and $v$ connect is given by the
+parameter $p_r$. The probability distribution of the network $G$ thus
+defined as:
 
@@
 distribution of bipartitions, from which a consensus tree is derived
-\citep{peel2014detecting}. We used $\beta$ distributions with the
-hyperparameters $\alpha=\beta=1$ to define priors for $p_r$.
+\citep{peel2014detecting}. ADD NEWMAN PAPER. We used $\beta$
+distributions with the hyperparameters $\alpha=\beta=1$ to define
+priors for $p_r$.
 
@@
 estimate the null distribution \citep{peel2014detecting}. The change
-point analysis was carried out using code published online by
-L.~Peel. Analyses we conducted in Python 3.4.
+point analysis was carried out using code published online by L~.Peel.
 
@@
 generalized linear model with Binomial error to test whether the
-probability of a change point occurring varied by site type. For the
-non-assembling hedgerows and weedy field margins, only sites with five
-or greater years of sampling was included in this analysis. All
-statistical analysis were conducted in R 3.2.3 \citep{R}.
+probability of a change point occurring varied by site type.
 
-\subsubsection*{Characteristics of species that contribute to change
-  points}
-
+\subsubsection*{Characteristics of ``core'' and ``peripheral'' communities}
 To further elucidate the nature of the change points, we examine the
 characteristics of the species that contributed the reorganization of
-interactions. Some species remain in relatively similar network
-positions through time, whereas others are more variable in their role
-and thus contribute more strongly to network reorganization. We test
-the that more persistent species with the highest degree are the most
-stable in their network position, as would be expected is the networks
-were assembling via preferential attachment.
+interactions. To do so, we first generate dendrograms using the GHRG
+model before and after each change point.  We then determined which
+species belonged to the ``core'' and ``peripheral'' network
+communities at each site. The ``core'' network communities contain the
+majority of species and are more basal than the more derived, less
+specious ``peripheral'' network communities. We next use a
+Permutational Multivariate Analysis of Variance (PERMANOVA)
+\citep{anderson-2013-557} to determine whether the species
+compositions of the species that belonged to the core and peripheral
+communities differed. We also test whether core or peripheral species
+had more variability in their species compositions \citep[i.e.,
+multivariate dispersion, ][]{anderson-2011-19,
+  anderson-2006-683}.
 
-We calculate species persistence as the proportion of the surveys a
-pollinator was observed. Species observed consistently within and
-between years are thus maximally persistent. Weighted species degree
-is calculated from interaction observations observed in more extensive
-data-set from Yolo County (approx.~ 18000 interaction records) that
-included both the data included in this study and additional data from
-sites where we collected flower visitors using the same methods
-\citep{mgonigle-2015-x, ponisio2015farm}.%% does this make sense to
-%% calculate from an independent data set?
-To represent the the variability of species within networks, we
-computed the coefficient of variation of weighted closeness at each
-site through time. Closeness describes the centrality of a species in
-the network by calculating path lengths to other vertices (species) in
-the graph. We used linear mixed models to test whether the variability
-of species closeness values was related to the persistence or degree
-of that species \citep{lme4, lmetest}. We included random effects for
-species, as well as site. We focused on the pollinator species because
-the hedgerow flowers are planted and thus are not directly
-assembling. Because degree and persistence were strongly correlated,
-($\rho = 0.84$, $p$-value $<$ $2*10^{-16}$), each explanatory variable
-was included in the model separately. Because a linear increase in
-closeness, as might be expected with assembly by preferential
-attachment, would lead to a high variability in closeness scores, we
-also test whether closeness increases through time.
-
-\section*{Temporal changes in interaction patterns}
-\subsection*{Network structure}
-The changing points in network structure may contribute to the
-reorganization of the assembling networks into predictable interaction
-patterns. Pollination networks exhibit two main topologies ---
-modularity \citep[e.g.,][]{Olesen2007} and nestedness
-\citep[e.g.,][]{Bascompte2006, Bascompte2003}. Modular community
-interactions are insular, occurring within separate groups or
-``modules'' more often than between modules. Communities in the
-network may fragment as the network assembles, enhancing
-modularity. Conversely, nested networks are like pyramid of
-interactions, where there are some species that interact with many
-species, other species that interact with a subset of those species,
-and so on. If species entering the network tend to interact with the
-generalist base of the network pyramid (i.e., preferential
-attachment), nestedness would increase through time. Lastly, if the
-network is accumulating specialist species or species are beginning to
-limit their interaction niche breath as the network assembles, this
-would lead to an increase in the network-level specialization
-\citep{bluthgen-2006-9}. To test whether network modularity,
-nestedness or specialization changed linearly with assembly, we used
-linear mixed models with the descriptive network metrics as the
-response variable, year of assembly as the explanatory variable, and
-random effects of site and year.
-
-We use NODF (weighted or unweighted) to evaluate network nestedness
-\citep{nodf}. NODF evaluates whether species with fewer partners
-interact with subsets of partners with which more connected species
-interact \citep{nodf}. To estimate modularity, we use a hierarchical
-clustering algorithm \citep{Newman2004, igraph}.  We calculated
-standardized $z$-scores so that nestedness and modularity metrics
-could be compared between communities. The $z$-scores were calculated
-by generating an ensemble of $999$ randomly assembled communities,
-subtracting the mean of the statistic calculated across these
-communities from the observed value, and then dividing by the standard
-deviation. To assemble random communities, we reshuffled the
-interactions between species but fixed the total number of
-interactions, species and the distribution of the interaction
-frequencies \citep{Galeano2009}. Lastly, Network specialization was
-measured using H2, which estimate the deviation of the observed
-interaction frequency between plants and pollinators from a null
-expectation where all partners interact in proportion to their
-abundances \citep{bluthgen-2006-9}. It ranges from 0 for generalized
-networks to 1 for specialized networks.
-
-\subsection*{Network robustness}
-
-Lastly, we test whether the changes in interaction patterns associated
-with network assembly affect the robustness. First, following
-\cite{Memmott2004}, we simulate the extinction of plant species the
-subsequent extinction cascades of pollinator species. Because the
-reproduction of plant species if facilitated by active restoration
-efforts, it is unlikely the extinction of pollinators would effect
-plant populations in the hedgerows. However, plants ceasing to bloom,
-for example in response to drought, will likely effect the pollinators
-that depend on them. Plants species were eliminated based on their
-degree or abundance, and the number of pollinators that secondarily
-went extinct is calculated. The area below the extinction curve is
-used as a measure of network robustness \citep{Memmott2004}.
-
-%% Marilia add perturbation methods here
-
-\section*{Results}
-\label{sec:results}
-
-Over eight years and $747$ samples, we collected and identified
-$19,547$ wild bees and syrphids comprising $173$ species from $50$
-genera. We observed $1,521$ unique interactions between plants and
-pollinators.
-
-\subsection*{Change point analysis}
-\subsubsection*{Identifying change points}
-
-The majority ($76\%$) of the sites tests underwent at least one
-significant reorganization of interactions (Fig.~
-\ref{fig:changePoints}). There were no consistent trends as to when
-change points occurred within assembling hedgerows or across all sites,
-except many site had changing points between year 2009 and 2011 (Fig.~
-\ref{fig:changePoints}). In California, 2011 marked the beginning of a
-multi year drought. In the assembling hedgerows were not sampled in
-2010, so disentangling whether the changing points are due to skipping
-a year of assembly or the drought is not impossible.
-
-All five of the assembling hedgerows experienced changing points,
-whereas only $40\%$ and $81\%$ of non-assembling hedgerows and field
-margins, respectively, underwent significant interaction
-reorganizations. Assembling hedgerows experienced significantly more
-changing points than the non-assembling networks (difference in the
-odds ratios between assembling and non-assembling networks $\pm$
-$95\%$ CI, $3.316 \pm 1.314 8.572$, $p$-value= $0.0117$). Network
-assembly following restoration is thus punctuated by more interaction
-reorganizations than would be expected by environmental shifts that
-would effect assembling and non-assembling networks equally.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.8\textwidth]{../analysis/changePoint/plotting/networks.pdf}
-  \caption{The network structure and changing points (vertical red
-    lines) in assembling hedgerows and a representative sample of
-    non-assembling hedgerows and weedy field margins. In each network,
-    plants and pollinators are represented by green and yellow
-    circles, respectively, weighted by their degree. Each species has
-    has a consistent position in the network across years. In the
-    assembling hedgerows, colored circles in the corner of each
-    network represent the years post restoration.}
-  \label{fig:changePoints}
-\end{figure}
-\clearpage
-
-\subsubsection*{Characteristics of species that contribute to change
-  points}
-
-In contradiction to the predictions of assembly by preferential
-attachment, both pollinator persistence and degree were positively
-related to network position variability. (ADD STATS IF KEEPING
-RESULT).
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.8\textwidth]{../analysis/variability/figures/cv/occ_degree.pdf}
-  \caption{The coefficient of variation of network position, as
-    represented by closeness, plotted against pollinator persistence
-    and degree. Points represents means for each species across sites.
-    The solid line indicates the mean slope estimate and the dashed
-    lines are the $95\%$ confidence intervals around the estimate. }
-  \label{fig:cv}
-\end{figure}
-\clearpage
-
-\section*{Temporal changes in interaction patterns}
-\subsection*{Network structure}
-Network nestedness significantly increased with assembly (estimate of
-the slope of nestedness through time $\pm$ standard error of the
-estimate, $1.834 \pm 0.6142$, $p$-value=$0.022$, Fig.~
-\ref{fig:baci}).  Modularity decreased (Fig.~ \ref{fig:baci}), though
-the slope was not significantly different from zero (estimate of the
-slope of modularity through time $\pm$ standard error of the estimate,
-$-0.524$ $\pm$ $0.295$, $p$-value=$0.124$). Specialization remained
-relatively constant (estimate of the slope of specialization through time
-$\pm$ standard error of the estimate, $0.003$ $\pm$ $0.015$,
-$p$-value=$0.827$).
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.7\textwidth]{../analysis/networkLevel/figures/baci.pdf}
-  \caption{The change in nestedness, modularity, specialization and
-    connectance as the networks assemble. The left panels represent
-    $z$-scores. Scores greater than $\sim 2$ or less than $\sim -2$
-    are significantly more or less structured than randomly assembled
-    networks. Points are the value for each site at each year of
-    assembly. The solid line indicates the mean slope estimate and the
-    dashed lines are the $95\%$ confidence intervals around the
-    estimate.}
-  \label{fig:baci}
-\end{figure}
-\clearpage
-
-\subsection*{Network robustness}
-Assembly did not effect the robustness of the networks to species
-extinction when species where removed incrementally by degree
-(estimate of the slope of robustness through time $\pm$ standard error
-of the estimate, $6*10^{-5} \pm 4*10^{-3}$, $p$-value=$0.987$) or
-abundance ($0.001 \pm 0.003$, $p$-value=$0.65$, Fig.~ \ref{fig:rob}).
-
-In contrast, the robustness of networks to perturbation, as measured
-by the algebraic connectivity of the network, increased as the network
-assembled (estimate of the slope of robustness through time $\pm$
-standard error of the estimate, $0.6814 \pm 0.272$, $p$-value=$0.042$,
-Fig.~ \ref{fig:rob}).
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.7\textwidth]{../analysis/networkLevel/figures/robustness.pdf}
-  \caption{The robustness of networks to species extinction and
-    perturbation. The robustness to species extinction is measured by
-    incrementally removing species by degree, through removing
-    species by abundance did not yield qualitatively different
-    results. Points are the value for each site at each year of
-    assembly. The solid line indicates the mean slope estimate and the
-    dashed lines are the $95\%$ confidence intervals around the
-    estimate.}
-  \label{fig:rob}
-\end{figure}
-\clearpage
-
-\section*{Discussion}
-\label{sec:discussion}
-
-\section*{Acknowledgments}
-\label{sec:acknowledge}
-
-We would like to thank Leto Peel and Aaron Clauset for their
-invaluable discussions and for help with the change point analysis.
-We thank the growers and land owners that allowed us to work on their
-property.  We also greatly appreciate the identification assistance of
-expert taxonomists Martin Hauser, Robbin Thorp and Jason Gibbs.  This
-work was supported by funding from the Army Research Office
-(W911NF-11-1-0361 to CK), the Natural Resources Conservation Service
-(CIG-69-3A75-12-253, CIG-69-3A75-9-142, CIG-68-9104-6-101 and
-WLF-69-7482-6-277 to The Xerces Society), the National Science
-Foundation (DEB-0919128 to CK), The U.S.  Department of Agriculture
-(USDA-NIFA 2012-51181-20105 to Michigan State University).  Funding
-for LCP was provided by an NSF Graduate Research Fellowship and the
-USDA NIFA Graduate Fellowship. FUNDING FOR MARILLIA.
-
-\bibliographystyle{ecol_let}
-
-\bibliography{refs}
-
-% \subsection*{Core and peripherial communities}
-
-% To classify species as we first generate dendrograms using the GHRG
-% model before and after each change point.  We then determined which
-% species belonged to the ``core'' and ``peripheral'' network
-% communities at each site. The ``core'' network communities contain the
-% majority of species and are more basil than the more derived, less
-% specious ``peripheral'' network communities. We next use a
-% Permutational Multivariate Analysis of Variance (PERMANOVA)
-% \citep{anderson-2013-557} to determine whether the species
-% compositions of the species that belonged to the core and peripheral
-% communities differed. We also test whether core or peripheral species
-% had more variability in their species compositions \citep[i.e.,
-% multivariate dispersion][]{anderson-2011-19, anderson-2006-683}.
+%% keep trait analysis?
+- I think we should mention it, but present it only in the supplementary, because the other results are more interesting, we already have lots fo results.
+And also, I remember Camille (Jason's student that publishes that traits paper in EcoLet with the horrible pics) had a really hard time with the analysis, so i doint think we should give too much emphasis on it.
 
 % Lastly, we explored the functional diversity of the core and
-% peripheral pollinator species
-
-% The syrphids in our study area have similar traits, so we focused on
-% the trait diversity of the bee subset of our interaction networks. We
-% focus on resource capture and use traits that collectively
-% \citep{kremen-2015-602} including resource specialization
-% \citep[quantitative, $d'$;][]{bluthgen-2006-9}, body size
-% \citep[quantitative, inter-tegular span, mm,][]{cane-1987-145},
-% sociality (categorical: eusocial, solitary, cleptoparasitic), nest
-% location (categorical: above ground, below ground or mix), and nest
-% construction \citep[categorical: excavator, mason or
-% renter;][]{williams-2010-2280} as described in more detail in
-% \cite{kremen-2015-602}.  Each trait has the same weight in trait
-% diversity metric estimation \citep{villeger-2008-2290,
-% schleuter-2010-469}.
+% peripheral pollinator species. The syrphids in our study area have
+% similar traits, so we focused on the trait diversity of the bee subset
+% of our interaction networks. We focus on resource capture and use
+% traits that collectively influence the distribution of bee species as
+% pollinators over space and time \citep{kremen-2015-602} including
+% resource specialization \citep[quantitative,
+% $d'$;][]{bluthgen-2006-9}, body size \citep[quantitative,
+% inter-tegular span, mm,][]{cane-1987-145}, sociality (categorical:
+% eusocial, solitary, cleptoparasitic), nest location (categorical:
+% above ground, below ground or mix), and nest construction
+% \citep[categorical: excavate or rent;][]{williams-2010-2280} as
+% described in more detail in \cite{kremen-2015-602}.  Each trait has
+% the same weight in trait diversity metric estimation
+% \citep{villeger-2008-2290, schleuter-2010-469}.  Pollinator
+% specialization was calculated using plant-pollinator interaction
+% observations from a more extensive data-set from Yolo County (18000
+% interaction records) that included both the data included in this
+% study and additional data from sites where we collected flower
+% visitors using the same methods \citep{mgonigle-2015-x}.  The
+% specialization metric measures the deviation of the observed
+% interaction frequency between a plant and pollinator from a null
+% expectation where all partners interact in proportion to their
+% abundances \citep{bluthgen-2006-9}.  It ranges from 0 for generalist
+% species to 1 for specialist species.
 
@@
 
+\subsection*{Network structure}
+
+The changing points in network structure may contribute to the
+reorganization of the assembling networks into predictable interaction
+patterns. Pollination networks exhibit two main structural patterns ---
+modularity \citep[e.g.,][]{Olesen2007} and nestedness
+\citep[e.g.,][]{Bascompte2006, Bascompte2003}. In modular community
+interactions are insular, occurring within separate groups or
+``modules'' more often than between modules. Communities in the
+network may fragment as the network assembles, enhancing
+modularity. Conversely, nested networks are like pyramid of
+interactions, where there are some species that interact with many
+species, other species that interact with a subset of those species,
+and so on. If species entering the network tend to interact with the
+generalist base of the network pyramid, nestedness would increase
+through time. Alternatively, if the network is accumulating specialist % I removed the lastly because I thought it was confusing, like a third point besides Mod and Nestedness. 
+species or if species are beginning to limit their interaction niche
+breath as the network assembles, this would lead to an increase in the
+network-level specialization \citep{bluthgen-2006-9} and nestedness would decrease
+through time. To test whether
+network modularity, nestedness or specialization changed linearly with
+assembly, we used linear mixed models with the descriptive network
+metrics as the response variable, year of assembly as the explanatory
+variable, and random effects of site and year.
+
+\subsection*{Network robustness}
+Lastly, we test the changes in interaction patterns associated with
+network assembly affect the robustness of the network to species
+loss. Following \cite{Memmott2004}, we simulate the extinction of
+plant species the subsequent extinction cascades of pollinator
+species. Because the reproduction of plant species if facilitated by
+active restoration efforts, it is unlikely the extinction of
+pollinators would affect plant populations in the hedgerows. However,
+plants ceasing to bloom, for example in response to drought, will
+likely affect the pollinators that depend on them. Plants species were
+eliminated based on their degree or abundance, and the number of
+pollinators that secondarily went extinct is calculated. The area
+below the extinction curve is used as a measure of network robustness. %Maybe cite the R function we used here?
+
+\section*{Results}
+\label{sec:results}
+\subsection*{Change point analysis}
+\subsubsection*{Identifying change points}
+
+% I love this figure, but it is hard to see the different collors of plants and pollinators (in print)... 
+\begin{figure}
+  \centering
+  \includegraphics[width=.8\textwidth]{../analysis/changePoint/plotting/networks.pdf}
+  \caption{XXX}
+  \label{fig:changePoints}
+\end{figure}
+\clearpage
+
+\subsubsection*{Characteristics of ``core'' and ``peripheral''
+  communities}
+
+%Maybe use the same color to all peripheral species communities? Are they giving us any extra information?
+\begin{figure}
+  \centering
+  \includegraphics[width=.8\textwidth]{../analysis/changePoint/plotting/communities.pdf}
+  \caption{XXX}
+  \label{fig:communities}
+\end{figure}
+\clearpage
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.7\textwidth]{../analysis/changePoint/plotting/figures/pcoa/pcoa.pdf}
+  \caption{XXX}
+  \label{fig:pcoa}
+\end{figure}
+\clearpage
+
+\subsection*{Network structure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.4\textwidth]{../analysis/networkLevel/figures/baci.pdf}
+  \caption{XXX}
+  \label{fig:baci}
+\end{figure}
+\clearpage
+
+\subsection*{Network robustness}
+
+%% no need for robustness figs bec not sig
+
+\section*{Discussion}
+\label{sec:discussion}
+
+\section*{Acknowledgments}
+\label{sec:acknowledge}
+
+% I think we should thank SF and CSSS somehow... 
+We would like to thank Leto Peel and Aaron Clauset for their
+invaluable discussions and for help with the change point analysis.
+We thank the growers and land owners that allowed us to work on their
+property.  We also appreciate the identification assistance of expert
+taxonomists Martin Hauser, Robbin Thorp and Jason Gibbs.  This work
+was supported by funding from the Army Research Office
+(W911NF-11-1-0361 to CK), the Natural Resources Conservation Service
+(CIG-69-3A75-12-253, CIG-69-3A75-9-142, CIG-68-9104-6-101 and
+WLF-69-7482-6-277 to The Xerces Society), the National Science
+Foundation (DEB-0919128 to CK), The U.S.  Department of Agriculture
+(USDA-NIFA 2012-51181-20105 to Michigan State University).  Funding
+for LCP was provided by an NSF Graduate Research Fellowship and the
+USDA NIFA Graduate Fellowship. Funding for MPG was provided by Sao Paulo Research Foundation (FAPESP, grant 2013/13319-5).
+
+\bibliographystyle{ecol_let}
+
+\bibliography{refs}
+
 \end{document}

### 2016-08-15 ms/ms_ma.tex
added the methods for algebraic connectivity

@@
 network assembly affect the robustness of the network to species
-loss. Following \cite{Memmott2004}, we simulate the extinction of
+loss and to cascading effects. Following \cite{Memmott2004}, we simulate the extinction of
 plant species the subsequent extinction cascades of pollinator
@@
 below the extinction curve is used as a measure of network robustness. %Maybe cite the R function we used here?
+We also explored how the robustness to cascading effects changed as community assembled, using algebraic connectivity--- the second smallest eigenvalue of the Laplacian matrix \citep{fiedler1973algebraic}--- as a proxy for network robustness (e.g.
+Gaiarsa et al., submitted).
+Algebraic connectivity was first used to describe spectral properties of complex graphs other than ecological networks, and it is related to how difficult it is to turn a network into completely disconnected groups of nodes, or species \citep{fiedler1973algebraic, costa2007characterization}, and also to flows in .
+The larger the algebraic connectivity, the more robust a network is to perturbations (e.g.
+Gaiarsa et al., submitted).
 

### 2016-08-15 ms/ms.tex
added marilia edits

@@
 networks to 1 for specialized networks.
+\subsection*{Network structure}
+
+The changing points in network structure may contribute to the
+reorganization of the assembling networks into predictable interaction
+patterns. Pollination networks exhibit two main structural patterns
+--- modularity \citep[e.g.,][]{Olesen2007} and nestedness
+\citep[e.g.,][]{Bascompte2006, Bascompte2003}. In modular community
+interactions are insular, occurring within separate groups or
+``modules'' more often than between modules. Communities in the
+network may fragment as the network assembles, enhancing
+modularity. Conversely, nested networks are like pyramid of
+interactions, where there are some species that interact with many
+species, other species that interact with a subset of those species,
+and so on. If species entering the network tend to interact with the
+generalist base of the network pyramid, nestedness would increase
+through time. Alternatively, if the network is accumulating specialist
+species or if species are beginning to limit their interaction niche
+breath as the network assembles, this would lead to an increase in the
+network-level specialization \citep{bluthgen-2006-9} and nestedness
+would decrease through time. To test whether network modularity,
+nestedness or specialization changed linearly with assembly, we used
+linear mixed models with the descriptive network metrics as the
+response variable, year of assembly as the explanatory variable, and
+random effects of site and year.
 
 \subsection*{Network robustness}
+Lastly, we test the changes in interaction patterns associated with
+network assembly affect the robustness of the network to species loss
+and to cascading effects. Following \cite{Memmott2004}, we simulate
+the extinction of plant species the subsequent extinction cascades of
+pollinator species. Because the reproduction of plant species if
+facilitated by active restoration efforts, it is unlikely the
+extinction of pollinators would affect plant populations in the
+hedgerows. However, plants ceasing to bloom, for example in response
+to drought, will likely affect the pollinators that depend on
+them. Plants species were eliminated based on their degree or
+abundance, and the number of pollinators that secondarily went extinct
+is calculated. The area below the extinction curve is used as a
+measure of network robustness \citep{bipartite}.
 
-Lastly, we test whether the changes in interaction patterns associated
-with network assembly affect the robustness. First, following
-\cite{Memmott2004}, we simulate the extinction of plant species the
-subsequent extinction cascades of pollinator species. Because the
-reproduction of plant species if facilitated by active restoration
-efforts, it is unlikely the extinction of pollinators would effect
-plant populations in the hedgerows. However, plants ceasing to bloom,
-for example in response to drought, will likely effect the pollinators
-that depend on them. Plants species were eliminated based on their
-degree or abundance, and the number of pollinators that secondarily
-went extinct is calculated. The area below the extinction curve is
-used as a measure of network robustness \citep{Memmott2004}.
-
-%% Marilia add perturbation methods here
+We also explored how the robustness to cascading effects changed as
+community assembled, using algebraic connectivity--- the second
+smallest eigenvalue of the Laplacian matrix
+\citep{fiedler1973algebraic}--- as a proxy for network robustness
+(e.g. Gaiarsa et al., submitted). Algebraic connectivity was first
+used to describe spectral properties of complex graphs other than
+ecological networks, and it is related to how difficult it is to turn
+a network into completely disconnected groups of nodes, or species
+\citep{fiedler1973algebraic, costa2007characterization}, and also to
+flows in . The larger the algebraic connectivity, the more robust a
+network is to perturbations (e.g. Gaiarsa et al., submitted).
 

### 2016-08-16 analysis/networkLevel/laplacian.R
updated ms

@@ -43,6 +43,8 @@ all.alg.Con.status$AlgCon <- as.numeric(as.character(
   all.alg.Con.status$AlgCon))
 all.alg.Con.status$EigenRatio <- as.numeric(as.character(
   all.alg.Con.status$EigenRatio))
+all.alg.Con.status$Ncomp <- as.numeric(as.character(
+  all.alg.Con.status$Ncomp))
 
 alg.con.mod <- lmer(AlgCon ~ SiteStatus +
                     (1|Site) + (1|Year),    
@@ -51,16 +53,19 @@ alg.con.mod <- lmer(AlgCon ~ SiteStatus +
 summary(alg.con.mod)
 
 # Algebrain connectivity x Ypr
-alg.con.mod.ypr <- lmer(AlgCon ~ ypr +
+alg.con.mod.ypr <- lmer(AlgCon ~ ypr + 
                         (1|Site) + (1|Year),    
                         data=all.alg.Con.status)
 summary(alg.con.mod.ypr)
 
 # Eigenvalue ratio  x Ypr
-EigenRatio.mod.ypr <- lmer(EigenRatio ~ ypr +
+EigenRatio.mod.ypr <- lmer(EigenRatio ~ ypr +  Ncomp +
                           (1|Site) + (1|Year),    
                         data=all.alg.Con.status)
 summary(EigenRatio.mod.ypr)
 
-save(alg.con.mod.ypr, all.alg.Con.status,
+save(alg.con.mod.ypr, all.alg.Con.status, EigenRatio.mod.ypr,
      file="saved/mods/AlgCon2.Rdata")
+
+
+## look at correlation between AlgCon and species richness

### 2016-08-17 ms/ms.tex
updated ms
First draft of the discussion

@@
 important contributors network reorganizations --- are less persistent
-and connected species. %% interaction beta div, time beta div??
-Lastly, we examine whether networks are assembling toward predictable
-interaction patterns, and the ramifications for the robustness of the
-networks to species extinction and perturbation.
+and connected species.  Lastly, we examine whether networks are
+assembling toward predictable interaction patterns, and the
+ramifications for the robustness of the networks to species extinction
+and perturbation.
 
@@
 
+We show that the temporal assembly of plant pollinator communities following restauration do not occur through preferential attachment.
+Instead, the community seems to undergo through a critical internal reorganization with major changes in species composition and interaction patterns.
+
+Changing points through time = more in the assembling communities than in the control ones, as would be expected given the species turnover.
+
+Preferential attachment is the most widely accepted theory explaining community assembly in mutualistic networks.
+Communities assembling through preferential attachment are formed by a core of generalist species, to which new species arriving to the comunity interact with.
+Thus, communities assembling through preferential attachmment exhibit an increase both in nestedness and in the overall level of species specialization, since specialist species are being added to the community.
+However, our changing point analysis results show that there is a critical reorganization in the interaction patterns as communities asssembly, which contradicts the preferential attatchment theory.
+Even though we encountered an increase in nestedness through time, specialization did not increase.
+Furthermore, this increase in nestedness could be explained by the increase in species richness.
+blablabla mechanistic models.
+
+Other interesting result is that even though nestedness increased through time, modularity and connectance did not.
+\cite{Fortuna2010} suggest that at low connectance levels, mutualistic networks would present high nestedness and high modularity, while at high connectance levels networks would present low nestedness and low modularity.
+However, our results show the contrary: early communities presented low nestedness, smaller richness, and higher levels of modularity and connectance, while more mature communities presented higher nestedness, higher richness, and smaller modularity and connectance.
+SO blabla...
+XX Temporal overlap.
+
+Evidence indicates that there is a positive relatioship between abundance and diet breadth, with the more generalist species having higher abundances.
+Because of the high generalism, species more abundat also present higher degrees.
+Thus, we expected that more generalist species would be part of the network core.
+Surprinsigly, the five most common species were also the ones that exhibited the most changes in interaction patterns.
+These species (names!) were always present in the periphery of the network.
+One hipothesys is that their high generalism allows them to explore the resources more boradly, allowing them to change their position in the network.
+XX mention the cloeseness variability and the species persistence.
+
+High nestedness has been related to an increase in the robustnsess of comminities to species extinctions because the core of generalist species would act as a buffer against species loss (REF).
+Similarly, communities with greater modularity would be more robust in general, because modules would act as a buffer against perturbation spreading and the pervasive effects of species loss.
+We expected that as communities assemble and speceis richness incrase, communities would become increasingly robust to species loss.
+however, we found  that robustness to species loss did not change through time.
+This is interestingly from the conservation biology point of view, because even early assemblig communities seem to be as robust as more mature communites, despite the richness insurance hipothesis.
+However, communities are not only subjected to species loss, but also to other types of perturbations that can cascade through the network.
+Gaiarsa et al.
+(submitted) suggest useing algebraic connectivity to explore how vulnerable ecoogical communities might be to cascading effects.
+Our results indicate that more mature communities are more robust to cascading effects than early on assemblages, and that this result is related to species richness.
+This migth be related to XX....
+
 \section*{Acknowledgments}
@@
 for LCP was provided by an NSF Graduate Research Fellowship and the
-USDA NIFA Graduate Fellowship. FUNDING FOR MARILLIA.
+USDA NIFA Graduate Fellowship. Funding for MPG was provided by Sao Paulo Research Foundation (FAPESP, grant 2013/13319-5).
 

### 2016-08-18 analysis/changePoint/core.R
added story board

@@ -70,6 +70,7 @@ plot.beta.div(list(comm.mats.yr$pollinator$comm,
                    comm.mats.yr$plant$status),
               fig.path,
               method= method)
+
 ## **********************************************************
 ## who are the core species? Species that are never in the core?
 ## pollinators
@@ -80,8 +81,6 @@ stayed.pol <-
 count.stayed.pol <- table(stayed.pol)
 count.stayed.pol <- count.stayed.pol[count.stayed.pol != 0]
 always.stayed.pol <- count.stayed.pol[count.stayed.pol == 5]
-
-hist(count.stayed.pol, breaks=0:5)
 
 left.pol <-
   site.cores$all$pollinator$GenusSpecies[site.cores$all$pollinator$comm ==

### 2016-08-18 analysis/networkLevel/baci.R

@@ -2,7 +2,7 @@ rm(list=ls())
 setwd('~/Dropbox/hedgerow_assembly/analysis/networkLevel')
 source('src/initialize.R')
 load('../../data/networks/baci_networks_years.Rdata')
-N <- 999
+N <- 99
 
 ## ## ************************************************************
 ## ## calculate metrics and zscores
@@ -11,26 +11,16 @@ N <- 999
 
 ## cor.dats <- prep.dat(mets,  spec)
 
-## ## ************************************************************
-## ## niche overlap
-## ## ************************************************************
-
-## no <- t(sapply(nets, networklevel, index="niche overlap"))
-
-## cor.dats$niche.overlap.pol <- no[, "niche.overlap.HL"][match(rownames(no),
-##                                      paste(cor.dats$Site,
-##                                      cor.dats$Year, sep="."))]
-## cor.dats$niche.overlap.plants <- no[, "niche.overlap.LL"][match(rownames(no),
-##                                      paste(cor.dats$Site,
-##                                      cor.dats$Year, sep="."))]
+## cor.dats$tot.rich <- cor.dats$number.of.species.LL +
+##   cor.dats$number.of.species.HL
 
 ## save(cor.dats, file='saved/corMets.Rdata')
 
 ## ************************************************************
 ## effect of years post restoration
 ## ************************************************************
-
 load(file='saved/corMets.Rdata')
+
 
 ## nestedness
 baci.nodf.mod <- lmer(zNODF ~ scale(ypr) +
@@ -53,6 +43,28 @@ baci.h2.mod <- lmer(H2 ~ scale(ypr) +
 summary(baci.h2.mod)
 save(baci.h2.mod, file='saved/mods/baci_h2.Rdata')
 
+
+## species richness pol
+baci.rich.hl.mod <- glmer(number.of.species.HL ~ scale(ypr) +
+                 (1|Site) + (1|Year), family="poisson",
+                 data=cor.dats[!is.na(cor.dats$ypr),])
+summary(baci.rich.hl.mod)
+save(baci.rich.hl.mod, file='saved/mods/baci_rich_hl.Rdata')
+
+## species richness plants
+baci.rich.ll.mod <- glmer(number.of.species.LL ~ scale(ypr) +
[…]

### 2016-08-18 analysis/networkLevel/laplacian.R

@@ -67,5 +67,16 @@ summary(EigenRatio.mod.ypr)
 save(alg.con.mod.ypr, all.alg.Con.status, EigenRatio.mod.ypr,
      file="saved/mods/AlgCon2.Rdata")
 
-
 ## look at correlation between AlgCon and species richness
+
+net.mets <- sapply(nets, function(x) sum(dim(empty(x))))
+
+all.alg.Con.status$tot.rich <- net.mets[match(names(net.mets),
+                                              paste(all.alg.Con.status$Site,
+                                                    all.alg.Con.status$Year,
+                                                    sep="."))]
+
+cor.test(all.alg.Con.status$tot.rich, all.alg.Con.status$AlgCon)
+
+cor.test(all.alg.Con.status$tot.rich[!is.na(all.alg.Con.status$ypr)],
+         all.alg.Con.status$AlgCon[!is.na(all.alg.Con.status$ypr)])

### 2016-08-18 analysis/speciesLevel/plotting/src/CIplotting.R

@@ -60,7 +60,7 @@ plot.predict.ypr <- function(new.dd,
                              extinction.method,
                              agg.col="Site"){
   plot.ci <- function(){
-    col.lines <-  brewer.pal(3, 'Dark2')[2]
+    col.lines <-  brewer.pal(4, "Greys")[3]
     col.fill <- add.alpha(col.lines, alpha=0.2)
     layout(matrix(1, ncol=1))
     par(oma=c(6, 5, 2, 1),

### 2016-08-18 analysis/variability/src/phyloIntBeta.R

@@ -93,11 +93,16 @@ plot.box <- function(ylabel,
     col.lines <- brewer.pal(3, "Dark2")
     col.fill <- add.alpha(col.lines, alpha=0.2)
     layout(matrix(1, ncol=1))
-    par(oma=c(2, 5, 2, 1),
-        mar=c(0.5, 0, 0.5, 1))
+    par(oma=c(1, 5, 2, 1),
+        mar=c(2, 0, 0.5, 1))
     boxplot(dats[,y1]~ dats$SiteStatus,
             col=col.lines,
-            names=c("Unrestored", "Maturing", "Mature"))
+             axes = F)
+    axis(1, at= 1:3, labels= c("", "", ""), lty= 1, las= .8)
+    mtext(c("Non-assembling \n field margin",
+              "Assembling \n  hedgerow",
+                     "Non-assembling \n hedgerow"),
+          side = 1, line= 3, at = 1:3)
     axis(2, pretty(dats[,y1]), labels=pretty(dats[,y1]))
     mtext(ylabel, 2, line=3.5, cex=1.5)
   }
@@ -106,6 +111,6 @@ plot.box <- function(ylabel,
                    sprintf("%s.pdf", paste(
                      gsub(" ", "", ylabel),
                      "box", sep="_"))),
-        width=5, height=4)
+        width=6, height=4)
 
 }

### 2016-08-18 ms/ms.tex
added story board
worked on discussion

--- 2016-08-15 ms/ms.tex
@@
 the most abundant and effective pollinators in the system (C. Kremen,
-A. Klein and L. Morandin, unpublished data). Bee and syrphid specimens
+A.~ Klein and L.~ Morandin, unpublished data). Bee and syrphid specimens
 were identified to species (or morpho-species for some bee specimens
@@
 significant reorganization of interactions (Fig.~
-\ref{fig:changePoints}). There were no consistent trends as to when
-change points occurred within assembling hedgerows or across all sites,
-except many site had changing points between year 2009 and 2011 (Fig.~
-\ref{fig:changePoints}). In California, 2011 marked the beginning of a
-multi year drought. In the assembling hedgerows were not sampled in
-2010, so disentangling whether the changing points are due to skipping
-a year of assembly or the drought is not impossible.
-
-All five of the assembling hedgerows experienced changing points,
-whereas only $40\%$ and $81\%$ of non-assembling hedgerows and field
-margins, respectively, underwent significant interaction
-reorganizations. Assembling hedgerows experienced significantly more
-changing points than the non-assembling networks (difference in the
-odds ratios between assembling and non-assembling networks, $3.316$,
-$95\%$ CI $[1.314, 8.572]$, $p$-value= $0.0117$). Network assembly
-following restoration is thus punctuated by more interaction
-reorganizations than would be expected by environmental shifts that
-would effect assembling and non-assembling networks equally.
+\ref{fig:changePoints}).  All five of the assembling hedgerows
+experienced changing points, whereas only $40\%$ and $81\%$ of
+non-assembling hedgerows and field margins, respectively, underwent
+significant interaction reorganizations. Assembling hedgerows
+experienced significantly more changing points than the non-assembling
+networks (difference in the odds ratios between assembling and
+non-assembling networks, $3.316$, $95\%$ CI $[1.314, 8.572]$,
+$p$-value= $0.0117$). Network assembly following restoration is thus
+punctuated by more interaction reorganizations than would be expected
+by environmental shifts that would effect assembling and
+non-assembling networks equally.
 
@@
     perturbation. The robustness to species extinction is measured by
-    incrementally removing species by degree, through removing
-    species by abundance did not yield qualitatively different
-    results. Points are the value for each site at each year of
-    assembly. The solid line indicates the mean slope estimate and the
-    dashed lines are the $95\%$ confidence intervals around the
-    estimate.}
+    incrementally removing species by degree, through removing species
+    by abundance did not yield qualitatively different results. Points
+    are the value for each site at each year of assembly. The solid
+    line indicates the mean slope estimate and the dashed lines are
+    the $95\%$ confidence intervals around the estimate.}
   \label{fig:rob}
@@
 
+We show that the temporal assembly of plant-pollinator networks
+following restoration is a highly dynamic process where patterns of
+interactions undergo several significant reorganizations. These
+network organizations are unlikley to be a project of environmental
+forces because the changing points in non-assembling communities are
+both less frequent and in different time points. The reorganizations
+are thus a product of the assembly process.
+
+There were no consistent trends as to when change points occurred
+within assembling hedgerows or across all sites, except many site had
+changing points between year 2009 and 2011 (Fig.~
+\ref{fig:changePoints}). In California, 2011 marked the beginning of a
+multi year drought. In the assembling hedgerows were not sampled in
+2010, so disentangling whether the changing points are due to skipping
+a year of assembly or the drought is not impossible. Interestingly,
+most of the assembling hedgerows
+
+The species that were most variable in their network position and thus
+contributed most to network reorganizations where species with the
+highest degrees (i.e., most generalized) and persistence.  For
+example, the five most ubiquitous species in our study landscape ---
+\textit{Halictus ligatus}, \textit{Halictus
+  tripartitus},\textit{Lasioglossum (Dialictus) incompletum}, and
+\textit{Toxomerus marginatus} --- where also the only species that
+changed what module (i.e., community) they were present in across
+years in all of the assembling hedgerows. Because species degree and
+persistence were strongly correlated, it is difficult to disentable
+the causal mechanism for why species with their chatacterists are so
+variable in their network position. Generalized species may be best
+able to exploit the limited floral resources the intensivly managed
+agriculture landscape, and thus also most persistent. These adaptable
+species will take advanatge of the most advantage resrouces, which my
+depend the other pollinator species that are present and the state of
+the floral resource. More persistent species also have longer
+phenologies, so they have the opportunity to visit many different
+flowers, resulting in a high degree.  Either way, our results suggest
+that highly persisitent and generalized species often change their
+network positions.
+
+Preferential attachment is the most widely accepted theory explaining
+community assembly in mutualistic networks. Communities assembling
+through preferential attachment are formed by a core of generalist
+species, to which new species arriving to the community interact
+with. Thus, communities assembling through preferential attachment
+exhibit an increase both in nestedness and in the overall level of
+species specialization, since specialist species are being added to
+the community. However, our changing point analysis results show that
+there is a critical reorganization in the interaction patterns as
+communities assembly, which contradicts the preferential attachment
+theory. Even though we encountered an increase in nestedness through
+time, specialization did not increase. Furthermore, this increase in
+nestedness could be explained by the increase in species
+richness. blablabla mechanistic models.
+
+%% not sure if we want to go here, since it is not really part of the
+%% assembly story. More just a "hmm networks are structured in odd
+%% ways sometimes"  
+% Other interesting result is that even though nestedness increased
+% through time, modularity and connectance did not. \cite{Fortuna2010}
+% suggest that at low connectance levels, mutualistic networks would
+% present high nestedness and high modularity, while at high connectance
+% levels networks would present low nestedness and low
+% modularity. However, our results show the contrary: early communities
+% presented low nestedness, smaller richness, and higher levels of
+% modularity and connectance, while more mature communities presented
+% higher nestedness, higher richness, and smaller modularity and
+% connectance. SO blabla... XX Temporal overlap.
+
+% Evidence indicates that there is a positive relationship between
+% abundance and diet breadth, with the more generalist species having
+% higher abundances. Because of the high generalism, species more
+% abundant also present higher degrees. Thus, we expected that more
+% generalist species would be part of the network core. Surprisingly,
+% the five most common species were also the ones that exhibited the
+% most changes in interaction patterns. These species (names!) were
+% always present in the periphery of the network. One hypotheses is that
+% their high generalism allows them to explore the resources more
+% broadly, allowing them to change their position in the network. XX
+% mention the closeness variability and the species persistence.
+
+High nestedness has been related to an increase in the robustness of
+communities to species extinctions because the core of generalist
+species would act as a buffer against species loss (REF). Similarly,
+communities with greater modularity would be more robust in general,
+because modules would act as a buffer against perturbation spreading
+and the pervasive effects of species loss. We expected that as
+communities assemble and species richness increase, communities would
+become increasingly robust to species loss. however, we found that
+robustness to species loss did not change through time. This is
+interestingly from the conservation biology point of view, because
+even early assembling communities seem to be as robust as more mature
+communities, despite the richness insurance hypothesis. However,
+communities are not only subjected to species loss, but also to other
+types of perturbations that can cascade through the network. Gaiarsa
+et al. (submitted) suggest using algebraic connectivity to explore
+how vulnerable ecoogical communities might be to cascading
+effects. Our results indicate that more mature communities are more
+robust to cascading effects than early on assemblages, and that this
+result is related to species richness. This migth be related to XX....
+
 \section*{Acknowledgments}
@@
 
-We would like to thank Leto Peel and Aaron Clauset for their
-invaluable discussions and for help with the change point analysis.
-We thank the growers and land owners that allowed us to work on their
-property.  We also greatly appreciate the identification assistance of
-expert taxonomists Martin Hauser, Robbin Thorp and Jason Gibbs.  This
-work was supported by funding from the Army Research Office
-(W911NF-11-1-0361 to CK), the Natural Resources Conservation Service
-(CIG-69-3A75-12-253, CIG-69-3A75-9-142, CIG-68-9104-6-101 and
+We would like to thank Paulo Guimar{\~a}es Leto Peel and Aaron Clauset for
+their invaluable discussions and for help with the change point
+analysis.  We thank the growers and land owners that allowed us to
+work on their property.  We also greatly appreciate the identification
+assistance of expert taxonomists Martin Hauser, Robbin Thorp and Jason
+Gibbs.  This work was supported by funding from the Army Research
+Office (W911NF-11-1-0361 to CK), the Natural Resources Conservation
+Service (CIG-69-3A75-12-253, CIG-69-3A75-9-142, CIG-68-9104-6-101 and
 WLF-69-7482-6-277 to The Xerces Society), the National Science
@@
 for LCP was provided by an NSF Graduate Research Fellowship and the
-USDA NIFA Graduate Fellowship. FUNDING FOR MARILLIA.
+USDA NIFA Graduate Fellowship. Funding for MPG was provided by S{\~a}o
+Paulo Research Foundation (FAPESP, grant 2013/13319-5).
 

### 2016-08-18 analysis/changePoint/plotting/src/communities.R

@@ -10,7 +10,7 @@ plotNet <- function(){
       cols <- c(rep("darkolivegreen", length(rownames(g))),
                 rep("gold", length(colnames(g))))
       V(gs)$color <- cols
-      importance <-  (c(rowSums(g) +0.1, colSums(g) + 0.1)/sum(g))*50
+      importance <-  (c(rowSums(g) +0.1, colSums(g) + 0.1)/sum(g))*100
       v.labs <- names(importance)
       v.labs[importance < 5] = ""
       V(gs)$size <- importance

### 2016-08-19 ms/ms.tex
worked on discussion

@@
 could result from a process like preferential attachment where
-specialist species attach to the well-connected, generalist core.
+specialist species attach to the well-connected, generalist core.  In
+addition, non-successional temporal dynamics also suggest a stable
+core of generalists persists despite high turnover of peripheral
+species \citep{fang2012relative, diaz2010changes, alarcon2008year}.
 
@@
 % \citep{eilers-2011-e21363}.
-
-% Even non-successional temporal dynamics suggest a stable core of
-% generalists persist despite high turnover of peripheral species
-% \citep{fang2012relative, diaz2010changes, alarcon2008year}.
 
@@
 
-There were no consistent trends as to when change points occurred
-within assembling hedgerows or across all sites, except many site had
+There were no consistent trends in when change points occurred within
+assembling hedgerows or across all sites, except many sites had
 changing points between year 2009 and 2011 (Fig.~
 \ref{fig:changePoints}). In California, 2011 marked the beginning of a
-multi year drought. In the assembling hedgerows were not sampled in
+multi-year drought. In the assembling hedgerows were not sampled in
 2010, so disentangling whether the changing points are due to skipping
 a year of assembly or the drought is not impossible. Interestingly,
-most of the assembling hedgerows
+most of the assembling hedgerows did not undergo a signficiant
+reorganization of interactions immediatly the hedgerow was planted
+(i.e., the transition from weedy field margin to hedgerow). This
+result is consistent with other studies finding that the gains in
+species richness of both the plant and pollinators takes several years
+post-restoration (Kremen and M'Gonigle, in prep).
 
-The species that were most variable in their network position and thus
-contributed most to network reorganizations where species with the
-highest degrees (i.e., most generalized) and persistence.  For
-example, the five most ubiquitous species in our study landscape ---
-\textit{Halictus ligatus}, \textit{Halictus
+Surprisingly, the species that were most variable in their network
+position and thus contributed most to network reorganizations where
+species with the highest degrees (i.e., most generalized) and
+persistence.  For example, the five most ubiquitous species in our
+study landscape --- \textit{Halictus ligatus}, \textit{Halictus
   tripartitus},\textit{Lasioglossum (Dialictus) incompletum}, and
@@
 agriculture landscape, and thus also most persistent. These adaptable
-species will take advanatge of the most advantage resrouces, which my
+species will take advanatge of the most advantage resources, which my
 depend the other pollinator species that are present and the state of
@@
 flowers, resulting in a high degree. Either way, our results suggest
-that highly persisitent and generalized species often change their
-network positions.
+that given the opportunity and ability to use different resources,
+species will often their network positions.
+
+%% node/species turnover
+
+The frequent changing points in network organization, dynamic nature
+of the location of species in networks, and turnover of species and
+nodes all point to an assembly mechanism other than preferential
+attachment. Nestedness did increase with years post restoration, as
+would be expected if colonizing, specialist species attached to a
+central, generalist core. With preferential attachment, however, we
+could also expect connectance and specialization to increase, and see
+neither. The increase in nestedness is thus more likely a consequence
+of the enhancement of species richness through time \flagged{DO WE
+  HAVE CITATIONS FOR THIS?}. In addition, though the colonization and
+persistence of floral specialists increases with assembly
+\citep{mgonigle-2015-x}, this must be accumpanied changes in the diet
+breat of the present species in order for network-level specialization
+to remain constant through time.
 
 Preferential attachment is the most widely accepted theory explaining
-community assembly in mutualistic networks. Communities assembling
-through preferential attachment are formed by a core of generalist
-species, to which new species arriving to the community interact
-with. Thus, communities assembling through preferential attachment
-exhibit an increase both in nestedness and in the overall level of
-species specialization, since specialist species are being added to
-the community. However, our changing point analysis results show that
-there is a critical reorganization in the interaction patterns as
-communities assembly, which contradicts the preferential attachment
-theory. Even though we encountered an increase in nestedness through
-time, specialization did not increase. Furthermore, this increase in
-nestedness could be explained by the increase in species
-richness. blablabla mechanistic models.
+network assembly, but few ecological studies have been conducted that
+test for this mechanism \citep{Olesen2008}. \flagged{other
+  citations??}. Though \cite{Olesen2008} found some evidence for an
+assmbly process like prefrential attachment governming the temporal
+build up of plant-pollinator networks within a season, the frequent
+reorganizations of the hedgerow communities suggest a very different
+mechanism is at play. Plant-pollinator networks in general are highly
+dynamic, with high turnover of species and interactions both within
+and between seasons \citep{Burkle2011}. Though our non-assembling
+communities experience fewer network reorganizations than the
+assembling hedgerows, $82\%$ of field marings and $40\%$ of mature
+hedgerows underwent at least one changing point in network
+structure. Few alternative models to perferential attachment that
+predict these network reorganizations have been explored, however.
+
+Complementarity between traits like tounge length and corolla depth as
+well as body and flower impose some biophysical limits to the
+interactions between plants and pollintors
+\citep{Vazquez2009evaluating, Stang2009, Stang2006,
+  Santamaria2007}. Similarly, the phenologies of plants and
+pollinators must also overlap \citep{Vazquez2009,
+  Vazquez2009evaluating}.
+
+ \citep{valdovinos2013adaptive}
+
+The changes in network patterns with assembly also increased the
+robustness of the networks to perturbation, but not to plant species
+loss.
+
+The increase in nestedness has been related to an increase in the
+robustness of communities to species extinctions because the core of
+generalist species would act as a buffer against species loss
+(REF). Similarly, communities with greater modularity would be more
+robust in general, because modules would act as a buffer against
+perturbation spreading and the pervasive effects of species loss. We
+expected that as communities assemble and species richness increase,
+communities would become increasingly robust to species loss. however,
+we found that robustness to species loss did not change through
+time. This is interestingly from the conservation biology point of
+view, because even early assembling communities seem to be as robust
+as more mature communities, despite the richness insurance
+hypothesis. However, communities are not only subjected to species
+loss, but also to other types of perturbations that can cascade
+through the network. Gaiarsa et al. (submitted) suggest using
+algebraic connectivity to explore how vulnerable ecoogical communities
+might be to cascading effects. Our results indicate that more mature
+communities are more robust to cascading effects than early on
+assemblages, and that this result is related to species richness. This
+migth be related to XX....
 
@@
 %% assembly story. More just a "hmm networks are structured in odd
-%% ways sometimes"  
+%% ways sometimes" and we dont' really have answer for why this is
+%% happening...
+
 % Other interesting result is that even though nestedness increased
@@
 % connectance. SO blabla... XX Temporal overlap.
-
-% Evidence indicates that there is a positive relationship between
-% abundance and diet breadth, with the more generalist species having
-% higher abundances. Because of the high generalism, species more
-% abundant also present higher degrees. Thus, we expected that more
-% generalist species would be part of the network core. Surprisingly,
-% the five most common species were also the ones that exhibited the
-% most changes in interaction patterns. These species (names!) were
-% always present in the periphery of the network. One hypotheses is that
-% their high generalism allows them to explore the resources more
-% broadly, allowing them to change their position in the network. XX
-% mention the closeness variability and the species persistence.
-
-High nestedness has been related to an increase in the robustness of
-communities to species extinctions because the core of generalist
-species would act as a buffer against species loss (REF). Similarly,
-communities with greater modularity would be more robust in general,
-because modules would act as a buffer against perturbation spreading
-and the pervasive effects of species loss. We expected that as
-communities assemble and species richness increase, communities would
-become increasingly robust to species loss. however, we found that
-robustness to species loss did not change through time. This is
-interestingly from the conservation biology point of view, because
-even early assembling communities seem to be as robust as more mature
-communities, despite the richness insurance hypothesis. However,
-communities are not only subjected to species loss, but also to other
-types of perturbations that can cascade through the network. Gaiarsa
-et al. (submitted) suggest using algebraic connectivity to explore
-how vulnerable ecoogical communities might be to cascading
-effects. Our results indicate that more mature communities are more
-robust to cascading effects than early on assemblages, and that this
-result is related to species richness. This migth be related to XX....
 
@@
 \bibliography{refs}
+
+% Communities assembling through preferential attachment are formed by a
+% core of generalist species, to which new species arriving to the
+% community interact with. Thus, communities assembling through
+% preferential attachment exhibit an increase both in nestedness and in
+% the overall level of species specialization, since specialist species
+% are being added to the community. However, our changing point analysis
+% results show that there is a critical reorganization in the
+% interaction patterns as communities assembly, which contradicts the
+% preferential attachment theory. Even though we encountered an increase
+% in nestedness through time, specialization did not
+% increase. Furthermore, this increase in nestedness could be explained
+% by the increase in species richness. blablabla mechanistic models.
+
+% Evidence indicates that there is a positive relationship between
+% abundance and diet breadth, with the more generalist species having
+% higher abundances. Because of the high generalism, species more
+% abundant also present higher degrees. Thus, we expected that more
+% generalist species would be part of the network core. Surprisingly,
+% the five most common species were also the ones that exhibited the
+% most changes in interaction patterns. These species (names!) were
+% always present in the periphery of the network. One hypotheses is that
+% their high generalism allows them to explore the resources more
+% broadly, allowing them to change their position in the network. XX
+% mention the closeness variability and the species persistence.
 

### 2016-08-20 ms/ms.tex
worked on disucssion

@@
 attachment, assembly can be punctuated by significant reorganizations
-of interactions \citep{peel2014detecting}. For example, as new species
-are added, resident species change their interaction partners to
-minimize competition, or become extinct. Such significant
+of interactions \citep{peel2014detecting}. Such significant
 reorganizations of interactions, or changing points, have been
-observed in networks responding to abrupt shifts in the behavior of
-interacters \citep{peel2014detecting}. No studies, however, have
-examined whether changing points occur during ecological network
-assembly.
+observed in social networks responding to abrupt shifts in the
+behavior of interactors \citep{peel2014detecting}.  In ecological
+communities, such shifts my occur if, as new species are added,
+resident species change their interaction partners to optimize their
+foraging effort. In plant-pollinator communities, theory predicts
+pollinators optimize their use of floral resources to reduce
+interspecific competition and improve resource-use efficiency
+\citep{pyke1984optimal, valdovinos2010consequences,
+  valdovinos2013adaptive, albrecht2010plant, Bluthgen2007}. No
+studies, however, have examined whether changing points occur during
+ecological network assembly and how they relate to the behavior of the
+interactors.
 
@@
 %% in ag. but ag is a bad place for pollinators
-Facilitating effective restoration of networks is particularly
+Facilitating effective restoration of networks is especially
 imperative in areas where species interactions provide essential
@@
 
-% Pollinators are one such key group of interactors: $75\%$ of all
-% crop species depend to some extent on pollinator visitation
-% \citep{klein-2007-303}, and animal-pollinated crops supply a large
-% proportion of essential nutrients to the human diet
-% \citep{eilers-2011-e21363}.
-
 \section*{Materials \& Methods}
@@
 the most abundant and effective pollinators in the system (C. Kremen,
-A.~ Klein and L.~ Morandin, unpublished data). Bee and syrphid specimens
+A.~Klein and L.~Morandin, unpublished data). Bee and syrphid specimens
 were identified to species (or morpho-species for some bee specimens
@@
 is calculated from interaction observations observed in more extensive
-data-set from Yolo County (approx.~ 18000 interaction records) that
+data-set from Yolo County (approx.~18000 interaction records) that
 included both the data included in this study and additional data from
@@
 \subsection*{Network structure}
-The changing points in network structure may contribute to the
+Any changing points in network structure may contribute to the
 reorganization of the assembling networks into predictable interaction
@@
 networks to 1 for specialized networks.
-\subsection*{Network structure}
-
-The changing points in network structure may contribute to the
-reorganization of the assembling networks into predictable interaction
-patterns. Pollination networks exhibit two main structural patterns
---- modularity \citep[e.g.,][]{Olesen2007} and nestedness
-\citep[e.g.,][]{Bascompte2006, Bascompte2003}. In modular community
-interactions are insular, occurring within separate groups or
-``modules'' more often than between modules. Communities in the
-network may fragment as the network assembles, enhancing
-modularity. Conversely, nested networks are like pyramid of
-interactions, where there are some species that interact with many
-species, other species that interact with a subset of those species,
-and so on. If species entering the network tend to interact with the
-generalist base of the network pyramid, nestedness would increase
-through time. Alternatively, if the network is accumulating specialist
-species or if species are beginning to limit their interaction niche
-breath as the network assembles, this would lead to an increase in the
-network-level specialization \citep{bluthgen-2006-9} and nestedness
-would decrease through time. To test whether network modularity,
-nestedness or specialization changed linearly with assembly, we used
-linear mixed models with the descriptive network metrics as the
-response variable, year of assembly as the explanatory variable, and
-random effects of site and year.
 
@@
 The majority ($76\%$) of the sites tests underwent at least one
-significant reorganization of interactions (Fig.~
-\ref{fig:changePoints}).  All five of the assembling hedgerows
+significant reorganization of interactions
+(Fig.~\ref{fig:changePoints}).  All five of the assembling hedgerows
 experienced changing points, whereas only $40\%$ and $81\%$ of
@@
     has a consistent position in the network across years. In the
-    assembling hedgerows, colored circles in the corner of each
+    assembling hedgerows, colored squares in the corner of each
     network represent the years post restoration.}
@@
 estimate, $1.834 \pm 0.6142$, $p$-value=$0.022$, Fig.~
-\ref{fig:baci}).  Modularity decreased (Fig.~ \ref{fig:baci}), though
+\ref{fig:baci}).  Modularity decreased (Fig.~\ref{fig:baci}), though
 the slope was not significantly different from zero (estimate of the
@@
 of the estimate, $6*10^{-5} \pm 4*10^{-3}$, $p$-value=$0.987$) or
-abundance ($0.001 \pm 0.003$, $p$-value=$0.65$, Fig.~ \ref{fig:rob}).
+abundance ($0.001 \pm 0.003$, $p$-value=$0.65$, Fig.~\ref{fig:rob}).
 
@@
 standard error of the estimate, $0.6814 \pm 0.272$, $p$-value=$0.042$,
-Fig.~ \ref{fig:rob}).
+Fig.~\ref{fig:rob}).
 
@@
 We show that the temporal assembly of plant-pollinator networks
-following restoration is a highly dynamic process where patterns of
-interactions undergo several significant reorganizations. These
-network organizations are unlikley to be a project of environmental
-forces because the changing points in non-assembling communities are
-both less frequent and in different time points. The reorganizations
-are thus a product of the assembly process.
+following restoration is a highly dynamic process where interactions
+often undergo significant reorganizations. These network organizations
+are unlikely to be a product of environmental forces alone because the
+network changing points in non-assembling communities are less
+frequent, and there are few consistent trends in when change points
+occurred across all sites. Several sites had changing points between
+year 2009 and 2011 (Fig.~\ref{fig:changePoints}). In California, 2011
+marked the beginning of a multi-year drought. In the assembling
+hedgerows were not sampled in 2010, so disentangling whether the
+changing points are due to skipping a year of assembly or the drought
+is not impossible. Interestingly, most of the assembling hedgerows did
+not undergo a significant reorganization of interactions immediately
+the hedgerow was planted (i.e., the transition from weedy field margin
+to hedgerow). This result is consistent with the finding that in our
+study system, hedgerow restoration takes several years to have a
+significant impact on the plant-pollintor communities (Kremen and
+M'Gonigle, in prep).
 
-There were no consistent trends in when change points occurred within
-assembling hedgerows or across all sites, except many sites had
-changing points between year 2009 and 2011 (Fig.~
-\ref{fig:changePoints}). In California, 2011 marked the beginning of a
-multi-year drought. In the assembling hedgerows were not sampled in
-2010, so disentangling whether the changing points are due to skipping
-a year of assembly or the drought is not impossible. Interestingly,
-most of the assembling hedgerows did not undergo a signficiant
-reorganization of interactions immediatly the hedgerow was planted
-(i.e., the transition from weedy field margin to hedgerow). This
-result is consistent with other studies finding that the gains in
-species richness of both the plant and pollinators takes several years
-post-restoration (Kremen and M'Gonigle, in prep).
-
-Surprisingly, the species that were most variable in their network
-position and thus contributed most to network reorganizations where
-species with the highest degrees (i.e., most generalized) and
-persistence.  For example, the five most ubiquitous species in our
-study landscape --- \textit{Halictus ligatus}, \textit{Halictus
-  tripartitus},\textit{Lasioglossum (Dialictus) incompletum}, and
-\textit{Toxomerus marginatus} --- where also the only species that
-changed what module (i.e., community) they were present in across
-years in all of the assembling hedgerows. Because species degree and
-persistence were strongly correlated, it is difficult to disentable
-the causal mechanism for why species with their chatacterists are so
-variable in their network position. Generalized species may be best
-able to exploit the limited floral resources the intensivly managed
-agriculture landscape, and thus also most persistent. These adaptable
-species will take advanatge of the most advantage resources, which my
-depend the other pollinator species that are present and the state of
-the floral resource. More persistent species also have longer
+TRANSITION... Surprisingly, the species that were most variable in
+their network position and thus contributed most to network
+reorganizations where species with the highest degrees (i.e., most
+generalized) and persistence.  For example, the five most ubiquitous
+species in our study landscape --- \textit{Halictus ligatus},
+\textit{Halictus tripartitus},\textit{Lasioglossum (Dialictus)
+  incompletum}, and \textit{Toxomerus marginatus} --- were also the
+only species that changed what module (i.e., community) they were
+present in across years in all of the assembling hedgerows. Because
+species degree and persistence were strongly correlated, it is
+difficult to disentangle the causal mechanism for why species with
+those characteristics are so variable in their network
+position. Generalized species may be best able to exploit the limited
+floral resources the intensively managed agriculture landscape, and
+thus also most persistent. More persistent species also have longer
 phenologies, so they have the opportunity to visit many different
-flowers, resulting in a high degree. Either way, our results suggest
-that given the opportunity and ability to use different resources,
-species will often their network positions.
+flowers, resulting in a higher degree. Either way, our result suggests
+that adaptable species can change their network position to utilize
+the most advantageous floral resources, which my depend the other
+pollinator species that are present and the state of the floral
+resource. Thus given the opportunity and ability to use different
+resources, species will often change their network positions.
 
@@
 would be expected if colonizing, specialist species attached to a
-central, generalist core. With preferential attachment, however, we
-could also expect connectance and specialization to increase, and see
-neither. The increase in nestedness is thus more likely a consequence
-of the enhancement of species richness through time \flagged{DO WE
-  HAVE CITATIONS FOR THIS?}. In addition, though the colonization and
-persistence of floral specialists increases with assembly
-\citep{mgonigle-2015-x}, this must be accumpanied changes in the diet
-breat of the present species in order for network-level specialization
-to remain constant through time.
-
-Preferential attachment is the most widely accepted theory explaining
-network assembly, but few ecological studies have been conducted that
-test for this mechanism \citep{Olesen2008}. \flagged{other
-  citations??}. Though \cite{Olesen2008} found some evidence for an
-assmbly process like prefrential attachment governming the temporal
-build up of plant-pollinator networks within a season, the frequent
-reorganizations of the hedgerow communities suggest a very different
-mechanism is at play. Plant-pollinator networks in general are highly
-dynamic, with high turnover of species and interactions both within
-and between seasons \citep{Burkle2011}. Though our non-assembling
-communities experience fewer network reorganizations than the
-assembling hedgerows, $82\%$ of field marings and $40\%$ of mature
-hedgerows underwent at least one changing point in network
-structure. Few alternative models to perferential attachment that
-predict these network reorganizations have been explored, however.
-
-Complementarity between traits like tounge length and corolla depth as
-well as body and flower impose some biophysical limits to the
-interactions between plants and pollintors
-\citep{Vazquez2009evaluating, Stang2009, Stang2006,
-  Santamaria2007}. Similarly, the phenologies of plants and
-pollinators must also overlap \citep{Vazquez2009,
-  Vazquez2009evaluating}.
-
- \citep{valdovinos2013adaptive}
+central, generalist core \cite{albrecht2010plant}. With preferential
+attachment, however, we would also expect connectance and
+specialization to increase, and see no such trends. The stable level
+of network-level specialization through the assembly process may be
+due to the increased colonization of specialized species
+\citep{mgonigle-2015-x} accompanied by an increase in the diet breath
+of resident species. This would be expected if resident species were
+able to minimize their foraging time by expanding their diet breath as
+plant diversity increases with hedgerow maturation \citep{Waser1996,
+  pyke1984optimal, Bluthgen2007, albrecht2010plant}. Such a change in
+pollinator behavior would also explain the increase in network
+nestedness.
 
 The changes in network patterns with assembly also increased the
-robustness of the networks to perturbation, but not to plant species
-loss.
+robustness of the networks to perturbation -- providing further
+evidence that hedgerows are valuable tool for promoting species
+conservation and ecosystem provision in agricultural areas
+\citep{mgonigle-2015-x, ponisio2015farm, kremen-2015-602}. Because the
+time it takes for a perturbation to spread through a network, as
+measured by algebraic connectivity, is correlated with species
+richness, the increase and plant and pollinator richness following
+restoration is at least partially responsible for enhancing network
+robustness. SOMETHING ABOUT DIVERSITY AND STABILITY
 
-The increase in nestedness has been related to an increase in the
-robustness of communities to species extinctions because the core of
-generalist species would act as a buffer against species loss
-(REF). Similarly, communities with greater modularity would be more
-robust in general, because modules would act as a buffer against
-perturbation spreading and the pervasive effects of species loss. We
-expected that as communities assemble and species richness increase,
-communities would become increasingly robust to species loss. however,
-we found that robustness to species loss did not change through
-time. This is interestingly from the conservation biology point of
-view, because even early assembling communities seem to be as robust
-as more mature communities, despite the richness insurance
-hypothesis. However, communities are not only subjected to species
-loss, but also to other types of perturbations that can cascade
-through the network. Gaiarsa et al. (submitted) suggest using
-algebraic connectivity to explore how vulnerable ecoogical communities
-might be to cascading effects. Our results indicate that more mature
-communities are more robust to cascading effects than early on
-assemblages, and that this result is related to species richness. This
-migth be related to XX....
+Interestingly, however, the robustness of hedgerow communities to
+species loss did not change with assembly. This is particularly
+surprising given the observed increase in nestedness, which is often
+associated with an enhanced in robustness to extinction
+\citep{Memmott2004}. It may be that assembling hedgerows have not yet
+reached a sufficient level of nestedness to realize its
+benefits. Nestedness of the assembling hedgerows, however, did not
+asymptote within the eight years following restoration that the sites
+were surveyed, so hedgerow networks may eventually reach sufficient
+levels of nestedness to gain the robustness advantage.
+
+Plant-pollinator networks in general are highly dynamic, with high
+turnover of species and interactions both within and between seasons
+\citep{Burkle2011}. Though our non-assembling communities experience
+fewer network reorganizations than the assembling hedgerows, $82\%$ of
+field margins and $40\%$ of mature hedgerows underwent at least one
+changing point in network structure. Pollinators are also highly
+opportunistic \citep{petanidou-2008-564, Vazquez2005b,
+  albrecht2010plant}, though trait complementarity such as tongue
+length and corolla depth impose some biophysical limits to the
+interactions between plants and pollinators
+\citep{Vazquez2009evaluating, Vazquez2009, Stang2009, Stang2006,
+  Santamaria2007}. Such opportunism may buffer plant-pollinator
+communities from global change \citep[e.g.,][]{ramos2012topological,
+  kaiser2010robustness}, but our limited understanding of the assembly
+of these communities impedes making such predictions
+\citep{Vazquez2009, Burkle2011}. Unlike in the broader food web
+literature, we have few mechanistic models of mutualistic network
+assembly \citep{valdovinos2013adaptive}. In addition, the few that
+have been developed often borrow their mechanisms from competitive
+interactions, leading to inaccurate biological assumptions
+\citep{holland2006comment}. We need further development of mechanistic
+models of mutualistic stystem to generate testable predictions, along
+with empirical exploration of network assembly. Plant-pollinator
+communities and mutualisms in general are vital for the maintenance of
+biodiversity and the provision of essential ecosystem services. We
+must therefore understand the processes underlying their assembly to
+facilitate restoration and conservation.
+
+% Though \cite{Olesen2008} found some evidence for an
+% assmbly process like prefrential attachment governing the temporal
+% build up of plant-pollinator networks within a season, the frequent
+% reorganizations of the hedgerow communities suggest a very different
+% mechanism is at play.
 
@@
 
-We would like to thank Paulo Guimar{\~a}es Leto Peel and Aaron Clauset for
-their invaluable discussions and for help with the change point
+We would like to thank Paulo Guimar{\~a}es and Aaron Clauset for their
+invaluable discussions, and Leto Peel for help with the change point
 analysis.  We thank the growers and land owners that allowed us to
@@
 % mention the closeness variability and the species persistence.
+
+% The increase in nestedness has been related to an increase in the
+% robustness of communities to species extinctions because the core of
+% generalist species would act as a buffer against species loss
+% (REF). Similarly, communities with greater modularity would be more
+% robust in general, because modules would act as a buffer against
+% perturbation spreading and the pervasive effects of species loss. We
+% expected that as communities assemble and species richness increase,
+% communities would become increasingly robust to species loss. however,
+% we found that robustness to species loss did not change through
+% time. This is interestingly from the conservation biology point of
+% view, because even early assembling communities seem to be as robust
+% as more mature communities, despite the richness insurance
+% hypothesis. However, communities are not only subjected to species
+% loss, but also to other types of perturbations that can cascade
+% through the network. Gaiarsa et al. (submitted) suggest using
+% algebraic connectivity to explore how vulnerable ecoogical communities
+% might be to cascading effects. Our results indicate that more mature
+% communities are more robust to cascading effects than early on
+% assemblages, and that this result is related to species richness. This
+% migth be related to XX....
 

### 2016-08-22 ms/ms.tex
changes the perturbation figures, modified the manuscript

@@
 used to describe spectral properties of complex graphs other than
-ecological networks, and it is related to how difficult it is to turn
+ecological networks, and it has been extensively studied \citep[e.g.,][]{fiedler1973algebraic, costa2007characterization, mohar1991laplacian, de2007old}.
+Algebraic connectivity is related to the mean diameter of a graph and to how difficult it is to turn
 a network into completely disconnected groups of nodes, or species
-\citep{fiedler1973algebraic, costa2007characterization}, and also to
-flows in . The larger the algebraic connectivity, the more robust a
-network is to perturbations (e.g. Gaiarsa et al., submitted).
+\citep{costa2007characterization}--- the larger the algebraic connectivity, the more robust a
+network is to cascading effects (e.g. Gaiarsa et al., submitted), and the harder it is to break the community into isolated groups of species.
 
@@
   \caption{The robustness of networks to species extinction and
-    perturbation. The robustness to species extinction is measured by
+    to cascading effects. The robustness to species extinction is measured by
     incrementally removing species by degree, through removing species
-    by abundance did not yield qualitatively different results. Points
+    by abundance did not yield qualitatively different results.
+The robustness of networks to cascading effects is measured as the algebraic connectivity, the second smallest eigenvalue of the Laplacian matrix.
+Points
     are the value for each site at each year of assembly. The solid
@@
 changing points are due to skipping a year of assembly or the drought
-is not impossible. Interestingly, most of the assembling hedgerows did
+is not possible, even though two changing points did occur in the non assembling hedgerow six between 2009 and 2010, and 2010 and 2011 (HR6, Fig.~\ref{fig:changePoints}) .
+Interestingly, most of the assembling hedgerows did
 not undergo a significant reorganization of interactions immediately
@@
 
-TRANSITION... Surprisingly, the species that were most variable in
+Given that changing ponts did occur, TRANSITION... Surprisingly, the species that were most variable in
 their network position and thus contributed most to network

### 2016-08-22 analysis/networkLevel/plotting/src/plotPanels_resilence.R

@@ -22,7 +22,7 @@ plot.panels <- function(){
                col.fill=col.fill,
                col.lines=col.lines,
                plotPoints=TRUE)
-    mtext("Robustness to \n perturbation", 2, line=4, cex=1.5)
+    mtext("Robustness to \n cascading effects", 2, line=4, cex=1.5)
 
     axis(1, pretty(res$ypr), labels=pretty(res$ypr))
     mtext("Year of assembly", 1, line=3.5, cex=1.5)

### 2016-08-22 ms/ms.tex
added bits to the discussion

@@
 changing points are due to skipping a year of assembly or the drought
-is not possible, even though two changing points did occur in the non assembling hedgerow six between 2009 and 2010, and 2010 and 2011 (HR6, Fig.~\ref{fig:changePoints}) .
+is not possible, even though two changing points did occur in the non assembling hedgerow six between 2009 and 2010, and 2010 and 2011 (HR6, Fig.~\ref{fig:changePoints}).
 Interestingly, most of the assembling hedgerows did
@@
 
-Given that changing ponts did occur, TRANSITION... Surprisingly, the species that were most variable in
+Given that changing ponts did occur, we then turned to explore how these changing points came to be, and what species were responsible for it.
+Ecolgical theory predicts that species abundance is positively correlated to number of interaction partners a species has (degree \citep{Bluthgen2007}).
+Thus, we expected that the more abundant species would be more persisten through network assembly, and would contribute the least to the changing points.
+Surprisingly however, we encountered the opposite: the species that were most variable in
 their network position and thus contributed most to network
@@
 \textit{Halictus tripartitus},\textit{Lasioglossum (Dialictus)
-  incompletum}, and \textit{Toxomerus marginatus} --- were also the
-only species that changed what module (i.e., community) they were
+  incompletum}, and \textit{Toxomerus marginatus} --- were the
+only species that changed what module (i.e., community), and they were also 
 present in across years in all of the assembling hedgerows. Because
@@
 those characteristics are so variable in their network
-position. Generalized species may be best able to exploit the limited
+position. Generalized species may be able to better exploit the limited
 floral resources the intensively managed agriculture landscape, and
@@
 that adaptable species can change their network position to utilize
-the most advantageous floral resources, which my depend the other
+the most advantageous floral resources available, which my depend the other
 pollinator species that are present and the state of the floral
@@
 
-%% node/species turnover
+%% node/species turnover  - I am not sure I get these results so there might be some nonsense in the below paragraph...
+Given the regional species pool it is possible that specialization migth be constant either because species are not changing through time, or because there is a species turnover, but new species have the same role/posotion as the species there are absent in the following time step.
+We found that node turover was higher in assembling and non-assembling hedgerows when compared to field margin, but there was no difference between them.
+Thus, maybe looking just at network level metrics, such as nestedness and modularity, when exploring the temporal assembly of ecological networks might be misleading, given that there is a node turnover happening at the species level.
 
@@
 would be expected if colonizing, specialist species attached to a
-central, generalist core \cite{albrecht2010plant}. With preferential
+central, pre existing generalist core \cite{albrecht2010plant}. With preferential
 attachment, however, we would also expect connectance and
-specialization to increase, and see no such trends. The stable level
+specialization to increase, and we found no such trends. The stable level
 of network-level specialization through the assembly process may be
@@
 \citep{mgonigle-2015-x, ponisio2015farm, kremen-2015-602}. Because the
-time it takes for a perturbation to spread through a network, as
+vulnerability of the network to cascading effects, as
 measured by algebraic connectivity, is correlated with species
@@
 restoration is at least partially responsible for enhancing network
-robustness. SOMETHING ABOUT DIVERSITY AND STABILITY
+robustness to cascading effects.
+This result contributes to the stability and diversity debate indicating that diversity might promote statibility, even though there are contrasting results in the literature \citep [e.g.][]{may1972will, pimm1984complexity, montoya2006ecological}.
+However, the robustness of ecological communities to disturbances  might depend on the identity of the species mostly being affected by the disturbance (REF).
+For example, networks might be more sensible to the disturbance of a generalist when compared to a specialist (REF).
 

### 2016-08-22 ms/ms.tex
modified discussion

@@
 ramifications for the robustness of the networks to species extinction
-and perturbation.
+and cascading perturbations.
 
@@
 We also explored how the robustness to cascading effects changed as
-community assembled, using algebraic connectivity--- the second
+community assembled, using algebraic connectivity --- the second
 smallest eigenvalue of the Laplacian matrix
-\citep{fiedler1973algebraic}--- as a proxy for network robustness
-(e.g. Gaiarsa et al., submitted). Algebraic connectivity was first
-used to describe spectral properties of complex graphs other than
-ecological networks, and it has been extensively studied \citep[e.g.,][]{fiedler1973algebraic, costa2007characterization, mohar1991laplacian, de2007old}.
-Algebraic connectivity is related to the mean diameter of a graph and to how difficult it is to turn
-a network into completely disconnected groups of nodes, or species
-\citep{costa2007characterization}--- the larger the algebraic connectivity, the more robust a
-network is to cascading effects (e.g. Gaiarsa et al., submitted), and the harder it is to break the community into isolated groups of species.
+\citep{fiedler1973algebraic} --- as a proxy for network robustness
+(Gaiarsa et al., submitted). Algebraic connectivity is related to the
+mean diameter of a graph and to how difficult it is to turn a network
+into completely disconnected groups of nodes, or species
+\citep{costa2007characterization} --- the larger the algebraic
+connectivity, the more robust a network is to cascading effects
+(Gaiarsa et al., submitted), and the harder it is to break the
+community into isolated groups of species.
+
+%% marilia:
+%% I'm not totally sure what this adds to the methods. Since most
+%% people reading this paper will not understand what "spectral
+%% properties" are (I don't in fact =/ ) it is a bit of a misdirection
+%% since we are not interested in it. I think it is enough to point to
+%% your paper which is using the Alg Con the way we want to.
+
+%% removed:
+% Algebraic connectivity is
+% used to describe spectral properties of complex graphs other than
+% ecological networks \citep[e.g.,][]{fiedler1973algebraic,
+%   costa2007characterization, mohar1991laplacian, de2007old}.
+
+% what's the mean diameter of a graph? Also can we use the
+%% term network instead of graph to be mote consistent with the rest
+%% of the methods? We use graph a bit in the change point section, but
+%% I think it is explained a bit better there?
+
+%% I also removed the e.g.'s because your paper is all we want to
+%% cite, and not "for example Gariarsa et al."
 
@@
 
-In contrast, the robustness of networks to perturbation, as measured
-by the algebraic connectivity of the network, increased as the network
-assembled (estimate of the slope of robustness through time $\pm$
-standard error of the estimate, $0.6814 \pm 0.272$, $p$-value=$0.042$,
-Fig.~\ref{fig:rob}).
+In contrast, the robustness of networks to cascading perturbations, as
+measured by the algebraic connectivity of the network, increased as
+the network assembled (estimate of the slope of robustness through
+time $\pm$ standard error of the estimate, $0.6814 \pm 0.272$,
+$p$-value=$0.042$, Fig.~\ref{fig:rob}).
 
@@
 changing points are due to skipping a year of assembly or the drought
-is not possible, even though two changing points did occur in the non assembling hedgerow six between 2009 and 2010, and 2010 and 2011 (HR6, Fig.~\ref{fig:changePoints}).
-Interestingly, most of the assembling hedgerows did
+is not possible. Interestingly, most of the assembling hedgerows did
 not undergo a significant reorganization of interactions immediately
@@
 
-Given that changing ponts did occur, we then turned to explore how these changing points came to be, and what species were responsible for it.
-Ecolgical theory predicts that species abundance is positively correlated to number of interaction partners a species has (degree \citep{Bluthgen2007}).
-Thus, we expected that the more abundant species would be more persisten through network assembly, and would contribute the least to the changing points.
-Surprisingly however, we encountered the opposite: the species that were most variable in
-their network position and thus contributed most to network
-reorganizations where species with the highest degrees (i.e., most
-generalized) and persistence. For example, the five most ubiquitous
-species in our study landscape --- \textit{Halictus ligatus},
-\textit{Halictus tripartitus},\textit{Lasioglossum (Dialictus)
-  incompletum}, and \textit{Toxomerus marginatus} --- were the
-only species that changed what module (i.e., community), and they were also 
-present in across years in all of the assembling hedgerows. Because
-species degree and persistence were strongly correlated, it is
-difficult to disentangle the causal mechanism for why species with
-those characteristics are so variable in their network
-position. Generalized species may be able to better exploit the limited
-floral resources the intensively managed agriculture landscape, and
-thus also most persistent. More persistent species also have longer
-phenologies, so they have the opportunity to visit many different
-flowers, resulting in a higher degree. Either way, our result suggests
-that adaptable species can change their network position to utilize
-the most advantageous floral resources available, which my depend the other
-pollinator species that are present and the state of the floral
-resource. Thus given the opportunity and ability to use different
-resources, species will often change their network positions.
+Given that several changing points in network organization occured
+during assembling, we next explored the species most likely
+responsible for the observed shifts in interaciton patterns. Based on
+a preferential attachment-like mechanism, we expect that the most
+persistent and high degree species would remain stable in the network
+core during assembly, and would thus contribute the least to the
+changing points in network organization. Surprisingly however, we
+encountered the opposite: the species that were most variable in their
+network position and thus contributed most to network reorganizations
+where species with the highest degrees (i.e., most generalized) and
+persistence. For example, the five most ubiquitous species in our
+study landscape --- \textit{Halictus ligatus}, \textit{Halictus
+  tripartitus}, \textit{Lasioglossum (Dialictus) incompletum}, and
+\textit{Toxomerus marginatus} --- were the only species that changed
+what module (i.e., community), and they were also present in across
+years in all of the assembling hedgerows. Because species degree and
+persistence were strongly correlated, it is difficult to disentangle
+the causal mechanism for why species with those characteristics are so
+variable in their network position. Generalized species may be able to
+better exploit the limited floral resources the intensively managed
+agriculture landscape, and thus also most persistent. More persistent
+species also have longer phenologies, so they have the opportunity to
+visit many different flowers, resulting in a higher degree. Either
+way, our result suggests that adaptable species can change their
+network position to utilize the most advantageous floral resources
+available, which my depend the other pollinator species that are
+present and the state of the floral resource. Thus given the
+opportunity and ability to use different resources, species will often
+change their network positions.
 
@@
 would be expected if colonizing, specialist species attached to a
-central, pre existing generalist core \cite{albrecht2010plant}. With preferential
+central, generalist core \cite{albrecht2010plant}. With preferential
 attachment, however, we would also expect connectance and
-specialization to increase, and we found no such trends. The stable level
-of network-level specialization through the assembly process may be
-due to the increased colonization of specialized species
+specialization to increase, and we found no such trends. The stable
+level of network-level specialization through the assembly process may
+be due to the increased colonization of specialized species
 \citep{mgonigle-2015-x} accompanied by an increase in the diet breath
@@
 The changes in network patterns with assembly also increased the
-robustness of the networks to perturbation -- providing further
-evidence that hedgerows are valuable tool for promoting species
-conservation and ecosystem provision in agricultural areas
+robustness of the networks to cascading perturbations -- providing
+further evidence that hedgerows are valuable tool for promoting
+species conservation and ecosystem provision in agricultural areas
 \citep{mgonigle-2015-x, ponisio2015farm, kremen-2015-602}. Because the
-vulnerability of the network to cascading effects, as
-measured by algebraic connectivity, is correlated with species
-richness, the increase and plant and pollinator richness following
-restoration is at least partially responsible for enhancing network
-robustness to cascading effects.
-This result contributes to the stability and diversity debate indicating that diversity might promote statibility, even though there are contrasting results in the literature \citep [e.g.][]{may1972will, pimm1984complexity, montoya2006ecological}.
-However, the robustness of ecological communities to disturbances  might depend on the identity of the species mostly being affected by the disturbance (REF).
-For example, networks might be more sensible to the disturbance of a generalist when compared to a specialist (REF).
+vulnerability of the network to cascading effects, as measured by
+algebraic connectivity, is correlated with species richness, the
+increase and plant and pollinator richness following restoration is at
+least partially responsible for enhancing network robustness to
+cascading effects. The relationship between diversity and stability in
+networks has been the subject of considerable debate
+\citep[e.g.,][]{may1972will, pimm1984complexity,
+  montoya2006ecological}. Our results provides one of the few
+empirical examples of how restoring species diversity contributes to
+enhancing network stability. Through these hedgerows were designed to
+promote floral resrouces across the growing season and not
+specifically to promote any specific network properties
+\citep{menz-2010-4}, the pollinators assembled into resilent
+interaction patterns. With prior knowledge of the floral preferences
+of pollinators, future restoration efforts may also be able to
+incorporate promoting network stability into planning efforts
+\citep{mgonigle2016tool}
+
+%% marilia: not totally sure where you are going here. I was thinking
+%% it could be expanded into a whole paragraph on the fact that we
+%% don't really understand what effects species persistence, and how
+%% that is problematic with understanding netwokr robustness. But
+%% maybe it is a bit too far removed from the main message of the
+%% paper? Maybe something could be added in the last papergraph about
+%% needed more mechanistic models?
+
+%  However, the robustness of ecological
+% communities to disturbances might depend on the identity of the
+% species mostly being affected by the disturbance (REF). For example,
+% networks might be more sensible to the disturbance of a generalist
+% when compared to a specialist (REF).
 

### 2016-08-22 ms/ms.tex
modified discussion

@@
 Given that several changing points in network organization occured
-during assembling, we next explored the species most likely
-responsible for the observed shifts in interaciton patterns. Based on
-a preferential attachment-like mechanism, we expect that the most
-persistent and high degree species would remain stable in the network
-core during assembly, and would thus contribute the least to the
-changing points in network organization. Surprisingly however, we
-encountered the opposite: the species that were most variable in their
-network position and thus contributed most to network reorganizations
-where species with the highest degrees (i.e., most generalized) and
+during assembly, we next explored the species most likely responsible
+for the shifts in interaciton patterns. Based on a preferential
+attachment-like mechanism, we expect that the most persistent and high
+degree species would remain stable in the network core during
+assembly, and would thus contribute the least to the changing points
+in network organization. Surprisingly however, we encountered the
+opposite: the species that were most variable in their network
+position and thus contributed most to network reorganizations where
+species with the highest degrees (i.e., most generalized) and
 persistence. For example, the five most ubiquitous species in our

### 2016-08-23 ms/ms.tex
modified discussion
more discussion modification

@@
 network assembly affect the robustness of the network to species loss
-and to cascading effects. Following \cite{Memmott2004}, we simulate
+and to cascading perturbations. Following \cite{Memmott2004}, we simulate
 the extinction of plant species the subsequent extinction cascades of
@@
 
-We also explored how the robustness to cascading effects changed as
+We also explored how the robustness to cascading perturbations changed as
 community assembled, using algebraic connectivity --- the second
@@
 \citep{fiedler1973algebraic} --- as a proxy for network robustness
-(Gaiarsa et al., submitted). Algebraic connectivity is related to the
-mean diameter of a graph and to how difficult it is to turn a network
+(Gaiarsa et al., submitted). Algebraic connectivity is related to how difficult it is to turn a network
 into completely disconnected groups of nodes, or species
 \citep{costa2007characterization} --- the larger the algebraic
-connectivity, the more robust a network is to cascading effects
+connectivity, the more robust a network is to cascading perturbations
 (Gaiarsa et al., submitted), and the harder it is to break the
 community into isolated groups of species.
-
-%% marilia:
-%% I'm not totally sure what this adds to the methods. Since most
-%% people reading this paper will not understand what "spectral
-%% properties" are (I don't in fact =/ ) it is a bit of a misdirection
-%% since we are not interested in it. I think it is enough to point to
-%% your paper which is using the Alg Con the way we want to.
-
-%% removed:
-% Algebraic connectivity is
-% used to describe spectral properties of complex graphs other than
-% ecological networks \citep[e.g.,][]{fiedler1973algebraic,
-%   costa2007characterization, mohar1991laplacian, de2007old}.
-
-% what's the mean diameter of a graph? Also can we use the
-%% term network instead of graph to be mote consistent with the rest
-%% of the methods? We use graph a bit in the change point section, but
-%% I think it is explained a bit better there?
-
-%% I also removed the e.g.'s because your paper is all we want to
-%% cite, and not "for example Gariarsa et al."
 
@@
   \caption{The robustness of networks to species extinction and
-    to cascading effects. The robustness to species extinction is measured by
+    to cascading perturbations. The robustness to species extinction is measured by
     incrementally removing species by degree, through removing species
     by abundance did not yield qualitatively different results.
-The robustness of networks to cascading effects is measured as the algebraic connectivity, the second smallest eigenvalue of the Laplacian matrix.
+The robustness of networks to cascading perturbations is measured as the algebraic connectivity, the second smallest eigenvalue of the Laplacian matrix.
 Points
@@
 
+%% marilia: I see where you are going witht he bluthgen reference, but
+%% that seemed like something that should be in the intro or
+%% methods. Otherwise it is beinging up a new concept we did not
+%% previously explain.
+
 Given that several changing points in network organization occured
@@
 variable in their network position. Generalized species may be able to
-better exploit the limited floral resources the intensively managed
+better exploit the limited floral resources in the intensively managed
 agriculture landscape, and thus also most persistent. More persistent
@@
 network position to utilize the most advantageous floral resources
-available, which my depend the other pollinator species that are
+available, which my depend on the other pollinator species that are
 present and the state of the floral resource. Thus given the
@@
 %% node/species turnover  - I am not sure I get these results so there might be some nonsense in the below paragraph...
-Given the regional species pool it is possible that specialization migth be constant either because species are not changing through time, or because there is a species turnover, but new species have the same role/posotion as the species there are absent in the following time step.
-We found that node turover was higher in assembling and non-assembling hedgerows when compared to field margin, but there was no difference between them.
-Thus, maybe looking just at network level metrics, such as nestedness and modularity, when exploring the temporal assembly of ecological networks might be misleading, given that there is a node turnover happening at the species level.
+Given the regional species pool it is possible that specialization moght be constant because species are not changing their interactions through time, and are thus always presenting the same degree.
+Alternatively, there might be a species turnover, but new species end up interacting with the same species, and thus, present the same degree as the absent species, and the overall specialization of the network does not change as community assembles.
+Thus, secies turnover probably happens, but new species might present the same role/posotion as the species previous species.
+We found that node turnover was higher in assembling and non-assembling hedgerows when compared to field margin, but there was no difference between them.
+Thus, perhaps looking only at network level metrics, such as nestedness and modularity, when exploring the temporal assembly of ecological networks might be misleading, given that we found node turnover happening at the species level.
 
@@
 \citep{mgonigle-2015-x, ponisio2015farm, kremen-2015-602}. Because the
-vulnerability of the network to cascading effects, as measured by
+vulnerability of the network to cascading perturbations, as measured by
 algebraic connectivity, is correlated with species richness, the
@@
 empirical examples of how restoring species diversity contributes to
-enhancing network stability. Through these hedgerows were designed to
-promote floral resrouces across the growing season and not
+enhancing network stability. Though these hedgerows were designed to
+promote floral resources across the growing season and not
 specifically to promote any specific network properties
@@
 \citep{mgonigle2016tool}
-
-%% marilia: not totally sure where you are going here. I was thinking
-%% it could be expanded into a whole paragraph on the fact that we
-%% don't really understand what effects species persistence, and how
-%% that is problematic with understanding netwokr robustness. But
-%% maybe it is a bit too far removed from the main message of the
-%% paper? Maybe something could be added in the last papergraph about
-%% needed more mechanistic models?
-
-%  However, the robustness of ecological
-% communities to disturbances might depend on the identity of the
-% species mostly being affected by the disturbance (REF). For example,
-% networks might be more sensible to the disturbance of a generalist
-% when compared to a specialist (REF).
 
@@
 %% happening...
-
-% Other interesting result is that even though nestedness increased
-% through time, modularity and connectance did not. \cite{Fortuna2010}
-% suggest that at low connectance levels, mutualistic networks would
-% present high nestedness and high modularity, while at high connectance
-% levels networks would present low nestedness and low
-% modularity. However, our results show the contrary: early communities
-% presented low nestedness, smaller richness, and higher levels of
-% modularity and connectance, while more mature communities presented
-% higher nestedness, higher richness, and smaller modularity and
-% connectance. SO blabla... XX Temporal overlap.
 

### 2016-08-23 analysis/speciesLevel/plotting/.#specMets.R
added plotting code for closeness panel, added to SI

@@ -1,1 +1,0 @@
-lcp@airbears2-10-142-17-35.airbears2.1918.berkeley.edu.1809

### 2016-08-23 analysis/speciesLevel/plotting/src/CIplotting.R

@@ -40,7 +40,7 @@ plot.panel <- function(dats,
        xlim=range(new.dd[,xs], na.rm=TRUE),
        ylim=range(c(0, new.dd$phi,  new.dd$plo,
          quantile(dats[,y1],
-         na.rm=TRUE)[c("25%","75%")]),
+         na.rm=TRUE)[c("1%","75%")]),
          na.rm=TRUE),
        xlab="",
        ylab="",

### 2016-08-25 ms/ms.tex
worked on discussion

@@
     Universidade de Sao Paulo\\
-    Sao Paulo, SP, Brazil\\
+    S{\~a}o Paulo, SP, Brazil\\
     05508-900\\
@@
 systems are particularly at risk from coextinction cascades. The
-nature of these coextinction cascades depends on the interaction
-patterns within a community \citep{Memmott2004, Rezende2007,
+nature of these cascades will depend on the interaction patterns
+within a community \citep{Memmott2004, Rezende2007,
   Bascompte2009}. Recovering the lost biodiversity and interactions
@@
 %% there is some evidence of that. (other studies on mutualisms?)
-The mostly widely explored mechanism of network assembly, preferential
+The most widely explored mechanism of network assembly, preferential
 attachment \citep{barabasi1999emergence}, predicts that a new species
@@
 that new species tended to interact with already well-connected
-species, likely because these species are either more abundant or more
-temporally persistent \citep{Olesen2008}. In addition, using a
+species, potentially because these species are either more abundant or
+more temporally persistent \citep{Olesen2008}. In addition, using a
 space-for-time substitution to study primary succession along a
 glacier foreland, \cite{albrecht2010plant} found some indication
-assembly was occurring through preferential attachment. Network
-nestedness, a pattern of interactions where a generalist core
-interacts with both specialist and generalist species, increased as
-the community aged \citep{albrecht2010plant}. Increasing nestedness
-could result from a process like preferential attachment where
-specialist species attach to the well-connected, generalist core.  In
-addition, non-successional temporal dynamics also suggest a stable
-core of generalists persists despite high turnover of peripheral
-species \citep{fang2012relative, diaz2010changes, alarcon2008year}.
+assembly occurred through preferential attachment. Network nestedness,
+a pattern of interactions where a generalist core interacts with both
+specialist and generalist species, increased as the community aged
+\citep{albrecht2010plant}. Increasing nestedness could result from
+preferential attachment process where specialist species attach to the
+well-connected, generalist core.  % In addition,
+% non-successional temporal dynamics also suggest a stable core of
+% generalists persists despite high turnover of peripheral species
+% \citep{fang2012relative, diaz2010changes, alarcon2008year}.
 
@@
 In contrast to the ordered network build-up described by preferential
-attachment, assembly can be punctuated by significant reorganizations
-of interactions \citep{peel2014detecting}. Such significant
-reorganizations of interactions, or changing points, have been
-observed in social networks responding to abrupt shifts in the
-behavior of interactors \citep{peel2014detecting}.  In ecological
-communities, such shifts my occur if, as new species are added,
-resident species change their interaction partners to optimize their
-foraging effort. In plant-pollinator communities, theory predicts
-pollinators optimize their use of floral resources to reduce
-interspecific competition and improve resource-use efficiency
-\citep{pyke1984optimal, valdovinos2010consequences,
-  valdovinos2013adaptive, albrecht2010plant, Bluthgen2007}. No
-studies, however, have examined whether changing points occur during
-ecological network assembly and how they relate to the behavior of the
-interactors.
+attachment, significant reorganizations of interactions can punctuate
+assembly \citep{peel2014detecting}. Such significant reorganizations
+of interactions, or changing points, are observed in social networks
+responding to abrupt shifts in the behavior of interactors
+\citep{peel2014detecting}.  In ecological communities, such shifts may
+occur if, as new species colonize, resident species change their
+interaction partners to optimize their foraging effort. In
+plant-pollinator communities, theory predicts pollinators optimize
+their use of floral resources to reduce interspecific competition and
+improve resource-use efficiency \citep{pyke1984optimal,
+  valdovinos2010consequences, valdovinos2013adaptive,
+  albrecht2010plant, Bluthgen2007}. No studies, however, have examined
+whether changing points occur during ecological network assembly, and
+how these changes relate to the behavior of the interactors.
 
@@
 \citep[e.g.,]{parker1997scale}. In pollination systems, the time since
-an area was restored has been shown to effect the structure of
+an area was restored has been shown to affect the structure of
 networks \citep{forup-2008-742, forup2008restoration,
-  devoto2012understanding}, suggesting interactions are evolving as the
-community develops. Understanding the mechanisms of network assembly
-will help to guide the restoration of particular communities.
+  devoto2012understanding}, suggesting interactions are evolving as
+the community develops. Understanding the mechanisms of network
+assembly will help to guide community restoration.
 
@@
 global decline \citep{neumann-2010-1, van-engelsdorp-2009-e6481}. In
-addition, native pollinators, which have the capacity to provide
+addition, native pollinators, which are capable of providing
 sufficient crop pollination \citep{kremen-2002-16816,
@@
 habitat for pollinators without removing arable land from
-production. Hedgerows have been shown to augment the richness,
-abundance and trait diversity of pollinators in agricultural
+production. Hedgerows augment the richness, abundance and trait
+diversity of pollinators in agricultural
 landscapes\citep{morandin-2013-829, mgonigle-2015-x, kremen-2015-602,
@@
 colonization of floral resource specialists
-\citep{mgonigle-2015-x}. Little is known however, about the assembly
+\citep{mgonigle-2015-x}. Little is known, however, about the assembly
 of the network following hedgerow restoration.
 
-Using a long-term data-set of plant-pollinator communities assembling
+Using a long-term dataset of plant-pollinator communities assembling
 following hedgerow restoration in the highly simplified and
@@
 attachment, or punctuated by significant reorganizations of
-interactions (i.e., changing points). Even with changing points in
-interaction organization, networks could still be assembling via
-preferential attachment if the network reorganizations were primarily
-driven the by peripheral, temporally variable species while a stable,
-well-connected core of species still persists. We thus examine whether
-the species are most variable in their network position --- and thus
-important contributors network reorganizations --- are less persistent
-and connected species.  Lastly, we examine whether networks are
-assembling toward predictable interaction patterns, and the
-ramifications for the robustness of the networks to species extinction
-and cascading perturbations.
+interactions (i.e., network changing points). Even with changing
+points in interaction organization, networks could still be assembling
+via preferential attachment if the network reorganizations were
+primarily driven the by peripheral, temporally variable species while
+a stable, well-connected core of species still persists. We test
+whether the species are most variable in their network position ---
+and thus important contributors network reorganizations --- are less
+persistent and connected species. To further explore the mechanisms
+underlying the temporal dynamics in the networks, we examine patterns
+in the species and interaction temporal turnover. Lastly, we
+investigate whether networks are assembling toward predictable
+interaction patterns, and the ramifications for the robustness of the
+networks to species extinction and cascading perturbations.
 
@@
 N=29). The sites were located in the Central Valley of California in
-Yolo, Colusa and Solano Counties. This area is comprised of
+Yolo, Colusa and Solano Counties. This area is comprosed of
 intensively managed agriculture -- primarily monocultures of
@@
 along field margins where they do not remove valuable land from
-production, and are ca. 3--6m wide and approximately 350m long and
-border large (ca.\ 30--hectare) crop fields. Hedgerows consist of
-native, perennial, shrub and tree plantings including \textit{Rosa
-  californica}, \textit{Cercis occidentalis}, \textit{Ceanothus spp.},
-\textit{Heteromeles arbutifolia}, \textit{Sambucus mexicana},
-\textit{Eriogonum spp.}, \textit{Baccharis spp.}, \textit{Salvia
-  spp}. and others \citep{kremen-2015-602, mgonigle-2015-x}. The mean
-distance between monitoring sites was 15 km, and the minimum distance
-between sites of the same type sampled in the same year was 2 km.  The
-entire area surveyed spanned almost 300 km$^2$. The crop fields
-adjacent to all sites were similarly managed as intensive, high-input
-monoculture.
+production. Plantings are between are ca. 3--6 m wide and
+approximately 350 m long, bordering large (ca.\ 30--hectare) crop
+fields. Hedgerows consist of native, perennial, shrub and tree
+plantings including \textit{Rosa californica}, \textit{Cercis
+  occidentalis}, \textit{Ceanothus spp.}, \textit{Heteromeles
+  arbutifolia}, \textit{Sambucus mexicana}, \textit{Eriogonum spp.},
+\textit{Baccharis spp.}, \textit{Salvia spp}. and others
+\citep{kremen-2015-602, mgonigle-2015-x}. The mean distance between
+monitoring sites was 15 km, and the minimum distance between sites of
+the same type sampled in the same year was 2 km.  The entire area
+surveyed spanned almost 300 km$^2$. The crop fields adjacent to all
+sites were similarly managed as intensive, high-input monoculture.
 
-Monitoring began in 2006 and continued through 2015. Sampling of the
+Monitoring began in 2006 and continued through 2015. Surveys of the
 assembling hedgerows began the year before the area was restored. For
@@
 in 2010. Sites were sampled between two and five times per year
-(Tables S1-S3). In each round of
-sampling, the order in which sites were sampled was
-randomized. Surveys were conducted under sunny conditions when the
-temperature was above $21^{\circ}\mathrm{C}$ and wind speed was below
-$2.5$ meters/second.
+(Tables S1-S3). In each round of sampling, the order in which sites
+were sampled was randomized. Surveys were conducted under sunny
+conditions when the temperature was above $21^{\circ}\mathrm{C}$ and
+wind speed was below $2.5$ meters/second.
 
@@
 handling specimens). Honeybees (\textit{Apis mellifera}) were not
-collected because their abundance is determined largely by the
-placement of hives throughout the region by bee-keepers. All other
-insect flower visitors that touched the reproductive parts of the
-flower were collected; however, here we focus only on wild bees and
-syrphids (representing 49 and 19 percent of records, respectively),
-the most abundant and effective pollinators in the system (C. Kremen,
-A.~Klein and L.~Morandin, unpublished data). Bee and syrphid specimens
-were identified to species (or morpho-species for some bee specimens
-in the genera \textit{Nomada} and \textit{Sphecodes}) by expert
-taxonomists.
+collected because their abundance is determined largely by hive
+placement by bee-keepers. All other insect flower visitors that
+touched the reproductive parts of the flower were collected; however,
+here we focus only on wild bees and syrphids --- the most abundant and
+effective pollinators in the system (representing 49 and 19 percent of
+records, respectively, C.~Kremen, A.~Klein and L.~Morandin,
+unpublished data). Bee and syrphid specimens were identified to
+species (or morpho-species for some bee specimens in the genera
+\textit{Nomada} and \textit{Sphecodes}) by expert taxonomists.
 
-Quantitative networks were generative for each site through time. To
-account for the unequal number of surveys between years, the mean of
-the interactions between a pair of plants and pollinators across
-surveys within a year was used as a representation of the frequency of
-interactions.
+Quantitative networks were generated for each site through time. To
+account for the unequal number of surveys between years (Tables
+S1-S3), we use the mean of the interactions between a pair of plants
+and pollinators across surveys within a year to represent intearction
+frequency.
 
@@
 We employed a change point detection method \citep{peel2014detecting}
-to identify fundamental changes in large-scale pattern of interactions
-of plants and pollinators. A change point is caused by a merge, split,
-fragmentation or formation of communities (also called modules or
-compartments). Change point detection methods have yet to be
-generalized to quantitative networks, so for this analysis we focused
-on qualitative (binary) networks. Following \cite{peel2014detecting},
-we first defined a probability distribution over the networks using
-the generalized hierarchical random graph model (GHRG). The GHRG model
-is able to capture both assortative and disassortative community
-structure patterns at all scales in the network
-\citep{peel2014detecting}. A network $G$ is composed of vertices $V$
-and edges $E \subseteq {V × V }$. The GHRG model decomposes the $N$
-vertices into a series of nested groups, the relationships among which
-are represented by the dendrogram $T$. The tips of $T$ are the
-vertices of $G$, and the probability that two vertices $u$ and $v$
-connect is given by the parameter $p_r$. The probability distribution
-of the network $G$ thus defined as:
+to identify fundamental reorganizations in large-scale interaction
+patterns. A change point is caused by a merge, split, fragmentation or
+formation of communities (also called modules or compartments). Change
+point detection methods have yet to be generalized to quantitative
+networks, so for this analysis we focused on qualitative (binary)
+networks. Following \cite{peel2014detecting}, we first defined a
+probability distribution over the networks using the generalized
+hierarchical random graph model (GHRG). The GHRG model is able to
+capture both assortative and disassortative community structure
+patterns at all scales in the network \citep{peel2014detecting}. A
+network $G$ is composed of vertices $V$ and edges $E \subseteq {V × V
+}$. The GHRG model decomposes the $N$ vertices into a series of nested
+groups, the relationships among which are represented by the
+dendrogram $T$. The tips of $T$ are the vertices of $G$, and the
+probability that two vertices $u$ and $v$ connect is given by the
+parameter $p_r$. The probability distribution of the network $G$ thus
+defined as:
 
@@
 \citep{peel2014detecting}. This is accomplished by using a Markov
-Chain Monte Carlo (MCMC) procedure to first sample the posterior
+chain Monte Carlo (MCMC) procedure to first sample the posterior
 distribution of bipartitions, from which a consensus tree is derived
-\citep{peel2014detecting}. We used $\beta$ distributions with the
+\citep{peel2014detecting}. We use $\beta$ distributions with the
 hyperparameters $\alpha=\beta=1$ to define priors for $p_r$.
@@
 a change point occurred between two time slices. To detect a change
-point, we compare the fit of two models -- one where a change point
-had occurred between two networks, and one where no change occurred --
+point, we compare the fit of two models --- one where a change point
+occurred between two networks, and one where no change occurred ---
 using posterior Bayes factors. We chose a sliding window of length,
@@
 the Bayes factors, we use parametric bootstrapping to numerically
-estimate the null distribution \citep{peel2014detecting}. The change
-point analysis was carried out using code published online by
-L.~Peel. Analyses we conducted in Python 3.4.
+estimate the null distribution \citep{peel2014detecting}. We emploied
+code published online by L.~Peel for the change point
+analysis. Analyses we conducted in Python 3.4.
 
@@
 generalized linear model with Binomial error to test whether the
-probability of a change point occurring varied by site type. For the
-non-assembling hedgerows and weedy field margins, only sites with five
-or greater years of sampling was included in this analysis. All
-statistical analysis were conducted in R 3.2.3 \citep{R}.
+probability of a change point occurring varied by site type. We used
+standard techniques to determine whether the assumptions of the models
+were met for this and all subsequent models. For the non-assembling
+hedgerows and weedy field margins, only sites with five or greater
+survey years were included in this analysis. All statistical analysis
+were conducted in R 3.2.3 \citep{R}.
 
@@
 To further elucidate the nature of the change points, we examine the
-characteristics of the species that contributed the reorganization of
-interactions. Some species remain in relatively similar network
+characteristics of the species that contributed interaction
+reorganization. Some species remain in relatively similar network
 positions through time, whereas others are more variable in their role
@@
 
-We calculate species persistence as the proportion of the surveys a
-pollinator was observed. Species observed consistently within and
+We calculate species persistence as the proportion of surveys a plant
+or pollinator is observed. Species observed consistently within and
 between years are thus maximally persistent. Weighted species degree
-is calculated from interaction observations observed in more extensive
-data-set from Yolo County (approx.~18000 interaction records) that
-included both the data included in this study and additional data from
-sites where we collected flower visitors using the same methods
-\citep{mgonigle-2015-x, ponisio2015farm}. %% does this make sense to
-%% calculate from an independent data set?
-To represent the the variability of species within networks, we
-computed the coefficient of variation of weighted closeness at each
-site through time. Closeness describes the centrality of a species in
-the network by calculating path lengths to other vertices (species) in
-the graph. We used linear mixed models to test whether the variability
-of species closeness values was related to the persistence or degree
-of that species \citep{lme4, lmetest}. We included random effects for
-species, as well as site. We focused on the pollinator species because
-the hedgerow flowers are planted and thus are not directly
-assembling. Because degree and persistence were strongly correlated,
-($\rho = 0.84$, $p$-value $<$ $2*10^{-16}$), each explanatory variable
-was included in the model separately. Because a linear increase in
-closeness, as might be expected with assembly by preferential
-attachment, would lead to a high variability in closeness scores, we
-also test whether closeness increases through time.
+is calculated from interaction observations from an extensive dataset
+from Yolo County (approx.~18000 interaction records) that included
+both the data included in this study and additional data from sites
+where we collected flower visitors using the same methods
+\citep{mgonigle-2015-x, ponisio2015farm}. To represent network
+position variability, we computed the coefficient of variation of
+weighted closeness at each site through time. Closeness describes the
+centrality of a species in the network by calculating path lengths to
+other vertices (species) in the graph. We use linear mixed models to
+test whether the species closeness variability (log) is related to the
+persistence or degree of that species \citep{lme4, lmetest}. We
+included random effects for species, as well as site. Because the
+degree and persistence of pollinators were strongly correlated, ($\rho
+= 0.84$, $p$-value $<$ $2*10^{-16}$), we include each explanatory
+variable in the model separately. Plant degree and persistent were not
+significantly correlated, but we use the same models as we did for the
+pollinators for consistency. A linear increase in closeness, as might
+be expected with assembly by preferential attachment, would also lead
+to a high variability in closeness scores, so we test whether
+closeness increases through time.
+
+\subsection*{Species and interaction turnover}
+
+Reorganizations of network structure can be the result of species
+turnover or species changing their interaction partners (i.e.,
+re-wiring). To better understand the mechanisms underlying the
+temporal dynamics of the assembling networks, we examined patterns of
+species and interaction turnover. Assembling networks may have higher
+rates of species turnover than non-assembling communities because new
+species are colonizing and establishing themselves
+\citep{mgonigle-2015-x}. Similarly, if pollinators try to maximize
+their foraging efficiency based on the species present, interactions
+may turnover more quickly than in established communities.
+
+To estimate the temporal turnover of species and interactions, we use
+an approach similar to calculating spatial $\beta$-diversity. Instead
+of calculating the variation in community composition across sites
+within a year, however, we estimated turnover across years at a
+site. We first calculated the pairwise dissimilarity of species and
+interactions between years within each site using a dissimilarity
+estimator that incorporates abundances, while also accounting for
+unobserved records \citep{chao-2005-148}. Dissimilarity estimates can
+be affected by the total number of species and individuals sampled at
+a site \citep[e.g.,][]{ponisio2015farm}. For example, the probability
+that two sites do not share any species is higher when there are few
+individuals at those sites. Following \cite{ponisio2015farm}, we use
+null models that constrained species richness to estimate the
+deviation of the observed dissimilarity from that which would be
+expected under a random community assembly process. With the corrected
+dissimilarity values, we then calculated the multivariate dispersion
+of community composition across years \citep{anderson-2011-19}. In
+order to test whether assembling hedgerows had more turnover of
+species and interactions than non-assembling communities, the species
+and interaction temporal turnover estimates were response variables in
+a linear mixed models with site type as an explanatory variable along
+with random effects for both year and site \citep{lme4, lmetest}.
+
+Though species may turnover across years, some groups of species may
+essentially replace each other if they fill similar roles in the
+network. At non-assembling communities, species turnover may
+overestimate the temporal changes in the networks if the interactions
+occurring in one year are similar to those in the next year when they
+are weighted by the similarity of their constituent species. We
+develope a method examine the temporal turnover of interactions
+weighted based on their similarity. We followed \cite{ahn2010link}
+algorithm to hierarchically cluster all the interactions (edges)
+across sites and years based on their similarity, and build a
+dendrogram. The interaction similarity is based how may plants and
+pollinators (vertices) two edges share \citep{ahn2010link,
+  kalinka2011linkcomm}. The more species edges shared in common, the
+shorter the branch length between them on the dendrogram.  We next
+calculated the temporal turnover of interactions weighted by their
+similarity, as approximated by ``phylogenetic'' distance
+\citep{graham2008phylogenetic, picante-2010-1463}. When then use
+linear models to test whether the weighted turnover of interactions
+varied between assembling and non-assembling networks \citep{lme4,
+  lmetest}.
 
 \section*{Temporal changes in interaction patterns}
+
 \subsection*{Network structure}
@@
 network may fragment as the network assembles, enhancing
-modularity. Conversely, nested networks are like pyramid of
+modularity. Conversely, nested networks are like a pyramid of
 interactions, where there are some species that interact with many
@@
 generalist base of the network pyramid (i.e., via preferential
-attachment), nestedness would increase through time. Lastly, if the
-network is accumulating specialist species or species are beginning to
-limit their interaction niche breath as the network assembles, this
-would lead to an increase in the network-level specialization
+attachment), nestedness would increase through time. The connectance
+of a network (the proportion of observed out of possible interaction),
+is realted to the
+
+The specialization 
+The colonization of specialist species or species are beginning to
+limit their interaction niche breath as the network assembles would
+lead to an increase in the network-level specialization
 \citep{bluthgen-2006-9}. To test whether network modularity,
-nestedness or specialization changed linearly with assembly, we used
+nestedness or specialization changed linearly with assembly, we use
 linear mixed models with the descriptive network metrics as the
@@
 interactions between species but fixed the total number of
-interactions, species and the distribution of the interaction
-frequencies \citep{Galeano2009}. Lastly, Network specialization was
-measured using H2, which estimate the deviation of the observed
-interaction frequency between plants and pollinators from a null
-expectation where all partners interact in proportion to their
-abundances \citep{bluthgen-2006-9}. It ranges from 0 for generalized
-networks to 1 for specialized networks.
+interactions, species and interaction frequency distributions
+\citep{Galeano2009}. Lastly, Network specialization was measured using
+H2, which estimate the deviation of the observed interaction frequency
+between plants and pollinators from a null expectation where all
+partners interact in proportion to their abundances
+\citep{bluthgen-2006-9}. It ranges from 0 for generalized networks to
+1 for specialized networks.
 
@@
 network assembly affect the robustness of the network to species loss
-and to cascading perturbations. Following \cite{Memmott2004}, we simulate
-the extinction of plant species the subsequent extinction cascades of
-pollinator species. Because the reproduction of plant species if
-facilitated by active restoration efforts, it is unlikely the
-extinction of pollinators would affect plant populations in the
+and to cascading perturbations. Following \cite{Memmott2004}, we
+simulate the extinction of plant species the subsequent extinction
+cascades of pollinator species. Because the reproduction of plant
+species if facilitated by active restoration efforts, it is unlikely
+the extinction of pollinators would affect plant populations in the
 hedgerows. However, plants ceasing to bloom, for example in response
@@
 abundance, and the number of pollinators that secondarily went extinct
-is calculated. The area below the extinction curve is used as a
-measure of network robustness \citep{bipartite}.
+is calculated. The area below the extinction curve is an estimate of
+network robustness \citep{bipartite}.
 
-We also explored how the robustness to cascading perturbations changed as
-community assembled, using algebraic connectivity --- the second
+We also explored how the robustness to cascading perturbations changed
+as community assembled, using algebraic connectivity --- the second
 smallest eigenvalue of the Laplacian matrix
 \citep{fiedler1973algebraic} --- as a proxy for network robustness
-(Gaiarsa et al., submitted). Algebraic connectivity is related to how difficult it is to turn a network
-into completely disconnected groups of nodes, or species
-\citep{costa2007characterization} --- the larger the algebraic
-connectivity, the more robust a network is to cascading perturbations
-(Gaiarsa et al., submitted), and the harder it is to break the
-community into isolated groups of species.
+(Gaiarsa et al., submitted). Algebraic connectivity is related to how
+difficult it is to turn a network into completely disconnected groups
+of vertexs, or species \citep{costa2007characterization} --- the larger
+the algebraic connectivity, the more robust a network is to cascading
+perturbations (Gaiarsa et al., submitted), and the harder it is to
+break the community into isolated groups of species.
 
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=.8\textwidth]{../analysis/changePoint/plotting/networks.pdf}
-  \caption{The network structure and changing points (vertical red
-    lines) in assembling hedgerows and a representative sample of
-    non-assembling hedgerows and weedy field margins. In each network,
-    plants and pollinators are represented by green and yellow
-    circles, respectively, weighted by their degree. Each species has
-    has a consistent position in the network across years. In the
-    assembling hedgerows, colored squares in the corner of each
-    network represent the years post restoration.}
-  \label{fig:changePoints}
-\end{figure}
-\clearpage
-
 \subsubsection*{Characteristics of species that contribute to change
@@
 attachment, both pollinator persistence and degree were positively
-related to network position variability. (ADD STATS IF KEEPING
-RESULT).
+related to network position variability (estimate of the slope of
+closeness variability and persistence $\pm$ standard error of the
+estimate, $0.653 \pm 0.225$, $p$-value=$0.009$; slope of closeness
+variability and degree, $0.008 \pm 0.002$, $p$-value='$0.002$). Plant
+persistence and degree were not significantly related to network
+position variability (estimate of the slope of closeness variability
+and persistence $\pm$ standard error of the estimate, $-2.063 \pm
+3.091$, $p$-value=$0.5$; slope of closeness variability and degree,
+$0.0018 \pm 0.002$, $p$-value=$0.3$). The variability of species
+network position was not the result of closeness linearly increasing
+through time, and, in fact, plant and pollinator closeness decreased
+decreased slightly through time (Fig.~S1, estimate of the slope of
+closeness through time $\pm$ standard error of the estimate,
+pollinators: $-0.0003 \pm 0.00005$, $p$-value=$2.7*10^{-12}$; plants
+$-0.007 \pm 0.001$, $p$-value=$1.4*10^{-6}$). Through statistically
+significant, the slopes are so slight they may not me biologically
+significant.
 
-\begin{figure}
-  \centering
-  \includegraphics[width=.8\textwidth]{../analysis/variability/figures/cv/occ_degree.pdf}
-  \caption{The coefficient of variation of network position, as
-    represented by closeness, plotted against pollinator persistence
-    and degree. Points represents means for each species across sites.
-    The solid line indicates the mean slope estimate and the dashed
-    lines are the $95\%$ confidence intervals around the estimate. }
-  \label{fig:cv}
-\end{figure}
-\clearpage
+\subsection*{Species and interaction turnover}
+The rates of species and interaction temporal turnover was consistent
+across assembling hedgerows, non-assembling hedgerows and field
+margins, though mature hedgerows has marginally significantly less
+species turnover than field margins (estimate $\pm$ standard error of
+the estimate of the difference in turnover between field margins and
+mature hedgerows, $-0.0498 \pm 0.026$, $p$-value=$0.058$). When
+interactions where weighted by their similarity, both assembling and
+mature hedgerows had high rates of turnover than field margins
+(estimate $\pm$ standard error of the estimate of the difference in
+turnover between field margins and assembling hedgerows, $0.115 \pm
+0.027$, $p$-value=$0.0002$; field margins and mature hedgerows, $0.082
+\pm 0.024$, $p$-value=$0.002$).
 
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=.7\textwidth]{../analysis/networkLevel/figures/baci.pdf}
-  \caption{The change in nestedness, modularity, specialization and
-    connectance as the networks assemble. The left panels represent
-    $z$-scores. Scores greater than $\sim 2$ or less than $\sim -2$
-    are significantly more or less structured than randomly assembled
-    networks. Points are the value for each site at each year of
-    assembly. The solid line indicates the mean slope estimate and the
-    dashed lines are the $95\%$ confidence intervals around the
-    estimate.}
-  \label{fig:baci}
-\end{figure}
-\clearpage
-
 \subsection*{Network robustness}
@@
 $p$-value=$0.042$, Fig.~\ref{fig:rob}).
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.7\textwidth]{../analysis/networkLevel/figures/robustness.pdf}
-  \caption{The robustness of networks to species extinction and
-    to cascading perturbations. The robustness to species extinction is measured by
-    incrementally removing species by degree, through removing species
-    by abundance did not yield qualitatively different results.
-The robustness of networks to cascading perturbations is measured as the algebraic connectivity, the second smallest eigenvalue of the Laplacian matrix.
-Points
-    are the value for each site at each year of assembly. The solid
-    line indicates the mean slope estimate and the dashed lines are
-    the $95\%$ confidence intervals around the estimate.}
-  \label{fig:rob}
-\end{figure}
-\clearpage
 
@@
 study system, hedgerow restoration takes several years to have a
-significant impact on the plant-pollintor communities (Kremen and
+significant impact on the plant-pollinator communities (Kremen and
 M'Gonigle, in prep).
@@
 
-Given that several changing points in network organization occured
+Given that several changing points in network organization occurred
 during assembly, we next explored the species most likely responsible
-for the shifts in interaciton patterns. Based on a preferential
+for the shifts in interaction patterns. Based on a preferential
 attachment-like mechanism, we expect that the most persistent and high
@@
 network position to utilize the most advantageous floral resources
-available, which my depend on the other pollinator species that are
+available, which may depend on the other pollinator species that are
 present and the state of the floral resource. Thus given the
@@
 
-%% node/species turnover  - I am not sure I get these results so there might be some nonsense in the below paragraph...
-Given the regional species pool it is possible that specialization moght be constant because species are not changing their interactions through time, and are thus always presenting the same degree.
-Alternatively, there might be a species turnover, but new species end up interacting with the same species, and thus, present the same degree as the absent species, and the overall specialization of the network does not change as community assembles.
-Thus, secies turnover probably happens, but new species might present the same role/posotion as the species previous species.
-We found that node turnover was higher in assembling and non-assembling hedgerows when compared to field margin, but there was no difference between them.
-Thus, perhaps looking only at network level metrics, such as nestedness and modularity, when exploring the temporal assembly of ecological networks might be misleading, given that we found node turnover happening at the species level.
+% %% vertex/species turnover  - I am not sure I get these results so there might be some nonsense in the below paragraph...
+% Given the regional species pool it is possible that specialization
+% might be constant because species are not changing their interactions
+% through time, and are thus always presenting the same
+% degree. Alternatively, there might be a species turnover, but new
+% species end up interacting with the same species, and thus, present
+% the same degree as the absent species, and the overall specialization
+% of the network does not change as community assembles. Thus, secies
+% turnover probably happens, but new species might present the same
+% role/posotion as the species previous species. We found that vertex
+% turnover was higher in assembling and non-assembling hedgerows when
+% compared to field margin, but there was no difference between
+% them. Thus, perhaps looking only at network level metrics, such as
+% nestedness and modularity, when exploring the temporal assembly of
+% ecological networks might be misleading, given that we found vertex
+% turnover happening at the species level.
 
@@
 of the location of species in networks, and turnover of species and
-nodes all point to an assembly mechanism other than preferential
-attachment. Nestedness did increase with years post restoration, as
-would be expected if colonizing, specialist species attached to a
-central, generalist core \cite{albrecht2010plant}. With preferential
-attachment, however, we would also expect connectance and
+interactions all point to an assembly mechanism other than
+preferential attachment. Nestedness did increase with years post
+restoration, as would be expected if colonizing, specialist species
+attached to a central, generalist core \cite{albrecht2010plant}. With
+preferential attachment, however, we would also expect connectance and
 specialization to increase, and we found no such trends. The stable
@@
 specifically to promote any specific network properties
-\citep{menz-2010-4}, the pollinators assembled into resilent
+\citep{menz-2010-4}, the pollinators assembled into resilient
 interaction patterns. With prior knowledge of the floral preferences
@@
 \citep{holland2006comment}. We need further development of mechanistic
-models of mutualistic stystem to generate testable predictions, along
+models of mutualistic system to generate testable predictions, along
 with empirical exploration of network assembly. Plant-pollinator
@@
 \bibliography{refs}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.8\textwidth]{../analysis/changePoint/plotting/networks.pdf}
+  \caption{Assembling hedgerow networks had more changing points
+    (vertical red lines) and non-assembling hedgerows and weedy field
+    margins (a represented sample of non-assembling sites are depicted
+    here). In each network, plants and pollinators are represented by
+    green and yellow circles, respectively, weighted by their
+    degree. Each species has has a consistent position in the network
+    across years. In the assembling hedgerows, colored squares in the
+    corner of each network represent the years post restoration.}
+  \label{fig:changePoints}
+\end{figure}
+\clearpage
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.8\textwidth]{../analysis/variability/figures/cv/occ_degree.pdf}
+  \caption{The coefficient of variation of network position, as
+    represented by closeness, plotted against pollinator persistence
+    and degree. Persistence and degree were positively related to
+    network position variability in pollinators, but unrelated in
+    plants. Points represents means for each species across sites.
+    The solid line indicates the mean slope estimate and the dashed
+    lines are the $95\%$ confidence intervals around the estimate. }
+  \label{fig:cv}
+\end{figure}
+\clearpage
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.6\textwidth]{../analysis/variability/figures/turnover_panels.pdf}
+  \caption{The pollinator species, interaction and weighted
+    interactions turnover of plant-pollinator networks at
+    non-assembling field margins sites, assembling hedgerows, and
+    non-assembling, mature hedgerows. Rates of species and interaction
+    turnover were similar between site types. However, when
+    interactions where weighted by their uniqueness, both hedgerow
+    types had higher turnover that unrestored field margins. Boxplots
+    represent medians (black horizontal line) first and third
+    quartiles (box perimeter) and extremes (whiskers). }
+  \label{fig:beta}
+\end{figure}
+\clearpage
+
+\begin{figure}
+  \centering
+  \includegraphics[width=1\textwidth]{../analysis/networkLevel/figures/baci.pdf}
+  \caption{The change in nestedness, modularity, specialization and
+    connectance as the networks assemble. The left panels represent
+    $z$-scores. Scores greater than $\sim 2$ or less than $\sim -2$
+    are significantly more or less structured than randomly assembled
+    networks. Points are the value for each site at each year of
+    assembly. The solid line indicates the mean slope estimate and the
+    dashed lines are the $95\%$ confidence intervals around the
+    estimate.}
+  \label{fig:baci}
+\end{figure}
+\clearpage
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.5\textwidth]{../analysis/networkLevel/figures/robustness.pdf}
+  \caption{The robustness of networks to species extinction and
+    to cascading perturbations. The robustness to species extinction is measured by
+    incrementally removing species by degree, through removing species
+    by abundance did not yield qualitatively different results.
+The robustness of networks to cascading perturbations is measured as the algebraic connectivity, the second smallest eigenvalue of the Laplacian matrix.
+Points
+    are the value for each site at each year of assembly. The solid
+    line indicates the mean slope estimate and the dashed lines are
+    the $95\%$ confidence intervals around the estimate.}
+  \label{fig:rob}
+\end{figure}
+\clearpage
 

### 2016-08-25 analysis/variability/beta-int.R

@@ -1,73 +1,0 @@
-rm(list=ls())
-setwd('~/Dropbox/hedgerow_assembly/analysis/variability')
-source('src/initialize.R')
-library(bipartite)
-source('src/calcDis.R')
-traits <- read.csv("../../data/traits.csv")
-## ************************************************************
-## the variability fo interactions of plants and pols
-## ************************************************************
-## for each species at each site, 1) create a interaction partner by
-## year matrix, 2) calculate the dissimilarity of interaction partners
-## between years 3) extract dispersion values for each species, site
-
-load('../../data/networks/expanded_networks.Rdata')
-
-sites <- sapply(strsplit(names(nets), "[.]"), function(x) x[1])
-
-specs.agg <- aggregate(list(k=specs$k,
-                            closeness=specs$weighted.closeness),
-                       list(GenusSpecies=specs$GenusSpecies), mean,
-                       na.rm=TRUE)
-
-plants <- getDis(sites, 1, nets, specs.agg, traits, spec)
-pols <- getDis(sites, 2, nets, specs.agg, traits, spec)
-
-## dipersion of interaction parteners by core/periphery (continuus
-## metric) (categoric metric didn'et make sense bec so few had values
-## > 1)
-mod.pols <- lmer(Dist ~ k + (1|Site) + (1|GenusSpecies),
-                 data=pols)
-
-## only two species with k >1
-hist(unique(data.frame(pols$GenusSpecies, pols$k))$pols.k)
-
-mod.plants <- lmer(Dist~k + (1|Site) + (1|GenusSpecies),
-                   data=plants)
-
-## only one species with k >1
-hist(unique(data.frame(plants$GenusSpecies, plants$k))$plants.k)
-
-## by specialization
-mod.pols.d <- lmer(Dist~d + (1|Site) + (1|GenusSpecies),
-                   data=pols)
-
-mod.plants.d <- lmer(Dist~d + (1|Site) + (1|GenusSpecies),
-                     data=plants)
-
-## pollinator date occurrence
-mod.pols.occ <- lmer(Dist~occ.date + (1|Site) + (1|GenusSpecies),
-                     data=pols)
-
-## closeness
-mod.pols.close <- lmer(Dist~ closeness + (1|Site) + (1|GenusSpecies),
-                 data=pols)
-
[…]

### 2016-08-25 analysis/variability/cv.R

@@ -121,13 +121,21 @@ summary(plants.degree.closeness.cv$lm.nss)
 
 
 ## check correlation of degree and occ 
-occ.k.sd$data$degree <-
-  degree$data$traits.ns[match(occ.k.sd$data$GenusSpecies,
-                              degree$data$GenusSpecies)]
+## pollinators
+check.pol <- unique(cbind(spec$degree,
+                          spec$occ.date))
+plot(check.pol)
 
-plot(occ.k.sd$data$traits.ns ~ occ.k.sd$data$degree)
+cor.test(check.pol[,1], check.pol[,2])
 
-cor.test(occ.k.sd$data$traits.ns, occ.k.sd$data$degree)
+
+check.plant <- unique(cbind(spec$plant.degree,
+                            spec$occ.plant.date))
+plot(check.plant)
+
+cor.test(check.plant[,1], check.plant[,2])
+
+
 
 
 ## ************************************************************

### 2016-08-25 analysis/variability/src/phyloIntBeta.R

@@ -3,22 +3,26 @@
 ## takes raw specimen data (spec.dat), the , 
 calcCommDis <- function(spec.dat, type, lc, abund.w=TRUE){
   ## prep community of sites and interactions
+ 
   prep.comm <- aggregate(spec.dat[, type],
                          list(site=spec.dat$Site,
                               year=spec.dat$Year,
                               status=spec.dat$SiteStatus,
                               sp=spec.dat[, type]),
                          length)
+  
   ## split by status then by site
   bystatus <- split(prep.comm, prep.comm$status)
   bysite1 <- lapply(bystatus, function(x){split(x, x$site)})
   bysite <- unlist(bysite1, recursive=FALSE)
+
   ## create a site by interaction community
   comm <- lapply(bysite, function(y) {
     samp2site.spp(y$year, y$sp, y$x)
   })
   ## drop sites that were only samples one year
-  comm <- comm[lapply(comm,nrow) > 1]
+  comm <- comm[lapply(comm, nrow) >= 3]
+
   ## calculate the number of years between sampling times
   years <- lapply(comm, rownames)
   dist.time <- as.matrix(dist(unique(spec.dat$Year), diag=TRUE))
@@ -42,7 +46,11 @@ calcCommDis <- function(spec.dat, type, lc, abund.w=TRUE){
                      assignCols(a,b),
                      a=comm,
                      b=node.order)
+
+  ## calculate "phylogenetic" distance between interactions
+  
   dist.tree <- cophenetic(lc$hclust)
+
   ## calculate the dissimilarity between years based on the dengrogram
   comm.dis <- lapply(comm.int, comdist,
                      dis=dist.tree,

### 2016-08-25 ms/ms.tex
deleated some sections marilia wrote that did not quite fit

@@
 temporal dynamics of the assembling networks, we examined patterns of
-species and interaction turnover. Assembling networks may have higher
-rates of species turnover than non-assembling communities because new
-species are colonizing and establishing themselves
+pollinator species and interaction turnover. Assembling networks may
+have higher rates of species turnover than non-assembling communities
+because new species are colonizing and establishing themselves
 \citep{mgonigle-2015-x}. Similarly, if pollinators try to maximize
@@
 
-To estimate the temporal turnover of species and interactions, we use
-an approach similar to calculating spatial $\beta$-diversity. Instead
-of calculating the variation in community composition across sites
-within a year, however, we estimated turnover across years at a
-site. We first calculated the pairwise dissimilarity of species and
+To estimate the temporal species and interaction turnover, we use an
+approach similar to calculating spatial $\beta$-diversity. Instead of
+calculating the variation in community composition across sites within
+a year, however, we estimated turnover across years at a site. We
+first calculated the pairwise dissimilarity of species and
 interactions between years within each site using a dissimilarity
@@
 attachment), nestedness would increase through time. The connectance
-of a network (the proportion of observed out of possible interaction),
-is realted to the
+--- the proportion of observed out of possible interactions --- is
+also expected to increase as species preferentially attach to the
+core. The network specialization may also change as the community
+assembles. If specialist species colonize the network or species begin
+to limit their interaction niche breath as the network assembles,
+network-level specialization will increase \citep{bluthgen-2006-9}.
 
-The specialization 
-The colonization of specialist species or species are beginning to
-limit their interaction niche breath as the network assembles would
-lead to an increase in the network-level specialization
-\citep{bluthgen-2006-9}. To test whether network modularity,
-nestedness or specialization changed linearly with assembly, we use
-linear mixed models with the descriptive network metrics as the
-response variable, year of assembly as the explanatory variable, and
-random effects of site and year.
-
-We use NODF (weighted or unweighted) to evaluate network nestedness
-\citep{nodf}. NODF evaluates whether species with fewer partners
+To evaluate network nestedness, we used the estimator weighted NODF
+citep{nodf}. NODF evaluates whether species with fewer partners
 interact with subsets of partners with which more connected species
@@
 interactions, species and interaction frequency distributions
-\citep{Galeano2009}. Lastly, Network specialization was measured using
-H2, which estimate the deviation of the observed interaction frequency
-between plants and pollinators from a null expectation where all
-partners interact in proportion to their abundances
-\citep{bluthgen-2006-9}. It ranges from 0 for generalized networks to
-1 for specialized networks.
+\citep{Galeano2009}. We evaluate network specialization with the
+metric H2, which estimate the deviation of the observed interaction
+frequency between plants and pollinators from a null expectation where
+all partners interact in proportion to their abundances
+\citep{bluthgen-2006-9}. It ranges from zero for generalized networks
+to one for specialized networks.
+
+To test whether network modularity, nestedness, connectance or
+specialization changed linearly with assembly, we use linear mixed
+models with the descriptive network metrics as the response variable,
+year of assembly as the explanatory variable, and random effects of
+site and year. The number of species in a network affects affects the
+patterns of interaction possible, so we also examined the change in
+plant and pollinator species richness through time. We employ
+generalized linear mixed models with Poisson error to model
+richness. We scaled explanatory variables.
 
@@
 and to cascading perturbations. Following \cite{Memmott2004}, we
-simulate the extinction of plant species the subsequent extinction
+simulate plant species extinction and the subsequent extinction
 cascades of pollinator species. Because the reproduction of plant
 species if facilitated by active restoration efforts, it is unlikely
-the extinction of pollinators would affect plant populations in the
-hedgerows. However, plants ceasing to bloom, for example in response
-to drought, will likely affect the pollinators that depend on
-them. Plants species were eliminated based on their degree or
-abundance, and the number of pollinators that secondarily went extinct
-is calculated. The area below the extinction curve is an estimate of
+the extinction of pollinator species would affect plant populations in
+the hedgerows. However, plants ceasing to bloom, for example in
+response to drought, will likely affect the pollinators that depend on
+them. We eliminate plants species based on their degree or abundance,
+and the number of pollinators that secondarily went extinct is
+calculated. The area below the extinction curve is an estimate of
 network robustness \citep{bipartite}.
@@
 \citep{fiedler1973algebraic} --- as a proxy for network robustness
-(Gaiarsa et al., submitted). Algebraic connectivity is related to how
+(Gaiarsa et al., submitted). Algebraic connectivity relates to how
 difficult it is to turn a network into completely disconnected groups
-of vertexs, or species \citep{costa2007characterization} --- the larger
-the algebraic connectivity, the more robust a network is to cascading
-perturbations (Gaiarsa et al., submitted), and the harder it is to
-break the community into isolated groups of species.
+of vertices, or species \citep{costa2007characterization} --- the
+larger the algebraic connectivity, the more robust a network is to
+cascading perturbations (Gaiarsa et al., submitted), and the harder it
+is to break the community into isolated groups of species.
 
@@
 
-The majority ($76\%$) of the sites tests underwent at least one
-significant reorganization of interactions
-(Fig.~\ref{fig:changePoints}).  All five of the assembling hedgerows
-experienced changing points, whereas only $40\%$ and $81\%$ of
-non-assembling hedgerows and field margins, respectively, underwent
-significant interaction reorganizations. Assembling hedgerows
-experienced significantly more changing points than the non-assembling
-networks (difference in the odds ratios between assembling and
-non-assembling networks, $3.316$, $95\%$ CI $[1.314, 8.572]$,
-$p$-value= $0.0117$). Network assembly following restoration is thus
-punctuated by more interaction reorganizations than would be expected
-by environmental shifts that would effect assembling and
-non-assembling networks equally.
+The majority ($76\%$) of the sites underwent at least one significant
+interaction reorganization (Fig.~\ref{fig:changePoints}).  All five of
+the assembling hedgerows experienced changing points, whereas only
+$40\%$ and $81\%$ of non-assembling hedgerows and field margins,
+respectively, underwent significant interaction
+reorganizations. Assembling hedgerows experienced significantly more
+changing points than the non-assembling networks (estimate of the
+difference in the odds ratios between assembling and non-assembling
+networks, $3.316$, $95\%$ CI $[1.314, 8.572]$, $p$-value=
+$0.0117$). Network assembly following restoration is thus punctuated
+by more interaction reorganizations than would be expected by
+environmental shifts that would effect assembling and non-assembling
+networks equally.
 
@@
 estimate, $0.653 \pm 0.225$, $p$-value=$0.009$; slope of closeness
-variability and degree, $0.008 \pm 0.002$, $p$-value='$0.002$). Plant
-persistence and degree were not significantly related to network
-position variability (estimate of the slope of closeness variability
-and persistence $\pm$ standard error of the estimate, $-2.063 \pm
-3.091$, $p$-value=$0.5$; slope of closeness variability and degree,
-$0.0018 \pm 0.002$, $p$-value=$0.3$). The variability of species
-network position was not the result of closeness linearly increasing
-through time, and, in fact, plant and pollinator closeness decreased
+variability and degree, $0.008 \pm 0.002$, $p$-value='$0.002$). In
+addition, plant persistence and degree were not significantly related
+to network position variability (estimate of the slope of closeness
+variability and persistence $\pm$ standard error of the estimate,
+$-2.063 \pm 3.091$, $p$-value=$0.5$; slope of closeness variability
+and degree, $0.0018 \pm 0.002$, $p$-value=$0.3$). The variability of
+species network position was not the result of closeness linearly
+increasing through time, and, in fact, plant and pollinator closeness
 decreased slightly through time (Fig.~S1, estimate of the slope of
@@
 \subsection*{Species and interaction turnover}
-The rates of species and interaction temporal turnover was consistent
-across assembling hedgerows, non-assembling hedgerows and field
-margins, though mature hedgerows has marginally significantly less
-species turnover than field margins (estimate $\pm$ standard error of
-the estimate of the difference in turnover between field margins and
+The rates of species and interaction temporal turnover similar across
+assembling hedgerows, non-assembling hedgerows and field margins,
+though mature hedgerows has marginally significantly less species
+turnover than field margins (estimate $\pm$ standard error of the
+estimate of the difference in turnover between field margins and
 mature hedgerows, $-0.0498 \pm 0.026$, $p$-value=$0.058$). When
 interactions where weighted by their similarity, both assembling and
-mature hedgerows had high rates of turnover than field margins
+mature hedgerows had higher rates of turnover than field margins
 (estimate $\pm$ standard error of the estimate of the difference in
@@
 0.027$, $p$-value=$0.0002$; field margins and mature hedgerows, $0.082
-\pm 0.024$, $p$-value=$0.002$).
+\pm 0.024$, $p$-value=$0.002$). The weighted interaction turnover at
+assembling hedgerows, however, was not significantly higher than in
+non-assembling, mature hedgerows.
 
@@
 the slope of nestedness through time $\pm$ standard error of the
-estimate, $1.834 \pm 0.6142$, $p$-value=$0.022$, Fig.~
-\ref{fig:baci}).  Modularity decreased (Fig.~\ref{fig:baci}), though
-the slope was not significantly different from zero (estimate of the
-slope of modularity through time $\pm$ standard error of the estimate,
-$-0.524$ $\pm$ $0.295$, $p$-value=$0.124$). Specialization remained
-relatively constant (estimate of the slope of specialization through time
-$\pm$ standard error of the estimate, $0.003$ $\pm$ $0.015$,
+estimate, $1.834 \pm 0.6142$, $p$-value=$0.022$, Fig.~\ref{fig:baci}).
+Modularity decreased (Fig.~\ref{fig:baci}), though the slope was not
+significantly different from zero (estimate of the slope of modularity
+through time $\pm$ standard error of the estimate, $-0.524$ $\pm$
+$0.295$, $p$-value=$0.124$). Connectance decreased as the community
+assembled (Fig.~\ref{fig:baci}, estimate of the slope of connectance
+through time $\pm$ standard error of the estimate, $-0.0434$ $\pm$
+$0.0152$, $p$-value=$0.03$). In contrast, specialization remained
+constant (estimate of the slope of specialization through time $\pm$
+standard error of the estimate, $0.003$ $\pm$ $0.015$,
 $p$-value=$0.827$).
+
+Both plant and pollinator species richness increased through time
+(Fig.~\ref{fig:baci}, estimate of the slope of richness through time
+$\pm$ standard error of the estimate, pollinators: $0.193$ $\pm$
+$0.0729$, $p$-value=$0.008$; plants:$0.212$ $\pm$ $0.0653$,
+$p$-value=$0.001$). Pollinator species are colonizing and persisting
+at the assembling hedgerows. Though some new plant species may
+eshablish themselves in the hedgerows, because the plant species
+richness is based on the flowers in the network and not the presence
+of a plant at a site, the increase in plant richness is likely due to
+plants attracting more visitors as they offer better rewards.
 
@@
 
-%% marilia: I see where you are going witht he bluthgen reference, but
-%% that seemed like something that should be in the intro or
-%% methods. Otherwise it is beinging up a new concept we did not
-%% previously explain.
-
 Given that several changing points in network organization occurred
@@
 change their network positions.
-
-% %% vertex/species turnover  - I am not sure I get these results so there might be some nonsense in the below paragraph...
-% Given the regional species pool it is possible that specialization
-% might be constant because species are not changing their interactions
-% through time, and are thus always presenting the same
-% degree. Alternatively, there might be a species turnover, but new
-% species end up interacting with the same species, and thus, present
-% the same degree as the absent species, and the overall specialization
-% of the network does not change as community assembles. Thus, secies
-% turnover probably happens, but new species might present the same
-% role/posotion as the species previous species. We found that vertex
-% turnover was higher in assembling and non-assembling hedgerows when
-% compared to field margin, but there was no difference between
-% them. Thus, perhaps looking only at network level metrics, such as
-% nestedness and modularity, when exploring the temporal assembly of
-% ecological networks might be misleading, given that we found vertex
-% turnover happening at the species level.
 

### 2016-08-25 ms/ms.tex
added last paragraph to discussion

@@
 is not possible. Interestingly, most of the assembling hedgerows did
-not undergo a significant reorganization of interactions immediately
-the hedgerow was planted (i.e., the transition from weedy field margin
-to hedgerow). This result is consistent with the finding that in our
+not undergo a significant interaction reorganization immediately the
+hedgerow was planted (i.e., the transition from weedy field margin to
+hedgerow). This result is consistent with the finding that in our
 study system, hedgerow restoration takes several years to have a
@@
 change their network positions.
+
+Interestingly, though assembling hedgerows had more network
+reorganizations than non-assembling communities, pollinator species
+and interaction turnover occurred at similar rates across site
+types. Only when interactions were weighted by their uniqueness did
+assembling hedgerows have higher turnover than non-assembling field
+margins. This is likely because though species and interactions are
+turning over at the field margins, but pollinator species and
+interactions that fill similar roles in the network are replacing each
+other. In contrast, at the assembling hedgerows, unique interactions
+are turning over as the networks continually
+reorganize. Non-assembling mature hedgerow communities, however, had
+similar rates of weighted interaction turnover as assembling hedgerows
+but also the lowest species turnover. Pollinator communities at mature
+hedgerows may be generally more stable, but rare and/or specialized
+pollinators could generate this pattern if they entered a community,
+formed interactions with plants that did not previously share
+pollinators, but did not have high persistence. ANYTHING ELSE? mature
+hedgerows are important for conservation or rare species? Maybe?
 

### 2016-08-26 ms/ms.tex
finished first full draft of manuscript, added plants to turnover analyses

--- 2016-08-25 ms/ms.tex
@@
 \begin{abstract}
-  TO BE RE-WRITTEN
   The structure of networks is related to ability of communities to
   maintain function in the face of species extinction. Understanding
-  network structure and how it relates to network disassembly,
-  therefore, is a priority for system-level conservation biology.  We
-  explore the assembly of plant-pollinator communities on native plant
-  restorations in the Central Valley of California. 
+  network structure and how it relates to network assembly, therefore,
+  is a priority for system-level conservation biology. We explore the
+  assembly of plant-pollinator communities on native plant
+  restorations in the Central Valley of California. Species are highly
+  dynamic in their positions in the networks, such that we find that
+  assembly is punctuated by significant reorganizations of
+  interactions. Non-assembling mature hedgerows and weedy field
+  margins did not undergo as many network reorganizations, suggesting
+  environmental factors alone cannot account for the changes observed
+  in assembling communities. Pollinator species were highly
+  opportunistic, and in fact the most persistent and generalized
+  species were most variable in their network positions --- the
+  opposite of what would be expected by preferential attachment. High
+  species and interaction turnover was ubiquitous across assembling
+  and non-assembling communities, though unique interactions turn over
+  at higher rates in assembling hedgerows as the networks continually
+  reorganize. Assembling networks also increased in nestedness through
+  time, though this did not make them more robust to simulated plant
+  extinctions. The resilience of networks to cascading perturbations,
+  however, increased as the communities assembled, at least partially
+  due to accumulating species richness. Our results elucidate some of
+  the mechanisms underlying plant-pollinator network assembly and
+  restoration, while providing further evidence that hedgerows are a
+  valuable tool for promoting species conservation and ecosystem
+  provision in agricultural areas.
 \end{abstract}
@@
 The most widely explored mechanism of network assembly, preferential
-attachment \citep{barabasi1999emergence}, predicts that a new species
-is more likely to interact with species that are already
-well-connected \citep[''the rich-get-richer''
+attachment \citep{barabasi1999emergence}, predicts that species
+entering a network are more likely to interact with species that are
+already well-connected \citep[''the rich-get-richer''
 principle,][]{barabasi1999emergence}. In pollination systems --- a
@@
 space-for-time substitution to study primary succession along a
-glacier foreland, \cite{albrecht2010plant} found some indication
+glacier foreland, \cite{albrecht2010plant} found some evidence that
 assembly occurred through preferential attachment. Network nestedness,
@@
 assembly \citep{peel2014detecting}. Such significant reorganizations
-of interactions, or changing points, are observed in social networks
-responding to abrupt shifts in the behavior of interactors
+of interactions, or network changing points, are observed in social
+networks responding to abrupt shifts in the behavior of interactors
 \citep{peel2014detecting}.  In ecological communities, such shifts may
@@
 whether changing points occur during ecological network assembly, and
-how these changes relate to the behavior of the interactors.
+how these changes relate to the species behavior.
 
@@
 restoration, which is essentially 'applied succession'
-\citep[e.g.,]{parker1997scale}. In pollination systems, the time since
-an area was restored has been shown to affect the structure of
+\citep[e.g.,][]{parker1997scale}. In pollination systems, the time
+since an area was restored has been shown to affect the structure of
 networks \citep{forup-2008-742, forup2008restoration,
@@
 %% in ag. but ag is a bad place for pollinators
-Facilitating effective restoration of networks is especially
-imperative in areas where species interactions provide essential
-ecosystem services, such as crop pollination. In intensively managed
-agricultural landscapes, the demand for pollination services is the
-greatest \citep{kremen-2008-10}. However, honey bees, managed
-extensively around the world to provide crop pollination, are in
-global decline \citep{neumann-2010-1, van-engelsdorp-2009-e6481}. In
-addition, native pollinators, which are capable of providing
-sufficient crop pollination \citep{kremen-2002-16816,
-  winfree-2007-1105, kremen-2004-1109}, are in short supply because
-these landscapes make poor habitats for pollinator populations
-\citep{kremen-2002-16816}. To ensure provision the continued provision
-of ecosystem services and curb biodiversity loss, effective
-restoration of pollinators and their interactions in agricultural
-landscapes is critical.
+Facilitating network restoration is especially imperative in areas
+where species interactions provide essential ecosystem services, such
+as crop pollination. In intensively managed agricultural landscapes,
+the demand for pollination services is the greatest
+\citep{kremen-2008-10}. However, honey bees, managed extensively
+around the world to provide crop pollination, are in global decline
+\citep{neumann-2010-1, van-engelsdorp-2009-e6481}. In addition, native
+pollinators, which are capable of providing sufficient crop
+pollination \citep{kremen-2002-16816, winfree-2007-1105,
+  kremen-2004-1109}, are in short supply because these landscapes make
+poor habitats for pollinator populations \citep{kremen-2002-16816}. To
+ensure provision the continued provision of ecosystem services and
+curb biodiversity loss, effective restoration of pollinators and their
+interactions in agricultural landscapes is critical.
 
 %% hedgerows!!  
-To promote pollinator services in agriculture, farmers are
-increasingly turning to the habitat restoration technique of planting
-strips of native plants along farm edges (hedgerows) to help provide
-habitat for pollinators without removing arable land from
+To promote pollinator services in agriculture, farmers may chose to
+plant strips of native plants along farm edges (hedgerows) to help
+provide habitat for pollinators without removing arable land from
 production. Hedgerows augment the richness, abundance and trait
-diversity of pollinators in agricultural
-landscapes\citep{morandin-2013-829, mgonigle-2015-x, kremen-2015-602,
+diversity of pollinators in agricultural landscapes
+\citep{morandin-2013-829, mgonigle-2015-x, kremen-2015-602,
   ponisio2015farm}. In addition, hedgerows promote the persistence and
 colonization of floral resource specialists
-\citep{mgonigle-2015-x}. Little is known, however, about the assembly
-of the network following hedgerow restoration.
+\citep{mgonigle-2015-x}. Little is known, however, about how new
+species are being incorporated into the network as the community
+assembles, or the consequences for interaction patterns and
+robustness.
 
@@
 primarily driven the by peripheral, temporally variable species while
-a stable, well-connected core of species still persists. We test
-whether the species are most variable in their network position ---
-and thus important contributors network reorganizations --- are less
-persistent and connected species. To further explore the mechanisms
-underlying the temporal dynamics in the networks, we examine patterns
-in the species and interaction temporal turnover. Lastly, we
-investigate whether networks are assembling toward predictable
-interaction patterns, and the ramifications for the robustness of the
-networks to species extinction and cascading perturbations.
+a stable, well-connected core of species persist. We test whether the
+species are most variable in their network position --- and thus
+important contributors network reorganizations --- are less persistent
+and connected species. To further explore the mechanisms underlying
+the temporal dynamics in the networks, we examine patterns in the
+species and interaction temporal turnover. Lastly, we investigate
+whether networks are assembling toward predictable interaction
+patterns, and the ramifications for the robustness of the networks to
+species extinction and cascading perturbations.
 
@@
 N=29). The sites were located in the Central Valley of California in
-Yolo, Colusa and Solano Counties. This area is comprosed of
-intensively managed agriculture -- primarily monocultures of
-conventional row crops, vineyards and orchards. Hedgerows we planted
-along field margins where they do not remove valuable land from
-production. Plantings are between are ca. 3--6 m wide and
+Yolo, Colusa and Solano Counties. This area is composed of intensively
+managed agriculture --- primarily monocultures of conventional row
+crops, vineyards and orchards. Hedgerows we planted along field
+margins where they do not remove valuable land from
+production. Plantings are between are ca.~3--6 m wide and
 approximately 350 m long, bordering large (ca.\ 30--hectare) crop
@@
 
-Monitoring began in 2006 and continued through 2015. Surveys of the
+Monitoring began in 2006 and continued through 2014. Surveys of the
 assembling hedgerows began the year before the area was restored. For
@@
 S1-S3), we use the mean of the interactions between a pair of plants
-and pollinators across surveys within a year to represent intearction
+and pollinators across surveys within a year to represent interaction
 frequency.
@@
 patterns at all scales in the network \citep{peel2014detecting}. A
-network $G$ is composed of vertices $V$ and edges $E \subseteq {V × V
-}$. The GHRG model decomposes the $N$ vertices into a series of nested
-groups, the relationships among which are represented by the
-dendrogram $T$. The tips of $T$ are the vertices of $G$, and the
-probability that two vertices $u$ and $v$ connect is given by the
-parameter $p_r$. The probability distribution of the network $G$ thus
-defined as:
+network $G$ is composed of vertices $V$ and edges $E$. The GHRG model
+decomposes the $N$ vertices into a series of nested groups, the
+relationships among which are represented by the dendrogram $T$. The
+tips of $T$ are the vertices of $G$, and the probability that two
+vertices $u$ and $v$ connect is given by the parameter $p_r$. The
+probability distribution of the network $G$ thus defined as:
 
@@
 for more gradual changes, and four was the maximum possible with our
-maximum of nine years of data. Lastly, to calculate a $p$-value for
+maximum of eight years of data. Lastly, to calculate a $p$-value for
 the Bayes factors, we use parametric bootstrapping to numerically
-estimate the null distribution \citep{peel2014detecting}. We emploied
+estimate the null distribution \citep{peel2014detecting}. We employed
 code published online by L.~Peel for the change point
@@
 and thus contribute more strongly to network reorganization. We test
-the that more persistent species with the highest degree are the most
-stable in their network position, as would be expected is the networks
-were assembling via preferential attachment.
+whether the more persistent species with the highest degree are the
+most stable in their network positions, as would be expected if the
+networks were assembling via preferential attachment.
 
@@
 temporal dynamics of the assembling networks, we examined patterns of
-pollinator species and interaction turnover. Assembling networks may
-have higher rates of species turnover than non-assembling communities
-because new species are colonizing and establishing themselves
+species and interaction turnover. Assembling networks may have higher
+rates of pollinator turnover than non-assembling communities because
+new pollinator species are colonizing and establishing themselves
 \citep{mgonigle-2015-x}. Similarly, if pollinators try to maximize
 their foraging efficiency based on the species present, interactions
-may turnover more quickly than in established communities.
+may turnover more quickly than in established communities. In
+addition, at assembling hedgerows, plants may ``colonize'' the
+networks as they become more attractive resources.
 
@@
 a year, however, we estimated turnover across years at a site. We
-first calculated the pairwise dissimilarity of species and
+first calculated the pairwise dissimilarity of plants, pollinators and
 interactions between years within each site using a dissimilarity
@@
 species and interactions than non-assembling communities, the species
-and interaction temporal turnover estimates were response variables in
-a linear mixed models with site type as an explanatory variable along
-with random effects for both year and site \citep{lme4, lmetest}.
+and interaction temporal turnover estimates were included as response
+variables in a linear mixed models with site type as an explanatory
+variable along with a random effect of site \citep{lme4, lmetest}.
 
@@
 are weighted by the similarity of their constituent species. We
-develope a method examine the temporal turnover of interactions
+develop a method examine the temporal turnover of interactions
 weighted based on their similarity. We followed \cite{ahn2010link}
@@
 interactions are insular, occurring within separate groups or
-``modules'' more often than between modules. Communities in the
-network may fragment as the network assembles, enhancing
+``modules'' more often than between modules. Modules in the network
+may fragment as the network assembles, enhancing
 modularity. Conversely, nested networks are like a pyramid of
@@
 and so on. If species entering the network tend to interact with the
-generalist base of the network pyramid (i.e., via preferential
-attachment), nestedness would increase through time. The connectance
---- the proportion of observed out of possible interactions --- is
-also expected to increase as species preferentially attach to the
-core. The network specialization may also change as the community
-assembles. If specialist species colonize the network or species begin
-to limit their interaction niche breath as the network assembles,
-network-level specialization will increase \citep{bluthgen-2006-9}.
+generalist base of the network pyramid as would be expected with
+preferential attachment, nestedness would increase through time. The
+connectance --- the proportion of observed out of possible
+interactions --- is also expected to increase as species
+preferentially attach to the core. \flagged{Is this true?} Lastly, the
+network specialization may change as the community assembles. If
+specialist species colonize the network or species begin to limit
+their interaction niche breath as the network assembles, network-level
+specialization will increase \citep{bluthgen-2006-9}.
 
 To evaluate network nestedness, we used the estimator weighted NODF
-citep{nodf}. NODF evaluates whether species with fewer partners
+\citep{nodf}. NODF evaluates whether species with fewer partners
 interact with subsets of partners with which more connected species
@@
 \citep{Galeano2009}. We evaluate network specialization with the
-metric H2, which estimate the deviation of the observed interaction
+metric H2, which estimates the deviation of the observed interaction
 frequency between plants and pollinators from a null expectation where
@@
 \subsection*{Network robustness}
-Lastly, we test the changes in interaction patterns associated with
-network assembly affect the robustness of the network to species loss
-and to cascading perturbations. Following \cite{Memmott2004}, we
+Lastly, we test whether the changes in interaction patterns associated
+with network assembly affect the robustness of the network to species
+loss and to cascading perturbations. Following \cite{Memmott2004}, we
 simulate plant species extinction and the subsequent extinction
@@
 them. We eliminate plants species based on their degree or abundance,
-and the number of pollinators that secondarily went extinct is
-calculated. The area below the extinction curve is an estimate of
-network robustness \citep{bipartite}.
+and then calculate the number of pollinators that secondarily went
+extinct. The area below the extinction curve is an estimate of network
+robustness \citep{Memmott2004, bipartite}.
 
@@
 difficult it is to turn a network into completely disconnected groups
-of vertices, or species \citep{costa2007characterization} --- the
-larger the algebraic connectivity, the more robust a network is to
-cascading perturbations (Gaiarsa et al., submitted), and the harder it
-is to break the community into isolated groups of species.
+of species \citep{costa2007characterization}. The larger the algebraic
+connectivity, the more robust a network is to cascading perturbations,
+and the harder it is to break the community into isolated groups of
+species (Gaiarsa et al., submitted).
 
@@
 interaction reorganization (Fig.~\ref{fig:changePoints}).  All five of
-the assembling hedgerows experienced changing points, whereas only
-$40\%$ and $81\%$ of non-assembling hedgerows and field margins,
+the assembling hedgerows experienced network changing points, whereas
+only $40\%$ and $81\%$ of non-assembling hedgerows and field margins,
 respectively, underwent significant interaction
-reorganizations. Assembling hedgerows experienced significantly more
-changing points than the non-assembling networks (estimate of the
-difference in the odds ratios between assembling and non-assembling
-networks, $3.316$, $95\%$ CI $[1.314, 8.572]$, $p$-value=
-$0.0117$). Network assembly following restoration is thus punctuated
-by more interaction reorganizations than would be expected by
-environmental shifts that would effect assembling and non-assembling
-networks equally.
+reorganizations. Assembling hedgerows had significantly more changing
+points than the non-assembling networks (estimate of the difference in
+the odds ratios between assembling and non-assembling networks,
+$3.316$, $95\%$ CI $[1.314, 8.572]$, $p$-value= $0.0117$). Network
+assembly following restoration is thus punctuated by more interaction
+reorganizations than would be expected by environmental shifts alone
+that would effect assembling and non-assembling networks equally.
 
@@
 estimate, $0.653 \pm 0.225$, $p$-value=$0.009$; slope of closeness
-variability and degree, $0.008 \pm 0.002$, $p$-value='$0.002$). In
+variability and degree, $0.008 \pm 0.002$, $p$-value=$0.002$). In
 addition, plant persistence and degree were not significantly related
@@
 \subsection*{Species and interaction turnover}
-The rates of species and interaction temporal turnover similar across
-assembling hedgerows, non-assembling hedgerows and field margins,
-though mature hedgerows has marginally significantly less species
+The rates of plant, pollinator and interaction temporal turnover were
+similar across assembling hedgerows, non-assembling hedgerows and
+field margins, though mature hedgerows has marginally significantly
+less pollinator turnover than field margins (estimate $\pm$ standard
+error of the estimate of the difference in turnover between field
+margins and mature hedgerows, $-0.0498 \pm 0.026$,
+$p$-value=$0.058$). When interactions where weighted by their
+similarity, both assembling and mature hedgerows had higher rates of
 turnover than field margins (estimate $\pm$ standard error of the
 estimate of the difference in turnover between field margins and
-mature hedgerows, $-0.0498 \pm 0.026$, $p$-value=$0.058$). When
-interactions where weighted by their similarity, both assembling and
-mature hedgerows had higher rates of turnover than field margins
-(estimate $\pm$ standard error of the estimate of the difference in
-turnover between field margins and assembling hedgerows, $0.115 \pm
-0.027$, $p$-value=$0.0002$; field margins and mature hedgerows, $0.082
-\pm 0.024$, $p$-value=$0.002$). The weighted interaction turnover at
-assembling hedgerows, however, was not significantly higher than in
+assembling hedgerows, $0.115 \pm 0.027$, $p$-value=$0.0002$; field
+margins and mature hedgerows, $0.082 \pm 0.024$,
+$p$-value=$0.002$). The weighted interaction turnover at assembling
+hedgerows, however, was not significantly higher than in
 non-assembling, mature hedgerows.
@@
 \subsection*{Network structure}
-Network nestedness significantly increased with assembly (estimate of
-the slope of nestedness through time $\pm$ standard error of the
-estimate, $1.834 \pm 0.6142$, $p$-value=$0.022$, Fig.~\ref{fig:baci}).
-Modularity decreased (Fig.~\ref{fig:baci}), though the slope was not
-significantly different from zero (estimate of the slope of modularity
-through time $\pm$ standard error of the estimate, $-0.524$ $\pm$
-$0.295$, $p$-value=$0.124$). Connectance decreased as the community
-assembled (Fig.~\ref{fig:baci}, estimate of the slope of connectance
-through time $\pm$ standard error of the estimate, $-0.0434$ $\pm$
-$0.0152$, $p$-value=$0.03$). In contrast, specialization remained
-constant (estimate of the slope of specialization through time $\pm$
-standard error of the estimate, $0.003$ $\pm$ $0.015$,
-$p$-value=$0.827$).
+Network nestedness significantly increased with assembly
+(Fig.~\ref{fig:baci}, estimate of the slope of nestedness through time
+$\pm$ standard error of the estimate, $1.834 \pm 0.6142$,
+$p$-value=$0.022$). Modularity decreased (Fig.~\ref{fig:baci}),
+though the slope was not significantly different from zero (estimate
+of the slope of modularity through time $\pm$ standard error of the
+estimate, $-0.524$ $\pm$ $0.295$, $p$-value=$0.124$). Connectance
+decreased as the community assembled (Fig.~\ref{fig:baci}, estimate of
+the slope of connectance through time $\pm$ standard error of the
+estimate, $-0.0434$ $\pm$ $0.0152$, $p$-value=$0.03$). In contrast,
+specialization remained constant (estimate of the slope of
+specialization through time $\pm$ standard error of the estimate,
+$0.003$ $\pm$ $0.015$, $p$-value=$0.827$).
 
@@
 $\pm$ standard error of the estimate, pollinators: $0.193$ $\pm$
-$0.0729$, $p$-value=$0.008$; plants:$0.212$ $\pm$ $0.0653$,
-$p$-value=$0.001$). Pollinator species are colonizing and persisting
-at the assembling hedgerows. Though some new plant species may
-eshablish themselves in the hedgerows, because the plant species
-richness is based on the flowers in the network and not the presence
-of a plant at a site, the increase in plant richness is likely due to
-plants attracting more visitors as they offer better rewards.
+$0.0729$, $p$-value=$0.008$; plants: $0.212$ $\pm$ $0.0653$,
+$p$-value=$0.001$). Unsurprisingly, pollinator species are colonizing
+and persisting at the assembling hedgerows. Though some new plant
+species may establish themselves in the hedgerows, because the plant
+species richness is based on the flowers in the network and not the
+presence of a plant at a site, the increase in plant richness is
+likely due to plants attracting visitors as they offer better rewards.
 
@@
 measured by the algebraic connectivity of the network, increased as
-the network assembled (estimate of the slope of robustness through
-time $\pm$ standard error of the estimate, $0.6814 \pm 0.272$,
-$p$-value=$0.042$, Fig.~\ref{fig:rob}).
+the network assembled (estimate of the slope of robustness to
+cascading perturbations through time $\pm$ standard error of the
+estimate, $0.6814 \pm 0.272$, $p$-value=$0.042$, Fig.~\ref{fig:rob}).
 
@@
 often undergo significant reorganizations. These network organizations
-are unlikely to be a product of environmental forces alone because the
+are unlikely a product of environmental forces alone because the
 network changing points in non-assembling communities are less
-frequent, and there are few consistent trends in when change points
-occurred across all sites. Several sites had changing points between
-year 2009 and 2011 (Fig.~\ref{fig:changePoints}). In California, 2011
-marked the beginning of a multi-year drought. In the assembling
-hedgerows were not sampled in 2010, so disentangling whether the
-changing points are due to skipping a year of assembly or the drought
-is not possible. Interestingly, most of the assembling hedgerows did
-not undergo a significant reorganization of interactions immediately
-the hedgerow was planted (i.e., the transition from weedy field margin
-to hedgerow). This result is consistent with the finding that in our
-study system, hedgerow restoration takes several years to have a
-significant impact on the plant-pollinator communities (Kremen and
-M'Gonigle, in prep).
+frequent, and few consistent trends in when change points occurred
+across all sites. Several sites had network changing points between
+years 2009 and 2011 (Fig.~\ref{fig:changePoints}). In California, 2011
+marked the beginning of a multi-year drought. The assembling hedgerows
+were not sampled in 2010, so disentangling whether the changing points
+are due to skipping a year of assembly or the drought is not
+possible. Interestingly, most assembling hedgerows did not undergo a
+significant interaction reorganization immediately after a hedgerow
+was planted (i.e., the transition from weedy field margin to
+hedgerow). This result is consistent with the finding that in our
+study system, hedgerow restoration takes several years to have an
+impact on the plant-pollinator communities (Kremen and M'Gonigle, in
+prep).
 
-Given that several changing points in network organization occurred
-during assembly, we next explored the species most likely responsible
-for the shifts in interaction patterns. Based on a preferential
-attachment-like mechanism, we expect that the most persistent and high
-degree species would remain stable in the network core during
-assembly, and would thus contribute the least to the changing points
-in network organization. Surprisingly however, we encountered the
-opposite: the species that were most variable in their network
-position and thus contributed most to network reorganizations where
-species with the highest degrees (i.e., most generalized) and
-persistence. For example, the five most ubiquitous species in our
-study landscape --- \textit{Halictus ligatus}, \textit{Halictus
-  tripartitus}, \textit{Lasioglossum (Dialictus) incompletum}, and
-\textit{Toxomerus marginatus} --- were the only species that changed
-what module (i.e., community), and they were also present in across
-years in all of the assembling hedgerows. Because species degree and
-persistence were strongly correlated, it is difficult to disentangle
-the causal mechanism for why species with those characteristics are so
-variable in their network position. Generalized species may be able to
-better exploit the limited floral resources in the intensively managed
-agriculture landscape, and thus also most persistent. More persistent
-species also have longer phenologies, so they have the opportunity to
-visit many different flowers, resulting in a higher degree. Either
-way, our result suggests that adaptable species can change their
-network position to utilize the most advantageous floral resources
-available, which may depend on the other pollinator species that are
-present and the state of the floral resource. Thus given the
-opportunity and ability to use different resources, species will often
-change their network positions.
+In addition to finding multiple network organization changing points
+during assembly, the species that contribute most to those
+reorganizations were not what is expected by preferential
+attachment. Based on a preferential attachment-like mechanism, we
+expect that the most persistent and high degree species would remain
+stable in the network core during assembly. Surprisingly, however, we
+encountered the opposite pattern. For example, the five most
+ubiquitous species in our study landscape --- \textit{Halictus
+  ligatus}, \textit{Halictus tripartitus}, \textit{Lasioglossum
+  (Dialictus) incompletum}, and \textit{Toxomerus marginatus} --- were
+the only species that changed what module (i.e., community) of which
+they were a member across years in all the assembling
+hedgerows. Because species degree and persistence were strongly
+correlated, it is difficult to disentangle the causal mechanism for
+why species with those characteristics are so variable in their
+network position. Generalized species may be able to better exploit
+the limited floral resources in the intensively managed agriculture
+landscape, and thus also most persistent. More persistent species also
+have longer phenologies, so they can visit many different flowers,
+resulting in a higher degree. Either way, our result suggests that
+adaptable species can change their network position to utilize the
+most advantageous floral resources available, which may depend on the
+other pollinator species that are present and the state of the plant
+community. Thus given the opportunity and ability to use different
+resources, species will often change their network positions.
+
+Interestingly, though assembling hedgerows had more network
+reorganizations than non-assembling communities, pollinator species
+and interaction turnover occurred at similar rates across site types.
+Assembling hedgerows have higher turnover than non-assembling field
+margins only when interactions were weighted by their similarity. This
+is likely because though species and interactions are turning over at
+the field margins, species and interactions that fill similar roles in
+the network are replacing each other. In contrast, at the assembling
+hedgerows, unique interactions are turning over as the networks
+continually reorganize. Non-assembling mature hedgerow communities,
+however, had similar rates of weighted interaction turnover as
+assembling hedgerows but also the lowest pollinator
+turnover. Pollinator communities at mature hedgerows may be generally
+more stable, but rare and/or specialized pollinators could generate
+this pattern if they entered a community, formed unique interactions
+with plants that did not previously share pollinators, but did not
+persist in the networks. These species would not contribute strongly
+to network reorganization or species turnover, but would enhance
+weighted interaction turnover. Mature hedgerows may thus both support
+more stable pollination communities and likely ecosystem services,
+while also providing resources for rare and/or specialized species
+\citep{kremen-2015-602, mgonigle-2015-x}.
 
 The frequent changing points in network organization, dynamic nature
-of the location of species in networks, and turnover of species and
+of species positions in the networks, and turnover of species and
 interactions all point to an assembly mechanism other than
-preferential attachment. Nestedness did increase with years post
-restoration, as would be expected if colonizing, specialist species
-attached to a central, generalist core \cite{albrecht2010plant}. With
-preferential attachment, however, we would also expect connectance and
-specialization to increase, and we found no such trends. The stable
-level of network-level specialization through the assembly process may
-be due to the increased colonization of specialized species
-\citep{mgonigle-2015-x} accompanied by an increase in the diet breath
-of resident species. This would be expected if resident species were
-able to minimize their foraging time by expanding their diet breath as
-plant diversity increases with hedgerow maturation \citep{Waser1996,
+preferential attachment. Nestedness did increase with years
+post-restoration, as would be expected if colonizing, specialist
+species attached to a central, generalist core
+\citep{albrecht2010plant}. With preferential attachment, however, we
+would also expect connectance and specialization to increase, and we
+found no such trends. \flagged{Wait if connectance increases doesn't
+  that mean specialization decreases? Which would we expect to change
+  with preferential attachment?} The stable level of network-level
+specialization through the assembly process may be due to the
+increased colonization of specialized species \citep{mgonigle-2015-x}
+accompanied by an increase in the diet breath of resident
+species. This would be expected if resident species were able to
+minimize their foraging time by expanding their diet breath as plant
+diversity increases with hedgerow maturation \citep{Waser1996,
   pyke1984optimal, Bluthgen2007, albrecht2010plant}. Such a change in
@@
 robustness of the networks to cascading perturbations -- providing
-further evidence that hedgerows are valuable tool for promoting
+further evidence that hedgerows are a valuable tool for promoting
 species conservation and ecosystem provision in agricultural areas
-\citep{mgonigle-2015-x, ponisio2015farm, kremen-2015-602}. Because the
-vulnerability of the network to cascading perturbations, as measured by
+\citep{mgonigle-2015-x, ponisio2015farm, kremen-2015-602}. Because
+network vulnerability to cascading perturbations, as measured by
 algebraic connectivity, is correlated with species richness, the
@@
 \citep[e.g.,][]{may1972will, pimm1984complexity,
-  montoya2006ecological}. Our results provides one of the few
-empirical examples of how restoring species diversity contributes to
-enhancing network stability. Though these hedgerows were designed to
-promote floral resources across the growing season and not
-specifically to promote any specific network properties
-\citep{menz-2010-4}, the pollinators assembled into resilient
-interaction patterns. With prior knowledge of the floral preferences
-of pollinators, future restoration efforts may also be able to
-incorporate promoting network stability into planning efforts
-\citep{mgonigle2016tool}
+  montoya2006ecological}. Our results provide one of the few empirical
+examples of how restoring species diversity contributes to enhancing
+network stability. Though these hedgerows were designed to promote
+floral resources across the growing season and not specifically to
+promote any specific network properties \citep{menz-2010-4}, the
+pollinators assembled into resilient interaction patterns. With prior
+knowledge of the floral preferences of pollinators, future restoration
+efforts may also be able to incorporate promoting network stability
+into planning efforts \citep{mgonigle2016tool}
 
@@
 associated with an enhanced in robustness to extinction
-\citep{Memmott2004}. It may be that assembling hedgerows have not yet
-reached a sufficient level of nestedness to realize its
-benefits. Nestedness of the assembling hedgerows, however, did not
+\citep{Memmott2004}. Assembling hedgerows have yet to reach sufficient
+levels of nestedness to realize the benefits nestedness
+confers. Nestedness of the assembling hedgerows, however, did not
 asymptote within the eight years following restoration that the sites
@@
 
-Plant-pollinator networks in general are highly dynamic, with high
+Plant-pollinator networks, in general, are highly dynamic, with high
 turnover of species and interactions both within and between seasons
@@
 literature, we have few mechanistic models of mutualistic network
-assembly \citep{valdovinos2013adaptive}. In addition, the few that
-have been developed often borrow their mechanisms from competitive
-interactions, leading to inaccurate biological assumptions
+assembly \citep{valdovinos2013adaptive}. In addition, the few
+developed often borrow their mechanisms from competitive interactions,
+leading to inaccurate biological assumptions
 \citep{holland2006comment}. We need further development of mechanistic
@@
 with empirical exploration of network assembly. Plant-pollinator
-communities and mutualisms in general are vital for the maintenance of
-biodiversity and the provision of essential ecosystem services. We
-must therefore understand the processes underlying their assembly to
+communities and mutualisms broadly are vital for biodiversity
+maintenance and essential ecosystem service provision. We must
+therefore understand the processes underlying their assembly to
 facilitate restoration and conservation.
-
-% Though \cite{Olesen2008} found some evidence for an
-% assmbly process like prefrential attachment governing the temporal
-% build up of plant-pollinator networks within a season, the frequent
-% reorganizations of the hedgerow communities suggest a very different
-% mechanism is at play.
-
-%% not sure if we want to go here, since it is not really part of the
-%% assembly story. More just a "hmm networks are structured in odd
-%% ways sometimes" and we dont' really have answer for why this is
-%% happening...
 
@@
 (USDA-NIFA 2012-51181-20105 to Michigan State University).  Funding
-for LCP was provided by an NSF Graduate Research Fellowship and the
-USDA NIFA Graduate Fellowship. Funding for MPG was provided by S{\~a}o
-Paulo Research Foundation (FAPESP, grant 2013/13319-5).
+for LCP was provided by an NSF Graduate Research Fellowship, the USDA
+NIFA Graduate Fellowship, and the Berkeley Institute for Data
+Science. Funding for MPG was provided by S{\~a}o Paulo Research
+Foundation (FAPESP, grant 2013/13319-5). We also appreciate the Santa
+Fe Institute for faciliting this international collaboration.
 
@@
   \caption{Assembling hedgerow networks had more changing points
-    (vertical red lines) and non-assembling hedgerows and weedy field
-    margins (a represented sample of non-assembling sites are depicted
-    here). In each network, plants and pollinators are represented by
-    green and yellow circles, respectively, weighted by their
-    degree. Each species has has a consistent position in the network
-    across years. In the assembling hedgerows, colored squares in the
-    corner of each network represent the years post restoration.}
+    (vertical red lines) than non-assembling hedgerows and weedy field
+    margins (a representative sample of non-assembling sites are
+    depicted here). In each network, plants and pollinators are
+    represented by green and yellow circles, respectively, weighted by
+    their degree. Each species has a consistent position in the
+    network across years. In the assembling hedgerows, colored squares
+    in the corner of each network represent the year of
+    assembly. Before hedgerows are installed (when they are still
+    field margins) the year of assembly is zero.}
   \label{fig:changePoints}
+\end{figure}
+\clearpage
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.8\textwidth]{../analysis/changePoint/plotting/communitiesSkanky.pdf}
+  \caption{The community (i.e., module or compartment) membership
+    between network changing points. Two representative assembling
+    hedgerows are depicted. In the top panel, species are grouped by
+    community. The bottom panels visualize the flow of species between
+    communities between changing points. }
+  \label{fig:changePoints2}
 \end{figure}
@@
     network position variability in pollinators, but unrelated in
-    plants. Points represents means for each species across sites.
-    The solid line indicates the mean slope estimate and the dashed
-    lines are the $95\%$ confidence intervals around the estimate. }
+    plants. Points represent means for each species across sites. The
+    solid line indicates the mean slope estimate and the dashed lines
+    are the $95\%$ confidence intervals around the estimate. }
   \label{fig:cv}
@@
   \centering
-  \includegraphics[width=.6\textwidth]{../analysis/variability/figures/turnover_panels.pdf}
-  \caption{The pollinator species, interaction and weighted
-    interactions turnover of plant-pollinator networks at
-    non-assembling field margins sites, assembling hedgerows, and
-    non-assembling, mature hedgerows. Rates of species and interaction
-    turnover were similar between site types. However, when
-    interactions where weighted by their uniqueness, both hedgerow
+  \includegraphics[width=.7\textwidth]{../analysis/variability/figures/turnover_panels.pdf}
+  \caption{The species, interaction and weighted interactions turnover
+    of plant-pollinator networks at non-assembling field margins
+    sites, assembling hedgerows, and non-assembling, mature
+    hedgerows. Rates of species and interaction turnover were similar
+    between site types, though mature hedgerows has marginally
+    significantly less pollinator turnover.  However, when
+    interactions where weighted by their similarity, both hedgerow
     types had higher turnover that unrestored field margins. Boxplots
@@
   \includegraphics[width=1\textwidth]{../analysis/networkLevel/figures/baci.pdf}
-  \caption{The change in nestedness, modularity, specialization and
-    connectance as the networks assemble. The left panels represent
-    $z$-scores. Scores greater than $\sim 2$ or less than $\sim -2$
-    are significantly more or less structured than randomly assembled
-    networks. Points are the value for each site at each year of
-    assembly. The solid line indicates the mean slope estimate and the
-    dashed lines are the $95\%$ confidence intervals around the
-    estimate.}
+  \caption{The nestedness, plant richness and pollinator richness
+    increased as the networks assembled. Specialization and modularity
+    remained consistent across years, while connectance decreased. The
+    left panels represent $z$-scores. Scores greater than $\sim 2$ or
+    less than $\sim -2$ are significantly more or less structured than
+    randomly assembled networks. Points are the metric value for each
+    site at each year of assembly. The solid line indicates the mean
+    slope estimate and the dashed lines are the $95\%$ confidence
+    intervals around the estimate.}
   \label{fig:baci}
@@
   \includegraphics[width=.5\textwidth]{../analysis/networkLevel/figures/robustness.pdf}
-  \caption{The robustness of networks to species extinction and
-    to cascading perturbations. The robustness to species extinction is measured by
-    incrementally removing species by degree, through removing species
-    by abundance did not yield qualitatively different results.
-The robustness of networks to cascading perturbations is measured as the algebraic connectivity, the second smallest eigenvalue of the Laplacian matrix.
-Points
-    are the value for each site at each year of assembly. The solid
-    line indicates the mean slope estimate and the dashed lines are
-    the $95\%$ confidence intervals around the estimate.}
+  \caption{The robustness of networks to species extinction did not
+    change with network assembly, but the resilience to to cascading
+    perturbation increased. The robustness to species extinction is
+    measured by incrementally removing species by degree, through
+    removing species by abundance did not yield qualitatively
+    different results. The robustness of networks to cascading
+    perturbations is measured as the algebraic connectivity, the
+    second smallest eigenvalue of the Laplacian matrix. Points are the
+    value for each site at each year of assembly. The solid line
+    indicates the mean slope estimate and the dashed lines are the
+    $95\%$ confidence intervals around the estimate.}
   \label{fig:rob}
@@
 \clearpage
-
-% Communities assembling through preferential attachment are formed by a
-% core of generalist species, to which new species arriving to the
-% community interact with. Thus, communities assembling through
-% preferential attachment exhibit an increase both in nestedness and in
-% the overall level of species specialization, since specialist species
-% are being added to the community. However, our changing point analysis
-% results show that there is a critical reorganization in the
-% interaction patterns as communities assembly, which contradicts the
-% preferential attachment theory. Even though we encountered an increase
-% in nestedness through time, specialization did not
-% increase. Furthermore, this increase in nestedness could be explained
-% by the increase in species richness. blablabla mechanistic models.
-
-% Evidence indicates that there is a positive relationship between
-% abundance and diet breadth, with the more generalist species having
-% higher abundances. Because of the high generalism, species more
-% abundant also present higher degrees. Thus, we expected that more
-% generalist species would be part of the network core. Surprisingly,
-% the five most common species were also the ones that exhibited the
-% most changes in interaction patterns. These species (names!) were
-% always present in the periphery of the network. One hypotheses is that
-% their high generalism allows them to explore the resources more
-% broadly, allowing them to change their position in the network. XX
-% mention the closeness variability and the species persistence.
-
-% The increase in nestedness has been related to an increase in the
-% robustness of communities to species extinctions because the core of
-% generalist species would act as a buffer against species loss
-% (REF). Similarly, communities with greater modularity would be more
-% robust in general, because modules would act as a buffer against
-% perturbation spreading and the pervasive effects of species loss. We
-% expected that as communities assemble and species richness increase,
-% communities would become increasingly robust to species loss. however,
-% we found that robustness to species loss did not change through
-% time. This is interestingly from the conservation biology point of
-% view, because even early assembling communities seem to be as robust
-% as more mature communities, despite the richness insurance
-% hypothesis. However, communities are not only subjected to species
-% loss, but also to other types of perturbations that can cascade
-% through the network. Gaiarsa et al. (submitted) suggest using
-% algebraic connectivity to explore how vulnerable ecoogical communities
-% might be to cascading effects. Our results indicate that more mature
-% communities are more robust to cascading effects than early on
-% assemblages, and that this result is related to species richness. This
-% migth be related to XX....
-
-% \subsection*{Core and peripherial communities}
-
-% To classify species as we first generate dendrograms using the GHRG
-% model before and after each change point.  We then determined which
-% species belonged to the ``core'' and ``peripheral'' network
-% communities at each site. The ``core'' network communities contain the
-% majority of species and are more basil than the more derived, less
-% specious ``peripheral'' network communities. We next use a
-% Permutational Multivariate Analysis of Variance (PERMANOVA)
-% \citep{anderson-2013-557} to determine whether the species
-% compositions of the species that belonged to the core and peripheral
-% communities differed. We also test whether core or peripheral species
-% had more variability in their species compositions \citep[i.e.,
-% multivariate dispersion][]{anderson-2011-19, anderson-2006-683}.
-
-% Lastly, we explored the functional diversity of the core and
-% peripheral pollinator species
-
-% The syrphids in our study area have similar traits, so we focused on
-% the trait diversity of the bee subset of our interaction networks. We
-% focus on resource capture and use traits that collectively
-% \citep{kremen-2015-602} including resource specialization
-% \citep[quantitative, $d'$;][]{bluthgen-2006-9}, body size
-% \citep[quantitative, inter-tegular span, mm,][]{cane-1987-145},
-% sociality (categorical: eusocial, solitary, cleptoparasitic), nest
-% location (categorical: above ground, below ground or mix), and nest
-% construction \citep[categorical: excavator, mason or
-% renter;][]{williams-2010-2280} as described in more detail in
-% \cite{kremen-2015-602}.  Each trait has the same weight in trait
-% diversity metric estimation \citep{villeger-2008-2290,
-% schleuter-2010-469}.
-
-% To characterize the functional diversity of the bee species in core
-% and peripherial communities, we computed three metrics that capture
-% diversity, uniqueness, and distribution of functional values in the
-% community: functional dispersion, divergence, and evenness
-% \citep{villeger-2008-2290, schleuter-2010-469}.  Functional
-% dispersion is a measure of trait diversity, corrected for species
-% richness \citep{schleuter-2010-469}; functional divergence measures
-% how species abundances are distributed within the trait space
-% \citep[i.e., a measure of functional
-% uniqueness,][]{villeger-2008-2290}; functional evenness measures the
-% regularity with which traits are distributed across functional
-% space, accounting for abundance \citep{villeger-2008-2290}.  In
-% combination, these metrics provide a relatively complete overview of
-% the different aspects of functional diversity
-% \citep{villeger-2008-2290, schleuter-2010-469}. To determine whether
-% trait evenness, dispersion, and divergence differed between controls
-% and hedgerows at different stages of maturation, we used the trait
-% diversity metrics as response variables in linear mixed models with
-% site type as a fixed effect and year and site as random effects
-% \citep{lme4,lmetest}.
 

### 2016-08-26 analysis/variability/beta-div.R

@@ -3,37 +3,37 @@ setwd('~/Dropbox/hedgerow_assembly/analysis/variability')
 binary <- FALSE
 alpha <- TRUE
 ## int or pols
-type <- "int"
+type <- "pols"
 source('src/initialize_beta.R')
 
 ## ************************************************************
 ## beta diversity as variation between years,
 ## centroid for each site
 ## ************************************************************
-dis.pols <- mapply(function(a, b, c, d)
-                   beta.status(comm= a,
-                               status= b,
-                               dis.method,
-                               nulls=c,
-                               occ=binary,
-                               years=d,
-                               sub="pols"),
-                   a=comm.pols$comm,
-                   b=comm.pols$status,
-                   c= nulls.pols,
-                   d= comm.pols$comm,
-                   SIMPLIFY=FALSE)
+dis <- mapply(function(a, b, c, d)
+              beta.status(comm= a,
+                          status= b,
+                          dis.method,
+                          nulls=c,
+                          occ=binary,
+                          years=d,
+                          sub=type),
+              a=comm$comm,
+              b=comm$status,
+              c= nulls,
+              d= comm$comm,
+              SIMPLIFY=FALSE)
 
-dats.pols <- data.frame(site=comm.pols$sites,
-                        status=unlist(comm.pols$status),
-                        year=unlist(comm.pols$years),
-                        dist=unlist(sapply(dis.pols, function(x)
-                          x$distances)))
-rownames(dats.pols) <- NULL
+dats <- data.frame(site=comm$sites,
+                   status=unlist(comm$status),
+                   year=unlist(comm$years),
+                   dist=unlist(sapply(dis, function(x)
+                     x$distances)))
+rownames(dats) <- NULL
 
-save(dats.pols, file= file.path('saved/speciesTurnover',
-                  sprintf('%s.pdf', paste(dis.method, alpha, occ, type,
-                                          sep='_'))))
+save(dats, file= file.path('saved/speciesTurnover',
+             sprintf('%s.pdf', paste(dis.method, alpha, occ, type,
+                                     sep='_'))))
 
 ## invid nulls all not sig, alpha nulls mature marginally sig less,
 ## occurrence nulls all not sig
@@ -44,17 +44,17 @@ load(file= file.path('saved/speciesTurnover', sprintf('%s.pdf',
[…]

### 2016-08-26 analysis/variability/src/initialize_beta.R

@@ -6,53 +6,39 @@ source('src/misc.R')
 source('src/beta.R')
 source('src/plotDists.R')
 
-if(type == "pols"){
-  if(!binary & alpha){
-    occ <- "abund"
-    dis.method <- "chao"
-    load(file='saved/communities/pols-abund.Rdata')
-    load(file='saved/nulls/pols-alpha.Rdata')
-  }
-
-  if(!binary & !alpha){
-    occ <- "indiv"
-    dis.method <- "chao"
-    load(file='saved/communities/pols-abund.Rdata')
-    load(file='saved/nulls/pols-indiv.Rdata')
-  }
-
-  if(binary){
-    occ <- "occ"
-    dis.method <- "jaccard"
-    load(file='saved/communities/pols-abund.Rdata')
-    load(file='saved/nulls/pols-occ.Rdata')
-  }
-  ylabel <- "Species turnover"
+if(!binary & alpha){
+  occ <- "abund"
+  dis.method <- "chao"
+  load(file=file.path('saved/communities',
+         sprintf('%s-abund.Rdata', type)))
+  load(file=file.path('saved/nulls',
+         sprintf('%s-alpha.Rdata', type)))
 }
 
+if(!binary & !alpha){
+  occ <- "indiv"
+  dis.method <- "chao"
+  load(file=file.path('saved/communities',
+         sprintf('%s-abund.Rdata', type)))
+  load(file=file.path('saved/nulls',
+         sprintf('%s-indiv.Rdata', type)))
+}
 
-if(type == "int"){
-  if(!binary & alpha){
-    occ <- "abund"
-    dis.method <- "chao"
-    load(file='saved/communities/int-abund.Rdata')
-    load(file='saved/nulls/int-alpha.Rdata')
-  }
+if(binary){
+  occ <- "occ"
+  dis.method <- "jaccard"
+  load(file=file.path('saved/communities',
+         sprintf('%s-abund.Rdata', type)))
+  load(file=file.path('saved/nulls',
+         sprintf('%s-occ.Rdata', type)))
+}
 
-  if(!binary & !alpha){
-    occ <- "indiv"
-    dis.method <- "chao"
-    load(file='saved/communities/int-abund.Rdata')
-    load(file='saved/nulls/int-indiv.Rdata')
-  }
-
-  if(binary){
-    occ <- "occ"
-    dis.method <- "jaccard"
-    load(file='saved/communities/int-abund.Rdata')
[…]

### 2016-08-26 ms/ms.tex
finished first full draft of manuscript, added plants to turnover analyses

@@
   \centering
-  \includegraphics[width=.8\textwidth]{../analysis/changePoint/plotting/communitiesSkanky.pdf}
+  \includegraphics[width=.9\textwidth]{../analysis/changePoint/plotting/communitiesSkanky.pdf}
   \caption{The community (i.e., module or compartment) membership
@@
   \centering
-  \includegraphics[width=.7\textwidth]{../analysis/variability/figures/turnover_panels.pdf}
+  \includegraphics[width=.8\textwidth]{../analysis/variability/figures/turnover_panels.pdf}
   \caption{The species, interaction and weighted interactions turnover

### 2016-08-26 ms/ms.tex
changed the final verison of the ms, added the svg files for the sanky plots figure

@@
 
-\begin{abstract}
+\begin{abstract} %% Lauren - I would invert the phrase just so it does not beging with metwork structure - "communities' ability to maintain function in the face of species extinction is related to the structure fo networks".
   The structure of networks is related to ability of communities to
@@
   network structure and how it relates to network assembly, therefore,
-  is a priority for system-level conservation biology. We explore the
-  assembly of plant-pollinator communities on native plant
+  is a priority for system-level conservation biology. We explore how the
+  assembly of plant-pollinator communities occured, on native plant
   restorations in the Central Valley of California. Species are highly
-  dynamic in their positions in the networks, such that we find that
-  assembly is punctuated by significant reorganizations of
+  dynamic in their positions in the networks, resulting in a community 
+  assembly punctuated by significant reorganizations of
   interactions. Non-assembling mature hedgerows and weedy field
-  margins did not undergo as many network reorganizations, suggesting
+  margins did not undergo as many significant network reorganizations, suggesting
   environmental factors alone cannot account for the changes observed
@@
   species and interaction turnover was ubiquitous across assembling
-  and non-assembling communities, though unique interactions turn over
+  and non-assembling communities, though unique interactions turnover
   at higher rates in assembling hedgerows as the networks continually
-  reorganize. Assembling networks also increased in nestedness through
-  time, though this did not make them more robust to simulated plant
+  reorganize. Nestedness of assembling networks also increased through
+  time, but there was no increase in robustness to simulated plant
   extinctions. The resilience of networks to cascading perturbations,
@@
 %% there is some evidence of that. (other studies on mutualisms?)
-The most widely explored mechanism of network assembly, preferential
-attachment \citep{barabasi1999emergence}, predicts that species
+Preferential attachment, the most widely explored mechanism of network assembly, \citep{barabasi1999emergence}, predicts that species
 entering a network are more likely to interact with species that are
@@
   klein-2007-303} --- some studies have found support for this
-mechanism of assembly. Investigating the day-to-day, temporal assembly
+assembly mechanism. Investigating the day-to-day, temporal assembly
 of a plant-pollinator network within a season, \cite{Olesen2008} found
@@
 space-for-time substitution to study primary succession along a
-glacier foreland, \cite{albrecht2010plant} found some evidence that
+glacier foreland, \cite{albrecht2010plant} also found some evidence that
 assembly occurred through preferential attachment. Network nestedness,
-a pattern of interactions where a generalist core interacts with both
+a pattern of interactions where a core formed by generalist species that interact with both
 specialist and generalist species, increased as the community aged
-\citep{albrecht2010plant}. Increasing nestedness could result from
+\citep{albrecht2010plant}. Thus, an increase in nestedness could result from
 preferential attachment process where specialist species attach to the
@@
 %% attchment, particularly vis big reorganizations of interactions
+%% Lauren I love this paragraph!
 In contrast to the ordered network build-up described by preferential
@@
 %% restoration
+%% Lauren, I would combine this paragraph with the next one, because is too short. I tried to think of stuff to add, but I coulnd't...
 Understanding network assembly is particularly relevant to ecological
@@
 poor habitats for pollinator populations \citep{kremen-2002-16816}. To
-ensure provision the continued provision of ecosystem services and
-curb biodiversity loss, effective restoration of pollinators and their
-interactions in agricultural landscapes is critical.
+ensure the continued provision of ecosystem services and
+curb biodiversity loss, it is critical to effective restore pollinators and their
+interactions in agricultural landscapes.
 
@@
 
-Using a long-term dataset of plant-pollinator communities assembling
-following hedgerow restoration in the highly simplified and
+%% Lauren I would put 10 year dataset; and I *loved* the way you organized the ideas here!! 
+We explore the process of network development using a long-term dataset of plant-pollinator communities assembling following hedgerow restoration in the highly simplified and
 intensively managed agricultural landscape of California's Central
-Valley, we explore the process of network development. We first
-determine whether the mechanism underlying network assembly is a
+Valley.  We first determine whether the mechanism underlying network assembly is a
 smooth build up of interactions as would be predicted by preferential
@@
 We surveyed plant-pollinator interaction networks of assembling
-hedgerows (N=5), as well as in two types of non-assembling communities
+hedgerows communities (N=5), as well as in two types of non-assembling communities
 to serve as controls: unrestored, weedy field margins (N=19) and
@@
 code published online by L.~Peel for the change point
-analysis. Analyses we conducted in Python 3.4.
+analysis. Analyses were conducted in Python 3.4.
 
@@
 To further elucidate the nature of the change points, we examine the
-characteristics of the species that contributed interaction
+characteristics of the species that contributed to interaction
 reorganization. Some species remain in relatively similar network
-positions through time, whereas others are more variable in their role
+positions through time, whereas others are more variable in their position
 and thus contribute more strongly to network reorganization. We test
-whether the more persistent species with the highest degree are the
+whether the more persistent species with the highest degree (number of interactions) are the
 most stable in their network positions, as would be expected if the
@@
 
-We calculate species persistence as the proportion of surveys a plant
+We calculate species persistence as the proportion of surveys in which a plant
 or pollinator is observed. Species observed consistently within and
@@
 = 0.84$, $p$-value $<$ $2*10^{-16}$), we include each explanatory
-variable in the model separately. Plant degree and persistent were not
+variable in the model separately. Plant degree and persistence were not
 significantly correlated, but we use the same models as we did for the
@@
 
+%%Lauren I really liked the way you explained this part =]
 Reorganizations of network structure can be the result of species
@@
 temporal dynamics of the assembling networks, we examined patterns of
-species and interaction turnover. Assembling networks may have higher
+species and interaction turnover. For example, assembling networks may have higher
 rates of pollinator turnover than non-assembling communities because
@@
 addition, at assembling hedgerows, plants may ``colonize'' the
-networks as they become more attractive resources.
+networks as they become more attractive resources, and establish new interactions with pollinators.
 
@@
 null models that constrained species richness to estimate the
-deviation of the observed dissimilarity from that which would be
+deviation of the observed dissimilarity from that, which would be
 expected under a random community assembly process. With the corrected
@@
 of community composition across years \citep{anderson-2011-19}. In
-order to test whether assembling hedgerows had more turnover of
-species and interactions than non-assembling communities, the species
+order to test whether assembling hedgerows had more species and interactions turnover than non-assembling communities, the species
 and interaction temporal turnover estimates were included as response
-variables in a linear mixed models with site type as an explanatory
-variable along with a random effect of site \citep{lme4, lmetest}.
+variables in a linear mixed model with site type as an explanatory
+variable along with site as a random effect \citep{lme4, lmetest}.
 
+%% Lauren I like this paragraph, but I think we could use a figure in the supplementary to help the reader understand... maybe a schematic representation? I'll try and think of something 
 Though species may turnover across years, some groups of species may
 essentially replace each other if they fill similar roles in the
-network. At non-assembling communities, species turnover may
+network, ocupying the same network position and interacting with similar species. At non-assembling communities, species turnover may
 overestimate the temporal changes in the networks if the interactions
@@
 connectance --- the proportion of observed out of possible
-interactions --- is also expected to increase as species
-preferentially attach to the core. \flagged{Is this true?} Lastly, the
-network specialization may change as the community assembles. If
+interactions --- would decrease as new, specialist species,
+preferentially attach to the core.
+\flagged{Is this true?
+- I think that the relationship between richness and connectance is not clear, i changed to decrease because it makes more sense via preferential attachment - the denominator increases because spp are being added, but the number of links does not, because they are specialists --- I checked with Paulo and he agrees} Lastly, the overall level of
+network specialization may change as the community assembles. Network-level
+specialization will increase if
 specialist species colonize the network or species begin to limit
-their interaction niche breath as the network assembles, network-level
-specialization will increase \citep{bluthgen-2006-9}.
+their interaction niche breath as the network assembles  \citep{bluthgen-2006-9}.
 
@@
 standardized $z$-scores so that nestedness and modularity metrics
-could be compared between communities. The $z$-scores were calculated
+could be compared across communities. The $z$-scores were calculated
 by generating an ensemble of $999$ randomly assembled communities,
@@
 of the estimate, $6*10^{-5} \pm 4*10^{-3}$, $p$-value=$0.987$) or
-abundance ($0.001 \pm 0.003$, $p$-value=$0.65$, Fig.~\ref{fig:rob}).
-
-In contrast, the robustness of networks to cascading perturbations, as
+abundance ($0.001 \pm 0.003$, $p$-value=$0.65$, Fig.~\ref{fig:rob}). In contrast, the robustness of networks to cascading perturbations, as
 measured by the algebraic connectivity of the network, increased as
@@
 following restoration is a highly dynamic process where interactions
-often undergo significant reorganizations. These network organizations
+often undergo significant reorganizations, the so called changing points. These network organizations
 are unlikely a product of environmental forces alone because the
@@
 In addition to finding multiple network organization changing points
-during assembly, the species that contribute most to those
-reorganizations were not what is expected by preferential
-attachment. Based on a preferential attachment-like mechanism, we
-expect that the most persistent and high degree species would remain
+during assembly, the way in which these reorganizations occur was different from what is expected by preferential attachment.
+Based on a preferential attachment mechanism \flagged{process instead of mechanism?}, we expect that the most persistent and high degree species would remain
 stable in the network core during assembly. Surprisingly, however, we
@@
   (Dialictus) incompletum}, and \textit{Toxomerus marginatus} --- were
-the only species that changed what module (i.e., community) of which
+the only species that changed which module (i.e., community)
 they were a member across years in all the assembling
@@
 the limited floral resources in the intensively managed agriculture
-landscape, and thus also most persistent. More persistent species also
+landscape, and thus also be the most persistent. More persistent species usually 
 have longer phenologies, so they can visit many different flowers,
@@
 most advantageous floral resources available, which may depend on the
-other pollinator species that are present and the state of the plant
+other pollinator species that are present, and the state of the plant
 community. Thus given the opportunity and ability to use different
@@
 
-The frequent changing points in network organization, dynamic nature
+When we look at the network-level metrics to explore the mechanism through which the community is assembling, we encounter that nestedness did increase with years
+post-restoration, as would be expected if colonizing, specialist
+species attached to a central, generalist core
+\citep{albrecht2010plant}, as predicted by preferential attachment.
+In addition, connectance decreased, as would be expected if the network is being colonized by specialist species and the overall mean number of interactions per species did not change.
+With preferential attachment, however, we
+would also expect specialization to increase, and we
+found no such trend. Furtheremore, the frequent changing points in network organization, dynamic nature
 of species positions in the networks, and turnover of species and
 interactions all point to an assembly mechanism other than
-preferential attachment. Nestedness did increase with years
-post-restoration, as would be expected if colonizing, specialist
-species attached to a central, generalist core
-\citep{albrecht2010plant}. With preferential attachment, however, we
-would also expect connectance and specialization to increase, and we
-found no such trends. \flagged{Wait if connectance increases doesn't
-  that mean specialization decreases? Which would we expect to change
-  with preferential attachment?} The stable level of network-level
+preferential attachment.  The stable level of network-level
 specialization through the assembly process may be due to the
@@
 associated with an enhanced in robustness to extinction
-\citep{Memmott2004}. Assembling hedgerows have yet to reach sufficient
+\citep{Memmott2004}. Perhaps assembling hedgerows have yet to reach sufficient
 levels of nestedness to realize the benefits nestedness
@@
   \centering
-  \includegraphics[width=.9\textwidth]{../analysis/changePoint/plotting/communitiesSkanky.pdf}
+  \includegraphics[width=.9\textwidth]{../analysis/changePoint/plotting/communitiesSkanky2.pdf}
   \caption{The community (i.e., module or compartment) membership
@@
   \includegraphics[width=.8\textwidth]{../analysis/variability/figures/cv/occ_degree.pdf}
-  \caption{The coefficient of variation of network position, as
+  \caption{The variation coefficient of network position, as
     represented by closeness, plotted against pollinator persistence
@@
   \includegraphics[width=1\textwidth]{../analysis/networkLevel/figures/baci.pdf}
-  \caption{The nestedness, plant richness and pollinator richness
+  \caption{Nestedness, plant richness and pollinator richness
     increased as the networks assembled. Specialization and modularity
@@
     perturbation increased. The robustness to species extinction is
-    measured by incrementally removing species by degree, through
+    measured by incrementally removing species by degree, though
     removing species by abundance did not yield qualitatively

### 2016-08-26 ms/ms.tex
made marilia's comments

@@
 
-\begin{abstract} %% Lauren - I would invert the phrase just so it does not beging with metwork structure - "communities' ability to maintain function in the face of species extinction is related to the structure fo networks".
-  The structure of networks is related to ability of communities to
-  maintain function in the face of species extinction. Understanding
-  network structure and how it relates to network assembly, therefore,
-  is a priority for system-level conservation biology. We explore how the
-  assembly of plant-pollinator communities occured, on native plant
-  restorations in the Central Valley of California. Species are highly
-  dynamic in their positions in the networks, resulting in a community 
-  assembly punctuated by significant reorganizations of
-  interactions. Non-assembling mature hedgerows and weedy field
-  margins did not undergo as many significant network reorganizations, suggesting
-  environmental factors alone cannot account for the changes observed
-  in assembling communities. Pollinator species were highly
-  opportunistic, and in fact the most persistent and generalized
+\begin{abstract}  
+  The ability of communities to maintain function in the face of
+  species extinction is related to the structure of
+  networks. Understanding network structure and how it relates to
+  network assembly, therefore, is a priority for system-level
+  conservation biology. Using a long-term data set comprising nearly
+  20,000 pollinator visitation records, we explore the assembly of
+  plant-pollinator communities at native plant restorations in the
+  Central Valley of California. Species are highly dynamic in their
+  positions in the networks, causing assembly to be punctuated by
+  significant reorganizations of interactions. Non-assembling mature
+  hedgerows and weedy field margins did not undergo as many
+  significant network reorganizations, suggesting environmental
+  factors alone cannot account for the changes observed in assembling
+  communities. Pollinator species were opportunistic in the flowers
+  they visited, and in fact the most persistent and generalized
   species were most variable in their network positions --- the
@@
 %% there is some evidence of that. (other studies on mutualisms?)
-Preferential attachment, the most widely explored mechanism of network assembly, \citep{barabasi1999emergence}, predicts that species
+Preferential attachment, the most widely explored mechanism of network
+assembly, \citep{barabasi1999emergence}, predicts that species
 entering a network are more likely to interact with species that are
@@
   klein-2007-303} --- some studies have found support for this
-assembly mechanism. Investigating the day-to-day, temporal assembly
-of a plant-pollinator network within a season, \cite{Olesen2008} found
+assembly mechanism. Investigating the day-to-day, temporal assembly of
+a plant-pollinator network within a season, \cite{Olesen2008} found
 that new species tended to interact with already well-connected
@@
 space-for-time substitution to study primary succession along a
-glacier foreland, \cite{albrecht2010plant} also found some evidence that
-assembly occurred through preferential attachment. Network nestedness,
-a pattern of interactions where a core formed by generalist species that interact with both
-specialist and generalist species, increased as the community aged
-\citep{albrecht2010plant}. Thus, an increase in nestedness could result from
-preferential attachment process where specialist species attach to the
-well-connected, generalist core.  % In addition,
+glacier foreland, \cite{albrecht2010plant} also found some evidence
+that assembly occurred through preferential attachment. Network
+nestedness, a pattern of interactions where a core formed by
+generalist species that interact with both specialist and generalist
+species, increased as the community aged
+\citep{albrecht2010plant}. Thus, an increase in nestedness could
+result from preferential attachment process where specialist species
+attach to the well-connected, generalist core.  % In addition,
 % non-successional temporal dynamics also suggest a stable core of
@@
 %% restoration
-%% Lauren, I would combine this paragraph with the next one, because is too short. I tried to think of stuff to add, but I coulnd't...
 Understanding network assembly is particularly relevant to ecological
@@
 the community develops. Understanding the mechanisms of network
-assembly will help to guide community restoration.
-
-%% restoration of ecosystem services is really important, particularly
-%% in ag. but ag is a bad place for pollinators
-Facilitating network restoration is especially imperative in areas
-where species interactions provide essential ecosystem services, such
-as crop pollination. In intensively managed agricultural landscapes,
-the demand for pollination services is the greatest
+assembly will help to guide community restoration. Facilitating
+network restoration is especially imperative in areas where species
+interactions provide essential ecosystem services, such as crop
+pollination. In intensively managed agricultural landscapes, the
+demand for pollination services is the greatest
 \citep{kremen-2008-10}. However, honey bees, managed extensively
@@
 poor habitats for pollinator populations \citep{kremen-2002-16816}. To
-ensure the continued provision of ecosystem services and
-curb biodiversity loss, it is critical to effective restore pollinators and their
+ensure the continued provision of ecosystem services and curb
+biodiversity loss, it is critical to restore pollinators and their
 interactions in agricultural landscapes.
@@
 
-%% Lauren I would put 10 year dataset; and I *loved* the way you organized the ideas here!! 
-We explore the process of network development using a long-term dataset of plant-pollinator communities assembling following hedgerow restoration in the highly simplified and
-intensively managed agricultural landscape of California's Central
-Valley.  We first determine whether the mechanism underlying network assembly is a
+%% Lauren I would put 10 year dataset; and I *loved* the way you
+%% organized the ideas here!!
+We explore the process of network development using a nine year
+dataset of plant-pollinator communities assembling following hedgerow
+restoration in the highly simplified and intensively managed
+agricultural landscape of California's Central Valley.  We first
+determine whether the mechanism underlying network assembly is a
 smooth build up of interactions as would be predicted by preferential
@@
 We surveyed plant-pollinator interaction networks of assembling
-hedgerows communities (N=5), as well as in two types of non-assembling communities
-to serve as controls: unrestored, weedy field margins (N=19) and
-established hedgerows (greater than 10 years since planting,
+hedgerows communities (N=5), as well as in two types of non-assembling
+communities to serve as controls: unrestored, weedy field margins
+(N=19) and mature hedgerows (greater than 10 years since planting,
 N=29). The sites were located in the Central Valley of California in
@@
 reorganization. Some species remain in relatively similar network
-positions through time, whereas others are more variable in their position
-and thus contribute more strongly to network reorganization. We test
-whether the more persistent species with the highest degree (number of interactions) are the
+positions through time, whereas others are more variable in their
+position and thus contribute more strongly to network
+reorganization. We test whether the more persistent species with the
+highest degree (number of different interaction parteners) are the
 most stable in their network positions, as would be expected if the
@@
 
-We calculate species persistence as the proportion of surveys in which a plant
-or pollinator is observed. Species observed consistently within and
-between years are thus maximally persistent. Weighted species degree
-is calculated from interaction observations from an extensive dataset
-from Yolo County (approx.~18000 interaction records) that included
-both the data included in this study and additional data from sites
-where we collected flower visitors using the same methods
+We calculate species persistence as the proportion of surveys in which
+a plant or pollinator is observed. Species observed consistently
+within and between years are thus maximally persistent. Weighted
+species degree is calculated from interaction observations from an
+extensive dataset from Yolo County (approx.~18000 interaction records)
+that included both the data included in this study and additional data
+from sites where we collected flower visitors using the same methods
 \citep{mgonigle-2015-x, ponisio2015farm}. To represent network
@@
 = 0.84$, $p$-value $<$ $2*10^{-16}$), we include each explanatory
-variable in the model separately. Plant degree and persistence were not
-significantly correlated, but we use the same models as we did for the
-pollinators for consistency. A linear increase in closeness, as might
-be expected with assembly by preferential attachment, would also lead
-to a high variability in closeness scores, so we test whether
+variable in the model separately. Plant degree and persistence were
+not significantly correlated, but we use the same models as we did for
+the pollinators for consistency. A linear increase in closeness, as
+might be expected with assembly by preferential attachment, would also
+lead to a high variability in closeness scores, so we test whether
 closeness increases through time.
@@
 temporal dynamics of the assembling networks, we examined patterns of
-species and interaction turnover. For example, assembling networks may have higher
-rates of pollinator turnover than non-assembling communities because
-new pollinator species are colonizing and establishing themselves
-\citep{mgonigle-2015-x}. Similarly, if pollinators try to maximize
-their foraging efficiency based on the species present, interactions
-may turnover more quickly than in established communities. In
-addition, at assembling hedgerows, plants may ``colonize'' the
-networks as they become more attractive resources, and establish new interactions with pollinators.
+species and interaction turnover. For example, assembling networks may
+have higher rates of pollinator turnover than non-assembling
+communities because new pollinator species are colonizing and
+establishing themselves \citep{mgonigle-2015-x}. Similarly, if
+pollinators try to maximize their foraging efficiency based on the
+species present, interactions may turnover more quickly than in
+established communities. In addition, at assembling hedgerows, plants
+may ``colonize'' the networks as they become more attractive
+resources, and establish new interactions with pollinators.
 
@@
 of community composition across years \citep{anderson-2011-19}. In
-order to test whether assembling hedgerows had more species and interactions turnover than non-assembling communities, the species
-and interaction temporal turnover estimates were included as response
+order to test whether assembling hedgerows had more species and
+interactions turnover than non-assembling communities, the species and
+interaction temporal turnover estimates were included as response
 variables in a linear mixed model with site type as an explanatory
@@
 
-%% Lauren I like this paragraph, but I think we could use a figure in the supplementary to help the reader understand... maybe a schematic representation? I'll try and think of something 
+%% Lauren I like this paragraph, but I think we could use a figure in
+%% the supplementary to help the reader understand... maybe a
+%% schematic representation? I'll try and think of something
 Though species may turnover across years, some groups of species may
 essentially replace each other if they fill similar roles in the
-network, ocupying the same network position and interacting with similar species. At non-assembling communities, species turnover may
+network, ocupying the same network position and interacting with
+similar species. At non-assembling communities, species turnover may
 overestimate the temporal changes in the networks if the interactions
@@
 interactions --- would decrease as new, specialist species,
-preferentially attach to the core.
-\flagged{Is this true?
-- I think that the relationship between richness and connectance is not clear, i changed to decrease because it makes more sense via preferential attachment - the denominator increases because spp are being added, but the number of links does not, because they are specialists --- I checked with Paulo and he agrees} Lastly, the overall level of
-network specialization may change as the community assembles. Network-level
-specialization will increase if
-specialist species colonize the network or species begin to limit
-their interaction niche breath as the network assembles  \citep{bluthgen-2006-9}.
+preferentially attach to the core.  Lastly, the overall level of
+network specialization may change as the community
+assembles. Network-level specialization will increase if specialist
+species colonize the network or species begin to limit their
+interaction niche breath as the network assembles
+\citep{bluthgen-2006-9}.
 
@@
 of the estimate, $6*10^{-5} \pm 4*10^{-3}$, $p$-value=$0.987$) or
-abundance ($0.001 \pm 0.003$, $p$-value=$0.65$, Fig.~\ref{fig:rob}). In contrast, the robustness of networks to cascading perturbations, as
-measured by the algebraic connectivity of the network, increased as
-the network assembled (estimate of the slope of robustness to
-cascading perturbations through time $\pm$ standard error of the
-estimate, $0.6814 \pm 0.272$, $p$-value=$0.042$, Fig.~\ref{fig:rob}).
+abundance ($0.001 \pm 0.003$, $p$-value=$0.65$,
+Fig.~\ref{fig:rob}). In contrast, the robustness of networks to
+cascading perturbations, as measured by the algebraic connectivity of
+the network, increased as the network assembled (estimate of the slope
+of robustness to cascading perturbations through time $\pm$ standard
+error of the estimate, $0.6814 \pm 0.272$, $p$-value=$0.042$,
+Fig.~\ref{fig:rob}).
 
@@
 following restoration is a highly dynamic process where interactions
-often undergo significant reorganizations, the so called changing points. These network organizations
-are unlikely a product of environmental forces alone because the
-network changing points in non-assembling communities are less
-frequent, and few consistent trends in when change points occurred
-across all sites. Several sites had network changing points between
-years 2009 and 2011 (Fig.~\ref{fig:changePoints}). In California, 2011
-marked the beginning of a multi-year drought. The assembling hedgerows
-were not sampled in 2010, so disentangling whether the changing points
-are due to skipping a year of assembly or the drought is not
+often undergo significant reorganizations, the so called changing
+points. These network reorganizations are unlikely a product of
+environmental forces alone because the network changing points in
+non-assembling communities are less frequent, and few consistent
+trends in when change points occurred across all sites. Several sites
+had network changing points between years 2009 and 2011
+(Fig.~\ref{fig:changePoints}). In California, 2011 marked the
+beginning of a multi-year drought. The assembling hedgerows were not
+sampled in 2010, so disentangling whether the changing points are due
+to skipping a year of assembly or the drought is not
 possible. Interestingly, most assembling hedgerows did not undergo a
@@
 In addition to finding multiple network organization changing points
-during assembly, the way in which these reorganizations occur was different from what is expected by preferential attachment.
-Based on a preferential attachment mechanism \flagged{process instead of mechanism?}, we expect that the most persistent and high degree species would remain
-stable in the network core during assembly. Surprisingly, however, we
-encountered the opposite pattern. For example, the five most
-ubiquitous species in our study landscape --- \textit{Halictus
-  ligatus}, \textit{Halictus tripartitus}, \textit{Lasioglossum
-  (Dialictus) incompletum}, and \textit{Toxomerus marginatus} --- were
-the only species that changed which module (i.e., community)
-they were a member across years in all the assembling
-hedgerows. Because species degree and persistence were strongly
-correlated, it is difficult to disentangle the causal mechanism for
-why species with those characteristics are so variable in their
-network position. Generalized species may be able to better exploit
-the limited floral resources in the intensively managed agriculture
-landscape, and thus also be the most persistent. More persistent species usually 
-have longer phenologies, so they can visit many different flowers,
-resulting in a higher degree. Either way, our result suggests that
-adaptable species can change their network position to utilize the
-most advantageous floral resources available, which may depend on the
-other pollinator species that are present, and the state of the plant
-community. Thus given the opportunity and ability to use different
-resources, species will often change their network positions.
+during assembly, the way in which these reorganizations occur was
+different from what is expected by preferential attachment. Based on a
+preferential attachment process, we expect that the most persistent
+and high degree species would remain stable in the network core during
+assembly. Surprisingly, however, we encountered the opposite
+pattern. For example, the five most ubiquitous species in our study
+landscape --- \textit{Halictus ligatus}, \textit{Halictus
+  tripartitus}, \textit{Lasioglossum (Dialictus) incompletum}, and
+\textit{Toxomerus marginatus} --- were the only species that changed
+which module (i.e., community) they were a member across years in all
+the assembling hedgerows. Because species degree and persistence were
+strongly correlated, it is difficult to disentangle the causal
+mechanism for why species with those characteristics are so variable
+in their network position. Generalized species may be able to better
+exploit the limited floral resources in the intensively managed
+agriculture landscape, and thus also be the most persistent. More
+persistent species usually have longer phenologies, so they can visit
+many different flowers, resulting in a higher degree. Either way, our
+result suggests that adaptable species can change their network
+position to utilize the most advantageous floral resources available,
+which may depend on the other pollinator species that are present, and
+the state of the plant community. Thus given the opportunity and
+ability to use different resources, species will often change their
+network positions.
 
@@
 
-When we look at the network-level metrics to explore the mechanism through which the community is assembling, we encounter that nestedness did increase with years
-post-restoration, as would be expected if colonizing, specialist
-species attached to a central, generalist core
-\citep{albrecht2010plant}, as predicted by preferential attachment.
-In addition, connectance decreased, as would be expected if the network is being colonized by specialist species and the overall mean number of interactions per species did not change.
-With preferential attachment, however, we
-would also expect specialization to increase, and we
-found no such trend. Furtheremore, the frequent changing points in network organization, dynamic nature
-of species positions in the networks, and turnover of species and
-interactions all point to an assembly mechanism other than
-preferential attachment.  The stable level of network-level
-specialization through the assembly process may be due to the
-increased colonization of specialized species \citep{mgonigle-2015-x}
-accompanied by an increase in the diet breath of resident
-species. This would be expected if resident species were able to
-minimize their foraging time by expanding their diet breath as plant
-diversity increases with hedgerow maturation \citep{Waser1996,
+When we explore the how the network-level interaction patterns changed
+through time, we found that nestedness did increase as the community
+assembled, as would be expected if colonizing, specialist species
+preferentially attached to a central, generalist core
+\citep{albrecht2010plant}. In addition, connectance decreased, as
+would be expected if the network is being colonized by specialist
+species and the overall mean number of interactions per species did
+not change. With preferential attachment, however, we would also
+expect specialization to increase, and we found no such
+trend. Furtheremore, the frequent changing points in network
+organization, dynamic nature of species positions in the networks, and
+turnover of species and interactions all point to an assembly
+mechanism other than preferential attachment.  The stable level of
+network-level specialization through the assembly process may be due
+to the increased colonization of specialized species
+\citep{mgonigle-2015-x} accompanied by an increase in the diet breath
+of resident species. This would be expected if resident species were
+able to minimize their foraging time by expanding their diet breath as
+plant diversity increases with hedgerow maturation \citep{Waser1996,
   pyke1984optimal, Bluthgen2007, albrecht2010plant}. Such a change in
 pollinator behavior would also explain the increase in network
-nestedness.
+nestedness. Because so many mechanisms give rise to the same patterns
+of interaction, additional tests are necessary assess the contribution
+of different mechanisms to community assembly.
 

### 2016-08-26 analysis/variability/src/phyloIntBeta.R

@@ -38,7 +38,7 @@ calcCommDis <- function(spec.dat, type, lc, abund.w=TRUE){
   node.order <- lapply(comm, function(x){
     node.num[,1][match(colnames(x), node.num[,2])]
   })
-  assignCols <- function(a,b){
+  assignCols <- function(a, b){
     colnames(a) <- b
     return(a)
   }

### 2016-08-29 ms/ms.tex
added conceptual figure to supplement

@@
 
-%% Lauren I like this paragraph, but I think we could use a figure in
-%% the supplementary to help the reader understand... maybe a
-%% schematic representation? I'll try and think of something
 Though species may turnover across years, some groups of species may
@@
 occurring in one year are similar to those in the next year when they
-are weighted by the similarity of their constituent species. We
-develop a method examine the temporal turnover of interactions
-weighted based on their similarity. We followed \cite{ahn2010link}
-algorithm to hierarchically cluster all the interactions (edges)
-across sites and years based on their similarity, and build a
-dendrogram. The interaction similarity is based how may plants and
-pollinators (vertices) two edges share \citep{ahn2010link,
+are weighted by the similarity of their constituent species
+(Fig.~S1). We develop a method examine the temporal turnover of
+interactions weighted based on their similarity. We followed
+\cite{ahn2010link} algorithm to hierarchically cluster all the
+interactions (edges) across sites and years based on their similarity,
+and build a dendrogram. The interaction similarity is based how may
+plants and pollinators (vertices) two edges share \citep{ahn2010link,
   kalinka2011linkcomm}. The more species edges shared in common, the
@@
 
-We would like to thank Paulo Guimar{\~a}es and Aaron Clauset for their
-invaluable discussions, and Leto Peel for help with the change point
-analysis.  We thank the growers and land owners that allowed us to
-work on their property. We also greatly appreciate the identification
-assistance of expert taxonomists Martin Hauser, Robbin Thorp and Jason
-Gibbs.  This work was supported by funding from the Army Research
-Office (W911NF-11-1-0361 to CK), the Natural Resources Conservation
-Service (CIG-69-3A75-12-253, CIG-69-3A75-9-142, CIG-68-9104-6-101 and
-WLF-69-7482-6-277 to The Xerces Society), the National Science
-Foundation (DEB-0919128 to CK), The U.S.  Department of Agriculture
-(USDA-NIFA 2012-51181-20105 to Michigan State University).  Funding
-for LCP was provided by an NSF Graduate Research Fellowship, the USDA
-NIFA Graduate Fellowship, and the Berkeley Institute for Data
-Science. Funding for MPG was provided by S{\~a}o Paulo Research
-Foundation (FAPESP, grant 2013/13319-5). We also appreciate the Santa
-Fe Institute for faciliting this international collaboration.
+We would like to thank Paulo Guimar{\~a}es, Aaron Clauset and Matthew
+Hutchinson for their invaluable discussions and comments, and Leto
+Peel for help with the change point analysis.  We thank the growers
+and land owners that allowed us to work on their property. We also
+greatly appreciate the identification assistance of expert taxonomists
+Martin Hauser, Robbin Thorp and Jason Gibbs.  This work was supported
+by funding from the Army Research Office (W911NF-11-1-0361 to CK), the
+Natural Resources Conservation Service (CIG-69-3A75-12-253,
+CIG-69-3A75-9-142, CIG-68-9104-6-101 and WLF-69-7482-6-277 to The
+Xerces Society), the National Science Foundation (DEB-0919128 to CK),
+The U.S.  Department of Agriculture (USDA-NIFA 2012-51181-20105 to
+Michigan State University).  Funding for LCP was provided by an NSF
+Graduate Research Fellowship, the USDA NIFA Graduate Fellowship, and
+the Berkeley Institute for Data Science. Funding for MPG was provided
+by S{\~a}o Paulo Research Foundation (FAPESP, grant 2013/13319-5). We
+also appreciate the Santa Fe Institute for faciliting this
+international collaboration.
 

### 2016-08-30 analysis/variability/plotting/src/plotPanels.R
made matt's comments

@@ -4,7 +4,7 @@ plot.panels <- function(){
     col.fill <- add.alpha(col.lines, alpha=0.2)
     names(col.lines) <- names(col.fill) <- "all"
     layout(matrix(1:4, nrow=2, byrow=TRUE))
-    par(oma=c(2, 5, 0.5, 1),
+    par(oma=c(2, 6.5, 0.5, 1),
         mar=c(2.5, 0, 0, 1), cex.axis=1.2)
     ## persistence pollinators
     plot.panel(new.dd=occ.pi.close,
@@ -17,6 +17,8 @@ plot.panels <- function(){
                plot.y=TRUE,
                plot.x=TRUE,
                treatments="all")
+    mtext("Pollinators", 2, line=5, cex=1.2)
+    
     ## degree pollinators
     plot.panel(new.dd=degree.pi,
                dats=degree.closeness.cv$data,
@@ -41,9 +43,12 @@ plot.panels <- function(){
                plot.y=TRUE,
                treatments="all")
     
+    mtext("Plants", 2, line=5, cex=1.2)
+    
     mtext("Persistence", 1, line=3, cex=1.2)
     mtext("Closeness variability (log)", 2,
           line=3.2, cex=1.2, adj=-0.75)
+
 
     ## degree plants
     plot.panel(new.dd=plants.degree.pi,

### 2016-08-30 ms/ms.tex
random stuff'

@@
   Central Valley of California. Species are highly dynamic in their
-  positions in the networks, causing assembly to be punctuated by
+  network position, causing assembly to be punctuated by
   significant reorganizations of interactions. Non-assembling mature
@@
   communities. Pollinator species were opportunistic in the flowers
-  they visited, and in fact the most persistent and generalized
-  species were most variable in their network positions --- the
-  opposite of what would be expected by preferential attachment. High
+  they visited. Indeed, the most persistent and generalized
+  species were also the most variable in their network positions --- the
+  opposite of which would be expected by preferential attachment. High
   species and interaction turnover was ubiquitous across assembling
@@
 
-Keywords: changing points, temporal networks, hedgerows, species
-interactions, preferential attachment, mutualisms
+\keywords
 
@@
   barnosky2011has}. Because mutualistic interactions are essential for
-maintaining the diversity their component guilds of species, these
-systems are particularly at risk from coextinction cascades. The
-nature of these cascades will depend on the interaction patterns
-within a community \citep{Memmott2004, Rezende2007,
-  Bascompte2009}. Recovering the lost biodiversity and interactions
-through ecological restoration has become increasingly imperative, and
-a key restoration aim is to facilitate assembly of robust interaction
-networks \citep{menz-2010-4}. We know little, however, about how to
-re-assemble interacting communities through restoration, or the
-process of ecological network assembly more generally.
+maintaining the diversity their component guilds, these systems are
+particularly at risk from coextinction cascades. The nature of these
+cascades will depend on the interaction patterns within a community
+\citep{Memmott2004, Rezende2007, Bascompte2009, Thebault2010}. To
+safeguard function it has become increasingly imperative to aid the
+recovery of lost biodiversity and interactions through ecological
+restoration, and a key restoration aim is to facilitate assembly of
+robust interaction networks \citep{menz-2010-4}. We know little,
+however, about how to re-assemble interacting communities through
+restoration, or the process of ecological network assembly more
+generally.
 
@@
 species, potentially because these species are either more abundant or
-more temporally persistent \citep{Olesen2008}. In addition, using a
-space-for-time substitution to study primary succession along a
-glacier foreland, \cite{albrecht2010plant} also found some evidence
-that assembly occurred through preferential attachment. Network
-nestedness, a pattern of interactions where a core formed by
-generalist species that interact with both specialist and generalist
-species, increased as the community aged
-\citep{albrecht2010plant}. Thus, an increase in nestedness could
-result from preferential attachment process where specialist species
-attach to the well-connected, generalist core.  % In addition,
+more temporally persistent. In addition, using a space-for-time
+substitution to study primary succession along a glacier foreland,
+\cite{albrecht2010plant} also found some evidence that assembly
+occurred through preferential attachment. Network nestedness, a
+pattern of interactions where a core formed by generalist species that
+interact with both specialist and generalist species, increased as the
+community aged \citep{albrecht2010plant}. Thus, an increase in
+nestedness could result from preferential attachment process where
+specialist species attach to the well-connected, generalist
+core. % In addition,
 % non-successional temporal dynamics also suggest a stable core of
@@
 biodiversity loss, it is critical to restore pollinators and their
-interactions in agricultural landscapes.
-
-%% hedgerows!!  
-To promote pollinator services in agriculture, farmers may chose to
-plant strips of native plants along farm edges (hedgerows) to help
-provide habitat for pollinators without removing arable land from
-production. Hedgerows augment the richness, abundance and trait
-diversity of pollinators in agricultural landscapes
-\citep{morandin-2013-829, mgonigle-2015-x, kremen-2015-602,
-  ponisio2015farm}. In addition, hedgerows promote the persistence and
-colonization of floral resource specialists
+interactions in agricultural landscapes. To promote pollinator
+services in agriculture, farmers may chose to plant strips of native
+plants along farm edges (hedgerows) to help provide habitat for
+pollinators without removing arable land from production. Hedgerows
+augment the richness, abundance and trait diversity of pollinators in
+agricultural landscapes \citep{morandin-2013-829, mgonigle-2015-x,
+  kremen-2015-602, ponisio2015farm}. In addition, hedgerows promote
+the persistence and colonization of floral resource specialists
 \citep{mgonigle-2015-x}. Little is known, however, about how new
@@
 We explore the process of network development using a nine year
-dataset of plant-pollinator communities assembling following hedgerow
+dataset of plant-pollinator communities assembly following hedgerow
 restoration in the highly simplified and intensively managed
@@
 a stable, well-connected core of species persist. We test whether the
-species are most variable in their network position --- and thus
-important contributors network reorganizations --- are less persistent
+species that are most variable in their network position --- and thus
+important contributors to network reorganizations --- are less persistent
 and connected species. To further explore the mechanisms underlying
-the temporal dynamics in the networks, we examine patterns in the
+the temporal dynamics of the networks, we examine patterns in the
 species and interaction temporal turnover. Lastly, we investigate
@@
 reorganization. We test whether the more persistent species with the
-highest degree (number of different interaction parteners) are the
+highest degree (number of different interaction partners) are the
 most stable in their network positions, as would be expected if the
@@
 first calculated the pairwise dissimilarity of plants, pollinators and
-interactions between years within each site using a dissimilarity
-estimator that incorporates abundances, while also accounting for
-unobserved records \citep{chao-2005-148}. Dissimilarity estimates can
-be affected by the total number of species and individuals sampled at
-a site \citep[e.g.,][]{ponisio2015farm}. For example, the probability
-that two sites do not share any species is higher when there are few
-individuals at those sites. Following \cite{ponisio2015farm}, we use
-null models that constrained species richness to estimate the
-deviation of the observed dissimilarity from that, which would be
-expected under a random community assembly process. With the corrected
-dissimilarity values, we then calculated the multivariate dispersion
-of community composition across years \citep{anderson-2011-19}. In
-order to test whether assembling hedgerows had more species and
-interactions turnover than non-assembling communities, the species and
-interaction temporal turnover estimates were included as response
-variables in a linear mixed model with site type as an explanatory
-variable along with site as a random effect \citep{lme4, lmetest}.
+interactions between years within each site using the Chao
+dissimilarity estimator that incorporates abundances, while also
+accounting for unobserved records \citep{chao-2005-148}. Dissimilarity
+estimates can be affected by the total number of species and
+individuals sampled at a site \citep[e.g.,][]{ponisio2015farm}. For
+example, the probability that two sites do not share any species is
+higher when there are few individuals at those sites. Following
+\cite{ponisio2015farm}, we use null models that constrained species
+richness to estimate the deviation of the observed dissimilarity from
+that, which would be expected under a random community assembly
+process. With the corrected dissimilarity values, we then calculated
+the multivariate dispersion of community composition across years
+\citep{anderson-2011-19}. In order to test whether assembling
+hedgerows had more species and interactions turnover than
+non-assembling communities, the species and interaction temporal
+turnover estimates were included as response variables in a linear
+mixed model with site type as an explanatory variable along with site
+as a random effect \citep{lme4, lmetest}.
 
@@
 are weighted by the similarity of their constituent species
-(Fig.~S1). We develop a method examine the temporal turnover of
-interactions weighted based on their similarity. We followed
+(Fig.~S1). We develop a method to examine the temporal turnover of
+interactions with weightings based on their similarity. We followed
 \cite{ahn2010link} algorithm to hierarchically cluster all the
@@
 The majority ($76\%$) of the sites underwent at least one significant
-interaction reorganization (Fig.~\ref{fig:changePoints}).  All five of
-the assembling hedgerows experienced network changing points, whereas
-only $40\%$ and $81\%$ of non-assembling hedgerows and field margins,
-respectively, underwent significant interaction
-reorganizations. Assembling hedgerows had significantly more changing
-points than the non-assembling networks (estimate of the difference in
-the odds ratios between assembling and non-assembling networks,
-$3.316$, $95\%$ CI $[1.314, 8.572]$, $p$-value= $0.0117$). Network
-assembly following restoration is thus punctuated by more interaction
-reorganizations than would be expected by environmental shifts alone
-that would effect assembling and non-assembling networks equally.
+interaction reorganization (Fig.~\ref{fig:changePoints},
+\ref{fig:changePoints2}).  All five of the assembling hedgerows
+experienced network changing points, whereas only $40\%$ and $81\%$ of
+non-assembling hedgerows and field margins, respectively, underwent
+significant interaction reorganizations. Assembling hedgerows had
+significantly more changing points than the non-assembling networks
+(estimate of the difference in the odds ratios between assembling and
+non-assembling networks, $3.316$, $95\%$ CI $[1.314, 8.572]$,
+$p$-value= $0.0117$). Network assembly following restoration is thus
+punctuated by more interaction reorganizations than would be expected
+by environmental shifts alone that would effect assembling and
+non-assembling networks equally.
 
@@
 attachment, both pollinator persistence and degree were positively
-related to network position variability (estimate of the slope of
-closeness variability and persistence $\pm$ standard error of the
-estimate, $0.653 \pm 0.225$, $p$-value=$0.009$; slope of closeness
-variability and degree, $0.008 \pm 0.002$, $p$-value=$0.002$). In
-addition, plant persistence and degree were not significantly related
-to network position variability (estimate of the slope of closeness
-variability and persistence $\pm$ standard error of the estimate,
-$-2.063 \pm 3.091$, $p$-value=$0.5$; slope of closeness variability
-and degree, $0.0018 \pm 0.002$, $p$-value=$0.3$). The variability of
-species network position was not the result of closeness linearly
-increasing through time, and, in fact, plant and pollinator closeness
-decreased slightly through time (Fig.~S1, estimate of the slope of
-closeness through time $\pm$ standard error of the estimate,
-pollinators: $-0.0003 \pm 0.00005$, $p$-value=$2.7*10^{-12}$; plants
-$-0.007 \pm 0.001$, $p$-value=$1.4*10^{-6}$). Through statistically
-significant, the slopes are so slight they may not me biologically
-significant.
+related to network position variability (Fig.~\ref{fig:cv}, estimate
+of the slope of closeness variability and persistence $\pm$ standard
+error of the estimate, $0.653 \pm 0.225$, $p$-value=$0.009$; slope of
+closeness variability and degree, $0.008 \pm 0.002$,
+$p$-value=$0.002$). In addition, plant persistence and degree were not
+significantly related to network position variability
+(Fig.~\ref{fig:cv}, estimate of the slope of closeness variability and
+persistence $\pm$ SE, $-2.063 \pm 3.091$,
+$p$-value=$0.5$; slope of closeness variability and degree, $0.0018
+\pm 0.002$, $p$-value=$0.3$). The variability of species network
+position was not the result of closeness linearly increasing through
+time, and, in fact, plant and pollinator closeness decreased slightly
+through time (Fig.~S2, estimate of the slope of closeness through time
+$\pm$ SE, pollinators: $-0.0003 \pm
+0.00005$, $p$-value=$2.7*10^{-12}$; plants $-0.007 \pm 0.001$,
+$p$-value=$1.4*10^{-6}$). Through statistically significant, the
+slopes are so slight they may not be biologically significant.
 
@@
 field margins, though mature hedgerows has marginally significantly
-less pollinator turnover than field margins (estimate $\pm$ standard
-error of the estimate of the difference in turnover between field
-margins and mature hedgerows, $-0.0498 \pm 0.026$,
-$p$-value=$0.058$). When interactions where weighted by their
+less pollinator turnover than field margins (Fig.~\ref{fig:beta},
+estimate $\pm$ SE of the difference in
+turnover between field margins and mature hedgerows, $-0.0498 \pm
+0.026$, $p$-value=$0.058$). When interactions where weighted by their
 similarity, both assembling and mature hedgerows had higher rates of
-turnover than field margins (estimate $\pm$ standard error of the
-estimate of the difference in turnover between field margins and
-assembling hedgerows, $0.115 \pm 0.027$, $p$-value=$0.0002$; field
-margins and mature hedgerows, $0.082 \pm 0.024$,
-$p$-value=$0.002$). The weighted interaction turnover at assembling
-hedgerows, however, was not significantly higher than in
+turnover than field margins (Fig.~\ref{fig:beta}, estimate $\pm$
+SE of the difference in turnover between
+field margins and assembling hedgerows, $0.115 \pm 0.027$,
+$p$-value=$0.0002$; field margins and mature hedgerows, $0.082 \pm
+0.024$, $p$-value=$0.002$). The weighted interaction turnover at
+assembling hedgerows, however, was not significantly higher than in
 non-assembling, mature hedgerows.
@@
 (Fig.~\ref{fig:baci}, estimate of the slope of nestedness through time
-$\pm$ standard error of the estimate, $1.834 \pm 0.6142$,
+$\pm$ SE, $1.834 \pm 0.6142$,
 $p$-value=$0.022$). Modularity decreased (Fig.~\ref{fig:baci}),
@@
 specialization remained constant (estimate of the slope of
-specialization through time $\pm$ standard error of the estimate,
+specialization through time $\pm$ SE,
 $0.003$ $\pm$ $0.015$, $p$-value=$0.827$).
@@
 (Fig.~\ref{fig:baci}, estimate of the slope of richness through time
-$\pm$ standard error of the estimate, pollinators: $0.193$ $\pm$
+$\pm$ SE, pollinators: $0.193$ $\pm$
 $0.0729$, $p$-value=$0.008$; plants: $0.212$ $\pm$ $0.0653$,
 $p$-value=$0.001$). Unsurprisingly, pollinator species are colonizing
-and persisting at the assembling hedgerows. Though some new plant
-species may establish themselves in the hedgerows, because the plant
-species richness is based on the flowers in the network and not the
-presence of a plant at a site, the increase in plant richness is
-likely due to plants attracting visitors as they offer better rewards.
+and persisting at the assembling hedgerows. Plant species richness is
+based on the flowers bbeing visiting by pollinators and not the
+presence of a particular plant species at a site. Thus, though some
+new plant species may establish themselves in the hedgerows, the
+increase in plant richness in the networks is likely due to plants
+attracting visitors, as they offer better rewards.
 
@@
 extinction when species where removed incrementally by degree
-(estimate of the slope of robustness through time $\pm$ standard error
-of the estimate, $6*10^{-5} \pm 4*10^{-3}$, $p$-value=$0.987$) or
-abundance ($0.001 \pm 0.003$, $p$-value=$0.65$,
-Fig.~\ref{fig:rob}). In contrast, the robustness of networks to
-cascading perturbations, as measured by the algebraic connectivity of
-the network, increased as the network assembled (estimate of the slope
-of robustness to cascading perturbations through time $\pm$ standard
-error of the estimate, $0.6814 \pm 0.272$, $p$-value=$0.042$,
-Fig.~\ref{fig:rob}).
+(Fig.~\ref{fig:rob}, estimate of the slope of robustness through time
+$\pm$ SE, $6*10^{-5} \pm 4*10^{-3}$, $p$-value=$0.987$) or abundance
+($0.001 \pm 0.003$, $p$-value=$0.65$). In contrast, the robustness of
+networks to cascading perturbations, as measured by the algebraic
+connectivity of the network, increased as the network assembled
+(Fig.~\ref{fig:rob}, estimate of the slope of robustness to cascading
+perturbations through time $\pm$ SE, $0.6814 \pm 0.272$,
+$p$-value=$0.042$).
 
@@
 often undergo significant reorganizations, the so called changing
-points. These network reorganizations are unlikely a product of
-environmental forces alone because the network changing points in
-non-assembling communities are less frequent, and few consistent
-trends in when change points occurred across all sites. Several sites
-had network changing points between years 2009 and 2011
-(Fig.~\ref{fig:changePoints}). In California, 2011 marked the
-beginning of a multi-year drought. The assembling hedgerows were not
-sampled in 2010, so disentangling whether the changing points are due
-to skipping a year of assembly or the drought is not
-possible. Interestingly, most assembling hedgerows did not undergo a
-significant interaction reorganization immediately after a hedgerow
-was planted (i.e., the transition from weedy field margin to
-hedgerow). This result is consistent with the finding that in our
-study system, hedgerow restoration takes several years to have an
-impact on the plant-pollinator communities (Kremen and M'Gonigle, in
-prep).
+points. If these network reorganizations were a product of
+environmental forces alone, we would explect to observe the same
+changing points at the same periods, consistently across all
+sites. However, network changing points in non-assembling communities
+are less frequent, and there are few consistent trends in when change
+points occurred across all sites. Several sites had network changing
+points between years 2009 and 2011 (Fig.~\ref{fig:changePoints}). In
+California, 2011 marked the beginning of a multi-year drought. The
+assembling hedgerows were not sampled in 2010, so disentangling
+whether the changing points are due to skipping a year of assembly or
+the drought is not possible. Interestingly, most assembling hedgerows
+did not undergo a significant interaction reorganization immediately
+after a hedgerow was planted (i.e., the transition from weedy field
+margin to hedgerow). This result is consistent with the finding that
+in our study system, hedgerow restoration takes several years to have
+an impact on the plant-pollinator communities (Kremen and M'Gonigle,
+in prep).
 
@@
 during assembly, the way in which these reorganizations occur was
-different from what is expected by preferential attachment. Based on a
+different from what is expected by preferential attachment Based on a
 preferential attachment process, we expect that the most persistent
@@
 assembly. Surprisingly, however, we encountered the opposite
-pattern. For example, the five most ubiquitous species in our study
+pattern. For example, the four most ubiquitous species in our study
 landscape --- \textit{Halictus ligatus}, \textit{Halictus
@@
 expect specialization to increase, and we found no such
-trend. Furtheremore, the frequent changing points in network
+trend. Furthermore, the frequent changing points in network
 organization, dynamic nature of species positions in the networks, and
@@
 nestedness. Because so many mechanisms give rise to the same patterns
-of interaction, additional tests are necessary assess the contribution
+of interaction, additional tests are necessary to assess the contribution
 of different mechanisms to community assembly.
@@
 literature, we have few mechanistic models of mutualistic network
-assembly \citep{valdovinos2013adaptive}. In addition, the few
-developed often borrow their mechanisms from competitive interactions,
-leading to inaccurate biological assumptions
+assembly \citep{valdovinos2013adaptive, Nuismer2013, Guimaraes2011}.
+In addition, the few developed often borrow their mechanisms from
+competitive interactions, leading to inaccurate biological assumptions
 \citep{holland2006comment}. We need further development of mechanistic
-models of mutualistic system to generate testable predictions, along
+models of mutualistic systems to generate testable predictions, along
 with empirical exploration of network assembly. Plant-pollinator
@@
   \centering
-  \includegraphics[width=.8\textwidth]{../analysis/variability/figures/turnover_panels.pdf}
+  \includegraphics[width=1\textwidth]{../analysis/variability/figures/turnover_panels.pdf}
   \caption{The species, interaction and weighted interactions turnover

### 2016-10-04 ms/ms.tex
made pailo's comments

@@
   plant-pollinator communities at native plant restorations in the
-  Central Valley of California. Species are highly dynamic in their
-  network position, causing assembly to be punctuated by
-  significant reorganizations of interactions. Non-assembling mature
-  hedgerows and weedy field margins did not undergo as many
-  significant network reorganizations, suggesting environmental
-  factors alone cannot account for the changes observed in assembling
-  communities. Pollinator species were opportunistic in the flowers
-  they visited. Indeed, the most persistent and generalized
-  species were also the most variable in their network positions --- the
-  opposite of which would be expected by preferential attachment. High
+  Central Valley of California. Across years, species are highly
+  dynamic in their network position, causing assembly to be punctuated
+  by significant reorganizations of interactions. In contrast, the
+  non-assembling networks did not restructure as frequently. Across
+  all communities, Pollinator species were opportunistic in the
+  flowers they visited. Indeed, the most persistent and generalized
+  species were also the most variable in their network positions. High
   species and interaction turnover was ubiquitous across assembling
@@
   time, but there was no increase in robustness to simulated plant
-  extinctions. The resilience of networks to cascading perturbations,
-  however, increased as the communities assembled, at least partially
-  due to accumulating species richness. Our results elucidate some of
-  the mechanisms underlying plant-pollinator network assembly and
-  restoration, while providing further evidence that hedgerows are a
-  valuable tool for promoting species conservation and ecosystem
-  provision in agricultural areas.
+  extinctions. Our results elucidate some of the mechanisms underlying
+  plant-pollinator network assembly and restoration, while providing
+  further evidence that hedgerows are a valuable tool for promoting
+  species conservation and ecosystem provision in agricultural areas.
 \end{abstract}
@@
 networks \citep{forup-2008-742, forup2008restoration,
-  devoto2012understanding}, suggesting interactions are evolving as
+  devoto2012understanding}, suggesting interactions are changing as
 the community develops. Understanding the mechanisms of network
@@
 agricultural landscape of California's Central Valley.  We first
-determine whether the mechanism underlying network assembly is a
-smooth build up of interactions as would be predicted by preferential
-attachment, or punctuated by significant reorganizations of
-interactions (i.e., network changing points). Even with changing
-points in interaction organization, networks could still be assembling
-via preferential attachment if the network reorganizations were
-primarily driven the by peripheral, temporally variable species while
-a stable, well-connected core of species persist. We test whether the
-species that are most variable in their network position --- and thus
-important contributors to network reorganizations --- are less persistent
-and connected species. To further explore the mechanisms underlying
-the temporal dynamics of the networks, we examine patterns in the
-species and interaction temporal turnover. Lastly, we investigate
-whether networks are assembling toward predictable interaction
-patterns, and the ramifications for the robustness of the networks to
-species extinction and cascading perturbations.
+determine whether the mechanism underlying network assembly is a build
+up of interactions as would be predicted by preferential attachment,
+or punctuated by significant reorganizations of interactions (i.e.,
+network changing points). Even with changing points in interaction
+organization, networks could still be assembling via preferential
+attachment if the network reorganizations were primarily driven the by
+peripheral, temporally variable species while a stable, well-connected
+core of species persist. We test whether the species that are most
+variable in their network position --- and thus important contributors
+to network reorganizations --- are less persistent and connected
+species. To further explore the mechanisms underlying the temporal
+dynamics of the networks, we examine patterns in the species and
+interaction temporal turnover. Lastly, we investigate whether networks
+are assembling toward predictable interaction patterns, and the
+ramifications for the robustness of the networks to species
+extinction.
 
@@
 
-We surveyed plant-pollinator interaction networks of assembling
-hedgerows communities (N=5), as well as in two types of non-assembling
-communities to serve as controls: unrestored, weedy field margins
-(N=19) and mature hedgerows (greater than 10 years since planting,
-N=29). The sites were located in the Central Valley of California in
-Yolo, Colusa and Solano Counties. This area is composed of intensively
-managed agriculture --- primarily monocultures of conventional row
-crops, vineyards and orchards. Hedgerows we planted along field
-margins where they do not remove valuable land from
+We surveyed plant-pollinator interaction networks of independent
+assembling hedgerows communities (N=5), as well as in two types of
+non-assembling communities to serve as controls: unrestored, weedy
+field margins (N=19) and mature hedgerows (greater than 10 years since
+planting, N=29). The sites were located in the Central Valley of
+California in Yolo, Colusa and Solano Counties. This area is composed
+of intensively managed agriculture --- primarily monocultures of
+conventional row crops, vineyards and orchards. Hedgerows we planted
+along field margins where they do not remove valuable land from
 production. Plantings are between are ca.~3--6 m wide and
@@
 patterns. A change point is caused by a merge, split, fragmentation or
-formation of communities (also called modules or compartments). Change
-point detection methods have yet to be generalized to quantitative
-networks, so for this analysis we focused on qualitative (binary)
+formation of modules (also called compartments). Change point
+detection methods have yet to be generalized to quantitative networks,
+so for this analysis we focused on qualitative (binary)
 networks. Following \cite{peel2014detecting}, we first defined a
@@
 hierarchical random graph model (GHRG). The GHRG model is able to
-capture both assortative and disassortative community structure
-patterns at all scales in the network \citep{peel2014detecting}. A
-network $G$ is composed of vertices $V$ and edges $E$. The GHRG model
-decomposes the $N$ vertices into a series of nested groups, the
-relationships among which are represented by the dendrogram $T$. The
-tips of $T$ are the vertices of $G$, and the probability that two
-vertices $u$ and $v$ connect is given by the parameter $p_r$. The
-probability distribution of the network $G$ thus defined as:
+capture both assortative and disassortative structure patterns at all
+scales in the network \citep{peel2014detecting}. A network $G$ is
+composed of vertices $V$ and edges $E$. The GHRG model decomposes the
+$N$ vertices into a series of nested groups, the relationships among
+which are represented by the dendrogram $T$. The tips of $T$ are the
+vertices of $G$, and the probability that two vertices $u$ and $v$
+connect is given by the parameter $p_r$. The probability distribution
+of the network $G$ thus defined as:
 
@@
 hedgerows and weedy field margins, only sites with five or greater
-survey years were included in this analysis. All statistical analysis
-were conducted in R 3.2.3 \citep{R}.
+survey years were included in this analysis (N=11). All statistical
+analysis were conducted in R 3.2.3 \citep{R}.
 
@@
 position variability, we computed the coefficient of variation of
-weighted closeness at each site through time. Closeness describes the
-centrality of a species in the network by calculating path lengths to
-other vertices (species) in the graph. We use linear mixed models to
-test whether the species closeness variability (log) is related to the
-persistence or degree of that species \citep{lme4, lmetest}. We
-included random effects for species, as well as site. Because the
-degree and persistence of pollinators were strongly correlated, ($\rho
-= 0.84$, $p$-value $<$ $2*10^{-16}$), we include each explanatory
-variable in the model separately. Plant degree and persistence were
-not significantly correlated, but we use the same models as we did for
-the pollinators for consistency. A linear increase in closeness, as
-might be expected with assembly by preferential attachment, would also
-lead to a high variability in closeness scores, so we test whether
-closeness increases through time.
+weighted closeness centrality \citep{freeman1978centrality} at each
+site through time. Closeness centrality describes the centrality of a
+species in the network by calculating path lengths to other vertices
+(species) in the graph \citep{freeman1978centrality}. We use linear
+mixed models to test whether the species closeness variability (log)
+is related to the persistence or degree of that species \citep{lme4,
+  lmetest}. We included random effects for species, as well as
+site. Because the degree and persistence of pollinators were strongly
+correlated, ($\rho = 0.84$, $p$-value $<$ $2*10^{-16}$), we include
+each explanatory variable in the model separately. Plant degree and
+persistence were not significantly correlated, but we use the same
+models as we did for the pollinators for consistency. An approximately
+logarithmic increase in closeness centrality, as would be expected
+with assembly by preferential attachment, would also lead to a high
+variability in closeness scores, so we test whether log closeness
+centrality increases through time.
 
@@
 species present, interactions may turnover more quickly than in
-established communities. In addition, at assembling hedgerows, plants
-may ``colonize'' the networks as they become more attractive
-resources, and establish new interactions with pollinators.
+established communities. In addition, at assembling hedgerows,
+unvisited plants may appear to ``colonize'' the networks as they
+establish new interactions with pollinators.
 
@@
 with network assembly affect the robustness of the network to species
-loss and to cascading perturbations. Following \cite{Memmott2004}, we
-simulate plant species extinction and the subsequent extinction
-cascades of pollinator species. Because the reproduction of plant
-species if facilitated by active restoration efforts, it is unlikely
-the extinction of pollinator species would affect plant populations in
-the hedgerows. However, plants ceasing to bloom, for example in
-response to drought, will likely affect the pollinators that depend on
-them. We eliminate plants species based on their degree or abundance,
-and then calculate the number of pollinators that secondarily went
-extinct. The area below the extinction curve is an estimate of network
-robustness \citep{Memmott2004, bipartite}.
+loss. Following \cite{Memmott2004}, we simulate plant species
+extinction and the subsequent extinction cascades of pollinator
+species. Because the reproduction of plant species if facilitated by
+active restoration efforts, it is unlikely the extinction of
+pollinator species would affect plant populations in the
+hedgerows. However, plants ceasing to bloom, for example in response
+to drought, will likely affect the pollinators that depend on them. We
+eliminate plants species based on their degree or abundance, and then
+calculate the number of pollinators that secondarily went extinct. The
+area below the extinction curve is an estimate of network robustness
+\citep{Memmott2004, bipartite}.
 
-We also explored how the robustness to cascading perturbations changed
-as community assembled, using algebraic connectivity --- the second
-smallest eigenvalue of the Laplacian matrix
-\citep{fiedler1973algebraic} --- as a proxy for network robustness
-(Gaiarsa et al., submitted). Algebraic connectivity relates to how
-difficult it is to turn a network into completely disconnected groups
-of species \citep{costa2007characterization}. The larger the algebraic
-connectivity, the more robust a network is to cascading perturbations,
-and the harder it is to break the community into isolated groups of
-species (Gaiarsa et al., submitted).
+% We also explored how the robustness to cascading perturbations changed
+% as community assembled, using algebraic connectivity --- the second
+% smallest eigenvalue of the Laplacian matrix
+% \citep{fiedler1973algebraic} --- as a proxy for network robustness
+% (Gaiarsa et al., submitted). Algebraic connectivity relates to how
+% difficult it is to turn a network into completely disconnected groups
+% of species \citep{costa2007characterization}. The larger the algebraic
+% connectivity, the more robust a network is to cascading perturbations,
+% and the harder it is to break the community into isolated groups of
+% species (Gaiarsa et al., submitted).
 
@@
 related to network position variability (Fig.~\ref{fig:cv}, estimate
-of the slope of closeness variability and persistence $\pm$ standard
-error of the estimate, $0.653 \pm 0.225$, $p$-value=$0.009$; slope of
-closeness variability and degree, $0.008 \pm 0.002$,
-$p$-value=$0.002$). In addition, plant persistence and degree were not
-significantly related to network position variability
+of the slope of closeness centrality variability and persistence $\pm$
+standard error of the estimate, $0.653 \pm 0.225$, $p$-value=$0.009$;
+slope of closeness centrality variability and degree, $0.008 \pm
+0.002$, $p$-value=$0.002$). In addition, plant persistence and degree
+were not significantly related to network position variability
 (Fig.~\ref{fig:cv}, estimate of the slope of closeness variability and
-persistence $\pm$ SE, $-2.063 \pm 3.091$,
-$p$-value=$0.5$; slope of closeness variability and degree, $0.0018
-\pm 0.002$, $p$-value=$0.3$). The variability of species network
-position was not the result of closeness linearly increasing through
-time, and, in fact, plant and pollinator closeness decreased slightly
-through time (Fig.~S2, estimate of the slope of closeness through time
-$\pm$ SE, pollinators: $-0.0003 \pm
-0.00005$, $p$-value=$2.7*10^{-12}$; plants $-0.007 \pm 0.001$,
-$p$-value=$1.4*10^{-6}$). Through statistically significant, the
-slopes are so slight they may not be biologically significant.
+persistence $\pm$ SE, $-2.063 \pm 3.091$, $p$-value=$0.5$; slope of
+closeness variability and degree, $0.0018 \pm 0.002$,
+$p$-value=$0.3$). The variability of species network position was not
+the result of closeness linearly increasing through time, and, in
+fact, plant and pollinator closeness decreased slightly through time
+(Fig.~S2, estimate of the slope of closeness through time $\pm$ SE,
+pollinators: $-0.0003 \pm 0.00005$, $p$-value=$2.7*10^{-12}$; plants
+$-0.007 \pm 0.001$, $p$-value=$1.4*10^{-6}$). Through statistically
+significant, the slopes are so slight they may not be biologically
+significant.
 
@@
 (Fig.~\ref{fig:baci}, estimate of the slope of nestedness through time
-$\pm$ SE, $1.834 \pm 0.6142$,
-$p$-value=$0.022$). Modularity decreased (Fig.~\ref{fig:baci}),
-though the slope was not significantly different from zero (estimate
-of the slope of modularity through time $\pm$ standard error of the
-estimate, $-0.524$ $\pm$ $0.295$, $p$-value=$0.124$). Connectance
-decreased as the community assembled (Fig.~\ref{fig:baci}, estimate of
-the slope of connectance through time $\pm$ standard error of the
-estimate, $-0.0434$ $\pm$ $0.0152$, $p$-value=$0.03$). In contrast,
-specialization remained constant (estimate of the slope of
-specialization through time $\pm$ SE,
-$0.003$ $\pm$ $0.015$, $p$-value=$0.827$).
+$\pm$ SE, $1.834 \pm 0.6142$, $p$-value=$0.022$). Modularity decreased
+(Fig.~\ref{fig:baci}), though the slope was not significantly
+different from zero (estimate of the slope of modularity through time
+$\pm$ standard error of the estimate, $-0.524$ $\pm$ $0.295$,
+$p$-value=$0.124$). In addition, none of the networks were
+significantly modular ($z$-scores $<$ 2, Fig.~\ref{fig:baci})
+Connectance decreased as the community assembled (Fig.~\ref{fig:baci},
+estimate of the slope of connectance through time $\pm$ standard error
+of the estimate, $-0.0434$ $\pm$ $0.0152$, $p$-value=$0.03$). In
+contrast, specialization remained constant (estimate of the slope of
+specialization through time $\pm$ SE, $0.003$ $\pm$ $0.015$,
+$p$-value=$0.827$).
 
@@
 extinction when species where removed incrementally by degree
-(Fig.~\ref{fig:rob}, estimate of the slope of robustness through time
-$\pm$ SE, $6*10^{-5} \pm 4*10^{-3}$, $p$-value=$0.987$) or abundance
-($0.001 \pm 0.003$, $p$-value=$0.65$). In contrast, the robustness of
-networks to cascading perturbations, as measured by the algebraic
-connectivity of the network, increased as the network assembled
-(Fig.~\ref{fig:rob}, estimate of the slope of robustness to cascading
-perturbations through time $\pm$ SE, $0.6814 \pm 0.272$,
-$p$-value=$0.042$).
+(estimate of the slope of robustness through time $\pm$ SE, $6*10^{-5}
+\pm 4*10^{-3}$, $p$-value=$0.987$) or abundance ($0.001 \pm 0.003$,
+$p$-value=$0.65$). % In contrast, the robustness of
+% networks to cascading perturbations, as measured by the algebraic
+% connectivity of the network, increased as the network assembled
+% (Fig.~\ref{fig:rob}, estimate of the slope of robustness to cascading
+% perturbations through time $\pm$ SE, $0.6814 \pm 0.272$,
+% $p$-value=$0.042$).
 
@@
 during assembly, the way in which these reorganizations occur was
-different from what is expected by preferential attachment Based on a
+different from what is expected by preferential attachment. Based on a
 preferential attachment process, we expect that the most persistent
 and high degree species would remain stable in the network core during
-assembly. Surprisingly, however, we encountered the opposite
-pattern. For example, the four most ubiquitous species in our study
-landscape --- \textit{Halictus ligatus}, \textit{Halictus
-  tripartitus}, \textit{Lasioglossum (Dialictus) incompletum}, and
-\textit{Toxomerus marginatus} --- were the only species that changed
-which module (i.e., community) they were a member across years in all
-the assembling hedgerows. Because species degree and persistence were
-strongly correlated, it is difficult to disentangle the causal
-mechanism for why species with those characteristics are so variable
-in their network position. Generalized species may be able to better
-exploit the limited floral resources in the intensively managed
-agriculture landscape, and thus also be the most persistent. More
-persistent species usually have longer phenologies, so they can visit
-many different flowers, resulting in a higher degree. Either way, our
+assembly \citep{barabasi1999emergence}. Surprisingly, however, we
+encountered the opposite pattern. For example, the four most
+ubiquitous species in our study landscape --- \textit{Halictus
+  ligatus}, \textit{Halictus tripartitus}, \textit{Lasioglossum
+  (Dialictus) incompletum}, and \textit{Toxomerus marginatus} --- were
+the only species that changed which module they were a member across
+years in all the assembling hedgerows. Because species degree and
+persistence were strongly correlated, it is difficult to disentangle
+the causal mechanism for why species with those characteristics are so
+variable in their network position. Generalized species may be able to
+better exploit the limited floral resources in the intensively managed
+agriculture landscape, and thus also be the most persistent
+\flagged{any citations? I can only find articles discussing
+  generalization in plants}. More persistent species usually have
+longer phenologies, so they can visit many different flowers,
+resulting in a higher degree \citep{Vazquez2009}. Either way, our
 result suggests that adaptable species can change their network
@@
 which may depend on the other pollinator species that are present, and
-the state of the plant community. Thus given the opportunity and
-ability to use different resources, species will often change their
-network positions.
+the state of the plant community \citep{macleod2016measuring,
+  gomez2006ecological, waser1996generalization}. Thus given the
+opportunity and ability to use different resources, species will often
+change their network positions \citep{macleod2016measuring}.
 
@@
 species and the overall mean number of interactions per species did
-not change. With preferential attachment, however, we would also
-expect specialization to increase, and we found no such
-trend. Furthermore, the frequent changing points in network
+not change. However, the frequent changing points in network
 organization, dynamic nature of species positions in the networks, and
@@
 nestedness. Because so many mechanisms give rise to the same patterns
-of interaction, additional tests are necessary to assess the contribution
-of different mechanisms to community assembly.
+of interaction, additional tests are necessary to assess the
+contribution of different mechanisms to community assembly.
 
@@
 associated with an enhanced in robustness to extinction
-\citep{Memmott2004}. Perhaps assembling hedgerows have yet to reach sufficient
-levels of nestedness to realize the benefits nestedness
+\citep{Memmott2004}. Perhaps assembling hedgerows have yet to reach
+sufficient levels of nestedness to realize the benefits nestedness
 confers. Nestedness of the assembling hedgerows, however, did not
@@
 \citep{Vazquez2009, Burkle2011}. Unlike in the broader food web
-literature, we have few mechanistic models of mutualistic network
+literature, we have few assembly models of mutualistic network
 assembly \citep{valdovinos2013adaptive, Nuismer2013, Guimaraes2011}.
-In addition, the few developed often borrow their mechanisms from
-competitive interactions, leading to inaccurate biological assumptions
-\citep{holland2006comment}. We need further development of mechanistic
-models of mutualistic systems to generate testable predictions, along
-with empirical exploration of network assembly. Plant-pollinator
-communities and mutualisms broadly are vital for biodiversity
-maintenance and essential ecosystem service provision. We must
-therefore understand the processes underlying their assembly to
-facilitate restoration and conservation.
+In addition, the few developed models often borrow their mechanisms
+from competitive interactions, leading to inaccurate biological
+assumptions \citep{holland2006comment}. We need further development of
+mechanistic models of mutualistic systems to generate testable
+predictions, along with empirical exploration of network
+assembly. Plant-pollinator communities and mutualisms broadly are
+vital for biodiversity maintenance and essential ecosystem service
+provision. We must therefore understand the processes underlying their
+assembly to facilitate restoration and conservation.
 
@@
 
-We would like to thank Paulo Guimar{\~a}es, Aaron Clauset and Matthew
+We would like to thank Paulo R.~Guimar{\~a}es Jr., Aaron Clauset and Matthew
 Hutchinson for their invaluable discussions and comments, and Leto

### 2016-10-04 analysis/speciesLevel/plotting/specMets.R

@@ -45,15 +45,15 @@ for(j in pp){
 
 ## closeness only
 
-ypr.pi.pol <- predict.int(mod= mods$pol$closeness,
+ypr.pi.pol <- predict.int(mod= mods$pol$closeness.log,
                       dd=dd.ypr,
-                      y="closeness",
+                      y="closeness.log",
                       family="gaussian")
 
 
-ypr.pi.plant <- predict.int(mod= mods$plant$closeness,
+ypr.pi.plant <- predict.int(mod= mods$plant$closeness.log,
                       dd=dd.ypr,
-                      y="closeness",
+                      y="closeness.log",
                       family="gaussian")
 
 plot.panels()

### 2016-10-04 analysis/speciesLevel/plotting/src/plotPanels.R

@@ -8,7 +8,7 @@ plot.panels <- function(){
     ## nodf
     plot.panel(new.dd=ypr.pi.pol,
                dats=specs,
-               y1="closeness",
+               y1="closeness.log",
                xs="ypr",
                col.fill=col.fill,
                col.lines=col.lines,
@@ -19,7 +19,7 @@ plot.panels <- function(){
 
     plot.panel(new.dd=ypr.pi.plant,
                dats=specs,
-               y1="closeness",
+               y1="closeness.log",
                xs="ypr",
                col.fill=col.fill,
                col.lines=col.lines,
@@ -27,7 +27,7 @@ plot.panels <- function(){
                agg.col="GenusSpecies")
     
     axis(1, pretty(specs$ypr), labels=pretty(specs$ypr))
-    mtext("Closeness", 2, line=4.5, cex=1.5, adj=1.75)
+    mtext("Closeness (log)", 2, line=4.5, cex=1.5, adj=4)
     mtext("Plants", 3, line=0.5, cex=1.5)
     mtext("Years of assembly", 1, line=3, cex=1.5)
 

### 2016-10-04 analysis/speciesLevel/specMets.R

@@ -11,12 +11,15 @@ load('../../data/networks/all_networks_years.Rdata')
 
 ## linear models
 load(file=file.path(save.path, 'specs.Rdata'))
+specs$closeness[specs$closeness == 0] <- 1*10^-6
+specs$closeness.log <- log(specs$closeness)
+
 ## SiteStatus or ypr
 xvar <- "ypr"
 
 ## anything outputted by specieslevel
 ys <- c("proportional.generality", "d", "degree", "betweenness",
-        "closeness")
+        "closeness.log")
 
 ## formulas <-lapply(ys, function(x) {
 ##   as.formula(paste(x, "~",

### 2016-10-04 ms/ms.tex
made pailo's comments

@@
 better exploit the limited floral resources in the intensively managed
-agriculture landscape, and thus also be the most persistent
-\flagged{any citations? I can only find articles discussing
-  generalization in plants}. More persistent species usually have
-longer phenologies, so they can visit many different flowers,
-resulting in a higher degree \citep{Vazquez2009}. Either way, our
-result suggests that adaptable species can change their network
+agriculture landscape, and thus also be the most persistent \citep[in
+ant-plant mutualisms, ][]{diaz2010changes}. More persistent species
+usually have longer phenologies, so they can visit many different
+flowers, resulting in a higher degree \citep{Vazquez2009}. Either way,
+our result suggests that adaptable species can change their network
 position to utilize the most advantageous floral resources available,
@@
 
-We would like to thank Paulo R.~Guimar{\~a}es Jr., Aaron Clauset and Matthew
-Hutchinson for their invaluable discussions and comments, and Leto
-Peel for help with the change point analysis.  We thank the growers
-and land owners that allowed us to work on their property. We also
-greatly appreciate the identification assistance of expert taxonomists
-Martin Hauser, Robbin Thorp and Jason Gibbs.  This work was supported
-by funding from the Army Research Office (W911NF-11-1-0361 to CK), the
-Natural Resources Conservation Service (CIG-69-3A75-12-253,
-CIG-69-3A75-9-142, CIG-68-9104-6-101 and WLF-69-7482-6-277 to The
-Xerces Society), the National Science Foundation (DEB-0919128 to CK),
-The U.S.  Department of Agriculture (USDA-NIFA 2012-51181-20105 to
-Michigan State University).  Funding for LCP was provided by an NSF
-Graduate Research Fellowship, the USDA NIFA Graduate Fellowship, and
-the Berkeley Institute for Data Science. Funding for MPG was provided
-by S{\~a}o Paulo Research Foundation (FAPESP, grant 2013/13319-5). We
-also appreciate the Santa Fe Institute for faciliting this
-international collaboration.
+We would like to thank Paulo R.~Guimar{\~a}es Jr., Aaron Clauset and
+Matthew Hutchinson for their invaluable discussions and comments, and
+Leto Peel for help with the change point analysis.  We thank the
+growers and land owners that allowed us to work on their property. We
+also greatly appreciate the identification assistance of expert
+taxonomists Martin Hauser, Robbin Thorp and Jason Gibbs.  This work
+was supported by funding from the Army Research Office
+(W911NF-11-1-0361 to CK), the Natural Resources Conservation Service
+(CIG-69-3A75-12-253, CIG-69-3A75-9-142, CIG-68-9104-6-101 and
+WLF-69-7482-6-277 to The Xerces Society), the National Science
+Foundation (DEB-0919128 to CK), The U.S.  Department of Agriculture
+(USDA-NIFA 2012-51181-20105 to Michigan State University).  Funding
+for LCP was provided by an NSF Graduate Research Fellowship, the USDA
+NIFA Graduate Fellowship, and the Berkeley Institute for Data
+Science. Funding for MPG was provided by S{\~a}o Paulo Research
+Foundation (FAPESP, grant 2013/13319-5). We also appreciate the Santa
+Fe Institute for faciliting this international collaboration.
 

### 2016-10-06 ms/ms.tex
marilia made small litte changes in the tex file after Paulos comments

@@
   non-assembling networks did not restructure as frequently. Across
-  all communities, Pollinator species were opportunistic in the
+  all communities, pollinator species were opportunistic in the
   flowers they visited. Indeed, the most persistent and generalized
@@
 species in the network by calculating path lengths to other vertices
-(species) in the graph \citep{freeman1978centrality}. We use linear
+(species) in the graph \citep{freeman1978centrality}. The smaller the mean path length of a species to other species, the closer they are, and the greater the species' centrality. We use linear
 mixed models to test whether the species closeness variability (log)

### 2016-10-28 ms/ms.tex
misc

--- 2016-10-04 ms/ms.tex
@@
   time, but there was no increase in robustness to simulated plant
-  extinctions. Our results elucidate some of the mechanisms underlying
-  plant-pollinator network assembly and restoration, while providing
-  further evidence that hedgerows are a valuable tool for promoting
-  species conservation and ecosystem provision in agricultural areas.
+  extinctions. The sensitivity of networks to cascading
+  perturbations, however, increased as the communities assembled, at
+  least partially due to accumulating species richness. We elucidate
+  some of the mechanisms underlying plant-pollinator network assembly
+  and restoration, while providing further evidence that hedgerows are
+  a valuable tool for promoting species conservation and ecosystem
+  provision in agricultural areas.
 \end{abstract}
@@
 %% Lauren I love this paragraph!
-In contrast to the ordered network build-up described by preferential
+In contrast to the network build-up described by preferential
 attachment, significant reorganizations of interactions can punctuate
@@
 since an area was restored has been shown to affect the structure of
-networks \citep{forup-2008-742, forup2008restoration,
-  devoto2012understanding}, suggesting interactions are changing as
-the community develops. Understanding the mechanisms of network
-assembly will help to guide community restoration. Facilitating
-network restoration is especially imperative in areas where species
-interactions provide essential ecosystem services, such as crop
-pollination. In intensively managed agricultural landscapes, the
-demand for pollination services is the greatest
-\citep{kremen-2008-10}. However, honey bees, managed extensively
-around the world to provide crop pollination, are in global decline
-\citep{neumann-2010-1, van-engelsdorp-2009-e6481}. In addition, native
-pollinators, which are capable of providing sufficient crop
-pollination \citep{kremen-2002-16816, winfree-2007-1105,
-  kremen-2004-1109}, are in short supply because these landscapes make
-poor habitats for pollinator populations \citep{kremen-2002-16816}. To
-ensure the continued provision of ecosystem services and curb
-biodiversity loss, it is critical to restore pollinators and their
-interactions in agricultural landscapes. To promote pollinator
-services in agriculture, farmers may chose to plant strips of native
-plants along farm edges (hedgerows) to help provide habitat for
-pollinators without removing arable land from production. Hedgerows
-augment the richness, abundance and trait diversity of pollinators in
-agricultural landscapes \citep{morandin-2013-829, mgonigle-2015-x,
-  kremen-2015-602, ponisio2015farm}. In addition, hedgerows promote
-the persistence and colonization of floral resource specialists
-\citep{mgonigle-2015-x}. Little is known, however, about how new
-species are being incorporated into the network as the community
-assembles, or the consequences for interaction patterns and
+networks \citep{forup-2008-742, devoto2012understanding}, suggesting
+interactions are changing as the community develops. Understanding the
+mechanisms of network assembly will help to guide community
+restoration. Facilitating network restoration is especially imperative
+in areas where species interactions provide essential ecosystem
+services, such as crop pollination. In intensively managed
+agricultural landscapes, the demand for pollination services is the
+greatest \citep{kremen-2008-10}. However, honey bees, managed
+extensively around the world to provide crop pollination, are in
+global decline \citep{neumann-2010-1, van-engelsdorp-2009-e6481}. In
+addition, native pollinators, which are capable of providing
+sufficient crop pollination \citep{kremen-2002-16816,
+  winfree-2007-1105, kremen-2004-1109}, are in short supply because
+these landscapes make poor habitats for pollinator populations
+\citep{kremen-2002-16816}. To ensure the continued provision of
+ecosystem services and curb biodiversity loss, it is critical to
+restore pollinators and their interactions in agricultural
+landscapes. To promote pollinator services in agriculture, farmers may
+chose to plant strips of native plants along farm edges (hedgerows) to
+help provide habitat for pollinators without removing arable land from
+production. Hedgerows augment the richness, abundance and trait
+diversity of pollinators in agricultural landscapes
+\citep{morandin-2013-829, mgonigle-2015-x, kremen-2015-602,
+  ponisio2015farm}. In addition, hedgerows promote the persistence and
+colonization of floral resource specialists
+\citep{mgonigle-2015-x}. It is important to further understand how
+these new species are being incorporated into the network as the
+community assembles, or the consequences for interaction patterns and
 robustness.
@@
 are assembling toward predictable interaction patterns, and the
-ramifications for the robustness of the networks to species
-extinction.
+ramifications for the robustness of the networks to species extinction
+and cascading perturbations.
 
@@
 \textit{Baccharis spp.}, \textit{Salvia spp}. and others
-\citep{kremen-2015-602, mgonigle-2015-x}. The mean distance between
-monitoring sites was 15 km, and the minimum distance between sites of
-the same type sampled in the same year was 2 km.  The entire area
-surveyed spanned almost 300 km$^2$. The crop fields adjacent to all
-sites were similarly managed as intensive, high-input monoculture.
+\citep[Fig.~S1][]{menz-2010-4, kremen-2015-602, mgonigle-2015-x}. The
+mean distance between monitoring sites was $15$ km, and the minimum
+distance between sites of the same type sampled in the same year was 2
+km.  The entire area surveyed spanned almost $300$ km$^2$. The crop
+fields adjacent to all sites were similarly managed as intensive,
+high-input monoculture.
 
@@
 species in the network by calculating path lengths to other vertices
-(species) in the graph \citep{freeman1978centrality}. We use linear
-mixed models to test whether the species closeness variability (log)
-is related to the persistence or degree of that species \citep{lme4,
-  lmetest}. We included random effects for species, as well as
-site. Because the degree and persistence of pollinators were strongly
-correlated, ($\rho = 0.84$, $p$-value $<$ $2*10^{-16}$), we include
-each explanatory variable in the model separately. Plant degree and
-persistence were not significantly correlated, but we use the same
-models as we did for the pollinators for consistency. An approximately
-logarithmic increase in closeness centrality, as would be expected
-with assembly by preferential attachment, would also lead to a high
-variability in closeness scores, so we test whether log closeness
-centrality increases through time.
+(species) in the graph \citep{freeman1978centrality}. The shorter the
+mean path length to other species, the higher is the closeness
+centrality. We use linear mixed models to test whether the species
+closeness variability (log) is related to the persistence or degree of
+that species \citep{lme4, lmetest}. We included random effects for
+species, as well as site. Because the degree and persistence of
+pollinators were strongly correlated, ($\rho = 0.84$, $p$-value $<$
+$2*10^{-16}$), we include each explanatory variable in the model
+separately. Plant degree and persistence were not significantly
+correlated, but we use the same models as we did for the pollinators
+for consistency. An approximately logarithmic increase in closeness
+centrality, as would be expected with assembly by preferential
+attachment, would also lead to a high variability in closeness scores,
+so we test whether log closeness centrality increases through time.
 
@@
 essentially replace each other if they fill similar roles in the
-network, ocupying the same network position and interacting with
+network, occupying the same network position and interacting with
 similar species. At non-assembling communities, species turnover may
@@
 are weighted by the similarity of their constituent species
-(Fig.~S1). We develop a method to examine the temporal turnover of
-interactions with weightings based on their similarity. We followed
-\cite{ahn2010link} algorithm to hierarchically cluster all the
-interactions (edges) across sites and years based on their similarity,
-and build a dendrogram. The interaction similarity is based how may
-plants and pollinators (vertices) two edges share \citep{ahn2010link,
-  kalinka2011linkcomm}. The more species edges shared in common, the
-shorter the branch length between them on the dendrogram.  We next
-calculated the temporal turnover of interactions weighted by their
-similarity, as approximated by ``phylogenetic'' distance
-\citep{graham2008phylogenetic, picante-2010-1463}. When then use
-linear models to test whether the weighted turnover of interactions
-varied between assembling and non-assembling networks \citep{lme4,
-  lmetest}.
+(Fig.~\ref{fig:methods}). We develop a method to examine the temporal
+turnover of interactions with weightings based on their similarity. We
+followed \cite{ahn2010link} algorithm to hierarchically cluster all
+the interactions (edges) across sites and years based on their
+similarity, and build a dendrogram. The interaction similarity is
+based how may plants and pollinators (vertices) two edges share
+\citep{ahn2010link, kalinka2011linkcomm}. The more species edges
+shared in common, the shorter the branch length between them on the
+dendrogram.  We next calculated the temporal turnover of interactions
+weighted by their similarity, as approximated by ``phylogenetic''
+distance \citep{graham2008phylogenetic, picante-2010-1463}. When then
+use linear models to test whether the weighted turnover of
+interactions varied between assembling and non-assembling networks
+\citep{lme4, lmetest}.
 
@@
 To evaluate network nestedness, we used the estimator weighted NODF
-\citep{nodf}. NODF evaluates whether species with fewer partners
-interact with subsets of partners with which more connected species
-interact \citep{nodf}. To estimate modularity, we use a hierarchical
-clustering algorithm \citep{Newman2004, igraph}.  We calculated
-standardized $z$-scores so that nestedness and modularity metrics
-could be compared across communities. The $z$-scores were calculated
-by generating an ensemble of $999$ randomly assembled communities,
-subtracting the mean of the statistic calculated across these
-communities from the observed value, and then dividing by the standard
-deviation. To assemble random communities, we reshuffled the
-interactions between species but fixed the total number of
-interactions, species and interaction frequency distributions
-\citep{Galeano2009}. We evaluate network specialization with the
-metric H2, which estimates the deviation of the observed interaction
-frequency between plants and pollinators from a null expectation where
-all partners interact in proportion to their abundances
-\citep{bluthgen-2006-9}. It ranges from zero for generalized networks
-to one for specialized networks.
+\citep{almeida-neto-2008-1227}. NODF evaluates whether species with
+fewer partners interact with subsets of partners with which more
+connected species interact \citep{almeida-neto-2008-1227}. To estimate
+modularity, we use a hierarchical clustering algorithm
+\citep{Newman2004, csardi-2006}. We evaluate network specialization with
+the metric H2, which estimates the deviation of the observed
+interaction frequency between plants and pollinators from a null
+expectation where all partners interact in proportion to their
+abundances \citep{bluthgen-2006-9}. It ranges from zero for
+generalized networks to one for specialized networks.  We calculated
+standardized $z$-scores so that nestedness, modularity and
+specialization metrics could be compared across communities. The
+$z$-scores were calculated by generating an ensemble of $999$ randomly
+assembled communities, subtracting the mean of the statistic
+calculated across these communities from the observed value, and then
+dividing by the standard deviation. To assemble random communities, we
+reshuffled the interactions between species but fixed the total number
+of interactions, species and interaction frequency distributions
+\citep{Galeano2009}.
 
@@
 with network assembly affect the robustness of the network to species
-loss. Following \cite{Memmott2004}, we simulate plant species
-extinction and the subsequent extinction cascades of pollinator
-species. Because the reproduction of plant species if facilitated by
-active restoration efforts, it is unlikely the extinction of
-pollinator species would affect plant populations in the
-hedgerows. However, plants ceasing to bloom, for example in response
-to drought, will likely affect the pollinators that depend on them. We
-eliminate plants species based on their degree or abundance, and then
-calculate the number of pollinators that secondarily went extinct. The
-area below the extinction curve is an estimate of network robustness
-\citep{Memmott2004, bipartite}.
+loss and cascading perturbations. Following \cite{Memmott2004}, we
+simulate plant species extinction and the subsequent extinction
+cascades of pollinator species. Because the reproduction of plant
+species if facilitated by active restoration efforts, it is unlikely
+the extinction of pollinator species would affect plant populations in
+the hedgerows. However, plants ceasing to bloom, for example in
+response to drought, will likely affect the pollinators that depend on
+them. We eliminate plants species based on their degree or abundance,
+and then calculate the number of pollinators that secondarily went
+extinct. The area below the extinction curve is an estimate of network
+robustness \citep{Memmott2004, bipartite}.
 
-% We also explored how the robustness to cascading perturbations changed
-% as community assembled, using algebraic connectivity --- the second
-% smallest eigenvalue of the Laplacian matrix
-% \citep{fiedler1973algebraic} --- as a proxy for network robustness
-% (Gaiarsa et al., submitted). Algebraic connectivity relates to how
-% difficult it is to turn a network into completely disconnected groups
-% of species \citep{costa2007characterization}. The larger the algebraic
-% connectivity, the more robust a network is to cascading perturbations,
-% and the harder it is to break the community into isolated groups of
-% species (Gaiarsa et al., submitted).
+We also explored how the robustness to cascading perturbations changed
+as community assembled, using algebraic connectivity --- the second
+smallest eigenvalue of the Laplacian matrix
+\citep{fiedler1973algebraic} --- as a proxy for network robustness
+(Gaiarsa et al., submitted). Algebraic connectivity relates to how
+difficult it is to turn a network into completely disconnected groups
+of species \citep{costa2007characterization}, and has implications to
+how integrated is the dynamics of the elements of the networks
+\citep{gibert2013spatial}. The larger the algebraic connectivity, the
+more robust a network is to cascading perturbations, and the easier it
+is to break the community into isolated groups of species (Gaiarsa et
+al., submitted).
 
@@
 (Fig.~\ref{fig:baci}, estimate of the slope of nestedness through time
-$\pm$ SE, $1.834 \pm 0.6142$, $p$-value=$0.022$). Modularity decreased
-(Fig.~\ref{fig:baci}), though the slope was not significantly
-different from zero (estimate of the slope of modularity through time
-$\pm$ standard error of the estimate, $-0.524$ $\pm$ $0.295$,
-$p$-value=$0.124$). In addition, none of the networks were
-significantly modular ($z$-scores $<$ 2, Fig.~\ref{fig:baci})
-Connectance decreased as the community assembled (Fig.~\ref{fig:baci},
-estimate of the slope of connectance through time $\pm$ standard error
-of the estimate, $-0.0434$ $\pm$ $0.0152$, $p$-value=$0.03$). In
-contrast, specialization remained constant (estimate of the slope of
-specialization through time $\pm$ SE, $0.003$ $\pm$ $0.015$,
-$p$-value=$0.827$).
+$\pm$ SE, $1.834 \pm 0.6142$, $p$-value=$0.022$). All of the networks
+were significantly nested ($z$-scores $>$ 2,
+Fig.~\ref{fig:baci}). Modularity decreased (Fig.~\ref{fig:baci}),
+though the slope was not significantly different from zero (estimate
+of the slope of modularity through time $\pm$ standard error of the
+estimate, $-0.524$ $\pm$ $0.295$, $p$-value=$0.124$). In addition,
+none of the networks were significantly modular ($z$-scores $<$ 2,
+Fig.~\ref{fig:baci}). Connectance decreased as the community assembled
+(Fig.~\ref{fig:baci}, estimate of the slope of connectance through
+time $\pm$ standard error of the estimate, $-0.0434$ $\pm$ $0.0152$,
+$p$-value=$0.03$). Specialization also decreased, though the slope was
+only marginally significantly different from zero (estimate of the
+slope of specialization through time $\pm$ SE, $-0.926$ $\pm$ $0.450$,
+$p$-value=$0.078$). Most communities were more generalized than
+expected when interactions were randomized (Fig.~\ref{fig:baci}).
 
@@
 (Fig.~\ref{fig:baci}, estimate of the slope of richness through time
-$\pm$ SE, pollinators: $0.193$ $\pm$
-$0.0729$, $p$-value=$0.008$; plants: $0.212$ $\pm$ $0.0653$,
-$p$-value=$0.001$). Unsurprisingly, pollinator species are colonizing
-and persisting at the assembling hedgerows. Plant species richness is
-based on the flowers bbeing visiting by pollinators and not the
-presence of a particular plant species at a site. Thus, though some
-new plant species may establish themselves in the hedgerows, the
-increase in plant richness in the networks is likely due to plants
-attracting visitors, as they offer better rewards.
+$\pm$ SE, pollinators: $0.193$ $\pm$ $0.0729$, $p$-value=$0.008$;
+plants: $0.212$ $\pm$ $0.0653$, $p$-value=$0.001$). Unsurprisingly,
+pollinator species are colonizing and persisting at the assembling
+hedgerows. Plant species richness is based on the flowers actually
+visited by pollinators and not the presence of a particular plant
+species at a site. Thus, though some new plant species may establish
+themselves in the hedgerows, the increase in plant richness in the
+networks is likely due to plants attracting visitors, as they offer
+better rewards.
 
@@
 \pm 4*10^{-3}$, $p$-value=$0.987$) or abundance ($0.001 \pm 0.003$,
-$p$-value=$0.65$). % In contrast, the robustness of
-% networks to cascading perturbations, as measured by the algebraic
-% connectivity of the network, increased as the network assembled
-% (Fig.~\ref{fig:rob}, estimate of the slope of robustness to cascading
-% perturbations through time $\pm$ SE, $0.6814 \pm 0.272$,
-% $p$-value=$0.042$).
+$p$-value=$0.65$). In contrast, the sensitivity of networks to
+cascading perturbations, as measured by the algebraic connectivity of
+the network, increased as the network assembled (Fig.~\ref{fig:rob},
+estimate of the slope of sensitivity to cascading perturbations
+through time $\pm$ SE, $0.6814 \pm 0.272$, $p$-value=$0.042$).
 
@@
 points. If these network reorganizations were a product of
-environmental forces alone, we would explect to observe the same
+environmental forces alone, we would expect to observe the same
 changing points at the same periods, consistently across all
@@
 usually have longer phenologies, so they can visit many different
-flowers, resulting in a higher degree \citep{Vazquez2009}. Either way,
-our result suggests that adaptable species can change their network
-position to utilize the most advantageous floral resources available,
-which may depend on the other pollinator species that are present, and
-the state of the plant community \citep{macleod2016measuring,
-  gomez2006ecological, waser1996generalization}. Thus given the
-opportunity and ability to use different resources, species will often
-change their network positions \citep{macleod2016measuring}.
+flowers, resulting in a higher degree \citep{Vazquez2009,
+  fort2016abundance}. Either way, our result suggests that adaptable
+species can change their network position to utilize the most
+advantageous floral resources available, which may depend on the other
+pollinator species that are present, and the state of the plant
+community \citep{macleod2016measuring, gomez2006ecological,
+  waser1996generalization}. Thus given the opportunity and ability to
+use different resources, species will often change their network
+positions \citep{macleod2016measuring}.
 
@@
 
-The changes in network patterns with assembly also increased the
-robustness of the networks to cascading perturbations -- providing
-further evidence that hedgerows are a valuable tool for promoting
-species conservation and ecosystem provision in agricultural areas
-\citep{mgonigle-2015-x, ponisio2015farm, kremen-2015-602}. Because
-network vulnerability to cascading perturbations, as measured by
-algebraic connectivity, is correlated with species richness, the
-increase and plant and pollinator richness following restoration is at
-least partially responsible for enhancing network robustness to
-cascading effects. The relationship between diversity and stability in
-networks has been the subject of considerable debate
-\citep[e.g.,][]{may1972will, pimm1984complexity,
-  montoya2006ecological}. Our results provide one of the few empirical
-examples of how restoring species diversity contributes to enhancing
-network stability. Though these hedgerows were designed to promote
-floral resources across the growing season and not specifically to
-promote any specific network properties \citep{menz-2010-4}, the
-pollinators assembled into resilient interaction patterns. With prior
-knowledge of the floral preferences of pollinators, future restoration
-efforts may also be able to incorporate promoting network stability
-into planning efforts \citep{mgonigle2016tool}
+Interestingly, however, the changes in network patterns with assembly
+the robustness of hedgerow communities to species loss did not change
+with assembly. This is particularly surprising given the observed
+increase in nestedness, which is often associated with an enhanced in
+robustness to extinction \citep{Memmott2004}. Perhaps assembling
+hedgerows have yet to reach sufficient levels of nestedness to realize
+the benefits nestedness confers. Nestedness of the assembling
+hedgerows, however, did not asymptote within the eight years following
+restoration that the sites were surveyed, so hedgerow networks may
+eventually reach sufficient levels of nestedness to gain the
+robustness advantage.
 
-Interestingly, however, the robustness of hedgerow communities to
-species loss did not change with assembly. This is particularly
-surprising given the observed increase in nestedness, which is often
-associated with an enhanced in robustness to extinction
-\citep{Memmott2004}. Perhaps assembling hedgerows have yet to reach
-sufficient levels of nestedness to realize the benefits nestedness
-confers. Nestedness of the assembling hedgerows, however, did not
-asymptote within the eight years following restoration that the sites
-were surveyed, so hedgerow networks may eventually reach sufficient
-levels of nestedness to gain the robustness advantage.
+Contrary to the general restoration goals, the susceptibility of the
+networks to cascading perturbations increased as the communities
+assembled. Because network vulnerability to cascading perturbations,
+as measured by algebraic connectivity, is correlated with species
+richness, the increase and plant and pollinator richness following
+restoration is at least partially responsible for the increase to
+cascading effects. Connectance is also positively related to algebraic
+connectivity \citep{gibert2013spatial}, but because we observed a
+decrease in connectance, topological characteristics of the networks
+beyond species richness and connectance are needed to explain the
+increased sensitivity to perturbations spreading. These hedgerows were
+designed to provide floral resources to the largest number of
+pollinators across the growing season \citep{menz-2010-4}. The
+generalized nature of the floral community may explain why the
+networks tended to be more generalized than expected if interactions
+were randomly distributed across species (Fig.~\ref{fig:baci}). In
+addition, the design of the hedgerow plantings may have facilitated
+the emergence of a single, highly connected module in all of the
+networks (see \ref{fig:changePoints2} for examples). This network
+configuration results in short path lengths (the distance between
+species in a network based on their shared partners), and thus, a
+perturbation in one species can more easily spread to other
+species. In order to promote more resilient communities, future
+restoration efforts should explore designing floral communities to
+promote more interaction partitioning using, for example, algorithms
+to optimize different network properties based on prior knowledge of
+pollinator floral preferences \citep{mgonigle2016tool}, and on desired
+network architectures that renders them more robust both to species
+loss and to cascading effects.
+
+% The relationship between diversity and stability in networks
+% has been the subject of considerable debate
+% \citep[e.g.,][]{may1972will, pimm1984complexity,
+%   montoya2006ecological}. Our results provide one of the few empirical
+% examples of how restoring species diversity contributes to enhancing
+% network stability.
+
+% providing further evidence that hedgerows are a valuable tool for
+% promoting species conservation and ecosystem provision in agricultural
+% areas \citep{mgonigle-2015-x, ponisio2015farm,
+%   kremen-2015-602}.
 
@@
 Foundation (FAPESP, grant 2013/13319-5). We also appreciate the Santa
-Fe Institute for faciliting this international collaboration.
+Fe Institute for facilitating this international collaboration.
 
@@
 \bibliography{refs}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.8\textwidth]{figures/scheme.pdf}
+  \caption{Diagram illustrating the analysis to examine the temporal
+    turnover of interactions weighted based on their similarity. A, B
+    and C are animal species, and Ps are plant species. The dendrogram
+    depicts the interaction similarity across years based on the
+    number of shared constituent species.}
+  \label{fig:methods}
+\end{figure}
+\clearpage
 
@@
     remained consistent across years, while connectance decreased. The
-    left panels represent $z$-scores. Scores greater than $\sim 2$ or
-    less than $\sim -2$ are significantly more or less structured than
-    randomly assembled networks. Points are the metric value for each
-    site at each year of assembly. The solid line indicates the mean
-    slope estimate and the dashed lines are the $95\%$ confidence
-    intervals around the estimate.}
+    nestedness, modularity and specialization scores represent
+    $z$-scores. Scores greater than $\sim 2$ or less than $\sim -2$
+    are significantly more or less structured than randomly assembled
+    networks. Points are the metric value for each site at each year
+    of assembly. The solid line indicates the mean slope estimate and
+    the dashed lines are the $95\%$ confidence intervals around the
+    estimate.}
   \label{fig:baci}
@@
   \caption{The robustness of networks to species extinction did not
-    change with network assembly, but the resilience to to cascading
-    perturbation increased. The robustness to species extinction is
+    change with network assembly, but the sensitivity to cascading
+    perturbations increased. The robustness to species extinction is
     measured by incrementally removing species by degree, though

### 2016-10-28 analysis/networkLevel/baci.R

@@ -37,7 +37,7 @@ summary(baci.mod.mod)
 save(baci.mod.mod, file='saved/mods/baci_mod.Rdata')
 
 ## h2
-baci.h2.mod <- lmer(H2 ~ scale(ypr) +
+baci.h2.mod <- lmer(zH2 ~ scale(ypr) +
                  (1|Site) + (1|Year),
                  data=cor.dats[!is.na(cor.dats$ypr),])
 summary(baci.h2.mod)

### 2016-10-28 analysis/networkLevel/plotting/baci.R

@@ -40,11 +40,11 @@ mod.pi <- predict.int(mod= baci.mod.mod,
 ## ************************************************************
 load(file='saved/mods/baci_h2.Rdata')
 
-dd.h2 <- cbind(dd, H2=0)
+dd.h2 <- cbind(dd, zH2=0)
 
 h2.pi <- predict.int(mod= baci.h2.mod,
                         dd=dd.h2,
-                        y="H2",
+                        y="zH2",
                         family="gaussian")
 
 ## ************************************************************

### 2016-10-28 analysis/networkLevel/plotting/src/plotPanels.R

@@ -30,7 +30,7 @@ plot.panels <- function(){
     ## specialization
     plot.panel(new.dd=h2.pi,
                dats=cor.dats,
-               y1="H2",
+               y1="zH2",
                xs="ypr",
                col.fill=col.fill,
                col.lines=col.lines,

### 2016-10-31 ms/ms.tex
added figures to ms

@@
   networks. Understanding network structure and how it relates to
-  network assembly, therefore, is a priority for system-level
-  conservation biology. Using a long-term data set comprising nearly
-  20,000 pollinator visitation records, we explore the assembly of
+  network assembly, therefore, is a priority for conservation
+  biology. Using a long-term data set comprising nearly $20,000$
+  pollinator visitation records, we explore the assembly of
   plant-pollinator communities at native plant restorations in the
   Central Valley of California. Across years, species are highly
-  dynamic in their network position, causing assembly to be punctuated
-  by significant reorganizations of interactions. In contrast, the
+  dynamic in their network position, causing community assembly to be
+  punctuated by major interaction reorganizations. In contrast, the
   non-assembling networks did not restructure as frequently. Across
-  all communities, Pollinator species were opportunistic in the
-  flowers they visited. Indeed, the most persistent and generalized
-  species were also the most variable in their network positions. High
-  species and interaction turnover was ubiquitous across assembling
-  and non-assembling communities, though unique interactions turnover
+  all communities, pollinator species were opportunistic in the
+  flowers they visited. The most persistent and generalized species
+  were also the most variable in their network positions. High species
+  and interaction turnover was ubiquitous across assembling and
+  non-assembling communities, though unique interactions turned over
   at higher rates in assembling hedgerows as the networks continually
-  reorganize. Nestedness of assembling networks also increased through
-  time, but there was no increase in robustness to simulated plant
-  extinctions. The sensitivity of networks to cascading
+  reorganized. Nestedness of assembling networks also increased
+  through time, but there was no increase in robustness to simulated
+  plant extinctions. The sensitivity of networks to cascading
   perturbations, however, increased as the communities assembled, at
@@
   a valuable tool for promoting species conservation and ecosystem
-  provision in agricultural areas.
+  service? provision in agricultural areas.
 \end{abstract}
@@
   barnosky2011has}. Because mutualistic interactions are essential for
-maintaining the diversity their component guilds, these systems are
+maintaining the diversity of their component guilds, these systems are
 particularly at risk from coextinction cascades. The nature of these
@@
 \citep{Memmott2004, Rezende2007, Bascompte2009, Thebault2010}. To
-safeguard function it has become increasingly imperative to aid the
-recovery of lost biodiversity and interactions through ecological
-restoration, and a key restoration aim is to facilitate assembly of
-robust interaction networks \citep{menz-2010-4}. We know little,
-however, about how to re-assemble interacting communities through
-restoration, or the process of ecological network assembly more
-generally.
+safeguard ecological function, it has become increasingly imperative
+to aid the recovery of lost interactions and component biodviersity
+through ecological restoration, and a key restoration aim is to
+facilitate assembly of robust interaction networks
+\citep{menz-2010-4}. We know little, however, about how to re-assemble
+interacting communities through restoration, or the process of
+ecological network assembly more generally.
 
@@
 a plant-pollinator network within a season, \cite{Olesen2008} found
-that new species tended to interact with already well-connected
-species, potentially because these species are either more abundant or
-more temporally persistent. In addition, using a space-for-time
-substitution to study primary succession along a glacier foreland,
-\cite{albrecht2010plant} also found some evidence that assembly
-occurred through preferential attachment. Network nestedness, a
-pattern of interactions where a core formed by generalist species that
-interact with both specialist and generalist species, increased as the
-community aged \citep{albrecht2010plant}. Thus, an increase in
-nestedness could result from preferential attachment process where
-specialist species attach to the well-connected, generalist
-core. % In addition,
-% non-successional temporal dynamics also suggest a stable core of
-% generalists persists despite high turnover of peripheral species
-% \citep{fang2012relative, diaz2010changes, alarcon2008year}.
+that phenologically new plant and pollinator species tended to
+interact with already well-connected species, potentially because
+these species are either more abundant or more temporally
+persistent. In addition, using a space-for-time substitution to study
+primary succession along a glacier foreland, \cite{albrecht2010plant}
+also found some evidence that assembly occurred through preferential
+attachment. Specifically, network nestedness (i.e, a core group of
+generalists interacts with both specialist and generalist species)
+increased as the community aged \citep{albrecht2010plant}.  An
+increase in nestedness could result from the preferential attachment
+process where specialist species attach to the well-connected,
+generalist core.
 
@@
 of interactions, or network changing points, are observed in social
-networks responding to abrupt shifts in the behavior of interactors
+networks that respond to abrupt shifts in the behavior of interactors
 \citep{peel2014detecting}. In ecological communities, such shifts may
@@
 interaction partners to optimize their foraging effort. In
-plant-pollinator communities, theory predicts pollinators optimize
-their use of floral resources to reduce interspecific competition and
-improve resource-use efficiency \citep{pyke1984optimal,
-  valdovinos2010consequences, valdovinos2013adaptive,
-  albrecht2010plant, Bluthgen2007}. No studies, however, have examined
-whether changing points occur during ecological network assembly, and
-how these changes relate to the species behavior.
+plant-pollinator communities, theory predicts that pollinators
+optimize their use of floral resources to reduce interspecific
+competition and improve resource-use efficiency
+\citep{pyke1984optimal, valdovinos2010consequences,
+  valdovinos2013adaptive, albrecht2010plant, Bluthgen2007}. No
+studies, however, have examined whether network changing points occur
+during ecological network assembly, and how these changes relate to
+the species behavior.
 
@@
 in areas where species interactions provide essential ecosystem
-services, such as crop pollination. In intensively managed
-agricultural landscapes, the demand for pollination services is the
-greatest \citep{kremen-2008-10}. However, honey bees, managed
-extensively around the world to provide crop pollination, are in
-global decline \citep{neumann-2010-1, van-engelsdorp-2009-e6481}. In
-addition, native pollinators, which are capable of providing
-sufficient crop pollination \citep{kremen-2002-16816,
-  winfree-2007-1105, kremen-2004-1109}, are in short supply because
-these landscapes make poor habitats for pollinator populations
-\citep{kremen-2002-16816}. To ensure the continued provision of
-ecosystem services and curb biodiversity loss, it is critical to
+services, such as crop pollination. To ensure the continued provision
+of ecosystem services and curb biodiversity loss, it is critical to
 restore pollinators and their interactions in agricultural
@@
 diversity of pollinators in agricultural landscapes
-\citep{morandin-2013-829, mgonigle-2015-x, kremen-2015-602,
-  ponisio2015farm}. In addition, hedgerows promote the persistence and
-colonization of floral resource specialists
-\citep{mgonigle-2015-x}. It is important to further understand how
+\citep{morandin-2013-829, kremen-2015-602, ponisio2015farm}, and
+promote the persistence and colonization of floral resource
+specialists \citep{mgonigle-2015-x}. It is important to understand how
 these new species are being incorporated into the network as the
-community assembles, or the consequences for interaction patterns and
+community assembles, and the consequences for interaction patterns and
 robustness.
@@
 We explore the process of network development using a nine year
-dataset of plant-pollinator communities assembly following hedgerow
+dataset of plant-pollinator community assembly following hedgerow
 restoration in the highly simplified and intensively managed
@@
 up of interactions as would be predicted by preferential attachment,
-or punctuated by significant reorganizations of interactions (i.e.,
-network changing points). Even with changing points in interaction
-organization, networks could still be assembling via preferential
-attachment if the network reorganizations were primarily driven the by
-peripheral, temporally variable species while a stable, well-connected
-core of species persist. We test whether the species that are most
-variable in their network position --- and thus important contributors
-to network reorganizations --- are less persistent and connected
-species. To further explore the mechanisms underlying the temporal
-dynamics of the networks, we examine patterns in the species and
-interaction temporal turnover. Lastly, we investigate whether networks
-are assembling toward predictable interaction patterns, and the
-ramifications for the robustness of the networks to species extinction
-and cascading perturbations.
+or instead is punctuated by significant reorganizations of
+interactions (i.e., network changing points). Even with changing
+points in interaction organization, networks could still be assembling
+via preferential attachment if the network reorganizations were
+primarily driven by peripheral, temporally variable species while a
+stable, well-connected core of species persist. We test whether the
+species that are most variable in their network position --- and thus
+important contributors to network reorganizations --- are less
+persistent and connected species. To further explore the mechanisms
+underlying the temporal dynamics of the networks, we examine patterns
+in the species and interaction temporal turnover. Lastly, we
+investigate whether networks are assembling toward predictable
+interaction patterns, and the ramifications for the robustness of the
+networks to species extinction and cascading perturbations.
 
@@
 We surveyed plant-pollinator interaction networks of independent
-assembling hedgerows communities (N=5), as well as in two types of
+assembling hedgerows communities (N=5), and of two types of
 non-assembling communities to serve as controls: unrestored, weedy
@@
 of intensively managed agriculture --- primarily monocultures of
-conventional row crops, vineyards and orchards. Hedgerows we planted
-along field margins where they do not remove valuable land from
-production. Plantings are between are ca.~3--6 m wide and
-approximately 350 m long, bordering large (ca.\ 30--hectare) crop
-fields. Hedgerows consist of native, perennial, shrub and tree
-plantings including \textit{Rosa californica}, \textit{Cercis
-  occidentalis}, \textit{Ceanothus spp.}, \textit{Heteromeles
-  arbutifolia}, \textit{Sambucus mexicana}, \textit{Eriogonum spp.},
-\textit{Baccharis spp.}, \textit{Salvia spp}. and others
-\citep[Fig.~S1][]{menz-2010-4, kremen-2015-602, mgonigle-2015-x}. The
-mean distance between monitoring sites was $15$ km, and the minimum
-distance between sites of the same type sampled in the same year was 2
-km.  The entire area surveyed spanned almost $300$ km$^2$. The crop
-fields adjacent to all sites were similarly managed as intensive,
-high-input monoculture.
+conventional row crops, vineyards and orchards. Hedgerows are ca.~3--6
+m wide and approximately 350 m long, bordering large (ca.\
+30--hectare) crop fields. Hedgerows consist of native, perennial,
+shrub and tree plantings including \textit{Rosa californica},
+\textit{Cercis occidentalis}, \textit{Ceanothus spp.},
+\textit{Heteromeles arbutifolia}, \textit{Sambucus mexicana},
+\textit{Eriogonum spp.}, \textit{Baccharis spp.}, \textit{Salvia
+  spp}. and others \citep[Fig.~S1][]{menz-2010-4, kremen-2015-602,
+  mgonigle-2015-x}. The mean distance between monitoring sites was
+$15$ km, and the minimum distance between sites sampled in the same
+year was 1 km.  The entire area surveyed spanned almost $300$
+km$^2$. The crop fields adjacent to all sites were similarly managed
+as intensive, high-input monoculture.
 
-Monitoring began in 2006 and continued through 2014. Surveys of the
-assembling hedgerows began the year before the area was restored. For
-logistical reasons, no sampling of assembling hedgerows was conducted
-in 2010. Sites were sampled between two and five times per year
-(Tables S1-S3). In each round of sampling, the order in which sites
-were sampled was randomized. Surveys were conducted under sunny
-conditions when the temperature was above $21^{\circ}\mathrm{C}$ and
-wind speed was below $2.5$ meters/second.
+Monitoring of assembling hedgerows began in 2006 and continued through
+2014. Surveys of these sites began the year before the area was
+restored. For logistical reasons, no sampling of assembling hedgerows
+was conducted in 2010. Sites were sampled between two and five times
+per year (Tables S1-S3, mean $3.4$ samples per year). In each round of
+sampling, the order in which sites were sampled was
+randomized. Surveys were conducted under sunny conditions when the
+temperature was above $21^{\circ}\mathrm{C}$ and wind speed was below
+$2.5$ meters/second.
 
@@
 connect is given by the parameter $p_r$. The probability distribution
-of the network $G$ thus defined as:
+of the network $G$ is thus defined as:
 
@@
 a change point occurred between two time slices. To detect a change
-point, we compare the fit of two models --- one where a change point
-occurred between two networks, and one where no change occurred ---
-using posterior Bayes factors. We chose a sliding window of length,
-$w$, of four, within which to find change points. Larger windows allow
-for more gradual changes, and four was the maximum possible with our
-maximum of eight years of data. Lastly, to calculate a $p$-value for
-the Bayes factors, we use parametric bootstrapping to numerically
-estimate the null distribution \citep{peel2014detecting}. We employed
-code published online by L.~Peel for the change point
-analysis. Analyses were conducted in Python 3.4.
+point, we use Bayes factors to compare the fit of two models --- one
+where a change point occurred between two networks, and one where no
+change occurred. We chose a sliding window of length, $w$, of four,
+within which to find change points. Larger windows allow for more
+gradual changes, and four was the maximum possible with our eight
+years of data. Lastly, to calculate a $p$-value for the Bayes factors,
+we use parametric bootstrapping to numerically estimate the null
+distribution \citep{peel2014detecting}. We employed code published
+online by L.~Peel for the change point analysis. Analyses were
+conducted in Python 3.4.
 
@@
 communities. We model the number of change points as successes and the
-total number of years each site was sampled as trails, and use a
+total number of years each site was sampled as trials, and use a
 generalized linear model with Binomial error to test whether the
@@
 survey years were included in this analysis (N=11). All statistical
-analysis were conducted in R 3.2.3 \citep{R}.
+analyses were conducted in R 3.2.3 \citep{R}.
 
@@
 weighted closeness centrality \citep{freeman1978centrality} at each
-site through time. Closeness centrality describes the centrality of a
-species in the network by calculating path lengths to other vertices
-(species) in the graph \citep{freeman1978centrality}. The shorter the
-mean path length to other species, the higher is the closeness
-centrality. We use linear mixed models to test whether the species
-closeness variability (log) is related to the persistence or degree of
-that species \citep{lme4, lmetest}. We included random effects for
-species, as well as site. Because the degree and persistence of
-pollinators were strongly correlated, ($\rho = 0.84$, $p$-value $<$
-$2*10^{-16}$), we include each explanatory variable in the model
-separately. Plant degree and persistence were not significantly
-correlated, but we use the same models as we did for the pollinators
-for consistency. An approximately logarithmic increase in closeness
-centrality, as would be expected with assembly by preferential
-attachment, would also lead to a high variability in closeness scores,
-so we test whether log closeness centrality increases through time.
+site through time. Closeness centrality represents the importance of a
+space by calcuating the path lengths to other vertices (species) in
+the network \citep{freeman1978centrality}. The shorter the mean path
+length to other species, the higher is the closeness centrality. We
+use linear mixed models to test whether the species closeness
+variability (log) is related to the persistence or degree of that
+species \citep{lme4, lmetest}. We included random effects for species
+and site. Because the degree and persistence of pollinators were
+strongly correlated, ($\rho = 0.84$, $p$-value $<$ $2*10^{-16}$), we
+include each explanatory variable in separate models. Plant degree and
+persistence were not significantly correlated, but we use the same
+models as we did for the pollinators for consistency.  Because an
+approximately logarithmic increase in closeness centrality, as would
+be expected with assembly by preferential attachment, would also lead
+to high variability in closeness scores, we also test whether log
+closeness centrality increases through time.
 
@@
 communities because new pollinator species are colonizing and
-establishing themselves \citep{mgonigle-2015-x}. Similarly, if
-pollinators try to maximize their foraging efficiency based on the
-species present, interactions may turnover more quickly than in
-established communities. In addition, at assembling hedgerows,
-unvisited plants may appear to ``colonize'' the networks as they
-establish new interactions with pollinators.
+establishing themselves \citep{mgonigle-2015-x}. Similarly, because
+species are turning over and pollinators are trying to maximize their
+foraging efficiency based on the species present, interactions may
+turnover more quickly than in established communities. In addition, at
+assembling hedgerows, plants that are unvisited in early years may
+appear to ``colonize'' the networks as they became more attractive
+resources and establish new interactions with pollinators.
 
@@
 calculating the variation in community composition across sites within
-a year, however, we estimated turnover across years at a site. We
-first calculated the pairwise dissimilarity of plants, pollinators and
+a year, we estimated turnover across years at a site. We first
+calculated the pairwise dissimilarity of plants, pollinators and
 interactions between years within each site using the Chao
@@
 richness to estimate the deviation of the observed dissimilarity from
-that, which would be expected under a random community assembly
+that which would be expected under a random community assembly
 process. With the corrected dissimilarity values, we then calculated
@@
 non-assembling communities, the species and interaction temporal
-turnover estimates were included as response variables in a linear
-mixed model with site type as an explanatory variable along with site
-as a random effect \citep{lme4, lmetest}.
+turnover estimates were modeled as responses in a linear mixed model
+with site type as an explanatory variable and site as a random effect
+\citep{lme4, lmetest}.
 
@@
 turnover of interactions with weightings based on their similarity. We
-followed \cite{ahn2010link} algorithm to hierarchically cluster all
-the interactions (edges) across sites and years based on their
-similarity, and build a dendrogram. The interaction similarity is
-based how may plants and pollinators (vertices) two edges share
+followed the algorithm of \cite{ahn2010link} to cluster all the
+interactions (edges) hierarchically across sites and years based on
+their similarity, and build a dendrogram. The interaction similarity
+is based how may plants and pollinators (vertices) two edges share
 \citep{ahn2010link, kalinka2011linkcomm}. The more species edges
@@
 weighted by their similarity, as approximated by ``phylogenetic''
-distance \citep{graham2008phylogenetic, picante-2010-1463}. When then
+distance \citep{graham2008phylogenetic, picante-2010-1463}. We then
 use linear models to test whether the weighted turnover of
@@
 To test whether network modularity, nestedness, connectance or
-specialization changed linearly with assembly, we use linear mixed
+specialization changed linearly with assembly, we used linear mixed
 models with the descriptive network metrics as the response variable,
 year of assembly as the explanatory variable, and random effects of
-site and year. The number of species in a network affects affects the
-patterns of interaction possible, so we also examined the change in
-plant and pollinator species richness through time. We employ
-generalized linear mixed models with Poisson error to model
-richness. We scaled explanatory variables.
+site and year. The number of species in a network affects the patterns
+of interaction possible, so we also examined the change in plant and
+pollinator species richness through time. We employ generalized linear
+mixed models with Poisson error to model richness. We scaled
+explanatory variables.
 
 \subsection*{Network robustness}
-Lastly, we test whether the changes in interaction patterns associated
-with network assembly affect the robustness of the network to species
-loss and cascading perturbations. Following \cite{Memmott2004}, we
-simulate plant species extinction and the subsequent extinction
-cascades of pollinator species. Because the reproduction of plant
-species if facilitated by active restoration efforts, it is unlikely
-the extinction of pollinator species would affect plant populations in
-the hedgerows. However, plants ceasing to bloom, for example in
-response to drought, will likely affect the pollinators that depend on
-them. We eliminate plants species based on their degree or abundance,
-and then calculate the number of pollinators that secondarily went
-extinct. The area below the extinction curve is an estimate of network
-robustness \citep{Memmott2004, bipartite}.
+Lastly, we tested whether the changes in interaction patterns
+associated with network assembly affect the robustness of the network
+to species loss and cascading perturbations. Following
+\cite{Memmott2004}, we simulated plant species extinction and the
+subsequent extinction cascades of pollinator species. Because the
+reproduction of plant species is facilitated by active restoration
+efforts, it is unlikely the extinction of pollinator species would
+affect plant populations in the hedgerows. However, plants ceasing to
+bloom, for example in response to drought, will likely affect the
+pollinators that depend on them. We eliminated plants species based on
+their degree or abundance, and then calculated the number of
+pollinators that secondarily went extinct. The area below the
+extinction curve is an estimate of network robustness
+\citep{Memmott2004, bipartite}.
 
 We also explored how the robustness to cascading perturbations changed
-as community assembled, using algebraic connectivity --- the second
-smallest eigenvalue of the Laplacian matrix
+as the community assembled, using algebraic connectivity --- the
+second smallest eigenvalue of the Laplacian matrix
 \citep{fiedler1973algebraic} --- as a proxy for network robustness
-(Gaiarsa et al., submitted). Algebraic connectivity relates to how
+(Gaiarsa et al., in prep). Algebraic connectivity relates to how
 difficult it is to turn a network into completely disconnected groups
-of species \citep{costa2007characterization}, and has implications to
-how integrated is the dynamics of the elements of the networks
-\citep{gibert2013spatial}. The larger the algebraic connectivity, the
-more robust a network is to cascading perturbations, and the easier it
-is to break the community into isolated groups of species (Gaiarsa et
-al., submitted).
+of species \citep{costa2007characterization, gibert2013spatial}. The
+larger the algebraic connectivity, the more sensitive a network is to
+cascading perturbations. Perturbations can have negative consequences
+for the species in the network such as the decrease in abundance of a
+plant or pollinator. For example, a pollinator could decrease in
+abundance, and the pollination services it provided to the plants it
+visited would be diminished. The effected plants would set less seeds,
+and decrease in abundance the subsequent year. The pollinators that
+depended on those plant species would suffer, and the cascade would
+continue to propagate throughout the network. A perturbation could
+have a positive effect, like the increase in abundance of a
+species. Similar to the negative perturbation, the impact of the
+positive perturbation would cascade through the network. The examples
+of negative perturbations (e.g., resource collapse, disease,
+parasites), however, outnumber possible positive perturbations. Both
+of these metrics of network resilience are static --- they do not
+account for the ability of species to alter their interaction
+depending on circumstances.
 
@@
 punctuated by more interaction reorganizations than would be expected
-by environmental shifts alone that would effect assembling and
-non-assembling networks equally.
+by environmental shifts alone that would effect all networks
+similarly.
 
@@
 pollinator species are colonizing and persisting at the assembling
-hedgerows. Plant species richness is based on the flowers actually
-visited by pollinators and not the presence of a particular plant
-species at a site. Thus, though some new plant species may establish
-themselves in the hedgerows, the increase in plant richness in the
-networks is likely due to plants attracting visitors, as they offer
-better rewards.
+hedgerows. Plant species richness in the networks is based on the
+flowers actually visited by pollinators and not the presence of a
+particular plant species at a site. Thus, though some new plant
+species may establish themselves in the hedgerows, the increase in
+plant richness in the networks is likely due to previously unvisited
+plants attracting visitors as they mature and offer better rewards.
 
@@
 Assembly did not effect the robustness of the networks to species
-extinction when species where removed incrementally by degree
-(estimate of the slope of robustness through time $\pm$ SE, $6*10^{-5}
-\pm 4*10^{-3}$, $p$-value=$0.987$) or abundance ($0.001 \pm 0.003$,
+extinction when species were removed incrementally by degree (estimate
+of the slope of robustness through time $\pm$ SE, $6*10^{-5} \pm
+4*10^{-3}$, $p$-value=$0.987$) or abundance ($0.001 \pm 0.003$,
 $p$-value=$0.65$). In contrast, the sensitivity of networks to
@@
 assembling hedgerows were not sampled in 2010, so disentangling
-whether the changing points are due to skipping a year of assembly or
-the drought is not possible. Interestingly, most assembling hedgerows
-did not undergo a significant interaction reorganization immediately
-after a hedgerow was planted (i.e., the transition from weedy field
-margin to hedgerow). This result is consistent with the finding that
-in our study system, hedgerow restoration takes several years to have
-an impact on the plant-pollinator communities (Kremen and M'Gonigle,
-in prep).
+whether the changing points are due to skipping a year of monitoring
+the assembly process or the drought, is not possible. Interestingly,
+most assembling hedgerows did not undergo a significant interaction
+reorganization immediately after a hedgerow was planted (i.e., the
+transition from weedy field margin to hedgerow). This result is
+consistent with the finding that in our study system, hedgerow
+restoration takes several years to have an impact on the
+plant-pollinator communities, and with the observation that hedgerows
+do not begin to produce many flowers until $3$--$5$ years following
+planting (Kremen and M'Gonigle, in prep).
 
@@
 during assembly, the way in which these reorganizations occur was
-different from what is expected by preferential attachment. Based on a
-preferential attachment process, we expect that the most persistent
-and high degree species would remain stable in the network core during
-assembly \citep{barabasi1999emergence}. Surprisingly, however, we
-encountered the opposite pattern. For example, the four most
-ubiquitous species in our study landscape --- \textit{Halictus
-  ligatus}, \textit{Halictus tripartitus}, \textit{Lasioglossum
-  (Dialictus) incompletum}, and \textit{Toxomerus marginatus} --- were
-the only species that changed which module they were a member across
-years in all the assembling hedgerows. Because species degree and
-persistence were strongly correlated, it is difficult to disentangle
-the causal mechanism for why species with those characteristics are so
-variable in their network position. Generalized species may be able to
-better exploit the limited floral resources in the intensively managed
-agriculture landscape, and thus also be the most persistent \citep[in
-ant-plant mutualisms, ][]{diaz2010changes}. More persistent species
-usually have longer phenologies, so they can visit many different
-flowers, resulting in a higher degree \citep{Vazquez2009,
+different from what would be expected from preferential
+attachment. With a preferential attachment process, we expect that the
+most persistent and high degree species would remain stable in the
+network core during assembly
+\citep{barabasi1999emergence}. Surprisingly, however, we encountered
+the opposite pattern. For example, the four most ubiquitous species in
+our study landscape --- \textit{Halictus ligatus}, \textit{Halictus
+  tripartitus}, \textit{Lasioglossum (Dialictus) incompletum}, and
+\textit{Toxomerus marginatus} --- were the only species that changed
+which module they were a member in across years in all the assembling
+hedgerows. Because species degree and persistence were strongly
+correlated, it is difficult to disentangle the causal mechanism for
+why species with those characteristics are so variable in their
+network position. Generalized species may be better able to exploit
+the limited floral resources in the intensively managed agriculture
+landscape, and thus also be the most persistent \citep[in ant-plant
+mutualisms, ][]{diaz2010changes}. More persistent species usually have
+longer phenologies, so they can visit many different flowers,
+resulting in a higher degree \citep{Vazquez2009,
   fort2016abundance}. Either way, our result suggests that adaptable
 species can change their network position to utilize the most
-advantageous floral resources available, which may depend on the other
-pollinator species that are present, and the state of the plant
+advantageous floral resources available, which may depend on both the
+other pollinator species that are present and the state of the plant
 community \citep{macleod2016measuring, gomez2006ecological,
@@
 weighted interaction turnover. Mature hedgerows may thus both support
-more stable pollination communities and likely ecosystem services,
-while also providing resources for rare and/or specialized species
-\citep{kremen-2015-602, mgonigle-2015-x}.
+more stable pollinator communities, while also providing resources for
+rare and/or specialized species \citep{kremen-2015-602,
+  mgonigle-2015-x}.
 
-When we explore the how the network-level interaction patterns changed
+When we explore the how network-level interaction patterns changed
 through time, we found that nestedness did increase as the community
@@
 
-Interestingly, however, the changes in network patterns with assembly
-the robustness of hedgerow communities to species loss did not change
-with assembly. This is particularly surprising given the observed
+Interestingly, however, the changes in network patterns associated
+with assembly did not effect the robustness of hedgerow communities to
+species loss. This is particularly surprising given the observed
 increase in nestedness, which is often associated with an enhanced in
@@
 as measured by algebraic connectivity, is correlated with species
-richness, the increase and plant and pollinator richness following
-restoration is at least partially responsible for the increase to
-cascading effects. Connectance is also positively related to algebraic
-connectivity \citep{gibert2013spatial}, but because we observed a
-decrease in connectance, topological characteristics of the networks
-beyond species richness and connectance are needed to explain the
-increased sensitivity to perturbations spreading. These hedgerows were
-designed to provide floral resources to the largest number of
+richness, the increase in and plant and pollinator richness following
+restoration is at least partially responsible for the increase in
+response to cascading effects. Connectance is also positively related
+to algebraic connectivity \citep{gibert2013spatial}, but because we
+observed a decrease in connectance, topological characteristics of the
+networks beyond species richness and connectance are needed to explain
+the increased sensitivity to perturbations spreading. These hedgerows
+were designed to provide floral resources to the largest number of
 pollinators across the growing season \citep{menz-2010-4}. The
@@
   \includegraphics[width=.9\textwidth]{../analysis/changePoint/plotting/communitiesSkanky2.pdf}
-  \caption{The community (i.e., module or compartment) membership
-    between network changing points. Two representative assembling
-    hedgerows are depicted. In the top panel, species are grouped by
-    community. The bottom panels visualize the flow of species between
-    communities between changing points. }
+  \caption{The module membership between network changing points. Two
+    representative assembling hedgerows are depicted. In the top
+    panel, species are grouped by community. The bottom panels
+    visualize the flow of species between communities between changing
+    points. }
   \label{fig:changePoints2}

### 2016-11-01 ms/ms.tex
made claire's edits

@@
 slope of closeness centrality variability and degree, $0.008 \pm
-0.002$, $p$-value=$0.002$). In addition, plant persistence and degree
-were not significantly related to network position variability
-(Fig.~\ref{fig:cv}, estimate of the slope of closeness variability and
-persistence $\pm$ SE, $-2.063 \pm 3.091$, $p$-value=$0.5$; slope of
-closeness variability and degree, $0.0018 \pm 0.002$,
-$p$-value=$0.3$). The variability of species network position was not
-the result of closeness linearly increasing through time, and, in
-fact, plant and pollinator closeness decreased slightly through time
-(Fig.~S2, estimate of the slope of closeness through time $\pm$ SE,
-pollinators: $-0.0003 \pm 0.00005$, $p$-value=$2.7*10^{-12}$; plants
-$-0.007 \pm 0.001$, $p$-value=$1.4*10^{-6}$). Through statistically
-significant, the slopes are so slight they may not be biologically
-significant.
+0.002$, $p$-value=$0.002$). The slope of these relationships remained
+significant when the species with the top $10$ persistence and degree
+scores were dropped. In addition, plant persistence and degree were
+not significantly related to network position variability
+(Fig.~\ref{fig:cv}). % estimate of the slope of closeness variability and
+% persistence $\pm$ SE, $-2.063 \pm 3.091$, $p$-value=$0.5$; slope of
+% closeness variability and degree, $0.0018 \pm 0.002$,
+% $p$-value=$0.3$).
+The variability of species network position was not the result of
+closeness linearly increasing through time, and, in fact, plant and
+pollinator closeness decreased slightly through time (Fig.~S2,
+estimate of the slope of closeness through time $\pm$ SE, pollinators:
+$-0.0003 \pm 0.00005$, $p$-value=$2.7*10^{-12}$; plants $-0.007 \pm
+0.001$, $p$-value=$1.4*10^{-6}$). Through statistically significant,
+the slopes are so slight they may not be biologically significant.
 
@@
 Fig.~\ref{fig:baci}). Modularity decreased (Fig.~\ref{fig:baci}),
-though the slope was not significantly different from zero (estimate
-of the slope of modularity through time $\pm$ standard error of the
-estimate, $-0.524$ $\pm$ $0.295$, $p$-value=$0.124$). In addition,
-none of the networks were significantly modular ($z$-scores $<$ 2,
-Fig.~\ref{fig:baci}). Connectance decreased as the community assembled
-(Fig.~\ref{fig:baci}, estimate of the slope of connectance through
-time $\pm$ standard error of the estimate, $-0.0434$ $\pm$ $0.0152$,
-$p$-value=$0.03$). Specialization also decreased, though the slope was
-only marginally significantly different from zero (estimate of the
-slope of specialization through time $\pm$ SE, $-0.926$ $\pm$ $0.450$,
-$p$-value=$0.078$). Most communities were more generalized than
-expected when interactions were randomized (Fig.~\ref{fig:baci}).
+though the slope was not significantly different from
+zero. % (estimate
+% of the slope of modularity through time $\pm$ standard error of the
+% estimate, $-0.524$ $\pm$ $0.295$, $p$-value=$0.124$).
+In addition, none of the networks were significantly modular
+($z$-scores $<$ 2, Fig.~\ref{fig:baci}). Connectance decreased as the
+community assembled (Fig.~\ref{fig:baci}, estimate of the slope of
+connectance through time $\pm$ standard error of the estimate,
+$-0.0434$ $\pm$ $0.0152$, $p$-value=$0.03$). Specialization also
+decreased, though the slope was only marginally significantly
+different from zero (estimate of the slope of specialization through
+time $\pm$ SE, $-0.926$ $\pm$ $0.450$, $p$-value=$0.078$). Most
+communities were more generalized than expected when interactions were
+randomized (Fig.~\ref{fig:baci}).
 
@@
 Assembly did not effect the robustness of the networks to species
-extinction when species were removed incrementally by degree (estimate
-of the slope of robustness through time $\pm$ SE, $6*10^{-5} \pm
-4*10^{-3}$, $p$-value=$0.987$) or abundance ($0.001 \pm 0.003$,
-$p$-value=$0.65$). In contrast, the sensitivity of networks to
+extinction when species were removed incrementally by
+degree % (estimate
+% of the slope of robustness through time $\pm$ SE, $6*10^{-5} \pm
+% 4*10^{-3}$, $p$-value=$0.987$)
+or abundance. % ($0.001 \pm 0.003$,
+% $p$-value=$0.65$)
+In contrast, the sensitivity of networks to
 cascading perturbations, as measured by the algebraic connectivity of

### 2016-11-01 analysis/networkLevel/plotting/baci.R

@@ -1,6 +1,7 @@
 library(RColorBrewer)
 library(lme4)
 library(lmerTest)
+setwd('~/Dropbox/hedgerow_assembly/analysis/networkLevel')
 source('plotting/src/predictIntervals.R')
 source('plotting/src/CIplotting.R')
 source('plotting/src/plotPanels.R')

### 2016-11-01 analysis/networkLevel/plotting/src/CIplotting.R

@@ -4,6 +4,7 @@ plot.panel <- function(dats,
                        xs,
                        col.lines,
                        col.fill,
+                       col.points=col.fill,
                        plotPoints=TRUE,
                        agg.col="Site",...){
   plotting.loop <- function(){
@@ -15,7 +16,7 @@ plot.panel <- function(dats,
       points(x=jitter(ys$x, factor=0.25),
              y=ys$y,
              pch=16,
-             col=col.fill,
+             col=col.points,
              cex=1.2)
     }
     lines(x=new.dd[,xs],

### 2016-11-01 analysis/networkLevel/plotting/src/plotPanels.R

@@ -1,5 +1,6 @@
 plot.panels <- function(){
   f <- function(){
+    col.white <- add.alpha("white", alpha=0)
     col.lines <- brewer.pal(4, "Greys")[3]
     col.fill <- add.alpha(col.lines, alpha=0.2)
     layout(matrix(1:6, ncol=3))
@@ -19,8 +20,9 @@ plot.panels <- function(){
                dats=cor.dats,
                y1="zmod.met.D",
                xs="ypr",
-               col.fill=col.fill,
-               col.lines=col.lines,
+               col.fill=col.white,
+               col.lines=col.white,
+               col.points=col.fill,
                plotPoints=TRUE)
     mtext("Modularity", 2, line=4, cex=1.5)
 

### 2016-11-01 analysis/networkLevel/plotting/src/plotPanels_resilence.R

@@ -1,20 +1,22 @@
 plot.panels <- function(){
   f <- function(){
+    col.white <- add.alpha("white", alpha=0)
     col.lines <- brewer.pal(4, "Greys")[3]
     col.fill <- add.alpha(col.lines, alpha=0.2)
     layout(matrix(1:2, ncol=1))
     par(oma=c(4, 1, 2, 1),
         mar=c(1, 6, 0.5, 1), cex.axis=1.5)
-    ## nodf
+    ## species extinction
     plot.panel(new.dd=ypr.pi,
                dats=res,
                y1="Robustness",
                xs="ypr",
-               col.fill=col.fill,
-               col.lines=col.lines,
+               col.fill=col.white,
+               col.lines=col.white,
+               col.points=col.fill,
                plotPoints=TRUE)
     mtext("Robustness to \n species extinction", 2, line=4, cex=1.5)
-    ## modularity
+    ## sensitive to perturbations
     plot.panel(new.dd=ypr.pi.alg,
                dats=all.alg.Con.status,
                y1="AlgCon",

### 2016-11-01 analysis/variability/cv.R

@@ -48,6 +48,7 @@ occ.closeness.cv <- cv.trait(spec,
                              na.rm=TRUE)
 summary(occ.closeness.cv$lm.nss)
 
+
 ## plants and closeness
 ## not sig
 plants.occ.closeness.cv <- cv.trait(spec,

### 2016-11-01 analysis/variability/plotting/src/CIplotting.R

@@ -6,6 +6,7 @@ plot.panel <- function(dats,
                        treatments,
                        col.lines,
                        col.fill,
+                       col.points=col.fill,
                        ylabel="",
                        ag.col="SiteStatus",
                        plot.x=TRUE,
@@ -26,7 +27,7 @@ plot.panel <- function(dats,
       ## plots means
       points(x=jitter(ys$x, factor=0.25),
              y=ys$y,
-             col=col.fill[treatments[i]],
+             col=col.points[treatments[i]],
              cex=1,
              pch=pchs[i])
       ## points(x=jitter(sub.dats[,xs], factor=0.25),

### 2016-11-01 ms/ms.tex
Marilia did some editing on the ms

@@
   flowers they visited. The most persistent and generalized species
-  were also the most variable in their network positions. High species
+  were also the most variable in their network positions, contrary to what is expected through prefereantial attachment theory. High species
   and interaction turnover was ubiquitous across assembling and
@@
 larger the algebraic connectivity, the more sensitive a network is to
-cascading perturbations. Perturbations can have negative consequences
-for the species in the network such as the decrease in abundance of a
-plant or pollinator. For example, a pollinator could decrease in
-abundance, and the pollination services it provided to the plants it
-visited would be diminished. The effected plants would set less seeds,
-and decrease in abundance the subsequent year. The pollinators that
-depended on those plant species would suffer, and the cascade would
-continue to propagate throughout the network. A perturbation could
-have a positive effect, like the increase in abundance of a
-species. Similar to the negative perturbation, the impact of the
-positive perturbation would cascade through the network. The examples
-of negative perturbations (e.g., resource collapse, disease,
-parasites), however, outnumber possible positive perturbations. Both
-of these metrics of network resilience are static --- they do not
+cascading perturbations. Perturbations, such as the decrease in abundance of a
+plant or pollinator, can have negative consequences
+for the species in the network. For example, pollination service provided to planst could decrease following a decrease in
+abundance of a pollinator. The affected plants would set less seeds,
+and decrease in abundance the subsequent year. Consequently, other pollinators that
+depended on those plant species would also be affected, and the effects of this perturbation would
+continue to propagate throughout the network, potentially affecting all species in the community. Alternatively, perturbations could
+also have a positive effect. For example, the increase in abundance of a
+plant species would lead to an increase in resource availability for the pollinators. The examples
+of negative perturbations (e.g., resource collapse, disease spreading,
+parasites), however, outnumber possible positive perturbations. It is important to point out that both robustness and algebraic connectivity assume that the network is static --- these metrics do not
 account for the ability of species to alter their interaction
-depending on circumstances.
+depending on circumstances and the resource availability.
 
@@
 whether the changing points are due to skipping a year of monitoring
-the assembly process or the drought, is not possible. Interestingly,
+the assembly process or the drought is not possible. Interestingly,
 most assembling hedgerows did not undergo a significant interaction

### 2016-11-01 ms/ms.tex
updated network figure

@@
 
-\def\title{The temporal assembly of plant-pollinator networks
-  following restoration}
-\def\author{Lauren C.\ Ponisio$^{1,2}$, Marilia P. Gaiarsa$^3$, Claire
-  Kremen$^1$}
+\def\title{Major interaction reorganizations punctuate the assembly of
+  pollination networks} \def\author{Lauren C.\ Ponisio$^{1,2,3}$,
+  Marilia P. Gaiarsa$^4$, Claire Kremen$^1$}
 
@@
     94720\\
+  \item Berkeley Institute for Data Science (BIDS)\\
+    University of California, Berkeley\\
+    190 Doe Library \\
+    Berkeley, California, USA\\
   \item Department of Entomology\\
@@
   network assembly, therefore, is a priority for conservation
-  biology. Using a long-term data set comprising nearly $20,000$
+  biology. Using a nine-year-dataset comprising nearly $20,000$
   pollinator visitation records, we explore the assembly of
@@
   flowers they visited. The most persistent and generalized species
-  were also the most variable in their network positions, contrary to what is expected through prefereantial attachment theory. High species
-  and interaction turnover was ubiquitous across assembling and
-  non-assembling communities, though unique interactions turned over
-  at higher rates in assembling hedgerows as the networks continually
-  reorganized. Nestedness of assembling networks also increased
-  through time, but there was no increase in robustness to simulated
-  plant extinctions. The sensitivity of networks to cascading
-  perturbations, however, increased as the communities assembled, at
-  least partially due to accumulating species richness. We elucidate
-  some of the mechanisms underlying plant-pollinator network assembly
-  and restoration, while providing further evidence that hedgerows are
-  a valuable tool for promoting species conservation and ecosystem
-  service? provision in agricultural areas.
+  were also the most variable in their network positions, contrary to
+  what is expected through prefereantial attachment theory. High
+  species and interaction turnover was ubiquitous across assembling
+  and non-assembling communities, though unique interactions turned
+  over at higher rates in assembling hedgerows as the networks
+  continually reorganized. Nestedness of assembling networks also
+  increased through time, but there was no increase in robustness to
+  simulated plant extinctions. The sensitivity of networks to
+  cascading perturbations, however, increased as the communities
+  assembled, at least partially due to accumulating species
+  richness. We elucidate some of the mechanisms underlying
+  plant-pollinator network assembly and restoration, while providing
+  further evidence that hedgerows are a valuable tool for promoting
+  species conservation and ecosystem service? provision in
+  agricultural areas.
 \end{abstract}
@@
   \centering
-  \includegraphics[width=.8\textwidth]{../analysis/changePoint/plotting/networks.pdf}
+  \includegraphics[width=1\textwidth]{../analysis/changePoint/plotting/networksv1.pdf}
   \caption{Assembling hedgerow networks had more changing points
@@
     their degree. Each species has a consistent position in the
-    network across years. In the assembling hedgerows, colored squares
-    in the corner of each network represent the year of
-    assembly. Before hedgerows are installed (when they are still
-    field margins) the year of assembly is zero.}
+    perimeter of the network across years. In the assembling
+    hedgerows, colored squares in the corner of each network represent
+    the year of assembly. Asterisks indicate the year the hegdgerow
+    was planted. Before that, the sites were weedy field margins.}
   \label{fig:changePoints}
@@
   \centering
-  \includegraphics[width=.9\textwidth]{../analysis/changePoint/plotting/communitiesSkanky2.pdf}
+  \includegraphics[width=.95\textwidth]{../analysis/changePoint/plotting/communitiesSkanky2.pdf}
   \caption{The module membership between network changing points. Two

### 2016-11-01 ms/ms.tex
made the .bib changes

@@
 community \citep{macleod2016measuring, gomez2006ecological,
-  waser1996generalization}. Thus given the opportunity and ability to
+  Waser1996}. Thus given the opportunity and ability to
 use different resources, species will often change their network
@@
 
+\clearpage
 \begin{figure}

### 2016-11-01 ms/ms.tex
updated box plot figures

@@
   \item Departamento de Ecologia\\
-    Universidade de Sao Paulo\\
+    Universidade de S{\~a}o Paulo\\
     S{\~a}o Paulo, SP, Brazil\\
@@
 are weighted by the similarity of their constituent species
-(Fig.~\ref{fig:methods}). We develop a method to examine the temporal
+(Fig.~ S2). We develop a method to examine the temporal
 turnover of interactions with weightings based on their similarity. We
@@
 larger the algebraic connectivity, the more sensitive a network is to
-cascading perturbations. Perturbations, such as the decrease in abundance of a
-plant or pollinator, can have negative consequences
-for the species in the network. For example, pollination service provided to planst could decrease following a decrease in
-abundance of a pollinator. The affected plants would set less seeds,
-and decrease in abundance the subsequent year. Consequently, other pollinators that
-depended on those plant species would also be affected, and the effects of this perturbation would
-continue to propagate throughout the network, potentially affecting all species in the community. Alternatively, perturbations could
-also have a positive effect. For example, the increase in abundance of a
-plant species would lead to an increase in resource availability for the pollinators. The examples
-of negative perturbations (e.g., resource collapse, disease spreading,
-parasites), however, outnumber possible positive perturbations. It is important to point out that both robustness and algebraic connectivity assume that the network is static --- these metrics do not
-account for the ability of species to alter their interaction
+cascading perturbations. Perturbations, such as the decrease in
+abundance of a plant or pollinator, can have negative consequences for
+the species in the network. For example, a decrease in abundance of a
+pollinator will diminish the pollination services it provides to
+plants. The affected plants would set less seeds, and decrease in
+abundance the subsequent year. Consequently, other pollinators that
+depended on those plant species would also be affected, and the
+effects of this perturbation would continue to propagate throughout
+the network. Alternatively, perturbations could also have a positive
+effect. For example, the increase in abundance of a plant species
+would lead to an increase in resource availability for the
+pollinators. The examples of negative perturbations (e.g., resource
+collapse, disease spreading, parasites), however, outnumber possible
+positive perturbations. It is important to note that both robustness
+and algebraic connectivity assume that the network is static. They do
+not account for the ability of species to alter their interaction
 depending on circumstances and the resource availability.
@@
 
-\clearpage
-\begin{figure}
-  \centering
-  \includegraphics[width=.8\textwidth]{figures/scheme.pdf}
-  \caption{Diagram illustrating the analysis to examine the temporal
-    turnover of interactions weighted based on their similarity. A, B
-    and C are animal species, and Ps are plant species. The dendrogram
-    depicts the interaction similarity across years based on the
-    number of shared constituent species.}
-  \label{fig:methods}
-\end{figure}
-\clearpage
-
 \begin{figure}
@@
     representative assembling hedgerows are depicted. In the top
-    panel, species are grouped by community. The bottom panels
-    visualize the flow of species between communities between changing
-    points. }
+    panel, species are grouped by module. The bottom panels visualize
+    the flow of species between modules between changing points. }
   \label{fig:changePoints2}

### 2016-11-01 analysis/variability/src/plotDists.R

@@ -144,9 +144,10 @@ plotDistPanels <- function(){
             xlab='', ylab='', ylim=c(0,1),
             names=c("","",""),
             las=1, xaxt="n")
-    beeswarm(dists ~ status, col = cols, add = TRUE)
+    beeswarm(dists ~ status, col = cols.fills,
+             add = TRUE, method="hex", corral="gutter")
     mtext("Pollinators",
-          3, line=0.9, cex=1.3)
+          3, line=0.9, cex=1.5)
 
     ## plant species turnover
     load(file= file.path('saved/speciesTurnover', sprintf('%s.pdf',
@@ -160,9 +161,10 @@ plotDistPanels <- function(){
             xlab='', ylab='',ylim=c(0,1),
             names=c("","",""),
             las=1, xaxt="n", yaxt="n")
-    beeswarm(dists ~ status, col = cols, add = TRUE)
+    beeswarm(dists ~ status, col = cols.fills, add = TRUE,
+             method="hex", corral="gutter")
     mtext("Plants",
-          3, line=0.9, cex=1.3)
+          3, line=0.9, cex=1.5)
     
     ## interaction turnover
     load(file= file.path('saved/speciesTurnover', sprintf('%s.pdf',
@@ -176,14 +178,15 @@ plotDistPanels <- function(){
             xlab='', ylab='',ylim=c(0,1),
             names=c("","",""),
             las=1)
-    beeswarm(dists ~ status, col = cols, add = TRUE)
+    beeswarm(dists ~ status, col = cols.fills, add = TRUE,
+             method="hex", corral="gutter")
     mtext("Interactions",
-          3, line=0.9, cex=1.3)
+          3, line=0.9, cex=1.5)
     mtext(c("Non-assembling \n field margin",
             "Assembling \n  hedgerow",
             "Non-assembling \n hedgerow"),
           side = 1, line= 2, at = 1:3, cex=0.9)
-    mtext("Turnover", 2, line=4, cex=1.3, adj=1.2)
+    mtext("Turnover", 2, line=4, cex=2, adj=1.35)
     
     ## weighted link turnover
     load(file="saved/phyloInt.Rdata")
@@ -197,8 +200,9 @@ plotDistPanels <- function(){
             "Assembling \n  hedgerow",
             "Non-assembling \n hedgerow"),
           side = 1, line= 2, at = 1:3, cex=0.9)
[…]

### 2016-11-01 ms/coverLetter/coverLetter.tex
added cover letter

\documentclass[12pt]{letter}
\usepackage[margin=2.5cm]{geometry}
\usepackage{mathpazo}
\usepackage[pdftex]{graphicx}
\usepackage{color}
\usepackage[normalem]{ulem}
\usepackage{times}
\usepackage{textcomp}

\address{Dept of Environmental Science,\\
  Policy, and Management\\
  130 Mulford Hall\\
  University of California, Berkeley\\
  Berkeley, California, 94720, USA} \raggedright
\begin{document}

\begin{letter}{}

  \opening{Dear editor,}

  Please find attached the manuscript \textit{Major interaction
    reorganizations punctuate the assembly of pollination networks}
  which we are submitting for consideration as a Letter in
  \textit{Ecology Letters}.

  When a species goes extinct, more than a species is lost. That
  species interacted with other species in its community, likely in
  ways that were important or even crucial to others' survival. As the
  world continues to lose species at an alarming rate, it has become
  increasingly imperative to aid the recovery of lost interactions and
  component biodviersity through ecological restoration. We know
  little, however, about how to re-assemble interacting communities
  through restoration, or the process of ecological network assembly
  more generally.

  Using extensive surveys of pollinators spanning eight years
  comprising \texttildelow $20,000$ pollinator visitation records, we
  explore the assembly of plant-pollinator communities at native plant
  restorations in the Central Valley of California. Employing newly
  developed methods for exmaining the temporal changes in networks, we
  find that species are highly dynamic in their network position,
  causing community assembly to be punctuated by major interaction
  reorganizations. The most persistent and generalized species were
  also the most variable in their network positions, contrary to what
  is expected through prefereantial attachment theory --- an assembly
  theory otherwise well-supported in the network literature. Our study
  is the first long-term study on the temporal assembly of ecological
  networks.

  Our manuscript is original and was carried out fully by the authors.
  All authors agree with the contents of the manuscript.  This
  manuscript is not published, nor is it in consideration for
  publication elsewhere.  All research not of the authors' is fully
  acknowledged.  The authors declare no conflict of interest. All
  appropriate ethical standards were followed.  Thank you for
  reviewing our manuscript and we hope you will find it suitable for
  publication.

  Regards,
  Lauren C. Ponisio, PhD
  Claire Kremen, Professor

\end{letter}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-PDF-mode: t
%%% End:

### 2016-11-01 ms/ms.tex
communities figure
added cover letter

@@
 \usepackage{url}
+\usepackage{rotating}
 \externaldocument{SI}
@@
   further evidence that hedgerows are a valuable tool for promoting
-  species conservation and ecosystem service? provision in
-  agricultural areas.
+  species conservation provision in agricultural areas.
 \end{abstract}
@@
 
-\begin{figure}
+\begin{sidewaysfigure}
   \centering
@@
   \label{fig:changePoints2}
-\end{figure}
+\end{sidewaysfigure}
 \clearpage

### 2016-11-02 ms/coverLetter/coverLetter.tex
added cover letter

@@
   increasingly imperative to aid the recovery of lost interactions and
-  component biodviersity through ecological restoration. We know
+  component biodiversity through ecological restoration. We know
   little, however, about how to re-assemble interacting communities
@@
   restorations in the Central Valley of California. Employing newly
-  developed methods for exmaining the temporal changes in networks, we
+  developed methods for examining temporal changes in networks, we
   find that species are highly dynamic in their network position,
@@
   also the most variable in their network positions, contrary to what
-  is expected through prefereantial attachment theory --- an assembly
+  is expected through preferential attachment theory --- an assembly
   theory otherwise well-supported in the network literature. Our study
   is the first long-term study on the temporal assembly of ecological
-  networks.
+  networks. It also challenges the hypothesis that mutualistic systems
+  assemble through preferential attachment (Bascompte and Stouffer,
+  2009).
 
@@
   Regards,
-  Lauren C. Ponisio, PhD
+  Lauren C. Ponisio, PhD\\
   Claire Kremen, Professor

### 2016-11-02 ms/ms.tex
added cover letter

@@
 \usepackage[pdftex]{graphicx}
-\usepackage{natbib} 
+\usepackage[numbers]{natbib} 
 \usepackage{color}
@@
   \item Submitted as a Letter
-  \item Abstract word count: 
-  \item Main text word count: 
-  \item Number of references: 
-  \item Number of figures:
+  \item Abstract word count: 148
+  \item Main text word count: 5152
+  \item Number of references: 65
+  \item Number of figures: 6
+  \item Author contributions: CK designed the study, LCP and CK
+    collected data, LCP and MPG analyzed output data.  LCP wrote the
+    first draft of the manuscript, and all authors contributed to
+    revisions.
+  \item Data accessibility statement: Data will de depositied in Dryad
+    upon publication
   \end{itemize}
@@
   dynamic in their network position, causing community assembly to be
-  punctuated by major interaction reorganizations. In contrast, the
-  non-assembling networks did not restructure as frequently. Across
-  all communities, pollinator species were opportunistic in the
-  flowers they visited. The most persistent and generalized species
+  punctuated by major interaction reorganizations.%  In contrast, the
+  % non-assembling networks did not restructure as frequently.
+  % Across all communities, pollinator species were opportunistic in the
+  % flowers they visited.
+  The most persistent and generalized species
   were also the most variable in their network positions, contrary to
-  what is expected through prefereantial attachment theory. High
-  species and interaction turnover was ubiquitous across assembling
-  and non-assembling communities, though unique interactions turned
-  over at higher rates in assembling hedgerows as the networks
-  continually reorganized. Nestedness of assembling networks also
-  increased through time, but there was no increase in robustness to
-  simulated plant extinctions. The sensitivity of networks to
-  cascading perturbations, however, increased as the communities
-  assembled, at least partially due to accumulating species
-  richness. We elucidate some of the mechanisms underlying
-  plant-pollinator network assembly and restoration, while providing
-  further evidence that hedgerows are a valuable tool for promoting
-  species conservation provision in agricultural areas.
+  what is expected through prefereantial attachment theory. % High
+  % species and interaction turnover was ubiquitous across assembling
+  % and non-assembling communities, though unique interactions turned
+  % over at higher rates in assembling hedgerows as the networks
+  % continually reorganized.
+  % Nestedness of assembling networks also increased through time, but
+  The sensitivity of networks to cascading perturbations,
+  increased with assembly, at least partially due to accumulating
+  species richness. We elucidate some of the mechanisms underlying
+  plant-pollinator network assembly and restoration --- challenging
+  the widely-held hypothesis that mutualistic networks assemble
+  through preferential attachment.
 \end{abstract}
@@
 second smallest eigenvalue of the Laplacian matrix
-\citep{fiedler1973algebraic} --- as a proxy for network robustness
-(Gaiarsa et al., in prep). Algebraic connectivity relates to how
+\citep{fiedler1973algebraic} --- as a proxy for network robustness.
+% (Gaiarsa et al., in prep)
+ Algebraic connectivity relates to how
 difficult it is to turn a network into completely disconnected groups
@@
 do not begin to produce many flowers until $3$--$5$ years following
-planting (Kremen and M'Gonigle, in prep).
+planting \citep{kremen-2015-602}.
 

### 2016-11-02 ms/coverLetter/coverLetter.tex
changed the sanky figure file and the ms, eddited the cover letter

--- 2016-11-01 ms/coverLetter/coverLetter.tex
@@
 
-  Please find attached the manuscript \textit{Major interaction
+  Please find attached the manuscript \textit{Major interaction% I like the "we are pleased to submit our manuscript" format...
     reorganizations punctuate the assembly of pollination networks}
@@
 
-  When a species goes extinct, more than a species is lost. That
-  species interacted with other species in its community, likely in
-  ways that were important or even crucial to others' survival. As the
+Our research deals with two fundamental aspects of ecological theory: to understand not only how species-rich communities assemble, but also how these assemblages change through time.
+Furthermore, as the
   world continues to lose species at an alarming rate, it has become
   increasingly imperative to aid the recovery of lost interactions and
-  component biodviersity through ecological restoration. We know
-  little, however, about how to re-assemble interacting communities
+  component biodiversity through ecological restoration.
+When a species goes extinct, not only a species is lost, but also its interactions.
+We know little, however, about how to re-assemble interacting communities
   through restoration, or the process of ecological network assembly
@@
   explore the assembly of plant-pollinator communities at native plant
-  restorations in the Central Valley of California. Employing newly
-  developed methods for exmaining the temporal changes in networks, we
+  restorations in the Central Valley of California. For the first time in the ecological litterature we employ a newly
+  developed method to examine the temporal changes in networks. Among other things, we
   find that species are highly dynamic in their network position,
   causing community assembly to be punctuated by major interaction
-  reorganizations. The most persistent and generalized species were
+  reorganizations. The most persistent and generalized species are
   also the most variable in their network positions, contrary to what
-  is expected through prefereantial attachment theory --- an assembly
+  is expected through preferential attachment theory --- an assembly
   theory otherwise well-supported in the network literature. Our study
@@
   networks.
+
+%I think this is an imporatnt paragraph - we have to be white males!!! I am not sure about the three fields....
+Our results are compelling and provide empirical evidence that widen our understading on how communities assembly and how species interactions changes through time.
+Furthermore, our results also contribute to the knowledge of how communities will be able to maintain function in the face of species extinctionAnd finally, our results challenge the view that communities assemble trhough preferential attachment.
+We believe that these exciting results that link three major ecological fields (interaction networks, community dinamics and restauration ecology) that will be of broad interest to the readership of \textit{Ecology Letters}.
 
@@
   Claire Kremen, Professor
-
+%is there a reason why I am not here? Just curious =0) 
 \end{letter}

### 2016-11-07 ms/coverLetter/coverLetter.tex
minor changes to cover letter

@@
 
-  Please find attached the manuscript \textit{Major interaction% I like the "we are pleased to submit our manuscript" format...
+  W are pleased to submit our manuscript \textit{Major interaction
     reorganizations punctuate the assembly of pollination networks}
-  which we are submitting for consideration as a Letter in
-  \textit{Ecology Letters}.
+  for consideration as a Letter in \textit{Ecology Letters}.
 
-Our research deals with two fundamental aspects of ecological theory: to understand not only how species-rich communities assemble, but also how these assemblages change through time.
-Furthermore, as the
-  world continues to lose species at an alarming rate, it has become
-  increasingly imperative to aid the recovery of lost interactions and
-  component biodiversity through ecological restoration.
-When a species goes extinct, not only a species is lost, but also its interactions.
-We know little, however, about how to re-assemble interacting communities
-  through restoration, or the process of ecological network assembly
-  more generally.
+  Our research deals with two fundamental aspects of ecological
+  theory: unerstanding how species-rich communities assemble, and how
+  these assemblages change through time. As the world continues to
+  lose species at an alarming rate, it has become increasingly
+  imperative to aid the recovery of lost interactions and component
+  biodiversity through ecological restoration. When a species goes
+  extinct, not only a species is lost, but also its interactions. We
+  know little, however, about how to re-assemble interacting
+  communities through restoration, or the process of ecological
+  network assembly more generally.
 
@@
   explore the assembly of plant-pollinator communities at native plant
-  restorations in the Central Valley of California. For the first time in the ecological litterature we employ a newly
-  developed method to examine the temporal changes in networks. Among other things, we
+  restorations in the Central Valley of California. For the first time
+  in the ecological litterature we employ a newly developed method to
+  examine the temporal changes in networks. Among other things, we
   find that species are highly dynamic in their network position,
@@
   is the first long-term study on the temporal assembly of ecological
-  networks.
+  networks.  Our results are compelling and provide empirical evidence
+  that widen our understading on how communities assembly and how
+  species interactions changes through time. Furthermore, our results
+  also contribute to the knowledge of how communities will be able to
+  maintain function in the face of species extinction. And finally,
+  our results challenge the view that communities assemble through
+  preferential attachment. We believe that these exciting results that
+  link three major ecological fields (interaction networks, community
+  dinamics and restauration ecology) that will be of broad interest to
+  the readership of \textit{Ecology Letters}.
 
-%I think this is an imporatnt paragraph - we have to be white males!!! I am not sure about the three fields....
-Our results are compelling and provide empirical evidence that widen our understading on how communities assembly and how species interactions changes through time.
-Furthermore, our results also contribute to the knowledge of how communities will be able to maintain function in the face of species extinctionAnd finally, our results challenge the view that communities assemble trhough preferential attachment.
-We believe that these exciting results that link three major ecological fields (interaction networks, community dinamics and restauration ecology) that will be of broad interest to the readership of \textit{Ecology Letters}.
-
-  Our manuscript is original and was carried out fully by the authors.
-  All authors agree with the contents of the manuscript.  This
-  manuscript is not published, nor is it in consideration for
-  publication elsewhere.  All research not of the authors' is fully
-  acknowledged.  The authors declare no conflict of interest. All
-  appropriate ethical standards were followed.  Thank you for
-  reviewing our manuscript and we hope you will find it suitable for
-  publication.
+  % Our manuscript is original and was carried out fully by the authors.
+  % All authors agree with the contents of the manuscript.  This
+  % manuscript is not published, nor is it in consideration for
+  % publication elsewhere.  All research not of the authors' is fully
+  % acknowledged.  The authors declare no conflict of interest. All
+  % appropriate ethical standards were followed. 
+  Thank you for reviewing our manuscript and we hope you will find it
+  suitable for publication.
 
   Regards,
-  Lauren C. Ponisio, PhD
+  Lauren C. Ponisio, PhD\\
   Claire Kremen, Professor

### 2016-11-10 ms/coverLetter/coverLetter.tex
added temepral traits

@@
 
-  Our research deals with two fundamental aspects of ecological
-  theory: unerstanding how species-rich communities assemble, and how
-  these assemblages change through time. As the world continues to
-  lose species at an alarming rate, it has become increasingly
-  imperative to aid the recovery of lost interactions and component
-  biodiversity through ecological restoration. When a species goes
-  extinct, not only a species is lost, but also its interactions. We
-  know little, however, about how to re-assemble interacting
+  As the world continues to lose species at an alarming rate, it has
+  become increasingly imperative to aid the recovery of lost
+  interactions and component biodiversity through ecological
+  restoration. When a species goes extinct, its interactions are also
+  lost. We know little, however, about how to re-assemble interacting
   communities through restoration, or the process of ecological
-  network assembly more generally.
+  network assembly more generally. Our research deals with two
+  fundamental aspects of the ecological theory that underlie
+  restoration ecology: understanding how species-rich communities
+  assemble, and how these assemblages change through time. Our work
+  uses a novel method to examine the temporal changes in networks,
+  represents the first long-term study of temporal assembly of
+  ecological networks, and challenges the currently accepted community
+  assembly theory of preferential attachment.
 
-  Using extensive surveys of pollinators spanning eight years
-  comprising \texttildelow $20,000$ pollinator visitation records, we
-  explore the assembly of plant-pollinator communities at native plant
-  restorations in the Central Valley of California. For the first time
-  in the ecological litterature we employ a newly developed method to
-  examine the temporal changes in networks. Among other things, we
-  find that species are highly dynamic in their network position,
-  causing community assembly to be punctuated by major interaction
-  reorganizations. The most persistent and generalized species are
-  also the most variable in their network positions, contrary to what
-  is expected through preferential attachment theory --- an assembly
-  theory otherwise well-supported in the network literature. Our study
-  is the first long-term study on the temporal assembly of ecological
-  networks.  Our results are compelling and provide empirical evidence
-  that widen our understading on how communities assembly and how
-  species interactions changes through time. Furthermore, our results
-  also contribute to the knowledge of how communities will be able to
-  maintain function in the face of species extinction. And finally,
-  our results challenge the view that communities assemble through
-  preferential attachment. We believe that these exciting results that
-  link three major ecological fields (interaction networks, community
-  dinamics and restauration ecology) that will be of broad interest to
-  the readership of \textit{Ecology Letters}.
+  Preferential attachment predicts that species entering a network are
+  more likely to interact with species that are already well-connected
+  --- and is well-supported in the network literature to date.  We
+  analyzed plant-pollinator interaction data comprising eight years
+  and \texttildelow $20,000$ records at native plant restorations in
+  the Central Valley of California. We find that species are highly
+  dynamic in their network position, causing community assembly to be
+  punctuated by major interaction reorganizations. The most persistent
+  and generalized species are also the most variable in their network
+  positions, challenging what is expected through preferential
+  attachment theory. Our results are compelling and provide empirical
+  evidence that fundamentally alter our understanding of how
+  communities assembly and how species interactions change through
+  time. Our insights will help to inform effrots to re-assembly robust
+  communities through restoration.  We believe that these exciting
+  results linking three major ecological fields (interaction networks,
+  community assembly dynamics and restoration ecology) will be of
+  broad interest to the readership of \textit{Ecology Letters}.
 
@@
   Thank you for reviewing our manuscript and we hope you will find it
-  suitable for publication.
+  suitable for consideration in \textit{Ecology Letters}.
 

### 2016-11-10 ms/ms.tex
changed the sanky figure file and the ms, eddited the cover letter
added temepral traits

@@
 % (Gaiarsa et al., in prep)
- Algebraic connectivity relates to how
-difficult it is to turn a network into completely disconnected groups
-of species \citep{costa2007characterization, gibert2013spatial}. The
-larger the algebraic connectivity, the more sensitive a network is to
-cascading perturbations. Perturbations, such as the decrease in
-abundance of a plant or pollinator, can have negative consequences for
-the species in the network. For example, a decrease in abundance of a
-pollinator will diminish the pollination services it provides to
-plants. The affected plants would set less seeds, and decrease in
-abundance the subsequent year. Consequently, other pollinators that
-depended on those plant species would also be affected, and the
-effects of this perturbation would continue to propagate throughout
-the network. Alternatively, perturbations could also have a positive
-effect. For example, the increase in abundance of a plant species
-would lead to an increase in resource availability for the
-pollinators. The examples of negative perturbations (e.g., resource
-collapse, disease spreading, parasites), however, outnumber possible
-positive perturbations. It is important to note that both robustness
-and algebraic connectivity assume that the network is static. They do
-not account for the ability of species to alter their interaction
-depending on circumstances and the resource availability.
+Algebraic connectivity relates to how difficult it is to turn a
+network into completely disconnected groups of species
+\citep{costa2007characterization, gibert2013spatial}. The larger the
+algebraic connectivity, the more sensitive a network is to cascading
+perturbations. Perturbations, such as the decrease in abundance of a
+plant or pollinator, can have negative consequences for the species in
+the network. For example, a decrease in abundance of a pollinator will
+diminish the pollination services it provides to plants. The affected
+plants would set less seeds, and decrease in abundance the subsequent
+year. Consequently, other pollinators that depended on those plant
+species would also be affected, and the effects of this perturbation
+would continue to propagate throughout the network. Alternatively,
+perturbations could also have a positive effect. For example, the
+increase in abundance of a plant species would lead to an increase in
+resource availability for the pollinators. The examples of negative
+perturbations (e.g., resource collapse, disease spreading, parasites),
+however, outnumber possible positive perturbations. It is important to
+note that both robustness and algebraic connectivity assume that the
+network is static. They do not account for the ability of species to
+alter their interaction depending on circumstances and the resource
+availability.
 
@@
   \centering
-  \includegraphics[width=.95\textwidth]{../analysis/changePoint/plotting/communitiesSkanky2.pdf}
+  \includegraphics[width=.95\textwidth]{../analysis/changePoint/plotting/communitiesSkanky3.pdf}
   \caption{The module membership between network changing points. Two

### 2016-11-10 dataPrep/src/misc.R

@@ -81,17 +81,27 @@ make.by.species <- function(spec,
       sr.sched$Date,
       sep=';')),
     GenusSpecies=unique(spec[,type]),
-    Occ=0)
+    ## Occ=0,
+    Abund=0)
   ## fill in the matrix of all combinations
   match.dates <- match(paste(all.sp$Site,
                        all.sp$Date,
-                       all.sp$GenusSpecies,
-                       sep=';'),
-                 paste(sp$SiteDate,
-                       sp$GenusSpecies,
-                       sep=';'))
-
-  sp$Occ[!is.na(match.dates)] <- 1
+                             all.sp$GenusSpecies,
+                             sep=';'),
+                       paste(sp$SiteDate,
+                             sp$GenusSpecies,
+                             sep=';'))
+  match.abund <- match(paste(sp$SiteDate,
+                             sp$GenusSpecies,
+                             sep=';'),
+                       paste(all.sp$Site,
+                       all.sp$Date,
+                             all.sp$GenusSpecies,
+                             sep=';'))
+  
+  sp$Abund <- all.sp$Abund[match.abund]
+  ## sp$Occ <- sp$Abund
+  ## sp$Occ[sp$Occ > 0] <- 1
   ## create site, date, genus etc. columns
   sp$Site <- sapply(strsplit(as.character(sp$SiteDate), ";"),
                     function(x) x[1])
@@ -113,7 +123,7 @@ make.by.species <- function(spec,
                     pollinator=as.vector(sp$GenusSpecies), 
                     var1=as.vector(sp$Site),
                     var2=sp$Date,
-                    occ=sp$Occ)
+                    occ=sp$Abund)
   
   sites <- rownames(mats[[1]])
   dates <- colnames(mats[[1]])
@@ -143,5 +153,6 @@ findOccPlant <- function(x){
 }
 
 calcOccArray <-  function(x){
+  x[x > 1] <- 1
   sum(x > 0, na.rm=TRUE)/sum(x >= 0, na.rm=TRUE)
 }

### 2016-11-14 ms/ms.tex
Marilia changed the ms accoring to some of Paulinhas suggestions

@@
 \begin{abstract}  
-  The ability of communities to maintain function in the face of
-  species extinction is related to the structure of
-  networks. Understanding network structure and how it relates to
+ % The ability of communities to maintain function in the face of
+  %species extinction is related to the structure of networks. Lauren - i changed the phrasing because structure of network and network structure were too close.
+  The structure of interaction networks is related to the ability of communities to maintain function in the face of
+  species extinction. Understanding network structure and how it relates to
   network assembly, therefore, is a priority for conservation
@@
   % flowers they visited.
-  The most persistent and generalized species
-  were also the most variable in their network positions, contrary to
+   Generalist species were the most persistent and 
+also the most variable in their network positions, contrary to
   what is expected through prefereantial attachment theory. % High
@@
 entering a network are more likely to interact with species that are
-already well-connected \citep[''the rich-get-richer''
+already well-connected \citep[the "rich-get-richer''
 principle,][]{barabasi1999emergence}. In pollination systems --- a
@@
 Understanding network assembly is particularly relevant to ecological
-restoration, which is essentially 'applied succession'
+restoration, which is essentially "applied succession"
 \citep[e.g.,][]{parker1997scale}. In pollination systems, the time
@@
 would continue to propagate throughout the network. Alternatively,
-perturbations could also have a positive effect. For example, the
-increase in abundance of a plant species would lead to an increase in
-resource availability for the pollinators. The examples of negative
+perturbations could also have a positive effect if, for example, the
+increase in the abundance of a plant species lead to an increase in
+resource availability for pollinators. The examples of negative
 perturbations (e.g., resource collapse, disease spreading, parasites),
@@
 \subsection*{Network robustness}
-Assembly did not effect the robustness of the networks to species
+Assembly did not affect network robustness to species
 extinction when species were removed incrementally by
@@
 different from what would be expected from preferential
-attachment. With a preferential attachment process, we expect that the
+attachment. In a preferential attachment process, we expect that the
 most persistent and high degree species would remain stable in the
@@
 Interestingly, however, the changes in network patterns associated
-with assembly did not effect the robustness of hedgerow communities to
+with assembly did not affect the robustness of hedgerow communities to
 species loss. This is particularly surprising given the observed
-increase in nestedness, which is often associated with an enhanced in
+increase in nestedness, which is often associated with an increase in
 robustness to extinction \citep{Memmott2004}. Perhaps assembling
@@
 
-Contrary to the general restoration goals, the susceptibility of the
-networks to cascading perturbations increased as the communities
-assembled. Because network vulnerability to cascading perturbations,
-as measured by algebraic connectivity, is correlated with species
-richness, the increase in and plant and pollinator richness following
+%Lauren - I changed the first part  of this paragraph but I really don't thinks it is that confusing.
+The problem is that it is indeed very theoretical, but it is impossible to make our point without being too technical..
+I tried to decrease the use of jargons.
+Contrary to the general restoration goals, networks' susceptibility 
+ to cascading perturbations increased as the communities assembled.
+We measured vulnerability to cascading perturbations as the algebraic connectivity of the network: the greater the algebraic connectivity, the harder it  is to turn a network into completely disconnected groups of species, and the more vulnerable a community is to cascading perturbations.
+However, algebraic connectivity is positively related not only to the number of species--- species richness--- but also to the number of interactions in the network--- connectance \citep{gibert2013spatial}.
+Therefore, the increase in and plant and pollinator richness following
 restoration is at least partially responsible for the increase in
-response to cascading effects. Connectance is also positively related
-to algebraic connectivity \citep{gibert2013spatial}, but because we
-observed a decrease in connectance, topological characteristics of the
+response to cascading effects. Additionally, despite the correlation between connectance and algebraic connectivity \citep{gibert2013spatial}, we
+observed a decrease in connectance. Thus, topological characteristics of the
 networks beyond species richness and connectance are needed to explain
@@
 
-We would like to thank Paulo R.~Guimar{\~a}es Jr., Aaron Clauset and
-Matthew Hutchinson for their invaluable discussions and comments, and
+We would like to thank Paulo R.~Guimar{\~a}es Jr., Aaron Clauset,
+Matthew Hutchinson and Ana Paula A. Assis for their invaluable discussions and comments, and
 Leto Peel for help with the change point analysis.  We thank the
@@
     their degree. Each species has a consistent position in the
-    perimeter of the network across years. In the assembling
-    hedgerows, colored squares in the corner of each network represent
-    the year of assembly. Asterisks indicate the year the hegdgerow
+    perimeter of the network across years. Asterisks indicate the year the hegdgerow
     was planted. Before that, the sites were weedy field margins.}
@@
   \includegraphics[width=.95\textwidth]{../analysis/changePoint/plotting/communitiesSkanky3.pdf}
-  \caption{The module membership between network changing points. Two
+  \caption{The species module membership between network changing points. Two
     representative assembling hedgerows are depicted. In the top
     panel, species are grouped by module. The bottom panels visualize
-    the flow of species between modules between changing points. }
+    the flow of species between modules between changing points. Each line represents the change of s species from one module to another. }
   \label{fig:changePoints2}
@@
   \includegraphics[width=1\textwidth]{../analysis/variability/figures/turnover_panels.pdf}
-  \caption{The species, interaction and weighted interactions turnover
+  \caption{Species interaction and weighted interactions turnover
     of plant-pollinator networks at non-assembling field margins
@@
     significantly less pollinator turnover.  However, when
-    interactions where weighted by their similarity, both hedgerow
-    types had higher turnover that unrestored field margins. Boxplots
+    interactions were weighted by their similarity, both hedgerow
+    types had higher turnover than unrestored field margins. Boxplots
     represent medians (black horizontal line) first and third

### 2016-11-16 ms/ms.tex
added main file that runs all analysis, fixed bug where the sum instead of the mean was taken to create networks

@@
 entering a network are more likely to interact with species that are
-already well-connected \citep[the "rich-get-richer''
+already well-connected \citep[the ``rich-get-richer''
 principle,][]{barabasi1999emergence}. In pollination systems --- a
@@
 Understanding network assembly is particularly relevant to ecological
-restoration, which is essentially "applied succession"
+restoration, which is essentially ``applied succession''
 \citep[e.g.,][]{parker1997scale}. In pollination systems, the time
@@
 generalized linear model with Binomial error to test whether the
-probability of a change point occurring varied by site type. We used
-standard techniques to determine whether the assumptions of the models
-were met for this and all subsequent models. For the non-assembling
-hedgerows and weedy field margins, only sites with five or greater
-survey years were included in this analysis (N=11). All statistical
-analyses were conducted in R 3.2.3 \citep{R}.
+probability of a change point occurring varied by site type
+\citep{lme4}. We used standard techniques to determine whether the
+assumptions of the models were met for this and all subsequent
+models. For the non-assembling hedgerows and weedy field margins, only
+sites with five or greater survey years were included in this analysis
+(N=11). All statistical analyses were conducted in R 3.2.3 \citep{R}.
 
@@
 the network \citep{freeman1978centrality}. The shorter the mean path
-length to other species, the higher is the closeness centrality. We
-use linear mixed models to test whether the species closeness
-variability (log) is related to the persistence or degree of that
-species \citep{lme4, lmetest}. We included random effects for species
-and site. Because the degree and persistence of pollinators were
-strongly correlated, ($\rho = 0.84$, $p$-value $<$ $2*10^{-16}$), we
-include each explanatory variable in separate models. Plant degree and
+length to other species, the higher the closeness centrality. We use
+linear mixed models to test whether the species closeness variability
+(log) is related to the persistence or degree of that species
+\citep{lme4, lmetest}. We included random effects for species and
+site. Because the degree and persistence of pollinators were strongly
+correlated, ($\rho = 0.84$, $p$-value $<$ $2*10^{-16}$), we include
+each explanatory variable in separate models. Plant degree and
 persistence were not significantly correlated, but we use the same
@@
 occurring in one year are similar to those in the next year when they
-are weighted by the similarity of their constituent species
-(Fig.~ S2). We develop a method to examine the temporal
-turnover of interactions with weightings based on their similarity. We
-followed the algorithm of \cite{ahn2010link} to cluster all the
-interactions (edges) hierarchically across sites and years based on
-their similarity, and build a dendrogram. The interaction similarity
-is based how may plants and pollinators (vertices) two edges share
+are weighted by the similarity of their constituent species (Fig.~
+S2). We develop a method to examine the temporal turnover of
+interactions with weightings based on their similarity. We followed
+the algorithm of \cite{ahn2010link} to cluster all the interactions
+(edges) hierarchically across sites and years based on their
+similarity, and build a dendrogram. The interaction similarity is
+based how may plants and pollinators (vertices) two edges share
 \citep{ahn2010link, kalinka2011linkcomm}. The more species edges
@@
 distance \citep{graham2008phylogenetic, picante-2010-1463}. We then
-use linear models to test whether the weighted turnover of
-interactions varied between assembling and non-assembling networks
-\citep{lme4, lmetest}.
+use linear mixed models to test whether the weighted turnover of
+interactions varied between assembling and non-assembling networks,
+including site as a random effect \citep{lme4, lmetest}.
 
@@
 pollinator species richness through time. We employ generalized linear
-mixed models with Poisson error to model richness. We scaled
-explanatory variables.
+mixed models with Poisson error to model richness \citep{lme4}. We
+scaled explanatory variables.
 
@@
 
+In order to test whether hedgerows changed in robustness as the
+communities assembled, the three measures of robustness were modeled
+as responses in a linear mixed model with site type as an explanatory
+variable and site as a random effect \citep{lme4, lmetest}.
+
 \section*{Results}
@@
 punctuated by more interaction reorganizations than would be expected
-by environmental shifts alone that would effect all networks
+by environmental shifts alone that would have affected all networks
 similarly.
@@
 less pollinator turnover than field margins (Fig.~\ref{fig:beta},
-estimate $\pm$ SE of the difference in
-turnover between field margins and mature hedgerows, $-0.0498 \pm
-0.026$, $p$-value=$0.058$). When interactions where weighted by their
-similarity, both assembling and mature hedgerows had higher rates of
-turnover than field margins (Fig.~\ref{fig:beta}, estimate $\pm$
-SE of the difference in turnover between
-field margins and assembling hedgerows, $0.115 \pm 0.027$,
+estimate $\pm$ SE of the difference in turnover between field margins
+and mature hedgerows, $-0.0498 \pm 0.026$, $p$-value=$0.058$). When
+interactions where weighted by their similarity, both assembling and
+mature hedgerows had higher rates of turnover than field margins
+(Fig.~\ref{fig:beta}, estimate $\pm$ SE of the difference in turnover
+between field margins and assembling hedgerows, $0.115 \pm 0.027$,
 $p$-value=$0.0002$; field margins and mature hedgerows, $0.082 \pm
@@
 % $p$-value=$0.65$)
-In contrast, the sensitivity of networks to
-cascading perturbations, as measured by the algebraic connectivity of
-the network, increased as the network assembled (Fig.~\ref{fig:rob},
-estimate of the slope of sensitivity to cascading perturbations
-through time $\pm$ SE, $0.6814 \pm 0.272$, $p$-value=$0.042$).
+In contrast, the sensitivity of networks to cascading perturbations,
+as measured by the algebraic connectivity of the network, increased as
+the network assembled (Fig.~\ref{fig:rob}, estimate of the slope of
+sensitivity to cascading perturbations through time $\pm$ SE, $0.6814
+\pm 0.272$, $p$-value=$0.042$).
 
@@
 during assembly, the way in which these reorganizations occur was
-different from what would be expected from preferential
-attachment. In a preferential attachment process, we expect that the
-most persistent and high degree species would remain stable in the
-network core during assembly
-\citep{barabasi1999emergence}. Surprisingly, however, we encountered
-the opposite pattern. For example, the four most ubiquitous species in
-our study landscape --- \textit{Halictus ligatus}, \textit{Halictus
-  tripartitus}, \textit{Lasioglossum (Dialictus) incompletum}, and
-\textit{Toxomerus marginatus} --- were the only species that changed
-which module they were a member in across years in all the assembling
-hedgerows. Because species degree and persistence were strongly
-correlated, it is difficult to disentangle the causal mechanism for
-why species with those characteristics are so variable in their
-network position. Generalized species may be better able to exploit
-the limited floral resources in the intensively managed agriculture
-landscape, and thus also be the most persistent \citep[in ant-plant
-mutualisms, ][]{diaz2010changes}. More persistent species usually have
-longer phenologies, so they can visit many different flowers,
-resulting in a higher degree \citep{Vazquez2009,
-  fort2016abundance}. Either way, our result suggests that adaptable
-species can change their network position to utilize the most
-advantageous floral resources available, which may depend on both the
-other pollinator species that are present and the state of the plant
-community \citep{macleod2016measuring, gomez2006ecological,
-  Waser1996}. Thus given the opportunity and ability to
-use different resources, species will often change their network
-positions \citep{macleod2016measuring}.
+different from what would be expected from preferential attachment. In
+a preferential attachment process, we expect that the most persistent
+and high degree species would remain stable in the network core during
+assembly \citep{barabasi1999emergence}. Surprisingly, however, we
+encountered the opposite pattern. For example, the four most
+ubiquitous species in our study landscape --- \textit{Halictus
+  ligatus}, \textit{Halictus tripartitus}, \textit{Lasioglossum
+  (Dialictus) incompletum}, and \textit{Toxomerus marginatus} --- were
+the only species that changed which module they were a member in
+across years in all the assembling hedgerows. Because species degree
+and persistence were strongly correlated, it is difficult to
+disentangle the causal mechanism for why species with those
+characteristics are so variable in their network position. Generalized
+species may be better able to exploit the limited floral resources in
+the intensively managed agriculture landscape, and thus also be the
+most persistent \citep[in ant-plant mutualisms,
+][]{diaz2010changes}. More persistent species usually have longer
+phenologies, so they can visit many different flowers, resulting in a
+higher degree \citep{Vazquez2009, fort2016abundance}. Either way, our
+result suggests that adaptable species can change their network
+position to utilize the most advantageous floral resources available,
+which may depend on both the other pollinator species that are present
+and the state of the plant community \citep{macleod2016measuring,
+  gomez2006ecological, Waser1996}. Thus given the opportunity and
+ability to use different resources, species will often change their
+network positions \citep{macleod2016measuring}.
 
@@
 
-%Lauren - I changed the first part  of this paragraph but I really don't thinks it is that confusing.
-The problem is that it is indeed very theoretical, but it is impossible to make our point without being too technical..
-I tried to decrease the use of jargons.
-Contrary to the general restoration goals, networks' susceptibility 
- to cascading perturbations increased as the communities assembled.
-We measured vulnerability to cascading perturbations as the algebraic connectivity of the network: the greater the algebraic connectivity, the harder it  is to turn a network into completely disconnected groups of species, and the more vulnerable a community is to cascading perturbations.
-However, algebraic connectivity is positively related not only to the number of species--- species richness--- but also to the number of interactions in the network--- connectance \citep{gibert2013spatial}.
-Therefore, the increase in and plant and pollinator richness following
-restoration is at least partially responsible for the increase in
-response to cascading effects. Additionally, despite the correlation between connectance and algebraic connectivity \citep{gibert2013spatial}, we
-observed a decrease in connectance. Thus, topological characteristics of the
-networks beyond species richness and connectance are needed to explain
-the increased sensitivity to perturbations spreading. These hedgerows
-were designed to provide floral resources to the largest number of
-pollinators across the growing season \citep{menz-2010-4}. The
-generalized nature of the floral community may explain why the
-networks tended to be more generalized than expected if interactions
-were randomly distributed across species (Fig.~\ref{fig:baci}). In
-addition, the design of the hedgerow plantings may have facilitated
-the emergence of a single, highly connected module in all of the
-networks (see \ref{fig:changePoints2} for examples). This network
-configuration results in short path lengths (the distance between
-species in a network based on their shared partners), and thus, a
-perturbation in one species can more easily spread to other
-species. In order to promote more resilient communities, future
-restoration efforts should explore designing floral communities to
-promote more interaction partitioning using, for example, algorithms
-to optimize different network properties based on prior knowledge of
-pollinator floral preferences \citep{mgonigle2016tool}, and on desired
-network architectures that renders them more robust both to species
-loss and to cascading effects.
+% Lauren - I changed the first part of this paragraph but I really
+% don't thinks it is that confusing. The problem is that it is indeed
+% very theoretical, but it is impossible to make our point without
+% being too technical.. I tried to decrease the use of jargons.
+Contrary to the general restoration goals, networks' susceptibility to
+cascading perturbations increased as the communities assembled. We
+measured vulnerability to cascading perturbations as the algebraic
+connectivity of the network: the greater the algebraic connectivity,
+the more vulnerable a community is to cascading
+perturbations. Algebraic connectivity is positively related to both
+the number of species and connectance of a network
+\citep{gibert2013spatial}. The increase in and plant and pollinator
+richness following restoration is thus at least partially responsible
+for the increase in response to cascading effects.  Algebraic
+connectivity is also positivly correlated with the number in
+interactions, or connectance of a network \citep{gibert2013spatial},
+We, however, observed a decrease in connectance, suggesting
+topological characteristics of the networks beyond just species
+richness and connectance are needed to explain the increased
+sensitivity to perturbations spreading. These hedgerows were designed
+to provide floral resources to the largest number of pollinators
+across the growing season \citep{menz-2010-4}. The generalized nature
+of the floral community may explain why the networks tended to be more
+generalized than expected if interactions were randomly distributed
+across species (Fig.~\ref{fig:baci}). In addition, the design of the
+hedgerow plantings may have facilitated the emergence of a single,
+highly connected module in all of the networks (see
+\ref{fig:changePoints2} for examples). This network configuration
+results in short path lengths (the distance between species in a
+network based on their shared partners), and thus, a perturbation in
+one species can more easily spread to other species. In order to
+promote more resilient communities, future restoration efforts should
+explore designing floral communities to promote more interaction
+partitioning using, for example, algorithms to optimize different
+network properties based on prior knowledge of pollinator floral
+preferences \citep{mgonigle2016tool}, and on desired network
+architectures that renders them more robust both to species loss and
+to cascading effects.
 
@@
 We would like to thank Paulo R.~Guimar{\~a}es Jr., Aaron Clauset,
-Matthew Hutchinson and Ana Paula A. Assis for their invaluable discussions and comments, and
-Leto Peel for help with the change point analysis.  We thank the
-growers and land owners that allowed us to work on their property. We
-also greatly appreciate the identification assistance of expert
-taxonomists Martin Hauser, Robbin Thorp and Jason Gibbs.  This work
-was supported by funding from the Army Research Office
-(W911NF-11-1-0361 to CK), the Natural Resources Conservation Service
-(CIG-69-3A75-12-253, CIG-69-3A75-9-142, CIG-68-9104-6-101 and
+Matthew Hutchinson and Ana Paula A. Assis for their invaluable
+discussions and comments, and Leto Peel for help with the change point
+analysis.  We thank the growers and land owners that allowed us to
+work on their property. We also greatly appreciate the identification
+assistance of expert taxonomists Martin Hauser, Robbin Thorp and Jason
+Gibbs.  This work was supported by funding from the Army Research
+Office (W911NF-11-1-0361 to CK), the Natural Resources Conservation
+Service (CIG-69-3A75-12-253, CIG-69-3A75-9-142, CIG-68-9104-6-101 and
 WLF-69-7482-6-277 to The Xerces Society), the National Science

### 2016-11-16 analysis/networkLevel/resilience.R

@@ -1,10 +1,10 @@
-rm(list=ls())
+
 setwd('~/Dropbox/hedgerow_assembly/analysis/networkLevel')
 source('src/initialize.R')
 load('../../data/networks/all_networks_years.Rdata')
 
 ## either "abund" or "degree"
-extinction.method <- "abund"
+## extinction.method <- "degree"
 
 ## **********************************************************
 ## robustness

### 2016-11-16 analysis/speciesLevel/plotting/specMets.R

@@ -44,16 +44,17 @@ for(j in pp){
 
 
 ## closeness only
+dd.ypr <- cbind(dd, closeness.log=0)
 
 ypr.pi.pol <- predict.int(mod= mods$pol$closeness.log,
-                      dd=dd.ypr,
-                      y="closeness.log",
-                      family="gaussian")
+                          dd=dd.ypr,
+                          y="closeness.log",
+                          family="gaussian")
 
 
 ypr.pi.plant <- predict.int(mod= mods$plant$closeness.log,
-                      dd=dd.ypr,
-                      y="closeness.log",
-                      family="gaussian")
+                            dd=dd.ypr,
+                            y="closeness.log",
+                            family="gaussian")
 
 plot.panels()

### 2016-11-16 analysis/speciesLevel/specMets.R

@@ -6,13 +6,8 @@ load('../../data/networks/all_networks_years.Rdata')
 ## **********************************************************
 ## species importance
 ## **********************************************************
-## specs <- calcSpec(nets, spec, spec.metric = "d", 0.3)
-## save(specs, file=file.path(save.path, 'specs.Rdata'))
-
 ## linear models
 load(file=file.path(save.path, 'specs.Rdata'))
-specs$closeness[specs$closeness == 0] <- 1*10^-6
-specs$closeness.log <- log(specs$closeness)
 
 ## SiteStatus or ypr
 xvar <- "ypr"

### 2016-11-16 analysis/speciesLevel/src/initialize.R

@@ -4,9 +4,7 @@ library(lme4)
 library(lmerTest)
 library(RColorBrewer)
 source('../../dataPrep/src/prepNets.R')
-source('src/CalcMetrics.R')
 source('src/misc.R')
-source("src/specialization.R")
 
 traits <- read.csv("../../data/traits.csv")
 load('../../data/networks/allSpecimens.Rdata')

### 2016-11-16 analysis/variability/beta-disp.R

@@ -5,7 +5,7 @@ library(bipartite)
 source('src/calcDis.R')
 traits <- read.csv("../../data/traits.csv")
 ## ************************************************************
-## the variability fo interactions of plants and pols
+## the variability of interactions of plants and pols
 ## ************************************************************
 ## for each species at each site, 1) create a interaction partner by
 ## year matrix, 2) calculate the dissimilarity of interaction partners

### 2016-11-16 analysis/variability/beta-div.R

@@ -1,9 +1,9 @@
-rm(list=ls())
+## rm(list=ls())
 setwd('~/Dropbox/hedgerow_assembly/analysis/variability')
 binary <- FALSE
 alpha <- TRUE
 ## int or pols
-type <- "pols"
+## type <- "plants"
 source('src/initialize_beta.R')
 
 ## ************************************************************
@@ -57,4 +57,4 @@ plot.coeffs(dis.method =dis.method, mod=summary(mod),
             sub= type, occ=occ)
 
 
-plotDistPanels()
+## plotDistPanels()

### 2016-11-16 analysis/variability/beta-link.R

@@ -12,10 +12,10 @@ edges.com <- aggregate(list(abund=spec$GenusSpecies),
                             PlantGenusSpecies=spec$PlantGenusSpecies),
                        length)
 
-## lc <- getLinkCommunities(edges.com,
-##                          hcmethod = "average",
-##                          bipartite=TRUE)
-## save(lc, file="saved/lc.Rdata")
+lc <- getLinkCommunities(edges.com,
+                         hcmethod = "average",
+                         bipartite=TRUE)
+save(lc, file="saved/lc.Rdata")
 
 ## ************************************************************
 ## turnover of phylo interactions through time
@@ -26,9 +26,9 @@ spec$Int <- paste(spec$GenusSpecies,
 phylo.int <- calcCommDis(spec, "Int", lc, abund.w=TRUE)
 save(phylo.int, file="saved/phyloInt.Rdata")
 
-## linear model of phylo int by years between samples
-## mature and maturing hedgerows have more intearction turnover
-## between years
+## linear model of phylo int by years between samples mature and
+## maturing hedgerows have more intearction turnover between years
+
 load(file="saved/phyloInt.Rdata")
 phylo.int.mod <- lmer(PhyloInt ~ SiteStatus +
                       (1|Site),
@@ -38,13 +38,13 @@ summary(phylo.int.mod)
 ## ## for comparing maturing and mature
 ## ## mature and maturing are not sig different
 
-## phylo.int$phylo.int$SiteStatus <-
-##   factor(phylo.int$phylo.int$SiteStatus,
-##          levels=c("mature", "control", "maturing"))
-## phylo.int.mod <- lmer(PhyloInt ~ SiteStatus +
-##                       (1|Site),
-##                       data=phylo.int$phylo.int)
-## summary(phylo.int.mod)
+phylo.int$phylo.int$SiteStatus <-
+  factor(phylo.int$phylo.int$SiteStatus,
+         levels=c("mature", "control", "maturing"))
+phylo.int.mod <- lmer(PhyloInt ~ SiteStatus +
+                      (1|Site),
+                      data=phylo.int$phylo.int)
+summary(phylo.int.mod)
 
 
 plot.node.box(ylabel="Weighted interaction turnover",

### 2016-11-16 analysis/variability/cv.R

@@ -197,8 +197,7 @@ save(itd, dprime, degree,
      dprime.k.cv, dprime.closeness.cv,
      occ.k.cv, occ.closeness.cv,
      plants.occ.k.cv, plants.occ.closeness.cv,
-     degree.k.cv, degree.closeness.cv,
-     plants.degree.k.cv, plants.degree.closeness.cv,
+     degree.k.cv, degree.closeness.cv, plants.degree.closeness.cv,
      file="saved/contMods.Rdata")
 
 ## ************************************************************

### 2016-11-16 analysis/variability/plotting/beta-int.R

@@ -26,20 +26,20 @@ plants.k.pi <- predict.int(mod= mod.plants,
                            y="Dist",
                            family="gaussian")
 
-plot.predict.div(new.dd=plants.k.pi,
-                 ylabel="Interaction dispersion",
-                 dats=plants,
-                 xs="k",
-                 y1="Dist",
-                 xlabel="Plant network position (k)",
-                 legend.loc="topright",
-                 height=5,
-                 width=5,
-                 x.adj=0.5,
-                 scaled=FALSE,
-                 treatments="all",
-                 col.lines="black",
-                 f.path = 'figures/intdisper')
+## plot.predict.div(new.dd=plants.k.pi,
+##                  ylabel="Interaction dispersion",
+##                  dats=plants,
+##                  xs="k",
+##                  y1="Dist",
+##                  xlabel="Plant network position (k)",
+##                  legend.loc="topright",
+##                  height=5,
+##                  width=5,
+##                  x.adj=0.5,
+##                  scaled=FALSE,
+##                  treatments="all",
+##                  col.lines="black",
+##                  f.path = 'figures/intdisper')
 
 
 ## ************************************************************
@@ -58,20 +58,20 @@ plants.closeness.pi <- predict.int(mod= mod.plants.close,
                                    y="Dist",
                                    family="gaussian")
 
-plot.predict.div(new.dd=plants.closeness.pi,
-                 ylabel="Interaction dispersion",
-                 dats=plants,
-                 xs="closeness",
-                 y1="Dist",
-                 xlabel="Plant closeness",
-                 legend.loc="topright",
-                 height=5,
-                 width=5,
-                 x.adj=0.5,
-                 scaled=FALSE,
-                 treatments="all",
-                 col.lines="black",
-                 f.path = 'figures/intdisper')
+## plot.predict.div(new.dd=plants.closeness.pi,
+##                  ylabel="Interaction dispersion",
+##                  dats=plants,
+##                  xs="closeness",
+##                  y1="Dist",
+##                  xlabel="Plant closeness",
+##                  legend.loc="topright",
+##                  height=5,
+##                  width=5,
+##                  x.adj=0.5,
+##                  scaled=FALSE,
+##                  treatments="all",
+##                  col.lines="black",
+##                  f.path = 'figures/intdisper')
 
 
 ## ************************************************************
@@ -91,20 +91,20 @@ plants.d.pi <- predict.int(mod= mod.plants.d,
[…]

### 2016-11-16 analysis/variability/src/plotDists.R

@@ -168,7 +168,7 @@ plotDistPanels <- function(){
     
     ## interaction turnover
     load(file= file.path('saved/speciesTurnover', sprintf('%s.pdf',
-           paste(dis.method, alpha, occ, type="int", sep='_'))))
+           paste(dis.method, alpha, occ, type="ints", sep='_'))))
     dists <- dats$dist
     status <- dats$status
     status <- factor(status,

### 2016-11-16 dataPrep/src/misc.R

@@ -2,7 +2,7 @@
 ## This functions takes site-species-abundance data and creates a
 ## matrix where the sites are columns and the rows are species.
 
-samp2site.spp <- function(site, spp, abund, FUN=mean) {
+samp2site.spp <- function(site, spp, abund, FUN=sum) {
   x <- tapply(abund, list(site = site, spp = spp), FUN)
   x[is.na(x)] <- 0
   return(x)

### 2016-11-16 dataPrep/src/prepNets.R

@@ -71,32 +71,45 @@ getVisitChange <- function(cut.off1, cut.off2, metric, type, type2){
 
 
 ## the purpose of this function is to break up data with many
-## sites/years and prepare it for network analysis.
+## sites/years and prepre it for network analysis.
 
 dropNet <- function(z){
   z[!sapply(z, FUN=function(q){
-    any(dim(q) < 3)
+    any(dim(q) < 5)
   })]
 }
 
 breakNet <- function(spec.dat, site, year){
   ## puts data together in a list and removes empty matrices
-  sites <- split(spec.dat, spec.dat[,site])
+  agg.spec <- aggregate(list(abund=spec.dat$GenusSpecies),
+                        list(GenusSpecies=spec.dat$GenusSpecies,
+                             SampleRound=spec.dat$SampleRound,
+                             Site=spec.dat[,site],
+                             Year=spec.dat[,year],
+                             PlantGenusSpecies=spec.dat$PlantGenusSpecies),
+                        length)
+  agg.spec <- aggregate(list(abund=agg.spec$abund),
+                        list(GenusSpecies=agg.spec$GenusSpecies,
+                             Site=agg.spec$Site,
+                             Year=agg.spec$Year,
+                             PlantGenusSpecies=agg.spec$PlantGenusSpecies),
+                        mean)
+  sites <- split(agg.spec, agg.spec[,site])
   networks <- lapply(sites, function(x){
-    lapply(split(x, f=x[,year]), as.matrix)
+    split(x, f=x[,year])
   })
   ## formats data matrices appropriate for network analysis
-  comms <- rapply(networks, function(y){
+  comms <- lapply(unlist(networks, recursive=FALSE), function(y){
     samp2site.spp(site=y[,"PlantGenusSpecies"],
                   spp=y[,"GenusSpecies"],
-                  abund=rep(1, nrow(y)))
-  }, how="replace")
-  adj.mat <- unlist(lapply(comms, dropNet), recursive=FALSE)
+                  abund=y[,"abund"])
+  })
+  comms <- dropNet(comms)
   if(year == "assem"){
-    names(adj.mat) <- sub("\\.", "_", names(adj.mat),
-                          perl=TRUE)
+    names(comms) <- sub("\\.", "_", names(comms),
+                        perl=TRUE)
   }
-  return(adj.mat)
+  return(comms)
 }
 
 

### 2016-11-16 main.R

@@ -1,0 +1,84 @@
+rm(list=ls())
+## prep data
+## 1) drop non bee/syrphid data
+## 2) create networks
+## 3) calculate network metrics for whole landscape
+## 4) add trait data to spec sheet
+source("~/Dropbox/hedgerow_assembly/dataPrep/dataPrep.R")
+
+##***************************************************************
+## change points
+##***************************************************************
+## create graphs for change point analysis
+source('~/Dropbox/hedgerow_assembly/analysis/changePoint/dataPrep.R')
+
+## run python code for calculating network change points
+## plot networks and network communities
+source('~/Dropbox/hedgerow_assembly/analysis/changePoint/plotting/networks.R')
+
+## binomial analysis in comparisons.R
+## exploratory analysis in  core.R
+
+##***************************************************************
+## species level metrics
+##***************************************************************
+## calcualte species level metrics, run linear models
+source('~/Dropbox/hedgerow_assembly/analysis/speciesLevel/specMets.R')
+## make species level metric plots
+source('~/Dropbox/hedgerow_assembly/analysis/speciesLevel/plotting/specMets.R')
+
+##***************************************************************
+## network level metrics
+##***************************************************************
+## network metrics through time
+source('~/Dropbox/hedgerow_assembly/analysis/networkLevel/baci.R')
+
+## robustness
+## either "abund" or "degree"
+extinction.method <- "degree"
+source('~/Dropbox/hedgerow_assembly/analysis/networkLevel/resilience.R')
+extinction.method <- "abund"
+source('~/Dropbox/hedgerow_assembly/analysis/networkLevel/resilience.R')
+## sensitivity to perturbation
+source('~/Dropbox/hedgerow_assembly/analysis/networkLevel/laplacian.R')
+
+## plotting
+source('~/Dropbox/hedgerow_assembly/analysis/networkLevel/plotting/baci.R')
+source('~/Dropbox/hedgerow_assembly/analysis/networkLevel/plotting/resilience.R')
+
+##***************************************************************
+## turnover metrics
+##***************************************************************
+## coefficient of variation
+source('~/Dropbox/hedgerow_assembly/analysis/variability/cv.R')
+source('~/Dropbox/hedgerow_assembly/analysis/variability/plotting/cv.R')
+
+## nulls for computing stats
[…]

### 2016-11-28 ms/ms.tex

@@
 site. Because the degree and persistence of pollinators were strongly
-correlated, ($\rho = 0.84$, $p$-value $<$ $2*10^{-16}$), we include
+correlated, ($\rho = 0.071$, $p$-value $<$ $2*10^{-16}$), we include
 each explanatory variable in separate models. Plant degree and
@@
 of the slope of closeness centrality variability and persistence $\pm$
-standard error of the estimate, $0.653 \pm 0.225$, $p$-value=$0.009$;
-slope of closeness centrality variability and degree, $0.008 \pm
-0.002$, $p$-value=$0.002$). The slope of these relationships remained
+standard error of the estimate, $0.496 \pm 0.182$, $p$-value=$0.016$;
+slope of closeness centrality variability and degree, $0.004 \pm
+0.002$, $p$-value=$0.056$). The slope of these relationships remained
 significant when the species with the top $10$ persistence and degree
@@
 estimate of the slope of closeness through time $\pm$ SE, pollinators:
-$-0.0003 \pm 0.00005$, $p$-value=$2.7*10^{-12}$; plants $-0.007 \pm
-0.001$, $p$-value=$1.4*10^{-6}$). Through statistically significant,
-the slopes are so slight they may not be biologically significant.
+$-0.082 \pm 0.021$, $p$-value=$0.003$; plants $-0.089 \pm
+0.044$, $p$-value=$0.04$). Though statistically significant,
+the models seems to be a poor fit to the data.
 
@@
 estimate $\pm$ SE of the difference in turnover between field margins
-and mature hedgerows, $-0.0498 \pm 0.026$, $p$-value=$0.058$). When
+and mature hedgerows, $-0.051 \pm 0.026$, $p$-value=$0.049$). The turnover of plants and interactions was not significantly different between site types. When
 interactions where weighted by their similarity, both assembling and
@@
 \subsection*{Network structure}
-Network nestedness significantly increased with assembly
-(Fig.~\ref{fig:baci}, estimate of the slope of nestedness through time
-$\pm$ SE, $1.834 \pm 0.6142$, $p$-value=$0.022$). All of the networks
-were significantly nested ($z$-scores $>$ 2,
-Fig.~\ref{fig:baci}). Modularity decreased (Fig.~\ref{fig:baci}),
-though the slope was not significantly different from
-zero. % (estimate
+%Network nestedness significantly increased with assembly
+%(Fig.~\ref{fig:baci}, estimate of the slope of nestedness through time
+%$\pm$ SE, $1.834 \pm 0.6142$, $p$-value=$0.022$). 
+Nestedness and modularity did not change in any predictable pattern with assembly (Fig.~\ref{fig:baci}).
+All of the networks were significantly nested ($z$-scores $>$ 2), but not modular ($z$-scores $<$ 2, Fig.~\ref{fig:baci}).
+
+%though the slope was not significantly different from
+%zero.(estimate
 % of the slope of modularity through time $\pm$ standard error of the
 % estimate, $-0.524$ $\pm$ $0.295$, $p$-value=$0.124$).
-In addition, none of the networks were significantly modular
-($z$-scores $<$ 2, Fig.~\ref{fig:baci}). Connectance decreased as the
+%In addition, none of the networks were significantly modular
+% Fig.~\ref{fig:baci}).
+
+Connectance decreased as the
 community assembled (Fig.~\ref{fig:baci}, estimate of the slope of
 connectance through time $\pm$ standard error of the estimate,
-$-0.0434$ $\pm$ $0.0152$, $p$-value=$0.03$). Specialization also
-decreased, though the slope was only marginally significantly
-different from zero (estimate of the slope of specialization through
-time $\pm$ SE, $-0.926$ $\pm$ $0.450$, $p$-value=$0.078$). Most
-communities were more generalized than expected when interactions were
+$-0.023$ $\pm$ $0.008$, $p$-value=$0.007$). Specialization also
+decreased, though the slope was not significantly
+different from zero. %(estimate of the slope of specialization through time $\pm$ SE, $-0.926$ $\pm$ $0.450$, $p$-value=$0.078$). 
+Most communities were more generalized than expected when interactions were
 randomized (Fig.~\ref{fig:baci}).
@@
 (Fig.~\ref{fig:baci}, estimate of the slope of richness through time
-$\pm$ SE, pollinators: $0.193$ $\pm$ $0.0729$, $p$-value=$0.008$;
-plants: $0.212$ $\pm$ $0.0653$, $p$-value=$0.001$). Unsurprisingly,
+$\pm$ SE, pollinators: $0.136$ $\pm$ $0.064$, $p$-value=$0.034$;
+plants: $0.140$ $\pm$ $0.060$, $p$-value=$0.012$). Unsurprisingly,
 pollinator species are colonizing and persisting at the assembling

### 2016-11-28 ms/ms.tex
make new resilence figs

@@
   % The ability of communities to maintain function in the face of
-  %species extinction is related to the structure of networks. Lauren - i changed the phrasing because structure of network and network structure were too close.
+  % species extinction is related to the structure of networks. Lauren
+  % - i changed the phrasing because structure of network and network
+  % structure were too close.
   The structure of interaction networks is related to the ability of communities to maintain function in the face of
@@
 networks that respond to abrupt shifts in the behavior of interactors
-\citep{peel2014detecting}. In ecological communities, such shifts may
-occur if, as new species colonize, resident species change their
-interaction partners to optimize their foraging effort. In
-plant-pollinator communities, theory predicts that pollinators
-optimize their use of floral resources to reduce interspecific
-competition and improve resource-use efficiency
-\citep{pyke1984optimal, valdovinos2010consequences,
-  valdovinos2013adaptive, albrecht2010plant, Bluthgen2007}. No
-studies, however, have examined whether network changing points occur
-during ecological network assembly, and how these changes relate to
-the species behavior.
-
-%% restoration
-Understanding network assembly is particularly relevant to ecological
-restoration, which is essentially ``applied succession''
-\citep[e.g.,][]{parker1997scale}. In pollination systems, the time
-since an area was restored has been shown to affect the structure of
-networks \citep{forup-2008-742, devoto2012understanding}, suggesting
-interactions are changing as the community develops. Understanding the
-mechanisms of network assembly will help to guide community
-restoration. Facilitating network restoration is especially imperative
-in areas where species interactions provide essential ecosystem
-services, such as crop pollination. To ensure the continued provision
-of ecosystem services and curb biodiversity loss, it is critical to
-restore pollinators and their interactions in agricultural
-landscapes. To promote pollinator services in agriculture, farmers may
-chose to plant strips of native plants along farm edges (hedgerows) to
-help provide habitat for pollinators without removing arable land from
-production. Hedgerows augment the richness, abundance and trait
-diversity of pollinators in agricultural landscapes
-\citep{morandin-2013-829, kremen-2015-602, ponisio2015farm}, and
-promote the persistence and colonization of floral resource
-specialists \citep{mgonigle-2015-x}. It is important to understand how
-these new species are being incorporated into the network as the
-community assembles, and the consequences for interaction patterns and
-robustness.
-
-%% Lauren I would put 10 year dataset; and I *loved* the way you
-%% organized the ideas here!!
-We explore the process of network development using a nine year
-dataset of plant-pollinator community assembly following hedgerow
-restoration in the highly simplified and intensively managed
-agricultural landscape of California's Central Valley.  We first
-determine whether the mechanism underlying network assembly is a build
-up of interactions as would be predicted by preferential attachment,
-or instead is punctuated by significant reorganizations of
-interactions (i.e., network changing points). Even with changing
-points in interaction organization, networks could still be assembling
-via preferential attachment if the network reorganizations were
-primarily driven by peripheral, temporally variable species while a
-stable, well-connected core of species persist. We test whether the
-species that are most variable in their network position --- and thus
-important contributors to network reorganizations --- are less
-persistent and connected species. To further explore the mechanisms
-underlying the temporal dynamics of the networks, we examine patterns
-in the species and interaction temporal turnover. Lastly, we
-investigate whether networks are assembling toward predictable
-interaction patterns, and the ramifications for the robustness of the
-networks to species extinction and cascading perturbations.
-
-%% objectives
-% 1) Are hedgerows assembling by preferential attachment, or are there
-% more abrupt reorganizations of network structure?
-% 2) Who are the species responsible for interaction reorganizations?
-% how to incorporate this? 
-% 3) Are the changing points a product of turnover in species
-% composition, or are more substantial changes in the verticles of the
-% networks?
-% 4) Is the structure of interactions changing? Does this change the
-% robustness of the network to species extinction and/or perturbation?
-
-\section*{Materials \& Methods}
-\label{sec:methods}
-
-\subsection*{Study sites and collection methods}
-\label{sec:study-sites}
-
-We surveyed plant-pollinator interaction networks of independent
-assembling hedgerows communities (N=5), and of two types of
-non-assembling communities to serve as controls: unrestored, weedy
-field margins (N=19) and mature hedgerows (greater than 10 years since
-planting, N=29). The sites were located in the Central Valley of
-California in Yolo, Colusa and Solano Counties. This area is composed
-of intensively managed agriculture --- primarily monocultures of
-conventional row crops, vineyards and orchards. Hedgerows are ca.~3--6
-m wide and approximately 350 m long, bordering large (ca.\
-30--hectare) crop fields. Hedgerows consist of native, perennial,
-shrub and tree plantings including \textit{Rosa californica},
-\textit{Cercis occidentalis}, \textit{Ceanothus spp.},
-\textit{Heteromeles arbutifolia}, \textit{Sambucus mexicana},
-\textit{Eriogonum spp.}, \textit{Baccharis spp.}, \textit{Salvia
-  spp}. and others \citep[Fig.~S1][]{menz-2010-4, kremen-2015-602,
-  mgonigle-2015-x}. The mean distance between monitoring sites was
-$15$ km, and the minimum distance between sites sampled in the same
-year was 1 km.  The entire area surveyed spanned almost $300$
-km$^2$. The crop fields adjacent to all sites were similarly managed
-as intensive, high-input monoculture.
-
-Monitoring of assembling hedgerows began in 2006 and continued through
-2014. Surveys of these sites began the year before the area was
-restored. For logistical reasons, no sampling of assembling hedgerows
-was conducted in 2010. Sites were sampled between two and five times
-per year (Tables S1-S3, mean $3.4$ samples per year). In each round of
-sampling, the order in which sites were sampled was
-randomized. Surveys were conducted under sunny conditions when the
-temperature was above $21^{\circ}\mathrm{C}$ and wind speed was below
-$2.5$ meters/second.
-
-Flower-visitors to plants in hedgerows and unrestored controls were
-netted for one hour of active search time (the timer was paused when
-handling specimens). Honeybees (\textit{Apis mellifera}) were not
-collected because their abundance is determined largely by hive
-placement by bee-keepers. All other insect flower visitors that
-touched the reproductive parts of the flower were collected; however,
-here we focus only on wild bees and syrphids --- the most abundant and
-effective pollinators in the system (representing 49 and 19 percent of
-records, respectively, C.~Kremen, A.~Klein and L.~Morandin,
-unpublished data). Bee and syrphid specimens were identified to
-species (or morpho-species for some bee specimens in the genera
-\textit{Nomada} and \textit{Sphecodes}) by expert taxonomists.
-
-Quantitative networks were generated for each site through time. To
-account for the unequal number of surveys between years (Tables
-S1-S3), we use the mean of the interactions between a pair of plants
-and pollinators across surveys within a year to represent interaction
-frequency.
-
-\subsection*{Change point analysis}
-\subsubsection*{Identifying change points}
-We employed a change point detection method \citep{peel2014detecting}
-to identify fundamental reorganizations in large-scale interaction
-patterns. A change point is caused by a merge, split, fragmentation or
-formation of modules (also called compartments). Change point
-detection methods have yet to be generalized to quantitative networks,
-so for this analysis we focused on qualitative (binary)
-networks. Following \cite{peel2014detecting}, we first defined a
-probability distribution over the networks using the generalized
-hierarchical random graph model (GHRG). The GHRG model is able to
-capture both assortative and disassortative structure patterns at all
-scales in the network \citep{peel2014detecting}. A network $G$ is
-composed of vertices $V$ and edges $E$. The GHRG model decomposes the
-$N$ vertices into a series of nested groups, the relationships among
-which are represented by the dendrogram $T$. The tips of $T$ are the
-vertices of $G$, and the probability that two vertices $u$ and $v$
-connect is given by the parameter $p_r$. The probability distribution
-of the network $G$ is thus defined as:
-
-\begin{equation}
-  \label{eq:lik}
-  P(G|T,{pr}) = p_r^{E_r}(1-p_r)^{N_r-E_r}
-\end{equation}
-% 
-Where $E_r$ is the observed number of edges between vertices with the
-common ancestor $r$, and $N_r$ is the total possible edges.
-
-Using Bayesian posterior inference and techniques from phylogenetic
-tree reconstruction, we fit the GHRG model to the networks
-\citep{peel2014detecting}. This is accomplished by using a Markov
-chain Monte Carlo (MCMC) procedure to first sample the posterior
-distribution of bipartitions, from which a consensus tree is derived
-\citep{peel2014detecting}. We use $\beta$ distributions with the
-hyperparameters $\alpha=\beta=1$ to define priors for $p_r$.
-
-Once the GHRG model has been fit to the networks, we determine whether
-a change point occurred between two time slices. To detect a change
-point, we use Bayes factors to compare the fit of two models --- one
-where a change point occurred between two networks, and one where no
-change occurred. We chose a sliding window of length, $w$, of four,
-within which to find change points. Larger windows allow for more
-gradual changes, and four was the maximum possible with our eight
-years of data. Lastly, to calculate a $p$-value for the Bayes factors,
-we use parametric bootstrapping to numerically estimate the null
-distribution \citep{peel2014detecting}. We employed code published
-online by L.~Peel for the change point analysis. Analyses were
-conducted in Python 3.4.
-
-We next test whether the change points occurring in maturing hedgerows
-were a component of the assembly process or a product of environmental
-shifts that lead to network reorganizations in all types of
-communities. We model the number of change points as successes and the
-total number of years each site was sampled as trials, and use a
-generalized linear model with Binomial error to test whether the
-probability of a change point occurring varied by site type
-\citep{lme4}. We used standard techniques to determine whether the
-assumptions of the models were met for this and all subsequent
-models. For the non-assembling hedgerows and weedy field margins, only
-sites with five or greater survey years were included in this analysis
-(N=11). All statistical analyses were conducted in R 3.2.3 \citep{R}.
-
-\subsubsection*{Characteristics of species that contribute to change
-  points}
-
-To further elucidate the nature of the change points, we examine the
-characteristics of the species that contributed to interaction
-reorganization. Some species remain in relatively similar network
-positions through time, whereas others are more variable in their
-position and thus contribute more strongly to network
-reorganization. We test whether the more persistent species with the
-highest degree (number of different interaction partners) are the
-most stable in their network positions, as would be expected if the
-networks were assembling via preferential attachment.
-
-We calculate species persistence as the proportion of surveys in which
-a plant or pollinator is observed. Species observed consistently
-within and between years are thus maximally persistent. Weighted
-species degree is calculated from interaction observations from an
-extensive dataset from Yolo County (approx.~18000 interaction records)
-that included both the data included in this study and additional data
-from sites where we collected flower visitors using the same methods
-\citep{mgonigle-2015-x, ponisio2015farm}. To represent network
-position variability, we computed the coefficient of variation of
-weighted closeness centrality \citep{freeman1978centrality} at each
-site through time. Closeness centrality represents the importance of a
-space by calcuating the path lengths to other vertices (species) in
-the network \citep{freeman1978centrality}. The shorter the mean path
-length to other species, the higher the closeness centrality. We use
-linear mixed models to test whether the species closeness variability
-(log) is related to the persistence or degree of that species
-\citep{lme4, lmetest}. We included random effects for species and
-site. Because the degree and persistence of pollinators were strongly
-correlated, ($\rho = 0.071$, $p$-value $<$ $2*10^{-16}$), we include
-each explanatory variable in separate models. Plant degree and
-persistence were not significantly correlated, but we use the same
-models as we did for the pollinators for consistency.  Because an
-approximately logarithmic increase in closeness centrality, as would
-be expected with assembly by preferential attachment, would also lead
-to high variability in closeness scores, we also test whether log
-closeness centrality increases through time.
-
-\subsection*{Species and interaction turnover}
-
-%%Lauren I really liked the way you explained this part =]
-Reorganizations of network structure can be the result of species
-turnover or species changing their interaction partners (i.e.,
-re-wiring). To better understand the mechanisms underlying the
-temporal dynamics of the assembling networks, we examined patterns of
-species and interaction turnover. For example, assembling networks may
-have higher rates of pollinator turnover than non-assembling
-communities because new pollinator species are colonizing and
-establishing themselves \citep{mgonigle-2015-x}. Similarly, because
-species are turning over and pollinators are trying to maximize their
-foraging efficiency based on the species present, interactions may
-turnover more quickly than in established communities. In addition, at
-assembling hedgerows, plants that are unvisited in early years may
-appear to ``colonize'' the networks as they became more attractive
-resources and establish new interactions with pollinators.
+\citep{peel2014detecting}. In ecolo% gical communities, such shifts may
+% occur if, as new species colonize, resident species change their
+% interaction partners to optimize their foraging effort. In
+% plant-pollinator communities, theory predicts that pollinators
+% optimize their use of floral resources to reduce interspecific
+% competition and improve resource-use efficiency
+% \citep{pyke1984optimal, valdovinos2010consequences,
+%   valdovinos2013adaptive, albrecht2010plant, Bluthgen2007}. No
+% studies, however, have examined whether network changing points occur
+% during ecological network assembly, and how these changes relate to
+% the species behavior.
+
+% %% restoration
+% Understanding network assembly is particularly relevant to ecological
+% restoration, which is essentially ``applied succession''
+% \citep[e.g.,][]{parker1997scale}. In pollination systems, the time
+% since an area was restored has been shown to affect the structure of
+% networks \citep{forup-2008-742, devoto2012understanding}, suggesting
+% interactions are changing as the community develops. Understanding the
+% mechanisms of network assembly will help to guide community
+% restoration. Facilitating network restoration is especially imperative
+% in areas where species interactions provide essential ecosystem
+% services, such as crop pollination. To ensure the continued provision
+% of ecosystem services and curb biodiversity loss, it is critical to
+% restore pollinators and their interactions in agricultural
+% landscapes. To promote pollinator services in agriculture, farmers may
+% chose to plant strips of native plants along farm edges (hedgerows) to
+% help provide habitat for pollinators without removing arable land from
+% production. Hedgerows augment the richness, abundance and trait
+% diversity of pollinators in agricultural landscapes
+% \citep{morandin-2013-829, kremen-2015-602, ponisio2015farm}, and
+% promote the persistence and colonization of floral resource
+% specialists \citep{mgonigle-2015-x}. It is important to understand how
+% these new species are being incorporated into the network as the
+% community assembles, and the consequences for interaction patterns and
+% robustness.
+
+% %% Lauren I would put 10 year dataset; and I *loved* the way you
+% %% organized the ideas here!!
+% We explore the process of network development using a nine year
+% dataset of plant-pollinator community assembly following hedgerow
+% restoration in the highly simplified and intensively managed
+% agricultural landscape of California's Central Valley.  We first
+% determine whether the mechanism underlying network assembly is a build
+% up of interactions as would be predicted by preferential attachment,
+% or instead is punctuated by significant reorganizations of
+% interactions (i.e., network changing points). Even with changing
+% points in interaction organization, networks could still be assembling
+% via preferential attachment if the network reorganizations were
+% primarily driven by peripheral, temporally variable species while a
+% stable, well-connected core of species persist. We test whether the
+% species that are most variable in their network position --- and thus
+% important contributors to network reorganizations --- are less
+% persistent and connected species. To further explore the mechanisms
+% underlying the temporal dynamics of the networks, we examine patterns
+% in the species and interaction temporal turnover. Lastly, we
+% investigate whether networks are assembling toward predictable
+% interaction patterns, and the ramifications for the robustness of the
+% networks to species extinction and cascading perturbations.
+
+% %% objectives
+% % 1) Are hedgerows assembling by preferential attachment, or are there
+% % more abrupt reorganizations of network structure?
+% % 2) Who are the species responsible for interaction reorganizations?
+% % how to incorporate this? 
+% % 3) Are the changing points a product of turnover in species
+% % composition, or are more substantial changes in the verticles of the
+% % networks?
+% % 4) Is the structure of interactions changing? Does this change the
+% % robustness of the network to species extinction and/or perturbation?
+
+% \section*{Materials \& Methods}
+% \label{sec:methods}
+
+% \subsection*{Study sites and collection methods}
+% \label{sec:study-sites}
+
+% We surveyed plant-pollinator interaction networks of independent
+% assembling hedgerows communities (N=5), and of two types of
+% non-assembling communities to serve as controls: unrestored, weedy
+% field margins (N=19) and mature hedgerows (greater than 10 years since
+% planting, N=29). The sites were located in the Central Valley of
+% California in Yolo, Colusa and Solano Counties. This area is composed
+% of intensively managed agriculture --- primarily monocultures of
+% conventional row crops, vineyards and orchards. Hedgerows are ca.~3--6
+% m wide and approximately 350 m long, bordering large (ca.\
+% 30--hectare) crop fields. Hedgerows consist of native, perennial,
+% shrub and tree plantings including \textit{Rosa californica},
+% \textit{Cercis occidentalis}, \textit{Ceanothus spp.},
+% \textit{Heteromeles arbutifolia}, \textit{Sambucus mexicana},
+% \textit{Eriogonum spp.}, \textit{Baccharis spp.}, \textit{Salvia
+%   spp}. and others \citep[Fig.~S1, ][]{menz-2010-4, kremen-2015-602,
+%   mgonigle-2015-x}. The mean distance between monitoring sites was
+% $15$ km, and the minimum distance between sites sampled in the same
+% year was 1 km.  The entire area surveyed spanned almost $300$
+% km$^2$. The crop fields adjacent to all sites were similarly managed
+% as intensive, high-input monoculture.
+
+% Monitoring of assembling hedgerows began in 2006 and continued through
+% 2014. Surveys of these sites began the year before the area was
+% restored. For logistical reasons, no sampling of assembling hedgerows
+% was conducted in 2010. Sites were sampled between two and five times
+% per year (Tables S1-S3, mean $3.4$ samples per year). In each round of
+% sampling, the order in which sites were sampled was
+% randomized. Surveys were conducted under sunny conditions when the
+% temperature was above $21^{\circ}\mathrm{C}$ and wind speed was below
+% $2.5$ meters/second.
+
+% Flower-visitors to plants in hedgerows and unrestored controls were
+% netted for one hour of active search time (the timer was paused when
+% handling specimens). Honeybees (\textit{Apis mellifera}) were not
+% collected because their abundance is determined largely by hive
+% placement by bee-keepers. All other insect flower visitors that
+% touched the reproductive parts of the flower were collected; however,
+% here we focus only on wild bees and syrphids --- the most abundant and
+% effective pollinators in the system (representing 49 and 19 percent of
+% records, respectively, C.~Kremen, A.~Klein and L.~Morandin,
+% unpublished data). Bee and syrphid specimens were identified to
+% species (or morpho-species for some bee specimens in the genera
+% \textit{Nomada} and \textit{Sphecodes}) by expert taxonomists.
+
+% Quantitative networks were generated for each site through time. To
+% account for the unequal number of surveys between years (Tables
+% S1-S3), we use the mean of the interactions between a pair of plants
+% and pollinators across surveys within a year to represent interaction
+% frequency.
+
+% \subsection*{Change point analysis}
+% \subsubsection*{Identifying change points}
+% We employed a change point detection method \citep{peel2014detecting}
+% to identify fundamental reorganizations in large-scale interaction
+% patterns. A change point is caused by a merge, split, fragmentation or
+% formation of modules (also called compartments). Change point
+% detection methods have yet to be generalized to quantitative networks,
+% so for this analysis we focused on qualitative (binary)
+% networks. Following \cite{peel2014detecting}, we first defined a
+% probability distribution over the networks using the generalized
+% hierarchical random graph model (GHRG). The GHRG model is able to
+% capture both assortative and disassortative structure patterns at all
+% scales in the network \citep{peel2014detecting}. A network $G$ is
+% composed of vertices $V$ and edges $E$. The GHRG model decomposes the
+% $N$ vertices into a series of nested groups, the relationships among
+% which are represented by the dendrogram $T$. The tips of $T$ are the
+% vertices of $G$, and the probability that two vertices $u$ and $v$
+% connect is given by the parameter $p_r$. The probability distribution
+% of the network $G$ is thus defined as:
+
+% \begin{equation}
+%   \label{eq:lik}
+%   P(G|T,{pr}) = p_r^{E_r}(1-p_r)^{N_r-E_r}
+% \end{equation}
+% % 
+% Where $E_r$ is the observed number of edges between vertices with the
+% common ancestor $r$, and $N_r$ is the total possible edges.
+
+% Using Bayesian posterior inference and techniques from phylogenetic
+% tree reconstruction, we fit the GHRG model to the networks
+% \citep{peel2014detecting}. This is accomplished by using a Markov
+% chain Monte Carlo (MCMC) procedure to first sample the posterior
+% distribution of bipartitions, from which a consensus tree is derived
+% \citep{peel2014detecting}. We use $\beta$ distributions with the
+% hyperparameters $\alpha=\beta=1$ to define priors for $p_r$.
+
+% Once the GHRG model has been fit to the networks, we determine whether
+% a change point occurred between two time slices. To detect a change
+% point, we use Bayes factors to compare the fit of two models --- one
+% where a change point occurred between two networks, and one where no
+% change occurred. We chose a sliding window of length, $w$, of four,
+% within which to find change points. Larger windows allow for more
+% gradual changes, and four was the maximum possible with our eight
+% years of data. Lastly, to calculate a $p$-value for the Bayes factors,
+% we use parametric bootstrapping to numerically estimate the null
+% distribution \citep{peel2014detecting}. We employed code published
+% online by L.~Peel for the change point analysis. Analyses were
+% conducted in Python 3.4.
+
+% We next test whether the change points occurring in maturing hedgerows
+% were a component of the assembly process or a product of environmental
+% shifts that lead to network reorganizations in all types of
+% communities. We model the number of change points as successes and the
+% total number of years each site was sampled as trials, and use a
+% generalized linear model with Binomial error to test whether the
+% probability of a change point occurring varied by site type
+% \citep{lme4}. We used standard techniques to determine whether the
+% assumptions of the models were met for this and all subsequent
+% models. For the non-assembling hedgerows and weedy field margins, only
+% sites with five or greater survey years were included in this analysis
+% (N=11). All statistical analyses were conducted in R 3.2.3 \citep{R}.
+
+% \subsubsection*{Characteristics of species that contribute to change
+%   points}
+
+% To further elucidate the nature of the change points, we examine the
+% characteristics of the species that contributed to interaction
+% reorganization. Some species remain in relatively similar network
+% positions through time, whereas others are more variable in their
+% position and thus contribute more strongly to network
+% reorganization. We test whether the more persistent species with the
+% highest degree (number of different interaction partners) are the
+% most stable in their network positions, as would be expected if the
+% networks were assembling via preferential attachment.
+
+% We calculate species persistence as the proportion of surveys in which
+% a plant or pollinator is observed. Species observed consistently
+% within and between years are thus maximally persistent. Weighted
+% species degree is calculated from interaction observations from an
+% extensive dataset from Yolo County (approx.~18000 interaction records)
+% that included both the data included in this study and additional data
+% from sites where we collected flower visitors using the same methods
+% \citep{mgonigle-2015-x, ponisio2015farm}. To represent network
+% position variability, we computed the coefficient of variation of
+% weighted closeness centrality \citep{freeman1978centrality} at each
+% site through time. Closeness centrality represents the importance of a
+% space by calcuating the path lengths to other vertices (species) in
+% the network \citep{freeman1978centrality}. The shorter the mean path
+% length to other species, the higher the closeness centrality. We use
+% linear mixed models to test whether the species closeness variability
+% (log) is related to the persistence or degree of that species
+% \citep{lme4, lmetest}. We included random effects for species and
+% site. Because the degree and persistence of pollinators were strongly
+% correlated, ($\rho = 0.071$, $p$-value $<$ $2*10^{-16}$), we include
+% each explanatory variable in separate models. Plant degree and
+% persistence were not significantly correlated, but we use the same
+% models as we did for the pollinators for consistency.  Because an
+% approximately logarithmic increase in closeness centrality, as would
+% be expected with assembly by preferential attachment, would also lead
+% to high variability in closeness scores, we also test whether log
+% closeness centrality increases through time.
+
+% \subsection*{Species and interaction turnover}
+
+% %%Lauren I really liked the way you explained this part =]
+% Reorganizations of network structure can be the result of species
+% turnover or species changing their interaction partners (i.e.,
+% re-wiring). To better understand the mechanisms underlying the
+% temporal dynamics of the assembling networks, we examined patterns of
+% species and interaction turnover. For example, assembling networks may
+% have higher rates of pollinator turnover than non-assembling
+% communities because new pollinator species are colonizing and
+% establishing themselves \citep{mgonigle-2015-x}. Similarly, because
+% species are turning over and pollinators are trying to maximize their
+% foraging efficiency based on the species present, interactions may
+% turnover more quickly than in established communities. In addition, at
+% assembling hedgerows, plants that are unvisited in early years may
+% appear to ``colonize'' the networks as they became more attractive
+% resources and establish new interactions with pollinators.
 
@@
 occurring in one year are similar to those in the next year when they
-are weighted by the similarity of their constituent species (Fig.~
-S2). We develop a method to examine the temporal turnover of
-interactions with weightings based on their similarity. We followed
-the algorithm of \cite{ahn2010link} to cluster all the interactions
-(edges) hierarchically across sites and years based on their
-similarity, and build a dendrogram. The interaction similarity is
-based how may plants and pollinators (vertices) two edges share
+are weighted by the similarity of their constituent species
+(Fig.~\ref{fig:methods}). We develop a method to examine the temporal
+turnover of interactions with weightings based on their similarity. We
+followed the algorithm of \cite{ahn2010link} to cluster all the
+interactions (edges) hierarchically across sites and years based on
+their similarity, and build a dendrogram. The interaction similarity
+is based how may plants and pollinators (vertices) two edges share
 \citep{ahn2010link, kalinka2011linkcomm}. The more species edges
@@
 %$\pm$ SE, $1.834 \pm 0.6142$, $p$-value=$0.022$). 
-Nestedness and modularity did not change in any predictable pattern with assembly (Fig.~\ref{fig:baci}).
-All of the networks were significantly nested ($z$-scores $>$ 2), but not modular ($z$-scores $<$ 2, Fig.~\ref{fig:baci}).
+Nestedness and modularity did not change in any predictable pattern
+with assembly (Fig.~\ref{fig:baci}). All of the networks were
+significantly nested ($z$-scores $>$ $1.96$), but not modular ($z$-scores
+$<$ 2, Fig.~\ref{fig:baci}).
 
@@
 
-Connectance decreased as the
-community assembled (Fig.~\ref{fig:baci}, estimate of the slope of
-connectance through time $\pm$ standard error of the estimate,
-$-0.023$ $\pm$ $0.008$, $p$-value=$0.007$). Specialization also
-decreased, though the slope was not significantly
-different from zero. %(estimate of the slope of specialization through time $\pm$ SE, $-0.926$ $\pm$ $0.450$, $p$-value=$0.078$). 
-Most communities were more generalized than expected when interactions were
-randomized (Fig.~\ref{fig:baci}).
+Connectance decreased as the community assembled (Fig.~\ref{fig:baci},
+estimate of the slope of connectance through time $\pm$ standard error
+of the estimate, $-0.023$ $\pm$ $0.008$,
+$p$-value=$0.007$). Specialization also decreased, though the slope
+was not significantly different from
+zero. %(estimate of the slope of specialization through time $\pm$ SE, $-0.926$ $\pm$ $0.450$, $p$-value=$0.078$).
+Most communities were more generalized than expected when interactions
+were randomized (Fig.~\ref{fig:baci}).
 
@@
 extinction when species were removed incrementally by
-degree % (estimate
+degree or abundance % (estimate
 % of the slope of robustness through time $\pm$ SE, $6*10^{-5} \pm
-% 4*10^{-3}$, $p$-value=$0.987$)
-or abundance. % ($0.001 \pm 0.003$,
+% 4*10^{-3}$, $p$-value=$0.987$) ($0.001 \pm 0.003$,
 % $p$-value=$0.65$)
@@
 We would like to thank Paulo R.~Guimar{\~a}es Jr., Aaron Clauset,
-Matthew Hutchinson and Ana Paula A. Assis for their invaluable
-discussions and comments, and Leto Peel for help with the change point
-analysis.  We thank the growers and land owners that allowed us to
-work on their property. We also greatly appreciate the identification
-assistance of expert taxonomists Martin Hauser, Robbin Thorp and Jason
-Gibbs.  This work was supported by funding from the Army Research
-Office (W911NF-11-1-0361 to CK), the Natural Resources Conservation
-Service (CIG-69-3A75-12-253, CIG-69-3A75-9-142, CIG-68-9104-6-101 and
-WLF-69-7482-6-277 to The Xerces Society), the National Science
-Foundation (DEB-0919128 to CK), The U.S.  Department of Agriculture
-(USDA-NIFA 2012-51181-20105 to Michigan State University).  Funding
-for LCP was provided by an NSF Graduate Research Fellowship, the USDA
-NIFA Graduate Fellowship, and the Berkeley Institute for Data
-Science. Funding for MPG was provided by S{\~a}o Paulo Research
-Foundation (FAPESP, grant 2013/13319-5). We also appreciate the Santa
-Fe Institute for facilitating this international collaboration.
+Matthew Hutchinson, Rebecca Brunner, and Ana Paula A. Assis for their
+invaluable discussions and comments, and Leto Peel for help with the
+change point analysis.  We thank the growers and land owners that
+allowed us to work on their property. We also greatly appreciate the
+identification assistance of expert taxonomists Martin Hauser, Robbin
+Thorp and Jason Gibbs.  This work was supported by funding from the
+Army Research Office (W911NF-11-1-0361 to CK), the Natural Resources
+Conservation Service (CIG-69-3A75-12-253, CIG-69-3A75-9-142,
+CIG-68-9104-6-101 and WLF-69-7482-6-277 to The Xerces Society), the
+National Science Foundation (DEB-0919128 to CK), The U.S.  Department
+of Agriculture (USDA-NIFA 2012-51181-20105 to Michigan State
+University).  Funding for LCP was provided by an NSF Graduate Research
+Fellowship, the USDA NIFA Graduate Fellowship, and the Berkeley
+Institute for Data Science. Funding for MPG was provided by S{\~a}o
+Paulo Research Foundation (FAPESP, grant 2013/13319-5). We also
+appreciate the Santa Fe Institute for facilitating this international
+collaboration.
 
@@
 \bibliography{refs}
+
+\clearpage
+\begin{figure}
+  \centering
+  \includegraphics[width=.8\textwidth]{figures/scheme.pdf}
+  \caption{Diagram illustrating the analysis to examine the temporal
+    turnover of interactions weighted based on their similarity. A, B
+    and C are animal species, and Ps are plant species. The dendrogram
+    depicts the interaction similarity across years based on the
+    number of shared constituent species.}
+  \label{fig:methods}
+\end{figure}
+\clearpage
 
@@
     their degree. Each species has a consistent position in the
-    perimeter of the network across years. Asterisks indicate the year the hegdgerow
-    was planted. Before that, the sites were weedy field margins.}
+    perimeter of the network across years. Asterisks indicate the year
+    the hegdgerow was planted. Before that, the sites were weedy field
+    margins.}
   \label{fig:changePoints}
@@
   \includegraphics[width=.95\textwidth]{../analysis/changePoint/plotting/communitiesSkanky3.pdf}
-  \caption{The species module membership between network changing points. Two
-    representative assembling hedgerows are depicted. In the top
-    panel, species are grouped by module. The bottom panels visualize
-    the flow of species between modules between changing points. Each line represents the change of s species from one module to another. }
+  \caption{The species module membership between network changing
+    points. Two representative assembling hedgerows are depicted. In
+    the top panel, species are grouped by module. The bottom panels
+    visualize the flow of species between modules between changing
+    points. Each line represents the change of s species from one
+    module to another. }
   \label{fig:changePoints2}

### 2016-11-28 analysis/networkLevel/plotting/src/plotPanels_resilence.R

@@ -21,8 +21,9 @@ plot.panels <- function(){
                dats=all.alg.Con.status,
                y1="AlgCon",
                xs="ypr",
-               col.fill=col.fill,
-               col.lines=col.lines,
+               col.fill=col.white,
+               col.lines=col.white,
+               col.points=col.fill,
                plotPoints=TRUE)
     mtext("Sensitivity to \n perturbations", 2, line=4, cex=1.5)
 

### 2016-12-01 ms/ms.tex
updated cover letter, ms

--- 2016-11-28 ms/ms.tex
@@
 \usepackage[pdftex]{graphicx}
-\usepackage[numbers]{natbib} 
+\usepackage[numbers,sort&compress]{natbib} 
 \usepackage{color}
@@
 \usepackage[parfill]{parskip}
-\usepackage{graphicx}
-\usepackage{longtable}
-\usepackage{times}
-\usepackage{textcomp}
 \usepackage{xr}
-\usepackage{etoolbox}
-\usepackage{filecontents}
+\externaldocument{SI}
 \usepackage{url}
 \usepackage{rotating}
-\externaldocument{SI}
+
 \hypersetup{colorlinks=true, linkcolor=black, citecolor=black}
@@
 
-\def\title{Major interaction reorganizations punctuate the assembly of
-  pollination networks} \def\author{Lauren C.\ Ponisio$^{1,2,3}$,
-  Marilia P. Gaiarsa$^4$, Claire Kremen$^1$}
+\def\title{Major network reorganizations punctuate the assembly of
+  plant-pollinator communities}
+
+\def\author{Lauren C.\
+  Ponisio$^{1,2,3}$, Marilia P. Gaiarsa$^4$, Claire Kremen$^1$}
 
 \def\runninghead{Plant-pollinator network assembly}
-\def\keywords{community assembly, change points, specialization,
+\def\keywords{community assembly, change points, robustness,
   nestedness, modularity, bipartite, preferential attachment}
@@
   \item Abstract word count: 148
-  \item Main text word count: 5152
+  \item Main text word count: 4967
   \item Number of references: 65
@@
 \begin{abstract}  
- % The ability of communities to maintain function in the face of
-  %species extinction is related to the structure of networks. Lauren - i changed the phrasing because structure of network and network structure were too close.
-  The structure of interaction networks is related to the ability of communities to maintain function in the face of
-  species extinction. Understanding network structure and how it relates to
-  network assembly, therefore, is a priority for conservation
-  biology. Using a nine-year-dataset comprising nearly $20,000$
-  pollinator visitation records, we explore the assembly of
-  plant-pollinator communities at native plant restorations in the
-  Central Valley of California. Across years, species are highly
-  dynamic in their network position, causing community assembly to be
-  punctuated by major interaction reorganizations.%  In contrast, the
-  % non-assembling networks did not restructure as frequently.
-  % Across all communities, pollinator species were opportunistic in the
-  % flowers they visited.
-   Generalist species were the most persistent and 
-also the most variable in their network positions, contrary to
-  what is expected through prefereantial attachment theory. % High
-  % species and interaction turnover was ubiquitous across assembling
-  % and non-assembling communities, though unique interactions turned
-  % over at higher rates in assembling hedgerows as the networks
-  % continually reorganized.
-  % Nestedness of assembling networks also increased through time, but
-  The sensitivity of networks to cascading perturbations,
-  increased with assembly, at least partially due to accumulating
-  species richness. We elucidate some of the mechanisms underlying
-  plant-pollinator network assembly and restoration --- challenging
-  the widely-held hypothesis that mutualistic networks assemble
-  through preferential attachment.
+  As the world continues to lose species at an alarming rate,
+  anticipating a community's ability to resist collapse will depend on
+  the strength of interaction networks. We know little, however, about
+  how to re-assemble interacting communities through restoration, or
+  the process of ecological network assembly more generally. Using an
+  eight-year dataset comprising nearly $20,000$ pollinator visitation
+  records, we explore the assembly of plant-pollinator communities at
+  native plant restorations sites in an agricultural landscape. We
+  find that species occupy highly dynamic network positions through
+  time, causing the community assembly process to be punctuated by
+  major network reorganizations. In contrast, the non-assembling
+  networks do not restructure as frequently. The most persistent and
+  generalized species are also the most variable in their network
+  positions, contrary to what is expected by assembly following
+  preferential attachment --- a widly supported assembly
+  mechanism. Our results fundamentally alter our understanding of how
+  communities assembly and how species interactions change through
+  time while helping to inform efforts to re-assembly robust
+  communities through restoration.
 \end{abstract}
@@
 %% introduce mutualisms, restoration, network assembly
-Global change has created a severe biodiversity crisis, and as species
-are lost, so are their interactions \citep{dunn2009sixth,
-  barnosky2011has}. Because mutualistic interactions are essential for
-maintaining the diversity of their component guilds, these systems are
-particularly at risk from coextinction cascades. The nature of these
-cascades will depend on the interaction patterns within a community
-\citep{Memmott2004, Rezende2007, Bascompte2009, Thebault2010}. To
-safeguard ecological function, it has become increasingly imperative
-to aid the recovery of lost interactions and component biodviersity
-through ecological restoration, and a key restoration aim is to
-facilitate assembly of robust interaction networks
-\citep{menz-2010-4}. We know little, however, about how to re-assemble
-interacting communities through restoration, or the process of
-ecological network assembly more generally.
+When a species goes extinct, more than a species is lost. That species
+interacted with other species in its community, likely in ways that
+were important or even crucial to others' survival. When a community
+--- a unique combination of interacting species --- loses one species,
+the system becomes weaker. Global change has created a severe
+biodiversity crisis, and as species are lost, so are their
+interactions \citep{dunn2009sixth, barnosky2011has}. A community's
+ability to resist collapse will depend on the strength of each
+interaction network \citep{Memmott2004, Rezende2007, Bascompte2009,
+  Thebault2010}. To safeguard ecological function, it has become
+increasingly imperative to aid the recovery of lost interactions and
+component biodiversity and by facilitating the assembly of robust
+interaction networks via ecological restoration
+\citep{menz-2010-4}. However, we know very little about how ecological
+networks assemble in general, rendering restoration of interacting
+communities quite difficult.
 
@@
 %% there is some evidence of that. (other studies on mutualisms?)
+
+Preferential attachment, the most widely explored mechanism of network
+assembly, predicts that species entering a network are more likely to
+interact with species that are already well-connected \citep[the
+``rich-get-richer'' principle,][]{barabasi1999emergence}. In
+pollination systems --- a particularly ubiquitous mutualism
+\citep{ollerton-2011-321, klein-2007-303} --- some studies have found
+support for this assembly mechanism. Investigating the day-to-day,
+temporal assembly of a plant-pollinator network within a season,
+\cite{Olesen2008} found that phenologically new plant and pollinator
+species tended to interact with already well-connected species,
+potentially because these species are either more abundant or
+temporally persistent. Using a space-for-time substitution to study
+primary succession, \cite{albrecht2010plant} also found evidence that
+assembly along a glacier foreland occurred through preferential
+attachment. Specifically, network nestedness (i.e, a core group of
+generalists interacts with both specialist and generalist species)
+increased as the community aged \citep{albrecht2010plant}. An increase
+in nestedness could have occurred via the preferential attachment
+process, whereby specialist species attach to the well-connected,
+generalist core.
+
 Preferential attachment, the most widely explored mechanism of network
@@
 In contrast to the network build-up described by preferential
-attachment, significant reorganizations of interactions can punctuate
-assembly \citep{peel2014detecting}. Such significant reorganizations
-of interactions, or network changing points, are observed in social
-networks that respond to abrupt shifts in the behavior of interactors
-\citep{peel2014detecting}. In ecological communities, such shifts may
-occur if, as new species colonize, resident species change their
-interaction partners to optimize their foraging effort. In
-plant-pollinator communities, theory predicts that pollinators
-optimize their use of floral resources to reduce interspecific
-competition and improve resource-use efficiency
-\citep{pyke1984optimal, valdovinos2010consequences,
+attachment, significant reorganizations of interactions can also
+punctuate assembly \citep{peel2014detecting}. Such significant
+reorganizations of interactions, or network changing points, are
+observed in social networks that respond to abrupt shifts in the
+behavior of interactors \citep{peel2014detecting}. A change point is
+caused by a merge, split, fragmentation or formation of modules within
+a network. In ecological communities, such shifts may occur if, as new
+species colonize, resident species change their interaction partners
+to optimize their foraging effort. In plant-pollinator communities,
+theory predicts that pollinators optimize their use of floral
+resources to reduce interspecific competition and improve resource-use
+efficiency \citep{pyke1984optimal, valdovinos2010consequences,
   valdovinos2013adaptive, albrecht2010plant, Bluthgen2007}. No
-studies, however, have examined whether network changing points occur
-during ecological network assembly, and how these changes relate to
-the species behavior.
+studies, however, have examined if network changing points occur
+during ecological network assembly and/or how these changes relate to
+species behavior.
 
@@
 restoration, which is essentially ``applied succession''
-\citep[e.g.,][]{parker1997scale}. In pollination systems, the time
-since an area was restored has been shown to affect the structure of
-networks \citep{forup-2008-742, devoto2012understanding}, suggesting
-interactions are changing as the community develops. Understanding the
-mechanisms of network assembly will help to guide community
-restoration. Facilitating network restoration is especially imperative
-in areas where species interactions provide essential ecosystem
-services, such as crop pollination. To ensure the continued provision
-of ecosystem services and curb biodiversity loss, it is critical to
-restore pollinators and their interactions in agricultural
-landscapes. To promote pollinator services in agriculture, farmers may
-chose to plant strips of native plants along farm edges (hedgerows) to
-help provide habitat for pollinators without removing arable land from
-production. Hedgerows augment the richness, abundance and trait
-diversity of pollinators in agricultural landscapes
-\citep{morandin-2013-829, kremen-2015-602, ponisio2015farm}, and
-promote the persistence and colonization of floral resource
-specialists \citep{mgonigle-2015-x}. It is important to understand how
-these new species are being incorporated into the network as the
-community assembles, and the consequences for interaction patterns and
-robustness.
+\citep[e.g.,][]{parker1997scale}.  In pollination systems, time has
+been shown to affect the structure of networks in restored areas
+\citep{forup-2008-742, devoto2012understanding}, suggesting that
+interactions change as a community develops. Facilitating network
+restoration is especially imperative in areas where species
+interactions provide essential ecosystem services, such as crop
+pollination in agricultural landscapes. To promote pollinator services
+in agriculture, some farmers plant strips of native plants along farm
+edges (hedgerows).  By providing habitat, hedgerows augment the
+richness, abundance and trait diversity of pollinators in agricultural
+landscapes \citep{morandin-2013-829, kremen-2015-602,
+  ponisio2015farm}, and promote the persistence and colonization of
+floral resource specialists \citep{mgonigle-2015-x}. As the community
+assembles, it is important to understand how these new species are
+incorporated into the network as well as the consequences of adding
+species for interaction patterns and robustness.
 
@@
 %% organized the ideas here!!
-We explore the process of network development using a nine year
+We explore the process of network development using a nine-year
 dataset of plant-pollinator community assembly following hedgerow
 restoration in the highly simplified and intensively managed
-agricultural landscape of California's Central Valley.  We first
-determine whether the mechanism underlying network assembly is a build
-up of interactions as would be predicted by preferential attachment,
-or instead is punctuated by significant reorganizations of
-interactions (i.e., network changing points). Even with changing
-points in interaction organization, networks could still be assembling
-via preferential attachment if the network reorganizations were
-primarily driven by peripheral, temporally variable species while a
-stable, well-connected core of species persist. We test whether the
-species that are most variable in their network position --- and thus
-important contributors to network reorganizations --- are less
-persistent and connected species. To further explore the mechanisms
-underlying the temporal dynamics of the networks, we examine patterns
-in the species and interaction temporal turnover. Lastly, we
-investigate whether networks are assembling toward predictable
-interaction patterns, and the ramifications for the robustness of the
-networks to species extinction and cascading perturbations.
+agricultural landscape of California’s Central Valley. We first
+determine whether the mechanism underlying network assembly represents
+a buildup of interactions as would be predicted by preferential
+attachment or significant reorganizations of interactions (i.e.,
+network changing points). Even with changing points in interaction
+organization, networks could still assemble via preferential
+attachment if the network reorganizations were primarily driven by
+peripheral, temporally variable species while a stable, well-connected
+core of species persist. Therefore, we tested whether the species that
+are most variable in their network position --- and thus important
+contributors to network reorganizations -- are less persistent and
+connected species. To further explore the mechanisms underlying the
+temporal dynamics of the networks, also we examine patterns in the
+species and interaction temporal turnover. Finally, we investigate
+whether networks assemble toward predictable interaction patterns, and
+the ramifications for the robustness of the networks.
 
@@
 
-We surveyed plant-pollinator interaction networks of independent
-assembling hedgerows communities (N=5), and of two types of
-non-assembling communities to serve as controls: unrestored, weedy
-field margins (N=19) and mature hedgerows (greater than 10 years since
-planting, N=29). The sites were located in the Central Valley of
-California in Yolo, Colusa and Solano Counties. This area is composed
-of intensively managed agriculture --- primarily monocultures of
-conventional row crops, vineyards and orchards. Hedgerows are ca.~3--6
-m wide and approximately 350 m long, bordering large (ca.\
-30--hectare) crop fields. Hedgerows consist of native, perennial,
-shrub and tree plantings including \textit{Rosa californica},
-\textit{Cercis occidentalis}, \textit{Ceanothus spp.},
-\textit{Heteromeles arbutifolia}, \textit{Sambucus mexicana},
-\textit{Eriogonum spp.}, \textit{Baccharis spp.}, \textit{Salvia
-  spp}. and others \citep[Fig.~S1][]{menz-2010-4, kremen-2015-602,
-  mgonigle-2015-x}. The mean distance between monitoring sites was
-$15$ km, and the minimum distance between sites sampled in the same
-year was 1 km.  The entire area surveyed spanned almost $300$
-km$^2$. The crop fields adjacent to all sites were similarly managed
-as intensive, high-input monoculture.
-
-Monitoring of assembling hedgerows began in 2006 and continued through
-2014. Surveys of these sites began the year before the area was
-restored. For logistical reasons, no sampling of assembling hedgerows
-was conducted in 2010. Sites were sampled between two and five times
-per year (Tables S1-S3, mean $3.4$ samples per year). In each round of
-sampling, the order in which sites were sampled was
-randomized. Surveys were conducted under sunny conditions when the
-temperature was above $21^{\circ}\mathrm{C}$ and wind speed was below
-$2.5$ meters/second.
-
-Flower-visitors to plants in hedgerows and unrestored controls were
-netted for one hour of active search time (the timer was paused when
-handling specimens). Honeybees (\textit{Apis mellifera}) were not
-collected because their abundance is determined largely by hive
-placement by bee-keepers. All other insect flower visitors that
-touched the reproductive parts of the flower were collected; however,
-here we focus only on wild bees and syrphids --- the most abundant and
+Study sites were located in the Central Valley of California in Yolo,
+Colusa and Solano Counties. This area is composed of intensively
+managed agriculture --- primarily monocultures of conventional row
+crops, vineyards and orchards. Hedgerows border large
+(ca. $30$-hectare) crop fields and measured between $3-6$ m wide and
+approximately $350$ m long. Hedgerows consist of native, perennial,
+shrub and tree plantings \citep[Fig.~S1, ][]{menz-2010-4,
+  kremen-2015-602, mgonigle-2015-x}.
+
+We selected five farm edges to be restored as hedgerows. We monitored
+those locations before the hedgerow was planted, and tracked the
+assembly of the community for up to seven years as the hedgerow
+matured (from $2006-2014$). For controls, we concurrently monitored
+non-assembling communities within mature hedgerows (greater than $10$
+years since planting, $N=29$) and unrestored, weedy field margins
+($N=19$). The mean distance between monitoring sites was $15$ km, and
+the minimum distance between sites sampled in the same year was $1$
+km. The entire area surveyed spanned almost $300$ $km^2$. The crop
+fields adjacent to all sites were similarly managed as intensive,
+high-input monoculture.
+
+Sites were sampled between two and five times per year (Tables S1-S3,
+mean $3.4$ samples per year). In each round of sampling, the order in
+which sites were sampled was randomized. Surveys were conducted under
+sunny conditions when the temperature was above $21^{\circ}\mathrm{C}$
+and wind speed was below $2.5$ meters/second.
+
+During each sampling round, flower-visitors to plants in assembling
+hedgerows and non-assembling controls were netted for one hour of
+active search time (the timer was paused when handling specimens). The
+identify of the plant being visited was recorded for each floral
+visitor. Honeybees (\textit{Apis mellifera}) were not collected
+because their abundance is determined largely by hive placement by
+bee-keepers. All other insect flower visitors that touched the
+reproductive parts of the flower were collected; however, here we
+focus only on wild bees and syrphids --- the most abundant and
 effective pollinators in the system (representing 49 and 19 percent of
@@
 
-Quantitative networks were generated for each site through time. To
-account for the unequal number of surveys between years (Tables
-S1-S3), we use the mean of the interactions between a pair of plants
-and pollinators across surveys within a year to represent interaction
+Quantitative networks were generated for each site through
+time. Because the nuber of sampling rounds varied between years
+(Tables S1-S3), we used the mean of the interactions between a pair of
+plants and pollinators within a year to represent interaction
 frequency.
@@
 to identify fundamental reorganizations in large-scale interaction
-patterns. A change point is caused by a merge, split, fragmentation or
-formation of modules (also called compartments). Change point
-detection methods have yet to be generalized to quantitative networks,
-so for this analysis we focused on qualitative (binary)
-networks. Following \cite{peel2014detecting}, we first defined a
-probability distribution over the networks using the generalized
-hierarchical random graph model (GHRG). The GHRG model is able to
-capture both assortative and disassortative structure patterns at all
+patterns. Change point detection methods have yet to be generalized to
+quantitative networks, so for this analysis we focused on qualitative
+(binary) networks. Following \cite{peel2014detecting}, we first
+defined a probability distribution over the networks using the
+generalized hierarchical random graph model (GHRG). The GHRG model
+captures both assortative and disassortative structure patterns at all
 scales in the network \citep{peel2014detecting}. A network $G$ is
@@
 Where $E_r$ is the observed number of edges between vertices with the
-common ancestor $r$, and $N_r$ is the total possible edges.
+common ancestor $r$, and $N_r$ is the total possible edges, and the
+rest of the notation is described above.
 
@@
 where a change point occurred between two networks, and one where no
-change occurred. We chose a sliding window of length, $w$, of four,
-within which to find change points. Larger windows allow for more
-gradual changes, and four was the maximum possible with our eight
-years of data. Lastly, to calculate a $p$-value for the Bayes factors,
-we use parametric bootstrapping to numerically estimate the null
-distribution \citep{peel2014detecting}. We employed code published
-online by L.~Peel for the change point analysis. Analyses were
-conducted in Python 3.4.
-
-We next test whether the change points occurring in maturing hedgerows
-were a component of the assembly process or a product of environmental
-shifts that lead to network reorganizations in all types of
-communities. We model the number of change points as successes and the
-total number of years each site was sampled as trials, and use a
-generalized linear model with Binomial error to test whether the
-probability of a change point occurring varied by site type
-\citep{lme4}. We used standard techniques to determine whether the
-assumptions of the models were met for this and all subsequent
-models. For the non-assembling hedgerows and weedy field margins, only
-sites with five or greater survey years were included in this analysis
-(N=11). All statistical analyses were conducted in R 3.2.3 \citep{R}.
+change occurred. We chose a sliding window of length, $w$, of four
+years to detect change points. Larger windows allow for more gradual
+changes, and four was the maximum possible with our eight years of
+data. Lastly, to calculate a $p$-value for the Bayes factors, we use
+parametric bootstrapping to numerically estimate the null distribution
+\citep{peel2014detecting}. We employed code published online by
+L.~Peel for the change point analysis. Analyses were conducted in
+Python 3.4.
+
+We next tested whether the change points occurring in maturing
+hedgerows were a component of the assembly process or a product of
+environmental shifts that lead to network reorganizations in all types
+of communities. We modeled the number of change points as successes
+and the total number of sample years at each site as trials, and used
+a generalized linear model with Binomial error to test whether the
+probability of a change point occurrence varied by site type. We used
+standard techniques to determine whether the assumptions of the models
+were met for this and all subsequent models. For the non-assembling
+controls, only sites with five or greater survey years were included
+in this analysis ($N=11$). All statistical analyses were conducted in
+R 3.2.3 \citep{R}.
 
@@
 characteristics of the species that contributed to interaction
-reorganization. Some species remain in relatively similar network
-positions through time, whereas others are more variable in their
-position and thus contribute more strongly to network
-reorganization. We test whether the more persistent species with the
-highest degree (number of different interaction partners) are the
+reorganization. Some species remained in relatively similar network
+positions through time, whereas others were more variable in their
+position and thus contributed more strongly to network
+reorganization. We tested whether the more persistent species with the
+highest degree (number of different interaction partners) were the
 most stable in their network positions, as would be expected if the
@@
 
-We calculate species persistence as the proportion of surveys in which
-a plant or pollinator is observed. Species observed consistently
-within and between years are thus maximally persistent. Weighted
+We calculated species persistence as the proportion of surveys in
+which a plant or pollinator is observed. Species observed consistently
+within and between years were thus maximally persistent. Weighted
 species degree is calculated from interaction observations from an
@@
 the network \citep{freeman1978centrality}. The shorter the mean path
-length to other species, the higher the closeness centrality. We use
+length to other species, the higher the closeness centrality. We used
 linear mixed models to test whether the species closeness variability
@@
 site. Because the degree and persistence of pollinators were strongly
-correlated, ($\rho = 0.071$, $p$-value $<$ $2*10^{-16}$), we include
+correlated, ($\rho = 0.071$, $p$-value $<$ $2*10^{-16}$), we included
 each explanatory variable in separate models. Plant degree and
-persistence were not significantly correlated, but we use the same
+persistence were not significantly correlated, but we used the same
 models as we did for the pollinators for consistency.  Because an
-approximately logarithmic increase in closeness centrality, as would
-be expected with assembly by preferential attachment, would also lead
-to high variability in closeness scores, we also test whether log
-closeness centrality increases through time.
+approximately logarithmic increase in closeness centrality --- as
+would be expected with assembly by preferential attachment --- we also
+tested whether log closeness centrality increases through time.
 
@@
 
-To estimate the temporal species and interaction turnover, we use an
+To estimate the temporal species and interaction turnover, we used an
 approach similar to calculating spatial $\beta$-diversity. Instead of
@@
 higher when there are few individuals at those sites. Following
-\cite{ponisio2015farm}, we use null models that constrained species
+\cite{ponisio2015farm}, we used null models that constrained species
 richness to estimate the deviation of the observed dissimilarity from
@@
 occurring in one year are similar to those in the next year when they
-are weighted by the similarity of their constituent species (Fig.~
-S2). We develop a method to examine the temporal turnover of
-interactions with weightings based on their similarity. We followed
-the algorithm of \cite{ahn2010link} to cluster all the interactions
-(edges) hierarchically across sites and years based on their
-similarity, and build a dendrogram. The interaction similarity is
-based how may plants and pollinators (vertices) two edges share
-\citep{ahn2010link, kalinka2011linkcomm}. The more species edges
-shared in common, the shorter the branch length between them on the
-dendrogram.  We next calculated the temporal turnover of interactions
-weighted by their similarity, as approximated by ``phylogenetic''
-distance \citep{graham2008phylogenetic, picante-2010-1463}. We then
-use linear mixed models to test whether the weighted turnover of
-interactions varied between assembling and non-assembling networks,
-including site as a random effect \citep{lme4, lmetest}.
+are weighted by the similarity of their constituent species
+(Fig.~\ref{fig:methods}). We developed a method to examine the
+temporal turnover of interactions with weightings based on their
+similarity. We followed the algorithm of \cite{ahn2010link} to cluster
+all the interactions (edges) hierarchically across sites and years
+based on their similarity, and build a dendrogram. The interaction
+similarity is based how many plants and pollinators (vertices) two
+edges share \citep{ahn2010link, kalinka2011linkcomm}. The more species
+edges shared in common, the shorter the branch length between them on
+the dendrogram.  We next calculated the temporal turnover of
+interactions weighted by their similarity, as approximated by
+``phylogenetic'' distance \citep{graham2008phylogenetic,
+  picante-2010-1463}. We then used linear mixed models to test whether
+the weighted turnover of interactions varied between assembling and
+non-assembling networks, including site as a random effect
+\citep{lme4, lmetest}.
 
@@
 interactions --- would decrease as new, specialist species,
-preferentially attach to the core. Lastly, the overall level of
+preferentially attach to the core. Finally, the overall level of
 network specialization may change as the community
@@
 connected species interact \citep{almeida-neto-2008-1227}. To estimate
-modularity, we use a hierarchical clustering algorithm
+modularity, we used a hierarchical clustering algorithm
 \citep{Newman2004, csardi-2006}. We evaluate network specialization with
@@
 To test whether network modularity, nestedness, connectance or
-specialization changed linearly with assembly, we used linear mixed
+specialization changed linearly with assembly, we usedd linear mixed
 models with the descriptive network metrics as the response variable,
@@
 \cite{Memmott2004}, we simulated plant species extinction and the
-subsequent extinction cascades of pollinator species. Because the
+subsequent extinction cascades of pollinator species. Becaused the
 reproduction of plant species is facilitated by active restoration
@@
 punctuated by more interaction reorganizations than would be expected
-by environmental shifts alone that would have affected all networks
-similarly.
+by external factors such as environmental shifts that would have
+affected all networks (assembling or non-assembling) similarly.
 
@@
 estimate of the slope of closeness through time $\pm$ SE, pollinators:
-$-0.082 \pm 0.021$, $p$-value=$0.003$; plants $-0.089 \pm
-0.044$, $p$-value=$0.04$). Though statistically significant,
-the models seems to be a poor fit to the data.
+$-0.082 \pm 0.021$, $p$-value=$0.003$; plants $-0.089 \pm 0.044$,
+$p$-value=$0.04$).
 
@@
 similar across assembling hedgerows, non-assembling hedgerows and
-field margins, though mature hedgerows has marginally significantly
-less pollinator turnover than field margins (Fig.~\ref{fig:beta},
-estimate $\pm$ SE of the difference in turnover between field margins
-and mature hedgerows, $-0.051 \pm 0.026$, $p$-value=$0.049$). The turnover of plants and interactions was not significantly different between site types. When
-interactions where weighted by their similarity, both assembling and
+field margins, though mature hedgerows had marginally less pollinator
+turnover than field margins (Fig.~\ref{fig:beta}, estimate $\pm$ SE of
+the difference in turnover between field margins and mature hedgerows,
+$-0.051 \pm 0.026$, $p$-value=$0.049$). The turnover of plants and
+interactions was not significantly different between site types. When
+interactions were weighted by their similarity, both assembling and
 mature hedgerows had higher rates of turnover than field margins
@@
 as measured by the algebraic connectivity of the network, increased as
-the network assembled (Fig.~\ref{fig:rob}, estimate of the slope of
+the network assembled (Fig.~S3), estimate of the slope of
 sensitivity to cascading perturbations through time $\pm$ SE, $0.6814
@@
 We show that the temporal assembly of plant-pollinator networks
-following restoration is a highly dynamic process where interactions
-often undergo significant reorganizations, the so called changing
+following restoration is a highly dynamic process in which
+interactions often undergo significant reorganizations, or changing
 points. If these network reorganizations were a product of
 environmental forces alone, we would expect to observe the same
-changing points at the same periods, consistently across all
+changing points at the same time periods, consistently across all
 sites. However, network changing points in non-assembling communities
-are less frequent, and there are few consistent trends in when change
-points occurred across all sites. Several sites had network changing
-points between years 2009 and 2011 (Fig.~\ref{fig:changePoints}). In
-California, 2011 marked the beginning of a multi-year drought. The
-assembling hedgerows were not sampled in 2010, so disentangling
-whether the changing points are due to skipping a year of monitoring
-the assembly process or the drought is not possible. Interestingly,
-most assembling hedgerows did not undergo a significant interaction
-reorganization immediately after a hedgerow was planted (i.e., the
-transition from weedy field margin to hedgerow). This result is
-consistent with the finding that in our study system, hedgerow
-restoration takes several years to have an impact on the
-plant-pollinator communities, and with the observation that hedgerows
-do not begin to produce many flowers until $3$--$5$ years following
-planting \citep{kremen-2015-602}.
-
-In addition to finding multiple network organization changing points
-during assembly, the way in which these reorganizations occur was
-different from what would be expected from preferential attachment. In
-a preferential attachment process, we expect that the most persistent
-and high degree species would remain stable in the network core during
+are less frequent, and there are few consistent trends in the years
+when change points occurred across all sites. Several sites had
+network changing points between years 2009 and 2011
+(Fig.~\ref{fig:changePoints}). In California, 2011 marked the
+beginning of a multi-year drought. The assembling hedgerows were not
+sampled in 2010, so we cannot disentangle whether the point changes
+are due to skipping a year of monitoring or the
+drought. Interestingly, most assembling hedgerows did not undergo a
+significant interaction reorganization immediately after planting
+(i.e., the transition from weedy field margin to hedgerow). This
+result is consistent with the finding that hedgerow restoration takes
+several years to have an impact on the plant-pollinator communities in
+our study system, as well as with the observation that hedgerows do
+not begin to produce many flowers until $3-5$ years following planting
+\citep{kremen-2015-602}.
+
+In a preferential attachment process, we expect that the most
+persistent and high degree species (the species with the most unique
+interaction partners) would remain stable in the network core during
 assembly \citep{barabasi1999emergence}. Surprisingly, however, we
@@
 the only species that changed which module they were a member in
-across years in all the assembling hedgerows. Because species degree
+across years in all the assembling hedgerows. Becaused species degree
 and persistence were strongly correlated, it is difficult to
 disentangle the causal mechanism for why species with those
-characteristics are so variable in their network position. Generalized
-species may be better able to exploit the limited floral resources in
+characteristics are so variable in their network position. Species
+that can interaction with a relativly high number of interaction
+partners may be better able to exploit the limited floral resources in
 the intensively managed agriculture landscape, and thus also be the
@@
 assembling hedgerows but also the lowest pollinator
-turnover. Pollinator communities at mature hedgerows may be generally
+turnover. Pollinator communities at mature hedgerows may generally be
 more stable, but rare and/or specialized pollinators could generate
@@
 to network reorganization or species turnover, but would enhance
-weighted interaction turnover. Mature hedgerows may thus both support
+weighted interaction turnover. Mature hedgerows therefore both support
 more stable pollinator communities, while also providing resources for
@@
 
-When we explore the how network-level interaction patterns changed
-through time, we found that nestedness did increase as the community
-assembled, as would be expected if colonizing, specialist species
-preferentially attached to a central, generalist core
-\citep{albrecht2010plant}. In addition, connectance decreased, as
+When we explored how network-level interaction patterns changed
+through time, we found few patterns. The connectance decreased, as
 would be expected if the network is being colonized by specialist
-species and the overall mean number of interactions per species did
-not change. However, the frequent changing points in network
-organization, dynamic nature of species positions in the networks, and
-turnover of species and interactions all point to an assembly
-mechanism other than preferential attachment.  The stable level of
-network-level specialization through the assembly process may be due
-to the increased colonization of specialized species
-\citep{mgonigle-2015-x} accompanied by an increase in the diet breath
-of resident species. This would be expected if resident species were
-able to minimize their foraging time by expanding their diet breath as
-plant diversity increases with hedgerow maturation \citep{Waser1996,
-  pyke1984optimal, Bluthgen2007, albrecht2010plant}. Such a change in
-pollinator behavior would also explain the increase in network
-nestedness. Because so many mechanisms give rise to the same patterns
-of interaction, additional tests are necessary to assess the
-contribution of different mechanisms to community assembly.
+species \citep{mgonigle-2015-x}. Network specialization, however, did
+not change predictably with assembly. Decreasing connectance without
+an accompanying increase in specialization would be possible if the
+increased colonization of specialized species was accompanied by an
+increase in the diet breath of resident species. This would be
+expected if resident species were able to minimize their foraging time
+by expanding their diet breath as plant diversity increases with
+hedgerow maturation \citep{Waser1996, pyke1984optimal, Bluthgen2007,
+  albrecht2010plant}.
 
 Interestingly, however, the changes in network patterns associated
-with assembly did not affect the robustness of hedgerow communities to
-species loss. This is particularly surprising given the observed
-increase in nestedness, which is often associated with an increase in
-robustness to extinction \citep{Memmott2004}. Perhaps assembling
-hedgerows have yet to reach sufficient levels of nestedness to realize
-the benefits nestedness confers. Nestedness of the assembling
-hedgerows, however, did not asymptote within the eight years following
-restoration that the sites were surveyed, so hedgerow networks may
-eventually reach sufficient levels of nestedness to gain the
-robustness advantage.
-
-% Lauren - I changed the first part of this paragraph but I really
-% don't thinks it is that confusing. The problem is that it is indeed
-% very theoretical, but it is impossible to make our point without
-% being too technical.. I tried to decrease the use of jargons.
-Contrary to the general restoration goals, networks' susceptibility to
-cascading perturbations increased as the communities assembled. We
-measured vulnerability to cascading perturbations as the algebraic
-connectivity of the network: the greater the algebraic connectivity,
-the more vulnerable a community is to cascading
-perturbations. Algebraic connectivity is positively related to both
-the number of species and connectance of a network
-\citep{gibert2013spatial}. The increase in and plant and pollinator
-richness following restoration is thus at least partially responsible
-for the increase in response to cascading effects.  Algebraic
-connectivity is also positivly correlated with the number in
-interactions, or connectance of a network \citep{gibert2013spatial},
-We, however, observed a decrease in connectance, suggesting
-topological characteristics of the networks beyond just species
-richness and connectance are needed to explain the increased
-sensitivity to perturbations spreading. These hedgerows were designed
-to provide floral resources to the largest number of pollinators
-across the growing season \citep{menz-2010-4}. The generalized nature
-of the floral community may explain why the networks tended to be more
-generalized than expected if interactions were randomly distributed
-across species (Fig.~\ref{fig:baci}). In addition, the design of the
-hedgerow plantings may have facilitated the emergence of a single,
-highly connected module in all of the networks (see
-\ref{fig:changePoints2} for examples). This network configuration
-results in short path lengths (the distance between species in a
-network based on their shared partners), and thus, a perturbation in
-one species can more easily spread to other species. In order to
-promote more resilient communities, future restoration efforts should
-explore designing floral communities to promote more interaction
-partitioning using, for example, algorithms to optimize different
-network properties based on prior knowledge of pollinator floral
-preferences \citep{mgonigle2016tool}, and on desired network
+with assembly did not affect the hedgerow network robustness to
+species extinction or susceptibility to cascading perturbations. The
+hedgerows were designed to provide floral resources for the largest
+number of pollinators across the growing season
+\citep{menz-2010-4}. The generalized nature of the floral community
+may explain why the networks tended to be more generalized than
+expected if interactions were randomly distributed across species
+(Fig.~\ref{fig:baci}). In addition, the design of the hedgerow
+plantings may have facilitated the emergence of a single, highly
+connected module in all of the networks (see \ref{fig:changePoints2}
+for examples). This network configuration results in short path
+lengths (the distance between species in a network based on their
+shared partners), and thus, a perturbation in one species can more
+easily spread to other species. These networks are also vulnerable to
+the extinction of highly connected species \citep{Burgos2007}. In
+order to promote more resilient communities, future restoration
+efforts should explore designing floral communities to promote more
+interaction partitioning using, for example, algorithms to optimize
+different network properties based on prior knowledge of pollinator
+floral preferences \citep{mgonigle2016tool}, and on desired network
 architectures that renders them more robust both to species loss and
@@
 
-Plant-pollinator networks, in general, are highly dynamic, with high
+In general, plant-pollinator networks are highly dynamic, with high
 turnover of species and interactions both within and between seasons
-\citep{Burkle2011}. Though our non-assembling communities experience
+\citep{Burkle2011}. Though our non-assembling communities experienced
 fewer network reorganizations than the assembling hedgerows, $82\%$ of
@@
 predictions, along with empirical exploration of network
-assembly. Plant-pollinator communities and mutualisms broadly are
-vital for biodiversity maintenance and essential ecosystem service
-provision. We must therefore understand the processes underlying their
-assembly to facilitate restoration and conservation.
+assembly. Plant-pollinator communities and mutualisms are vital for
+biodiversity maintenance and essential ecosystem service provision. We
+must therefore understand the processes underlying their assembly to
+facilitate restoration and conservation.
 
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=.5\textwidth]{../analysis/networkLevel/figures/robustness.pdf}
-  \caption{The robustness of networks to species extinction did not
-    change with network assembly, but the sensitivity to cascading
-    perturbations increased. The robustness to species extinction is
-    measured by incrementally removing species by degree, though
-    removing species by abundance did not yield qualitatively
-    different results. The robustness of networks to cascading
-    perturbations is measured as the algebraic connectivity, the
-    second smallest eigenvalue of the Laplacian matrix. Points are the
-    value for each site at each year of assembly. The solid line
-    indicates the mean slope estimate and the dashed lines are the
-    $95\%$ confidence intervals around the estimate.}
-  \label{fig:rob}
-\end{figure}
-\clearpage
-
 \end{document}

### 2016-12-01 ms/ms.tex
changed the abstract and the first paragraph of intro

@@
 \begin{abstract}
-  As the world continues to lose species at an alarming rate,
-  anticipating a community's ability to resist collapse will depend on
-  the strength of interaction networks. We know little, however, about
-  how to re-assemble interacting communities through restoration, or
-  the process of ecological network assembly more generally. Using an
-  eight-year dataset comprising nearly $20,000$ pollinator visitation
-  records, we explore the assembly of plant-pollinator communities at
-  native plant restorations sites in an agricultural landscape. We
-  find that species occupy highly dynamic network positions through
-  time, causing the community assembly process to be punctuated by
-  major network reorganizations. In contrast, the non-assembling
-  networks do not restructure as frequently. The most persistent and
-  generalized species are also the most variable in their network
-  positions, contrary to what is expected by assembly following
-  preferential attachment --- a widly supported assembly
-  mechanism. Our results fundamentally alter our understanding of how
-  communities assembly and how species interactions change through
-  time while helping to inform efforts to re-assembly robust
-  communities through restoration.
+
+Species are being lost at an alarming rate and it is imperative to understand how communities assemble if we are to prevent their collapse or restore lost interactions.
+We know little, however, about how to re-assemble interacting communities through restoration, or the process of ecological network assembly more generally.
+Using an eight-year dataset comprising nearly $20,000$ pollinator visitation records, we explore the assembly of plant-pollinator communities at native plant restorations sites in an agricultural landscape.
+We find that species occupy highly dynamic network positions through time, causing the assembly process to be punctuated by major network reorganizations.
+The most persistent and generalized species are also the most variable in their network positions, contrary to what assembly through preferential attachment predicts.
+Our results fundamentally alter our understanding of how communities assembly and how species interactions change through time while helping to inform efforts to re-assemble robust communities through restoration.
+%It was re-assembly but I think it is re-assemble no?
+
 \end{abstract}
@@
 %% introduce mutualisms, restoration, network assembly
-When a species goes extinct, more than a species is lost. That species
-interacted with other species in its community, likely in ways that
-were important or even crucial to others' survival. When a community
---- a unique combination of interacting species --- loses one species,
-the system becomes weaker. Global change has created a severe
+%When a species goes extinct, more than a species is lost. That species
+%interacted with other species in its community, likely in ways that
+%were important or even crucial to others' survival. When a community
+%--- a unique combination of interacting species --- loses one species,
+%the system becomes weaker. 
+Species in ecological communities are linked through ecological interactions. Global change, however, has created a severe
 biodiversity crisis, and as species are lost, so are their
 interactions \citep{dunn2009sixth, barnosky2011has}. A community's
-ability to resist collapse will depend on the strength of each
-interaction network \citep{Memmott2004, Rezende2007, Bascompte2009,
-  Thebault2010}. To safeguard ecological function, it has become
+ability to resist collapse will depend on the community's robustness to species and interaction loss \citep{Memmott2004, Rezende2007, Bascompte2009,
+  Thebault2010}. Furthermore, it has been show that biodiversity has a pivotal role in maintaining ecosystems services \citep{cardinale-2012-59}. Thus, to safeguard ecological function, it has become
 increasingly imperative to aid the recovery of lost interactions and
@@
 process, whereby specialist species attach to the well-connected,
-generalist core.
-
-Preferential attachment, the most widely explored mechanism of network
-assembly, \citep{barabasi1999emergence}, predicts that species
-entering a network are more likely to interact with species that are
-already well-connected \citep[the ``rich-get-richer''
-principle,][]{barabasi1999emergence}. In pollination systems --- a
-particularly ubiquitous mutualism \citep{ollerton-2011-321,
-  klein-2007-303} --- some studies have found support for this
-assembly mechanism. Investigating the day-to-day, temporal assembly of
-a plant-pollinator network within a season, \cite{Olesen2008} found
-that phenologically new plant and pollinator species tended to
-interact with already well-connected species, potentially because
-these species are either more abundant or more temporally
-persistent. In addition, using a space-for-time substitution to study
-primary succession along a glacier foreland, \cite{albrecht2010plant}
-also found some evidence that assembly occurred through preferential
-attachment. Specifically, network nestedness (i.e, a core group of
-generalists interacts with both specialist and generalist species)
-increased as the community aged \citep{albrecht2010plant}. An
-increase in nestedness could result from the preferential attachment
-process where specialist species attach to the well-connected,
 generalist core.

### 2016-12-02 ms/coverLetter/coverLetter.tex
updated cover letter, ms
version sent to ecol let

@@
 
-  W are pleased to submit our manuscript \textit{Major interaction
+  We are pleased to submit our manuscript \textit{Major interaction
     reorganizations punctuate the assembly of pollination networks}
@@
   communities assembly and how species interactions change through
-  time. Our insights will help to inform effrots to re-assembly robust
+  time. Our insights will help to inform efforts to re-assemble robust
   communities through restoration.  We believe that these exciting
@@
 
-  % Our manuscript is original and was carried out fully by the authors.
-  % All authors agree with the contents of the manuscript.  This
-  % manuscript is not published, nor is it in consideration for
-  % publication elsewhere.  All research not of the authors' is fully
-  % acknowledged.  The authors declare no conflict of interest. All
-  % appropriate ethical standards were followed. 
   Thank you for reviewing our manuscript and we hope you will find it
@@
   Lauren C. Ponisio, PhD\\
+  Marilia Gaiarsa, PhD candidate\\
   Claire Kremen, Professor
-%is there a reason why I am not here? Just curious =0) 
+
 \end{letter}

### 2016-12-02 ms/hardcoded/ms.tex
version sent to ecol let

--- 2016-12-01 ms/ms.tex
@@
 \usepackage[pdftex]{graphicx}
-\usepackage[numbers,sort&compress]{natbib} 
+\usepackage[numbers]{natbib}
+% \usepackage{natbib}
 \usepackage{color}
@@
 \usepackage{lscape}
+\usepackage{array}
+\usepackage{paralist}
 \usepackage{fancyhdr}
-\usepackage[normalem]{ulem} 
-\usepackage{hyperref}
-\usepackage[parfill]{parskip}
+\usepackage{graphicx}
+\usepackage{longtable}
+\usepackage{times}
+\usepackage{textcomp}
 \usepackage{xr}
-\externaldocument{SI}
+\usepackage{etoolbox}
+\usepackage{filecontents}
 \usepackage{url}
 \usepackage{rotating}
-
-\hypersetup{colorlinks=true, linkcolor=black, citecolor=black}
+\externaldocument{SI}
+
+\makeatletter
+\def\@biblabel#1{#1.\\}
+\makeatother
+
+\parindent=2em
+\setlength{\parskip}{1ex}
+
 \RequirePackage{lineno}
-
-\newcommand{\flagged}[1] {
-  \textcolor{blue}{#1}
-}
 
@@
 \def\keywords{community assembly, change points, robustness,
-  nestedness, modularity, bipartite, preferential attachment}
+  nestedness, modularity, restoration, mutualism, preferential attachment}
 
@@
   \item Submitted as a Letter
-  \item Abstract word count: 148
-  \item Main text word count: 4967
-  \item Number of references: 65
+  \item Abstract word count: 136
+  \item Main text word count: 4680
+  \item Number of references: 67
   \item Number of figures: 6
@@
 \begin{abstract}  
-
-Species are being lost at an alarming rate and it is imperative to understand how communities assemble if we are to prevent their collapse or restore lost interactions.
-We know little, however, about how to re-assemble interacting communities through restoration, or the process of ecological network assembly more generally.
-Using an eight-year dataset comprising nearly $20,000$ pollinator visitation records, we explore the assembly of plant-pollinator communities at native plant restorations sites in an agricultural landscape.
-We find that species occupy highly dynamic network positions through time, causing the assembly process to be punctuated by major network reorganizations.
-The most persistent and generalized species are also the most variable in their network positions, contrary to what assembly through preferential attachment predicts.
-Our results fundamentally alter our understanding of how communities assembly and how species interactions change through time while helping to inform efforts to re-assemble robust communities through restoration.
-%It was re-assembly but I think it is re-assemble no?
-
+  Species and interactions are being lost at alarming rates and it is
+  imperative to understand how communities assemble if we are to
+  prevent their collapse and restore lost interactions. Using an
+  eight-year dataset comprising nearly $20,000$ pollinator visitation
+  records, we explore the assembly of plant-pollinator communities at
+  native plant restorations sites in an agricultural landscape. We
+  find that species occupy highly dynamic network positions through
+  time, causing the assembly process to be punctuated by major network
+  reorganizations. The most persistent and generalized pollinator
+  species are also the most variable in their network positions,
+  contrary to what preferential attachment --- the most widely studied
+  theory of network assembly --- predicts. Our results fundamentally
+  alter our understanding of how communities assembly and how species
+  interactions change through time while helping to inform efforts to
+  re-assemble robust communities through restoration.
 \end{abstract}
@@
 
-%% introduce mutualisms, restoration, network assembly
-%When a species goes extinct, more than a species is lost. That species
-%interacted with other species in its community, likely in ways that
-%were important or even crucial to others' survival. When a community
-%--- a unique combination of interacting species --- loses one species,
-%the system becomes weaker. 
-Species in ecological communities are linked through ecological interactions. Global change, however, has created a severe
-biodiversity crisis, and as species are lost, so are their
-interactions \citep{dunn2009sixth, barnosky2011has}. A community's
-ability to resist collapse will depend on the community's robustness to species and interaction loss \citep{Memmott2004, Rezende2007, Bascompte2009,
-  Thebault2010}. Furthermore, it has been show that biodiversity has a pivotal role in maintaining ecosystems services \citep{cardinale-2012-59}. Thus, to safeguard ecological function, it has become
-increasingly imperative to aid the recovery of lost interactions and
-component biodiversity and by facilitating the assembly of robust
-interaction networks via ecological restoration
+Species in ecological communities are linked through ecological
+interactions. As the world continues to lose species at an alarming
+rate \citep{dunn2009sixth, barnosky2011has}, anticipating a
+community's ability to resist collapse will depend on the strength of
+each interaction network and the species within it \citep{Memmott2004,
+  Rezende2007, Bascompte2009, Thebault2010}. Furthermore, it has been
+show that biodiversity has a pivotal role in maintaining ecosystems
+services \citep{cardinale-2012-59}. Thus, to safeguard ecological
+function, it has become increasingly imperative to aid the recovery of
+lost interactions and component biodiversity and by facilitating the
+assembly of robust interaction networks via ecological restoration
 \citep{menz-2010-4}. However, we know very little about how ecological
@@
 communities quite difficult.
-
-%% most people think network assemble via preferential attachment, and
-%% there is some evidence of that. (other studies on mutualisms?)
 
@@
 
-%% there are other ways to assemble that are not preferentia;
-%% attchment, particularly vis big reorganizations of interactions
-%% Lauren I love this paragraph!
-In contrast to the network build-up described by preferential
-attachment, significant reorganizations of interactions can also
-punctuate assembly \citep{peel2014detecting}. Such significant
-reorganizations of interactions, or network changing points, are
-observed in social networks that respond to abrupt shifts in the
-behavior of interactors \citep{peel2014detecting}. A change point is
+Significant reorganizations of interactions --- change points --- can
+also punctuate assembly \citep{peel2014detecting}. A change point is
 caused by a merge, split, fragmentation or formation of modules within
-a network. In ecological communities, such shifts may occur if, as new
-species colonize, resident species change their interaction partners
-to optimize their foraging effort. In plant-pollinator communities,
-theory predicts that pollinators optimize their use of floral
-resources to reduce interspecific competition and improve resource-use
-efficiency \citep{pyke1984optimal, valdovinos2010consequences,
-  valdovinos2013adaptive, albrecht2010plant, Bluthgen2007}. No
-studies, however, have examined if network changing points occur
-during ecological network assembly and/or how these changes relate to
-species behavior.
-
-%% restoration
+a network. Such interaction reorganizations are observed in social
+networks responding to abrupt shifts in the behavior of interactors
+\citep{peel2014detecting}. In ecological communities, such shifts may
+occur if, as new species colonize, resident species change their
+interaction partners to optimize their foraging effort
+\citep{pyke1984optimal, Bluthgen2007, valdovinos2010consequences,
+  albrecht2010plant, valdovinos2013adaptive}. For preferential
+attachment to be at play, change points would need to be driven
+primarily by peripheral, temporally variable species while a stable,
+well-connected core of species remained stable in their
+interactions. No studies, however, have examined if network changing
+points occur during ecological network assembly and/or how these
+changes relate to species behavior.
+
 Understanding network assembly is particularly relevant to ecological
@@
 
-%% Lauren I would put 10 year dataset; and I *loved* the way you
-%% organized the ideas here!!
-We explore the process of network development using a nine-year
+We explore the process of network development using a eight-year
 dataset of plant-pollinator community assembly following hedgerow
-restoration in the highly simplified and intensively managed
-agricultural landscape of California’s Central Valley. We first
-determine whether the mechanism underlying network assembly represents
-a buildup of interactions as would be predicted by preferential
-attachment or significant reorganizations of interactions (i.e.,
-network changing points). Even with changing points in interaction
-organization, networks could still assemble via preferential
-attachment if the network reorganizations were primarily driven by
-peripheral, temporally variable species while a stable, well-connected
-core of species persist. Therefore, we tested whether the species that
-are most variable in their network position --- and thus important
-contributors to network reorganizations -- are less persistent and
-connected species. To further explore the mechanisms underlying the
-temporal dynamics of the networks, also we examine patterns in the
-species and interaction temporal turnover. Finally, we investigate
-whether networks assemble toward predictable interaction patterns, and
-the ramifications for the robustness of the networks.
-
-%% objectives
-% 1) Are hedgerows assembling by preferential attachment, or are there
-% more abrupt reorganizations of network structure?
-% 2) Who are the species responsible for interaction reorganizations?
-% how to incorporate this? 
-% 3) Are the changing points a product of turnover in species
-% composition, or are more substantial changes in the verticles of the
-% networks?
-% 4) Is the structure of interactions changing? Does this change the
-% robustness of the network to species extinction and/or perturbation?
+installment in the highly simplified and intensively managed
+agricultural landscape of California's Central Valley. We first
+determine whether network assembly was punctuated by significant
+reorganizations of interactions. We next tested whether the species
+that are most variable in their network position --- and thus
+important contributors to network reorganizations --- are less
+persistent and connected species, as would be expected by preferential
+attachment. To further explore the mechanisms underlying the temporal
+dynamics of the networks, also we examine patterns in the species and
+interaction temporal turnover. Finally, we investigate whether
+networks assemble toward predictable interaction patterns, and the
+ramifications for the robustness of the networks.
 
@@
 Quantitative networks were generated for each site through
-time. Because the nuber of sampling rounds varied between years
+time. Because the number of sampling rounds varied between years
 (Tables S1-S3), we used the mean of the interactions between a pair of
@@
 tree reconstruction, we fit the GHRG model to the networks
-\citep{peel2014detecting}. This is accomplished by using a Markov
-chain Monte Carlo (MCMC) procedure to first sample the posterior
-distribution of bipartitions, from which a consensus tree is derived
-\citep{peel2014detecting}. We use $\beta$ distributions with the
-hyperparameters $\alpha=\beta=1$ to define priors for $p_r$.
-
-Once the GHRG model has been fit to the networks, we determine whether
-a change point occurred between two time slices. To detect a change
-point, we use Bayes factors to compare the fit of two models --- one
+\citep{peel2014detecting}. We used a Markov chain Monte Carlo (MCMC)
+procedure to first sample the posterior distribution of bipartitions,
+from which a consensus tree was derived \citep{peel2014detecting}. We
+used $\beta$ distributions with the hyperparameters $\alpha=\beta=1$
+to define priors for $p_r$.
+
+Once the GHRG model was fit to the networks, we determined whether a
+change point occurred between two time slices. To detect a change
+point, we used Bayes factors to compare the fit of two models --- one
 where a change point occurred between two networks, and one where no
@@
 changes, and four was the maximum possible with our eight years of
-data. Lastly, to calculate a $p$-value for the Bayes factors, we use
+data. Lastly, to calculate a $p$-value for the Bayes factors, we used
 parametric bootstrapping to numerically estimate the null distribution
@@
 
-To further elucidate the nature of the change points, we examine the
+To further elucidate the nature of the change points, we examined the
 characteristics of the species that contributed to interaction
-reorganization. Some species remained in relatively similar network
-positions through time, whereas others were more variable in their
-position and thus contributed more strongly to network
 reorganization. We tested whether the more persistent species with the
@@
 which a plant or pollinator is observed. Species observed consistently
-within and between years were thus maximally persistent. Weighted
-species degree is calculated from interaction observations from an
-extensive dataset from Yolo County (approx.~18000 interaction records)
-that included both the data included in this study and additional data
-from sites where we collected flower visitors using the same methods
+within and between years were thus maximally persistent. Species
+degree was calculated from interaction observations from an extensive
+dataset from Yolo County (approx.~18000 interaction records) that
+included both the data included in this study and additional data from
+sites where we collected flower visitors using the same methods
 \citep{mgonigle-2015-x, ponisio2015farm}. To represent network
@@
 site through time. Closeness centrality represents the importance of a
-space by calcuating the path lengths to other vertices (species) in
+species by calculating the path lengths to other vertices (species) in
 the network \citep{freeman1978centrality}. The shorter the mean path
@@
 would be expected with assembly by preferential attachment --- we also
-tested whether log closeness centrality increases through time.
+tested whether log closeness centrality increased through time.
 
@@
 
-%%Lauren I really liked the way you explained this part =]
 Reorganizations of network structure can be the result of species
@@
 all the interactions (edges) hierarchically across sites and years
-based on their similarity, and build a dendrogram. The interaction
+based on their similarity, and built a dendrogram. The interaction
 similarity is based how many plants and pollinators (vertices) two
@@
 connectance --- the proportion of observed out of possible
-interactions --- would decrease as new, specialist species,
-preferentially attach to the core. Finally, the overall level of
-network specialization may change as the community
-assembles. Network-level specialization will increase if specialist
-species colonize the network or species begin to limit their
-interaction niche breath as the network assembles
-\citep{bluthgen-2006-9}.
+interactions --- would also decrease as new, specialist species,
+preferentially attach to the core. Finally, network-level
+specialization will increase if specialist species colonize the
+network or species begin to limit their interaction niche breath as
+the network assembles \citep{bluthgen-2006-9}.
 
@@
 modularity, we used a hierarchical clustering algorithm
-\citep{Newman2004, csardi-2006}. We evaluate network specialization with
-the metric H2, which estimates the deviation of the observed
+\citep{Newman2004, csardi-2006}. We evaluated network specialization
+with the metric H2, which estimates the deviation of the observed
 interaction frequency between plants and pollinators from a null
@@
 To test whether network modularity, nestedness, connectance or
-specialization changed linearly with assembly, we usedd linear mixed
+specialization changed linearly with assembly, we used linear mixed
 models with the descriptive network metrics as the response variable,
@@
 of interaction possible, so we also examined the change in plant and
-pollinator species richness through time. We employ generalized linear
-mixed models with Poisson error to model richness \citep{lme4}. We
-scaled explanatory variables.
+pollinator species richness through time. We employed generalized
+linear mixed models with Poisson error to model richness
+\citep{lme4}. We scaled explanatory variables.
 
 \subsection*{Network robustness}
-Lastly, we tested whether the changes in interaction patterns
+Finally, we tested whether the changes in interaction patterns
 associated with network assembly affect the robustness of the network
@@
 \cite{Memmott2004}, we simulated plant species extinction and the
-subsequent extinction cascades of pollinator species. Becaused the
+subsequent extinction cascades of pollinator species. Because the
 reproduction of plant species is facilitated by active restoration
@@
 extinction curve is an estimate of network robustness
-\citep{Memmott2004, bipartite}.
+\citep{Memmott2004, Burgos2007, bipartite}.
 
@@
 \citep{fiedler1973algebraic} --- as a proxy for network robustness.
-% (Gaiarsa et al., in prep)
 Algebraic connectivity relates to how difficult it is to turn a
@@
 note that both robustness and algebraic connectivity assume that the
-network is static. They do not account for the ability of species to
-alter their interaction depending on circumstances and the resource
+network is static --- they do not account for the ability of species
+to alter their interaction depending on circumstances and the resource
 availability.
@@
 not significantly related to network position variability
-(Fig.~\ref{fig:cv}). % estimate of the slope of closeness variability and
-% persistence $\pm$ SE, $-2.063 \pm 3.091$, $p$-value=$0.5$; slope of
-% closeness variability and degree, $0.0018 \pm 0.002$,
-% $p$-value=$0.3$).
-The variability of species network position was not the result of
-closeness linearly increasing through time, and, in fact, plant and
-pollinator closeness decreased slightly through time (Fig.~S2,
-estimate of the slope of closeness through time $\pm$ SE, pollinators:
-$-0.082 \pm 0.021$, $p$-value=$0.003$; plants $-0.089 \pm 0.044$,
-$p$-value=$0.04$).
+(Fig.~\ref{fig:cv}).  The variability of species network position was
+not the result of closeness linearly increasing through time, and, in
+fact, plant and pollinator closeness decreased slightly through time
+(Fig.~S2, estimate of the slope of closeness through time $\pm$ SE,
+pollinators: $-0.082 \pm 0.021$, $p$-value=$0.003$; plants $-0.089 \pm
+0.044$, $p$-value=$0.04$).
 
@@
 \subsection*{Network structure}
-%Network nestedness significantly increased with assembly
-%(Fig.~\ref{fig:baci}, estimate of the slope of nestedness through time
-%$\pm$ SE, $1.834 \pm 0.6142$, $p$-value=$0.022$). 
-Nestedness and modularity did not change in any predictable pattern
-with assembly (Fig.~\ref{fig:baci}). All of the networks were
-significantly nested ($z$-scores $>$ $1.96$), but not modular ($z$-scores
-$<$ 2, Fig.~\ref{fig:baci}).
-
-%though the slope was not significantly different from
-%zero.(estimate
-% of the slope of modularity through time $\pm$ standard error of the
-% estimate, $-0.524$ $\pm$ $0.295$, $p$-value=$0.124$).
-%In addition, none of the networks were significantly modular
-% Fig.~\ref{fig:baci}).
-
-Connectance decreased as the community assembled (Fig.~\ref{fig:baci},
-estimate of the slope of connectance through time $\pm$ standard error
-of the estimate, $-0.023$ $\pm$ $0.008$,
-$p$-value=$0.007$). Specialization also decreased, though the slope
-was not significantly different from
-zero. %(estimate of the slope of specialization through time $\pm$ SE, $-0.926$ $\pm$ $0.450$, $p$-value=$0.078$).
-Most communities were more generalized than expected when interactions
-were randomized (Fig.~\ref{fig:baci}).
+
+Nestedness, modularity and specialization did not change predictably
+assembly (Fig.~\ref{fig:baci}). All of the networks were significantly
+nested ($z$-scores $>$ $2$), but not modular ($z$-scores $<$ 2,
+Fig.~\ref{fig:baci}). Most communities were more generalized than
+expected when interactions were randomized
+(Fig.~\ref{fig:baci}). Connectance decreased as the community
+assembled (Fig.~\ref{fig:baci}, estimate of the slope of connectance
+through time $\pm$ standard error of the estimate, $-0.023$ $\pm$
+$0.008$, $p$-value=$0.007$).
 
@@
 \subsection*{Network robustness}
-Assembly did not affect network robustness to species
-extinction when species were removed incrementally by
-degree or abundance % (estimate
-% of the slope of robustness through time $\pm$ SE, $6*10^{-5} \pm
-% 4*10^{-3}$, $p$-value=$0.987$) ($0.001 \pm 0.003$,
-% $p$-value=$0.65$)
-In contrast, the sensitivity of networks to cascading perturbations,
-as measured by the algebraic connectivity of the network, increased as
-the network assembled (Fig.~S3), estimate of the slope of
-sensitivity to cascading perturbations through time $\pm$ SE, $0.6814
-\pm 0.272$, $p$-value=$0.042$).
+Assembly did not affect network robustness to species extinction when
+species were removed incrementally by degree or abundance
+(Fig.~S3). Similarly, the sensitivity of networks to cascading
+perturbations, as measured by the algebraic connectivity of the
+network, did not change predictably as the community assembled
+(Fig.~S3).
 
@@
 the only species that changed which module they were a member in
-across years in all the assembling hedgerows. Becaused species degree
+across years in all the assembling hedgerows. Because species degree
 and persistence were strongly correlated, it is difficult to
@@
 characteristics are so variable in their network position. Species
-that can interaction with a relativly high number of interaction
-partners may be better able to exploit the limited floral resources in
-the intensively managed agriculture landscape, and thus also be the
-most persistent \citep[in ant-plant mutualisms,
-][]{diaz2010changes}. More persistent species usually have longer
-phenologies, so they can visit many different flowers, resulting in a
-higher degree \citep{Vazquez2009, fort2016abundance}. Either way, our
-result suggests that adaptable species can change their network
-position to utilize the most advantageous floral resources available,
-which may depend on both the other pollinator species that are present
-and the state of the plant community \citep{macleod2016measuring,
-  gomez2006ecological, Waser1996}. Thus given the opportunity and
-ability to use different resources, species will often change their
-network positions \citep{macleod2016measuring}.
+that can interact with a relatively high number of partners may be
+better able to exploit the limited floral resources in the intensively
+managed agriculture landscape, and thus also be the most persistent
+\citep[in ant-plant mutualisms, ][]{diaz2010changes}. More persistent
+species usually have longer phenologies, so they can visit many
+different flowers, resulting in a higher degree \citep{Vazquez2009,
+  fort2016abundance}. Either way, our result suggests that adaptable
+species can change their network position to utilize the most
+advantageous floral resources available, which may depend on both the
+other pollinator species that are present and the state of the plant
+community \citep{Waser1996, gomez2006ecological,
+  macleod2016measuring}. Thus given the opportunity and ability to use
+different resources, species will often change their network positions
+\citep{macleod2016measuring}.
 
@@
 by expanding their diet breath as plant diversity increases with
-hedgerow maturation \citep{Waser1996, pyke1984optimal, Bluthgen2007,
+hedgerow maturation \citep{pyke1984optimal, Waser1996, Bluthgen2007,
   albrecht2010plant}.
@@
 
-% The relationship between diversity and stability in networks
-% has been the subject of considerable debate
-% \citep[e.g.,][]{may1972will, pimm1984complexity,
-%   montoya2006ecological}. Our results provide one of the few empirical
-% examples of how restoring species diversity contributes to enhancing
-% network stability.
-
-% providing further evidence that hedgerows are a valuable tool for
-% promoting species conservation and ecosystem provision in agricultural
-% areas \citep{mgonigle-2015-x, ponisio2015farm,
-%   kremen-2015-602}.
-
 In general, plant-pollinator networks are highly dynamic, with high
@@
 field margins and $40\%$ of mature hedgerows underwent at least one
-changing point in network structure. Pollinators are also highly
+changing point in network structure. Pollinators are highly
 opportunistic \citep{petanidou-2008-564, Vazquez2005b,
@@
 length and corolla depth impose some biophysical limits to the
-interactions between plants and pollinators
-\citep{Vazquez2009evaluating, Vazquez2009, Stang2009, Stang2006,
-  Santamaria2007}. Such opportunism may buffer plant-pollinator
-communities from global change \citep[e.g.,][]{ramos2012topological,
-  kaiser2010robustness}, but our limited understanding of the assembly
-of these communities impedes making such predictions
-\citep{Vazquez2009, Burkle2011}. Unlike in the broader food web
-literature, we have few assembly models of mutualistic network
-assembly \citep{valdovinos2013adaptive, Nuismer2013, Guimaraes2011}.
-In addition, the few developed models often borrow their mechanisms
-from competitive interactions, leading to inaccurate biological
-assumptions \citep{holland2006comment}. We need further development of
-mechanistic models of mutualistic systems to generate testable
-predictions, along with empirical exploration of network
-assembly. Plant-pollinator communities and mutualisms are vital for
-biodiversity maintenance and essential ecosystem service provision. We
-must therefore understand the processes underlying their assembly to
-facilitate restoration and conservation.
+interactions between plants and pollinators \citep{Stang2006,
+  Santamaria2007, Vazquez2009evaluating, Vazquez2009, Stang2009}. Such
+opportunism may buffer plant-pollinator communities from global change
+\citep[e.g.,][]{ramos2012topological, kaiser2010robustness}, but our
+limited understanding of the assembly of these communities impedes
+making such predictions \citep{Vazquez2009, Burkle2011}. Unlike in the
+broader food web literature, we have few assembly models of
+mutualistic network assembly \citep{valdovinos2013adaptive,
+  Nuismer2013, Guimaraes2011}.  In addition, the few developed models
+often borrow their mechanisms from competitive interactions, leading
+to inaccurate biological assumptions \citep{holland2006comment}. We
+need further development of mechanistic models of mutualistic systems
+to generate testable predictions, along with empirical exploration of
+network assembly. Plant-pollinator communities and mutualisms are
+vital for biodiversity maintenance and essential ecosystem service
+provision. We must therefore understand the processes underlying their
+assembly to facilitate restoration and conservation.
 
@@
 
-\bibliographystyle{ecol_let}
-
-\bibliography{refs}
+\begin{thebibliography}{67}
+\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
+
+\bibitem[{Ahn \emph{et~al.}(2010)Ahn, Bagrow \& Lehmann}]{ahn2010link}
+\hspace{0em}\\Ahn, Y.Y., Bagrow, J.P. \& Lehmann, S. (2010).
+\newblock Link communities reveal multiscale complexity in networks.
+\newblock \emph{Nature}, 466, 761--764.
+
+\bibitem[{Albrecht \emph{et~al.}(2010)Albrecht, Riesen \&
+  Schmid}]{albrecht2010plant}
+\hspace{0em}\\Albrecht, M., Riesen, M. \& Schmid, B. (2010).
+\newblock Plant--pollinator network assembly along the chronosequence of a
+  glacier foreland.
+\newblock \emph{Oikos}, 119, 1610--1624.
+
+\bibitem[{Almeida-neto \emph{et~al.}(2008)Almeida-neto, Guimar\~{a}es,
+  Guimar{\~a}es, Loyola \& Ulrich}]{almeida-neto-2008-1227}
+\hspace{0em}\\Almeida-neto, M., Guimar\~{a}es, P., Guimar{\~a}es, P., Loyola,
+  R. \& Ulrich, W. (2008).
+\newblock A consistent metric for nestedness analysis in ecological systems:
+  reconciling concept and measurement.
+\newblock \emph{Oikos}, 117, 1227--1239.
+
+\bibitem[{Anderson \emph{et~al.}(2011)Anderson, Crist, Chase, Vellend, Inouye,
+  Freestone, Sanders, Cornell, Comita, Davies \emph{et~al.}}]{anderson-2011-19}
+\hspace{0em}\\Anderson, M.J., Crist, T.O., Chase, J.M., Vellend, M., Inouye,
+  B.D., Freestone, A.L., Sanders, N.J., Cornell, H.V., Comita, L.S., Davies,
+  K.F. \emph{et~al.} (2011).
+\newblock Navigating the multiple meanings of $\beta$ diversity: a roadmap for
+  the practicing ecologist.
+\newblock \emph{Ecol. Lett.}, 14, 19--28.
+
+\bibitem[{Barab{\'a}si \& Albert(1999)}]{barabasi1999emergence}
+\hspace{0em}\\Barab{\'a}si, A.L. \& Albert, R. (1999).
+\newblock Emergence of scaling in random networks.
+\newblock \emph{Science}, 286, 509--512.
+
+\bibitem[{Barnosky \emph{et~al.}(2011)Barnosky, Matzke, Tomiya, Wogan, Swartz,
+  Quental, Marshall, McGuire, Lindsey, Maguire \emph{et~al.}}]{barnosky2011has}
+\hspace{0em}\\Barnosky, A.D., Matzke, N., Tomiya, S., Wogan, G.O., Swartz, B.,
+  Quental, T.B., Marshall, C., McGuire, J.L., Lindsey, E.L., Maguire, K.C.
+  \emph{et~al.} (2011).
+\newblock Has the earth's sixth mass extinction already arrived?
+\newblock \emph{Nature}, 471, 51--57.
+
+\bibitem[{Bascompte \emph{et~al.}(2003)Bascompte, Jordano, Meli\'an \&
+  Olesen}]{Bascompte2003}
+\hspace{0em}\\Bascompte, J., Jordano, P., Meli\'an, C.J. \& Olesen, J.M.
+  (2003).
+\newblock The nested assembly of plant-animal mutualistic networks.
+\newblock \emph{Proc. Natl. Acad. Sci. USA}, 100, 9383--9387.
+
+\bibitem[{Bascompte \emph{et~al.}(2006)Bascompte, Jordano \&
+  Olesen}]{Bascompte2006}
+\hspace{0em}\\Bascompte, J., Jordano, P. \& Olesen, J.M. (2006).
+\newblock Asymmetric coevolutionary networks facilitate biodiversity
+  maintenance.
+\newblock \emph{Science}, 312, 431--433.
+
+\bibitem[{Bascompte \& Stouffer(2009)}]{Bascompte2009}
+\hspace{0em}\\Bascompte, J. \& Stouffer, D.B. (2009).
+\newblock The assembly and disassembly of ecological networks.
+\newblock \emph{Phil. Trans. R. Soc. B}, 364, 1781.
+
+\bibitem[{Bates \emph{et~al.}(2014)Bates, Maechler, Bolker \& Walker}]{lme4}
+\hspace{0em}\\Bates, D., Maechler, M., Bolker, B. \& Walker, S. (2014).
+\newblock \emph{lme4: Linear mixed-effects models using Eigen and S4}.
+\newblock R package version 1.1-7.
+
+\bibitem[{Bl{\"u}thgen \emph{et~al.}(2006)Bl{\"u}thgen, Menzel \&
+  Bl{\"u}thgen}]{bluthgen-2006-9}
+\hspace{0em}\\Bl{\"u}thgen, N., Menzel, F. \& Bl{\"u}thgen, N. (2006).
+\newblock Measuring specialization in species interaction networks.
+\newblock \emph{BMC Ecol.}, 6, 9.
+
+\bibitem[{Bl{\"u}thgen \emph{et~al.}(2007)Bl{\"u}thgen, Menzel, Hovestadt,
+  Fiala \& Bl{\"u}thgen}]{Bluthgen2007}
+\hspace{0em}\\Bl{\"u}thgen, N., Menzel, F., Hovestadt, T., Fiala, B. \&
+  Bl{\"u}thgen, N. (2007).
+\newblock Specialization, constraints, and conflicting interests in mutualistic
+  networks.
+\newblock \emph{Curr. Biol.}, 17, 341--346.
+
+\bibitem[{Burgos \emph{et~al.}(2007)Burgos, Ceva, Perazzo, Devoto, Medan,
+  Zimmermann \& Delbue}]{Burgos2007}
+\hspace{0em}\\Burgos, E., Ceva, H., Perazzo, R.P., Devoto, M., Medan, D.,
+  Zimmermann, M. \& Delbue, A.M. (2007).
+\newblock Why nestedness in mutualistic networks?
+\newblock \emph{J. Theor. Biol.}, 249, 307--313.
+
+\bibitem[{Burkle \& Alarc{\'o}n(2011)}]{Burkle2011}
+\hspace{0em}\\Burkle, L.A. \& Alarc{\'o}n, R. (2011).
+\newblock The future of plant-pollinator diversity: Understanding interaction
+  networks across time, space, and global change.
+\newblock \emph{Am. J. Bot.}, 98, 528.
+
+\bibitem[{Cardinale \emph{et~al.}(2012)Cardinale, Duffy, Gonzalez, Hooper,
+  Perrings, Venail, Narwani, Mace, Tilman, Wardle
+  \emph{et~al.}}]{cardinale-2012-59}
+\hspace{0em}\\Cardinale, B.J., Duffy, J.E., Gonzalez, A., Hooper, D.U.,
+  Perrings, C., Venail, P., Narwani, A., Mace, G.M., Tilman, D., Wardle, D.A.
+  \emph{et~al.} (2012).
+\newblock Biodiversity loss and its impact on humanity.
+\newblock \emph{Nature}, 486, 59--67.
+
+\bibitem[{Chao \emph{et~al.}(2005)Chao, Chazdon, Colwell \&
+  Shen}]{chao-2005-148}
+\hspace{0em}\\Chao, A., Chazdon, R.L., Colwell, R.K. \& Shen, T.J. (2005).
+\newblock A new statistical approach for assessing similarity of species
+  composition with incidence and abundance data.
+\newblock \emph{Ecol. Lett.}, 8, 148--159.
+
+\bibitem[{Costa \emph{et~al.}(2007)Costa, Rodrigues, Travieso \&
+  Villas~Boas}]{costa2007characterization}
+\hspace{0em}\\Costa, L.d.F., Rodrigues, F.A., Travieso, G. \& Villas~Boas, P.R.
+  (2007).
+\newblock Characterization of complex networks: A survey of measurements.
+\newblock \emph{Adv. Phys.}, 56, 167--242.
+
+\bibitem[{Csardi \& Nepusz(2006)}]{csardi-2006}
+\hspace{0em}\\Csardi, G. \& Nepusz, T. (2006).
+\newblock The igraph software package for complex network research.
+\newblock \emph{InterJournal}, p. 1695.
+
+\bibitem[{Devoto \emph{et~al.}(2012)Devoto, Bailey, Craze \&
+  Memmott}]{devoto2012understanding}
+\hspace{0em}\\Devoto, M., Bailey, S., Craze, P. \& Memmott, J. (2012).
+\newblock Understanding and planning ecological restoration of
+  plant--pollinator networks.
+\newblock \emph{Ecol. Lett.}, 15, 319--328.
+
+\bibitem[{D{\'\i}az-Castelazo \emph{et~al.}(2010)D{\'\i}az-Castelazo,
+  Guimar{\~a}es~Jr, Jordano, Thompson, Marquis \& Rico-Gray}]{diaz2010changes}
+\hspace{0em}\\D{\'\i}az-Castelazo, C., Guimar{\~a}es~Jr, P.R., Jordano, P.,
+  Thompson, J.N., Marquis, R.J. \& Rico-Gray, V. (2010).
+\newblock Changes of a mutualistic network over time: reanalysis over a 10-year
+  period.
+\newblock \emph{Ecology}, 91, 793--801.
+
+\bibitem[{Dormann \emph{et~al.}(2008)Dormann, Gruber \& Fr{\"u}nd}]{bipartite}
+\hspace{0em}\\Dormann, C., Gruber, B. \& Fr{\"u}nd, J. (2008).
+\newblock Introducing the bipartite package: analysing ecological networks.
+\newblock \emph{R News}, 8, 8.
+
+\bibitem[{Dunn \emph{et~al.}(2009)Dunn, Harris, Colwell, Koh \&
+  Sodhi}]{dunn2009sixth}
+\hspace{0em}\\Dunn, R.R., Harris, N.C., Colwell, R.K., Koh, L.P. \& Sodhi, N.S.
+  (2009).
+\newblock The sixth mass coextinction: are most endangered species parasites
+  and mutualists?
+\newblock \emph{Proc. R. Soc. Lond. B}, 276, 3037--3045.
+
+\bibitem[{Fiedler(1973)}]{fiedler1973algebraic}
+\hspace{0em}\\Fiedler, M. (1973).
+\newblock Algebraic connectivity of graphs.
+\newblock \emph{Czech. Math. J.}, 23, 298--305.
+
+\bibitem[{Fort \emph{et~al.}(2016)Fort, V{\'a}zquez \& Lan}]{fort2016abundance}
+\hspace{0em}\\Fort, H., V{\'a}zquez, D.P. \& Lan, B.L. (2016).
+\newblock Abundance and generalisation in mutualistic networks: solving the
+  chicken-and-egg dilemma.
+\newblock \emph{Ecol. Lett.}, 19, 4--11.
+
+\bibitem[{Forup \emph{et~al.}(2008)Forup, Henson, Craze \&
+  Memmott}]{forup-2008-742}
+\hspace{0em}\\Forup, M., Henson, K., Craze, P. \& Memmott, J. (2008).
+\newblock The restoration of ecological interactions: plant-pollinator networks
+  on ancient and restored heathlands.
+\newblock \emph{J. Appl. Ecol.}, 45, 742--752.
+
+\bibitem[{Freeman(1978)}]{freeman1978centrality}
+\hspace{0em}\\Freeman, L.C. (1978).
+\newblock Centrality in social networks conceptual clarification.
+\newblock \emph{Soc. Networks}, 1, 215--239.
+
+\bibitem[{Galeano \emph{et~al.}(2009)Galeano, Pastor \& Iriondo}]{Galeano2009}
+\hspace{0em}\\Galeano, J., Pastor, J.M. \& Iriondo, J.M. (2009).
+\newblock Weighted-interaction nestedness estimator (wine): a new estimator to
+  calculate over frequency matrices.
+\newblock \emph{Environ. Model. Softw.}, 24, 1342--1346.
+
+\bibitem[{Gibert \emph{et~al.}(2013)Gibert, Pires, Thompson \&
+  Guimar{\~a}es~Jr}]{gibert2013spatial}
+\hspace{0em}\\Gibert, J.P., Pires, M.M., Thompson, J.N. \& Guimar{\~a}es~Jr,
+  P.R. (2013).
+\newblock The spatial structure of antagonistic species affects coevolution in
+  predictable ways.
+\newblock \emph{Am. Nat.}, 182, 578--591.
+
+\bibitem[{G{\'o}mez \& Zamora(2006)}]{gomez2006ecological}
+\hspace{0em}\\G{\'o}mez, J.M. \& Zamora, R. (2006).
+\newblock Ecological factors that promote the evolution of generalization in
+  pollination systems.
+\newblock \emph{Plant-pollinator interactions, NM Waser, J. Ollerton (eds.).
+  The University of Chicago Press, Chicago}, pp. 145--166.
+
+\bibitem[{Graham \& Fine(2008)}]{graham2008phylogenetic}
+\hspace{0em}\\Graham, C.H. \& Fine, P.V. (2008).
+\newblock Phylogenetic beta diversity: linking ecological and evolutionary
+  processes across space in time.
+\newblock \emph{Ecol. Lett.}, 11, 1265--1277.
+
+\bibitem[{Guimar{\~a}es \emph{et~al.}(2011)Guimar{\~a}es, Jordano \&
+  Thompson}]{Guimaraes2011}
+\hspace{0em}\\Guimar{\~a}es, P.R., Jordano, P. \& Thompson, J.N. (2011).
+\newblock Evolution and coevolution in mutualistic networks.
+\newblock \emph{Ecol. Lett.}, 14, 877--885.
+
+\bibitem[{Holland \emph{et~al.}(2006)Holland, Okuyama \&
+  DeAngelis}]{holland2006comment}
+\hspace{0em}\\Holland, J.N., Okuyama, T. \& DeAngelis, D.L. (2006).
+\newblock Comment on ``asymmetric coevolutionary networks facilitate
+  biodiversity maintenance''.
+\newblock \emph{Science}, 313, 1887b--1887b.
+
+\bibitem[{Kaiser-Bunbury \emph{et~al.}(2010)Kaiser-Bunbury, Muff, Memmott,
+  M{\"u}ller \& Caflisch}]{kaiser2010robustness}
+\hspace{0em}\\Kaiser-Bunbury, C.N., Muff, S., Memmott, J., M{\"u}ller, C.B. \&
+  Caflisch, A. (2010).
+\newblock The robustness of pollination networks to the loss of species and
+  interactions: a quantitative approach incorporating pollinator behaviour.
+\newblock \emph{Ecol. Lett.}, 13, 442--452.
+
+\bibitem[{Kalinka \& Tomancak(2011)}]{kalinka2011linkcomm}
+\hspace{0em}\\Kalinka, A.T. \& Tomancak, P. (2011).
+\newblock linkcomm: an r package for the generation, visualization, and
+  analysis of link communities in networks of arbitrary size and type.
+\newblock \emph{Bioinformatics}, 27.
+
+\bibitem[{Kembel \emph{et~al.}(2010)Kembel, Cowan, Helmus, Cornwell, Morlon,
+  Ackerly, Blomberg \& Webb}]{picante-2010-1463}
+\hspace{0em}\\Kembel, S., Cowan, P., Helmus, M., Cornwell, W., Morlon, H.,
+  Ackerly, D., Blomberg, S. \& Webb, C. (2010).
+\newblock Picante: {R} tools for integrating phylogenies and ecology.
+\newblock \emph{Bioinformatics}, 26, 1463--1464.
+
+\bibitem[{Klein \emph{et~al.}(2007)Klein, Vaissi\`{e}re, Cane,
+  Steffan-Dewenter, Cunningham, Kremen \& Tscharntke}]{klein-2007-303}
+\hspace{0em}\\Klein, A., Vaissi\`{e}re, B., Cane, J., Steffan-Dewenter, I.,
+  Cunningham, S., Kremen, C. \& Tscharntke, T. (2007).
+\newblock Importance of pollinators in changing landscapes for world crops.
+\newblock \emph{Proc. R. Soc. B}, 274, 303--313.
+
+\bibitem[{Kremen \& M'Gonigle(2015)}]{kremen-2015-602}
+\hspace{0em}\\Kremen, C. \& M'Gonigle, L.K. (2015).
+\newblock Small-scale restoration in intensive agricultural landscapes supports
+  more specialized and less mobile pollinator species.
+\newblock \emph{J. Appl. Ecol.}, 52, 602--610.
+
+\bibitem[{Kuznetsova \emph{et~al.}(2014)Kuznetsova, {Bruun Brockhoff} \& {Haubo
+  Bojesen Christensen}}]{lmetest}
+\hspace{0em}\\Kuznetsova, A., {Bruun Brockhoff}, P. \& {Haubo Bojesen
+  Christensen}, R. (2014).
+\newblock \emph{{lmerTest}: Tests for random and fixed effects for linear mixed
+  effect models (lmer objects of lme4 package).}
+\newblock R package version 2.0-11.
+
+\bibitem[{MacLeod \emph{et~al.}(2016)MacLeod, Genung, Ascher \&
+  Winfree}]{macleod2016measuring}
+\hspace{0em}\\MacLeod, M., Genung, M.A., Ascher, J.S. \& Winfree, R. (2016).
+\newblock Measuring partner choice in plant--pollinator networks: Using null
+  models to separate rewiring and fidelity from chance.
+\newblock \emph{Ecology}.
+
+\bibitem[{Memmott \emph{et~al.}(2004)Memmott, Waser \& Price}]{Memmott2004}
+\hspace{0em}\\Memmott, J., Waser, N.M. \& Price, M.V. (2004).
+\newblock Tolerance of pollination networks to species extinctions.
+\newblock \emph{Proc. R. Soc. Lond. B}, 271, 2605--2611.
+
+\bibitem[{Menz \emph{et~al.}(2010)Menz, Phillips, Winfree, Kremen, Aizen,
+  Johnson \& Dixon}]{menz-2010-4}
+\hspace{0em}\\Menz, M., Phillips, R., Winfree, R., Kremen, C., Aizen, M.,
+  Johnson, S. \& Dixon, K. (2010).
+\newblock Reconnecting plants and pollinators: challenges in the restoration of
+  pollination mutualisms.
+\newblock \emph{Trends Plant Sci.}, 16, 4--12.
+
+\bibitem[{M'Gonigle \emph{et~al.}(2015)M'Gonigle, Ponisio, Cutler \&
+  Kremen}]{mgonigle-2015-x}
+\hspace{0em}\\M'Gonigle, L., Ponisio, L., Cutler, K. \& Kremen, C. (2015).
+\newblock Habitat restoration promotes pollinator persistence and colonization
+  in intensively-managed agriculture.
+\newblock \emph{Ecol. Appl.}, 25, 1557--1565.
+
+\bibitem[{M'Gonigle \emph{et~al.}(2016)M'Gonigle, Williams, Lonsdorf \&
+  Kremen}]{mgonigle2016tool}
+\hspace{0em}\\M'Gonigle, L.K., Williams, N.M., Lonsdorf, E. \& Kremen, C.
+  (2016).
+\newblock A tool for selecting plants when restoring habitat for pollinators.
+\newblock \emph{Conserv. Lett.}
+
+\bibitem[{Morandin \& Kremen(2013)}]{morandin-2013-829}
+\hspace{0em}\\Morandin, L. \& Kremen, C. (2013).
+\newblock Hedgerow restoration promotes pollinator populations and exports
+  native bees to adjacent fields.
+\newblock \emph{Ecol. Appl.}, 23, 829--839.
+
+\bibitem[{Newman \& Girvan(2004)}]{Newman2004}
+\hspace{0em}\\Newman, M.E.J. \& Girvan, M. (2004).
+\newblock Finding and evaluating community structure in networks.
+\newblock \emph{Phys. Rev. E}, 69, 026113.
+
+\bibitem[{Nuismer \emph{et~al.}(2013)Nuismer, Jordano \&
+  Bascompte}]{Nuismer2013}
+\hspace{0em}\\Nuismer, S.L., Jordano, P. \& Bascompte, J. (2013).
+\newblock Coevolution and the architecture of mutualistic networks.
+\newblock \emph{Evolution}, 67, 338--354.
+
+\bibitem[{Olesen \emph{et~al.}(2007)Olesen, Bascompte, Dupont \&
+  Jordano}]{Olesen2007}
+\hspace{0em}\\Olesen, J.M., Bascompte, J., Dupont, Y. \& Jordano, P. (2007).
+\newblock The modularity of pollination networks.
+\newblock \emph{Proc. Natl. Acad. Sci. USA}, 104, 19891--19896.
+
+\bibitem[{Olesen \emph{et~al.}(2008)Olesen, Bascompte, Elberling \&
+  Jordano}]{Olesen2008}
+\hspace{0em}\\Olesen, J.M., Bascompte, J., Elberling, H. \& Jordano, P. (2008).
+\newblock Temporal dynamics in a pollination network.
+\newblock \emph{Ecology}, 89, 1573.
+
+\bibitem[{Ollerton \emph{et~al.}(2011)Ollerton, Winfree \&
+  Tarrant}]{ollerton-2011-321}
+\hspace{0em}\\Ollerton, J., Winfree, R. \& Tarrant, S. (2011).
+\newblock How many flowering plants are pollinated by animals?
+\newblock \emph{Oikos}, 120, 321--326.
+
+\bibitem[{Parker(1997)}]{parker1997scale}
+\hspace{0em}\\Parker, V.T. (1997).
+\newblock The scale of successional models and restoration objectives.
+\newblock \emph{Restor. Ecol.}, 5, 301--306.
+
+\bibitem[{Peel \& Clauset(2014)}]{peel2014detecting}
+\hspace{0em}\\Peel, L. \& Clauset, A. (2014).
+\newblock Detecting change points in the large-scale structure of evolving
+  networks.
+\newblock \emph{arXiv preprint arXiv:1403.0989}.
+
+\bibitem[{Petanidou \emph{et~al.}(2008)Petanidou, Kallimanis, Tzanopoulos,
+  Sgardelis \& Pantis}]{petanidou-2008-564}
+\hspace{0em}\\Petanidou, T., Kallimanis, S., Tzanopoulos, J., Sgardelis, S. \&
+  Pantis, J. (2008).
+\newblock Long-term observation of a pollination network: fluctuation in
+  species and interactions, relative invariance of network structure and
+  implications for estimates of specialization.
+\newblock \emph{Ecol. Lett.}, 11, 564--575.
+
+\bibitem[{Ponisio \emph{et~al.}(2016)Ponisio, M'Gonigle \&
+  Kremen}]{ponisio2015farm}
+\hspace{0em}\\Ponisio, L.C., M'Gonigle, L.K. \& Kremen, C. (2016).
+\newblock On-farm habitat restoration counters biotic homogenization in
+  intensively managed agriculture.
+\newblock \emph{Global Change Biol.}, 22, 704--715.
+
+\bibitem[{Pyke(1984)}]{pyke1984optimal}
+\hspace{0em}\\Pyke, G.H. (1984).
+\newblock Optimal foraging theory: a critical review.
+\newblock \emph{Annu. Rev. Ecol. Evol. Syst.}, pp. 523--575.
+
+\bibitem[{{R Core Team}(2015)}]{R}
+\hspace{0em}\\{R Core Team} (2015).
+\newblock \emph{R: A Language and Environment for Statistical Computing}.
+\newblock R Foundation for Statistical Computing, Vienna, Austria.
+
+\bibitem[{Ramos-Jiliberto \emph{et~al.}(2012)Ramos-Jiliberto, Valdovinos,
+  Moisset~de Espan{\'e}s \& Flores}]{ramos2012topological}
+\hspace{0em}\\Ramos-Jiliberto, R., Valdovinos, F.S., Moisset~de Espan{\'e}s, P.
+  \& Flores, J.D. (2012).
+\newblock Topological plasticity increases robustness of mutualistic networks.
+\newblock \emph{J. Anim. Ecol.}, 81, 896--904.
+
+\bibitem[{Rezende \emph{et~al.}(2007)Rezende, Lavabre, Guimar{\~a}es, Jordano
+  \& Bascompte}]{Rezende2007}
+\hspace{0em}\\Rezende, E.L., Lavabre, J.E., Guimar{\~a}es, P.R., Jordano, P. \&
+  Bascompte, J. (2007).
+\newblock Non-random coextinctions in phylogenetically structured mutualistic
+  networks.
+\newblock \emph{Nature}, 448, 925--928.
+
+\bibitem[{Santamar{\'\i}a \& Rodr{\'\i}guez-Giron{\'e}s(2007)}]{Santamaria2007}
+\hspace{0em}\\Santamar{\'\i}a, L. \& Rodr{\'\i}guez-Giron{\'e}s, M.A. (2007).
+\newblock Linkage rules for plant--pollinator networks: Trait complementarity
+  or exploitation barriers?
+\newblock \emph{PLoS Biol.}, 5, e31.
+
+\bibitem[{Stang \emph{et~al.}(2009)Stang, Klinkhamer, Waser, Stang \& van
+  der}]{Stang2009}
+\hspace{0em}\\Stang, M., Klinkhamer, P., Waser, N.M., Stang, I. \& van der,
+  M.E. (2009).
+\newblock Size-specific interaction patterns and size matching in a
+  plant-pollinator interaction web.
+\newblock \emph{Ann. Bot.}, 103.
+
+\bibitem[{Stang \emph{et~al.}(2006)Stang, Klinkhamer \& Van
+  Der~Meijden}]{Stang2006}
+\hspace{0em}\\Stang, M., Klinkhamer, P.G. \& Van Der~Meijden, E. (2006).
+\newblock Size constraints and flower abundance determine the number of
+  interactions in a plant--flower visitor web.
+\newblock \emph{Oikos}, 112, 111--121.
+
+\bibitem[{Th{\'e}bault \& Fontaine(2010)}]{Thebault2010}
+\hspace{0em}\\Th{\'e}bault, E. \& Fontaine, C. (2010).
+\newblock Stability of ecological communities and the architecture of
+  mutualistic and trophic networks.
+\newblock \emph{Science}, 329, 853--856.
+
+\bibitem[{Valdovinos \emph{et~al.}(2013)Valdovinos, Moisset~de Espan{\'e}s,
+  Flores \& Ramos-Jiliberto}]{valdovinos2013adaptive}
+\hspace{0em}\\Valdovinos, F.S., Moisset~de Espan{\'e}s, P., Flores, J.D. \&
+  Ramos-Jiliberto, R. (2013).
+\newblock Adaptive foraging allows the maintenance of biodiversity of
+  pollination networks.
+\newblock \emph{Oikos}, 122, 907--917.
+
+\bibitem[{Valdovinos \emph{et~al.}(2010)Valdovinos, Ramos-Jiliberto,
+  Garay-Narv{\'a}ez, Urbani \& Dunne}]{valdovinos2010consequences}
+\hspace{0em}\\Valdovinos, F.S., Ramos-Jiliberto, R., Garay-Narv{\'a}ez, L.,
+  Urbani, P. \& Dunne, J.A. (2010).
+\newblock Consequences of adaptive behaviour for the structure and dynamics of
+  food webs.
+\newblock \emph{Ecol. Lett.}, 13, 1546--1559.
+
+\bibitem[{V\'azquez(2005)}]{Vazquez2005b}
+\hspace{0em}\\V\'azquez, D.P. (2005).
+\newblock Degree distribution in plant-animal mutualistic networks: forbidden
+  links or random interactions?
+\newblock \emph{Oikos}, 108, 421--426.
+
+\bibitem[{V\'azquez \emph{et~al.}(2009)V\'azquez, Bl{\"u}thgen, Cagnolo \&
+  Chacoff}]{Vazquez2009}
+\hspace{0em}\\V\'azquez, D.P., Bl{\"u}thgen, N., Cagnolo, L. \& Chacoff, N.P.
+  (2009).
+\newblock Uniting pattern and process in plant-animal mutualistic networks: a
+  review.
+\newblock \emph{Ann. of Bot.}, 103, 1445--1457.
+
+\bibitem[{V{\'a}zquez \emph{et~al.}(2009)V{\'a}zquez, Chacoff \&
+  Cagnolo}]{Vazquez2009evaluating}
+\hspace{0em}\\V{\'a}zquez, D.P., Chacoff, N.P. \& Cagnolo, L. (2009).
+\newblock Evaluating multiple determinants of the structure of plant-animal
+  mutualistic networks.
+\newblock \emph{Ecology}, 90, 2039--2046.
+
+\bibitem[{Waser \emph{et~al.}(1996)Waser, Chittka, Price, Williams \&
+  Ollerton}]{Waser1996}
+\hspace{0em}\\Waser, N.M., Chittka, L., Price, M.V., Williams, N.M. \&
+  Ollerton, J. (1996).
+\newblock Generalization in pollination systems, and why it matters.
+\newblock \emph{Ecology}, 77, 1043.
+
+\end{thebibliography}
 
@@
   \centering
-  \includegraphics[width=1\textwidth]{../analysis/changePoint/plotting/networksv1.pdf}
+  \includegraphics[width=1\textwidth]{figures/networksv1.pdf}
   \caption{Assembling hedgerow networks had more changing points
@@
     perimeter of the network across years. Asterisks indicate the year
-    the hegdgerow was planted. Before that, the sites were weedy field
+    the hedgerow was planted. Before that, the sites were weedy field
     margins.}
@@
   \centering
-  \includegraphics[width=.95\textwidth]{../analysis/changePoint/plotting/communitiesSkanky3.pdf}
+  \includegraphics[width=.95\textwidth]{figures/communitiesSkanky3.pdf}
   \caption{The species module membership between network changing
@@
     the top panel, species are grouped by module. The bottom panels
-    visualize the flow of species between modules between changing
-    points. Each line represents the change of s species from one
-    module to another. }
+    visualize the flow of species to different modules between change
+    points. }
   \label{fig:changePoints2}
@@
   \centering
-  \includegraphics[width=.8\textwidth]{../analysis/variability/figures/cv/occ_degree.pdf}
+  \includegraphics[width=.8\textwidth]{figures/occ_degree.jpeg}
   \caption{The variation coefficient of network position, as
@@
   \centering
-  \includegraphics[width=1\textwidth]{../analysis/variability/figures/turnover_panels.pdf}
+  \includegraphics[width=1\textwidth]{figures/turnover_panels.jpeg}
   \caption{Species interaction and weighted interactions turnover
@@
   \centering
-  \includegraphics[width=1\textwidth]{../analysis/networkLevel/figures/baci.pdf}
-  \caption{Nestedness, plant richness and pollinator richness
-    increased as the networks assembled. Specialization and modularity
-    remained consistent across years, while connectance decreased. The
-    nestedness, modularity and specialization scores represent
-    $z$-scores. Scores greater than $\sim 2$ or less than $\sim -2$
-    are significantly more or less structured than randomly assembled
-    networks. Points are the metric value for each site at each year
-    of assembly. The solid line indicates the mean slope estimate and
-    the dashed lines are the $95\%$ confidence intervals around the
-    estimate.}
+  \includegraphics[width=1\textwidth]{figures/baci.jpeg}
+  \caption{Plant richness and pollinator richness increased as the
+    networks assembled. Nestedness, modularity and specialization did
+    not change predictable across years, while connectance
+    decreased. The nestedness, modularity and specialization scores
+    represent $z$-scores. Scores greater than $\sim 2$ or less than
+    $\sim -2$ are significantly more or less structured than randomly
+    assembled networks. Points are the metric value for each site at
+    each year of assembly. The solid line indicates the mean slope
+    estimate and the dashed lines are the $95\%$ confidence intervals
+    around the estimate.}
   \label{fig:baci}

### 2016-12-02 ms/ms.tex
version sent to ecol let

@@
 \usepackage{lscape}
+\usepackage{array}
+\usepackage{paralist}
 \usepackage{fancyhdr}
-\usepackage[normalem]{ulem} 
-\usepackage{hyperref}
-\usepackage[parfill]{parskip}
+\usepackage{graphicx}
+\usepackage{longtable}
+\usepackage{times}
+\usepackage{textcomp}
 \usepackage{xr}
-\externaldocument{SI}
+\usepackage{etoolbox}
+\usepackage{filecontents}
 \usepackage{url}
 \usepackage{rotating}
+\externaldocument{SI}
 
-\hypersetup{colorlinks=true, linkcolor=black, citecolor=black}
+\makeatletter
+\def\@biblabel#1{#1.\\}
+\makeatother
+
+\bibliographystyle{ecology_let}
+
+\parindent=2em
+\setlength{\parskip}{1ex}
+
 \RequirePackage{lineno}
-
-\newcommand{\flagged}[1] {
-  \textcolor{blue}{#1}
-}
 
@@
 
-%% there are other ways to assemble that are not preferentia;
-%% attchment, particularly vis big reorganizations of interactions
-%% Lauren I love this paragraph!
-In contrast to the network build-up described by preferential
-attachment, significant reorganizations of interactions can also
-punctuate assembly \citep{peel2014detecting}. Such significant
-reorganizations of interactions, or network changing points, are
-observed in social networks that respond to abrupt shifts in the
-behavior of interactors \citep{peel2014detecting}. A change point is
+Significant reorganizations of interactions --- change points --- can
+also punctuate assembly \citep{peel2014detecting}. A change point is
 caused by a merge, split, fragmentation or formation of modules within
-a network. In ecological communities, such shifts may occur if, as new
-species colonize, resident species change their interaction partners
-to optimize their foraging effort. In plant-pollinator communities,
-theory predicts that pollinators optimize their use of floral
-resources to reduce interspecific competition and improve resource-use
-efficiency \citep{pyke1984optimal, valdovinos2010consequences,
-  valdovinos2013adaptive, albrecht2010plant, Bluthgen2007}. No
-studies, however, have examined if network changing points occur
-during ecological network assembly and/or how these changes relate to
-species behavior.
+a network. Such interaction reorganizations are observed in social
+networks responding to abrupt shifts in the behavior of interactors
+\citep{peel2014detecting}. In ecological communities, such shifts may
+occur if, as new species colonize, resident species change their
+interaction partners to optimize their foraging effort
+\citep{pyke1984optimal, Bluthgen2007, valdovinos2010consequences,
+  albrecht2010plant, valdovinos2013adaptive}. For preferential
+attachment to be at play, change points would need to be driven
+primarily by peripheral, temporally variable species while a stable,
+well-connected core of species remained stable in their
+interactions. No studies, however, have examined if network changing
+points occur during ecological network assembly and/or how these
+changes relate to species behavior.
 
-%% restoration
 Understanding network assembly is particularly relevant to ecological
@@
 tree reconstruction, we fit the GHRG model to the networks
-\citep{peel2014detecting}. This is accomplished by using a Markov
-chain Monte Carlo (MCMC) procedure to first sample the posterior
-distribution of bipartitions, from which a consensus tree is derived
-\citep{peel2014detecting}. We use $\beta$ distributions with the
-hyperparameters $\alpha=\beta=1$ to define priors for $p_r$.
+\citep{peel2014detecting}. We used a Markov chain Monte Carlo (MCMC)
+procedure to first sample the posterior distribution of bipartitions,
+from which a consensus tree was derived \citep{peel2014detecting}. We
+used $\beta$ distributions with the hyperparameters $\alpha=\beta=1$
+to define priors for $p_r$.
 
-Once the GHRG model has been fit to the networks, we determine whether
-a change point occurred between two time slices. To detect a change
-point, we use Bayes factors to compare the fit of two models --- one
+Once the GHRG model was fit to the networks, we determined whether a
+change point occurred between two time slices. To detect a change
+point, we used Bayes factors to compare the fit of two models --- one
 where a change point occurred between two networks, and one where no
@@
 \subsection*{Network structure}
-%Network nestedness significantly increased with assembly
-%(Fig.~\ref{fig:baci}, estimate of the slope of nestedness through time
-%$\pm$ SE, $1.834 \pm 0.6142$, $p$-value=$0.022$). 
-Nestedness and modularity did not change in any predictable pattern
-with assembly (Fig.~\ref{fig:baci}). All of the networks were
-significantly nested ($z$-scores $>$ $1.96$), but not modular ($z$-scores
-$<$ 2, Fig.~\ref{fig:baci}).
 
-%though the slope was not significantly different from
-%zero.(estimate
-% of the slope of modularity through time $\pm$ standard error of the
-% estimate, $-0.524$ $\pm$ $0.295$, $p$-value=$0.124$).
-%In addition, none of the networks were significantly modular
-% Fig.~\ref{fig:baci}).
-
-Connectance decreased as the community assembled (Fig.~\ref{fig:baci},
-estimate of the slope of connectance through time $\pm$ standard error
-of the estimate, $-0.023$ $\pm$ $0.008$,
-$p$-value=$0.007$). Specialization also decreased, though the slope
-was not significantly different from
-zero. %(estimate of the slope of specialization through time $\pm$ SE, $-0.926$ $\pm$ $0.450$, $p$-value=$0.078$).
-Most communities were more generalized than expected when interactions
-were randomized (Fig.~\ref{fig:baci}).
+Nestedness, modularity and specialization did not change predictably
+assembly (Fig.~\ref{fig:baci}). All of the networks were significantly
+nested ($z$-scores $>$ $2$), but not modular ($z$-scores $<$ 2,
+Fig.~\ref{fig:baci}). Most communities were more generalized than
+expected when interactions were randomized
+(Fig.~\ref{fig:baci}). Connectance decreased as the community
+assembled (Fig.~\ref{fig:baci}, estimate of the slope of connectance
+through time $\pm$ standard error of the estimate, $-0.023$ $\pm$
+$0.008$, $p$-value=$0.007$).
 
@@
 collaboration.
-
-\bibliographystyle{ecol_let}
 
@@
   \includegraphics[width=1\textwidth]{../analysis/networkLevel/figures/baci.pdf}
-  \caption{Nestedness, plant richness and pollinator richness
-    increased as the networks assembled. Specialization and modularity
-    remained consistent across years, while connectance decreased. The
-    nestedness, modularity and specialization scores represent
-    $z$-scores. Scores greater than $\sim 2$ or less than $\sim -2$
-    are significantly more or less structured than randomly assembled
-    networks. Points are the metric value for each site at each year
-    of assembly. The solid line indicates the mean slope estimate and
-    the dashed lines are the $95\%$ confidence intervals around the
-    estimate.}
+  \caption{Plant richness and pollinator richness increased as the
+    networks assembled. Nestedness, modularity and specialization did
+    not change predictable across years, while connectance
+    decreased. The nestedness, modularity and specialization scores
+    represent $z$-scores. Scores greater than $\sim 2$ or less than
+    $\sim -2$ are significantly more or less structured than randomly
+    assembled networks. Points are the metric value for each site at
+    each year of assembly. The solid line indicates the mean slope
+    estimate and the dashed lines are the $95\%$ confidence intervals
+    around the estimate.}
   \label{fig:baci}

### 2016-12-02 analysis/variability/plotting/src/CIplotting.R

@@ -60,8 +60,8 @@ plot.panel <- function(dats,
   plot(NA,
        xlim=range(new.dd[,xs], na.rm=TRUE),
        ylim=range(c(new.dd$phi,  new.dd$plo, y2,
-          quantile(dats[,y1],
-         na.rm=TRUE)[c("25%","75%")]),
+         quantile(dats[,y1],
+                  na.rm=TRUE, probs=c(0.04, 0.975))[c("4%","97.5%")]),
          na.rm=TRUE),
        xlab="",
        ylab="",
@@ -70,8 +70,11 @@ plot.panel <- function(dats,
        las=1)
   if(plot.y){
     axis(2,
-         pretty(range(c(new.dd$phi,  new.dd$plo, y2), na.rm=TRUE),
-                       min.n=3),
+         pretty(range(c(new.dd$phi,  new.dd$plo, y2,
+                        quantile(dats[,y1],
+                                 na.rm=TRUE, probs=c(0.04, 0.975))[c("4%","97.5%")]),
+                      na.rm=TRUE),
+                min.n=3),
          las=1)
     mtext(ylabel, 2, line=3, cex=1.5)
   }

### 2016-12-14 analysis/changePoint/hedgerows.py
misc

@@ -1,46 +1,34 @@
 
-## BACI
-python runNetworkChangePoint.py graph-names.lut 4 'Barger_2006.pairs Barger_2007.pairs Barger_2008.pairs Barger_2009.pairs Barger_2011.pairs Barger_2012.pairs Barger_2013.pairs Barger_2014.pairs Barger_2015.pairs' -p 'baci/'
+## if this were in R I would list all the files then divide by using
+## grep on the names.... then loop over the name groups 
 
-python runNetworkChangePoint.py graph-names.lut 4 'Butler_2006.pairs Butler_2007.pairs Butler_2008.pairs Butler_2009.pairs Butler_2011.pairs Butler_2012.pairs Butler_2013.pairs Butler_2014.pairs Butler_2015.pairs' -p 'baci/'
+import glob
+from subprocess import call
 
-python runNetworkChangePoint.py graph-names.lut 4 'Hrdy_2007.pairs Hrdy_2008.pairs Hrdy_2009.pairs Hrdy_2011.pairs Hrdy_2012.pairs Hrdy_2013.pairs Hrdy_2014.pairs' -p 'baci/'
+## find .pair files
+l = glob.glob("baci/*_*.pairs")
 
-python runNetworkChangePoint.py graph-names.lut 4 'MullerB_2006.pairs MullerB_2007.pairs MullerB_2008.pairs MullerB_2009.pairs MullerB_2011.pairs MullerB_2012.pairs MullerB_2013.pairs MullerB_2014.pairs MullerB_2015.pairs' -p baci/
+prefixes = []
+for file in l:
+    s = file.split("_")
+    fname = s[0]
+    fname = fname[5:]
+    prefixes.append(fname)
 
-python runNetworkChangePoint.py graph-names.lut 4 'Sperandio_2007.pairs Sperandio_2008.pairs Sperandio_2009.pairs Sperandio_2011.pairs Sperandio_2012.pairs Sperandio_2013.pairs Sperandio_2014.pairs' -p 'baci/'
+prefixes = set(prefixes)
 
-## Controls
-python runNetworkChangePoint.py graph-names.lut 4 'BC2_2007.pairs BC2_2008.pairs BC2_2009.pairs BC2_2011.pairs BC2_2012.pairs BC2_2013.pairs BC2_2014.pairs' -p 'baci/'
-
[…]

### 2016-12-14 analysis/changePoint/hedgerows.sh

@@ -1,0 +1,62 @@
+
+## if this were in R I would list all the files then divide by using
+## grep on the names.... then loop over the name groups 
+
+import glob 
+from subprocess import call
+
+
+pairs = glob.glob("*_.pairs")
+
+
+
+
+
+
+
+
+## BACI
+python runNetworkChangePoint.py graph-names.lut 4 'Barger_2006.pairs Barger_2007.pairs Barger_2008.pairs Barger_2009.pairs Barger_2011.pairs Barger_2012.pairs Barger_2013.pairs Barger_2014.pairs Barger_2015.pairs' -p 'baci/'
+
+python runNetworkChangePoint.py graph-names.lut 4 'Butler_2006.pairs Butler_2007.pairs Butler_2008.pairs Butler_2009.pairs Butler_2011.pairs Butler_2012.pairs Butler_2013.pairs Butler_2014.pairs Butler_2015.pairs' -p 'baci/'
+
+python runNetworkChangePoint.py graph-names.lut 4 'Hrdy_2007.pairs Hrdy_2008.pairs Hrdy_2009.pairs Hrdy_2011.pairs Hrdy_2012.pairs Hrdy_2013.pairs Hrdy_2014.pairs' -p 'baci/'
+
+python runNetworkChangePoint.py graph-names.lut 4 'MullerB_2006.pairs MullerB_2007.pairs MullerB_2008.pairs MullerB_2009.pairs MullerB_2011.pairs MullerB_2012.pairs MullerB_2013.pairs MullerB_2014.pairs MullerB_2015.pairs' -p baci/
+
+python runNetworkChangePoint.py graph-names.lut 4 'Sperandio_2007.pairs Sperandio_2008.pairs Sperandio_2009.pairs Sperandio_2011.pairs Sperandio_2012.pairs Sperandio_2013.pairs Sperandio_2014.pairs' -p 'baci/'
+
+## Controls
+python runNetworkChangePoint.py graph-names.lut 4 'BC2_2007.pairs BC2_2008.pairs BC2_2009.pairs BC2_2011.pairs BC2_2012.pairs BC2_2013.pairs BC2_2014.pairs' -p 'baci/'
+
[…]

### 2016-12-14 analysis/networkLevel/src/resilience.R

@@ -1,6 +1,6 @@
 
 ## calculates the robustness of a network using Memmot et al.'s method
-## takes the adjacency matric, whether to drop species by abundance or
+## takes the adjacency matrix, whether to drop species by abundance or
 ## degree and whther to drop the "higer" or "lower" level of the
 ## network
 ## returns a data frame with the site, robustness score and YPR

### 2016-12-14 analysis/variability/beta-div.R

@@ -10,6 +10,7 @@ source('src/initialize_beta.R')
 ## beta diversity as variation between years,
 ## centroid for each site
 ## ************************************************************
+
 dis <- mapply(function(a, b, c, d)
               beta.status(comm= a,
                           status= b,
@@ -31,7 +32,8 @@ dats <- data.frame(site=comm$sites,
                      x$distances)))
 rownames(dats) <- NULL
 
-save(dats, file= file.path('saved/speciesTurnover',
+
+save(dats, file= file.path('saved/speciesTurnover', 
              sprintf('%s.pdf', paste(dis.method, alpha, occ, type,
                                      sep='_'))))
 

### 2016-12-15 ms/hardcoded/ms.tex
misc
made some suggestions to the intro

@@
 \def\extras{
+\clearpage
   \begin{itemize}
@@
   \paragraph{Running head:} \textsc{\runninghead}
-  % \parindent=0pt
-  \begin{center}%
-    {\LARGE \title \par}%
-    \vskip 3em%
+  \parindent=0pt
+  \begin{center}
+    {\LARGE \title \par}
+    \vskip 3em
     {\large
-      \lineskip .75em%
-      \begin{tabular}[t]{c}%
+      \lineskip .75em
+      \begin{tabular}[t]{c}
         \author
-      \end{tabular}\par}%
-    \vskip 1.5em%
+      \end{tabular}\par}
+    \vskip 1.5em
   \end{center}\par
   \affiliation
+  \extras
 }
@@
 \label{sec:introduction}
+
+1) Stability x persistence x community assembly - maybe the easiest part to put a theory. [H4]
+
+2) Link assembly to restoration and ecosystem services, biodiversity crisis (input the former first paragraph). [H4]
+
+3) Focus on specific species in this paragraph, maybe link to niche theory?
+Mention \cite{kaiser2010robustness, macleod2016measuring} to link species change in interactions (rewire) to their persistence and temporal variation; say that we want to restore habitats that are robust and persistent through time.
+However, the relationship between species persistence across years and their role in community assembly is still unclear.
+[H3 and H4]
+
+4) Paragraph that starts with "Understanding network assembly..." [H1 and H2]
+
+5) Paragraph on changing points; mention the control and the non-assembly hedgerows here [H1 and H2]
+
+6) Maybe rephrase the last paragraph to more question driven statements?
+We tested four theoretically motivated hypothesis exploring the process of community assembly using a eight-year
+dataset of plant-pollinator network development following hedgerow
+installment in the highly simplified and intensively managed
+agricultural landscape of California's Central Valley. First,
+we determine whether network assembly was punctuated by significant
+reorganizations of interactions, and if assembling networks presented more change than the control, non-assembly hedgerows. We next tested whether the species
+that are most variable in their network position --- and thus
+important contributors to network reorganizations --- are less
+persistent and connected species, as would be expected by preferential
+attachment. To further explore the mechanisms underlying the temporal
+dynamics of the networks, also we examine patterns in the species and
+interaction temporal turnover. Finally, we investigate whether
+networks assemble toward predictable interaction patterns, and the
+ramifications for the robustness of the networks.
 

### 2016-12-18 ms/ms.tex
small changes to mc

@@
 each interaction network and the species within it \citep{Memmott2004,
-  Rezende2007, Bascompte2009, Thebault2010}. Furthermore, it has been
-show that biodiversity has a pivotal role in maintaining ecosystems
-services \citep{cardinale-2012-59}. Thus, to safeguard ecological
-function, it has become increasingly imperative to aid the recovery of
-lost interactions and component biodiversity and by facilitating the
+  Rezende2007, Bascompte2009, Thebault2010}. Furthermore, biodiversity
+play a pivotal role in maintaining ecosystems services
+\citep{cardinale-2012-59}. Thus, to safeguard ecological function, it
+has become increasingly imperative to aid the recovery of lost
+interactions and component biodiversity and by facilitating the
 assembly of robust interaction networks via ecological restoration
-\citep{menz-2010-4}. However, we know very little about how ecological
-networks assemble in general, rendering restoration of interacting
-communities quite difficult.
+\citep{menz-2010-4}. We know very little, however, about how
+ecological networks assemble in general, rendering restoration of
+interacting communities quite difficult.
 
@@
 interact with species that are already well-connected \citep[the
-``rich-get-richer'' principle,][]{barabasi1999emergence}. In
+``rich-get-richer'' principle,][]{barabasi1999emergence}.
+
+In
 pollination systems --- a particularly ubiquitous mutualism
@@
 incorporated into the network as well as the consequences of adding
-species for interaction patterns and robustness.
+xspecies for interaction patterns and robustness.
 
@@
 different flowers, resulting in a higher degree \citep{Vazquez2009,
-  fort2016abundance}. Either way, our result suggests that adaptable
+  fort2016abundance}. Either way, our results suggests that adaptable
 species can change their network position to utilize the most

### 2017-01-20 ms/hardcoded/ms.tex
added the paragraph in the intor about the oppotunistic attachment

@@
 
-1) Stability x persistence x community assembly - maybe the easiest part to put a theory. [H4]
-
-2) Link assembly to restoration and ecosystem services, biodiversity crisis (input the former first paragraph). [H4]
+(First and second paragraphs as they are)
+
+However, properties that enable a species to interact with a broad range of partners (\textit{i.e.}, generalized, high degree species), may switch partners depending on community context.
+The adaptive foraging theory predicts that there is a trade-off between foraging efficiency and resource availability; thus, if competing species are present, it is expected that there will be a decrease in niche overlap among species present in a given area.
+In this sense, when other, competing species are present, generalist species can explore different resources, switching their interaction partners.
+This alternative assembly mechanism can be described as an opportunistic attachment process.
+In the opportunistic attachment process generalist species, instead of being part of a persistent species' core, are actually the species that explore, opportunistically, the most available resources, that is, the resources that are being use the less.
+Thus, when you consider the assembly process in a time series, the most persistent species are also the ones that, through time, exhibit the highest degree, but are also the most variable in their network position at each particular time.
+Species may have multiple interaction partners, but only interact when opportunity arises.
 
@@
 
-4) Paragraph that starts with "Understanding network assembly..." [H1 and H2]
+4) important to understand assembly because of habitat restoration and ecosystem services, biodiversity crisis (input the former first paragraph).
+[H4] (Paragraph that starts with "Understanding network assembly...")  [H1 and H2]
 
@@
 6) Maybe rephrase the last paragraph to more question driven statements?
+
 We tested four theoretically motivated hypothesis exploring the process of community assembly using a eight-year

### 2017-04-19 analysis/writeup/writeup.tex
added sh master file

\documentclass{article}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}
\usepackage{natbib}
\usepackage[unicode=true]{hyperref}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{color}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{verbatim}
\usepackage{mathpazo}
\usepackage{setspace}
\usepackage{multirow}
\usepackage{fullpage}
\usepackage{lscape}
\usepackage{fancyhdr}
\usepackage{wrapfig,lipsum,booktabs}
\usepackage[normalem]{ulem}
\usepackage[parfill]{parskip}
\usepackage{multirow}
\geometry{tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

\bibliographystyle{rspublicnatsort}

%% for inline R code: if the inline code is not correctly parsed, you will see a message
\newcommand{\rinline}[1]{SOMETHING WRONG WITH knitr}

\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
\title{Diversification practices reduce organic to conventional yield
gap: A walkthrough of the anlaysis}
\author{Lauren Ponisio}

\maketitle
\section{Overview}
\label{sec:overview}

In our study
\href{http://nature.berkeley.edu/~lponisio/wp-content/uploads/2014/08/Ponisio2014.pdf}{\textcolor{blue}
{Ponisio
et al.~2015}} we compared the yields or organic and conventional
agriculture using a custom-build meta-analytic model. Here I explain
the analytic differences between our study and others, example our
modeling decisions, and walk through our code. Below is the output of
the final meta-analytic model, which will be described in detail in
this document.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{print}\hlstd{(out}\hlopt{$}\hlstd{bugs,} \hlkwc{dig}\hlstd{=}\hlnum{3}\hlstd{)}
\end{alltt}
\begin{verbatim}
## Inference for Bugs model at "model.jags", fit using jags,
##  3 chains, each with 1e+05 iterations (first 100 discarded), n.thin = 10
##  n.sims = 29970 iterations saved
##            mu.vect sd.vect      2.5%       25%       50%       75%
## cv.obs       1.154   0.134     0.909     1.061     1.149     1.241
## exp.mu       0.808   0.019     0.771     0.795     0.808     0.821
## sigma        0.189   0.023     0.145     0.172     0.188     0.204
## deviance -1293.537  36.474 -1362.740 -1318.528 -1294.309 -1269.421
##              97.5%  Rhat n.eff
## cv.obs       1.434 1.001  5100
## exp.mu       0.846 1.001 27000
## sigma        0.236 1.001  9400
## deviance -1219.532 1.001 20000
## 
## For each parameter, n.eff is a crude measure of effective sample size,
## and Rhat is the potential scale reduction factor (at convergence, Rhat=1).
## 
## DIC info (using the rule, pD = var(deviance)/2)
## pD = 665.1 and DIC = -628.4
## DIC is an estimate of expected predictive error (lower deviance is better).
\end{verbatim}
\end{kframe}
\end{knitrout}

\section{Where it all began, Seufert et al.~2011}
\label{sec:beginning}

How organic agriculture may contribute to world food production has
been subject to vigorous debate over the past decade. Early reviews
comparing organic to conventional agriculture found yield gaps of
$8-9$\% in developed countries \citep{Stanhill1990, Badgley2007} but
yield gains of as much as $180$\% in developing countries. Two recent
meta-analyses, however, found organic yields to be $20-25$\% lower
than conventional yields \citep{dePonti2012, Seufert2012}. These
studies used different criteria for selecting the data to be included,
but importantly each of the above studies used different analytical
methods to combine the data across the different sub-studies. The
studies comparing organic and conventional yields systems often
reported yields from multiple crops across several years. In addition,
they tended to compare multiple treatments (usually organic) to one
control treatment (usually conventional). For example from
\cite{denison2004crop}:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{samp}
\end{alltt}
\begin{verbatim}
##       study   crop conv org year
## 100 study13  maize  CMT OMT 2001
## 101 study13  maize  CMT OMT 1999
## 102 study13 tomato  CWT OMT 2002
## 103 study13  maize  LMT OMT 2000
## 104 study13  maize  CMT OMT 1996
## 105 study13  maize  LMT OMT 2001
## 106 study13 tomato  LMT OMT 1999
## 107 study13  maize  CMT OMT 1997
## 108 study13 tomato  CMT OMT 2001
## 109 study13 tomato  CMT OMT 1995
## 110 study13 tomato  LMT OMT 1995
## 111 study13  maize  LMT OMT 1999
## 112 study13  maize  LMT OMT 2002
## 113 study13  maize  LMT OMT 1994
## 114 study13  maize  CMT OMT 1994
## 115 study13  maize  CMT OMT 2002
## 116 study13 tomato  LMT OMT 1998
## 117 study13 tomato  CMT OMT 1994
## 118 study13 tomato  CMT OMT 1999
## 119 study13 tomato  CWT OMT 1994
## 120 study13 tomato  CWT OMT 1995
## 121 study13 tomato  CMT OMT 1996
## 122 study13 tomato  LMT OMT 1996
## 123 study13 tomato  CWT OMT 1996
## 124 study13  maize  LMT OMT 1995
## 125 study13  maize  CMT OMT 1995
## 126 study13  maize  LMT OMT 1996
## 127 study13 tomato  LMT OMT 1994
## 128 study13 tomato  CWT OMT 1998
## 129 study13 tomato  CMT OMT 1998
## 130 study13 tomato  CMT OMT 1997
## 131 study13 tomato  LMT OMT 1997
## 132 study13 tomato  LMT OMT 2001
## 133 study13 tomato  CWT OMT 2001
## 134 study13 tomato  CWT OMT 1997
## 135 study13  maize  LMT OMT 1997
\end{verbatim}
\end{kframe}
\end{knitrout}

Using the organic to conventional yield comparisons without taking
into account the underlying data structure can lead to potential
pseudo-replication and an understated Type 1 error rate. We used a
randomization test to estimate the Type I error rate of the Seufert et
al.~analysis.  We forced the null hypothesis to be true by randomly
re-assigning the 'organic' and 'conventional' labels for each study
and then using the R package {\tt Metafor} \citep{metafor} to
implement a random effects meta-analysis on each randomized dataset.
Repeating this procedure $10^5$ times enabled us to determine the Type
I error rate (false rejection) resulting from not accounting for the
hierarchical structure of the data.  In over $50\%$ of simulations,
the null hypothesis was rejected using a nominal Type I error rate of
0.05 (Fig.~\ref{fig:permutation}).  In other words, even if organic
and conventional yields are known not to be different, applying the
model used by \cite{Seufert2012} for these data would lead to the
conclusion that they are significantly different in over $50\%$ of
cases. This means that the actual Type I error rate is inflated
relative to what was reported, leading to the following related
statements: the significance levels were overstated; the confidence
intervals were underestimated; the uncertainty was not fully accounted
for.

\begin{figure} \centering
\includegraphics[width=0.5\textwidth]{figure/permutation.pdf}
\caption{The distribution of $p$-values when the null hypothesis was
forced to be true using the data and analysis type present in
Seufert et al.\  \cite{Seufert2012}.  If the analysis procedure was
valid for these data, the distribution of P-values should be
uniform between 0 and 1.  Instead it is sharply shifted toward low
P-values.  In over $50\%$ of simulations, the null hypothesis was
rejected using a nominal Type I error rate of 0.05 (red region
above).}
\label{fig:permutation}
\end{figure}

The de Ponti et al.\ \citep{dePonti2012} study had similar issues with
pseudo-replication. Additionally they did not account for the sampling
variance within studies, which is the recommended practice to deal
with unequal variances in the sample of studies \citep{Gurevitch1999}.

Given these methodological and data-related critiques, a new study was
needed to produce a more robust estimate of the gap between organic
and conventional yields.  We developed a hierarchical meta-analytic
framework that overcomes the methodological pitfalls of previous
studies by accounting for both the multi-level nature of the data and
the yield variation within studies.  Furthermore, via a literature
search we compiled a more extensive and up-to-date meta-dataset,
comprising 1071 organic to conventional yield comparisons from 115
studies --- over three times the number of observations of any of the
previous analyses. Our meta-dataset includes studies from 38 countries
and 52 crop species over a span of 35 years.

\section{Building our meta-analytic model}
\label{sec:model-building}

We built a hierarchical meta-analytic model to generate an estimate of
the yield gap. Following standard practice, we compared the natural
log of the ratios between organic and conventional yields (the
``response ratio'') across studies \citep{Hedges1999,
Seufert2012}. The response ratio is more normally distributed than the
raw ratio and independent of the units of measurement used within a
study and, thus, comparable across studies \citep{Hedges1999}.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{plot}\hlstd{(}\hlkwd{density}\hlstd{(meta.dat}\hlopt{$}\hlstd{lnR),} \hlkwc{main}\hlstd{=}\hlstr{"Distribution of log response ratios"}\hlstd{)}
\end{alltt}
\end{kframe}
\includegraphics[width=\maxwidth]{figure/RR_density-1}

\end{knitrout}

We constructed a hierarchical regression model to account for
the dependencies in the yield data. We expanded on the traditional
random effects model \citep{Hedges1985} by considering three
additional sources of random variation (i.e., random effects): 1)
between studies, 2) within a study between years, and 3) within a year
between response ratios (e.g., across replicated trials of a crop
planted at different times in the season). We also considered whether
the variances of the random effect distributions for 2) and 3) were
shared across studies, or study-specific.

The full possible model, prior to model selection, with all sources of
random variation is

\begin{equation}
  \begin{split}
    \label{eq:fullModel}
    y_{ijk} &= \mu + \alpha_i + \beta_{ij} +  \eta_{ijk} +
    \epsilon_{ijk}  \\
    \alpha_{i} &\sim N(0, \sigma^2_{\alpha})\\
    \beta_{ij} &\sim N(0, \sigma^2_{\beta}[i])\\
    \eta_{ijk} &\sim N(0, \sigma^2_{\eta}[i])\\
    \epsilon_{ijk} &\sim N(0, S_{ijk}])\\
    \sigma^2_{\beta}[i] &\sim \Gamma(CV_\beta, scale_\beta)\\
    \sigma^2_{\eta}[i] &\sim \Gamma(CV_\eta, scale_\eta)
  \end{split}
\end{equation}
% where $y_{ijk}$ is the observed magnitude of the $k^{\mathrm{th}}$
response ratio from the $j^{\mathrm{th}}$ year of the
$i^{\mathrm{th}}$ study, $\mu$ is the mean response ratio across
studies, $\alpha_{i}$ is the effect of $i^{\mathrm{th}}$ study,
$\beta_{ij}$ is the effect of $j^{\mathrm{th}}$ year of the
$i^{\mathrm{th}}$ study, $\eta_{ijk}$ is the effect of the
$k^{\mathrm{th}}$ response ratio of the $j^{\mathrm{th}}$ year of
$i^{\mathrm{th}}$ study, and $\epsilon_{ijk}$ is the
residual. $\sigma^2_{\alpha}$ is the between study variance,
$\sigma^2_{\beta}[i]$ is the between year variance of study $i$,
$\sigma^2_{\eta}[i]$ is the within year, between response ratio
variance of study $i$, and $S_{ijk}$ is the variance of response ratio
$ijk$ as reported by its study. $CV_\beta$ and $CV_\eta$ and
$scale_\beta$ and $scale_\eta$ are the coefficient of variation and
scale parameters of the gamma distributions of the study-specific
between- and within-year variances.  When response ratios that shared
a common control were combined, $y_{ijk}$ corresponds to the aggregate
within-study response ratio \citep[Eq.~3,][]{Lajeunesse2011} and
$S_{ijk}$ is its pooled variance \citep[Eq.~8,][]{Lajeunesse2011}.

\subsection{Parameter inclusion}
To determine the levels of hierarchy supported by the data, we
sequentially added random effects and examined the posteriors of the
parameters to determine the support for their inclusion. We also
confirmed our selection with Deviance Information Criterion (DIC). The
DIC can be problematic for hierarchical models because the effective
number of parameters is not clearly defined \citep{gelman2006data,
kery2012bayesian}. The DIC was therefore used in combination with a
visual examination of the posterior distributions of the parameters to
select the best supported model.

For the variance within and between year random effect distributions,
we considered two parameterizations: 1) the variance terms, denoted
$\sigma^2_{\eta}$ and $\sigma^2_{\beta}$, respectively, were shared
across all studies each with a Uniform(0,100) prior, and 2) the
variance terms were study-specific (i.e., $\sigma^2_{\eta}[i]$ and
$\sigma^2_{\beta}[i]$). In the latter case, the study-specific
precision terms (1/variance) were assumed to be distributed according
to a gamma distribution whose parameters were
estimated. Uniform(0,100) priors were used for the coefficient of
variation ($1/\sqrt{shape}$) and the square root of the scale.

We first added a random effect of study and examined the posterior for
$\sigma_{\alpha}$ (the standard deviation of the common distribution
from which the study effects are drawn). The posterior was clearly
differentiated from zero (Fig.~\ref{fig:posteriors}a). We next added
random variation within a year and examined $\sigma_{\eta}$. We found
it was also clearly different from zero
(Fig.~\ref{fig:posteriors}b). The DIC was also smaller than when only
a random effect of study was included (Tab.~\ref{table:params}). Next
we allowed the within-year precisions to be study-specific and follow
a gamma distribution. We examined the coefficient of variation of the
gamma distribution and found it was clearly differentiated from zero
(Fig.~\ref{fig:posteriors}c). The DIC was also smaller than when a
single within year effect was shared across studies
(Tab.~\ref{table:params}). Lastly, we added a between year random
effect and examined $\sigma_{\beta}$. The posterior was concentrated
at zero (Fig.~\ref{fig:posteriors}d) so we concluded there was
insufficient support for including it in the model. The estimate of
the yield gap and its uncertainty did not differ substantially from
when no between year effect was included
(Fig.~\ref{fig:model_selection}), the DIC, however, was marginally
smaller then when no between year random effect was included
(Tab.~\ref{table:params}).

\clearpage

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figure/posteriors.pdf}
\caption{The posterior distributions for the random effect of a)
study ($\sigma_{\alpha}$); b) response ratios within a year
($\sigma_{\eta}$); c) response ratios within a year where the
within year variance is study-specific, $CV_{\sigma_{\eta}}$ is
the coefficient of variation ($1/\sqrt{shape}$) of the gamma
distribution (this model is most supported by the data); and d)
between year ($\sigma_{\beta}$). Including a between-year variance
term was not supported by the data (the posterior for
($\sigma_{\beta}$ is not differentiated from zero ).}
\label{fig:posteriors}
\end{figure}
\clearpage

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figure/model_selection.pdf}
\caption{The sensitivity of the yield gap to including different
levels of hierarchy in the model. The random effects included in
the model are: a) study ($\sigma_{\alpha}$); b) study and response
ratios within a year ($\sigma_{\eta}$); c) study and response
ratios within a year where the within year variance is
study-specific ($\sigma_{\eta}[i]$) (this model is most supported
by the data); and d) study, study-specific within-year variance,
and between year ($\sigma_{\beta}$). Including a between-year
variance term was not supported by the data (the posterior for
($\sigma_{\beta}$ is not differentiated from zero ). Values are
the posterior mean with 95\% credible intervals.}
\label{fig:model_selection}
\end{figure}
\clearpage

\begin{table}
  \renewcommand*\arraystretch{1.25}
  \centering
  \caption{Parameter posteriors for models without explanatory
    variables. $\mu$ is the true mean response ratio across years and
    studies, $\sigma_{\alpha}$ is the standard deviation of the
    distribution from which the study random effects are drawn;
    $\sigma_{\eta}$ is the standard deviation of the distribution from
    which the within year random effects are drawn; $CV_{\sigma_\eta}$ is the
    coefficient of variation of the gamma from which the
    study-specific within-year variance are drawn; and $\sigma_\beta$
    is the standard deviation of the distribution of random between year
    effects. Values of Rhat $< 1.1$ indicate convergence. Lower
    Deviance Information Criterion (DIC) indicates better model fit to
    the data.}

  \vspace{20pt}
  \label{table:params}
  \begin{tabular}{|c|c|c|c|c|} \hline
    Parameter & Posterior mean & Posterior standard deviation
    & 95\% CI & Rhat \\ \hline
    \multicolumn{5}{|l|}{Study random effect, DIC=1684.8} \\ \hline
    \multirow{1}{*}{$\mu$} &
    $0.795$ & $0.027$ & $0.742 - 0.848$ & $1.001$ \\ \hline
    \multirow{1}{*}{$\sigma_{\alpha}$} &
    $0.341$ & $0.026$ & $0.294 - 0.396$ & $1.001$ \\ \hline

    \multicolumn{5}{|l|}{Study and within year random effects,
      DIC= -565.9} \\ \hline
    \multirow{1}{*}{$\mu$} &
    $0.788$ & $0.021$ & $0.749 - 0.829$ & $1.001$ \\ \hline
    \multirow{1}{*}{$\sigma_{\alpha}$} &
    $0.188$ & $0.024$ & $0.144 - 0.239$ & $1.001$ \\ \hline
    \multirow{1}{*}{$\sigma_{\eta}$} &
    $0.312$ & $0.011$ & $0.291 - 0.333$ & $1.001$ \\ \hline

    \multicolumn{5}{|l|}{Study and study-specific within year random
      effects, DIC= -618.0} \\ \hline
    \multirow{1}{*}{$\mu$} &
    $0.808$ & $0.019$ & $0.771 - 0.845$ & $1.001$ \\ \hline
    \multirow{1}{*}{$\sigma_{\alpha}$} &
    $0.189$ & $0.023$ & $0.145 - 0.237$ & $1.001$ \\ \hline
   \multirow{1}{*}{$CV_{\sigma_\eta}$} &
    $1.155$ & $0.135$ & $0.907 - 1.436$ & $1.001$ \\ \hline

    \multicolumn{5}{|l|}{Study, study-specific within year,
      and between year random effects, DIC= -621.2} \\ \hline
     \multirow{1}{*}{$\mu$} &
    $0.808$ & $0.019$ & $0.770 - 0.846$ & $1.001$ \\ \hline
    \multirow{1}{*}{$\sigma_{\alpha}$} &
    $0.186$ & $0.024$ & $0.142 - 0.234$ & $1.001$ \\ \hline
   \multirow{1}{*}{$CV_{\sigma_\eta}$} &
    $1.157$ & $0.136$ & $0.907 - 1.440$ & $1.001$ \\ \hline
   \multirow{1}{*}{$\sigma_{\beta}$} &
    $0.041$ & $0.027$ & $0.002 - 0.098$ & $1.001$ \\ \hline

  \end{tabular}
\end{table}
\clearpage

The best supported model, coded in JAGS, is

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{verbatim}
## model {
##      for(study in 1:Nstudy) {
##        for(year in 1:Nyear[study]) {
##          for(obs in 1:Nobs[study,year]) {
##            P[study,year,obs] <- 1/V[study,year,obs]
##            RR[study,year,obs] ~ dnorm(mu.RR[study,year,obs],
##                                       P[study,year,obs])
##            mu.RR[study,year,obs] ~ dnorm(mu.study[study],
##                                      tau.yr.obs[study])
##          }
##        }
##        mu.study[study] ~ dnorm(mu, tau)
##        tau.yr.obs[study] ~ dgamma(shape.obs, scale.obs)
##      }
## 
##      mu ~ dnorm(0, 1e-4)
##      exp.mu <- exp(mu)
##      tau <- 1 / (sigma * sigma)
##      sigma ~ dunif(0, 100)
## 
##      shape.obs <- (1/cv.obs)^2
##      cv.obs ~ dunif(0, 100)
##      scale.obs <- (1/in.scale.obs)^2
##      in.scale.obs ~ dunif(0, 100)
##    }
\end{verbatim}
\end{kframe}
\end{knitrout}

Where the parameters names match the notation in
Equ. \ref{eq:fullModel}.

\section{Analysis}
\label{sec:analysis}

After deciding on the model parameterization, the most difficult part
of running a model in JAGS in getting tha data in the right format (in
my opinion). Our situation was particularly difficult because we
needed to build in flexibility to pool different combinations of
response ratios using the method presented in Lajeunesse
\cite{Lajeunesse2011} depending on the explanatory variable being
considered (when response ratios that share a control split between
different levels of an explanatory variable, they can be left
un-pooled). Below I focus on the best supported model without
explanatory variables.

First we have a function to calculate pooled RR and their variances
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{## applies the Lajeunesse (2011) method for combining RR with shared}
\hlcom{## controls. Returns a pooled RR and variance}
\hlstd{poolCommonControl} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{dd}\hlstd{) \{}

  \hlcom{## function to make covariance matrix:}
  \hlstd{covMat} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{x}\hlstd{)\{}
    \hlcom{## case 1: organic is repeated}
    \hlkwa{if}\hlstd{(}\hlkwd{all}\hlstd{(}\hlkwd{diff}\hlstd{(}\hlkwd{as.numeric}\hlstd{(x[,}\hlstr{"mean.org"}\hlstd{]))}\hlopt{==}\hlnum{0}\hlstd{))}
      \hlstd{vals} \hlkwb{<-} \hlkwd{as.numeric}\hlstd{(x[}\hlnum{1}\hlstd{,} \hlkwd{c}\hlstd{(}\hlstr{"sd.org"}\hlstd{,} \hlstr{"n.organic"}\hlstd{,} \hlstr{"mean.org"}\hlstd{)])}
    \hlkwa{if}\hlstd{(}\hlkwd{all}\hlstd{(}\hlkwd{diff}\hlstd{(}\hlkwd{as.numeric}\hlstd{(x[,}\hlstr{"mean.conv"}\hlstd{]))}\hlopt{==}\hlnum{0}\hlstd{))}
      \hlstd{vals} \hlkwb{<-} \hlkwd{as.numeric}\hlstd{(x[}\hlnum{1}\hlstd{,} \hlkwd{c}\hlstd{(}\hlstr{"sd.conv"}\hlstd{,} \hlstr{"n.conv"}\hlstd{,} \hlstr{"mean.conv"}\hlstd{)])}

    \hlstd{m} \hlkwb{<-} \hlkwd{matrix}\hlstd{(vals[}\hlnum{1}\hlstd{]}\hlopt{^}\hlnum{2}\hlopt{/}\hlstd{(vals[}\hlnum{2}\hlstd{]}\hlopt{*}\hlstd{vals[}\hlnum{3}\hlstd{]}\hlopt{^}\hlnum{2}\hlstd{),}
                \hlkwc{ncol}\hlstd{=}\hlkwd{nrow}\hlstd{(dd),} \hlkwc{nrow}\hlstd{=}\hlkwd{nrow}\hlstd{(dd))}
    \hlkwd{diag}\hlstd{(m)} \hlkwb{<-} \hlkwd{as.numeric}\hlstd{(x[,}\hlstr{"varlnR"}\hlstd{])}
    \hlstd{m}
  \hlstd{\}}

  \hlcom{## covariance matrix and its inverse:}
  \hlstd{V} \hlkwb{<-} \hlkwd{covMat}\hlstd{(dd)}
  \hlstd{V.inv} \hlkwb{<-} \hlkwd{solve}\hlstd{(V)}

  \hlstd{X} \hlkwb{<-} \hlkwd{matrix}\hlstd{(}\hlnum{1}\hlstd{,} \hlkwd{nrow}\hlstd{(V))}
  \hlstd{E} \hlkwb{<-} \hlkwd{matrix}\hlstd{(dd[,}\hlstr{"lnR"}\hlstd{],} \hlkwd{nrow}\hlstd{(V))}
  \hlcom{## compute and return RR_bar and variance(RR_bar), as defined in}
  \hlcom{## Eq. 3 and subsequent text in Lajeunesse (2011):}
  \hlkwd{return}\hlstd{(}\hlkwd{c}\hlstd{(}\hlkwc{obs}\hlstd{=dd}\hlopt{$}\hlstd{obs[}\hlnum{1}\hlstd{],}
           \hlkwc{lnR}\hlstd{=}\hlkwd{as.numeric}\hlstd{(}\hlkwd{solve}\hlstd{(}\hlkwd{t}\hlstd{(X)} \hlopt{%*%} \hlstd{V.inv} \hlopt{%*%} \hlstd{X)} \hlopt{%*%}
                          \hlstd{(}\hlkwd{t}\hlstd{(X)} \hlopt{%*%} \hlstd{V.inv} \hlopt{%*%} \hlstd{E)),}
           \hlkwc{varlnR}\hlstd{=}\hlkwd{as.numeric}\hlstd{(}\hlkwd{solve}\hlstd{(}\hlkwd{t}\hlstd{(X)} \hlopt{%*%} \hlstd{V.inv} \hlopt{%*%} \hlstd{X))))}
\hlstd{\}}
\end{alltt}
\end{kframe}
\end{knitrout}

Which is called by this function which combines the meta-data based on
keys

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{## takes raw meta-data and a vector of covariates and combines data}
\hlcom{## using lajeunesse (2011). Returns a dataframe with keys, the}
\hlcom{## observation number, the pooled RR and its variance}

\hlstd{poolRR} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{D}\hlstd{,} \hlkwc{covariates}\hlstd{) \{}
  \hlcom{## first split the data up into relevant groups:}
  \hlkwa{if}\hlstd{(}\hlkwd{length}\hlstd{(covariates)} \hlopt{==} \hlnum{1}\hlstd{) \{}
    \hlstd{dd} \hlkwb{<-} \hlkwd{split}\hlstd{(D,} \hlkwd{paste}\hlstd{(D}\hlopt{$}\hlstd{study,}
                         \hlstd{D[,covariates],}
                         \hlstd{D}\hlopt{$}\hlstd{year_conv,}
                         \hlstd{D}\hlopt{$}\hlstd{mult_yrs,}
                         \hlstd{D}\hlopt{$}\hlstd{duplicate_trt_dat,}
                         \hlkwc{sep}\hlstd{=}\hlstr{';'}\hlstd{))}
  \hlstd{\}}
  \hlkwa{if}\hlstd{(}\hlkwd{length}\hlstd{(covariates)} \hlopt{>} \hlnum{1}\hlstd{) \{}
    \hlstd{dd} \hlkwb{<-} \hlkwd{split}\hlstd{(D,} \hlkwd{paste}\hlstd{(D}\hlopt{$}\hlstd{study,}
                         \hlkwd{apply}\hlstd{(D[,covariates],} \hlnum{1}\hlstd{, paste,} \hlkwc{collapse}\hlstd{=}\hlstr{':'}\hlstd{),}
                         \hlstd{D}\hlopt{$}\hlstd{year_conv,}
                         \hlstd{D}\hlopt{$}\hlstd{mult_yrs,}
                         \hlstd{D}\hlopt{$}\hlstd{duplicate_trt_dat,}
                         \hlkwc{sep}\hlstd{=}\hlstr{';'}\hlstd{))}
  \hlstd{\}}

  \hlcom{## combine response ratios for studies that share treatments using}
  \hlcom{## the method given in Lajeunesse (2011):}
  \hlstd{share.trts} \hlkwb{<-} \hlkwd{which}\hlstd{(}\hlkwd{sapply}\hlstd{(}\hlkwd{strsplit}\hlstd{(}\hlkwd{names}\hlstd{(dd),} \hlkwc{split}\hlstd{=}\hlstr{';'}\hlstd{),}
                             \hlkwa{function}\hlstd{(}\hlkwc{x}\hlstd{)} \hlkwd{rev}\hlstd{(x)[[}\hlnum{1}\hlstd{]])} \hlopt{==} \hlstr{'yes'}\hlstd{)}

  \hlstd{dd[share.trts]} \hlkwb{<-} \hlkwd{lapply}\hlstd{(dd[share.trts],}
                           \hlstd{poolCommonControl)}

  \hlstd{dd[}\hlopt{-}\hlstd{share.trts]} \hlkwb{<-}
    \hlkwd{lapply}\hlstd{(dd[}\hlopt{-}\hlstd{share.trts],} \hlkwa{function}\hlstd{(}\hlkwc{x}\hlstd{) x[,}\hlkwd{c}\hlstd{(}\hlstr{'obs'}\hlstd{,} \hlstr{'lnR'}\hlstd{,} \hlstr{'varlnR'}\hlstd{)])}
  \hlkwd{return}\hlstd{(}\hlkwd{do.call}\hlstd{(rbind, dd))}
\hlstd{\}}
\end{alltt}
\end{kframe}
\end{knitrout}

This is all brought together by this long, unwieldy function that
returns the data structure necessary for the JAGS model including the
indexes for looping and arrays of RR, variances and covariates is
applicable

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{## makes data arrays for Bayesian analysis; takes the raw}
\hlcom{## meta-data set, whether to apply the lajeunesse method to combine RR}
\hlcom{## with shared controls, and a vector of explanatory variables. The}
\hlcom{## default for covariates is "Crop.species" so RR are never combined}
\hlcom{## between crop species. Returns a list of the number of studies}
\hlcom{## (Nstudy), the number of years for each study (Nyear), a matrix}
\hlcom{## (study, year) of the number of observations in each year (Nobs),}
\hlcom{## and an array (study, year, observation) of the RR and their}
\hlcom{## variances, and an array of covariates if applicable}

\hlstd{makeData} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{meta.dat}\hlstd{,}
                     \hlkwc{lajeunesse} \hlstd{=} \hlnum{TRUE}\hlstd{,}
                     \hlkwc{covariates}\hlstd{=}\hlkwd{c}\hlstd{(}\hlstr{'Crop.species'}\hlstd{)) \{}

  \hlstd{convert2Int} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{x}\hlstd{)} \hlkwd{seq_along}\hlstd{(x)[}\hlkwd{match}\hlstd{(x,} \hlkwd{unique}\hlstd{(x))]}

  \hlkwa{if}\hlstd{(lajeunesse) \{}
    \hlstd{dd} \hlkwb{<-} \hlkwd{poolRR}\hlstd{(meta.dat,}
                 \hlkwc{covariates}\hlstd{=covariates)}
    \hlstd{keys.split} \hlkwb{<-} \hlkwd{sapply}\hlstd{(}\hlkwd{rownames}\hlstd{(dd), strsplit,} \hlkwc{split}\hlstd{=}\hlstr{';'}\hlstd{)}
    \hlstd{get} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{i}\hlstd{)} \hlkwd{as.vector}\hlstd{(}\hlkwd{sapply}\hlstd{(keys.split,} \hlkwa{function}\hlstd{(}\hlkwc{x}\hlstd{) x[i]))}
    \hlstd{study} \hlkwb{<-} \hlkwd{get}\hlstd{(}\hlnum{1}\hlstd{)}
    \hlstd{year} \hlkwb{<-} \hlkwd{get}\hlstd{(}\hlnum{3}\hlstd{)}
  \hlstd{\}} \hlkwa{else} \hlstd{\{}
    \hlstd{dd} \hlkwb{<-} \hlstd{meta.dat}
    \hlstd{study} \hlkwb{<-} \hlkwd{as.character}\hlstd{(meta.dat}\hlopt{$}\hlstd{study)}
    \hlstd{year} \hlkwb{<-} \hlstd{meta.dat}\hlopt{$}\hlstd{year_conv}
  \hlstd{\}}

  \hlcom{## make study index}
  \hlstd{ind.study} \hlkwb{<-} \hlkwd{convert2Int}\hlstd{(study)}

  \hlcom{## make year index}
  \hlstd{ind.year} \hlkwb{<-} \hlkwd{rep}\hlstd{(}\hlnum{NA}\hlstd{,} \hlkwd{nrow}\hlstd{(dd))}
  \hlkwa{for}\hlstd{(i} \hlkwa{in} \hlkwd{unique}\hlstd{(ind.study))}
    \hlstd{ind.year[ind.study}\hlopt{==}\hlstd{i]} \hlkwb{<-} \hlkwd{convert2Int}\hlstd{(year[ind.study}\hlopt{==}\hlstd{i])}

  \hlcom{## make obs index}
  \hlstd{ind.obs} \hlkwb{<-} \hlkwd{rep}\hlstd{(}\hlnum{NA}\hlstd{,} \hlkwd{nrow}\hlstd{(dd))}
  \hlkwa{for}\hlstd{(i} \hlkwa{in} \hlkwd{unique}\hlstd{(ind.study))}
    \hlkwa{for}\hlstd{(j} \hlkwa{in} \hlkwd{unique}\hlstd{(ind.year[ind.study}\hlopt{==}\hlstd{i]))}
      \hlstd{ind.obs[ind.year}\hlopt{==}\hlstd{j} \hlopt{&} \hlstd{ind.study}\hlopt{==}\hlstd{i]} \hlkwb{<-}
        \hlkwd{seq_len}\hlstd{(}\hlkwd{sum}\hlstd{(ind.year}\hlopt{==}\hlstd{j} \hlopt{&} \hlstd{ind.study}\hlopt{==}\hlstd{i))}

  \hlcom{## make response ratio and vi matrices}
  \hlstd{makeMat} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{d.vec}\hlstd{) \{}
    \hlstd{mat} \hlkwb{<-} \hlkwd{array}\hlstd{(}\hlnum{NA}\hlstd{,} \hlkwc{dim}\hlstd{=}\hlkwd{c}\hlstd{(}\hlkwd{max}\hlstd{(ind.study),}
                           \hlkwd{max}\hlstd{(ind.year),}
                           \hlkwd{max}\hlstd{(ind.obs)))}
    \hlstd{mat[}\hlkwd{cbind}\hlstd{(ind.study, ind.year, ind.obs)]} \hlkwb{<-} \hlstd{d.vec}
    \hlkwd{return}\hlstd{(mat)}
  \hlstd{\}}
  \hlstd{RR.mat} \hlkwb{<-} \hlkwd{makeMat}\hlstd{(dd}\hlopt{$}\hlstd{lnR)}
  \hlstd{vi.mat} \hlkwb{<-} \hlkwd{makeMat}\hlstd{(dd}\hlopt{$}\hlstd{varlnR)}

  \hlcom{## make index bounds}
  \hlstd{nstudy} \hlkwb{<-} \hlkwd{max}\hlstd{(ind.study)}
  \hlstd{nyear}  \hlkwb{<-} \hlkwd{apply}\hlstd{(RR.mat[,,}\hlnum{1}\hlstd{],}  \hlnum{1}\hlstd{,} \hlkwa{function}\hlstd{(}\hlkwc{x}\hlstd{)} \hlkwd{sum}\hlstd{(}\hlopt{!}\hlkwd{is.na}\hlstd{(x)))}
  \hlstd{nobs}   \hlkwb{<-} \hlkwd{apply}\hlstd{(RR.mat[,,],} \hlnum{1}\hlopt{:}\hlnum{2}\hlstd{,} \hlkwa{function}\hlstd{(}\hlkwc{x}\hlstd{)} \hlkwd{sum}\hlstd{(}\hlopt{!}\hlkwd{is.na}\hlstd{(x)))}

  \hlstd{covs} \hlkwb{<-} \hlnum{NA}
  \hlstd{num.cats} \hlkwb{<-} \hlnum{NA}
  \hlcom{## covariates (only for lajeunesse)}
  \hlkwa{if}\hlstd{(lajeunesse) \{}
    \hlstd{num.covs} \hlkwb{<-} \hlkwd{length}\hlstd{(covariates)}
    \hlkwa{if}\hlstd{(num.covs} \hlopt{>} \hlnum{1}\hlstd{) \{}
      \hlstd{covs} \hlkwb{<-} \hlkwd{sapply}\hlstd{(}\hlkwd{get}\hlstd{(}\hlnum{2}\hlstd{), strsplit,} \hlkwc{split}\hlstd{=}\hlstr{':'}\hlstd{)}
      \hlstd{getCov} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{cov}\hlstd{)}
        \hlkwd{as.numeric}\hlstd{(}\hlkwd{as.factor}\hlstd{(}\hlkwd{sapply}\hlstd{(covs,} \hlkwa{function}\hlstd{(}\hlkwc{x}\hlstd{) x[cov])))}
      \hlstd{covs} \hlkwb{<-} \hlkwd{lapply}\hlstd{(}\hlnum{1}\hlopt{:}\hlstd{num.covs,} \hlkwa{function}\hlstd{(}\hlkwc{cov}\hlstd{)}
        \hlkwd{make.mat}\hlstd{(}\hlkwd{getCov}\hlstd{(cov)))}
      \hlkwd{names}\hlstd{(covs)} \hlkwb{<-} \hlstd{covariates}
      \hlstd{num.cats} \hlkwb{<-} \hlkwd{sapply}\hlstd{(covs, max,} \hlkwc{na.rm}\hlstd{=}\hlnum{TRUE}\hlstd{)}
    \hlstd{\}}
  \hlstd{\}}

  \hlkwd{return}\hlstd{(}\hlkwd{list}\hlstd{(}\hlkwc{Nstudy}\hlstd{=nstudy,}
              \hlkwc{Nyear}\hlstd{=nyear,}
              \hlkwc{Nobs}\hlstd{=nobs,}
              \hlkwc{V}\hlstd{=vi.mat,}
              \hlkwc{RR}\hlstd{=RR.mat,}
              \hlkwc{covariates}\hlstd{=covs,}
              \hlkwc{num.cats}\hlstd{=num.cats))}
\hlstd{\}}
\end{alltt}
\end{kframe}
\end{knitrout}

We then pacakges everything up in a function that also passes the JAGS
parameters (init values, thinning rate etc.) and runs the model.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{## takes preped data and a scale parameter and packages the data and}
\hlcom{## parameters for JAGS, then runs the analysis. Returns the JAGS}
\hlcom{## summary output}
\hlstd{runAnalysis} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{dd}\hlstd{,} \hlkwc{scale}\hlstd{) \{}

  \hlstd{my.inits} \hlkwb{<-} \hlkwa{function}\hlstd{() \{}
    \hlkwd{list}\hlstd{()}
  \hlstd{\}}
  \hlstd{my.params} \hlkwb{<-} \hlkwd{get.params}\hlstd{()}

  \hlstd{dd} \hlkwb{<-} \hlkwd{list}\hlstd{(}\hlkwc{data}\hlstd{=dd,} \hlkwc{inits}\hlstd{=my.inits,} \hlkwc{params}\hlstd{=}\hlkwd{get.params}\hlstd{())}

  \hlstd{bugs} \hlkwb{<-} \hlkwd{run.model}\hlstd{(dd,}
                    \hlkwc{n.thin}\hlstd{=scale,}
                    \hlkwc{n.iter}\hlstd{=(}\hlnum{1e4}\hlstd{)}\hlopt{*}\hlstd{scale,}
                    \hlkwc{n.burnin}\hlstd{=}\hlnum{1e1}\hlopt{*}\hlstd{scale,}
                    \hlkwc{n.chains}\hlstd{=}\hlnum{3}\hlstd{)}

  \hlkwd{return}\hlstd{(}\hlkwd{list}\hlstd{(}\hlkwc{data}\hlstd{=dd,}
       \hlkwc{bugs}\hlstd{=bugs,}
       \hlkwc{summary}\hlstd{=bugs}\hlopt{$}\hlstd{BUGSoutput}\hlopt{$}\hlstd{summary))}
\hlstd{\}}
\end{alltt}
\end{kframe}
\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{print}\hlstd{(out}\hlopt{$}\hlstd{bugs,} \hlkwc{dig}\hlstd{=}\hlnum{3}\hlstd{)}
\end{alltt}
\begin{verbatim}
## Inference for Bugs model at "model.jags", fit using jags,
##  3 chains, each with 1e+05 iterations (first 100 discarded), n.thin = 10
##  n.sims = 29970 iterations saved
##            mu.vect sd.vect      2.5%       25%       50%       75%
## cv.obs       1.154   0.134     0.909     1.061     1.149     1.241
## exp.mu       0.808   0.019     0.771     0.795     0.808     0.821
## sigma        0.189   0.023     0.145     0.172     0.188     0.204
## deviance -1293.537  36.474 -1362.740 -1318.528 -1294.309 -1269.421
##              97.5%  Rhat n.eff
## cv.obs       1.434 1.001  5100
## exp.mu       0.846 1.001 27000
## sigma        0.236 1.001  9400
## deviance -1219.532 1.001 20000
## 
## For each parameter, n.eff is a crude measure of effective sample size,
## and Rhat is the potential scale reduction factor (at convergence, Rhat=1).
## 
## DIC info (using the rule, pD = var(deviance)/2)
## pD = 665.1 and DIC = -628.4
## DIC is an estimate of expected predictive error (lower deviance is better).
\end{verbatim}
\end{kframe}
\end{knitrout}

exp.mu is the mean ratio of organic and conventional yields, which is
around $80.1\%$ with $95\%$ credible intervals ranging from around
$77--85\%$

Sigma is the between study variance, and cv.obs is the coefficient of
variation of $\gamma$ distribution the within-study variance
parameters are drawn from.

Values of Rhat < 1.1 indicate convergence, whcih we We can see from
inspecting the posteriors

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{ggs_density}\hlstd{(inspect.out)}
\end{alltt}
\end{kframe}
\includegraphics[width=\maxwidth]{figure/densityPlots-1}

\end{knitrout}

The chains look wonderfully grassy

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{ggs_traceplot}\hlstd{(inspect.out)}
\end{alltt}
\end{kframe}
\includegraphics[width=\maxwidth]{figure/traceplots-1}

\end{knitrout}

\section{Explanatory variables}
\label{sec:variables}
We also extended this model in order to accommodate analyses of study
characteristics such as crop type and management practices. We analyze
these additional explanatory variables one at a time because not all
studies reported all explanatory variables. In these analyses, for
cases where multiple organic treatments represented different
categories for a specific explanatory variable, they could not be
combined using Lajeunesse's method \citep{Lajeunesse2011}. The
potential bias resulting from non-independence of the response ratios
in these cases, however, would be minimized by the fact that they are
not pooled together in the analysis \citep{Lajeunesse2011}.

Letting $h$ index the categories for a particular explanatory variable
(e.g., crop species), we then have:
%
\begin{equation}
  \begin{split}
    \label{eq:fullModel_cov}
    y_{hijk} &=
    \mu + \gamma_h + \alpha_i + \eta_{ijk} +  \beta_{ij} + \epsilon_{ijk}\\
  \end{split}
\end{equation}
%

where $\gamma_h$ is the effect of the $h^{\mathrm{th}}$ category, and
the rest of the model parallels that given in Eq.~\ref{eq:fullModel}.

The model including explanatory variables in JAGS is

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{verbatim}
## model {
##     for(study in 1:Nstudy) {
##       for(year in 1:Nyear[study]) {
##         for(obs in 1:Nobs[study,year]) {
##           P[study,year,obs] <- 1/V[study,year,obs]
##           RR[study,year,obs] ~ dnorm(mu.RR[study,year,obs], P[study,year,obs])
##           mu.RR[study,year,obs] ~ dnorm(mu.study[study,year,obs],
##                                         tau.obs[study])
##           mu.study[study,year,obs] ~ dnorm(mu[cov[study,year,obs]], tau)
##         }
##       }
##       tau.obs[study] ~ dgamma(shape.obs, scale.obs)
##     }
## 
##     ## set priors on covariates
##     for(covs in 1:Ncov) {
##       mu[covs] ~ dnorm(0, 1E-3)
##       exp.mu[covs] <- exp(mu[covs])
##     }
##     tau <- 1 / (sigma * sigma)
##     sigma ~ dunif(0, 100)
## 
##     shape.obs <- (1/cv)^2
##     cv ~ dunif(0, 100)
##     scale.obs <- (1/in.scale)^2
##     in.scale ~ dunif(0, 100)
## 
##   }
\end{verbatim}
\end{kframe}
\end{knitrout}

Where the parameters names match the notation in
Equ.~\ref{eq:fullModel} and cov is a matrix of explanatory variables
assiciated with each response ratio.

To re-run the models with explanatory variables, we just added each
variable one-by-one to the covariates argument of the makeData()
function.

\bibliography{org_v_conv} \clearpage

\end{document}

### 2017-04-19 analysis/mainR.R

@@ -1,31 +1,16 @@
 rm(list=ls())
-## prep data
-## 1) drop non bee/syrphid data
-## 2) create networks with mean taken over sampling rounds in each
-## year
-## 3) calculate network metrics for whole landscape
-## 4) add trait data to specimen data
 analysis.dir <- "~/Dropbox/hedgerow_assembly/"
-
-source(file.path(analysis.dir, "dataPrep/dataPrep.R"))
 
 ##***************************************************************
 ## change points
 ##***************************************************************
 
-## create graphs for change point analysis
-source(file.path(analysis.dir,
-                 'analysis/changePoint/dataPrep.R'))
-
-## run python code for calculating network change points
-## plot networks and network communities
-source(file.path(analysis.dir,
-                 'analysis/changePoint/plotting/networks.R'))
+## data is preped and change point analysis is run through the bash
+## script mainChangePoint.sh
 
 ## binomial analysis in comparisons.R, no resulting plots
 source(file.path(analysis.dir,
                  'analysis/changePoint/comparisons.R'))
-
 
 ## exploratory analysis in  src/core.R
 

### 2017-04-19 analysis/variability/plotting/beta-int.R

@@ -6,41 +6,8 @@ source("plotting/src/CIplotting.R")
 source('../networkLevel/src/misc.R', chdir = TRUE)
 load('saved/dissMods.Rdata')
 
-## ************************************************************
-## interaction dispersion, k, plants
-## ************************************************************
 plants$SiteStatus <- "all"
 pols$SiteStatus <- "all"
-
-dd.plants.k <- expand.grid(k=seq(
-                             from= min(plants$k),
-                             to= max(plants$k),
-                             length=10),
-                           ## SiteStatus= c("control", "maturing",
-                           ## "mature"),
-                           SiteStatus= c("all"),
-                           Dist= 0)
-
-plants.k.pi <- predict.int(mod= mod.plants,
-                           dd=dd.plants.k,
-                           y="Dist",
-                           family="gaussian")
-
-## plot.predict.div(new.dd=plants.k.pi,
-##                  ylabel="Interaction dispersion",
-##                  dats=plants,
-##                  xs="k",
-##                  y1="Dist",
-##                  xlabel="Plant network position (k)",
-##                  legend.loc="topright",
-##                  height=5,
-##                  width=5,
-##                  x.adj=0.5,
-##                  scaled=FALSE,
-##                  treatments="all",
-##                  col.lines="black",
-##                  f.path = 'figures/intdisper')
-
 
 ## ************************************************************
 ## interaction dispersion, closeness, plants
@@ -57,21 +24,6 @@ plants.closeness.pi <- predict.int(mod= mod.plants.close,
                                    dd=dd.plants.closeness,
                                    y="Dist",
                                    family="gaussian")
-
-## plot.predict.div(new.dd=plants.closeness.pi,
-##                  ylabel="Interaction dispersion",
-##                  dats=plants,
-##                  xs="closeness",
-##                  y1="Dist",
-##                  xlabel="Plant closeness",
-##                  legend.loc="topright",
-##                  height=5,
-##                  width=5,
-##                  x.adj=0.5,
-##                  scaled=FALSE,
-##                  treatments="all",
-##                  col.lines="black",
-##                  f.path = 'figures/intdisper')
 
 
 ## ************************************************************

### 2017-04-19 dataPrep/dataPrep.R

@@ -37,8 +37,9 @@ spec <- spec[!spec$SiteStatus %in% to.drop.status,]
 
 
 tot.spec <- nrow(spec)
-
+print("proportion syrphids")
 nrow(spec[spec$Family == 'Syrphidae',])/tot.spec
+print("proprotion bee")
 nrow(spec[spec$BeeNonbee == 'bee',])/tot.spec
 
 ## subset to just bees and syrphids
@@ -50,19 +51,25 @@ spec$SiteStatus[spec$SiteStatus == "restored"] <- "maturing"
 
 
 ## total specimens
+print("total specimens")
 nrow(spec)
 
 ## total species
+print("total species")
 length(unique(spec$GenusSpecies))
 
 ## sampling dates
+print("sampling dates")
 length(unique(paste(spec$Site, spec$Date)))
 
 ## families and genera
+print("total families")
 length(unique(spec$Family))
+print("total genera")
 length(unique(spec$Genus))
 
 ## interactions
+print("total interactions")
 length(unique(paste(spec$GenusSpecies, spec$PlantGenusSpecies)))
 
 ## *************************************************

### 2017-04-21 analysis/writeup/writeup.tex
debugging change point \analysis

@@
 \newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+\let\hlipl\hlkwb
 
@@
 
-\bibliographystyle{rspublicnatsort}
-
 %% for inline R code: if the inline code is not correctly parsed, you will see a message
@@
 \begin{document}
-\title{Diversification practices reduce organic to conventional yield
-gap: A walkthrough of the anlaysis}
+\title{Opportunistic attachment assembles plant-pollinator networks: A walkthrough of the anlaysis}
 \author{Lauren Ponisio}
@@
 
-In our study
-\href{http://nature.berkeley.edu/~lponisio/wp-content/uploads/2014/08/Ponisio2014.pdf}{\textcolor{blue}
-{Ponisio
-et al.~2015}} we compared the yields or organic and conventional
-agriculture using a custom-build meta-analytic model. Here I explain
-the analytic differences between our study and others, example our
-modeling decisions, and walk through our code. Below is the output of
-the final meta-analytic model, which will be described in detail in
-this document.
+In our study we examing the temporal dynamcis of plant-pollinator
+network assembly using vareity of different methods include 1) network
+change point detection 2) node/species-level position variation, 3)
+species and interaction turnover 4) network-level metrics and 5)
+extinction simulations. We are commited to reproducible science and
+all analytical code will be maintained on github, along wiht this
+write up (as included as a supplment to the publication).
+
+The entire analysis is executable from the main.sh file
+
+The change point analysis which is a mix of python and R. hedgerows.py
+(the meat of the change point analysis) runs in paralell on two cores
+(which can be modified in the script if needed), but will likely take
+a few hours, depending on your machine.
 
@@
 \begin{alltt}
-\hlkwd{print}\hlstd{(out}\hlopt{$}\hlstd{bugs,} \hlkwc{dig}\hlstd{=}\hlnum{3}\hlstd{)}
+ RScript changePoint/dataPrep.R
+ python changePoint/hedgerows.py
+ RScript changePoint/prepChangePointOutput.R
+ python changePoint/postChangePoint.py saved/consensus.txt
+ python changePoint/convertfiles.py
+ RScript changePoint/plotting/networks.R
 \end{alltt}
-\begin{verbatim}
-## Inference for Bugs model at "model.jags", fit using jags,
-##  3 chains, each with 1e+05 iterations (first 100 discarded), n.thin = 10
-##  n.sims = 29970 iterations saved
-##            mu.vect sd.vect      2.5%       25%       50%       75%
-## cv.obs       1.154   0.134     0.909     1.061     1.149     1.241
-## exp.mu       0.808   0.019     0.771     0.795     0.808     0.821
-## sigma        0.189   0.023     0.145     0.172     0.188     0.204
-## deviance -1293.537  36.474 -1362.740 -1318.528 -1294.309 -1269.421
-##              97.5%  Rhat n.eff
-## cv.obs       1.434 1.001  5100
-## exp.mu       0.846 1.001 27000
-## sigma        0.236 1.001  9400
-## deviance -1219.532 1.001 20000
-## 
-## For each parameter, n.eff is a crude measure of effective sample size,
-## and Rhat is the potential scale reduction factor (at convergence, Rhat=1).
-## 
-## DIC info (using the rule, pD = var(deviance)/2)
-## pD = 665.1 and DIC = -628.4
-## DIC is an estimate of expected predictive error (lower deviance is better).
-\end{verbatim}
 \end{kframe}
@@
 
-\section{Where it all began, Seufert et al.~2011}
-\label{sec:beginning}
-
-How organic agriculture may contribute to world food production has
-been subject to vigorous debate over the past decade. Early reviews
-comparing organic to conventional agriculture found yield gaps of
-$8-9$\% in developed countries \citep{Stanhill1990, Badgley2007} but
-yield gains of as much as $180$\% in developing countries. Two recent
-meta-analyses, however, found organic yields to be $20-25$\% lower
-than conventional yields \citep{dePonti2012, Seufert2012}. These
-studies used different criteria for selecting the data to be included,
-but importantly each of the above studies used different analytical
-methods to combine the data across the different sub-studies. The
-studies comparing organic and conventional yields systems often
-reported yields from multiple crops across several years. In addition,
-they tended to compare multiple treatments (usually organic) to one
-control treatment (usually conventional). For example from
-\cite{denison2004crop}:
+And all of the subsequent R analyses (takes several hours, mostly for
+building null communities for standardizing the network level metrics)
 
@@
 \begin{alltt}
-\hlstd{samp}
-\end{alltt}
-\begin{verbatim}
-##       study   crop conv org year
-## 100 study13  maize  CMT OMT 2001
-## 101 study13  maize  CMT OMT 1999
-## 102 study13 tomato  CWT OMT 2002
-## 103 study13  maize  LMT OMT 2000
-## 104 study13  maize  CMT OMT 1996
-## 105 study13  maize  LMT OMT 2001
-## 106 study13 tomato  LMT OMT 1999
-## 107 study13  maize  CMT OMT 1997
-## 108 study13 tomato  CMT OMT 2001
-## 109 study13 tomato  CMT OMT 1995
-## 110 study13 tomato  LMT OMT 1995
-## 111 study13  maize  LMT OMT 1999
-## 112 study13  maize  LMT OMT 2002
-## 113 study13  maize  LMT OMT 1994
-## 114 study13  maize  CMT OMT 1994
-## 115 study13  maize  CMT OMT 2002
-## 116 study13 tomato  LMT OMT 1998
-## 117 study13 tomato  CMT OMT 1994
-## 118 study13 tomato  CMT OMT 1999
-## 119 study13 tomato  CWT OMT 1994
-## 120 study13 tomato  CWT OMT 1995
-## 121 study13 tomato  CMT OMT 1996
-## 122 study13 tomato  LMT OMT 1996
-## 123 study13 tomato  CWT OMT 1996
-## 124 study13  maize  LMT OMT 1995
-## 125 study13  maize  CMT OMT 1995
-## 126 study13  maize  LMT OMT 1996
-## 127 study13 tomato  LMT OMT 1994
-## 128 study13 tomato  CWT OMT 1998
-## 129 study13 tomato  CMT OMT 1998
-## 130 study13 tomato  CMT OMT 1997
-## 131 study13 tomato  LMT OMT 1997
-## 132 study13 tomato  LMT OMT 2001
-## 133 study13 tomato  CWT OMT 2001
-## 134 study13 tomato  CWT OMT 1997
-## 135 study13  maize  LMT OMT 1997
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-
-Using the organic to conventional yield comparisons without taking
-into account the underlying data structure can lead to potential
-pseudo-replication and an understated Type 1 error rate. We used a
-randomization test to estimate the Type I error rate of the Seufert et
-al.~analysis.  We forced the null hypothesis to be true by randomly
-re-assigning the 'organic' and 'conventional' labels for each study
-and then using the R package {\tt Metafor} \citep{metafor} to
-implement a random effects meta-analysis on each randomized dataset.
-Repeating this procedure $10^5$ times enabled us to determine the Type
-I error rate (false rejection) resulting from not accounting for the
-hierarchical structure of the data.  In over $50\%$ of simulations,
-the null hypothesis was rejected using a nominal Type I error rate of
-0.05 (Fig.~\ref{fig:permutation}).  In other words, even if organic
-and conventional yields are known not to be different, applying the
-model used by \cite{Seufert2012} for these data would lead to the
-conclusion that they are significantly different in over $50\%$ of
-cases. This means that the actual Type I error rate is inflated
-relative to what was reported, leading to the following related
-statements: the significance levels were overstated; the confidence
-intervals were underestimated; the uncertainty was not fully accounted
-for.
-
-\begin{figure} \centering
-\includegraphics[width=0.5\textwidth]{figure/permutation.pdf}
-\caption{The distribution of $p$-values when the null hypothesis was
-forced to be true using the data and analysis type present in
-Seufert et al.\  \cite{Seufert2012}.  If the analysis procedure was
-valid for these data, the distribution of P-values should be
-uniform between 0 and 1.  Instead it is sharply shifted toward low
-P-values.  In over $50\%$ of simulations, the null hypothesis was
-rejected using a nominal Type I error rate of 0.05 (red region
-above).}
-\label{fig:permutation}
-\end{figure}
-
-The de Ponti et al.\ \citep{dePonti2012} study had similar issues with
-pseudo-replication. Additionally they did not account for the sampling
-variance within studies, which is the recommended practice to deal
-with unequal variances in the sample of studies \citep{Gurevitch1999}.
-
-Given these methodological and data-related critiques, a new study was
-needed to produce a more robust estimate of the gap between organic
-and conventional yields.  We developed a hierarchical meta-analytic
-framework that overcomes the methodological pitfalls of previous
-studies by accounting for both the multi-level nature of the data and
-the yield variation within studies.  Furthermore, via a literature
-search we compiled a more extensive and up-to-date meta-dataset,
-comprising 1071 organic to conventional yield comparisons from 115
-studies --- over three times the number of observations of any of the
-previous analyses. Our meta-dataset includes studies from 38 countries
-and 52 crop species over a span of 35 years.
-
-\section{Building our meta-analytic model}
-\label{sec:model-building}
-
-We built a hierarchical meta-analytic model to generate an estimate of
-the yield gap. Following standard practice, we compared the natural
-log of the ratios between organic and conventional yields (the
-``response ratio'') across studies \citep{Hedges1999,
-Seufert2012}. The response ratio is more normally distributed than the
-raw ratio and independent of the units of measurement used within a
-study and, thus, comparable across studies \citep{Hedges1999}.
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{plot}\hlstd{(}\hlkwd{density}\hlstd{(meta.dat}\hlopt{$}\hlstd{lnR),} \hlkwc{main}\hlstd{=}\hlstr{"Distribution of log response ratios"}\hlstd{)}
-\end{alltt}
-\end{kframe}
-\includegraphics[width=\maxwidth]{figure/RR_density-1}
-
-\end{knitrout}
-
-We constructed a hierarchical regression model to account for
-the dependencies in the yield data. We expanded on the traditional
-random effects model \citep{Hedges1985} by considering three
-additional sources of random variation (i.e., random effects): 1)
-between studies, 2) within a study between years, and 3) within a year
-between response ratios (e.g., across replicated trials of a crop
-planted at different times in the season). We also considered whether
-the variances of the random effect distributions for 2) and 3) were
-shared across studies, or study-specific.
-
-The full possible model, prior to model selection, with all sources of
-random variation is
-
-\begin{equation}
-  \begin{split}
-    \label{eq:fullModel}
-    y_{ijk} &= \mu + \alpha_i + \beta_{ij} +  \eta_{ijk} +
-    \epsilon_{ijk}  \\
-    \alpha_{i} &\sim N(0, \sigma^2_{\alpha})\\
-    \beta_{ij} &\sim N(0, \sigma^2_{\beta}[i])\\
-    \eta_{ijk} &\sim N(0, \sigma^2_{\eta}[i])\\
-    \epsilon_{ijk} &\sim N(0, S_{ijk}])\\
-    \sigma^2_{\beta}[i] &\sim \Gamma(CV_\beta, scale_\beta)\\
-    \sigma^2_{\eta}[i] &\sim \Gamma(CV_\eta, scale_\eta)
-  \end{split}
-\end{equation}
-% where $y_{ijk}$ is the observed magnitude of the $k^{\mathrm{th}}$
-response ratio from the $j^{\mathrm{th}}$ year of the
-$i^{\mathrm{th}}$ study, $\mu$ is the mean response ratio across
-studies, $\alpha_{i}$ is the effect of $i^{\mathrm{th}}$ study,
-$\beta_{ij}$ is the effect of $j^{\mathrm{th}}$ year of the
-$i^{\mathrm{th}}$ study, $\eta_{ijk}$ is the effect of the
-$k^{\mathrm{th}}$ response ratio of the $j^{\mathrm{th}}$ year of
-$i^{\mathrm{th}}$ study, and $\epsilon_{ijk}$ is the
-residual. $\sigma^2_{\alpha}$ is the between study variance,
-$\sigma^2_{\beta}[i]$ is the between year variance of study $i$,
-$\sigma^2_{\eta}[i]$ is the within year, between response ratio
-variance of study $i$, and $S_{ijk}$ is the variance of response ratio
-$ijk$ as reported by its study. $CV_\beta$ and $CV_\eta$ and
-$scale_\beta$ and $scale_\eta$ are the coefficient of variation and
-scale parameters of the gamma distributions of the study-specific
-between- and within-year variances.  When response ratios that shared
-a common control were combined, $y_{ijk}$ corresponds to the aggregate
-within-study response ratio \citep[Eq.~3,][]{Lajeunesse2011} and
-$S_{ijk}$ is its pooled variance \citep[Eq.~8,][]{Lajeunesse2011}.
-
-\subsection{Parameter inclusion}
-To determine the levels of hierarchy supported by the data, we
-sequentially added random effects and examined the posteriors of the
-parameters to determine the support for their inclusion. We also
-confirmed our selection with Deviance Information Criterion (DIC). The
-DIC can be problematic for hierarchical models because the effective
-number of parameters is not clearly defined \citep{gelman2006data,
-kery2012bayesian}. The DIC was therefore used in combination with a
-visual examination of the posterior distributions of the parameters to
-select the best supported model.
-
-For the variance within and between year random effect distributions,
-we considered two parameterizations: 1) the variance terms, denoted
-$\sigma^2_{\eta}$ and $\sigma^2_{\beta}$, respectively, were shared
-across all studies each with a Uniform(0,100) prior, and 2) the
-variance terms were study-specific (i.e., $\sigma^2_{\eta}[i]$ and
-$\sigma^2_{\beta}[i]$). In the latter case, the study-specific
-precision terms (1/variance) were assumed to be distributed according
-to a gamma distribution whose parameters were
-estimated. Uniform(0,100) priors were used for the coefficient of
-variation ($1/\sqrt{shape}$) and the square root of the scale.
-
-We first added a random effect of study and examined the posterior for
-$\sigma_{\alpha}$ (the standard deviation of the common distribution
-from which the study effects are drawn). The posterior was clearly
-differentiated from zero (Fig.~\ref{fig:posteriors}a). We next added
-random variation within a year and examined $\sigma_{\eta}$. We found
-it was also clearly different from zero
-(Fig.~\ref{fig:posteriors}b). The DIC was also smaller than when only
-a random effect of study was included (Tab.~\ref{table:params}). Next
-we allowed the within-year precisions to be study-specific and follow
-a gamma distribution. We examined the coefficient of variation of the
-gamma distribution and found it was clearly differentiated from zero
-(Fig.~\ref{fig:posteriors}c). The DIC was also smaller than when a
-single within year effect was shared across studies
-(Tab.~\ref{table:params}). Lastly, we added a between year random
-effect and examined $\sigma_{\beta}$. The posterior was concentrated
-at zero (Fig.~\ref{fig:posteriors}d) so we concluded there was
-insufficient support for including it in the model. The estimate of
-the yield gap and its uncertainty did not differ substantially from
-when no between year effect was included
-(Fig.~\ref{fig:model_selection}), the DIC, however, was marginally
-smaller then when no between year random effect was included
-(Tab.~\ref{table:params}).
-
-\clearpage
-
-\begin{figure}
-\centering
-\includegraphics[width=0.8\textwidth]{figure/posteriors.pdf}
-\caption{The posterior distributions for the random effect of a)
-study ($\sigma_{\alpha}$); b) response ratios within a year
-($\sigma_{\eta}$); c) response ratios within a year where the
-within year variance is study-specific, $CV_{\sigma_{\eta}}$ is
-the coefficient of variation ($1/\sqrt{shape}$) of the gamma
-distribution (this model is most supported by the data); and d)
-between year ($\sigma_{\beta}$). Including a between-year variance
-term was not supported by the data (the posterior for
-($\sigma_{\beta}$ is not differentiated from zero ).}
-\label{fig:posteriors}
-\end{figure}
-\clearpage
-
-\begin{figure}
-\centering
-\includegraphics[width=0.8\textwidth]{figure/model_selection.pdf}
-\caption{The sensitivity of the yield gap to including different
-levels of hierarchy in the model. The random effects included in
-the model are: a) study ($\sigma_{\alpha}$); b) study and response
-ratios within a year ($\sigma_{\eta}$); c) study and response
-ratios within a year where the within year variance is
-study-specific ($\sigma_{\eta}[i]$) (this model is most supported
-by the data); and d) study, study-specific within-year variance,
-and between year ($\sigma_{\beta}$). Including a between-year
-variance term was not supported by the data (the posterior for
-($\sigma_{\beta}$ is not differentiated from zero ). Values are
-the posterior mean with 95\% credible intervals.}
-\label{fig:model_selection}
-\end{figure}
-\clearpage
-
-\begin{table}
-  \renewcommand*\arraystretch{1.25}
-  \centering
-  \caption{Parameter posteriors for models without explanatory
-    variables. $\mu$ is the true mean response ratio across years and
-    studies, $\sigma_{\alpha}$ is the standard deviation of the
-    distribution from which the study random effects are drawn;
-    $\sigma_{\eta}$ is the standard deviation of the distribution from
-    which the within year random effects are drawn; $CV_{\sigma_\eta}$ is the
-    coefficient of variation of the gamma from which the
-    study-specific within-year variance are drawn; and $\sigma_\beta$
-    is the standard deviation of the distribution of random between year
-    effects. Values of Rhat $< 1.1$ indicate convergence. Lower
-    Deviance Information Criterion (DIC) indicates better model fit to
-    the data.}
-
-  \vspace{20pt}
-  \label{table:params}
-  \begin{tabular}{|c|c|c|c|c|} \hline
-    Parameter & Posterior mean & Posterior standard deviation
-    & 95\% CI & Rhat \\ \hline
-    \multicolumn{5}{|l|}{Study random effect, DIC=1684.8} \\ \hline
-    \multirow{1}{*}{$\mu$} &
-    $0.795$ & $0.027$ & $0.742 - 0.848$ & $1.001$ \\ \hline
-    \multirow{1}{*}{$\sigma_{\alpha}$} &
-    $0.341$ & $0.026$ & $0.294 - 0.396$ & $1.001$ \\ \hline
-
-    \multicolumn{5}{|l|}{Study and within year random effects,
-      DIC= -565.9} \\ \hline
-    \multirow{1}{*}{$\mu$} &
-    $0.788$ & $0.021$ & $0.749 - 0.829$ & $1.001$ \\ \hline
-    \multirow{1}{*}{$\sigma_{\alpha}$} &
-    $0.188$ & $0.024$ & $0.144 - 0.239$ & $1.001$ \\ \hline
-    \multirow{1}{*}{$\sigma_{\eta}$} &
-    $0.312$ & $0.011$ & $0.291 - 0.333$ & $1.001$ \\ \hline
-
-    \multicolumn{5}{|l|}{Study and study-specific within year random
-      effects, DIC= -618.0} \\ \hline
-    \multirow{1}{*}{$\mu$} &
-    $0.808$ & $0.019$ & $0.771 - 0.845$ & $1.001$ \\ \hline
-    \multirow{1}{*}{$\sigma_{\alpha}$} &
-    $0.189$ & $0.023$ & $0.145 - 0.237$ & $1.001$ \\ \hline
-   \multirow{1}{*}{$CV_{\sigma_\eta}$} &
-    $1.155$ & $0.135$ & $0.907 - 1.436$ & $1.001$ \\ \hline
-
-    \multicolumn{5}{|l|}{Study, study-specific within year,
-      and between year random effects, DIC= -621.2} \\ \hline
-     \multirow{1}{*}{$\mu$} &
-    $0.808$ & $0.019$ & $0.770 - 0.846$ & $1.001$ \\ \hline
-    \multirow{1}{*}{$\sigma_{\alpha}$} &
-    $0.186$ & $0.024$ & $0.142 - 0.234$ & $1.001$ \\ \hline
-   \multirow{1}{*}{$CV_{\sigma_\eta}$} &
-    $1.157$ & $0.136$ & $0.907 - 1.440$ & $1.001$ \\ \hline
-   \multirow{1}{*}{$\sigma_{\beta}$} &
-    $0.041$ & $0.027$ & $0.002 - 0.098$ & $1.001$ \\ \hline
-
-  \end{tabular}
-\end{table}
-\clearpage
-
-The best supported model, coded in JAGS, is
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-## model {
-##      for(study in 1:Nstudy) {
-##        for(year in 1:Nyear[study]) {
-##          for(obs in 1:Nobs[study,year]) {
-##            P[study,year,obs] <- 1/V[study,year,obs]
-##            RR[study,year,obs] ~ dnorm(mu.RR[study,year,obs],
-##                                       P[study,year,obs])
-##            mu.RR[study,year,obs] ~ dnorm(mu.study[study],
-##                                      tau.yr.obs[study])
-##          }
-##        }
-##        mu.study[study] ~ dnorm(mu, tau)
-##        tau.yr.obs[study] ~ dgamma(shape.obs, scale.obs)
-##      }
-## 
-##      mu ~ dnorm(0, 1e-4)
-##      exp.mu <- exp(mu)
-##      tau <- 1 / (sigma * sigma)
-##      sigma ~ dunif(0, 100)
-## 
-##      shape.obs <- (1/cv.obs)^2
-##      cv.obs ~ dunif(0, 100)
-##      scale.obs <- (1/in.scale.obs)^2
-##      in.scale.obs ~ dunif(0, 100)
-##    }
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-
-Where the parameters names match the notation in
-Equ. \ref{eq:fullModel}.
-
-\section{Analysis}
-\label{sec:analysis}
-
-After deciding on the model parameterization, the most difficult part
-of running a model in JAGS in getting tha data in the right format (in
-my opinion). Our situation was particularly difficult because we
-needed to build in flexibility to pool different combinations of
-response ratios using the method presented in Lajeunesse
-\cite{Lajeunesse2011} depending on the explanatory variable being
-considered (when response ratios that share a control split between
-different levels of an explanatory variable, they can be left
-un-pooled). Below I focus on the best supported model without
-explanatory variables.
-
-First we have a function to calculate pooled RR and their variances
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlcom{## applies the Lajeunesse (2011) method for combining RR with shared}
-\hlcom{## controls. Returns a pooled RR and variance}
-\hlstd{poolCommonControl} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{dd}\hlstd{) \{}
-
-  \hlcom{## function to make covariance matrix:}
-  \hlstd{covMat} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{x}\hlstd{)\{}
-    \hlcom{## case 1: organic is repeated}
-    \hlkwa{if}\hlstd{(}\hlkwd{all}\hlstd{(}\hlkwd{diff}\hlstd{(}\hlkwd{as.numeric}\hlstd{(x[,}\hlstr{"mean.org"}\hlstd{]))}\hlopt{==}\hlnum{0}\hlstd{))}
-      \hlstd{vals} \hlkwb{<-} \hlkwd{as.numeric}\hlstd{(x[}\hlnum{1}\hlstd{,} \hlkwd{c}\hlstd{(}\hlstr{"sd.org"}\hlstd{,} \hlstr{"n.organic"}\hlstd{,} \hlstr{"mean.org"}\hlstd{)])}
-    \hlkwa{if}\hlstd{(}\hlkwd{all}\hlstd{(}\hlkwd{diff}\hlstd{(}\hlkwd{as.numeric}\hlstd{(x[,}\hlstr{"mean.conv"}\hlstd{]))}\hlopt{==}\hlnum{0}\hlstd{))}
-      \hlstd{vals} \hlkwb{<-} \hlkwd{as.numeric}\hlstd{(x[}\hlnum{1}\hlstd{,} \hlkwd{c}\hlstd{(}\hlstr{"sd.conv"}\hlstd{,} \hlstr{"n.conv"}\hlstd{,} \hlstr{"mean.conv"}\hlstd{)])}
-
-    \hlstd{m} \hlkwb{<-} \hlkwd{matrix}\hlstd{(vals[}\hlnum{1}\hlstd{]}\hlopt{^}\hlnum{2}\hlopt{/}\hlstd{(vals[}\hlnum{2}\hlstd{]}\hlopt{*}\hlstd{vals[}\hlnum{3}\hlstd{]}\hlopt{^}\hlnum{2}\hlstd{),}
-                \hlkwc{ncol}\hlstd{=}\hlkwd{nrow}\hlstd{(dd),} \hlkwc{nrow}\hlstd{=}\hlkwd{nrow}\hlstd{(dd))}
-    \hlkwd{diag}\hlstd{(m)} \hlkwb{<-} \hlkwd{as.numeric}\hlstd{(x[,}\hlstr{"varlnR"}\hlstd{])}
-    \hlstd{m}
-  \hlstd{\}}
-
-  \hlcom{## covariance matrix and its inverse:}
-  \hlstd{V} \hlkwb{<-} \hlkwd{covMat}\hlstd{(dd)}
-  \hlstd{V.inv} \hlkwb{<-} \hlkwd{solve}\hlstd{(V)}
-
-  \hlstd{X} \hlkwb{<-} \hlkwd{matrix}\hlstd{(}\hlnum{1}\hlstd{,} \hlkwd{nrow}\hlstd{(V))}
-  \hlstd{E} \hlkwb{<-} \hlkwd{matrix}\hlstd{(dd[,}\hlstr{"lnR"}\hlstd{],} \hlkwd{nrow}\hlstd{(V))}
-  \hlcom{## compute and return RR_bar and variance(RR_bar), as defined in}
-  \hlcom{## Eq. 3 and subsequent text in Lajeunesse (2011):}
-  \hlkwd{return}\hlstd{(}\hlkwd{c}\hlstd{(}\hlkwc{obs}\hlstd{=dd}\hlopt{$}\hlstd{obs[}\hlnum{1}\hlstd{],}
-           \hlkwc{lnR}\hlstd{=}\hlkwd{as.numeric}\hlstd{(}\hlkwd{solve}\hlstd{(}\hlkwd{t}\hlstd{(X)} \hlopt{%*%} \hlstd{V.inv} \hlopt{%*%} \hlstd{X)} \hlopt{%*%}
-                          \hlstd{(}\hlkwd{t}\hlstd{(X)} \hlopt{%*%} \hlstd{V.inv} \hlopt{%*%} \hlstd{E)),}
-           \hlkwc{varlnR}\hlstd{=}\hlkwd{as.numeric}\hlstd{(}\hlkwd{solve}\hlstd{(}\hlkwd{t}\hlstd{(X)} \hlopt{%*%} \hlstd{V.inv} \hlopt{%*%} \hlstd{X))))}
-\hlstd{\}}
+ RScript mainR.R
 \end{alltt}
@@
 
-Which is called by this function which combines the meta-data based on
-keys
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlcom{## takes raw meta-data and a vector of covariates and combines data}
-\hlcom{## using lajeunesse (2011). Returns a dataframe with keys, the}
-\hlcom{## observation number, the pooled RR and its variance}
-
-\hlstd{poolRR} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{D}\hlstd{,} \hlkwc{covariates}\hlstd{) \{}
-  \hlcom{## first split the data up into relevant groups:}
-  \hlkwa{if}\hlstd{(}\hlkwd{length}\hlstd{(covariates)} \hlopt{==} \hlnum{1}\hlstd{) \{}
-    \hlstd{dd} \hlkwb{<-} \hlkwd{split}\hlstd{(D,} \hlkwd{paste}\hlstd{(D}\hlopt{$}\hlstd{study,}
-                         \hlstd{D[,covariates],}
-                         \hlstd{D}\hlopt{$}\hlstd{year_conv,}
-                         \hlstd{D}\hlopt{$}\hlstd{mult_yrs,}
-                         \hlstd{D}\hlopt{$}\hlstd{duplicate_trt_dat,}
-                         \hlkwc{sep}\hlstd{=}\hlstr{';'}\hlstd{))}
-  \hlstd{\}}
-  \hlkwa{if}\hlstd{(}\hlkwd{length}\hlstd{(covariates)} \hlopt{>} \hlnum{1}\hlstd{) \{}
-    \hlstd{dd} \hlkwb{<-} \hlkwd{split}\hlstd{(D,} \hlkwd{paste}\hlstd{(D}\hlopt{$}\hlstd{study,}
-                         \hlkwd{apply}\hlstd{(D[,covariates],} \hlnum{1}\hlstd{, paste,} \hlkwc{collapse}\hlstd{=}\hlstr{':'}\hlstd{),}
-                         \hlstd{D}\hlopt{$}\hlstd{year_conv,}
-                         \hlstd{D}\hlopt{$}\hlstd{mult_yrs,}
-                         \hlstd{D}\hlopt{$}\hlstd{duplicate_trt_dat,}
-                         \hlkwc{sep}\hlstd{=}\hlstr{';'}\hlstd{))}
-  \hlstd{\}}
-
-  \hlcom{## combine response ratios for studies that share treatments using}
-  \hlcom{## the method given in Lajeunesse (2011):}
-  \hlstd{share.trts} \hlkwb{<-} \hlkwd{which}\hlstd{(}\hlkwd{sapply}\hlstd{(}\hlkwd{strsplit}\hlstd{(}\hlkwd{names}\hlstd{(dd),} \hlkwc{split}\hlstd{=}\hlstr{';'}\hlstd{),}
-                             \hlkwa{function}\hlstd{(}\hlkwc{x}\hlstd{)} \hlkwd{rev}\hlstd{(x)[[}\hlnum{1}\hlstd{]])} \hlopt{==} \hlstr{'yes'}\hlstd{)}
-
-  \hlstd{dd[share.trts]} \hlkwb{<-} \hlkwd{lapply}\hlstd{(dd[share.trts],}
-                           \hlstd{poolCommonControl)}
-
-  \hlstd{dd[}\hlopt{-}\hlstd{share.trts]} \hlkwb{<-}
-    \hlkwd{lapply}\hlstd{(dd[}\hlopt{-}\hlstd{share.trts],} \hlkwa{function}\hlstd{(}\hlkwc{x}\hlstd{) x[,}\hlkwd{c}\hlstd{(}\hlstr{'obs'}\hlstd{,} \hlstr{'lnR'}\hlstd{,} \hlstr{'varlnR'}\hlstd{)])}
-  \hlkwd{return}\hlstd{(}\hlkwd{do.call}\hlstd{(rbind, dd))}
-\hlstd{\}}
-\end{alltt}
-\end{kframe}
-\end{knitrout}
-
-This is all brought together by this long, unwieldy function that
-returns the data structure necessary for the JAGS model including the
-indexes for looping and arrays of RR, variances and covariates is
-applicable
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlcom{## makes data arrays for Bayesian analysis; takes the raw}
-\hlcom{## meta-data set, whether to apply the lajeunesse method to combine RR}
-\hlcom{## with shared controls, and a vector of explanatory variables. The}
-\hlcom{## default for covariates is "Crop.species" so RR are never combined}
-\hlcom{## between crop species. Returns a list of the number of studies}
-\hlcom{## (Nstudy), the number of years for each study (Nyear), a matrix}
-\hlcom{## (study, year) of the number of observations in each year (Nobs),}
-\hlcom{## and an array (study, year, observation) of the RR and their}
-\hlcom{## variances, and an array of covariates if applicable}
-
-\hlstd{makeData} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{meta.dat}\hlstd{,}
-                     \hlkwc{lajeunesse} \hlstd{=} \hlnum{TRUE}\hlstd{,}
-                     \hlkwc{covariates}\hlstd{=}\hlkwd{c}\hlstd{(}\hlstr{'Crop.species'}\hlstd{)) \{}
-
-  \hlstd{convert2Int} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{x}\hlstd{)} \hlkwd{seq_along}\hlstd{(x)[}\hlkwd{match}\hlstd{(x,} \hlkwd{unique}\hlstd{(x))]}
-
-  \hlkwa{if}\hlstd{(lajeunesse) \{}
-    \hlstd{dd} \hlkwb{<-} \hlkwd{poolRR}\hlstd{(meta.dat,}
-                 \hlkwc{covariates}\hlstd{=covariates)}
-    \hlstd{keys.split} \hlkwb{<-} \hlkwd{sapply}\hlstd{(}\hlkwd{rownames}\hlstd{(dd), strsplit,} \hlkwc{split}\hlstd{=}\hlstr{';'}\hlstd{)}
-    \hlstd{get} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{i}\hlstd{)} \hlkwd{as.vector}\hlstd{(}\hlkwd{sapply}\hlstd{(keys.split,} \hlkwa{function}\hlstd{(}\hlkwc{x}\hlstd{) x[i]))}
-    \hlstd{study} \hlkwb{<-} \hlkwd{get}\hlstd{(}\hlnum{1}\hlstd{)}
-    \hlstd{year} \hlkwb{<-} \hlkwd{get}\hlstd{(}\hlnum{3}\hlstd{)}
-  \hlstd{\}} \hlkwa{else} \hlstd{\{}
-    \hlstd{dd} \hlkwb{<-} \hlstd{meta.dat}
-    \hlstd{study} \hlkwb{<-} \hlkwd{as.character}\hlstd{(meta.dat}\hlopt{$}\hlstd{study)}
-    \hlstd{year} \hlkwb{<-} \hlstd{meta.dat}\hlopt{$}\hlstd{year_conv}
-  \hlstd{\}}
-
-  \hlcom{## make study index}
-  \hlstd{ind.study} \hlkwb{<-} \hlkwd{convert2Int}\hlstd{(study)}
-
-  \hlcom{## make year index}
-  \hlstd{ind.year} \hlkwb{<-} \hlkwd{rep}\hlstd{(}\hlnum{NA}\hlstd{,} \hlkwd{nrow}\hlstd{(dd))}
-  \hlkwa{for}\hlstd{(i} \hlkwa{in} \hlkwd{unique}\hlstd{(ind.study))}
-    \hlstd{ind.year[ind.study}\hlopt{==}\hlstd{i]} \hlkwb{<-} \hlkwd{convert2Int}\hlstd{(year[ind.study}\hlopt{==}\hlstd{i])}
-
-  \hlcom{## make obs index}
-  \hlstd{ind.obs} \hlkwb{<-} \hlkwd{rep}\hlstd{(}\hlnum{NA}\hlstd{,} \hlkwd{nrow}\hlstd{(dd))}
-  \hlkwa{for}\hlstd{(i} \hlkwa{in} \hlkwd{unique}\hlstd{(ind.study))}
-    \hlkwa{for}\hlstd{(j} \hlkwa{in} \hlkwd{unique}\hlstd{(ind.year[ind.study}\hlopt{==}\hlstd{i]))}
-      \hlstd{ind.obs[ind.year}\hlopt{==}\hlstd{j} \hlopt{&} \hlstd{ind.study}\hlopt{==}\hlstd{i]} \hlkwb{<-}
-        \hlkwd{seq_len}\hlstd{(}\hlkwd{sum}\hlstd{(ind.year}\hlopt{==}\hlstd{j} \hlopt{&} \hlstd{ind.study}\hlopt{==}\hlstd{i))}
-
-  \hlcom{## make response ratio and vi matrices}
-  \hlstd{makeMat} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{d.vec}\hlstd{) \{}
-    \hlstd{mat} \hlkwb{<-} \hlkwd{array}\hlstd{(}\hlnum{NA}\hlstd{,} \hlkwc{dim}\hlstd{=}\hlkwd{c}\hlstd{(}\hlkwd{max}\hlstd{(ind.study),}
-                           \hlkwd{max}\hlstd{(ind.year),}
-                           \hlkwd{max}\hlstd{(ind.obs)))}
-    \hlstd{mat[}\hlkwd{cbind}\hlstd{(ind.study, ind.year, ind.obs)]} \hlkwb{<-} \hlstd{d.vec}
-    \hlkwd{return}\hlstd{(mat)}
-  \hlstd{\}}
-  \hlstd{RR.mat} \hlkwb{<-} \hlkwd{makeMat}\hlstd{(dd}\hlopt{$}\hlstd{lnR)}
-  \hlstd{vi.mat} \hlkwb{<-} \hlkwd{makeMat}\hlstd{(dd}\hlopt{$}\hlstd{varlnR)}
-
-  \hlcom{## make index bounds}
-  \hlstd{nstudy} \hlkwb{<-} \hlkwd{max}\hlstd{(ind.study)}
-  \hlstd{nyear}  \hlkwb{<-} \hlkwd{apply}\hlstd{(RR.mat[,,}\hlnum{1}\hlstd{],}  \hlnum{1}\hlstd{,} \hlkwa{function}\hlstd{(}\hlkwc{x}\hlstd{)} \hlkwd{sum}\hlstd{(}\hlopt{!}\hlkwd{is.na}\hlstd{(x)))}
-  \hlstd{nobs}   \hlkwb{<-} \hlkwd{apply}\hlstd{(RR.mat[,,],} \hlnum{1}\hlopt{:}\hlnum{2}\hlstd{,} \hlkwa{function}\hlstd{(}\hlkwc{x}\hlstd{)} \hlkwd{sum}\hlstd{(}\hlopt{!}\hlkwd{is.na}\hlstd{(x)))}
-
-  \hlstd{covs} \hlkwb{<-} \hlnum{NA}
-  \hlstd{num.cats} \hlkwb{<-} \hlnum{NA}
-  \hlcom{## covariates (only for lajeunesse)}
-  \hlkwa{if}\hlstd{(lajeunesse) \{}
-    \hlstd{num.covs} \hlkwb{<-} \hlkwd{length}\hlstd{(covariates)}
-    \hlkwa{if}\hlstd{(num.covs} \hlopt{>} \hlnum{1}\hlstd{) \{}
-      \hlstd{covs} \hlkwb{<-} \hlkwd{sapply}\hlstd{(}\hlkwd{get}\hlstd{(}\hlnum{2}\hlstd{), strsplit,} \hlkwc{split}\hlstd{=}\hlstr{':'}\hlstd{)}
-      \hlstd{getCov} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{cov}\hlstd{)}
-        \hlkwd{as.numeric}\hlstd{(}\hlkwd{as.factor}\hlstd{(}\hlkwd{sapply}\hlstd{(covs,} \hlkwa{function}\hlstd{(}\hlkwc{x}\hlstd{) x[cov])))}
-      \hlstd{covs} \hlkwb{<-} \hlkwd{lapply}\hlstd{(}\hlnum{1}\hlopt{:}\hlstd{num.covs,} \hlkwa{function}\hlstd{(}\hlkwc{cov}\hlstd{)}
-        \hlkwd{make.mat}\hlstd{(}\hlkwd{getCov}\hlstd{(cov)))}
-      \hlkwd{names}\hlstd{(covs)} \hlkwb{<-} \hlstd{covariates}
-      \hlstd{num.cats} \hlkwb{<-} \hlkwd{sapply}\hlstd{(covs, max,} \hlkwc{na.rm}\hlstd{=}\hlnum{TRUE}\hlstd{)}
-    \hlstd{\}}
-  \hlstd{\}}
-
-  \hlkwd{return}\hlstd{(}\hlkwd{list}\hlstd{(}\hlkwc{Nstudy}\hlstd{=nstudy,}
-              \hlkwc{Nyear}\hlstd{=nyear,}
-              \hlkwc{Nobs}\hlstd{=nobs,}
-              \hlkwc{V}\hlstd{=vi.mat,}
-              \hlkwc{RR}\hlstd{=RR.mat,}
-              \hlkwc{covariates}\hlstd{=covs,}
-              \hlkwc{num.cats}\hlstd{=num.cats))}
-\hlstd{\}}
-\end{alltt}
-\end{kframe}
-\end{knitrout}
-
-We then pacakges everything up in a function that also passes the JAGS
-parameters (init values, thinning rate etc.) and runs the model.
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlcom{## takes preped data and a scale parameter and packages the data and}
-\hlcom{## parameters for JAGS, then runs the analysis. Returns the JAGS}
-\hlcom{## summary output}
-\hlstd{runAnalysis} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{dd}\hlstd{,} \hlkwc{scale}\hlstd{) \{}
-
-  \hlstd{my.inits} \hlkwb{<-} \hlkwa{function}\hlstd{() \{}
-    \hlkwd{list}\hlstd{()}
-  \hlstd{\}}
-  \hlstd{my.params} \hlkwb{<-} \hlkwd{get.params}\hlstd{()}
-
-  \hlstd{dd} \hlkwb{<-} \hlkwd{list}\hlstd{(}\hlkwc{data}\hlstd{=dd,} \hlkwc{inits}\hlstd{=my.inits,} \hlkwc{params}\hlstd{=}\hlkwd{get.params}\hlstd{())}
-
-  \hlstd{bugs} \hlkwb{<-} \hlkwd{run.model}\hlstd{(dd,}
-                    \hlkwc{n.thin}\hlstd{=scale,}
-                    \hlkwc{n.iter}\hlstd{=(}\hlnum{1e4}\hlstd{)}\hlopt{*}\hlstd{scale,}
-                    \hlkwc{n.burnin}\hlstd{=}\hlnum{1e1}\hlopt{*}\hlstd{scale,}
-                    \hlkwc{n.chains}\hlstd{=}\hlnum{3}\hlstd{)}
-
-  \hlkwd{return}\hlstd{(}\hlkwd{list}\hlstd{(}\hlkwc{data}\hlstd{=dd,}
-       \hlkwc{bugs}\hlstd{=bugs,}
-       \hlkwc{summary}\hlstd{=bugs}\hlopt{$}\hlstd{BUGSoutput}\hlopt{$}\hlstd{summary))}
-\hlstd{\}}
-\end{alltt}
-\end{kframe}
-\end{knitrout}
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{print}\hlstd{(out}\hlopt{$}\hlstd{bugs,} \hlkwc{dig}\hlstd{=}\hlnum{3}\hlstd{)}
-\end{alltt}
-\begin{verbatim}
-## Inference for Bugs model at "model.jags", fit using jags,
-##  3 chains, each with 1e+05 iterations (first 100 discarded), n.thin = 10
-##  n.sims = 29970 iterations saved
-##            mu.vect sd.vect      2.5%       25%       50%       75%
-## cv.obs       1.154   0.134     0.909     1.061     1.149     1.241
-## exp.mu       0.808   0.019     0.771     0.795     0.808     0.821
-## sigma        0.189   0.023     0.145     0.172     0.188     0.204
-## deviance -1293.537  36.474 -1362.740 -1318.528 -1294.309 -1269.421
-##              97.5%  Rhat n.eff
-## cv.obs       1.434 1.001  5100
-## exp.mu       0.846 1.001 27000
-## sigma        0.236 1.001  9400
-## deviance -1219.532 1.001 20000
-## 
-## For each parameter, n.eff is a crude measure of effective sample size,
-## and Rhat is the potential scale reduction factor (at convergence, Rhat=1).
-## 
-## DIC info (using the rule, pD = var(deviance)/2)
-## pD = 665.1 and DIC = -628.4
-## DIC is an estimate of expected predictive error (lower deviance is better).
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-
-exp.mu is the mean ratio of organic and conventional yields, which is
-around $80.1\%$ with $95\%$ credible intervals ranging from around
-$77--85\%$
-
-Sigma is the between study variance, and cv.obs is the coefficient of
-variation of $\gamma$ distribution the within-study variance
-parameters are drawn from.
-
-Values of Rhat < 1.1 indicate convergence, whcih we We can see from
-inspecting the posteriors
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{ggs_density}\hlstd{(inspect.out)}
-\end{alltt}
-\end{kframe}
-\includegraphics[width=\maxwidth]{figure/densityPlots-1}
-
-\end{knitrout}
-
-The chains look wonderfully grassy
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{ggs_traceplot}\hlstd{(inspect.out)}
-\end{alltt}
-\end{kframe}
-\includegraphics[width=\maxwidth]{figure/traceplots-1}
-
-\end{knitrout}
-
-\section{Explanatory variables}
-\label{sec:variables}
-We also extended this model in order to accommodate analyses of study
-characteristics such as crop type and management practices. We analyze
-these additional explanatory variables one at a time because not all
-studies reported all explanatory variables. In these analyses, for
-cases where multiple organic treatments represented different
-categories for a specific explanatory variable, they could not be
-combined using Lajeunesse's method \citep{Lajeunesse2011}. The
-potential bias resulting from non-independence of the response ratios
-in these cases, however, would be minimized by the fact that they are
-not pooled together in the analysis \citep{Lajeunesse2011}.
-
-Letting $h$ index the categories for a particular explanatory variable
-(e.g., crop species), we then have:
-%
-\begin{equation}
-  \begin{split}
-    \label{eq:fullModel_cov}
-    y_{hijk} &=
-    \mu + \gamma_h + \alpha_i + \eta_{ijk} +  \beta_{ij} + \epsilon_{ijk}\\
-  \end{split}
-\end{equation}
-%
-
-where $\gamma_h$ is the effect of the $h^{\mathrm{th}}$ category, and
-the rest of the model parallels that given in Eq.~\ref{eq:fullModel}.
-
-The model including explanatory variables in JAGS is
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-## model {
-##     for(study in 1:Nstudy) {
-##       for(year in 1:Nyear[study]) {
-##         for(obs in 1:Nobs[study,year]) {
-##           P[study,year,obs] <- 1/V[study,year,obs]
-##           RR[study,year,obs] ~ dnorm(mu.RR[study,year,obs], P[study,year,obs])
-##           mu.RR[study,year,obs] ~ dnorm(mu.study[study,year,obs],
-##                                         tau.obs[study])
-##           mu.study[study,year,obs] ~ dnorm(mu[cov[study,year,obs]], tau)
-##         }
-##       }
-##       tau.obs[study] ~ dgamma(shape.obs, scale.obs)
-##     }
-## 
-##     ## set priors on covariates
-##     for(covs in 1:Ncov) {
-##       mu[covs] ~ dnorm(0, 1E-3)
-##       exp.mu[covs] <- exp(mu[covs])
-##     }
-##     tau <- 1 / (sigma * sigma)
-##     sigma ~ dunif(0, 100)
-## 
-##     shape.obs <- (1/cv)^2
-##     cv ~ dunif(0, 100)
-##     scale.obs <- (1/in.scale)^2
-##     in.scale ~ dunif(0, 100)
-## 
-##   }
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-
-Where the parameters names match the notation in
-Equ.~\ref{eq:fullModel} and cov is a matrix of explanatory variables
-assiciated with each response ratio.
-
-To re-run the models with explanatory variables, we just added each
-variable one-by-one to the covariates argument of the makeData()
-function.
-
-\bibliography{org_v_conv} \clearpage
-
 \end{document}

### 2017-04-21 analysis/changePoint/convertConsensus.py

@@ -1,7 +1,8 @@
 #!/usr/bin/env python
+import os
 import sys
 import subprocess as sp
-import os
+import convertConsensus as cc
 
 paradigms_file = sys.argv[1]
 fh = open(paradigms_file, 'r')
@@ -10,9 +11,9 @@ paradigms = [l.strip() for l in fh]
 
 os.chdir('cptPeel')
 
-
 for paradigm in paradigms:
-    command = "python runConsensus.py graph-names.lut {} {} -p 'baci/'"
+    command = "python cc.writeToFile {} {} 'gml'"
     command = command.format(len(paradigm.split()), paradigm)
     job = sp.Popen(command, shell=True)
     job.wait()
+

### 2017-04-21 analysis/changePoint/convertfiles.py

@@ -1,28 +1,0 @@
-
-import convertConsensus as cc
-## three change points
-cc.writeToFile('baci/Barger_2008_joint_s1-consensus.tree','gml')
-cc.writeToFile('baci/Barger_2009_joint_s1-consensus.tree','gml')
-cc.writeToFile('baci/Barger_2012_joint_s1-consensus.tree','gml')
-cc.writeToFile('baci/Barger_2014_joint_s1-consensus.tree','gml')
-
-## 1 change point
-cc.writeToFile('baci/Butler_2009_joint_s1-consensus.tree','gml')
-cc.writeToFile('baci/Butler_2015_joint_s1-consensus.tree','gml')
-
-## three change points
-cc.writeToFile('baci/Hrdy_2007_joint_s1-consensus.tree','gml')
-cc.writeToFile('baci/Hrdy_2009_joint_s1-consensus.tree','gml')
-cc.writeToFile('baci/Hrdy_2011_joint_s1-consensus.tree','gml')
-cc.writeToFile('baci/Hrdy_2014_joint_s1-consensus.tree','gml')
-
-## three change points
-cc.writeToFile('baci/MullerB_2006_joint_s1-consensus.tree','gml')
-cc.writeToFile('baci/MullerB_2009_joint_s1-consensus.tree','gml')
-cc.writeToFile('baci/MullerB_2012_joint_s1-consensus.tree','gml')
-cc.writeToFile('baci/MullerB_2015_joint_s1-consensus.tree','gml')
-
-## two change points
-cc.writeToFile('baci/Sperandio_2009_joint_s1-consensus.tree','gml')
-cc.writeToFile('baci/Sperandio_2014_joint_s1-consensus.tree','gml')
-

### 2017-04-21 analysis/changePoint/mainChangePoint.sh

@@ -1,7 +1,12 @@
 #!/usr/bin/env bash
 RScript dataPrep.R
+## remove old resuts because the change point analysis appends them
+## otherwise
+rm cptPeel/LogLs_4.txt cptPeel/results_4.txt
+## runs in parallel on 2 cores
 python hedgerows.py
 RScript prepChangePointOutput.R
 python postChangePoint.py saved/consensus.txt
-python convertfiles.py
+python convertConsensus.py saved/lastyr_consensus.txt
+# python convertfiles.py
 RScript plotting/networks.R

### 2017-05-18 analysis/writeup/writeup.tex
lots of small changes to script to integrate with wrieup

@@
 
+\let\hlesc\hlstd \let\hlpps\hlstd \let\hllin\hlstd \let\hlslc\hlcom \let\hlppc\hlcom
 \usepackage{alltt}
@@
 
+\bibliographystyle{ecology_let}
+
 %% for inline R code: if the inline code is not correctly parsed, you will see a message
@@
 \begin{document}
-\title{Opportunistic attachment assembles plant-pollinator networks: A walkthrough of the anlaysis}
+\title{Opportunistic attachment assembles plant-pollinator networks: A walk through of the analysis}
 \author{Lauren Ponisio}
@@
 \maketitle
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=1.1\textwidth]{figure/SankeyIDb303b2e99ce.pdf}
+  \label{fig:methods}
+\end{figure}
+\clearpage
+
 \section{Overview}
@@
 
-In our study we examing the temporal dynamcis of plant-pollinator
-network assembly using vareity of different methods include 1) network
+In our study we examining the temporal dynamics of plant-pollinator
+network assembly using variety of different methods include 1) network
 change point detection 2) node/species-level position variation, 3)
 species and interaction turnover 4) network-level metrics and 5)
-extinction simulations. We are commited to reproducible science and
-all analytical code will be maintained on github, along wiht this
-write up (as included as a supplment to the publication).
+extinction simulations. We are committed to reproducible science and
+all analytical code will be maintained on github, along with this
+write up.
 
-The entire analysis is executable from the main.sh file
+The entire analysis is executable from the main.sh file. All of the
+packages needed to run the analyses are listed in the packages.sh
+file, including the specific versions of the python dendropy library,
+which is very finicky. We have had no problems with back compatibility
+for the R scripts (in terms of R versions from v.~2-3.3 or packages)
+but the change point analysis must be run in python2 (not easily
+convertible to python3, though this is potentially in progress).
+
+Navigate to the analysis folder within the github repo
+(hedgerow\_assembly) then the main.sh file can be selected and
+run. I would not recommend doing this unless you are prepared to have two
+cores of your machine running for days to weeks (depending on how many
+of iterations of the change point analysis specific in
+mainChangePoint.sh loop), but you could run of the analyses in the
+study by running this line in BASH.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\noindent
+\ttfamily
+\hlstd{\ bash\ main.sh}\hspace*{\fill}
+\mbox{}
+\normalfont
+\end{kframe}
+\end{knitrout}
+
+This will somewhat helpfully print the results of each analysis
+and re-create any accompanying figures.
+
+We will walk through each analysis individually.
+
+\section{Change Point Analysis}
 
 The change point analysis which is a mix of python and R. hedgerows.py
-(the meat of the change point analysis) runs in paralell on two cores
-(which can be modified in the script if needed), but will likely take
-a few hours, depending on your machine.
+(the meat of the change point analysis) runs in parallel on two cores
+(which can be modified in the hedgerows.py script by toggling the
+ncores object), but will likely take a many hours (a day on running on
+my 2.5 GHz Intel Core i7 mac pro) depending on your machine.
+
+There is some stochasticity in the change point analysis in the
+hypothesis testing step where samples are drawn from the null
+distribution and the likelihood of a change point occurring is
+calculated from those samples. Because of this, we re-ran the change
+point detection analysis 1000 times (this was done on a cluster) and
+used the change points that were identified in 95\% of runs in
+subsequent analyses.
+
+mainChangePoint.sh compiles the python code, then creates the .pairs
+files needed for the change point analysis (dataPrep.R). Nodes names
+need to be consistent across years within a site, and for simplicity
+we made them consistent across all sites as well. This enabled us to
+provide a single .lut master list of the nodes to the analysis.
+
+Next, the change point detection loop is run. This function begins by
+fitting the generalized hierarchical random graph model (GHRG) which
+outputs consensus trees that are used by the change point detection
+algorithm. There is only slight variability in the fitting of the
+GHRG between runs, so once the first loop is run, the consensus trees
+are used in subsequent runs.
+
+The output of the change point analysis is saved as a list of the
+change points identified and like likelihood scores (formatted by
+prepChangePointOutput.R).
+
+After the loop, consensusChangePoints.R drops the change points that
+were not identified by $95\%$ of the runs and then creates consensus
+trees for each chunk of years between change point in python
+(postChangePoint.py). The outputs of the change point analysis were
+very specialized to that analysis, so we had to convert them to .glm
+files for plotting and further exploration in R
+(convertConsensusTrees.py)
+
+The plotting and statistical analyses are in R in different
+scripts. The plotting function will create the network, community and
+flow plots (networks.R). The binomial linear mixed models are run in
+the comparison.R script, along with addition statistical analysis
+requested by the reviewers testing whether change points tended to
+occur between years with a larger difference in samples.
 
@@
 \begin{alltt}
- RScript changePoint/dataPrep.R
- python changePoint/hedgerows.py
- RScript changePoint/prepChangePointOutput.R
- python changePoint/postChangePoint.py saved/consensus.txt
- python changePoint/convertfiles.py
- RScript changePoint/plotting/networks.R
+\hlstd{mod.chpt} \hlkwb{<-} \hlkwd{glm}\hlstd{(}\hlkwd{cbind}\hlstd{(chpt.trial}\hlopt{$}\hlstd{chpts,}
+                      \hlstd{chpt.trial}\hlopt{$}\hlstd{trial} \hlopt{-} \hlstd{chpt.trial}\hlopt{$}\hlstd{chpts)} \hlopt{~}
+    \hlstd{chpt.trial}\hlopt{$}\hlstd{status,} \hlkwc{family}\hlstd{=}\hlstr{"binomial"}\hlstd{)}
+\hlkwd{print}\hlstd{(}\hlkwd{summary}\hlstd{(mod.chpt))}
+\end{alltt}
+\begin{verbatim}
+## 
+## Call:
+## glm(formula = cbind(chpt.trial$chpts, chpt.trial$trial - chpt.trial$chpts) ~ 
+##     chpt.trial$status, family = "binomial")
+## 
+## Deviance Residuals: 
+##      Min        1Q    Median        3Q       Max  
+## -1.67612  -0.46233  -0.42803  -0.09731   1.67955  
+## 
+## Coefficients:
+##                          Estimate Std. Error z value Pr(>|z|)    
+## (Intercept)               -1.5041     0.4513  -3.333 0.000861 ***
+## chpt.trial$statusmature   -0.8473     0.8668  -0.978 0.328315    
+## chpt.trial$statuscontrol  -2.6703     1.1041  -2.418 0.015585 *  
+## ---
+## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
+## 
+## (Dispersion parameter for binomial family taken to be 1)
+## 
+##     Null deviance: 21.029  on 20  degrees of freedom
+## Residual deviance: 12.035  on 18  degrees of freedom
+## AIC: 31.654
+## 
+## Number of Fisher Scoring iterations: 6
+\end{verbatim}
+\begin{alltt}
+\hlkwd{print}\hlstd{(}\hlkwd{exp}\hlstd{(}\hlkwd{cbind}\hlstd{(}\hlkwd{coef}\hlstd{(mod.chpt),} \hlkwd{confint}\hlstd{(mod.chpt))))}
+\end{alltt}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Waiting for profiling to be done...}}\begin{verbatim}
+##                                           2.5 %    97.5 %
+## (Intercept)              0.22222222 0.082931016 0.5027608
+## chpt.trial$statusmature  0.42857143 0.058657189 2.0812211
+## chpt.trial$statuscontrol 0.06923077 0.003574624 0.4311991
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+In the data, site status "maturing" corresponds to the assembling
+hedgerows (i.e., they are maturing). "Mature" corresponds to the
+non-assembling hedgerows $>$ 10 years old, and "control" corresponds
+to the non-assembling, weedy field margins.
+
+In this analysis the assembling hedgerows are set as the intercept, to
+weedy field margins ("control") have significantly less change points
+than assembling hedgerows ("maturing"). Mature and maturing are about
+the same.
+
+\section{Characteristics of species that contribute to change
+points}
+
+The degree of each species was calculated from a larger dataset
+including observations across the northern central valley
+(approx.~18000 interaction records). These scores were rarefied by
+treating plants like sites in a traditional rarefaction
+\citep{winfree2014species}. This helped to disentangle degree with the
+number of observations of each species. The number of observations of
+a species/the total number of samples at each site was then calculated
+as an estimate of persistence. For each species, its weighted closeness
+in the network is also calculated. All of these calculations were done
+from the dataPrep/dataPrep.R file which needs access to some
+non-publically available data, though the data needed for these
+analysis are all in github.
+
+cv.R calculates the cv of closeness for each species at each site, and
+appends the values for persistence and rarefied degree. For the
+pollinators:
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## ************************************************************}
+\hlcom{## pollinators}
+\hlcom{## ************************************************************}
+
+\hlstd{pol.cv} \hlkwb{<-} \hlkwd{calcCvTrait}\hlstd{(spec,} \hlcom{## specimen data}
+                   \hlstd{specs[specs}\hlopt{$}\hlstd{speciesType} \hlopt{==}\hlstr{"pollinator"}\hlstd{,],} \hlcom{## network metrics}
+                   \hlkwc{trait1}\hlstd{=}\hlstr{"occ.date"}\hlstd{,} \hlcom{## first trait of interest, persistence}
+                   \hlkwc{trait2}\hlstd{=}\hlstr{"r.degree"}\hlstd{,} \hlcom{## second trait of interest}
+                   \hlcom{## rarefied degree}
+                   \hlkwc{time.col}\hlstd{=}\hlstr{"assem"}\hlstd{,} \hlcom{## name of the time column}
+                   \hlkwc{abund.col}\hlstd{=}\hlstr{"weighted.closeness"}\hlstd{,} \hlcom{## network metric}
+                   \hlcom{## of interest}
+                   \hlkwc{cv.function}\hlstd{=cv,} \hlcom{## function to use for calculating cv}
+                   \hlkwc{zero2na}\hlstd{=}\hlnum{TRUE}\hlstd{,} \hlcom{## don't convert NAs to zeros in cv calc}
+                   \hlkwc{standard.cv}\hlstd{=}\hlnum{TRUE}\hlstd{,} \hlcom{## log the cv}
+                   \hlkwc{na.rm}\hlstd{=}\hlnum{TRUE}\hlstd{,}
+                   \hlkwc{species.type}\hlstd{=}\hlstr{"GenusSpecies"}\hlstd{)}
 \end{alltt}
@@
 
-And all of the subsequent R analyses (takes several hours, mostly for
-building null communities for standardizing the network level metrics)
+We can then use a linear mixed model to regress the cv of weighted
+closeness by rarefied degree and persistence. Random effects of site
+and species are also included.
 
@@
 \begin{alltt}
- RScript mainR.R
+\hlstd{pol.mod} \hlkwb{<-} \hlkwd{lmer}\hlstd{(formula.cv,} \hlkwc{data}\hlstd{=pol.cv}\hlopt{$}\hlstd{lm.data)}
+\hlkwd{print}\hlstd{(}\hlkwd{summary}\hlstd{(pol.mod))}
+\end{alltt}
+\begin{verbatim}
+## Linear mixed model fit by REML t-tests use Satterthwaite approximations
+##   to degrees of freedom [lmerMod]
+## Formula: cv ~ occ.date + r.degree + (1 | Site) + (1 | GenusSpecies)
+##    Data: pol.cv$lm.data
+## 
+## REML criterion at convergence: 1043.5
+## 
+## Scaled residuals: 
+##     Min      1Q  Median      3Q     Max 
+## -5.0709 -0.3940  0.1942  0.6762  1.6799 
+## 
+## Random effects:
+##  Groups       Name        Variance Std.Dev.
+##  GenusSpecies (Intercept) 0.01056  0.1028  
+##  Site         (Intercept) 0.05128  0.2265  
+##  Residual                 0.61513  0.7843  
+## Number of obs: 427, groups:  GenusSpecies, 66; Site, 20
+## 
+## Fixed effects:
+##              Estimate Std. Error        df t value Pr(>|t|)    
+## (Intercept)  3.696191   0.094045 36.350000  39.302  < 2e-16 ***
+## occ.date     0.741596   0.246251 15.420000   3.012  0.00855 ** 
+## r.degree    -0.002708   0.001831 17.770000  -1.479  0.15669    
+## ---
+## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
+## 
+## Correlation of Fixed Effects:
+##          (Intr) occ.dt
+## occ.date -0.095       
+## r.degree -0.451 -0.674
+\end{verbatim}
+\begin{alltt}
+\hlkwd{vif.mer}\hlstd{(pol.mod)}
+\end{alltt}
+\begin{verbatim}
+## occ.date r.degree 
+## 1.834127 1.834127
+\end{verbatim}
+\begin{alltt}
+\hlcom{## variance inflation factors < 2, so okay!!! (Zurr et al. 2010)}
 \end{alltt}
@@
 
+There is a significant position relationship between cv of closeness
+and persistence ("occ.date") but not rarified degree ("r.degree"). The
+variance inflation factors are both before 2, which
+\cite{zuur2010protocol} suggests is acceptable.
+
+We see the same pattern in the plant data.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## ************************************************************}
+\hlcom{## plants}
+\hlcom{## ************************************************************}
+
+\hlstd{plant.cv} \hlkwb{<-} \hlkwd{calcCvTrait}\hlstd{(spec,}
+                     \hlstd{specs[specs}\hlopt{$}\hlstd{speciesType} \hlopt{==}\hlstr{"plant"}\hlstd{,],}
+                     \hlkwc{trait1}\hlstd{=}\hlstr{"occ.plant.date"}\hlstd{,}
+                     \hlkwc{trait2}\hlstd{=}\hlstr{"plant.r.degree"}\hlstd{,}
+                     \hlkwc{time.col}\hlstd{=}\hlstr{"assem"}\hlstd{,}
+                     \hlkwc{abund.col}\hlstd{=}\hlstr{"weighted.closeness"}\hlstd{,}
+                     \hlkwc{cv.function}\hlstd{=corCv,}
+                     \hlkwc{zero2na}\hlstd{=}\hlnum{TRUE}\hlstd{,}
+                     \hlkwc{standard.cv}\hlstd{=}\hlnum{TRUE}\hlstd{,}
+                     \hlkwc{na.rm}\hlstd{=}\hlnum{TRUE}\hlstd{,}
+                     \hlkwc{species.type}\hlstd{=}\hlstr{"PlantGenusSpecies"}\hlstd{)}
+
+\hlstd{plant.mod} \hlkwb{<-} \hlkwd{lmer}\hlstd{(formula.plant.cv,} \hlkwc{data}\hlstd{=plant.cv}\hlopt{$}\hlstd{lm.data)}
+\hlkwd{print}\hlstd{(}\hlkwd{summary}\hlstd{(plant.mod))}
+\end{alltt}
+\begin{verbatim}
+## Linear mixed model fit by REML t-tests use Satterthwaite approximations
+##   to degrees of freedom [lmerMod]
+## Formula: 
+## cv ~ occ.plant.date + plant.r.degree + (1 | Site) + (1 | GenusSpecies)
+##    Data: plant.cv$lm.data
+## 
+## REML criterion at convergence: 99.7
+## 
+## Scaled residuals: 
+##     Min      1Q  Median      3Q     Max 
+## -3.6898 -0.3189  0.0615  0.4319  1.9013 
+## 
+## Random effects:
+##  Groups       Name        Variance Std.Dev.
+##  Site         (Intercept) 0.1071   0.3273  
+##  GenusSpecies (Intercept) 0.0000   0.0000  
+##  Residual                 0.2777   0.5270  
+## Number of obs: 50, groups:  Site, 18; GenusSpecies, 15
+## 
+## Fixed effects:
+##                  Estimate Std. Error         df t value Pr(>|t|)    
+## (Intercept)     4.865e+00  2.107e-01  4.614e+01  23.088   <2e-16 ***
+## occ.plant.date -3.180e-01  6.459e-01  4.342e+01  -0.492    0.625    
+## plant.r.degree -6.915e-05  2.736e-03  4.299e+01  -0.025    0.980    
+## ---
+## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
+## 
+## Correlation of Fixed Effects:
+##             (Intr) occ.p.
+## occ.plnt.dt -0.022       
+## plant.r.dgr -0.607 -0.684
+\end{verbatim}
+\begin{alltt}
+\hlkwd{vif.mer}\hlstd{(plant.mod)}
+\end{alltt}
+\begin{verbatim}
+## occ.plant.date plant.r.degree 
+##       1.878196       1.878196
+\end{verbatim}
+\begin{alltt}
+\hlcom{## variance inflation factors <1, so okay (Zurr et al. 2010)}
+
+\hlcom{## ************************************************************}
+\hlcom{## the reviewers wanted a quantile regression, can only include on}
+\hlcom{## random effect at a time}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+You can play with using different network metrics to calculate the cv
+of. You can also change the function used for calculating the cv, or
+whatever function you would like to calculate between years.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+ \hlcom{## network metric options}
+ \hlkwd{colnames}\hlstd{(specs)[}\hlnum{1}\hlopt{:}\hlnum{24}\hlstd{]}
+\end{alltt}
+\begin{verbatim}
+##  [1] "degree"                        "normalised.degree"            
+##  [3] "species.strength"              "interaction.push.pull"        
+##  [5] "nestedrank"                    "PDI"                          
+##  [7] "resource.range"                "species.specificity.index"    
+##  [9] "PSI"                           "node.specialisation.index.NSI"
+## [11] "betweenness"                   "weighted.betweenness"         
+## [13] "closeness"                     "weighted.closeness"           
+## [15] "Fisher.alpha"                  "partner.diversity"            
+## [17] "effective.partners"            "proportional.generality"      
+## [19] "proportional.similarity"       "d"                            
+## [21] "tot.int"                       "mean.k"                       
+## [23] "sd.k"                          "k"
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+The reviewers suggested we also do quantile regressions, which can only
+include on random effect at a time, but yielded similar results.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## ************************************************************}
+\hlcom{## the reviewers wanted a quantile regression, can only include on}
+\hlcom{## random effect at a time}
+
+\hlcom{## occ.date is significant}
+\hlstd{pol.lm.cv.quant} \hlkwb{<-} \hlkwd{lqmm}\hlstd{(}\hlkwc{fixed}\hlstd{=cv} \hlopt{~} \hlstd{occ.date} \hlopt{+} \hlstd{r.degree,} \hlkwc{random}\hlstd{=}\hlopt{~}\hlnum{1}\hlstd{,}
+                        \hlkwc{group}\hlstd{= GenusSpecies,}
+                        \hlkwc{data}\hlstd{=pol.cv}\hlopt{$}\hlstd{lm.data,}
+                        \hlkwc{control}\hlstd{=}\hlkwd{list}\hlstd{(}\hlkwc{LP_max_iter}\hlstd{=}\hlnum{10}\hlopt{^}\hlnum{4}\hlstd{))}
+\hlstd{pol.sum.boot.quant} \hlkwb{<-} \hlkwd{summary.boot.lqmm}\hlstd{(}\hlkwd{boot}\hlstd{(pol.lm.cv.quant,}
+                                             \hlkwc{R}\hlstd{=}\hlnum{100}\hlstd{))}
+\end{alltt}
+\begin{verbatim}
+## Quantile 0.5 
+##                  Value Std. Error lower bound upper bound  Pr(>|t|)    
+## (Intercept)  3.7160065  0.0920970   3.5332661      3.8987 < 2.2e-16 ***
+## occ.date     0.7040012  0.2676157   0.1729936      1.2350  0.009884 ** 
+## r.degree    -0.0010938  0.0032020  -0.0074472      0.0053  0.733375    
+## scale        0.2869339  0.0191323   0.2489713      0.3249 < 2.2e-16 ***
+## ---
+## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+\section{Species and interaction turnover}
+\subsection{Pollinators, plants and interactions}
+
+To calculate species and interaction turnover over, we borrow methods
+from the $\beta$-diversity world \citep{ponisio2015farm,
+chase-2011-art24, anderson-2006-683, anderson-2011-19,
+anderson-2006-245}.  The goal is to calculate the pairwise
+dissimilarity between years within each site of the dataset using a
+dissimilarity estimator that incorporates species abundances, while
+also accounting for unobserved species \citep{chao-2005-148}, and then
+following \citep{chase-2011-art24}, correct our estimates of
+$\beta$-diversity using null models.
+
+The nulls.R script prepares null communities in variety of ways to
+calculate an expected distribution by generating randomized
+communities and calculating the dissimilarity of these communities.
+To do this, we defined the species pool within each site as the
+species and number of individuals present across all samples from that
+site.  We then generated $999$ random communities by constraining
+either 1) the total number of individuals caught at each site
+(individual nulls) or 2) the species richness at each site (alpha div
+nulls). We also had a binary option that constrained but marginal
+sums (occurrence nulls).
+
+Next, in the function calcBetaStatus.R executed by beta-div.R, for the
+observed and nulls communities, we calculated the pair-wise
+dissimilarity between sites.  We then used the null community
+dissimilarities to calculate the expected $\beta$-diversity when
+communities are randomly assembled but constrained so that they have
+either the same 1) number of individuals or 2) species richness as the
+observed communities and with species drawn from a meta-community with
+the same species abundance distributions.  In order to do this, we
+followed \cite{chase-2011-art24}.  Specifically, we calculated the
+fraction of randomly assembled communities with dissimilarity values
+less than (and half of those equal to) that of the observed community.
+We used this fraction as a ``corrected dissimilarity score'' for our
+observed data.  Corrected dissimilarity values near one indicate that
+our observed communities exhibit more species turnover between sites
+than expected under a random assembly process while values near $0.5$
+indicate that our observed communities exhibit levels of turnover more
+in line with the null expectation. We calculated the corrected
+dissimilarities for each type of randomized community. We then
+generated principle coordinate axes (PCoA) based on the corrected
+pair-wise dissimilarities \citep{oksanen-2013, anderson-2006-683,
+anderson-2011-19, anderson-2006-245}.  We calculated dispersion for
+each site type by finding the centroid in PCoA space for that site
+type and then calculating the distances from sites of that type to
+that centroid. The centroid is the point that minimizes the sum of
+these distances.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{dis} \hlkwb{<-} \hlkwd{mapply}\hlstd{(}\hlkwa{function}\hlstd{(}\hlkwc{a}\hlstd{,} \hlkwc{b}\hlstd{,} \hlkwc{c}\hlstd{,} \hlkwc{d}\hlstd{)}
+    \hlkwd{calcBetaStatus}\hlstd{(}\hlkwc{comm}\hlstd{= a,} \hlcom{## observed communities}
+                   \hlkwc{status}\hlstd{= b,} \hlcom{## vector of site types}
+                   \hlstd{dis.method,} \hlcom{## dissimilarity metric}
+                   \hlkwc{nulls}\hlstd{=c,} \hlcom{## null communities}
+                   \hlkwc{occ}\hlstd{=binary,} \hlcom{## binary or abundance weighted?}
+                   \hlkwc{years}\hlstd{=d,} \hlcom{## calculate beta div within?}
+                   \hlkwc{sub}\hlstd{=type,}
+                   \hlkwc{zscore}\hlstd{=}\hlnum{FALSE}\hlstd{),} \hlcom{## use Chase method not zscores}
+    \hlkwc{a}\hlstd{=comm}\hlopt{$}\hlstd{comm,}
+    \hlkwc{b}\hlstd{=comm}\hlopt{$}\hlstd{status,}
+    \hlkwc{c}\hlstd{= nulls,}
+    \hlkwc{d}\hlstd{= comm}\hlopt{$}\hlstd{comm,}
+    \hlkwc{SIMPLIFY}\hlstd{=}\hlnum{FALSE}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+Changing zscore to TRUE calculates dissimilarity zscores instead of
+\cite{chase-2011-art24}'s pvalue-like method.
+
+Toggling alpha and beta in the beta-div.R script changes between the
+null model options. When binary is TRUE, the binary null models are
+used and the dissimilarity method is "jaccard" (all set up in the
+initialize\_beta.R). When alpha is TRUE, species richness but not
+individual abundances are constrained, and the reverse when alpha is
+FALSE. The chao dissimilarity measure is used in both
+abundance-weighted cases.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+ \hlstd{binary} \hlkwb{<-} \hlnum{FALSE}
+ \hlstd{alpha} \hlkwb{<-} \hlnum{TRUE}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+The dispersion values were then used in linear mixed-effect models to
+investigate the effect of different site types on
+"$\beta$-diversity". Site is included as a random effect. For
+pollinators:
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{mod} \hlkwb{<-} \hlkwd{lmer}\hlstd{(dist} \hlopt{~} \hlstd{status} \hlopt{+}  \hlstd{(}\hlnum{1}\hlopt{|}\hlstd{site),}
+            \hlkwc{data}\hlstd{=dats)}
+\hlkwd{print}\hlstd{(}\hlkwd{summary}\hlstd{(mod))}
+\end{alltt}
+\begin{verbatim}
+## Linear mixed model fit by REML t-tests use Satterthwaite approximations
+##   to degrees of freedom [lmerMod]
+## Formula: dist ~ status + (1 | site)
+##    Data: dats
+## 
+## REML criterion at convergence: -173.6
+## 
+## Scaled residuals: 
+##     Min      1Q  Median      3Q     Max 
+## -2.5340 -0.7712  0.0072  0.6896  2.3481 
+## 
+## Random effects:
+##  Groups   Name        Variance  Std.Dev. 
+##  site     (Intercept) 5.471e-20 2.339e-10
+##  Residual             2.268e-02 1.506e-01
+## Number of obs: 198, groups:  site, 38
+## 
+## Fixed effects:
+##                  Estimate Std. Error         df t value Pr(>|t|)    
+## (Intercept)      0.388962   0.012819 195.000000  30.342   <2e-16 ***
+## statusmature    -0.050733   0.025639 195.000000  -1.979   0.0493 *  
+## statusmaturing  -0.008949   0.042240 195.000000  -0.212   0.8324    
+## ---
+## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
+## 
+## Correlation of Fixed Effects:
+##             (Intr) statsmtr
+## statusmatur -0.500         
+## statusmtrng -0.303  0.152
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+There is little difference between any of the options, except mature
+sites have significantly less turnover in pollinators with the alpha
+null options, which is likely most appropriate.
+
+Changing "type" to "pols", "plants", or "Ints" calculates turnover of
+pollinators, plants and interactions respectively. All of these are
+executed in the main.sh script.
+
+We also included a linear model that took temporal auto-correlation
+into account, with the same results but not a large gain in fit given
+the additional parameters (tested using AIC)
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{mod.spatial} \hlkwb{<-} \hlkwd{lme}\hlstd{(dist} \hlopt{~} \hlstd{status,}
+                   \hlkwc{random} \hlstd{=} \hlopt{~} \hlnum{1} \hlopt{+} \hlstd{cYear} \hlopt{|} \hlstd{site,}
+                   \hlkwc{correlation}\hlstd{=}\hlkwd{corAR1}\hlstd{(}\hlkwc{form}\hlstd{=}\hlopt{~}\hlstd{cYear),}
+                   \hlkwc{data}\hlstd{=dats,}
+                   \hlkwc{control}\hlstd{=}\hlkwd{list}\hlstd{(}\hlkwc{maxIter}\hlstd{=}\hlnum{10}\hlopt{^}\hlnum{5}\hlstd{,} \hlkwc{niterEM}\hlstd{=}\hlnum{10}\hlopt{^}\hlnum{5}\hlstd{))}
+
+\hlkwd{summary}\hlstd{(mod.spatial)}
+\end{alltt}
+\begin{verbatim}
+## Linear mixed-effects model fit by REML
+##  Data: dats 
+##         AIC       BIC   logLik
+##   -160.7266 -134.5426 88.36329
+## 
+## Random effects:
+##  Formula: ~1 + cYear | site
+##  Structure: General positive-definite, Log-Cholesky parametrization
+##             StdDev     Corr  
+## (Intercept) 0.05875859 (Intr)
+## cYear       0.01568352 -1    
+## Residual    0.14390959       
+## 
+## Correlation Structure: ARMA(1,0)
+##  Formula: ~cYear | site 
+##  Parameter estimate(s):
+##         Phi1 
+## -0.007385481 
+## Fixed effects: dist ~ status 
+##                     Value  Std.Error  DF   t-value p-value
+## (Intercept)     0.3881307 0.01264276 160 30.699827  0.0000
+## statusmature   -0.0478766 0.02679690  35 -1.786649  0.0827
+## statusmaturing  0.0025782 0.04266169  35  0.060434  0.9522
+##  Correlation: 
+##                (Intr) statsmtr
+## statusmature   -0.472         
+## statusmaturing -0.296  0.140  
+## 
+## Standardized Within-Group Residuals:
+##         Min          Q1         Med          Q3         Max 
+## -2.76201315 -0.71766223  0.03396789  0.67383510  2.20624889 
+## 
+## Number of Observations: 198
+## Number of Groups: 38
+\end{verbatim}
+\begin{alltt}
+\hlkwd{AIC}\hlstd{(mod)} \hlopt{-}\hlkwd{AIC}\hlstd{(mod.spatial)}
+\end{alltt}
+\begin{verbatim}
+## [1] -2.82759
+\end{verbatim}
+\begin{alltt}
+\hlcom{## spatial auto-correlation has a larger AIC}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+\subsection{Weighted interaction turnover}
+
+To calculate weighted interaction turnover we borrowed concepts from
+calculating phylogenetic $\beta$-diversity. We began by creating a
+"phylogenetic" tree of interactions based on their position in the
+network using the getLinkCommunities function in the linkcomm package
+\citep{kalinka2011linkcomm}. For bi-partite networks, the set of
+neighbors is used to count nodes for the edge similarity metric
+because node i and node j cannot share an edge in a bi-partite
+network. The partition density for bi-partite networks is calculated
+as:
+
+\begin{equation}
+ (2/M)*\sum_c m_c*(m_c + 1-n_c)/(2*n_{c_0}*n_{c_1} - 2*(n_c - 1))
+\end{equation}
+
+where M is the total number of edges, $m_c$ is the number of edges in
+subset $c$, $n_c$ is the number of nodes in subset c, $n_{c_0}$ is the
+number of nodes in partition 0, and $n_{c_1}$ is the number of nodes
+in partition 1.
+
+Use the resulting dendrogram and interaction dissimilarity matrices,
+we use the comdist function in the picante package
+\citep{picante-2010-1463} to calculate expected "phylogenetic"
+distance separating two interactions drawn randomly from different
+communities. This gives us a measure of interaction turnover weighted
+by how interaction similarity. This is all wrapped up in the function
+CalcCommDis
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{phylo.int} \hlkwb{<-} \hlkwd{calcCommDis}\hlstd{(spec,} \hlstr{"Int"}\hlstd{, lc,} \hlkwc{abund.w}\hlstd{=}\hlnum{TRUE}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+We use a linear mixed model to regress weighted interaction turnover
+and site type and find no statistically significant differences.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{phylo.int.mod} \hlkwb{<-} \hlkwd{lmer}\hlstd{(PhyloInt} \hlopt{~} \hlstd{SiteStatus} \hlopt{+}
+                          \hlstd{(}\hlnum{1}\hlopt{|}\hlstd{Site),}
+                      \hlkwc{data}\hlstd{=phylo.int}\hlopt{$}\hlstd{phylo.int)}
+\hlkwd{print}\hlstd{(}\hlkwd{summary}\hlstd{(phylo.int.mod))}
+\end{alltt}
+\begin{verbatim}
+## Linear mixed model fit by REML t-tests use Satterthwaite approximations
+##   to degrees of freedom [lmerMod]
+## Formula: PhyloInt ~ SiteStatus + (1 | Site)
+##    Data: phylo.int$phylo.int
+## 
+## REML criterion at convergence: -1087.4
+## 
+## Scaled residuals: 
+##     Min      1Q  Median      3Q     Max 
+## -4.4271 -0.3718  0.1392  0.5481  2.9588 
+## 
+## Random effects:
+##  Groups   Name        Variance Std.Dev.
+##  Site     (Intercept) 0.003341 0.05781 
+##  Residual             0.004470 0.06686 
+## Number of obs: 459, groups:  Site, 38
+## 
+## Fixed effects:
+##                    Estimate Std. Error       df t value Pr(>|t|)    
+## (Intercept)         0.90778    0.01965 37.00000  46.198  < 2e-16 ***
+## SiteStatuscontrol  -0.08206    0.02414 35.13000  -3.400  0.00169 ** 
+## SiteStatusmaturing  0.03339    0.03065 33.20000   1.089  0.28385    
+## ---
+## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
+## 
+## Correlation of Fixed Effects:
+##             (Intr) StSttsc
+## StSttscntrl -0.814        
+## StSttsmtrng -0.641  0.522
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+\section{Temporal changes in interaction patterns}
+\subsection{Network structure}
+
+The baci.R script (before-after-control-impact, baci, the orginal
+design of the study, vs.~ the word for kisses in Italian) calculates
+various network metrics for the assembling hedgerows, standardizes
+them using null communities, then include then as response variables
+in linear models. The number of null communities can be changed by
+setting N at the top of the script.
+
+Calculate and standardize the network metrics. CalcNetworkMetrics
+takes an argument "networkLevel.mets" which is passed to networkLevel
+within the package bipartite \citep{bipartite} if other metrics are
+desired. Unlike the turnover analysis, the networks include
+non-integers because the mean was taken over sampling rounds within a
+year. We therefore use the null model proposed by \citep{Galeano2009}
+to randomize the communities while fixing the total number of
+interactions, species and interaction frequency distributions (though
+any function that takes a single interaction matrix as an argument can
+be substituted using the null.fun argument of the calcNullStat
+function.
+
+Calculate and pretty network metrics.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+ \hlstd{mets} \hlkwb{<-} \hlkwd{lapply}\hlstd{(nets, calcNetworkMetrics, N)}
+ \hlstd{cor.dats} \hlkwb{<-} \hlkwd{prepDat}\hlstd{(mets,  spec)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+Then regress the nestedness, modularity, specialization and
+connectance by time ("ypr" in the data, years post restoration).
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{lapply}\hlstd{(mods, summary)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+As per the suggestion of reviewers, we also tested the same hypotheses
+with linear mixed models that included temporal auto-correlation, and
+splines. Like the turnover models, the temporal auto-correlation
+models showed similar results, but the fit was not substantially
+improved by adding the additional parameters. The spline chose to fit
+a line, indicating that the linear models were sufficient. They are
+time consuming to run so the results are not included here.
+
+\subsection{Network robustness}
+
+The resilience.R script calculates resilience to species extinction
+using the method of \citep{Memmott2004}. Species can be dropped by
+abundance ( extinction.method = "abund") or degree (extinction.method
+="degree"). From the helpfile: "The procedure of this function is
+simple. For example imagine the web to represent a pollination web, in
+which pollinators die one by one. Set all entries of a column to zero,
+see how may rows are now also all-zero (i.e., species that are now not
+pollinated any more), and count these as secondary extinctions of the
+primary exterminated pollinator." The area under the curve is then
+calculated.
+
+This area is then regressed against time ("ypr" again).
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## no effect of ypr on robustness}
+\hlstd{mod.ypr} \hlkwb{<-} \hlkwd{lmer}\hlstd{(Robustness} \hlopt{~} \hlstd{ypr}
+                \hlopt{+} \hlstd{(}\hlnum{1}\hlopt{|}\hlstd{Site)} \hlopt{+} \hlstd{(}\hlnum{1}\hlopt{|}\hlstd{Year),}
+                \hlkwc{data}\hlstd{=res[}\hlopt{!}\hlkwd{is.na}\hlstd{(res}\hlopt{$}\hlstd{ypr),])}
+\hlkwd{summary}\hlstd{(mod.ypr)}
+\end{alltt}
+\begin{verbatim}
+## Linear mixed model fit by REML t-tests use Satterthwaite approximations
+##   to degrees of freedom [lmerMod]
+## Formula: Robustness ~ ypr + (1 | Site) + (1 | Year)
+##    Data: res[!is.na(res$ypr), ]
+## 
+## REML criterion at convergence: -167.6
+## 
+## Scaled residuals: 
+##      Min       1Q   Median       3Q      Max 
+## -2.03661 -0.73640 -0.02933  0.62103  2.40910 
+## 
+## Random effects:
+##  Groups   Name        Variance  Std.Dev.
+##  Site     (Intercept) 0.0019852 0.04456 
+##  Year     (Intercept) 0.0001163 0.01078 
+##  Residual             0.0048546 0.06967 
+## Number of obs: 80, groups:  Site, 14; Year, 8
+## 
+## Fixed effects:
+##              Estimate Std. Error        df t value Pr(>|t|)    
+## (Intercept)  0.331489   0.016356 11.040000  20.267 4.36e-10 ***
+## ypr          0.001166   0.004144 34.880000   0.281     0.78    
+## ---
+## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
+## 
+## Correlation of Fixed Effects:
+##     (Intr)
+## ypr -0.395
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+\clearpage
+\bibliography{refs}
+\clearpage
+
 \end{document}

### 2017-05-18 analysis/changePoint/comparisons.R

@@ -3,7 +3,6 @@ library(lme4)
 
 setwd("changePoint")
 ## setwd("~/Dropbox/hedgerow_assembly/analysis/changePoint")
-
 source('../../dataPrep/src/misc.R')
 load('cptPeel/baci/graphs.Rdata')
 load('../../data/networks/allSpecimens.Rdata')
@@ -17,6 +16,7 @@ dats <- dats[dats$num.runs >= 0.8,]
 ## **********************************************************
 ## binomial tests
 ## **********************************************************
+
 chpt.trial <- aggregate(spec$Year, list(Site=spec$Site),
                            FUN=function(x) length(unique(x)))
 chpt.trial$x <- chpt.trial$x - 1
@@ -43,13 +43,15 @@ chpt.trial <- chpt.trial[order(chpt.trial$status),]
 ## binomial model with change points as successes
 chpt.trial$status <- factor(chpt.trial$status,
                             levels=c("maturing", "mature", "control"))
-
+## @knitr external_binomialreg
 mod.chpt <- glm(cbind(chpt.trial$chpts,
                       chpt.trial$trial - chpt.trial$chpts) ~
     chpt.trial$status, family="binomial")
 print(summary(mod.chpt))
 
 print(exp(cbind(coef(mod.chpt), confint(mod.chpt))))
+
+## @knitr external_otherCalc
 
 # controls with change points
 nrow(chpt.trial[chpt.trial$status == "control" &
@@ -70,8 +72,7 @@ nrow(chpt.trial[chpt.trial$chpts != 0,])/
   nrow(chpt.trial)
 
 
-## maturing has more successes than mature and controls, and mature
-## and controls have about the same
+## proportions
 
 chpt.trial$prop <- chpt.trial$chpts/chpt.trial$trial
 tapply(chpt.trial$prop, chpt.trial$status, mean)

### 2017-05-18 analysis/variability/beta-div.R

@@ -1,12 +1,18 @@
 ## rm(list=ls())
 ## setwd('~/Dropbox/hedgerow_assembly/analysis/variability')
 setwd('variability')
+
 args <- commandArgs(trailingOnly=TRUE)
 
 binary <- FALSE
 alpha <- TRUE
 ## ints or pols
-type <- args[1]
+
+if(length(args) == 0){
+    type <- "pols"
+} else{
+    type <- args[1]
+}
 source('src/initialize_beta.R')
 
 ## ************************************************************
@@ -14,31 +20,34 @@ source('src/initialize_beta.R')
 ## centroid for each site
 ## ************************************************************
 
+## @knitr external_beta_div
 dis <- mapply(function(a, b, c, d)
-              beta.status(comm= a,
-                          status= b,
-                          dis.method,
-                          nulls=c,
-                          occ=binary,
-                          years=d,
-                          sub=type),
-              a=comm$comm,
-              b=comm$status,
-              c= nulls,
-              d= comm$comm,
-              SIMPLIFY=FALSE)
+    calcBetaStatus(comm= a, ## observed communities
+                   status= b, ## vector of site types
+                   dis.method, ## dissimilarity metric
+                   nulls=c, ## null communities
+                   occ=binary, ## binary or abundance weighted?
+                   years=d, ## calculate beta div within?
+                   sub=type,
+                   zscore=FALSE), ## use Chase method not zscores
+    a=comm$comm,
+    b=comm$status,
+    c= nulls,
+    d= comm$comm,
+    SIMPLIFY=FALSE)
+## @knitr external_beta_div_end
 
 dats <- data.frame(site=comm$sites,
                    status=unlist(comm$status),
                    year=unlist(comm$years),
                    dist=unlist(sapply(dis, function(x)
-                     x$distances)))
+                       x$distances)))
 rownames(dats) <- NULL
 
 
 save(dats, file= file.path('saved/speciesTurnover',
-             sprintf('%s.pdf', paste(dis.method, alpha, occ, type,
-                                     sep='_'))))
+                           sprintf('%s.pdf', paste(dis.method, alpha, occ, type,
+                                                   sep='_'))))
 
 ## ************************************************************
 ## invid nulls all not sig, alpha nulls mature marginally sig less,
[…]

### 2017-05-18 analysis/variability/cv.R

@@ -3,26 +3,34 @@ setwd('variability')
 
 source('src/initialize.R')
 
+## @knitr external_cv
 ## ************************************************************
 ## pollinators
 ## ************************************************************
 
-pol.cv <- cv.trait(spec,
-                   specs[specs$speciesType =="pollinator",],
-                   trait1="occ.date",
-                   trait2="r.degree",
-                   method= "time", time.col="assem",
-                   abund.col="weighted.closeness",
-                   cv.function=cv,
-                   zero2na=TRUE,
-                   standard.cv=TRUE,
-                   na.rm=TRUE)
+pol.cv <- calcCvTrait(spec, ## specimen data
+                   specs[specs$speciesType =="pollinator",], ## network metrics
+                   trait1="occ.date", ## first trait of interest, persistence
+                   trait2="r.degree", ## second trait of interest
+                   ## rarefied degree
+                   time.col="assem", ## name of the time column
+                   abund.col="weighted.closeness", ## network metric
+                   ## of interest
+                   cv.function=cv, ## function to use for calculating cv
+                   zero2na=TRUE, ## don't convert NAs to zeros in cv calc
+                   standard.cv=TRUE, ## log the cv
+                   na.rm=TRUE,
+                   species.type="GenusSpecies")
+
+## @knitr external_cv_lm
 
 pol.mod <- lmer(formula.cv, data=pol.cv$lm.data)
 print(summary(pol.mod))
 vif.mer(pol.mod)
 
 ## variance inflation factors < 2, so okay!!! (Zurr et al. 2010)
+
+## @knitr external_cv_quantreg
 
 ## ************************************************************
 ## the reviewers wanted a quantile regression, can only include on
@@ -36,16 +44,18 @@ pol.lm.cv.quant <- lqmm(fixed=cv ~ occ.date + r.degree, random=~1,
 pol.sum.boot.quant <- summary.boot.lqmm(boot(pol.lm.cv.quant,
                                              R=100))
 
+## @knitr external_cv_plants
 ## ************************************************************
-## plant
+## plants
 ## ************************************************************
-plant.cv <- cv.trait(spec,
+
+plant.cv <- calcCvTrait(spec,
                      specs[specs$speciesType =="plant",],
                      trait1="occ.plant.date",
                      trait2="plant.r.degree",
[…]

### 2017-05-18 analysis/variability/linkComm.R

@@ -1,299 +1,0 @@
-function (network, hcmethod = "average", use.all.edges = FALSE,
-    edglim = 10^4, directed = FALSE, dirweight = 0.5, bipartite = FALSE,
-    dist = NULL, plot = TRUE, check.duplicates = TRUE, removetrivial = TRUE,
-    verbose = TRUE)
-{
-    if (is.character(network) && !is.matrix(network)) {
-        if (file.access(network) == -1) {
-            stop(cat("\nfile not found: \"", network, "\"\n",
-                sep = ""))
-        }
-        else {
-            network <- read.table(file = network, header = FALSE)
-        }
-    }
-    x <- network
-    rm(network)
-    if (ncol(x) == 3) {
-        wt <- as.numeric(as.character(x[, 3]))
-        if (length(which(is.na(wt) == TRUE)) > 0) {
-            stop("\nedge weights must be numerical values\n")
-        }
-        x <- cbind(as.character(x[, 1]), as.character(x[, 2]))
-    }
-    else if (ncol(x) == 2) {
-        x <- cbind(as.character(x[, 1]), as.character(x[, 2]))
-        wt <- NULL
-    }
-    else {
-        stop("\ninput data must be an edge list with 2 or 3 columns\n")
-    }
-    if (check.duplicates) {
-        dret <- edge.duplicates(x, verbose = verbose)
-        x <- dret$edges
-        if (!is.null(wt)) {
-            if (length(dret$inds) > 0) {
-                wt <- wt[-dret$inds]
-            }
-        }
-        rm(dret)
-    }
-    el <- x
-    rm(x)
-    len <- nrow(el)
-    nnodes <- length(unique(c(as.character(el[, 1]), as.character(el[,
-        2]))))
-    intel <- integer.edgelist(el)
-    edges <- intel$edges
-    node.names <- names(intel$nodes)
-    numnodes <- length(node.names)
-    if (bipartite) {
-        big <- graph.edgelist(as.matrix(el), directed = directed)
-        bip.test <- bipartite.mapping(big)
-        if (!bip.test$res) {
-            stop("\nnetwork is not bi-partite; change bipartite argument to FALSE\n")
-        }
-        bip <- rep(1, length(bip.test$type))
-        bip[which(bip.test$type == FALSE)] <- 0
-        names(bip) <- V(big)$name
-        bip <- bip[match(node.names, names(bip))]
-        rm(big, bip.test)
-    }
-    else {
[…]

### 2017-05-18 analysis/variability/nulls.R

@@ -4,7 +4,7 @@ source('src/initialize.R')
 source('src/commPrep.R')
 save.dir.comm <- "saved/communities"
 save.dir.nulls <- "saved/nulls"
-nnull <- 99
+nnull <- 999
 
 src.dir2 <- 'analysis/beta-div'
 source(file.path(src.dir2, 'src/misc.R'))

### 2017-05-18 analysis/variability/src/beta.R

@@ -1,12 +1,12 @@
 
-beta.status <- function(comm,
+calcBetaStatus <- function(comm,
                         status,
                         dis.method,
                         nulls,
                         sub= "pol",
                         occ= FALSE,
                         years,
-                        zscore=FALSE){
+                        zscore=FALSE){ ## calculate zscores?
     ## computes dispersion of community matrices, returns output of
     ## vegan function betadisper
 

### 2017-05-18 analysis/variability/src/cvCalc.R

@@ -4,11 +4,10 @@ corCv <- function(x,...){
     cv(x)*(1 + (1/4*length(x)))
 }
 
-cv.trait <- function(spec.dat,
+calcCvTrait <- function(spec.dat,
                      byType,
                      trait1,
                      trait2,
-                     method,
                      time.col,
                      abund.col,
                      status.order=c("control", "maturing", "mature"),
@@ -16,9 +15,20 @@ cv.trait <- function(spec.dat,
                      zero2na =FALSE,
                      standard.cv=TRUE,
                      species.type="GenusSpecies",...){
-    ## cv.function can be corCv, cv, or sd
-    ## standard cv devides cv by 100 to avoid too large of numbers for the
-    ## lm
+    ##spec.dat: specimens data
+    ## byType: table of species level network metrics
+    ## trait1: first trait of interest (to regress cv
+    ## against)
+    ## trait2: second trait of interest
+    ## time.col: name of the time column
+    ## abund.col: network metric column name
+    ## cv.function: function to use for calculating cv. cv.function
+    ## can be corCv, cv, or sd
+
+    ## zero2na: converts NAs to zeros, but NAs are not
+    ## inclcued in the cv calculations whereas zeros are, so
+    ## recommended to keep as FALSE
+    ## standard cv logs the cv
     byStatus <- split(byType, byType$SiteStatus)
     bySite <- lapply(byStatus, function(x) {split(x, x$Site)})
     bySite <- unlist(bySite, recursive=FALSE)
@@ -47,9 +57,9 @@ cv.trait <- function(spec.dat,
     dats <- cbind(dats, spec.dat[, c(trait1, trait2)][match(dats$GenusSpecies,
                                                             spec.dat[,
                                                                      species.type]),])
-   lm.dats <- dats[!is.na(dats$cv) & !is.na(dats[,trait1]) & !is.na(dats[,trait2]),]
+    lm.dats <- dats[!is.na(dats$cv) & !is.na(dats[,trait1]) & !is.na(dats[,trait2]),]
     return(list(data=dats,
-           lm.data =lm.dats))
+                lm.data =lm.dats))
 }
 
 ## variance inflation factor

### 2017-05-18 dataPrep/dataPrep.R

@@ -181,8 +181,11 @@ spec$plant.r.degree <- traits$r.degree[match(spec$PlantGenusSpecies,
                                          traits$GenusSpecies)]
 
 ## occurence
+## load site by sample by species array
 load('~/Dropbox/hedgerow/data_sets/matrices/net/bee.syr.RData')
 
+## calculate the proportion of samples a species was present at for
+## each site
 occ <- apply(mat, c(3,1), calcOccArray)
 spec$occ.date <- apply(spec, 1, findOcc)
 traits$occ.date <- spec$occ.date[match(traits$GenusSpecies,

### 2017-05-18 analysis/writeup/writeup.tex
lots of small cahnges for reproducibility

@@
 
-In our study we examining the temporal dynamics of plant-pollinator
-network assembly using variety of different methods include 1) network
+In our study we examin the temporal dynamics of plant-pollinator
+network assembly using variety of different methods including 1) network
 change point detection 2) node/species-level position variation, 3)
@@
 file, including the specific versions of the python dendropy library,
-which is very finicky. We have had no problems with back compatibility
-for the R scripts (in terms of R versions from v.~2-3.3 or packages)
-but the change point analysis must be run in python2 (not easily
-convertible to python3, though this is potentially in progress).
+which is very finicky. We have had no problems with backward
+compatibility for the R scripts (in terms of R versions from v.~2-3.4
+or packages) but the change point analysis must be run in python2 (not
+easily convertible to python3, though this is potentially in
+progress).
 
@@
 
-We will walk through each analysis individually.
+We will walk through each the main script for each analysis
+individually.
 
@@
 The output of the change point analysis is saved as a list of the
-change points identified and like likelihood scores (formatted by
+change points identified and likelihood scores (formatted by
 prepChangePointOutput.R).
@@
 
-In this analysis the assembling hedgerows are set as the intercept, to
+In this analysis the assembling hedgerows are set as the intercept, so
 weedy field margins ("control") have significantly less change points
@@
 closeness by rarefied degree and persistence. Random effects of site
-and species are also included.
+and species are included. Because degrees of freedom are not
+well-dinfed for a mixed model Satterthwaite approximations implemented
+in the lmeTest package are used so that p-values can be calculated for
+the linear mixed models \citep{lmetest}.
 
@@
 
-We see the same pattern in the plant data.
+We found no statistically significant relationships in the plant data.
 
@@
 ##                  Value Std. Error lower bound upper bound Pr(>|t|)    
-## (Intercept)  3.7160065  0.0920970   3.5332661      3.8987 < 2.2e-16 ***
-## occ.date     0.7040012  0.2676157   0.1729936      1.2350  0.009884 ** 
-## r.degree    -0.0010938  0.0032020  -0.0074472      0.0053  0.733375    
-## scale        0.2869339  0.0191323   0.2489713      0.3249 < 2.2e-16 ***
+## (Intercept)  3.7160065  0.0854433   3.5464685      3.8855  < 2e-16 ***
+## occ.date     0.7040012  0.2454156   0.2170433      1.1910  0.00504 ** 
+## r.degree    -0.0010938  0.0032266  -0.0074960      0.0053  0.73533    
+## scale        0.2869339  0.0185467   0.2501333      0.3237  < 2e-16 ***
 ## ---
@@
 type and then calculating the distances from sites of that type to
-that centroid. The centroid is the point that minimizes the sum of
-these distances.
+that centroid using the betadisper function in the vegan packages
+\citep{oksanen-2013}. The centroid is the point that minimizes the sum
+of these distances.
 
@@
 FALSE. The chao dissimilarity measure is used in both
-abundance-weighted cases.
+abundance-weighted cases. There is little difference between any of
+the options, except mature sites have significantly less turnover in
+pollinators with the alpha null options, which is likely most
+appropriate.
 
@@
 \end{knitrout}
+
+Changing "type" to "pols", "plants", or "Ints" calculates turnover of
+pollinators, plants and interactions respectively. All of these
+options are executed in the main.sh script.
 
@@
 
-There is little difference between any of the options, except mature
-sites have significantly less turnover in pollinators with the alpha
-null options, which is likely most appropriate.
-
-Changing "type" to "pols", "plants", or "Ints" calculates turnover of
-pollinators, plants and interactions respectively. All of these are
-executed in the main.sh script.
-
-We also included a linear model that took temporal auto-correlation
-into account, with the same results but not a large gain in fit given
-the additional parameters (tested using AIC)
+Follow a reviwer's suggestion, we also included a linear model that
+took temporal auto-correlation into account, with the same results but
+not a large gain in fit given the additional parameters (tested using
+AIC)
 
@@
 network using the getLinkCommunities function in the linkcomm package
-\citep{kalinka2011linkcomm}. For bi-partite networks, the set of
+\citep{kalinka2011linkcomm}. For bipartite networks, the set of
 neighbors is used to count nodes for the edge similarity metric
-because node i and node j cannot share an edge in a bi-partite
-network. The partition density for bi-partite networks is calculated
+because node i and node j cannot share an edge in a bipartite
+network. The partition density for bipartite networks is calculated
 as:
@@
 number of nodes in partition 0, and $n_{c_1}$ is the number of nodes
-in partition 1.
+in partition 1 (from the getLinkCommunities helpfile).
 
-Use the resulting dendrogram and interaction dissimilarity matrices,
+With the resulting dendrogram and interaction dissimilarity matrices,
 we use the comdist function in the picante package
@@
 
-We use a linear mixed model to regress weighted interaction turnover
-and site type and find no statistically significant differences.
+abund.w = TRUE weighted dissimilarity scores by species abundances.
+
+We use a linear mixed model to regress weighted interaction turnover. When
+interactions were weighted by their similarity, both assembling and
+mature hedgerows had higher rates of turnover than field margins.
 
@@
 ##                   Estimate Std. Error       df t value Pr(>|t|)    
-## (Intercept)         0.90778    0.01965 37.00000  46.198  < 2e-16 ***
-## SiteStatuscontrol  -0.08206    0.02414 35.13000  -3.400  0.00169 ** 
-## SiteStatusmaturing  0.03339    0.03065 33.20000   1.089  0.28385    
+## (Intercept)        0.94117    0.02352 30.88000  40.014  < 2e-16 ***
+## SiteStatuscontrol -0.11545    0.02738 31.12000  -4.216 0.000198 ***
+## SiteStatusmature  -0.03339    0.03065 33.20000  -1.089 0.283853    
 ## ---
@@
 ##             (Intr) StSttsc
-## StSttscntrl -0.814        
-## StSttsmtrng -0.641  0.522
+## StSttscntrl -0.859        
+## SiteSttsmtr -0.767  0.659
 \end{verbatim}
@@
 The baci.R script (before-after-control-impact, baci, the orginal
-design of the study, vs.~ the word for kisses in Italian) calculates
+design of the study, vs.~the word for kisses in Italian) calculates
 various network metrics for the assembling hedgerows, standardizes
@@
 
-Calculate and standardize the network metrics. CalcNetworkMetrics
-takes an argument "networkLevel.mets" which is passed to networkLevel
-within the package bipartite \citep{bipartite} if other metrics are
-desired. Unlike the turnover analysis, the networks include
-non-integers because the mean was taken over sampling rounds within a
-year. We therefore use the null model proposed by \citep{Galeano2009}
-to randomize the communities while fixing the total number of
-interactions, species and interaction frequency distributions (though
-any function that takes a single interaction matrix as an argument can
-be substituted using the null.fun argument of the calcNullStat
-function.
-
-Calculate and pretty network metrics.
+CalcNetworkMetrics takes an argument "index" which is passed to
+networkLevel within the package bipartite \citep{bipartite} if other
+metrics are desired. Unlike the turnover analysis, the networks
+include non-integers because the mean was taken over sampling rounds
+within a year. We therefore use the null model proposed by
+\citep{Galeano2009} to randomize the communities while fixing the
+total number of interactions, species and interaction frequency
+distributions (though any function that takes a single interaction
+matrix as an argument can be substituted using the null.fun argument
+of the calcNullStat function.
 
@@
 Then regress the nestedness, modularity, specialization and
-connectance by time ("ypr" in the data, years post restoration).
+connectance, plant/pollinator species richness by time ("ypr" in the
+data, years post restoration).
 
@@
 \begin{alltt}
+\hlcom{## ************************************************************}
+\hlcom{## linear mixed models}
+\hlstd{ys} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlstr{"zNODF"}\hlstd{,} \hlstr{"zmod.met.D"}\hlstd{,} \hlstr{"zH2"}\hlstd{,} \hlstr{"connectance"}\hlstd{)}
+
+\hlstd{formulas} \hlkwb{<-}\hlkwd{lapply}\hlstd{(ys,} \hlkwa{function}\hlstd{(}\hlkwc{x}\hlstd{) \{}
+    \hlkwd{as.formula}\hlstd{(}\hlkwd{paste}\hlstd{(x,} \hlstr{"~"}\hlstd{,}
+                     \hlkwd{paste}\hlstd{(}\hlstr{"scale(ypr)"}\hlstd{,}
+                           \hlstr{"(1|Site)"}\hlstd{,}
+                           \hlstr{"(1|Year)"}\hlstd{,}
+                           \hlkwc{sep}\hlstd{=}\hlstr{"+"}\hlstd{)))}
+\hlstd{\})}
+
+\hlstd{formulas.spac} \hlkwb{<-}\hlkwd{lapply}\hlstd{(ys,} \hlkwa{function}\hlstd{(}\hlkwc{x}\hlstd{) \{}
+    \hlkwd{as.formula}\hlstd{(}\hlkwd{paste}\hlstd{(x,} \hlstr{"~"}\hlstd{,}
+                     \hlkwd{paste}\hlstd{(}\hlstr{"scale(ypr)"}\hlstd{,}
+                           \hlkwc{sep}\hlstd{=}\hlstr{"+"}\hlstd{)))}
+\hlstd{\})}
+
+\hlstd{mods} \hlkwb{<-} \hlkwd{lapply}\hlstd{(formulas,} \hlkwa{function}\hlstd{(}\hlkwc{x}\hlstd{)\{}
+    \hlkwd{lmer}\hlstd{(x,}
+         \hlkwc{data}\hlstd{=baci)}
+\hlstd{\})}
+
+\hlkwd{names}\hlstd{(mods)} \hlkwb{<-} \hlstd{ys}
+\hlcom{## results}
 \hlkwd{lapply}\hlstd{(mods, summary)}
 \end{alltt}
+\begin{verbatim}
+## $zNODF
+## Linear mixed model fit by REML t-tests use Satterthwaite approximations
+##   to degrees of freedom [lmerMod]
+## Formula: zNODF ~ scale(ypr) + (1 | Site) + (1 | Year)
+##    Data: baci
+## 
+## REML criterion at convergence: 186.2
+## 
+## Scaled residuals: 
+##     Min      1Q  Median      3Q     Max 
+## -1.7060 -0.5585 -0.1869  0.4507  2.2506 
+## 
+## Random effects:
+##  Groups   Name        Variance Std.Dev.
+##  Year     (Intercept) 1.477    1.215   
+##  Site     (Intercept) 6.427    2.535   
+##  Residual             8.241    2.871   
+## Number of obs: 36, groups:  Year, 8; Site, 5
+## 
+## Fixed effects:
+##             Estimate Std. Error     df t value Pr(>|t|)   
+## (Intercept)   6.1644     1.3091 4.7280   4.709   0.0061 **
+## scale(ypr)    1.2424     0.6717 6.2160   1.850   0.1121   
+## ---
+## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
+## 
+## Correlation of Fixed Effects:
+##            (Intr)
+## scale(ypr) 0.008 
+## 
+## $zmod.met.D
+## Linear mixed model fit by REML t-tests use Satterthwaite approximations
+##   to degrees of freedom [lmerMod]
+## Formula: zmod.met.D ~ scale(ypr) + (1 | Site) + (1 | Year)
+##    Data: baci
+## 
+## REML criterion at convergence: 116.8
+## 
+## Scaled residuals: 
+##     Min      1Q  Median      3Q     Max 
+## -1.8149 -0.7167 -0.2496  0.7042  2.8513 
+## 
+## Random effects:
+##  Groups   Name        Variance Std.Dev.
+##  Year     (Intercept) 0.000    0.000   
+##  Site     (Intercept) 0.000    0.000   
+##  Residual             1.471    1.213   
+## Number of obs: 36, groups:  Year, 8; Site, 5
+## 
+## Fixed effects:
+##             Estimate Std. Error      df t value Pr(>|t|)  
+## (Intercept)  -0.4738     0.2022 34.0000  -2.344   0.0251 *
+## scale(ypr)    0.2363     0.2050 34.0000   1.152   0.2572  
+## ---
+## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
+## 
+## Correlation of Fixed Effects:
+##            (Intr)
+## scale(ypr) 0.000 
+## 
+## $zH2
+## Linear mixed model fit by REML t-tests use Satterthwaite approximations
+##   to degrees of freedom [lmerMod]
+## Formula: zH2 ~ scale(ypr) + (1 | Site) + (1 | Year)
+##    Data: baci
+## 
+## REML criterion at convergence: 189.7
+## 
+## Scaled residuals: 
+##     Min      1Q  Median      3Q     Max 
+## -2.0587 -0.4511 -0.1303  0.5563  2.0243 
+## 
+## Random effects:
+##  Groups   Name        Variance Std.Dev.
+##  Year     (Intercept) 3.509    1.873   
+##  Site     (Intercept) 3.461    1.860   
+##  Residual             9.107    3.018   
+## Number of obs: 36, groups:  Year, 8; Site, 5
+## 
+## Fixed effects:
+##             Estimate Std. Error      df t value Pr(>|t|)   
+## (Intercept)  -5.9515     1.1889  6.2850  -5.006  0.00213 **
+## scale(ypr)   -1.4434     0.8571  6.8770  -1.684  0.13680   
+## ---
+## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
+## 
+## Correlation of Fixed Effects:
+##            (Intr)
+## scale(ypr) 0.025 
+## 
+## $connectance
+## Linear mixed model fit by REML t-tests use Satterthwaite approximations
+##   to degrees of freedom [lmerMod]
+## Formula: connectance ~ scale(ypr) + (1 | Site) + (1 | Year)
+##    Data: baci
+## 
+## REML criterion at convergence: -99.5
+## 
+## Scaled residuals: 
+##      Min       1Q   Median       3Q      Max 
+## -1.62761 -0.54457 -0.04675  0.38328  2.08088 
+## 
+## Random effects:
+##  Groups   Name        Variance  Std.Dev.
+##  Year     (Intercept) 0.0000000 0.00000 
+##  Site     (Intercept) 0.0009775 0.03126 
+##  Residual             0.0021487 0.04635 
+## Number of obs: 36, groups:  Year, 8; Site, 5
+## 
+## Fixed effects:
+##              Estimate Std. Error        df t value Pr(>|t|)    
+## (Intercept)  0.210108   0.015979  3.937000   13.15 0.000213 ***
+## scale(ypr)  -0.022972   0.008004 30.733000   -2.87 0.007359 ** 
+## ---
+## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
+## 
+## Correlation of Fixed Effects:
+##            (Intr)
+## scale(ypr) -0.003
+\end{verbatim}
 \end{kframe}
@@
 improved by adding the additional parameters. The spline chose to fit
-a line, indicating that the linear models were sufficient. They are
-time consuming to run so the results are not included here.
+a line, indicating that the linear models were sufficient.
+
+And the splines:
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## ************************************************************}
+\hlcom{## splines}
+
+\hlcom{## for gaussian models}
+\hlstd{mods.splines} \hlkwb{<-} \hlkwd{lapply}\hlstd{(ys, calcSpine)}
+\end{alltt}
+\begin{verbatim}
+## $info
+## [1] 0
+## 
+## $smoothingParameters
+##    mu     v 
+## 10000 10000 
+## 
+## $info
+## [1] 0
+## 
+## $smoothingParameters
+##    mu     v 
+## 11000 10000 
+## 
+## $info
+## [1] 0
+## 
+## $smoothingParameters
+##    mu     v 
+## 10000 10000 
+## 
+## $info
+## [1] 0
+## 
+## $smoothingParameters
+##    mu     v 
+## 21437  1750
+\end{verbatim}
+\begin{alltt}
+\hlkwd{names}\hlstd{(mods.splines)} \hlkwb{<-} \hlstd{ys}
+
+\hlcom{## for poisson models}
+\hlstd{mods.splines.log} \hlkwb{<-} \hlkwd{lapply}\hlstd{(poi.ys, calcSpine,} \hlkwc{calc.log}\hlstd{=}\hlnum{TRUE}\hlstd{)}
+\end{alltt}
+\begin{verbatim}
+## $info
+## [1] 0
+## 
+## $smoothingParameters
+##    mu     v 
+## 11000 10000 
+## 
+## $info
+## [1] 0
+## 
+## $smoothingParameters
+##    mu     v 
+## 11000 10000 
+## 
+## $info
+## [1] 0
+## 
+## $smoothingParameters
+##    mu     v 
+## 11000 10000
+\end{verbatim}
+\begin{alltt}
+\hlkwd{names}\hlstd{(mods.splines.log)} \hlkwb{<-} \hlstd{poi.ys}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+The smoothing parameters indicate lines were fit.
 
@@
 
-This area is then regressed against time ("ypr" again).
+This area is then regressed against time ("ypr") Random effects of
+year and site are included.
 
@@
 ## 
-## REML criterion at convergence: -167.6
+## REML criterion at convergence: -166.9
 ## 
@@
 ##      Min       1Q   Median       3Q      Max 
-## -2.03661 -0.73640 -0.02933  0.62103  2.40910 
+## -1.99009 -0.62116 -0.07289  0.56358  2.37929 
 ## 
@@
 ##  Groups   Name        Variance  Std.Dev.
-##  Site     (Intercept) 0.0019852 0.04456 
-##  Year     (Intercept) 0.0001163 0.01078 
-##  Residual             0.0048546 0.06967 
+##  Site     (Intercept) 2.395e-03 0.048943
+##  Year     (Intercept) 5.091e-05 0.007135
+##  Residual             4.842e-03 0.069583
 ## Number of obs: 80, groups:  Site, 14; Year, 8
@@
 ##              Estimate Std. Error        df t value Pr(>|t|)    
-## (Intercept)  0.331489   0.016356 11.040000  20.267 4.36e-10 ***
-## ypr          0.001166   0.004144 34.880000   0.281     0.78    
+## (Intercept)  0.333299   0.016943 10.950000  19.671 6.85e-10 ***
+## ypr          0.000965   0.004133 32.430000   0.233    0.817    
 ## ---
@@
 ##     (Intr)
-## ypr -0.395
+## ypr -0.374
 \end{verbatim}
@@
 \end{knitrout}
+
+The next estimate of resilience we use is based on how quickly
+information spreads through a network as estimated by algebraic
+connectivity.
 

### 2017-05-18 analysis/networkLevel/baci.R

@@ -12,10 +12,6 @@ N <- 999
 ## ************************************************************
 ## mets <- lapply(nets, calcNetworkMetrics, N)
 ## cor.dats <- prepDat(mets,  spec)
-
-## cor.dats$tot.rich <- cor.dats$number.of.species.LL +
-##     cor.dats$number.of.species.HL
-
 ## save(cor.dats, file='saved/corMets.Rdata')
 
 ## ************************************************************
@@ -26,9 +22,10 @@ load(file='saved/corMets.Rdata')
 baci <- cor.dats[!is.na(cor.dats$ypr),]
 baci$cYear <- as.numeric(baci$Year)-min(as.numeric(baci$Year))
 
-## ************************************************************
-## linear models
 
+## @knitr external_mets_lm
+## ************************************************************
+## linear mixed models
 ys <- c("zNODF", "zmod.met.D", "zH2", "connectance")
 
 formulas <-lapply(ys, function(x) {
@@ -52,11 +49,34 @@ mods <- lapply(formulas, function(x){
 
 names(mods) <- ys
 ## results
-## @knitr external_mets_lm
 lapply(mods, summary)
 ## @knitr external_mets_lm_end
 aic.mods <- sapply(mods, AIC)
 
+## ************************************************************
+## generalized linear mixed models
+poi.ys <- c("number.of.species.HL", "number.of.species.LL")
+
+formulas.poi <-lapply(poi.ys, function(x) {
+    as.formula(paste(x, "~",
+                     paste("scale(ypr)",
+                           "(1|Site)",
+                           "(1|Year)",
+                           sep="+")))
+})
+
+mods.poi <- lapply(formulas.poi, function(x){
+    glmer(x,
+          data=baci, family="poisson")
+})
+names(mods.poi) <- poi.ys
+
+## results
+lapply(mods.poi, summary)
+## @knitr external_mets_lm_end
+
+
+## @knitr external_mets_lm_autocorr
 ## ************************************************************
 ## standard discrete-time autocorrelation first order model
 ## takes a while to optimize
@@ -82,44 +102,32 @@ lapply(mods.spatial, summary)
 aic.sp.mods <- sapply(mods.spatial, AIC)
 
 aic.mods-aic.sp.mods
+## spatial autocorrelation models have a larger AIC value
+## @knitr external_mets_lm_autocorr
 
+
+
+## @knitr external_mets_lm_spline
[…]

### 2017-05-18 analysis/networkLevel/src/CalcMetrics.R

@@ -1,8 +1,5 @@
 
-calcMetric <- function(dat.web, H2_integer=FALSE,
-                        networkLevel.mets= c("H2", "connectance",
-                          "weighted connectance", "niche overlap",
-                          "number of species")) {
+calcMetric <- function(dat.web, ...) {
   ## calculates modularity
   calc.mod <- function(dat.web){
     ## converts a p-a matrix to a graph for modularity computation
@@ -49,9 +46,7 @@ calcMetric <- function(dat.web, H2_integer=FALSE,
                        weighted=TRUE,
                        wbinary=TRUE)$statistic["NODF"]
     mets <-  c(nodf,
-               networklevel(dat.web,
-                            index= networkLevel.mets,
-                            H2_integer=H2_integer))
+               networklevel(dat.web, ...))
   }
   mod.met <- calc.mod(dat.web)
   return(c(mets, mod.met= mod.met))
@@ -105,14 +100,19 @@ calcProbNull <- function(M) {
 }
 
 ## function to simulate 1 null, and calculate statistics on it
-calcNullStat <- function(dat.web, null.fun= calcProbNull) {
+calcNullStat <- function(dat.web,
+                         null.fun= calcProbNull,...) {
   sim.web <- null.fun(dat.web)
-  return(calcMetric(sim.web))
+  return(calcMetric(sim.web, ...))
 }
 
 ##  function that computes summary statistics on simulated null matrices
 ##  (nulls simulated from web N times)
-calcNetworkMetrics <- function (dat.web, N) {
+calcNetworkMetrics <- function (dat.web, N,
+                                H2_integer=FALSE,
+                                index= c("H2", "connectance",
+                          "weighted connectance", "niche overlap",
+                          "number of species")) {
   ## calculate pvalues
   pvals <- function(stats, nnull){
     rowSums(stats >= stats[, rep(1, ncol(stats))])/(nnull + 1)
@@ -137,7 +137,8 @@ calcNetworkMetrics <- function (dat.web, N) {
         null.stat <- replicate(N, calcNullStat(dat.web),
                                simplify=TRUE)
         ## calculate metrics from data
-        true.stat <- calcMetric(dat.web)
+          true.stat <- calcMetric(dat.web, H2_integer=H2_integer,
+                                  index=index)
         out.mets <- cbind(true.stat, null.stat)
         ## compute z scores
         zvalues <- zvals(out.mets)

### 2017-05-18 analysis/variability/cv.R

@@ -89,38 +89,6 @@ cor.test(pol.cv$lm.data$r.degree, pol.cv$lm.data$occ.date)
 cor.test(plant.cv$lm.data$plant.r.degree,
          plant.cv$lm.data$occ.plant.date)
 
-## ************************************************************
-## @knitr external_cv_resids
-## residual plots
-
-resid.plot <- function(){
-    layout(matrix(1:6, nrow=2, byrow=TRUE))
-    plot(fitted(pol.mod), residuals(pol.mod),
-         xlab = "Fitted Values", ylab = "Residuals",
-         main="Pollinators")
-    abline(h=0, lty=2)
-    lines(smooth.spline(fitted(pol.mod),
-                        residuals(pol.mod)))
-
-    qqnorm(residuals(pol.mod), abline(0,1))
-
-    plot(density(residuals(pol.mod)))
-
-    plot(fitted(plant.mod),
-         residuals(plant.mod),
-         xlab = "Fitted Values", ylab = "Residuals",
-         main="Plants")
-    abline(h=0, lty=2)
-    lines(smooth.spline(fitted(plant.mod),
-                        residuals(plant.mod)))
-    qqnorm(residuals(plant.mod), abline(0,1))
-
-    plot(density(residuals(plant.mod)))
-}
-
-## passable
-resid.plot()
-
 ## save
 save(pol.cv, pol.mod, plant.cv, plant.mod,
      file="saved/contMods.Rdata")

### 2017-05-18 analysis/variability/plotting/cv.R

@@ -88,3 +88,36 @@ plant.degree.pi.close <- predict.int(mod= plant.mod,
 plot.panels()
 
 
+## ************************************************************
+## residual plots
+
+resid.plot <- function(){
+    layout(matrix(1:6, nrow=2, byrow=TRUE))
+    plot(fitted(pol.mod), residuals(pol.mod),
+         xlab = "Fitted Values", ylab = "Residuals",
+         main="Pollinators")
+    abline(h=0, lty=2)
+    lines(smooth.spline(fitted(pol.mod),
+                        residuals(pol.mod)))
+
+    qqnorm(residuals(pol.mod), abline(0,1))
+
+    plot(density(residuals(pol.mod)))
+
+    plot(fitted(plant.mod),
+         residuals(plant.mod),
+         xlab = "Fitted Values", ylab = "Residuals",
+         main="Plants")
+    abline(h=0, lty=2)
+    lines(smooth.spline(fitted(plant.mod),
+                        residuals(plant.mod)))
+    qqnorm(residuals(plant.mod), abline(0,1))
+
+    plot(density(residuals(plant.mod)))
+}
+
+## passable
+resid.plot()
+
+
+

### 2017-05-18 analysis/writeup/writeup.tex
lots of small cahnges for reproducibility

@@
 ##                  Value Std. Error lower bound upper bound Pr(>|t|)    
-## (Intercept)  3.7160065  0.0854433   3.5464685      3.8855  < 2e-16 ***
-## occ.date     0.7040012  0.2454156   0.2170433      1.1910  0.00504 ** 
-## r.degree    -0.0010938  0.0032266  -0.0074960      0.0053  0.73533    
-## scale        0.2869339  0.0185467   0.2501333      0.3237  < 2e-16 ***
+## (Intercept)  3.7160065  0.0767061   3.5638049      3.8682  < 2e-16 ***
+## occ.date     0.7040012  0.2686421   0.1709569      1.2370  0.01016 *  
+## r.degree    -0.0010938  0.0026191  -0.0062906      0.0041  0.67712    
+## scale        0.2869339  0.0191043   0.2490269      0.3248  < 2e-16 ***
 ## ---
@@
 ## 11000 10000
-## 
-## $info
-## [1] 0
-## 
-## $smoothingParameters
-##    mu     v 
-## 11000 10000
 \end{verbatim}
@@
 ## 
-## REML criterion at convergence: -166.9
+## REML criterion at convergence: -165.6
 ## 
@@
 ##     Min      1Q  Median      3Q     Max 
-## -1.99009 -0.62116 -0.07289  0.56358  2.37929 
+## -1.9965 -0.6661 -0.0787  0.5873  2.3210 
 ## 
@@
 ##  Groups   Name        Variance  Std.Dev.
-##  Site     (Intercept) 2.395e-03 0.048943
-##  Year     (Intercept) 5.091e-05 0.007135
-##  Residual             4.842e-03 0.069583
+##  Site     (Intercept) 0.0022317 0.04724 
+##  Year     (Intercept) 0.0001305 0.01143 
+##  Residual             0.0049176 0.07013 
 ## Number of obs: 80, groups:  Site, 14; Year, 8
@@
 ##              Estimate Std. Error        df t value Pr(>|t|)    
-## (Intercept)  0.333299   0.016943 10.950000  19.671 6.85e-10 ***
-## ypr          0.000965   0.004133 32.430000   0.233    0.817    
+## (Intercept) 3.320e-01  1.703e-02 1.127e+01  19.501 4.91e-10 ***
+## ypr         8.384e-04  4.231e-03 3.627e+01   0.198    0.844    
 ## ---
@@
 ##     (Intr)
-## ypr -0.374
+## ypr -0.386
 \end{verbatim}
